### 2021-04-13 papers/paper.tex
add the paper

\documentclass[12pt]{article}

% Figures
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{{../plots/}{../tikz/}}
\usepackage[section]{placeins}

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, fleqn}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage{libertine}

% page size and margins
\usepackage{geometry}
\geometry{letterpaper,margin=1in}
\usepackage{fancyhdr}

% better tables
\usepackage{booktabs}
\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

% better lists
\usepackage{enumitem}
\setlist{nosep}

% authors
\usepackage{authblk}
\title{Mainstreaming Adaptive Resilience: how high to elevate a house given multiple PDFs of future flood hazard?}
\author[1]{James Doss-Gollin}
\author[2,3]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Department of Geosciences, the Pennsylvania State University}
\affil[3]{Earth and Environmental Systems Institute, the Pennsylvania State University}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{iid}{IID}{independent and identically distributed}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{mordm}{MORDM}{multiobjective \gls{rdm}}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\newcommand{\usd}[1]{\SI{#1}[\$]{}}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{../_bibliography/library.bib}

% load this last
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

% up to 1250 words
\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
    In response to evolving real and perceived flood risk, homeowners, often subsidized by federal and local programs, may invest in building-scale floodproofing measures such as house elevation.
    Federal guidance as to how high to elevate a building relies on floodplain maps that reflect current hazard, suggesting that in regions where flood risk is fast changing, such as along coastlines, incorporating nonstationary flood risk into decision support might lead to dramatically different guidance.
    However, future flood hazard is subject to large and deep uncertainties driven by climate forcings, sea level response, and economics.
    Using a case study of elevating a single home in Norfolk, VA, we first show that decisions tailored to a particular set of assumptions may perform poorly under another.
    This illustrates that even when deep uncertainties are modeled, decision makers must make subjective design choices.
    We present a framework for robust planning under deep uncertainty that transparently maps subjective beliefs to stochastic optimization, and that facilitates computationally cheap \emph{a posteriori} robustness checks to evaluate performance under alternative subjective beliefs.
    We find that ignoring uncertainty between \gls{rcp} scenarios [\texttt{leads to bad outcomes}\ldots].
\end{abstract}
\subsection*{Submission Plan}
We should revisit this!
\begin{enumerate}
    \item Climatic Change
    \item IOP Environmental Research: Infrastructure and Sustainability (new journal)
    \item Earth's Future
    \item Risk analysis
\end{enumerate}
\subsection*{Internal peer review}
\begin{enumerate}
    \item TBD
\end{enumerate}

\clearpage
\section{Introduction}
\begin{enumerate}
    \item Home elevation is an important problem
          \begin{enumerate}
              \item Floodproofing and building-scale vulnerability reduction measures can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn}
              \item \citet{mobley_mitigation:2020} shows wide potential for elevation to reduce flood losses
              \item \citet{cardenas_elevation:2018} shows that rich Houston residents are already elevating their houses
              \item \citet{aerts_cost:2018} looks at costs of elevation
          \end{enumerate}
    \item Prior studies have illuminated a lot
          \begin{enumerate}
              \item Existing engineering guidance and policy recommend elevating to the \gls{bfe} plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014}
              \item \citet{xian_elevation:2017} minimizes discounted expected lifetime costs to show that the optimal elevation can differ substantially from \gls{fema} recommendations
              \item \citet{zarekarizi_suboptimal:2020} expands on this analysis to consider multiple objectives (including those that capture the hard to quantify disruptions of experiencing a flood) and multiple sources of uncertainty (\eg, discount rate, flood probability, and house lifetime)
              \item However, these analyses are silent on the question of how to address deep uncertainties in nonstationary flood hazard
          \end{enumerate}
    \item Multiple \glspl{pdf} present a challenge for coastal adaptation
          \begin{enumerate}
              \item Coastal floods are often decomposed into changes in local sea level \citep{kopp_evolving:2017,kopp_probabilistic:2014,wong_brick0.2:2017} plus storm surge \citep{garner_slrise:2018,cagigal_emulator:2020,rueda_surge:2016}
              \item The choices of \gls{rcp} scenario and model structure matter a lot for sea level \citep{kopp_probabilistic:2014,kopp_evolving:2017,wong_brick0.2:2017,ruckert_coastal:2019,wong_nola:2017}
              \item The choice of statistical representation of storm surge risk matters for future surge risk \citep{wong_nola:2017,wong_structural:2020}
              \item \citet{wong_structural:2020} illustrates the equifinality problem: there may be one or more models that fit the historical data equally well, but give very different results into the future.
          \end{enumerate}
    \item How can robust decision support transparently balance the goals of (i) sampling the tails of plausible future conditions; (ii) integrating information from multiple \gls{rcp} scenarios and models; and (iii) exploiting probabilistic beliefs about the future?
          \begin{enumerate}
              \item \citet{bankes:1993,lempert_shaping:2003,Brown:2012kb} emphasize the importance of using models to explore possibilities and consequences rather than trying to consolidate all information into a single prediction
              \item Exploratory modeling enables the use of non-optimization analytic methods like scenario discovery \citep{kwakkel:2019,lamontagne_discovery:2018}
              \item While we agree with the importance of exploratory modeling, it defers to the decision maker all choices about how to make decisions given a map of possibilities, decisions, and consequences
              \item This motivates a search for ``robust'' approaches, which \citet{herman:2015} reviews how others have conceived of robustness, focusing primarily on identifying the fraction of a parameter space over which some constraints are met
              \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
              \item \citet{mcphail_robustness:2019} elaborates on robustness metrics and their implications
              \item However, these methods are silent on the question of what the distribution or set of \glspl{sow} over which these metrics are calculated should be, and what the robustness of inference is to alternatives.
          \end{enumerate}
\end{enumerate}

In this paper we \ldots

\section{Methods}\label{sec:methods}

We first present a general framework for decision making under uncertainty (\cref{sec:methods-framework}).
This approach is designed for decision problems with a multiplicity of possible futures and outcomes whose likelihood of occurring can be represented quantatively, but necessarily using highly subjective representations.
We then illustrate this method through a didactic case study of deciding if, and how high, to elevate a single house in Norfolk, VA (\cref{sec:methods-case}).
We finally compare our approach to two other robustness metrics (\cref{sec:methods-alternative}).

\subsection{Decision framework}\label{sec:methods-framework}

\begin{figure}
    \includegraphics[width=\textwidth]{conceptual.pdf}
    \caption{
        Schematic of general decision framework and definitions.
        Nodes in green are modeling inputs, those in yellow are derived quantities, and those in gray describe opportunities for iterative model critique and examination to support decision-making.
    }\label{fig:framework}
\end{figure}
Our decision framework, illustrated in \cref{fig:framework}, draws from three main bodies of literature.

First, the literature on \gls{dmdu} emphasizes the importance of identifying decisions that perform well over a wide range of possible futures.
Exploratory modeling can demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
Exploring many possible futures is particularly important when facing deep and epistemic uncertainties \citep{lempert_shaping:2003,wong_nola:2017,oddo_coastal:2017}
Many tools used for \gls{dmdu} including decision scaling \citep{Brown:2012kb}, robust decision making \citep{lempert:2007,kasprzyk:2013,kasprzyk_denovo:2012}, sensitivity analysis \citep{saltelli_comment:2019}, and scenario discovery \citep{kwakkel:2019,lamontagne_discovery:2018}, build on the idea of exploratory modeling, and we incorporate this philosophy into our approach.

The observation that many of these mechanisms cannot be represented by a single objective \gls{pdf} has motivated many criticisms of the application of \gls{bdt} to planning problems.
Yet \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
\citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.

Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}.

\subsubsection{Exploratory modeling}

Following \citet{mcphail_robustness:2019}, we consider a set of decisions $x_1, \ldots, x_m \in \mathcal{X}$ as well as a set of plausible future conditions (scenarios) $s_1, \ldots, s_n \in \Omega$.
We refer to $\mathcal{X}$ as the decision space and $\Omega$ as the event space.
To evaluate a given decision $x_i$, we use an arbitrary system model $f: \mathcal{X} \times \mathcal{S} \rightarrow \mathcal{U}$ to compute a scalar or vector performance metric $u \in \mathcal{U}$ for each scenario.
We define the performance of the $i$th decision on the $j$th scenario as $u_{ij} = f(x_i, s_j)$ and call the $u_{ij}$ ``scenario metrics.''
The full set $u_{i1}, \ldots, u_{in}$ can be used in a fully exploratory context \citep[\eg, scenario discovery;][]{kwakkel:2019,lamontagne_discovery:2018} to identify the conditions under which a given decision $x_i$ performs well or poorly.
These consequences can also be shared with stakeholders to improve the model, assess the completeness of scenarios considered, or expand the decision set.

\subsubsection{Probabilistic modeling}

If the decision space is small, then exploratory modeling may provide sufficient information for decision making.
However, as the decision space grows large it becomes helpful to develop ``decision metrics'' $\hat{m}(x_i)$, that summarize performance across all scenarios.

A general approach is define the decision metrics as a weighted sum of the scenario metrics: $\hat{m}(x_i) = \sum_{j=1}^n w_j u_{ij}$.
Many commonly used heuristics can be described using this framework.
In the most extreme case, maximum and minimum metrics set all $w_j$ but one to zero, conveying information only about worst-case performance.
In the other extreme, setting $wj \propto 1$ conveys information about all scenarios, but is sensitive to the set of scenarios considered.
A more stable alternative is to set the weights using a \gls{pdf} to describe subjective belief about the likelihood of each scenario, $p_\text{belief}(\Sigma)$, and then to use this \gls{pdf} to set the scenario weights.
Given a probabilistic belief model on the sample space $p_\text{belief}(x)$, the weight assigned to the $i$th scenario is
\begin{equation}\label{eq:scenario-weight}
    w_i \propto \frac{p_\text{belief}(x_i)}{\hat{p}_\text{sampling}(x_i)},
\end{equation}
where $\hat{p}_\text{sampling}$ is the sampling distribution estimated using, \eg, kernel density estimation.
\Cref{eq:scenario-weight} is equivalent to the definition of rejection sampling, except that instead of accepting draws from a proposal distribution (here $\hat{p}_\text{sampling}$) with probability $w_i$ to generate \gls{iid} draws from the target distribution (here $p_\text{belief}$), we retain and weight all scenarios.

Setting the weights from a \gls{pdf} has several advantages:
\begin{enumerate}
    \item we gain a formal justification for reasoning about the \emph{distribution} of outcomes, $\hat{p}(y | x_i)$, rather than merely a \emph{set} of possible outcomes;
    \item the decision metric $\hat{m}(x_i)$ is stable if the sampling distribution is changed (\eg, if new scenarios are added); and
    \item the subjective assumption on $p_\text{belief}$ is highly transparent and can be critiqued, \eg by interpreting $p_\text{belief}$ as a prior in a Bayesian framework \citep{Gelman:2014tc,McElreath:2016vu} and conducting prior predictive checks to identify, improve, and communicate the joint implications of the model and prior \citep[see][section 2.4]{gelman_workflow:2020}.
\end{enumerate}
Since the choice of \gls{pdf} to represent the analyst's subjective belief about the likelihood of different scenario is necessarily subjective, it is advisable to formulate several plausible \glspl{pdf} and compare performance across them.
The computational effort required for this is quite low: using a different numerator in \cref{eq:scenario-weight}, the weights assigned to each scenario are updated, but the system model does not need to be re-run.
Thus, answering the question ``what does the distribution''
We illustrate this approach in the house elevation case study (\cref{sec:methods-case}).

\subsubsection{Iterative model critique and robustness checks}

Drawing from iterative model building philosophies of the statistical and decision science communities, the fundamental goal of our decision analysis is to use a model to transparently examine the implications of various assumptions.
\Cref{fig:framework} illustrates several opportunities for iterative validation, critique, and improvement of the full decision problem.

We next turn to the problem posed by the intrinsically subjective choice of $p_\text{belief}$.
The \gls{dmdu} literature emphasizes the importance of identifying decisions that are robust, often defined as trading some performance in the most likely scenario for performance in other plausible futures.
In

Since our $p_\text{belief}$ is subjective, we define robustness in terms of alternative beliefs that other reasonable analysts might hold.
Obviously the full space of all possible beliefs cannot be computed any more than the full set of all possible models or datasets or problem framings can be defined, and so such a comparison is necessarily qualitative.
However, many alternative beliefs can easily be computed because only the $w_i$ need to be recomputed to evaluate performance under another belief; the $y_{ij}$ are unchanged and thus, the model does not need to be re-run.

\subsection{Case study: house elevation}\label{sec:methods-case}

\begin{table}
    \centering
    \begin{tabular}{llr}
        \toprule
        Uncertainty           & Treatment                                    & Discussed in                      \\
        \midrule
        Sea Level             & Deep uncertainty: 4 \acrshort{rcp} scenarios & \cref{sec:methods-lsl}            \\
        Storm Surge           & Parametric uncertainty                       & \cref{sec:methods-surge}          \\
        Discount Rate         & Fixed constant                               & \cref{sec:methods-cost}           \\
        Planning Horizon      & Fixed constant                               & \cref{sec:methods-cost}           \\
        Depth-Damage Function & Deterministic and known                      & \cref{sec:methods-cost-damage}    \\
        Cost of Elevation     & Deterministic and known                      & \cref{sec:methods-cost-elevation} \\
        \bottomrule
    \end{tabular}
    \caption{
        Treatment of potential sources of uncertainty.
        As indicated, strong assumptions are made as to some potential sources of uncertainty in order to maintain interpretability of findings.
    }\label{sec:tab-uncertainties}
\end{table}

\subsection{Sea Level}\label{sec:methods-lsl}

We use BRICK \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019}, which is sufficient to include both model structure uncertainty (fast or slow dynamics) and human factors (\gls{rcp} 2.6, 4.5, 6.0, or 8.5).

\subsection{Storm Surge}\label{sec:methods-surge}

Define $\mu^*(t)=\mu_0 + \beta \qty(t-t_0)$, where $t_0$ is 2020. Then $y'(t) \sim \text{GEV}(\mu^*(t), \rho \mu^*(t), \xi)$, where $\rho$ is a coefficient of variation. We can put together a single model for storm surge with informative priors, and note that there are other models that we could look at in future work.

\subsection{Costs}\label{sec:methods-cost}

Our objective is to minimize discounted total costs (initial cost of construction plus expected future damages), taken as a weighted average over $N$ modeled realizations of the future, over a fixed 50 year period.
$$\min C_e(\Delta h) + \sum_{i=1}^N \sum_{t=2020}^{2070} \gamma_t w_i D(y_{i,t}, \Delta h) $$
where $D$ is the loss or damage function and $w_i$ is a weight applied to the $i$th scenario such that $\sum_{i=1}^N w_i = 1$.
We can keep the discount rate fixed for concision: $\gamma_t = 0.97^t$ is a 3\% rate -- we can vary this number but let's keep it simple.

\subsubsection{Flood Damage}\label{sec:methods-cost-damage}

Flood damage as \citet{zarekarizi_suboptimal:2020}.

\subsubsection{Elevation Cost}\label{sec:methods-cost-elevation}
Elevation Cost as \citet{zarekarizi_suboptimal:2020}.

\subsubsection{Belief model}

\begin{table}
    \centering
    \footnotesize
    \begin{tabular}{L{1in}L{0.75in}L{0.5in}L{3.5in}}
        \toprule
        Variable                     &
        distribution                 &
        University                   &
        Discussion                                                                                                                                           \\
        \midrule
        LSL in 2100                  &
        $\mathcal{N}(1, 1)$          &
        \si{m}                       &
        Express a belief about sea level in 2100, which is well studied, and treat all scenarios with the same $\overline{y}_{2100}$ as equally likely       \\
                                     &
        $\mathcal{N}(1.25, 1.25)$    &
                                     &
        Different prior belief                                                                                                                               \\
        RCP scenario                 &
        $\mathcal{N}(4.5, 1.5)$      &
        \si{\watt\per\meter\squared} &
        Express a prior belief that RCP 4.5 is most likely. This amounts to treating each scenario as IID given the RCP scenario and then weighting the PDFs \\
                                     &
        $\mathcal{N}(6.0, 1.5)$      &
                                     &
        As above, but belief that RCP 6 is most likely.                                                                                                      \\
        Uniform                      &
        $w_i \propto 1$              &
                                     &
        Treat all scenarios as equally likely, implying that $\hat{p}_\text{sampling} = p_\text{belief}$.                                                    \\
        \bottomrule
    \end{tabular}
    \caption{Prior beliefs}
\end{table}

\subsection{Decision Model}\label{sec:methods-decision}

\begin{enumerate}
    \item We sample $N$ scenarios from the set of plausible future outcomes, consistent with ideals of exploratory modeling \citep{bankes:1993} \citep[indeed we can use samples for scenario discovery;][]{kwakkel:2019}
    \item If $w_i = \frac{1}{N} \, \forall \, i \in 1, \ldots, N$ then all scenarios are treated equally; this is what is typically used in robustness metrics (though instead of a continuous objective function a binary satisficing constraint is often used; see \citealt{herman:2015} or \citealt{kwakkel:2019})
    \item However, this reveals a tension: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision
    \item To see how this works in high dimensions, if we make the (admittedly silly) assumption that the true distribution is uniform within a sphere and we sample the enclosing hypercube, the fraction of samples that are likely is
          $$ \frac{\pi^\frac{n}{2}}{\Gamma(\frac{n}{2} + 1)} \frac{1}{2^n}$$
          which if $n=7$ is about 3.7\%. (This is just an illustration of how high dimensions work)
    \item Similarly, we want decisions that do not change drastically if we change the sampling dimension. For example, if we explore \gls{rcp} 2.6, 4.5, 6, and 8.5, and then we consider \gls{rcp} 4.6 (just like 4.5 but very very slightly modified), our decisions should not change dramatically
    \item We therefore have two key steps:
          \begin{enumerate}
              \item A function $g: \mathcal{X} \rightarrow \mathcal{Z}$ that maps a realization of the future $x \in \mathcal{X}$ to a low-dimension representation $z \in \mathcal{Z}$. In our case, $\mathcal{Z}$ is just the sea level in 2100
              \item A subjective joint probability distribution over $\mathcal{Z}$: $p(\mathcal{Z})$
          \end{enumerate}
    \item This is conceptually similar to the approach of \citet{garner_slrise:2018}
    \item the steps are thus:
          \begin{enumerate}
              \item Start with a set of $N$ realizations of the future
              \item Define a low dimension space $\mathcal{Z}$ and a subjective probability belief $p(\mathcal{Z})$
              \item Compute the sampling distribution $p_\text{sampling}(\mathcal{Z})$ using, \eg, kernel density estimation
              \item Assign each realization a weight $w_i = \frac{p(\mathcal{Z})}{p_\text{sampling}(\mathcal{Z})}$.
              \item Run the optimization
              \item Optionally (but highly encouraged!), coonduct robustness checks by altering $p_\text{sampling}(\mathcal{Z})$ and recomputing expected outcomes. Since outcomes have already been calculated for each realization $x_i \forall i=1,\ldots,N$, outcomes can be recomputed with new weights with no need to run any additional calculations
          \end{enumerate}
\end{enumerate}

\subsection{Alternative decision frameworks}\label{sec:methods-alternative}

\subsubsection{Minimax regret}

\subsubsection{Hypervolume satisficing}

We follow \citet{zarekarizi_suboptimal:2020} exactly.

\section{Results}

\begin{enumerate}
    \item Show a scenario map, fixing $x_i$ to be FEMA recommendation (elevate to the 100 year flood estimated from a historic GEV plus freeboard). The two axes can represent mean sea level in 2070 and the 99th percentile of storm surge in 2070. Each dot can be a scenario.
\end{enumerate}

\section{Discussion}

\begin{enumerate}
    \item There are other objectives that might be useful for real-world decision support
          \begin{enumerate}
              \item We focused on cost minimization for simplicity but practical decision support might consider many objectives
              \item Real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house,
          \end{enumerate}
    \item Similarly, there are other considerations and sources of uncertainty that we could incorporate
          \begin{enumerate}
              \item Damage functions, cost of construction, house lifespan, discount rate, \etc
              \item We could have a sequential decision with learning (about sea level rise, cost of construction, and more)
              \item Data on cost of elevation and depth-damage is insufficiently recent, localized, and accurate.
              \item Optimization is for a single house, not many different structures and locations.
          \end{enumerate}
    \item Comparison to other definitions and understandings of robustness
          \begin{enumerate}
              \item One could pick a single sea level rise time series and optimize for it, then look at how well this meets some satisficing criteria for other plausible sea level rise time series. But how do you pick your best guess? How do you measure satisficing? How do you sample alternative SOWs against which to calculate this satisficing?
              \item One could do full stochastic optimization, \ie Bayesian Decision Theory. But how do you pick the ``objective'' probability distribution to use? In practice this often means choosing a model and RCP scenario and then drawing scenario-conditional samples.
              \item Our approach draws from both perspectives: we sample across deep uncertainties (like RCP scenario) using a subjective probability distribution, then assess robustness by quantifying how well this decision performs if other subjective distributions are used.
          \end{enumerate}
    \item There are links to a literature on model averaging
          \begin{enumerate}
              \item Stacking \citep{Yao:2018bu}
              \item Bayesian model averaging \citep{massoud_bma:2020,bhat_bma:2011}
              \item Importantly, we are not attempting to use historical data to weight models, but rather using transparent assumptions about the future to weight realizations of models.
          \end{enumerate}
    \item Our belief model $g : \mathcal{X} \rightarrow \mathcal{Z}$ is limited because $\mathcal{Z} \in \mathbb{R}$, but other measures beyond sea level in 2100 might be useful. For our approach to be valid, we implicitly assume that all candidate realizations are equally likely, conditional on the information contained in $\mathcal{Z}$. In practice we might wish to encode beliefs about future greenhouse forcing, temperature, sea level response to temperature, or other measures.
    \item The benefits of this approach:
          \begin{enumerate}
              \item A goal is to make -- intrinsically flawed and subjective -- modeling choices as transparent as possible. Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions
              \item Because of how we weight realizations of the future, you can apply a different weighting function and immediately assess performance. Thus qualitative and quantitative sensitivity analyses are simple.
              \item Of course the choice of $p(\mathcal{Z})$ is subjective, but so is every other aspect of how we frame the problem. Modeling is intrinsically subjective. The idea of embracing subjective choices -- and making them explicit and open to critique -- draws heavily from philosophies of iterative workflow in statistics \citep{box:1976,gelman_workflow:2020,gelman_philosophy:2013}.
              \item This approach can be used in a multiobjective context and is an alternative way to measure robustness in \gls{rdm} and \gls{mordm}
          \end{enumerate}
    \item This approach can be used in many other contexts
          \begin{enumerate}
              \item Stormwater management \citep{sharma_stormwater:2021,lopez-cantu:2018}
              \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}
          \end{enumerate}
\end{enumerate}

\section{End Matter}

\subsection{Acknowledgements}

\begin{enumerate}
    \item MARISA (details!) for JDG funding
    \item PSIRC (details!) for JDG funding
\end{enumerate}

\subsection{Author Contributions}

See journal requirements and format

\subsection{Supplementary Materials}

\begin{enumerate}
    \item Supplementary figures and analysis online
    \item Link to live repository on GitHub
    \item Link to code DOI on Zenodo
\end{enumerate}

\printbibliography
\end{document}

### 2021-04-13 tikz/conceptual.tex
add the paper

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}

\usepackage{physics}

\usepackage{xcolor}
\definecolor{TurquoiseGreen}{HTML}{B8D5B8}
\definecolor{TealBlue}{HTML}{598392}
\definecolor{YellowOrange}{HTML}{FFB140}
\definecolor{RichBlack}{HTML}{011627}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{function} = [circle, draw, fill=TurquoiseGreen, text=RichBlack, text width=4.5em, text centered]
\tikzstyle{block} = [rectangle, draw, fill=TurquoiseGreen, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=YellowOrange!33, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
\tikzstyle{check} = [rectangle, draw, fill=TealBlue!20, text=RichBlack, minimum width=4em, text centered,  minimum height=3em, align=center]

\tikzstyle{line} = [draw, very thick, color=RichBlack, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    \node [block] (decision) at (0, 2.5) {decision\\{\color{TealBlue}$x_i$}};
    \node [block] at (0, -2.5) (scenarios) {many scenarios\\{\color{TealBlue}$\qty{s_1, \ldots, s_n} \in \Omega$}};
    \node [function] (model) at (1, 0) {system model\\{\color{TealBlue}$f(x, s)$}};
    \node [derived] at (5, 1.25) (smetrics) {scenario metrics\\{\color{TealBlue}$u_{ij} = f(x_i, s_j)$}};
    \path [line] (decision) -- (model);
    \path [line] (scenarios) -- (model);
    \path [line] (model) to[out=22.5, in=180] (smetrics.west);

    \draw [color=TealBlue] (-2, -3.5) -- (6.75, -3.5) -- (6.75, 3.5) -- (-2, 3.5) -- (-2, -3.5);
    \node [rectangle, text=TealBlue, rotate=90] at (-1.75, 1) {Exploratory Modeling};

    \node [block] (priortop) at (10.3, -2.2) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
    \node [block] at (10.2, -2.3) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
    \node [block] at (10.1, -2.4) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
    \node [block] (prior) at (10, -2.5) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};

    \node [derived] (weights) at (9, 0) {weights\\{\color{TealBlue}eq.~2}};
    \path [line] (scenarios) to [out=0, in=180] (weights);
    \path [line] (prior) -- (weights);

    \node [derived] (combined) at (10, 2.5) {distribution of outcomes\\{\color{TealBlue}$\hat{p}(u | x_i)$}};
    \path[line] (smetrics.east) to [out=0, in=180] (combined);
    \path[line] (weights) -- (combined);

    \node [derived] (dmetrics) at (15, 2.5) {decision metrics\\{\color{TealBlue}$\hat{m}(x_i)$}};
    \path[line] (combined) -- (dmetrics);

    \draw [color=TealBlue] (7.2, -3.5) -- (16.5, -3.5) -- (16.5, 3.5) -- (7.25, 3.5) -- (7.25, -3.5);
    \node [rectangle, text=TealBlue, rotate=270] at (16.25, -1) {Probabilistic Modeling};

    \draw [color=TealBlue] (-2, -4) -- (16.5, -4) -- (16.5, -6.5) -- (-2, -6.5) -- (-2, -4);
    \node [rectangle, text=TealBlue, rotate=0, align=left] at (-1, -5.25) {Critique};
    \node [check] (discovery) at (3, -5.25) {scenario discovery};
    \node [check] (ppc) at (7, -5.25) {prior checks};
    \node [check] (posterior) at (11, -5.25) {posterior checks};
    \node [check] (compare) at (15, -5.25) {compare\\{\color{TealBlue}$\hat{m}(x_{\neq i})$}};
    \path [line, dashed, color=TealBlue] (prior) to[out=-157.5, in=67.5] (ppc);
    \path [line, dashed, color=TealBlue] (smetrics) to[out=-120, in=90] (discovery);
    \path [line, dashed, color=TealBlue] (combined.-22.5) to [out=-60, in=60] (posterior.60);
    \path [line, dashed, color=TealBlue] (dmetrics) -- (compare);

\end{tikzpicture}

\end{document}

### 2021-04-18 scripts/surge_prior_checks.jl
Can put the prior on GEV quantiles w/o doing the Stephenson math

@@ -37,7 +37,9 @@ The Jacobian for the prior
 
 See the unnumbered equation following equation 13.7 of
 
-Stephenson, A. (2015). Bayesian inference for extreme value modeling. In D. K. Dey & J. Yan, Extreme value modeling and risk analysis: methods and applications. Philadelphia, PA: CRC Press LLC.
+Stephenson, A. (2015). Bayesian inference for extreme value modeling.
+In D. K. Dey & J. Yan, Extreme value modeling and risk analysis:
+methods and applications. Philadelphia, PA: CRC Press LLC.
 """
 function jacobian_coles(θ, x)
     μ, σ, ξ = θ

### 2021-04-18 scripts/turing_example_prior.jl

@@ -1,0 +1,42 @@
+#=
+This script illustrates how to set a prior on quantiles of the outcome
+rather than on the parameters themselves. This is a very useful thing to
+do as it is often easier to elicit priors on the outcome variable than on
+the parameters themselves.
+
+See the following discussion with Tor Erlend Fjelde (https://github.com/torfjelde)
+https://discourse.julialang.org/t/turing-jl-prior-on-quantiles/59510/3
+=#
+
+using DrWatson
+@quickactivate "2021-elevation-robustness"
+
+using Turing, Distributions, DynamicHMC
+using DynamicPPL: @addlogprob!
+
+
+@model function Demo(y)
+
+    # define the parameters as vaguely as possible
+    μ ~ Flat()
+    σ ~ FlatPos(0.0)
+
+    # we're going to use this distribution repeatedly
+    dist = Normal(μ, σ)
+
+    # define quantiles -- these are deterministic
+    q50 = quantile(dist, 0.50)
+    q90 = quantile(dist, 0.90)
+
+    # prior distribution on quantiles - pretend this is "from experts"
+    DynamicPPL.@addlogprob! logpdf(Normal(3, 0.1), q50) # our prior on the median is about 5
+    DynamicPPL.@addlogprob! logpdf(Normal(10, 0.1), q90) # our prior on the 90th percentile is about 10
+
+    y .~ dist
+end
+
+y = [0]
+prior = sample(Demo(y), DynamicNUTS(), 100_000)
+yhat_prior = rand.(Normal.(prior[:μ], prior[:σ]))[:]
+@show quantile(yhat_prior, 0.50)
+@show quantile(yhat_prior, 0.90)

### 2021-10-12 papers/operationalizing-bayesian-checking.tex
Requesting feedback from Klaus

\documentclass[12pt]{article}

% Figures
\usepackage{graphicx}
\usepackage{wrapfig}
\graphicspath{{../plots/}{../tikz/}{../img/}}
\usepackage[section]{placeins} % require floats to appear in the section they are defined

% fonts and appearance
\usepackage{amsmath, amsfonts, physics, siunitx, fleqn, nicefrac}
\usepackage[american]{babel} 
\usepackage[T1]{fontenc} % improved font encoding
\usepackage[ttscale=0.8]{libertine}

% page size and margins
\usepackage{geometry}
\geometry{letterpaper,margin=1in}

% footer
\usepackage{fancyhdr}
\usepackage{lastpage}
\usepackage[en-US]{datetime2}
\fancyhf{}
\fancyhead[L]{DRAFT IN PROGRESS -- Doss-Gollin \& Keller}
\fancyhead[R]{\DTMnow}
\fancyfoot[R]{page~\thepage~of~\pageref{LastPage}}
\pagestyle{fancy}

% better tables
\usepackage{booktabs}
\usepackage{array}
\newcommand{\PreserveBackslash}[1]{\let\temp=\\#1\let\\=\temp}
\newcolumntype{C}[1]{>{\PreserveBackslash\centering}p{#1}}
\newcolumntype{R}[1]{>{\PreserveBackslash\raggedleft}p{#1}}
\newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}

% better lists
\usepackage[inline]{enumitem}
\setlist{nosep}

% authors
\usepackage{authblk}
\title{Which Scenario Should We Design For? Insights from House Elevation on the Multiple PDF Problem}
\author[1]{James Doss-Gollin}
\author[2,3,4]{Klaus Keller}
\affil[1]{Department of Civil and Environmental Engineering, Rice University}
\affil[2]{Department of Geosciences, the Pennsylvania State University}
\affil[3]{Earth and Environmental Systems Institute, the Pennsylvania State University}
\affil[4]{Dartmouth?}
\renewcommand*{\Affilfont}{\normalsize\normalfont}

% ACRONYMS
\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
\newacronym{bdt}{BDT}{Bayesian decision theory}
\newacronym{bfe}{BFE}{base flood elevation}
\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
\newacronym{gcm}{GCM}{general circulation model}
\newacronym{lsl}{LSL}{local mean sea level}
\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
\newacronym{mordm}{MORDM}{multiobjective \gls{rdm}}
\newacronym{pdf}{PDF}{probability density function}
\newacronym{rcp}{RCP}{representative concentration pathway}
\newacronym{rslr}{RSLR}{relative sea level rise}
\newacronym{rdm}{RDM}{robust decision making}
\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
\newacronym{tc}{TC}{tropical cyclone}

\usepackage{xspace}
\makeatletter
\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
\newcommand{\usd}[1]{\SI{#1}[\$]{}}
\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}

% use biblatex
\usepackage{csquotes}
\usepackage[
  backend=biber,
  doi=true,
  url=false,
  isbn=false,
  style=authoryear-comp,
  natbib=true,
  backref=false,
  maxbibnames=10,
  maxcitenames=2,
  uniquename=false,
  uniquelist=false,
  sorting=nyt,
]{biblatex}
\renewbibmacro{in:}{}
\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
\addbibresource{../_bibliography/library.bib}

% load this last
\usepackage[hidelinks]{hyperref}
\usepackage{cleveref}

% up to 1250 words
\begin{document}
\maketitle
\thispagestyle{empty}

\begin{abstract}
    Many homeowners in the coastal zone elevate their houses to manage flood risk.
    Federal guidance for this decision relies on floodplain maps that are silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards.
    This uncertainty poses challenges for the design of decision-support systems.

    Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
    Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
    We introduce an approach to bridge this divide through a computationally efficient method that weights each state of the world based on a probability distribution over possible futures.
    Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
    This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.
\end{abstract}

\subsection*{Submission Plan}
We should revisit this!
\begin{enumerate}
    \item Climatic Change
    \item IOP Environmental Research: Infrastructure and Sustainability (new journal)
    \item Earth's Future
    \item Risk analysis
\end{enumerate}
\subsection*{Internal peer review}
\begin{enumerate}
    \item TBD
\end{enumerate}

\clearpage
\section{Introduction}

\paragraph{Why this problem is important}
Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face far into the future.
On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_codes_txtreme:2021}.
On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
For example, (\texttt{megaproject example})\ldots.

\paragraph{Present the Multiple PDF Problem}
While classical utility- or regret-based frameworks offer a structure for weighing these risks, they are not designed to handle uncertainties that are \emph{deep} in the sense that stakeholders may not agree upon their probabilities \citep{lempert_complex:2002} or \emph{irreducible} in the sense that there is not an objective truth that better models and data can identify \citep{DossGollin:2019}.\footnote{Unless we believe everything is pre-determined, in which case why bother with climate adaptation\ldots}
Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S. federal government will continue to subsidize flood insurance for homeowners.
In response, a wide range of fields have developed scenario-based methods\ldots.
This leads to the ``Multiple PDF Problem'' \citep{sharma_rcp:2021}\ldots.
A key question, then, facing the designers, builders, regulators, owners, and users of built facilities is \emph{to which scenario should infrastructure be designed?}

\paragraph{Set up house elevation as an example}
One example is building codes and guidelines\ldots.
Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
\citet{zarekarizi_suboptimal:2020} showed that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.

\paragraph{Nonstationarity motivates the case study}
Nonstationarity is particularly relevant in coastal communities where \gls{rslr} is expected to drive future flood risk.
As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch}, given stochastic flood peaks and deeply uncertain \gls{rslr}.
\Gls{rslr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
Two key uncertainties identified by previous studies are
\begin{enumerate*}[label=(\roman*)]
    \item which emissions pathway (\eg, \gls{rcp} scenario) will we experience and
    \item how will polar ice sheets respond to warming \citep{wong_brick0.2:2017,ruckert_coastal:2019,wong_nola:2017,deconto_antarctica:2016}.
\end{enumerate*}
Simulations of \gls{rslr} are available to decision makers grouped by \gls{rcp} scenario and model structure.
For a given combination of \gls{rcp} scenario and model, results are typically presented and interpreted as probabilistic, but this is not the same as giving decision-makers a single \gls{pdf}.
This is the ``Multiple \gls{pdf} Problem'' described in, \eg, \citet{sharma_rcp:2021}.
Thus, \emph{which scenario/\gls{pdf} should we design for?}

\paragraph{Needs work}
Lacking a crystal ball capable of telling us precisely which \gls{pdf} we need to design for, we have to consider several as plausible.
Any decision will perform differently in different scenarios.
There is no objective answer to this question.
Since we can't be ojectively correct, we might as well be transparent and make our assumptions explicit (and subject to critique) rather than implicit.

\paragraph{Roadmap}
We first present a general framework for\ldots.
We then illustrate this method through a didactic case study of deciding if, and how high, to elevate a single house in Norfolk, VA.

\begin{figure}
    \centering
    \includegraphics[height=2in]{uconn-elevateA}
    \caption{
        Conceptual sketch illustrating the path from historical data to FEMA's guidance.
        This figure SHOULD BE MODIFIED to have three panels, of which this is the third.
        (a): A time series of historical floods.
        (b): A \gls{pdf} of the flood peak.
        (c): as shown, but with waves eliminated.
        This sketch figure is from \citet{patten_elevate:2016}.
    }\label{fig:house-sketch}
\end{figure}

\section{Decision framework}\label{sec:framework}

In this section we present a general framework for decision making under deep uncertainty.
We begin with terms and definitions that are broadly consistent with the rest of the literature.
In \cref{sec:methods-weight} we present a Bayesian model for weighting \glspl{sow}.
In \cref{sec:methods-iterative} we discuss how this framework can be integrated into frameworks for iterative engagement with experts and stakeholders to expand and frame the model.

\begin{figure}
    \centering
    \includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
    \caption{
        Schematic of our decision support framework.
        Nodes in green are modeling choices that the modeler must make while those in yellow are computed quantities.
        Teal text illustrates a non-exhaustive set of opportunities to perform model checks and/or engage with stakeholders to iterate on design of the decision support system.
        Feedbacks between these questions and the green nodes are omitted for clarity.
        Our approach could be extended by coupling the evaluation framework to an optimization or policy search tool that updates the candidate decisions.
        This approach could also be extended to a stochastic system model.
    }\label{fig:framework}
\end{figure}

We start by modeling the performance of a given decision $x \in \mathcal{X}$.
To do this, we will use a system model $f: \mathcal{X}, \Omega \rightarrow \mathcal{U}$ which takes as inputs the decision $x_i$ and a state of the world $s$ and returns one or more metrics of interest, $u$ \citep[we generally follow the notation of][]{mcphail_robustness:2019}.
We assume $f$ to be deterministic; if $f$ is a stochastic function we can use a technique like Monte Carlo sampling to estimate expectated values of the metrics, \ie $u_{ij} = \mathbb{E}\qty[u | x=x_i, s=s_j]$ (in the limit where the expectation can be calculated exactly, $f$ behaves like a deterministic function).

The literature on decision making under deep uncertainty \citep[see][for a review]{marchau:2019} emphasizes the importance of using system models like $f$ in an exploratory context by sampling widely from plausible futures $s \in \Omega$ and studying their implications \citep{bankes:1993,lempert_shaping:2003,kwakkel:2019,lamontagne_discovery:2018}.
This is often described as ``bottom-up'' modeling \citep{Brown:2012kb} and has parallels to model-based stress tests and vulnerability assessments.

Given our system model $f$ and a sampling of $J$ possible states of the world $s_1, s_2, \ldots, s_J$, we can calculate a set of \emph{\gls{sow} metrics} $u_{ij}$ that tell us ``how will decision $x_i$ perform under state of the world $s_j$'' for many $s_j$, and one or more $x_i$.
For an apples to apples comparison of how different decisions perform on the $j$th \gls{sow}, we can directly compare $u_{1j}, u_{2j}, \ldots, u_{Ij}$.
However, if we wish to compare the performance of two or more decisions across \emph{all} scenarios, we will need some way of aggregating the \gls{sow} metrics into \emph{decision metrics} $m_1, m_2, \ldots, m_I$ such that for every $x_i$ there is a single (scalar or vector) $m_i$.\footnote{I think the link back to our original motivation -- the multiple PDF problem}

\citet{mcphail_robustness:2019} split the problem of aggregating \gls{sow} metrics into decision metrics into two steps: ``subsetting'' (\eg, single value, all values, subset of values) and ``metric calculation'' (\eg, mean, variance, skew, kurtosis).
We generalize this to
\begin{equation}
    m_i = g(\qty{u_{i1}, u_{i2}, \ldots, u_{iJ}}, \qty{w_1, w_2, \ldots, w_J}),
\end{equation}
where $g$ is an \emph{aggregating function} describing the ``metric calculation'' step and the $w_j$ describe weights given to each \gls{sow}.
A optional, but likely useful, interpretation of the weighted samples $\qty{w_j, u_{ij} \forall j=1, \ldots, J}$ is an empirical \gls{pdf}, in which case $g$ maps a \gls{pdf} of future outcomes to the metrics $m_i$.\footnote{perhaps this belongs in the Discussion}
This is a flexible framework; \cref{tab:literature-weighting} shows some methods for robust decision making from the literature using our notation.

\begin{table}
    \centering
    \footnotesize
    \begin{tabular}{llll}
        \toprule
        Citation        & Explanation                     & $w_j$                                                                             & $g(u, w)$          \\
        \midrule
        Many            & All $s_j$ are equally plausible & $w_j = \nicefrac{1}{J}$                                                           & any                \\
        Many            & Worst case                      & $w_j = \begin{cases}1 & u_{ij} = \min_{j'} u_{ij'} \\ 0 & \text{else}\end{cases}$ & $u$                \\
        TBD             & Average over \glspl{sow}        & $w_j = \nicefrac{1}{J}$                                                           & $\sum_j w_j u_{j}$ \\
        Characklis/Reed & 95th percentile                 & $w_j = \nicefrac{1}{J}$                                                           & weighted quantile  \\
        \bottomrule
    \end{tabular}
    \caption{
        A non-exhaustive summary of methods for setting $w_j$ and $g$ from the literature.
        (\emph{What other examples should we add?})
    }\label{tab:literature-weighting}
\end{table}

In the next section we propose an alternative scheme for weighting scenarios.

\paragraph{To think  about}
I have two flavors of this idea presented here!
\Cref{fig:framework} takes a more Bayesian focus: the goal is to reason about properties of the \emph{distribution} over outcomes $p(u | x=x_i)$.
The text follows \citet{mcphail_robustness:2019} more.
I'm not sure which is easier to understand.

\subsection{Weighting \glsentryplural{sow}}\label{sec:methods-weight}

Settinget all $w_j \propto 1$ as is often done (\cref{tab:literature-weighting}) reveals a tension: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision.

Some aims:
\begin{enumerate}
    \item We should be invariant to the parameterization of $s$ -- if we sample population uniformly from 1 million to 10 million or $\log_{10}$ population uniformly from \num6 to 7, we should get the same answer
    \item We should be invariant to the sampling of $s$:
          \begin{enumerate}
              \item if we include scientifically interesting but extremely low-probability scenarios (\eg, if someone created \gls{rcp} 45), our esimtates of $m_i$ should not change meaningfully
              \item if we include nearly redundant scenarios (\eg, if we are studying \gls{rcp} 2.6, 4.5, 6.0, and 8.5 and we then add results from a \gls{rcp} 6.01 experiment), our estimates of $m_i$ should not change meaningfully
          \end{enumerate}
    \item Our assumptions should be transparent and open to critique
\end{enumerate}
We can do this with probability if we have a target distribution $p(s)$.
We return to the question of how to choose this in \cref{sec:methods-prior}.

\subsection{Melding prior over \glsentryplural{sow}}\label{sec:methods-prior}

Often (including in the case study described in \cref{sec:case}) our \glspl{sow} are the outputs of complicated model chains.
To set our prior $p(s)$ we might like to incorporate prior external knowledge about both the

\subsection{Opportunities for iterative model expansion}\label{sec:methods-iterative}

\paragraph{Robustness checks}
The computationally expensive piece is computing the $u_{ij}$.
Once you do that, you can look at several $p(\Omega)$ and compute the $m_i$ multiple times.

\paragraph{Links to DMDU Methods}
You can use scenario discovery on the $(x_i, \mu_{ij}$).
(Others to note?)

\paragraph{Bayesian model checks}
Let's just cite \citet{gelman_workflow:2020} and \citet{gabry_visualization:2019} and point out that you can and should conduct sample-based checks on your prior $p(\Omega)$ and your posterior $p(u | x=x_i)$.\footnote{See discussion above about whether we want to emphasize that we are treating this as a posterior or not}

\section{Case study: house elevation}\label{sec:case}

We start with a bare-bones version of the \citet{zarekarizi_suboptimal:2020} model.
For interpretability, we fix the discount rate, house lifetime, and depth-damage curve to illustrate how our framework can be used to model nonstationary flood risk.
Real-world decision makers should consider additional uncertainties and stakeholder-identified metrics.

\begin{figure}
    \centering
    \includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
    \caption{
        XLRM diagram.
        Red nodes represent uncertainties ($X$).
        Yellow node represents the lever ($L$).
        The gray node shows system relationships ($R$).
        Blue nodes show metrics ($M$).
    }\label{fig:xlrm}
\end{figure}

\subsection{Scenarios}

Coastal floods are often decomposed into changes in local sea level \citep{kopp_evolving:2017,kopp_probabilistic:2014,wong_brick0.2:2017} plus storm surge \citep{garner_slrise:2018,cagigal_emulator:2020,rueda_surge:2016}.
We neglect waves and hydrodynamics because (FILL IN)\ldots.

\gls{rslr}:
\begin{enumerate}
    \item We have 2 models of ice sheet dynamics (3? Vivek and Tony are trying to tell me what models the data from \citet{ruckert_coastal:2019} actually comes from)
    \item We have 4 \gls{rcp} scenario
    \item We use the BRICK \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019} to define $J$ \glspl{sow}.
\end{enumerate}
Storm surge:
\begin{enumerate}
    \item We have a $N$ year observational record
    \item We separate out storm surge from the background trend in some standard way
    \item We model storm surge as a stationary GEV model
    \item We apply clever priors: we place a prior on the return level (data space) rather than a prior on the parameters themselves (except the shape parameter -- we put a restrictive prior for it to be positive)\footnote{arguably we should stick to boring priors here -- the question being whether it's better to say we use iterative checks to improve the priors, or whether we want to keep things simple}
    \item We draw $K$ sets of parameters $\qty(\mu_k, \sigma_k, \xi_k)$
\end{enumerate}
Thus a \gls{sow} is defined just an 81 year time series of \gls{lsl}.
We will see what we do with the $K$ draws of storm surge parameters shortly.
\begin{figure}
    \caption{
        INSERT HERE: figure showing time series of observed surge and \gls{lsl} with notable storms indicated, as well as projections of future \gls{lsl} under all eight models (two BRICK, four \gls{rcp})
    }
\end{figure}

\subsection{Decisions}

We consider only one lever: how high to elevate the house today.
This is the same lever considered in \citet{zarekarizi_suboptimal:2020}

\subsection{Metrics}

Three scenario metrics $u$:
\begin{enumerate}
    \item Construction cost
    \item Discounted expected flood damages
    \item Total cost (construction plus flood)
\end{enumerate}
The decision metrics $m$ are just the expected value of $u$.

\subsection{System model}

\begin{enumerate}
    \item Costs of elevating the house are the same as \citet{zarekarizi_suboptimal:2020}
    \item Depth-damage: use HAZUS model from \citet{zarekarizi_suboptimal:2020}
    \item Discount: fix at some appropriate rate (2\%?)
    \item House lifetime: fix at some reasonable amount (80 years)
\end{enumerate}

Recall that we have $K$ simulations of the storm surge parameters.
For each year of each \gls{sow} we would like to estimate the expected value of the convolution of hazard (storm surge plus \gls{lsl}) with fragility (depth-damage).
A simple algorithm to estimate this is, for each \gls{pdf} of storm surge $k \in 1, \ldots, K$, to draw $N$ Monte Carlo sample values of storm surge, calculate the damage associated with each, and take the expectation with respect to $N$.
Then, take the expectation again with respect to the $K$ storm surge \glspl{pdf}.
Since this is computationally expensive, we build an emulator that estimates expected damage as a function of mean sea level only by fitting a spline to the result of systematically calculating expected damages for different values of \gls{lsl}.\footnote{A supplemental figure would be good here. This is probably a bit dense of a discussion.}

\begin{figure}
    \caption{
        INSERT HERE: figure showing the convolution of storm surge hazard through the depth-damage function.
        We fix \gls{lsl} constant but have many \glspl{pdf} of \gls{lsl}.
        This is the ``hazard, damage, risk'' graph that Klaus has sketched.
    }
\end{figure}

\subsection{Belief model}

\paragraph{Prior on outputs only}
Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{rslr} in 2100.

\paragraph{Prior on inputs only}
Alternatively, we can place a prior over scenarios.
Let's say that we think that emissions forcing in 2100 is has some distribution.
This means we are placing a weight over \gls{rcp} scenarios.
We can assume both BRICK models are equally likely.

\paragraph{Melding}
We can do cool stuff with Bayesian melding \citep{poole_melding:2000,sevcikova_melding:2007}.
This combines the prior over inputs and outputs geometrically.
\begin{equation}\label{eq:melding}
    \text{melding equation here}
\end{equation}
This lets us express information about some inputs being more or less likely than others, and also some outputs being more or less likely than others.
THIS IS USEFUL because often we would like to incorporate relevant knowledge from many different sources on both inputs and outputs (\eg, IPCC reports, National Climate Assessments, \etc).

\paragraph{Implementation}
Code will be available on GitHub.
We use the Julia programming language.

\section{Results}\label{sec:results}

We will need a figure or two showing our prior and posterior checks, as well as some results.
Results should show the tradeoff between cost and damage under different assumptions.

\section{Discussion}

Most fundamentally, we are thinking about robustness a bit differently than usual.
The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.
We are thinking about robustness to different modeling structures and assumptions.
This has links to the ``rival framings'' literature that Pat has been working on.
Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).

\begin{enumerate}
    \item Links to DMDU literature
          \begin{enumerate}
              \item We should also cite \citet{quinn_exploratory:2020}, which makes the point that it matters how you sample \glspl{sow}.
              \item Exploratory modeling can demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
              \item You can do scenario discovery
              \item We are in effect calculating robustness metrics \citep{mcphail_robustness:2019,herman:2015}
          \end{enumerate}
    \item Philosophy of Bayesian stats / decision making
          \begin{enumerate}
              \item The observation that many of these mechanisms cannot be represented by a single objective \gls{pdf} has motivated many criticisms of the application of \gls{bdt} to planning problems.
                    Yet \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
                    \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
                    The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
                    More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
                    Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
                    Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
                    Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
                    More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
                    Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}\footnote{this paragraph is copied from my dissertation, needs to be re-worked!}
          \end{enumerate}
    \item There are interesting links to Bayesian model averaging
          \begin{enumerate}
              \item Lots of examples to cite if we want
              \item I'm partial to something like stacking \citep{Yao:2018bu}
              \item The main difference is we are using only the prior to average the models!
          \end{enumerate}
    \item Limitations of the case study
          \begin{enumerate}
              \item Objectives: real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house, and more
              \item More uncertainty (damage functions, cost of construction, lifespan, discount rate, \etc)
              \item Better data on cost of elevation and depth-damage
              \item Robustness: ptimize separately for different kinds of house structures and locations
              \item Timing
          \end{enumerate}
    \item Limitations of the method
          \begin{enumerate}
              \item Our prior $p(s)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
              \item We neglect true ignorance \citep{knight_risk:1921}.
              \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
          \end{enumerate}
    \item The benefits of this approach:
          \begin{enumerate}
              \item A goal is to make -- intrinsically flawed and subjective -- modeling choices as transparent as possible. Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions
              \item Because of how we weight realizations of the future, you can apply a different weighting function and immediately assess performance. Thus qualitative and quantitative sensitivity analyses are simple.
              \item Of course the choice of $p(\mathcal{Z})$ is subjective, but so is every other aspect of how we frame the problem. Modeling is intrinsically subjective. The idea of embracing subjective choices -- and making them explicit and open to critique -- draws heavily from philosophies of iterative workflow in statistics \citep{box:1976,gelman_workflow:2020,gelman_philosophy:2013}.
              \item This approach can be used in a multiobjective context and is an alternative way to measure robustness in \gls{rdm} and \gls{mordm}
          \end{enumerate}
    \item This approach can be used in many other contexts
          \begin{enumerate}
              \item Direct
                    \begin{enumerate}
                        \item Stormwater management \citep{sharma_stormwater:2021,lopez-cantu:2018}
                        \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}
                    \end{enumerate}
              \item Indirect
                    \begin{enumerate}
                        \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!
                        \item Model structure uncertainties
                    \end{enumerate}
          \end{enumerate}
\end{enumerate}.

\section{Conclusions}

\begin{enumerate}
    \item We started with the problem of ``which scenario should we design to?'' We then devised a cop-out: design to a weighted combination of all the scenarios. (So how should we wight the scenarios? Come up with a model $p(\Omega)$ that tells you.)
    \item There is no objective way to choose $p(\Omega)$ because we don't have data on the future!
    \item We can't get around from choosing $p(\Omega)$ implicitly\footnote{I'm not sure we have demonstrated this rigorously, or whether we want to argue it, but I think there's a way to frame this point that is valid and helpful}
    \item Ultimately, engineering design needs to consider not only how to weight different future scenarios, but also which metrics are important, how risk aversion should be addressed, \etc.
    \item We used a didactic case study to present a toolkit (decision framework, melding priors, prior over return levels) for making transparent and critique-friendly assessments.
\end{enumerate}

\section*{End Matter}

\subsection*{Acknowledgements}

\begin{enumerate}
    \item MARISA (details!) for JDG funding
    \item PSIRC (details!) for JDG funding
    \item Rice University (do I need to thank startup?)
    \item Tor Erlend Fjelde for helpful advice on implementing the GEV prior using \texttt{Turing.jl}.
\end{enumerate}

\subsection*{Author Contributions}

See journal requirements and format

\subsection*{Supplementary Materials}

\begin{enumerate}
    \item Supplementary figures and analysis online
    \item Link to live repository on GitHub
    \item Link to code DOI on Zenodo
\end{enumerate}

\printbibliography
\end{document}

### 2021-10-12 tikz/conceptual.tex
Requesting feedback from Klaus

@@
 
-\tikzstyle{function} = [circle, draw, fill=TurquoiseGreen, text=RichBlack, text width=4.5em, text centered]
+\tikzstyle{function} = [circle, draw, fill=TurquoiseGreen, text=RichBlack, text width=5.5em, text centered]
 \tikzstyle{block} = [rectangle, draw, fill=TurquoiseGreen, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
 \tikzstyle{derived} = [rectangle, draw, rounded corners, fill=YellowOrange!33, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
-\tikzstyle{check} = [rectangle, draw, fill=TealBlue!20, text=RichBlack, minimum width=4em, text centered,  minimum height=3em, align=center]
+\tikzstyle{check} = [text=TealBlue, align=center]
 
@@
 
-    \node [block] (decision) at (0, 2.5) {decision\\{\color{TealBlue}$x_i$}};
-    \node [block] at (0, -2.5) (scenarios) {many scenarios\\{\color{TealBlue}$\qty{s_1, \ldots, s_n} \in \Omega$}};
-    \node [function] (model) at (1, 0) {system model\\{\color{TealBlue}$f(x, s)$}};
-    \node [derived] at (5, 1.25) (smetrics) {scenario metrics\\{\color{TealBlue}$u_{ij} = f(x_i, s_j)$}};
+    \node [block] (decision) at (4, 3) {Candidate decisions\\{\color{TealBlue}$\qty{x_1, x_2, \ldots, x_I}$}};
+    \node [block] at (-4, 3) (scenarios) {Many possible futures\\{\color{TealBlue}$\qty{s_1, \ldots, s_J} \in \Omega$}};
+    \node [function] (model) at (0, 0) {Deterministic system model\\{\color{TealBlue}$f(x, s)$}};
+    \node [derived] at (0, -3) (smetrics) {Scenario metrics\\{\color{TealBlue}$u_{ij} = f(x_i, s_j)$}};
+    \node [block] (prior) at (-4, -6) {Prior over scenarios\\{\color{TealBlue}$p(s)$}};\
+    \node [derived] (weights) at (0, -6) {Scenario weights\\{\color{TealBlue}eq.~XX}};
+    \node [derived] (combined) at (4, -6) {Distribution over\\outcomes\\{\color{TealBlue}$\hat{p}(u | x_i)$}};
+    \node [derived] (dmetrics) at (8, -6) {Decision metrics\\{\color{TealBlue}$\hat{m}(x_i)$}};
+
     \path [line] (decision) -- (model);
     \path [line] (scenarios) -- (model);
-    \path [line] (model) to[out=22.5, in=180] (smetrics.west);
+    \path [line] (model) to (smetrics);
+    \path [line] (scenarios.south) to[out=-90, in=135] (weights.north west);
+    \path [line] (prior) -- (weights);
+    \path [line] (smetrics) to (combined);
+    \path [line] (weights) -- (combined);
+    \path [line] (combined) to (dmetrics);
 
-    \draw [color=TealBlue] (-2, -3.5) -- (6.75, -3.5) -- (6.75, 3.5) -- (-2, 3.5) -- (-2, -3.5);
-    \node [rectangle, text=TealBlue, rotate=90] at (-1.75, 1) {Exploratory Modeling};
-
-    \node [block] (priortop) at (10.3, -2.2) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
-    \node [block] at (10.2, -2.3) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
-    \node [block] at (10.1, -2.4) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
-    \node [block] (prior) at (10, -2.5) {subjective belief\\{\color{TealBlue}$p_\text{belief}(x)$}};
-
-    \node [derived] (weights) at (9, 0) {weights\\{\color{TealBlue}eq.~2}};
-    \path [line] (scenarios) to [out=0, in=180] (weights);
-    \path [line] (prior) -- (weights);
-
-    \node [derived] (combined) at (10, 2.5) {distribution of outcomes\\{\color{TealBlue}$\hat{p}(u | x_i)$}};
-    \path[line] (smetrics.east) to [out=0, in=180] (combined);
-    \path[line] (weights) -- (combined);
-
-    \node [derived] (dmetrics) at (15, 2.5) {decision metrics\\{\color{TealBlue}$\hat{m}(x_i)$}};
-    \path[line] (combined) -- (dmetrics);
-
-    \draw [color=TealBlue] (7.2, -3.5) -- (16.5, -3.5) -- (16.5, 3.5) -- (7.25, 3.5) -- (7.25, -3.5);
-    \node [rectangle, text=TealBlue, rotate=270] at (16.25, -1) {Probabilistic Modeling};
-
-    \draw [color=TealBlue] (-2, -4) -- (16.5, -4) -- (16.5, -6.5) -- (-2, -6.5) -- (-2, -4);
-    \node [rectangle, text=TealBlue, rotate=0, align=left] at (-1, -5.25) {Critique};
-    \node [check] (discovery) at (3, -5.25) {scenario discovery};
-    \node [check] (ppc) at (7, -5.25) {prior checks};
-    \node [check] (posterior) at (11, -5.25) {posterior checks};
-    \node [check] (compare) at (15, -5.25) {compare\\{\color{TealBlue}$\hat{m}(x_{\neq i})$}};
-    \path [line, dashed, color=TealBlue] (prior) to[out=-157.5, in=67.5] (ppc);
-    \path [line, dashed, color=TealBlue] (smetrics) to[out=-120, in=90] (discovery);
-    \path [line, dashed, color=TealBlue] (combined.-22.5) to [out=-60, in=60] (posterior.60);
-    \path [line, dashed, color=TealBlue] (dmetrics) -- (compare);
+    \node [check] at (-9, 3) (span) {Do scenarios span\\plausible outcomes?};
+    \path [line, dashed] (scenarios) -- (span);
+    \node [check] at (9, 3) (represented) {Are relevant levers\\represented?};
+    \path [line, dashed] (decision) -- (represented);
+    \node [check] at (5, 0) (interpretable) {Is the model credible?\\Interpretable?};
+    \path [line, dashed] (model) -- (interpretable);
+    \node [check] at (5, -3) (discovery) {What are relevant regions\\of the $\qty(s, x)$ space?};
+    \path [line, dashed] (smetrics) -- (discovery);
+    \node [check] at (-4, -8) (priorcheck) {Are these consistent\\with prior knowledge?};
+    \path [line, dashed] (prior) -- (priorcheck);
+    \node [check] at (0, -8) (weightcheck) {Are computational\\issues apparent?};
+    \path [line, dashed] (weights) -- (weightcheck);
+    \node [check] at (4, -8) (ppc) {Are these consistent\\with our expectations?};
+    \path [line, dashed] (combined) -- (ppc);
 

### 2021-10-12 tikz/xlrm.tex
Requesting feedback from Klaus

\documentclass[tikz, border=1pt]{standalone}

\usepackage{tikz}
\usepackage[ttscale=0.8]{libertine}
\usepackage{physics}

\usepackage{xcolor}
\definecolor{TurquoiseGreen}{HTML}{B8D5B8}
\definecolor{TealBlue}{HTML}{598392}
\definecolor{YellowOrange}{HTML}{FFB140}
\definecolor{RichBlack}{HTML}{011627}

\usetikzlibrary{shapes,arrows,positioning,calc,decorations.text}
\begin{document}
\pagestyle{empty}

\tikzstyle{uncertainty} = [rectangle, draw, fill=red!30, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{lever} = [rectangle, draw, fill=yellow!20, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{relation} = [rectangle, draw, fill=gray!20, text=RichBlack, text width=6.5em, text centered]
\tikzstyle{metric} = [rectangle, draw, fill=blue!20, text=RichBlack, text width=6.5em, text centered]

\tikzstyle{line} = [draw, very thick, color=RichBlack, ->]
\tikzset{>=latex}

\def\myshift#1{\raisebox{-2.5ex}}

\begin{tikzpicture}[auto]

    \node [relation] (relationships) at (0, 0) {System\\Relationships\\(R)};
    \node [uncertainty] (surge) at (-3.5, 2.5) {Storm Surge\\{\footnotesize ($J$ draws from 1 posterior)}};
    \path [line] (surge) to (relationships);
    \node [uncertainty] (slr) at (-3.5, -1) {Sea Level Rise\\{\footnotesize ($J$ Simulations)}};
    \path [line] (slr) to (relationships);
    \node [lever] (elevate) at (0, 2.5) {How high to elevate?};
    \path [line] (elevate) to (relationships);
    \node [metric] (ccost) at (3.5, 1.25) {Construction Costs};
    \path [line] (relationships) to (ccost);
    \node [metric] (fcost) at (3.5, -1.25) {Flood Damages};
    \path [line] (relationships) to (fcost);

    \node[uncertainty] (rcp) at (-4.5, -2.5) {RCP Scenario\\{\footnotesize (4 Scenarios)}};
    \node[uncertainty] (ice) at (-4.5, 0.75) {Ice Sheet Dynamics\\{\footnotesize (3 Models)}};
    \path [line] (rcp) to (slr);
    \path [line] (ice) to (slr);

\end{tikzpicture}

\end{document}

### 2021-10-13 papers/operationalizing-bayesian-checking.tex
Updates from Overleaf

@@
 \affil[3]{Earth and Environmental Systems Institute, the Pennsylvania State University}
-\affil[4]{Dartmouth?}
+\affil[4]{Thayer School of Engineering, Dartmouth College}
 \renewcommand*{\Affilfont}{\normalsize\normalfont}
@@
 \AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
-\addbibresource{../_bibliography/library.bib}
+%\addbibresource{../_bibliography/library.bib}
+\addbibresource{library.bib}
 
@@
 \begin{enumerate}
+    \item Earth's Future
     \item Climatic Change
     \item IOP Environmental Research: Infrastructure and Sustainability (new journal)
-    \item Earth's Future
     \item Risk analysis
@@
 On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_codes_txtreme:2021}.
+For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021}.
 On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
@@
                     \begin{enumerate}
-                        \item Stormwater management \citep{sharma_stormwater:2021,lopez-cantu:2018}
+                        \item Stormwater management \citep{sharma_rcp:2021,lopez-cantu:2018}
                         \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}

### 2021-10-13 papers/operationalizing-bayesian-checking.tex
add some notes highlighting Klaus's comments

@@
 \usepackage{geometry}
-\geometry{letterpaper,margin=1in}
+\geometry{letterpaper,top=1in, bottom=1in, left=0.5in, right=2in}
 
@@
 \pagestyle{fancy}
+
+% TO DO NOTES
+\usepackage{xcolor} % list of colors at https://en.wikibooks.org/wiki/LaTeX/Colors
+\definecolor{giallo}{HTML}{F0BC42} % https://teamcolorcodes.com/a-s-roma-color-codes/
+\definecolor{rosso}{HTML}{8E1F2F}
+\definecolor{grigio}{HTML}{CACACC}
+\definecolor{nero}{HTML}{000000}
+\usepackage[textsize=scriptsize]{todonotes}
+\setlength{\marginparwidth}{1.75in}
+\newcommand{\james}[1]{\todo[color=giallo, textcolor=nero]{\textbf{ATTN James:~}#1}} % if desired create a custom command for each author
+\newcommand{\klaus}[1]{\todo[color=rosso, textcolor=grigio]{\textbf{ATTN Klaus:~}#1}}
 
@@
     Federal guidance for this decision relies on floodplain maps that are silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards.
-    This uncertainty poses challenges for the design of decision-support systems.
+    This uncertainty poses challenges for the design of decision-support systems.\james{If this is about RDM then let's lead with that.}
 
@@
 \end{abstract}
+
+\section*{Front Matter}
 
@@
 \begin{enumerate}
-    \item TBD
+    \item Vivek
+    \item Courtney
+    \item Sanjib
+    \item Kelsey to replicate code
 \end{enumerate}
@@
 On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021}.
+For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
 On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
-For example, (\texttt{megaproject example})\ldots.
+For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}.}
 
-\paragraph{Present the Multiple PDF Problem}
+\paragraph{Set up the ``Multiple PDF Problem''}
 While classical utility- or regret-based frameworks offer a structure for weighing these risks, they are not designed to handle uncertainties that are \emph{deep} in the sense that stakeholders may not agree upon their probabilities \citep{lempert_complex:2002} or \emph{irreducible} in the sense that there is not an objective truth that better models and data can identify \citep{DossGollin:2019}.\footnote{Unless we believe everything is pre-determined, in which case why bother with climate adaptation\ldots}
-Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S. federal government will continue to subsidize flood insurance for homeowners.
+Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S.
+federal government will continue to subsidize flood insurance for homeowners\james{Add some references -- Klaus suggests \citet{savage:1954} or similar}.
 In response, a wide range of fields have developed scenario-based methods\ldots.
@@
 For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
-\citet{zarekarizi_suboptimal:2020} showed that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
+\citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
 Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
@@
 Nonstationarity is particularly relevant in coastal communities where \gls{rslr} is expected to drive future flood risk.
-As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch}, given stochastic flood peaks and deeply uncertain \gls{rslr}.
-\Gls{rslr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
+As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch},\james{Need to revisit this picture} given stochastic flood peaks and deeply uncertain \gls{rslr}.
+\Gls{rslr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots\klaus{You wrote something here that I can't read -- I see something that looks like ICM.
+I assume Tony and Kelsey papers are relevant as well?}) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
 Two key uncertainties identified by previous studies are
@@
         Conceptual sketch illustrating the path from historical data to FEMA's guidance.
-        This figure SHOULD BE MODIFIED to have three panels, of which this is the third.
+        This figure \emph{needs to be revised} (KATERINA?) to have three panels, of which this is the third.
         (a): A time series of historical floods.
@@
 In this section we present a general framework for decision making under deep uncertainty.
-We begin with terms and definitions that are broadly consistent with the rest of the literature.
+We begin with terms and definitions that are broadly consistent with common adaptive frameworks\james{add refs}\klaus{did I read this right?}.
 In \cref{sec:methods-weight} we present a Bayesian model for weighting \glspl{sow}.
@@
 
-We start by modeling the performance of a given decision $x \in \mathcal{X}$.
-To do this, we will use a system model $f: \mathcal{X}, \Omega \rightarrow \mathcal{U}$ which takes as inputs the decision $x_i$ and a state of the world $s$ and returns one or more metrics of interest, $u$ \citep[we generally follow the notation of][]{mcphail_robustness:2019}.
-We assume $f$ to be deterministic; if $f$ is a stochastic function we can use a technique like Monte Carlo sampling to estimate expectated values of the metrics, \ie $u_{ij} = \mathbb{E}\qty[u | x=x_i, s=s_j]$ (in the limit where the expectation can be calculated exactly, $f$ behaves like a deterministic function).
+We start by modeling the performance of a decision $x \in \mathcal{X}$.
+Outcomes depend both on the decision and on a state of the world $s \in \Omega$ that describes all relevant uncertainties.
+To model outcomes $u \in \mathcal{U}$, we use a system model $f: \mathcal{X}, \Omega \rightarrow \mathcal{U}$ \citep[we generally follow the notation of][]{mcphail_robustness:2019} which we assume to be deterministic.
+If $f$ is a stochastic function, we can define its metrics in terms of expectations, \eg $u_{ij} = \mathbb{E}\qty[u | x=x_i, s=s_j]$, which can be approximated through techniques like Monte Carlo sampling.
 
@@
 
-\citet{mcphail_robustness:2019} split the problem of aggregating \gls{sow} metrics into decision metrics into two steps: ``subsetting'' (\eg, single value, all values, subset of values) and ``metric calculation'' (\eg, mean, variance, skew, kurtosis).
+\citet{mcphail_robustness:2019} split the problem of aggregating \gls{sow} metrics into decision metrics into two steps: ``subsetting'' (\eg, single value, all values, subset of values) and ``metric calculation'' (\eg, mean, variance, skew, kurtosis).\klaus{You say something about see discussion in \citet{mcinerney_robust:2012} -- please elaborate}
 We generalize this to
@@
 A optional, but likely useful, interpretation of the weighted samples $\qty{w_j, u_{ij} \forall j=1, \ldots, J}$ is an empirical \gls{pdf}, in which case $g$ maps a \gls{pdf} of future outcomes to the metrics $m_i$.\footnote{perhaps this belongs in the Discussion}
-This is a flexible framework; \cref{tab:literature-weighting} shows some methods for robust decision making from the literature using our notation.
+This is a flexible framework; \cref{tab:literature-weighting} shows some methods for robust decision making from the literature using our notation.\james{Add some references to the table}
 
@@
 I have two flavors of this idea presented here!
-\Cref{fig:framework} takes a more Bayesian focus: the goal is to reason about properties of the \emph{distribution} over outcomes $p(u | x=x_i)$.
+\Cref{fig:framework}\james{Updates for \cref{fig:framework}: Don't need to say here that $f$ is deterministic.
+Larger text and darker color for questions.
+Rename $p(s)$ as ``prior over parameters and model structures''} takes a more Bayesian focus: the goal is to reason about properties of the \emph{distribution} over outcomes $p(u | x=x_i)$.
 The text follows \citet{mcphail_robustness:2019} more.
@@
 
-Settinget all $w_j \propto 1$ as is often done (\cref{tab:literature-weighting}) reveals a tension: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision.
+If the common practice of weighting all scenarios equally is applied ($w_j \propto 1$ in \cref{tab:literature-weighting}), a tension emerges: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision.
 
@@
 \begin{enumerate}
+    \item Assumptions should be as transparent and visible as possible to facilitate critique and dialogue\klaus{I moved this up}
     \item We should be invariant to the parameterization of $s$ -- if we sample population uniformly from 1 million to 10 million or $\log_{10}$ population uniformly from \num6 to 7, we should get the same answer
@@
           \begin{enumerate}
-              \item if we include scientifically interesting but extremely low-probability scenarios (\eg, if someone created \gls{rcp} 45), our esimtates of $m_i$ should not change meaningfully
-              \item if we include nearly redundant scenarios (\eg, if we are studying \gls{rcp} 2.6, 4.5, 6.0, and 8.5 and we then add results from a \gls{rcp} 6.01 experiment), our estimates of $m_i$ should not change meaningfully
+              \item if we include scientifically interesting but extremely low-probability scenarios (\eg, if someone created \gls{rcp} 45), our esimtates of $m_i$ should not change meaningfully.\james{Explain why}
+              \item if we include nearly redundant scenarios (\eg, if we are studying \gls{rcp} 2.6, 4.5, 6.0, and 8.5 and we then add results from a \gls{rcp} 6.01 experiment), our estimates of $m_i$ should not change meaningfully.\james{Explain why}
           \end{enumerate}
-    \item Our assumptions should be transparent and open to critique
 \end{enumerate}
@@
     \caption{
-        XLRM diagram.
-        Red nodes represent uncertainties ($X$).
-        Yellow node represents the lever ($L$).
-        The gray node shows system relationships ($R$).
-        Blue nodes show metrics ($M$).
+        Diagram showing uncertainties ($X$; red), levers ($L$; yellow), system relationships ($R$; gray), and metrics ($M$; blue).
     }\label{fig:xlrm}
@@
 Coastal floods are often decomposed into changes in local sea level \citep{kopp_evolving:2017,kopp_probabilistic:2014,wong_brick0.2:2017} plus storm surge \citep{garner_slrise:2018,cagigal_emulator:2020,rueda_surge:2016}.
-We neglect waves and hydrodynamics because (FILL IN)\ldots.
+We neglect waves and hydrodynamics for simplicity.
 
@@
     \item We have 4 \gls{rcp} scenario
-    \item We use the BRICK \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019} to define $J$ \glspl{sow}.
+    \item We use the BRICK model \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019} to define $J$ \glspl{sow}.
 \end{enumerate}
@@
     \item We model storm surge as a stationary GEV model
-    \item We apply clever priors: we place a prior on the return level (data space) rather than a prior on the parameters themselves (except the shape parameter -- we put a restrictive prior for it to be positive)\footnote{arguably we should stick to boring priors here -- the question being whether it's better to say we use iterative checks to improve the priors, or whether we want to keep things simple}
+    \item We place a prior on the return level (data space) rather than a prior on the parameters themselves (except the shape parameter -- we put a restrictive prior for it to be positive)\james{arguably we should stick to boring priors here -- the question being whether it's better to say we use iterative checks to improve the priors, or whether we want to keep things simple}
     \item We draw $K$ sets of parameters $\qty(\mu_k, \sigma_k, \xi_k)$
@@
 
-\subsection{System model}
+\subsection{A simplified didactic system model}
 
@@
 \paragraph{Prior on outputs only}
-Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{rslr} in 2100.
+Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{rslr} in 2100.\klaus{You suggest looking at another paper.
+Are you referring to this?\\\fullcite{sriver_sealevel:2018}}
 
@@
                     The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
-                    More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
+                    More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
                     Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
@@
               \item A goal is to make -- intrinsically flawed and subjective -- modeling choices as transparent as possible.
-Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions
+Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions\james{Focus on this!}
               \item Because of how we weight realizations of the future, you can apply a different weighting function and immediately assess performance.
@@
 \begin{enumerate}
-    \item MARISA (details!) for JDG funding
-    \item PSIRC (details!) for JDG funding
-    \item Rice University (do I need to thank startup?)
+    \item MARISA\james{details}
+    \item PSIRC\james{details}
+    \item Rice University (for startup)
     \item Tor Erlend Fjelde for helpful advice on implementing the GEV prior using \texttt{Turing.jl}.
+    \item For many journals, a code and data statement will go here.
 \end{enumerate}

### 2021-10-21 papers/operationalizing-bayesian-checking.tex
Klaus Edits

@@
 \usepackage[textsize=scriptsize]{todonotes}
-\setlength{\marginparwidth}{1.75in}
+\setlength{\marginparwidth}{1.5in}
 \newcommand{\james}[1]{\todo[color=giallo, textcolor=nero]{\textbf{ATTN James:~}#1}} % if desired create a custom command for each author
@@
 Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S.
-federal government will continue to subsidize flood insurance for homeowners\james{Add some references -- Klaus suggests \citet{savage:1954} or similar}.
-In response, a wide range of fields have developed scenario-based methods\ldots.
+federal government will continue to subsidize flood insurance for homeowners.\james{Add some references}
+In response, a wide range of fields have developed scenario-based methods\ldots.\klaus{Which arguments in \citet{savage:1954} did you think should be cited here?}
 This leads to the ``Multiple PDF Problem'' \citep{sharma_rcp:2021}\ldots.

### 2021-10-28 tikz/bayes-rdm.tex
a major reorganization is underway, this is just an intermediate saved point

--- 2021-10-12 tikz/conceptual.tex
@@
 \usepackage{physics}
+\usepackage{libertine}
 
 \usepackage{xcolor}
-\definecolor{TurquoiseGreen}{HTML}{B8D5B8}
-\definecolor{TealBlue}{HTML}{598392}
-\definecolor{YellowOrange}{HTML}{FFB140}
-\definecolor{RichBlack}{HTML}{011627}
+% FORZA GRANDE ROMA!!! https://teamcolorcodes.com/a-s-roma-color-codes/
+\definecolor{giallo}{HTML}{F0BC42}
+\definecolor{rosso}{HTML}{8E1F2F}
+\definecolor{grigio}{HTML}{CACACC}
+\definecolor{nero}{HTML}{000000}
+\definecolor{thirds}{HTML}{31467A}
 
@@
 
-\tikzstyle{function} = [circle, draw, fill=TurquoiseGreen, text=RichBlack, text width=5.5em, text centered]
-\tikzstyle{block} = [rectangle, draw, fill=TurquoiseGreen, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
-\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=YellowOrange!33, text=RichBlack, minimum width=6em, text centered,  minimum height=4em, align=center]
-\tikzstyle{check} = [text=TealBlue, align=center]
+\tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, text width=5.5em, text centered]
+\tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=center]
+\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=rosso, text=grigio, minimum width=6em, text centered,  minimum height=4em, align=center]
+\tikzstyle{check} = [text=thirds, align=center]
 
-\tikzstyle{line} = [draw, very thick, color=RichBlack, ->]
+\tikzstyle{line} = [draw, very thick, color=nero, ->]
 \tikzset{>=latex}
@@
 
-    \node [block] (decision) at (4, 3) {Candidate decisions\\{\color{TealBlue}$\qty{x_1, x_2, \ldots, x_I}$}};
-    \node [block] at (-4, 3) (scenarios) {Many possible futures\\{\color{TealBlue}$\qty{s_1, \ldots, s_J} \in \Omega$}};
-    \node [function] (model) at (0, 0) {Deterministic system model\\{\color{TealBlue}$f(x, s)$}};
-    \node [derived] at (0, -3) (smetrics) {Scenario metrics\\{\color{TealBlue}$u_{ij} = f(x_i, s_j)$}};
-    \node [block] (prior) at (-4, -6) {Prior over scenarios\\{\color{TealBlue}$p(s)$}};\
-    \node [derived] (weights) at (0, -6) {Scenario weights\\{\color{TealBlue}eq.~XX}};
-    \node [derived] (combined) at (4, -6) {Distribution over\\outcomes\\{\color{TealBlue}$\hat{p}(u | x_i)$}};
-    \node [derived] (dmetrics) at (8, -6) {Decision metrics\\{\color{TealBlue}$\hat{m}(x_i)$}};
+    \node [block] (decision) at (4, 3) {Candidate decisions\\{$\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$}};
+    \node [block] at (-4, 3) (scenarios) {Many possible futures\\{$\qty{s_1, \ldots, s_J} \in \Omega$}};
+    \node [function] (model) at (0, 0) {System model\\{$f(x, s)$}};
+    \node [derived] at (0, -3) (smetrics) {Outcomes\\{$u_{ij} = f(x_i, s_j)$}};
+    \node [block] (prior) at (-4, -6) {Prior over scenarios\\and model structures\\{$p(s)$}};\
+    \node [derived] (weights) at (0, -6) {Scenario weights\\{eq.~XX}};
+    \node [derived] (combined) at (4, -6) {Distribution over\\outcomes\\{$\hat{p}(u | x_i)$}};
+    \node [derived] (dmetrics) at (8, -6) {Metrics and\\tradeoffs\\{$\hat{m}(x_i)$}};
 
@@
     \path [line, dashed] (decision) -- (represented);
-    \node [check] at (5, 0) (interpretable) {Is the model credible?\\Interpretable?};
+    \node [check] at (9, 0) (interpretable) {Is the model credible?\\Interpretable?};
     \path [line, dashed] (model) -- (interpretable);
-    \node [check] at (5, -3) (discovery) {What are relevant regions\\of the $\qty(s, x)$ space?};
+    \node [check] at (9, -3) (discovery) {What are relevant regions\\of the $\qty(\Omega, \mathcal{X})$ space?};
     \path [line, dashed] (smetrics) -- (discovery);

### 2021-10-28 tikz/fema-approach.tex
a major reorganization is underway, this is just an intermediate saved point

\documentclass[tikz, border=1pt]{standalone}

\usepackage{libertine}

\usepackage{tikz}
\usetikzlibrary{
  shapes,
  arrows,
  shapes.geometric,
  positioning,
}

% Define block styles
\tikzstyle{model} = [
  rectangle,
  align=left,
]
\tikzstyle{arrow} = [draw, -latex']

% set consistent font size
\tikzset{
    every picture/.style={font issue=\small},
    font issue/.style={execute at begin picture={#1\selectfont}}
 }

\begin{document}
\begin{tikzpicture}[auto, family=\sffamily]
  \node[model] (a) at (0,0) {
    (a)\\
    \includegraphics[height=2.5in]{../plots/historic_surge.pdf}
  };
  \node[model] (b) at (8.5,0) {
    (b)\\
    \includegraphics[height=2.5in]{../plots/surge_cdf.pdf}
  };
\end{tikzpicture}
\end{document}

### 2021-10-28 notebooks/01-surge-modeling.jl

@@ -1,0 +1,907 @@
+### A Pluto.jl notebook ###
+# v0.16.4
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 50a61c27-768a-4cf8-8494-e9f99da073c6
+begin
+	# we start by loading in the packages we use
+	using Pkg
+	Pkg.activate("..")
+	using Base: @kwdef
+	using ColorSchemes
+	using CSV
+	using DataFrames
+	using Dates
+	using Distributions
+	using Downloads
+	using DrWatson
+	using DynamicPPL
+	using ExtremeStats
+	using PlutoUI
+	using Plots
+	using Plots: mm
+	using ProgressBars
+	using Random
+	using TimeSeries
+	using Turing
+	using Statistics
+	using StatsBase: pacf, corkendall
+	using StatsPlots
+	using Unitful
+	using UnitfulRecipes
+	TableOfContents()
+end
+
+# ╔═╡ 37b93f78-3750-11ec-2245-4d3c8e4a5b39
+md"""
+# Notebook 01: Surge Modeling
+
+> Doss-Gollin and Keller, 2021, In Prep. **Do not circulate!**
+
+In this notebook we'll explore the storm surge data and fit an extreme value distribution to it.
+"""
+
+# ╔═╡ 57cc380e-c124-488b-af10-a63acab149b0
+md"""
+## About Pluto
+
+If you've never seen Pluto notebooks before, check out [this video](https://www.youtube.com/watch?v=HiI4jgDyDhY).
+Every cell in this notebook (including this Markdown cell!) is produced by code, but sometimes the code is hidden.
+To see it, click the eyeball icon to the left.
+"""
+
+# ╔═╡ e33670da-89a8-4feb-8b0c-35cead899baf
+md"""
+## Get data
+
+We're going to work with the time series of annual maximum storm surges, after subtracting historic mean sea level.
+We start by reading in the hourly data (see helper functions below for more on how this actually works!)
+"""
+
+# ╔═╡ ee7a0e99-3e1b-4413-ac40-0628fa1d5d0b
+md"""
+We can plot this annual data
+"""
+
+# ╔═╡ 140cf28e-e0ce-41f2-a61e-51907805ad40
+colors = ColorSchemes.okabe_ito; # colorblind friendly
+
+# ╔═╡ 80915922-3f67-4d2f-853d-6e50c1e8706e
+md"""
+## Maximum Likelihood Fit
+
[…]

### 2021-10-28 notebooks/01_basic_plots.jl

@@ -1,155 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.14.9
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 0c531ffe-381a-448a-90d2-ca1bd42d28c6
-begin
-    using DrWatson
-	using PlutoUI
-    using Plots
-    using Unitful
-    using ColorSchemes
-    using UnitfulRecipes
-
-    using NorfolkFloods
-    using HouseElevation
-
-    PlutoUI.TableOfContents()
-end
-
-
-# ╔═╡ fcfb26dd-1767-49f7-ae97-259e09790be5
-md"# Raw Data Plots
-
-In this notebook we will plot some constants and fixed quantities from our models
-
-## Front Matter
-
-Start with package imports and definitions of constants
-"
-
-# ╔═╡ 0422680b-afb4-4e5e-a1fe-421562b1d66f
-colors = ColorSchemes.okabe_ito;
-
-# ╔═╡ 785eb441-acb7-46dd-9ce3-61847a789bcf
-md"""
-## Elevation Cost
-
-Now let's visualize the cost of elevating a house -- this is another fixed quantity that we can examine.
-This plot should reproduce supplemental figure S3 of Zarekarizi et al (2020).
-
-> Zarekarizi, M., Srikrishnan, V., & Keller, K. (2020). Neglecting uncertainties biases house-elevation decisions to manage riverine flood risks. Nature Communications, 11(1), 5361. https://doi.org/10.1038/s41467-020-19188-9
-"""
-
-# ╔═╡ b02ce2d7-2128-4d1d-93ee-0781ea61fbb9
-function plot_elevation_cost()
-
-    areas = [3_000, 2_000, 1_000]u"ft^2"
-    elevations = (0:0.1:14)u"ft"
-
-    p = plot(
-        legend = :bottomright,
-        xlabel = "Height Increase",
-        ylabel = "Elevation Cost (x1000 \$USD)",
-        ymin = 0,
-        palette = :tab10,
-        xticks = 0:2:14,
-        size = (600, 400),
-    )
-    for area in areas
-        house = HouseStructure(area, 185_000, 5u"ft") # last 2 don't matter here
-        plot!(
-            p,
-            elevations,
-            [elevation_cost(house, Δh) for Δh in elevations] ./ 1_000,
-            label = "$area house",
-            linewidth = 3,
-        )
[…]

### 2021-10-28 notebooks/02-mean-sea-level.jl

@@ -1,0 +1,376 @@
+### A Pluto.jl notebook ###
+# v0.16.4
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 6eaeeaaf-5fe8-44b2-8957-b47918f26b17
+begin
+	# we start by loading in the packages we use
+	using Pkg
+	Pkg.activate("..")
+	using ColorSchemes
+	using CSV
+	using DataFrames
+	using DrWatson
+	using NetCDF
+	using PlutoUI
+	using Plots
+	using Statistics
+	using StatsBase
+	using StatsPlots
+	using Unitful
+	using UnitfulRecipes
+	TableOfContents()
+end
+
+# ╔═╡ 49cd8998-378d-11ec-3ceb-811bf9c58697
+md"""
+# Notebook 02: Mean Sea Level
+
+> Doss-Gollin and Keller, 2021, In Prep. **Do not circulate!**
+
+In this notebook we'll explore simulations of mean sea level.
+
+*If you have not already run Notebook 01, stop and do that!*
+"""
+
+# ╔═╡ ab567401-943f-4877-8572-cfa8ffb531d4
+md"""
+## Read Data
+"""
+
+# ╔═╡ c259b7d3-0873-4502-9d4b-a5d18c5adcd8
+md"""
+The main projections of sea level rise that we use come from the BRICK model with two parameterizations of ice sheet dynamics ('fast' and 'slow') and four RCP scenarios.
+For a detailed comparison of these simulations to other model structures and assumptions, see
+
+> Ruckert, K. L., Srikrishnan, V., & Keller, K. (2019). Characterizing the deep uncertainties surrounding coastal flood hazard projections: A case study for Norfolk, VA. Scientific Reports, 9(1), 1–12. https://doi.org/10.1038/s41598-019-47587-6
+"""
+
+# ╔═╡ 5048e15a-d382-4a3c-9aa6-d0d34cb2b297
+md"""
+## Plots
+
+Let's get straight into the analysis
+"""
+
+# ╔═╡ a4a6d097-0d5c-48c7-b8bc-34baa3c910cc
+colors = ColorSchemes.okabe_ito; # colorblind friendly
+
+# ╔═╡ f33c8247-c110-412d-a7d4-4b8aa8f89f34
+md"""
[…]

### 2021-10-28 notebooks/03-sow-outcomes.jl

@@ -1,0 +1,129 @@
+### A Pluto.jl notebook ###
+# v0.16.4
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 6eaeeaaf-5fe8-44b2-8957-b47918f26b17
+begin
+	# we start by loading in the packages we use
+	using Pkg
+	Pkg.activate("..")
+	using ColorSchemes
+	using DataFrames
+	using Distributions
+	using DrWatson
+	using PlutoUI
+	using Plots
+	using Unitful
+	using Turing
+	TableOfContents()
+end
+
+# ╔═╡ 49cd8998-378d-11ec-3ceb-811bf9c58697
+md"""
+# Notebook 03: SOWs and OUtcomes
+
+> Doss-Gollin and Keller, 2021, In Prep. **Do not circulate!**
+
+In this notebook we'll explore simulations of mean sea level.
+
+*If you have not already run Notebooks 01 and 02, stop and do that!*
+"""
+
+# ╔═╡ ab567401-943f-4877-8572-cfa8ffb531d4
+md"""
+## Read Data
+
+We start by loading in the previously defined data
+"""
+
+# ╔═╡ 86eb1175-f48e-456a-a4d4-1ec5d0c0a55e
+posterior_df = DrWatson.load(datadir("processed", "surge_posterior.jld2"), "posterior") |> DataFrame
+
+# ╔═╡ 04640130-791c-4aac-8af2-450fbcbf7bdb
+lsl_trajs = DrWatson.load(datadir("processed", "msl_brick.jld2"), "trajectories")
+
+# ╔═╡ 4450ffd7-679b-49f0-8128-8f2ff2beb906
+md"""
+## Create SOWs
+
+Now we need to create a database of SOWs
+"""
+
+# ╔═╡ 5d8adb71-5e3a-41dd-aaad-c76cf3d43c61
+n_SOWs = 100_000
+
+# ╔═╡ d70e47ab-33d8-4e63-ac5d-45c126656d5d
+SOWs
+
+# ╔═╡ f180e72b-d2c5-46d7-a2ef-84c012e59cf4
+md"""
+## Outcomes
+"""
+
[…]

### 2021-10-28 notebooks/04_expected_annual_damage.jl

@@ -1,101 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.14.9
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 82c386f3-c1a1-480e-b034-14034dc85ffe
-begin
-	using DrWatson
-	using Plots
-	using Unitful
-
-	using DecisionProblem
-end
-
-# ╔═╡ 58422108-e041-11eb-10ce-13568cddf9bf
-md"""
-# Emulate Damages
-
-This notebook gives a brief explanation of the model used to calculate expected annual damage as a function of ((house elevation) - (mean sea level)).
-This model is already implemented in the `DecisionProblem` local package.
-"""
-
-# ╔═╡ f176c2b4-e441-4878-9753-b42ea6e4084b
-md"""
-The expected damages $d$ as a function of $h$, the difference between the elevation of the house and mean sea level, is the convolution of the probability density of storm surge $p(y')$ and the depth-damage function $d(y' - h)$:
-
-$$\mathbb{E}(d | h) = \int p(y') \, d(y') \, dy'$$
-
-This can be approximated as
-
-$$\frac{1}{N} \sum_{i=1}^N d(y^*_i)$$
-
-where $y^*_1, \ldots, y^*_N$ are draws from $p(y')$.
-Approximating this integral many times over (eg, for each trajectory of sea level rise and each year) may be computationally costly.
-Instead, we develop an emulator for this by calculating $\mathbb{E}(d | h)$ for many values of $h$ and then fitting a curve to this data.
-"""
-
-# ╔═╡ edf06404-68e7-476e-af48-9550f9cfd696
-surge_mc = DecisionProblem.draw_surges(1_000_000);
-
-# ╔═╡ caa24b9c-d816-4aaa-af3a-fb93d03a5897
-depths = collect(-25:0.5:20)u"ft";
-
-# ╔═╡ 411af8d9-362c-46af-8a01-42c89da9cc1d
-xp_dmg = [
-	# this function calculates the Monte Carlo integral
-	DecisionProblem.expected_damage_frac(depth, surge_mc; key = :hazus)
-	for depth in depths
-];
-
-# ╔═╡ 5f96d337-19df-4763-911b-577559fbf969
-md"and calculate the expected (per year) structural damage for each"
-
-# ╔═╡ b4fe8c62-c074-4369-aa86-36a7bf9b0564
[…]

### 2021-10-28 scripts/02_surge_model.jl

@@ -1,473 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.14.9
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ f0b9ecec-e017-11eb-3f88-c72476818de3
-begin
-    using PlutoUI
-    PlutoUI.TableOfContents(title = "Outline")
-end
-
-# ╔═╡ 38e43be9-d161-42ef-bd06-5649f52fce97
-begin
-    using ColorSchemes
-    using CSV
-    using DataFrames
-    using DecisionProblem
-	using DrWatson
-    using Plots
-    using StatsBase
-    using StatsPlots
-    using Turing
-    using Unitful
-
-    using StatsBase: pacf, corkendall
-    using Printf: @sprintf
-	using Plots: mm
-
-    using NorfolkFloods
-    using NorfolkBRICK
-end
-
-# ╔═╡ fe66d55d-0565-4740-b0a4-ce82d69cc0e9
-md"load some required packages and set up the table of contents"
-
-# ╔═╡ e502336a-d13e-468d-8ff1-b0404e20883e
-md"
-# Storm Surge Modeling
-
-In this notebook, we will develop and check our model of storm surge for Norfolk, VA.
-"
-
-# ╔═╡ b779c4e8-4c80-4958-be3c-28f589555cc3
-md"""
-## Setup
-
-Let's start with some setup.
-First, let's pick a consistent color scheme.
-We'll choose one that is colorblind-friendly according to [https://juliagraphics.github.io/ColorSchemes.jl/dev/catalogue/](https://juliagraphics.github.io/ColorSchemes.jl/dev/catalogue/).
-"""
-
-# ╔═╡ 760742a2-da3e-4ce4-a565-4f54c36ec005
-colors = ColorSchemes.okabe_ito;
-
-# ╔═╡ dd48b4f1-37b5-4698-8b15-bf709ce32dc3
-md"Next, we load in raw data and choose how many samples to create and how many chains to run"
-
-# ╔═╡ 56067de6-2d51-48ec-afc3-c42b94642508
-obs = NorfolkFloods.get_norfolk_annual(); # observed MSL and surge
-
-# ╔═╡ 69b193fa-0d8a-4c4e-81ad-17a776ba7e68
[…]

### 2021-10-28 scripts/03_plot_historical.jl

@@ -1,401 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.14.9
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 61c22ae1-ec2d-480c-b515-9cff34f3c4c3
-begin
-	using ColorSchemes
-	using DataFrames
-	using Distributions
-	using DrWatson
-	using Plots
-	using StatsPlots
-	using Turing: Chains
-	using Unitful
-
-	using Plots: mm
-	using Printf: @sprintf
-	
-	using NorfolkBRICK
-	using NorfolkFloods
-end
-
-# ╔═╡ 46688379-adb6-4ea7-8b2e-7d0864c652fc
-md"""
-load some required packages and set up the table of contents
-"""
-
-# ╔═╡ f0c4c942-ccb2-4776-a958-53bca97fc15f
-md"""
-# Observations
-
-In this notebook we'll create a plot showing:
-
-1. Historic floods
-1. Return period estimates of storm surge
-1. Projections of mean sea level (MSL) from the BRICK model (Wong et al, 2017) using both fast and slow ice sheet dynamics
-
-> Wong, T. E., Bakker, A. M. R., Ruckert, K., Applegate, P., Slangen, A. B. A., & Keller, K. (2017). BRICK v0.2, a simple, accessible, and transparent model framework for climate and regional sea-level projections. Geoscientific Model Development, 10(7), 2741–2760. https://doi.org/10.5194/gmd-10-2741-2017
-
-"""
-
-# ╔═╡ 66ac51c6-78ca-41bc-83ac-322867bec456
-obs = NorfolkFloods.get_norfolk_annual(); # observed MSL and surge
-
-# ╔═╡ 50174bdc-9c64-446a-a6b4-2866af053630
-md"now define some constants"
-
-# ╔═╡ b2864283-ce0b-45a5-9614-1fbbf9dcab26
-colors = ColorSchemes.okabe_ito # colorblind friendly
-
-# ╔═╡ 5e57a1a2-d4bc-4188-a4be-d47d1a39cddb
-begin
-	SYEAR = maximum(obs.year) + 1
-	EYEAR = SYEAR + 100	
-end;
-
-# ╔═╡ 5e0fff62-acec-4b0f-860e-6fec91abb8f1
-struct HistoricSurge
-    name::String
-    year::Int
-    is_tc::Bool
-end
-
-# ╔═╡ 2236c1aa-ebf6-4138-ae8b-fa315d11938e
-md"""
-## Historic Floods
-
[…]

### 2021-10-28 scripts/05_probabilistic_msl.jl

@@ -1,154 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.14.9
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ d32837b2-e05e-11eb-0fb7-1fc68c7f97bf
-begin
-	using ColorSchemes
-	using Distributions
-	using DrWatson
-	using KernelDensity
-	using PlutoUI
-	using Plots
-	using StatsBase
-	using StatsPlots
-	using Unitful
-	using Plots: mm
-
-    using NorfolkBRICK
-
-    PlutoUI.TableOfContents()
-end
-
-# ╔═╡ cbc1f922-4e33-4bf9-bb2c-ee93e7086c68
-md"""
-# Probabilistic MSL Analysis
-"""
-
-# ╔═╡ 50c162f8-b90a-43a6-8c9e-0074329dbbe6
-colors = ColorSchemes.okabe_ito;
-
-# ╔═╡ c981f860-7dbe-4313-8d6f-7268b4249ec7
-md"""
-We combine melding and rejection sampling
-"""
-
-# ╔═╡ a52e2861-1717-424d-b8f7-1d43f6c8efac
-trajectories = get_norfolk_brick()
-
-# ╔═╡ 7af03bed-6341-48bb-b04d-ac9f584af58b
-get_lsl_2100(trajs) = ustrip.(u"ft", get_year_data(trajs, 2100));
-
-# ╔═╡ f2c4ee04-04a8-4eaa-bfae-082b850c685d
-function normalize(w::Vector{<:Real})
-	return StatsBase.pweights(w ./ sum(w))
-end
-
-# ╔═╡ 315a762a-fce2-444b-bf2d-395d3b736457
-function rejection_sampling(trajs, target)
-	y = get_lsl_2100(trajs)
-	target_trunc = truncated(target, quantile(y, 0.0025), quantile(y, 0.9975))
-	sampling = kde(y)
-	w = pdf(target_trunc, y) ./ pdf(sampling, y)
-	return w
-end
-
-# ╔═╡ 9a83aa2b-f2cd-4c73-a2a5-20fbed0278e7
-function get_unit_weights(trajs::Vector{<:BRICKtrajectory})
-	N = length(trajs)
-	w = ones(N) ./ N
-	return normalize(w)
-end
-
-# ╔═╡ 6a361969-3851-44b6-b2fe-53f295cc2dde
[…]

### 2021-10-28 scripts/main.jl

@@ -1,51 +1,0 @@
-using Revise
-using DrWatson
-@quickactivate "2021-elevation-robustness"
-
-using PlutoSliderServer
-using Pluto
-
-function clean_outputs()
-    for dir in [datadir("processed"), plotsdir(), projectdir("notebooks")]
-        if isdir(dir)
-            rm(dir; recursive = true)
-        end
-        mkdir(dir)
-    end
-end
-
-function get_all_scripts()
-    return scriptsdir.([
-        "01_basic_plots.jl",
-        "02_surge_model.jl",
-        "03_plot_historical.jl",
-        "04_expected_annual_damage.jl",
-        "05_probabilistic_msl.jl",
-    ])
-end
-
-function run_pluto()
-    Pluto.run()
-end
-
-function convert_notebooks()
-    notebooks = get_all_scripts()
-
-    PlutoSliderServer.export_notebook.(
-        notebooks;
-        Export_output_dir = projectdir("notebooks"),
-    )
-end
-
-function source_scripts()
-    scripts = get_all_scripts()
-    for script in scripts
-        @info "Now in $script"
-        include(script)
-    end
-end
-
-# run one of the commented commands
-# source_scripts()
-# run_pluto()
-# convert_notebooks()

### 2021-10-28 src/DecisionProblem/src/DecisionProblem.jl

@@ -1,6 +1,0 @@
-module DecisionProblem
-
-include("emulator.jl")
-
-export get_expected_damage_emulator
-end

### 2021-10-28 src/DecisionProblem/src/emulator.jl

@@ -1,82 +1,0 @@
-import Interpolations
-import JLD2
-import Unitful
-import DrWatson
-using StatsBase: mean
-
-import HouseElevation
-import NorfolkFloods
-
-data_dir = abspath(joinpath(@__DIR__, "..", "data"))
-cache_dir = abspath(joinpath(@__DIR__, "..", "..", "..", "data", "processed"))
-
-const ft = Unitful.u"ft"
-
-"Get a bunch of synthetic storm surges!"
-function draw_surges(n_samples::Int)
-    obs = NorfolkFloods.get_norfolk_annual()
-    N = Int(ceil(n_samples / 100_000))
-    y = Unitful.ustrip.(ft, obs.surge)
-    model = NorfolkFloods.GEVModel(y)
-    posterior = NorfolkFloods.get_fits(model, "stationary", 100_000; n_chains = 4)
-    y_hat = NorfolkFloods.sample_predictive(posterior, N)
-    return vcat(y_hat...) .* 1ft
-end
-
-"""
-Use Monte Carlo approximation to estimate annual expected damages given  house's elevation relative to mean sea level
-
-The expected damages ``d`` as a function of ``h``, the difference between the elevation of the house and mean sea level, is the convolution of the probability density of storm surge ``p(y')`` and the depth-damage function ``d(y' - h)``:
-
-```math
-\\mathbb{E}(d | h) = \\int p(y') d(y') dy'
-````
-
-This can be approximated as
-
-```math
-\\frac{1}{N} \\sum_{i=1}^N d(y^*_i)
-````
-
-where ``y^*_1, \\ldots, y^*_N`` are draws from ``p(y')``.
-"""
-function expected_damage_frac(h::Unitful.Length, s::Vector{<:Unitful.Length}; key = :hazus)
-    house = HouseElevation.HouseStructure(2000ft^2, 150_000, h)
-    dmg = [HouseElevation.depth_damage_frac(house, si, key) for si in s]
-    return mean(dmg)
-end
-
-"Fit an interpolation function that estimates expected annual damage given difference between house elevation and MSL"
-function fit_expected_damage_emulator(key::Symbol)
-    depths = collect(-30:0.5:30)ft
-    surge_mc = draw_surges(1_000_000)
-    xp_dmg = [expected_damage_frac(depth, surge_mc; key = key) for depth in depths]
-    interp_fn = Interpolations.LinearInterpolation(
-        HouseElevation.to_feet.(depths),
-        xp_dmg,
[…]

### 2021-10-28 src/HouseElevation/src/HouseElevation.jl

@@ -1,9 +1,0 @@
-module HouseElevation
-
-include("util.jl")
-include("core.jl")
-include("deterministic_depth_damage.jl")
-
-export HouseStructure, copy, elevation_cost, depth_damage_frac, depth_damage
-
-end # module

### 2021-10-28 src/HouseElevation/src/core.jl

@@ -1,44 +1,0 @@
-import Base: copy
-import Interpolations, Unitful
-
-"All information needed to compute costs for a particular house"
-mutable struct HouseStructure{P<:Unitful.Area,T<:Real,L<:Unitful.Length}
-    A::P # the house area, in square feet
-    V::T # the structural value of the house, in USD2020
-    h::L # the height of the house relative to the gauge
-end
-
-"copy a house structure"
-function copy(hs::HouseStructure)
-    return HouseStructure(hs.A, hs.V, hs.h)
-end
-
-
-elevation_thresholds = [0 5.0 8.5 12.0 14.0][:] # cost is piecewise linear
-elevation_rates = [80.36 82.5 86.25 103.75 113.75][:] # cost per unit increase per area
-elevation_itp = Interpolations.LinearInterpolation(elevation_thresholds, elevation_rates)
-
-"""
-    elevation_cost(house, Δh)
-
-Compute the cost of elevating a particular HouseStructure `house` by `Δh`, where `house` is a `HouseStructure` and `Δh` is a length.
-
-This cost function follows Zarekarizi et al (2020), which is in turn based on the CLARA model. The valid domain of `Δh` is [0, 14ft].
-
-> Zarekarizi, M., Srikrishnan, V., & Keller, K. (2020). Neglecting uncertainties biases house-elevation decisions to manage riverine flood risks. Nature Communications, 11(1), 5361. https://doi.org/10.1038/s41467-020-19188-9
-"""
-function elevation_cost(house::HouseStructure, Δh::T) where {T<:Unitful.Length}
-    area_ft2 = to_sq_feet(house.A)
-    base_cost = (10000 + 300 + 470 + 4300 + 2175 + 3500) # all costs are in dollars
-    if Δh < 0.0ft
-        throw(DomainError(Δh, "Cannot lower the house"))
-    elseif Δh ≈ 0.0ft
-        cost = 0.0
-    elseif 0.0ft < Δh <= 14.0ft
-        rate = elevation_itp(to_feet(Δh))
-        cost = base_cost + area_ft2 * rate
-    else
-        throw(DomainError(Δh, "Cannot elevate >14ft"))
-    end
-    return cost
-end

### 2021-10-28 src/HouseElevation/src/deterministic_depth_damage.jl

@@ -1,91 +1,0 @@
-import CSV
-import DataFrames
-import Interpolations
-import Unitful
-
-data_dir = abspath(joinpath(@__DIR__, "..", "data"))
-
-"""
-Parse the file containing the parameters of the Europa depth-damage model
-
-See [https://github.com/mahkam/Zarekarizi-flood-home-elavate/blob/master/Source_Code/S03_Construction_Cost.R](https://github.com/mahkam/Zarekarizi-flood-home-elavate/blob/master/Source_Code/S03_Construction_Cost.R) for details.
-"""
-function parse_europa()
-    fname = joinpath(data_dir, "fragility_europa.csv")
-    dat = DataFrames.DataFrame(CSV.File(fname))
-    depth = dat[!, "depth_feet"]ft
-    damage_frac = dat[!, "damage_frac"]
-    return depth, damage_frac
-end
-
-"""
-Parse the file containing the parameters of the HAZUS depth-damage model
-
-See [https://github.com/mahkam/Zarekarizi-flood-home-elavate/blob/master/Source_Code/S03_Construction_Cost.R](https://github.com/mahkam/Zarekarizi-flood-home-elavate/blob/master/Source_Code/S03_Construction_Cost.R) for details.
-"""
-function parse_hazus()
-    fname = joinpath(data_dir, "fragility_hazus.csv")
-    dat = DataFrames.DataFrame(CSV.File(fname))
-    depth = (dat[!, "depth_feet"] .+ 0.0)ft
-    damage_frac = dat[!, "damage_frac"]
-    return depth, damage_frac
-end
-
-"Generic function to parse the function given a key"
-function parse_data(key)
-    if key == :europa
-        depth, damage_frac = parse_europa()
-    elseif key == :hazus
-        depth, damage_frac = parse_hazus()
-    else
-        throw("Invalid key")
-    end
-    return depth, damage_frac
-end
-
-"Get the depth-damage interpolation function"
-function get_depth_damage(key)
-    depth, damage_frac = parse_data(key)
-    prepend!(depth, minimum(depth) - 0.1ft)
-    prepend!(damage_frac, 0)
-    interp_fn = Interpolations.LinearInterpolation(
-        to_feet.(depth),
-        damage_frac,
-        extrapolation_bc = Interpolations.Flat(),
-    )
-    damage_fn = function (depth::T) where {T<:Unitful.Length}
-        return interp_fn(to_feet(depth))
-    end
-    return damage_fn
-end
-
[…]

### 2021-10-28 src/HouseElevation/src/util.jl

@@ -1,15 +1,0 @@
-#=
-some functions for deailng with units
-=#
-
-import Unitful
-
-const meter = Unitful.u"m"
-const ft = Unitful.u"ft"
-
-function to_feet(length::T) where {T<:Unitful.Length}
-    return Unitful.ustrip(Unitful.uconvert(ft, length))
-end
-function to_sq_feet(area::T) where {T<:Unitful.Area}
-    return Unitful.ustrip(Unitful.uconvert(ft^2, area))
-end

### 2021-10-28 src/NorfolkBRICK/src/NorfolkBRICK.jl

@@ -1,5 +1,0 @@
-module NorfolkBRICK
-
-include("data.jl")
-export BRICKtrajectory, get_norfolk_brick, get_year_data
-end # module

### 2021-10-28 src/NorfolkBRICK/src/data.jl

@@ -1,73 +1,0 @@
-using NetCDF
-using Unitful
-
-data_dir = abspath(joinpath(@__DIR__, "..", "data"))
-
-"""
-A single BRICK trajectory has the following information: time (defined by start and end year), local sea level, an RCP scenario, and a BRICK dynamics model (fast or slow)
-
-> Ruckert, K. L., Srikrishnan, V., & Keller, K. (2019). Characterizing the deep uncertainties surrounding coastal flood hazard projections: A case study for Norfolk, VA. Scientific Reports, 9(1), 1–12. https://doi.org/10/ggfsbt =#
-"""
-struct BRICKtrajectory{I<:Integer,T<:AbstractFloat}
-    years::UnitRange{I}
-    lsl_m::Vector{<:Unitful.Length}
-    rcp::T
-    dynamics::AbstractString
-end
-
-"""
-Parse the naive BRICK sea level simulations. It's very quick so no need to cache. Note that the default value for `fname` assumes the particular directory structure of this project; if you are using this for another project you will need to specify the filename explicitly or else change this argument manually.
-"""
-function get_norfolk_brick(
-    fname::AbstractString = joinpath(data_dir, "BRICK.nc");
-    syear::Union{Int,Missing} = missing,
-    eyear::Union{Int,Missing} = missing,
-)
-
-    @assert isfile(fname) "$fname is not a valid filename"
-
-    rcp_scenarios = NetCDF.ncread(fname, "RCP")
-    dynamics = NetCDF.ncread(fname, "dynamics")
-    years = Int.(NetCDF.ncread(fname, "time_proj"))
-    sims = NetCDF.ncread(fname, "simulation")
-    lsl = Array{typeof(rcp_scenarios[1]),4}(NetCDF.ncread(fname, "lsl_m"))Unitful.u"m" # meters
-
-    nsim = length(sims)
-    nrcp = length(rcp_scenarios)
-    ndynam = length(dynamics)
-
-    if ismissing(syear)
-        syear = minimum(years)
-    else
-        @assert minimum(years) <= syear
-    end
-    if ismissing(eyear)
-        eyear = maximum(years)
-    else
-        @assert eyear <= maximum(years)
-    end
-
-    idx_eyear = findfirst(years .== eyear)
-    idx_syear = findfirst(years .== syear)
-    trajectories = [
-        BRICKtrajectory(
-            syear:eyear,
[…]

### 2021-10-28 src/NorfolkFloods/src/NorfolkFloods.jl

@@ -1,9 +1,0 @@
-module NorfolkFloods
-
-include("hourly.jl")
-include("annual.jl")
-include("stationary.jl")
-
-export get_norfolk_annual,
-    AnnualGageRecord, get_GEV_priors, GEVModel, sample_predictive, get_fits
-end # module

### 2021-10-28 src/NorfolkFloods/src/annual.jl

@@ -1,45 +1,0 @@
-using Base: @kwdef
-import DataFrames
-using Statistics: mean
-using DataFrames: DataFrame # to extend
-
-
-
-"Create a structure for annual gauge records"
-@kwdef struct AnnualGageRecord
-    year::Vector{<:Int}
-    surge::Vector{<:Unitful.Length}
-    msl::Vector{<:Unitful.Length}
-    gage_name::String
-    gage_id::String
-end
-
-function AnnualGageRecord(annual_df::DataFrames.AbstractDataFrame)
-    return AnnualGageRecord(
-        year = annual_df[!, :year],
-        surge = annual_df[!, :surge],
-        msl = annual_df[!, :msl],
-        gage_name = annual_df[!, :gage_name][1],
-        gage_id = annual_df[!, :gage_id][1],
-    )
-end
-
-function AnnualGageRecord(hourly::HourlyGageRecord)
-    hourly_df = DataFrames.DataFrame(hourly.data)
-    DataFrames.dropmissing!(hourly_df)
-    hourly_df[!, :year] = [Dates.year(date) for date in hourly_df[!, :timestamp]]
-    annual_df = DataFrames.combine(
-        DataFrames.groupby(hourly_df, :year),
-        :sl => maximum => :surge,
-        :sl => mean => :msl,
-    )
-    annual_df[!, :gage_id] .= hourly.gage_id
-    annual_df[!, :gage_name] .= hourly.gage_name
-    annual_df[!, :surge] = annual_df[!, :surge] - annual_df[!, :msl]
-    return AnnualGageRecord(annual_df)
-end
-
-"Get the annual stage data"
-function get_norfolk_annual()
-    return AnnualGageRecord(get_norfolk_hourly())
-end

### 2021-10-28 src/NorfolkFloods/src/hourly.jl

@@ -1,89 +1,0 @@
-import CSV
-import DataFrames
-import Dates
-import Downloads
-import DrWatson
-import ProgressBars
-import TimeSeries
-import Unitful
-
-"data structure to hold hourly data"
-struct HourlyGageRecord
-    data::TimeSeries.TimeArray
-    gage_name::String
-    gage_id::String
-end
-
-"""
-Get the URL for the NOAA file containing tides and currents data
-
-Some things have been hard-coded:
-    datum=NAVD
-    station=8638610
-    time_zone=GMT
-    units=metric
-"""
-function get_noaa_url(year::Int)
-    return "https://api.tidesandcurrents.noaa.gov/api/prod/datagetter?product=hourly_height&application=NOS.COOPS.TAC.WL&begin_date=$(year)0101&end_date=$(year)1231&datum=NAVD&station=8638610&time_zone=GMT&units=metric&format=csv"
-end
-
-"Get the path of local CSV file where to save a single year of data"
-function get_csv_fname(year::Int)
-    cache_dir = abspath(joinpath(@__DIR__, "..", "..", "..", "data", "processed", "gauge"))
-    if !isdir(cache_dir)
-        mkdir(cache_dir)
-    end
-    return joinpath(cache_dir, "$(year).csv")
-end
-
-"Parse one of the CSV files downloaded from NOAA"
-function parse_noaa_csv_file(fname)
-    surge = DataFrames.DataFrame(CSV.File(fname))[[1, 2]]
-    DataFrames.rename!(surge, [:datetime, :sl])
-    surge[!, :datetime] =
-        map(x -> Dates.DateTime(x, "YYYY-mm-dd HH:MM"), surge[!, :datetime])
-    surge[!, :sl] = surge[!, :sl]Unitful.u"m" # explicitly specify unit to avoid issues
-    surge[!, :sl] = Unitful.uconvert.(Unitful.u"ft", surge[!, :sl])
-    DataFrames.dropmissing!(surge)
-    return surge
-end
-
-"Get a DataFrame of hourly sea level readings"
-function get_hourly_obs(year::Int)
-    fname = get_csv_fname(year)
-    try
-        ts = parse_noaa_csv_file(fname)
-        return ts
-    catch
-        url = get_noaa_url(year)
-        @info "downloading data for $year"
-        Downloads.download(url, fname)
-        ts = parse_noaa_csv_file(fname)
-        return ts
-    end
-end
-
-"Get all hourly sea level data"
[…]

### 2021-10-28 src/NorfolkFloods/src/stationary.jl

@@ -1,81 +1,0 @@
-using Distributions
-using JLD2
-using Turing
-
-data_dir = abspath(joinpath(@__DIR__, "..", "data"))
-cache_dir = abspath(joinpath(@__DIR__, "..", "..", "..", "data", "processed"))
-
-"Specifies priors over the return levels. See `notebooks/surge_model.jl` for visualization and description."
-function get_GEV_priors()
-    return [
-        (0.9, Distributions.LogNormal(1.8, 0.4)), # 10 year flood
-        (0.99, Distributions.LogNormal(2, 0.4)), # 100 year flood
-        (0.999, Distributions.LogNormal(2.2, 0.4)), # 1000 year flood
-    ]
-end
-
-# this is not exported
-gev_priors = get_GEV_priors()
-
-"A stationary GEV model. See `get_GEV_priors` for info on the priors."
-@model function GEVModel(y)
-
-    # completely flat priors (albeit positive -- see docstring)
-    μ ~ Turing.FlatPos(0.0)
-    σ ~ Turing.FlatPos(0.0)
-    ξ ~ Turing.FlatPos(0.0)
-
-    # we're going to use this distribution repeatedly
-    dist = Distributions.GeneralizedExtremeValue(μ, σ, ξ)
-
-    # implement the prior on quantile levels
-    for (prob, prior_dist) in gev_priors
-        rl = Distributions.quantile(dist, prob)
-        Turing.@addlogprob!(logpdf(prior_dist, rl))
-    end
-
-    # data model
-    if !any(ismissing.(y))
-        y .~ dist
-    end
-end
-
-"Draws samples from the posterior distribution of fit. This will simulate a vector of `N` samples for each  posterior draw in `fit`."
-function sample_predictive(fit::Chains, N::Int)
-    yhat = [
-        rand(Distributions.GeneralizedExtremeValue(μ, σ, ξ), N) for
-        (μ, σ, ξ) in zip(fit[:μ][:], fit[:σ][:], fit[:ξ][:])
-    ]
-    return yhat
-end
-
-"Cache the sampling outputs to avoid constant re-running"
-function get_fits(
-    model::DynamicPPL.Model,
-    model_name::String,
-    n_samples::Int;
-    n_chains::Int = 1,
-    drop_warmup::Bool = true,
-    overwrite::Bool = false,
-)
-    cachename =
-        joinpath(cache_dir, "surge_models", "stationary_$(model_name)_$(n_samples).jld2")
[…]

### 2021-11-02 tikz/priors-tradeoffs.tex
draft figures to meet with Klaus

--- 2021-10-28 tikz/fema-approach.tex
@@
     (a)\\
-    \includegraphics[height=2.5in]{../plots/historic_surge.pdf}
+    \includegraphics[height=2.5in]{../plots/lsl_priors.pdf}
   };
-  \node[model] (b) at (8.5,0) {
+  \node[model] (b) [right=0.1in of a] {
     (b)\\
-    \includegraphics[height=2.5in]{../plots/surge_cdf.pdf}
+    \includegraphics[height=2.5in]{../plots/tradeoffs_height_damages_byprior.pdf}
+  };
+  \node[model] (c) [right=0.1in of b]{
+    (c)\\
+    \includegraphics[height=2.5in]{../plots/tradeoffs_height_totalcost_byprior.pdf}
+  };
+  \node[model] (d) [right=0.1in of c]{
+    (d)\\
+    \includegraphics[height=2.5in]{../plots/tradeoffs_height_iqr_byprior.pdf}
   };

### 2021-11-02 notebooks/01-surge-modeling.jl

@@ -797,6 +797,7 @@ begin
 			legend = :topleft,
 			xticks = (xticks, string.(xticks)),
 			dpi=250, # for saving
+			size=(500, 500),
 		)
 		
 		# start with Bayesian posterior draws
@@ -1042,7 +1043,7 @@ end
 # ╟─97cb91a4-abde-4752-a13a-343ae69a2d5b
 # ╟─d9d8c5b1-97b9-4a95-b8e3-34e2285c4156
 # ╟─128fccc1-fab6-4a57-9b9c-76ce851ba503
-# ╟─a3a7d738-36ea-4057-b9c0-f3293e5aa539
+# ╠═a3a7d738-36ea-4057-b9c0-f3293e5aa539
 # ╟─2ddb440a-812b-4766-b4f1-73e5bb270ead
 # ╟─61a65bef-9e9b-4b00-8c00-1e735b847237
 # ╟─5586c372-498e-4649-a455-b7054ca33404

### 2021-11-02 notebooks/02-mean-sea-level.jl

@@ -165,36 +165,35 @@ all_trajs = get_norfolk_brick(syear=2020, eyear=2100)
 
 # ╔═╡ c73ddb13-50d4-4e6d-ac54-c171c66b3786
 begin
-	function plot_lsl_models()
+	function plot_lsl_models(;ub = 0.95, lb = 0.05)
 		p = plot(
 			legend=:topleft,
 			xlabel="Year",
 			ylabel="Mean Sea Level at Norfolk, VA [ft]",
 			size = (500, 500),
 		)
-		models = ((2.6, "slow"), (4.5, "fast"), (8.5, "fast"))
+		models = ((2.6, "slow"), (4.5, "slow"), (6.0, "fast"), (8.5, "fast"))
 		for (model, color) in zip(models, colors)
 			rcp, dyn = model
 			trajs = [
 				traj for
-					traj in all_trajs if (traj.rcp == rcp) & (traj.dynamics == dyn)
-					]
-			traj = first(trajs)
+				traj in all_trajs if (traj.rcp == rcp) & (traj.dynamics == dyn)
+			]
 			sims = ustrip.(u"ft", hcat([traj.lsl for traj in trajs]...))
 			N = size(sims)[1]
-
-			upper = [quantile(sims[i, :], 0.95) for i = 1:N]
-			lower = [quantile(sims[i, :], 0.05) for i = 1:N]
+			upper = [quantile(sims[i, :], ub) for i = 1:N]
+			lower = [quantile(sims[i, :], lb) for i = 1:N]
 			plot!(
 				p,
-				traj.years,
+				first(trajs).years,
 				upper,
 				fillrange = lower,
 				fillalpha = 0.45,
-				label = "RCP $rcp / $dyn, 90% CI",
+				label = "RCP $rcp / BRICK $dyn",
 				color = color,
-				linewidth = 0,
+				linewidth = 1,
 			)
+			plot!(p, first(trajs).years, lower, label=false, color=color, linewidth=1)
 		end
 		return p
 	end

### 2021-11-02 notebooks/03-sow-outcomes.jl

@@ -20,8 +20,10 @@ begin
 	using UnitfulRecipes
 	using PlutoUI
 	using Plots
+	using Plots: mm
 	using ProgressBars
 	using Tullio
+	using StaticArrays
 	using StatsPlots
 	TableOfContents()
 end
@@ -84,7 +86,7 @@ md"""
 We'll start by defining a house.
 We'll use the same house for all our calculations.
 A house is characterized by its floor plan, in square feet, its dollar value, and its initial elevation relative to the gauge.
-Like Zarekarizi et al (2020), we'll consider a house that is below the BFE.
+Like Zarekarizi et al (2020), we'll consider a house that was at the BFE in 2000.
 """
 
 # ╔═╡ fcb1ddc0-e6b8-4847-b994-fce762046f8e
@@ -136,8 +138,8 @@ md"""
 Following our paper's notation, we need to define a function 𝑓 (`\itf`) that takes in a scenario s and a decision x and spits out an outcome, which is just a vector of length $K$.
 """
 
-# ╔═╡ 0fbff6fd-69b8-4aa9-904a-cf6b004428ba
-K = 4;
+# ╔═╡ 5e9e8516-cf5d-4a62-8617-ba7bca1c6580
+K = 4; # number of outcomes to track
 
 # ╔═╡ f180e72b-d2c5-46d7-a2ef-84c012e59cf4
 md"""
@@ -148,7 +150,7 @@ We'll use some notation magic with the `Tullio.jl` package to create an array of
 
 # ╔═╡ 0bd6b397-59ad-429d-bc84-c5ab6bc91ba2
 md"""
-## Scenario Discovery
+## Scenario visualizations
 """
 
 # ╔═╡ adefc367-4f7e-41b1-8666-b716c346cbd1
@@ -190,48 +192,6 @@ begin
 		push!(𝐬, sow)
 	end
 	first(𝐬)
-end
-
-# ╔═╡ 31b3c3af-32d6-410d-bca2-f200e0326801
-y = zeros(Float64, length(𝐬), length(𝐱), K);
-
-# ╔═╡ 3f3a888b-4cc9-4415-9e50-ef941b7fdd95
-DrWatson.wsave(datadir("processed", "sows.jld2"), Dict("y" => y));
[…]

### 2021-11-02 notebooks/04-more-msl.jl

@@ -1,56 +1,0 @@
-### A Pluto.jl notebook ###
-# v0.16.4
-
-using Markdown
-using InteractiveUtils
-
-# ╔═╡ 62956480-37bc-11ec-1901-13f49754cd1d
-begin
-	using Pkg
-	Pkg.activate("..")
-	using CSV
-	using DataFrames
-	using Downloads
-	using DrWatson
-end
-
-# ╔═╡ 1db7b953-25b2-4969-8b1f-d4af37b9223c
-md"Some more scenarios"
-
-# ╔═╡ 5e18bac7-5217-4e42-b1fa-0630449173d9
-function get_file(fname::String, url::String)
-    if !isfile(fname)
-		Downloads.download(url, fname)
-	end
-	return 1
-end;
-
-# ╔═╡ efa86325-f600-484c-abac-53ecc4bded59
-sources = [
-	(
-		url = "https://github.com/scrim-network/local-coastal-flood-risk/raw/master/Data/LSLproj_MC_299_rcp26.csv",
-		fname=datadir("processed", "MC_200_rcp26.csv")
-	)
-]
-
-# ╔═╡ d54110c9-975e-4192-bab0-e44dc7baf6be
-[get_file(source.fname, source.url) for source in sources]
-
-# ╔═╡ b45d4406-4b4b-49c4-a77f-8d2ab5c05d7f
-df = CSV.File(first(sources).fname) |> DataFrame;
-
-# ╔═╡ 4d18f67e-0376-4a89-9013-d9a035e8b882
-df[!, :sample] = 1:nrow(df)
-
-# ╔═╡ c34e10df-739b-49b9-9ca4-caaea5e632d6
-df
-
-# ╔═╡ Cell order:
-# ╠═62956480-37bc-11ec-1901-13f49754cd1d
-# ╠═1db7b953-25b2-4969-8b1f-d4af37b9223c
-# ╠═5e18bac7-5217-4e42-b1fa-0630449173d9
-# ╠═efa86325-f600-484c-abac-53ecc4bded59
-# ╠═d54110c9-975e-4192-bab0-e44dc7baf6be
-# ╠═b45d4406-4b4b-49c4-a77f-8d2ab5c05d7f
-# ╠═4d18f67e-0376-4a89-9013-d9a035e8b882
-# ╠═c34e10df-739b-49b9-9ca4-caaea5e632d6

### 2021-11-02 notebooks/04-outcomes-tradeoffs.jl

@@ -1,0 +1,345 @@
+### A Pluto.jl notebook ###
+# v0.17.0
+
+using Markdown
+using InteractiveUtils
+
+# ╔═╡ 62956480-37bc-11ec-1901-13f49754cd1d
+begin
+	using Pkg
+	Pkg.activate("..")
+	using ColorSchemes
+	using CSV
+	using DataFrames
+	using Distributions
+	using Downloads
+	using DrWatson
+	using KernelDensity
+	using PlutoUI
+	using StaticArrays
+	using StatsBase
+	using StatsPlots
+	using Unitful
+	using Turing
+	TableOfContents()
+end
+
+# ╔═╡ 3c2d6a02-1f68-48ef-9c9c-3b29a653b2c5
+md"""
+# Notebook 04: Outcomes and Tradeoffs
+
+> Doss-Gollin and Keller, 2021, In Prep. **Do not circulate!**
+
+*If you have not already run Notebooks 01, 02, and 03 stop and do that!*
+"""
+
+# ╔═╡ c57017ea-d32d-4fb0-8cf7-88ffc45edc93
+colors = ColorSchemes.okabe_ito;
+
+# ╔═╡ 24a51c14-9046-42d0-8999-2b0c5e6ee519
+md"""
+## Load data
+
+We start by loading the scenarios (𝐬; `\bfs`), decisions (𝐱; `\bfx`), and outcomes (y).
+"""
+
+# ╔═╡ d92e3059-9bc5-4993-a38d-6dfad3854ae5
+begin
+	cachefile = datadir("processed", "sows_outcomes.jld2")
+	u = DrWatson.load(cachefile, "u")
+	𝐬 = DrWatson.load(cachefile, "𝐬")
+	𝐱 = DrWatson.load(cachefile, "𝐱")
+end;
+
+# ╔═╡ 4544a2aa-0926-45c9-9082-5f706f7b9273
+md"""
+## Bayesian weighting
+
+>TODO add some explanation
+"""
+
+# ╔═╡ 37b2cda6-7d57-404a-a235-fb4ec04f26c4
+begin
+	lsl_2100 = ustrip.(u"ft", map(s -> s.lsl[findfirst(s.years .== 2100)], 𝐬))
+	lb = minimum(lsl_2100)#quantile(lsl_2100, 0.0025)
+	ub = maximum(lsl_2100)#quantile(lsl_2100, 0.9975)
+	(lb, ub)
+end
+
[…]

### 2021-11-02 scripts/runall.jl

@@ -1,5 +1,5 @@
 using DrWatson
-using PlutoSliderServer
+using PlutoSliderServer: export_notebook
 using Pluto
 
 NB_DIR = projectdir("notebooks")
@@ -29,8 +29,8 @@ function get_all_notebooks()
 end
 
 """run Pluto"""
-function run_pluto()
-    Pluto.run()
+function run_pluto(; port = 8888)
+    Pluto.run(port = port)
 end
 
 """run all notebooks and convert to HTML files"""

### 2021-12-10 papers/agu21/case-study.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{case study: norfolk, va}
    We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard, following the approach of \cite{zarekarizi_suboptimal:2020}.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{historic_surge.pdf}
        \caption{
            Time series of annual maximum of storm surge (after subtracting mean sea level) at Sewells Point, VA (NAVD datum).
        }
        \label{fig:observations}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{hazard-tradeoffs.pdf}
        \caption{
            Neglecting hydrodynamics, we add sea level rise (a; see ref.~\cite{ruckert_coastal:2019}) to a Bayesian GEV model of storm surge (b) to get flood hazard.
            The convolution of hazard and fragility (c; see ref.~\cite{zarekarizi_suboptimal:2020}) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
        }
    \end{figure}
\end{block}

### 2021-12-10 papers/agu21/combining.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{synthesizing pdfs for decision relevance}
    If there are only a few PDFs, then qualitative comparisions (\eg, \cref{fig:tradeoffs}) may be sufficient.
    With \emph{many} PDFs (\eg, \cref{fig:boxplots}), further synthesis is needed.

    We consider the \emph{common case} of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{bayes-rdm.pdf}
        \caption{
            Sketch of the ``Prior Model Averaging'' framework.
        }\label{fig:flowchart}
    \end{figure}
\end{block}

### 2021-12-10 papers/agu21/cut-paste.tex
Some results for AGU,  hope I haven't broken anything too badly!

Flood elevation is a widely used strategy for managing household flood risk.
Yet official guidance from \gls{fema}, \textbf{and regulated through local building codes} in many areas, uses a ``one size fits all'' approach that is silent on important questions.
\begin{framed}
    \begin{description}
        \item[Q1] How to adapt recommendations to a particular house's structure and a particular household's objectives?
        \item[Q2] How to incorporate nonstationary flood risk into decision support and building codes?
        \item[Q3] Which scenarios and model structures should we design for?
    \end{description}
\end{framed}
\textbf{These issues challenge efforts to incorporate climate change into engineering codes more generally.}

### 2021-12-10 papers/agu21/doesnt-fit-all.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{one size doesn't fit all!}
    \begin{itemize}
        \item Tailoring guidance to specific building characteristics can improve outcomes
        \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \cite{xian_elevation:2017,zarekarizi_suboptimal:2020}
        \item Both over- and under-building can be costly \cite{ansar_bigisfragile:2017,DossGollin:2019}
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.475\textwidth]{5.5.pdf}%
            \includegraphics[width=0.475\textwidth]{7.5.pdf}
            \caption{
                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (L) \SI{2.5}{ft} and (R) \SI{0.5}{ft} below the \gls{bfe} under \gls{rcp} 6.0 with fast BRICK dynamics.
                Note $y$-axes are not equal.
            }
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/house-elevation.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{house elevation}
    Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \cite{zarekarizi_suboptimal:2020,xian_elevation:2017}.
    \begin{framed}
        \begin{description}
            \item[Q1] How to adapt guidance to building characteristics or household preferences?
            \item[Q2] How does nonstationary hazard change guidance?
            \item[Q3] Which model of nonstationary hazard should be used?
        \end{description}
    \end{framed}
    \begin{figure}
        \centering
        \subcaptionbox{Selinsgrove, PA}{\includegraphics[height=3.35in]{img_selinsgrove.jpeg}}%
        \hfill
        \subcaptionbox{Houston, TX}{\includegraphics[height=3.35in]{img_braeswood.png}}\\
        \subcaptionbox{Plaquemines Parish, LA}{\includegraphics[height=3.3in]{epstein-2020-la.jpeg}}%
        \hfill
        \subcaptionbox{Dare County, NC}{\includegraphics[height=3.3in]{nichols-ocracoke.jpeg}}%
        \caption{
            \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols.
        }
        \label{fig:images}
    \end{figure}
\end{block}

### 2021-12-10 papers/agu21/multiple-pdf.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{the multiple pdf problem}
    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie the \textbf{\itshape{Multiple PDF Problem}}.
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=\textwidth]{msl_boxplots.pdf}
            \caption{
                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017}.
                We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
            }\label{fig:boxplots}
        \end{figure}
    \end{framed}
    Considering other models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} would \emph{compound} this problem.
    \textbf{Which PDF should long-lasting infrastructure like housing be designed for?}
\end{block}

### 2021-12-10 papers/agu21/not-equally-plausible.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{we're always }
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{lsl_priors.pdf}
        \caption{
            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution.
            This may or may not accurately reflect available information.
        }
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[width=\textwidth]{inference_weights.pdf}
        \caption{
            The prior model averaging approach (\cref{fig:flowchart}) re-weights PDFs based on their distance, and that of other PDFs available, from the (ideally coproduced!) prior.
        }
    \end{figure}
\end{block}

### 2021-12-10 papers/agu21/other-applications.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{looking ahead}
    This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
    \textbf{\scshape let's collaborate on:}
    \begin{itemize}
        \item More complex models to capture more relevant metrics
        \item Better priors over nonstationary hazard
        \item Interacting, sequential decisions
        \item Inclusive assessment of ethical questions around scenario weighting
    \end{itemize}
    \begin{framed}
        \begin{figure}
            \centering
            \subcaptionbox{Stormwater}{\includegraphics[height=3.5in]{culvert.jpeg}}%
            \hfill
            \subcaptionbox{Bridge Clearance}{\includegraphics[height=3.5in]{bridge_clearance.jpg}}
            \caption{
                House elevation is just one of many design problems where nonstationary hazards, subject to scenario and model structure uncertainties, drive outcomes.
                \faIcon{camera}: (a) JDG (b) D. Howard / Wikipedia.
            }
            \label{fig:images}
        \end{figure}
    \end{framed}
\end{block}

### 2021-12-10 papers/agu21/scenario-matters.tex
Some results for AGU,  hope I haven't broken anything too badly!

\begin{block}{it matters which pdf you choose!}
    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes (see ref.~\cite{gelman_philosophy:2013} for a more philosophical discussion).
    \begin{framed}
        \begin{figure}
            \centering
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byscenario.pdf}%
            \includegraphics[width=0.48\textwidth]{tradeoffs_height_iqr_byscenario.pdf}
            \caption{
                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the model selected.
            }
            \label{fig:tradeoffs}
        \end{figure}
    \end{framed}
    Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement
\end{block}

### 2021-12-10 tikz/bayes-rdm.tex
Some results for AGU,  hope I haven't broken anything too badly!

@@
 
-\tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, text width=5.5em, text centered]
+\tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, minimum width=5.5em, text centered, align=center]
 \tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=center]
@@
 
-    \node [block] (decision) at (4, 3) {Candidate decisions\\{$\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$}};
-    \node [block] at (-4, 3) (scenarios) {Many possible futures\\{$\qty{s_1, \ldots, s_J} \in \Omega$}};
-    \node [function] (model) at (0, 0) {System model\\{$f(x, s)$}};
-    \node [derived] at (0, -3) (smetrics) {Outcomes\\{$u_{ij} = f(x_i, s_j)$}};
-    \node [block] (prior) at (-4, -6) {Prior over scenarios\\and model structures\\{$p(s)$}};\
-    \node [derived] (weights) at (0, -6) {Scenario weights\\{eq.~XX}};
-    \node [derived] (combined) at (4, -6) {Distribution over\\outcomes\\{$\hat{p}(u | x_i)$}};
-    \node [derived] (dmetrics) at (8, -6) {Metrics and\\tradeoffs\\{$\hat{m}(x_i)$}};
+    \node [block] (decisions) at (0, 3) {
+        Candidate decisions\\
+        $\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$
+    };
+    \node [block] at (-4, 3) (models) {
+        Multiple PDFs\\
+        $\qty{M_1, \ldots, M_K} \in \mathcal{M}$
+    };
+    \node [block] at (-4, 0) (simulations) {
+        Simulated SOWs\\
+        $\qty{s_1, \ldots, s_{N_k}}_k$
+    };
+    \path [line] (models) -- (simulations);
+    \node [function] (model) at (0, 0) {
+        System model\\
+        $u_{ij} = f(x_i, s_j)$
+    };
+    \path [line] (simulations) -- (model);
+    \path [line] (decisions) -- (model);
 
-    \path [line] (decision) -- (model);
-    \path [line] (scenarios) -- (model);
-    \path [line] (model) to (smetrics);
-    \path [line] (scenarios.south) to[out=-90, in=135] (weights.north west);
+    \node [derived] (combined) at (4, -3) {
+        Conditional distributions\\
+        over outcomes\\
+        $\hat{p} \qty(u | x_i, M_k)$
+    };
+    \path [line] (model) -- (combined);
+
+    \node [function] (prior) at (-4, -3) {
+        Prior\\
+        over SOWs\\
+        $p\qty(s)$
+    };
+    \node [derived] (weights) at (0, -3) {
+        Model weights\\
+        $w_1, \ldots, w_K$
+    };
+    \path [line] (simulations) -- (weights);
     \path [line] (prior) -- (weights);
-    \path [line] (smetrics) to (combined);
-    \path [line] (weights) -- (combined);
-    \path [line] (combined) to (dmetrics);
 
-    \node [check] at (-9, 3) (span) {Do scenarios span\\plausible outcomes?};
-    \path [line, dashed] (scenarios) -- (span);
-    \node [check] at (9, 3) (represented) {Are relevant levers\\represented?};
-    \path [line, dashed] (decision) -- (represented);
-    \node [check] at (9, 0) (interpretable) {Is the model credible?\\Interpretable?};
-    \path [line, dashed] (model) -- (interpretable);
-    \node [check] at (9, -3) (discovery) {What are relevant regions\\of the $\qty(\Omega, \mathcal{X})$ space?};
-    \path [line, dashed] (smetrics) -- (discovery);
-    \node [check] at (-4, -8) (priorcheck) {Are these consistent\\with prior knowledge?};
-    \path [line, dashed] (prior) -- (priorcheck);
-    \node [check] at (0, -8) (weightcheck) {Are computational\\issues apparent?};
-    \path [line, dashed] (weights) -- (weightcheck);
-    \node [check] at (4, -8) (ppc) {Are these consistent\\with our expectations?};
-    \path [line, dashed] (combined) -- (ppc);
-
+    \node [derived] (aggregated) at (4, -6) {
+        Aggregated distribution\\
+        over outcomes\\
+        $\hat{p}(u | x_i)$
+    };
+    \path [line] (combined) -- (aggregated);
+    \path [line] (weights) -- (aggregated);
 \end{tikzpicture}

### 2021-12-10 notebooks/01-surge-modeling.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.0
+# v0.17.3
 
 using Markdown
 using InteractiveUtils
@@ -39,7 +39,7 @@ end
 md"""
 # Notebook 01: Surge Modeling
 
-> Doss-Gollin and Keller, 2021, In Prep. **Do not circulate!**
+> Doss-Gollin, James and Klaus Keller (In Prep). "*Which Scenario Should We Design For? Insights from House Elevation on the Multiple PDF Problem.*" **Do not circulate!**
 
 In this notebook we'll explore the storm surge data and fit an extreme value distribution to it.
 """
@@ -563,7 +563,8 @@ begin
 			markerstrokewidth = 0,
 			titlefontsize = 10,
 			labelfontsize = 9,
-			size = (750, 500),
+			size = (750, 250),
+			left_margin = 5mm,
 		)
 
 		storms = [
@@ -1001,7 +1002,7 @@ begin
 end
 
 # ╔═╡ Cell order:
-# ╟─37b93f78-3750-11ec-2245-4d3c8e4a5b39
+# ╠═37b93f78-3750-11ec-2245-4d3c8e4a5b39
 # ╠═50a61c27-768a-4cf8-8494-e9f99da073c6
 # ╟─57cc380e-c124-488b-af10-a63acab149b0
 # ╟─e33670da-89a8-4feb-8b0c-35cead899baf
@@ -1009,7 +1010,7 @@ end
 # ╠═f9a2c638-2988-4a70-b774-5dc6babc3aaf
 # ╟─ee7a0e99-3e1b-4413-ac40-0628fa1d5d0b
 # ╠═140cf28e-e0ce-41f2-a61e-51907805ad40
-# ╟─d694f959-4946-488a-a505-5df4ef43297d
+# ╠═d694f959-4946-488a-a505-5df4ef43297d
 # ╟─ea8110ab-bf2b-4f66-aae5-382695a97607
 # ╠═96a9bfa1-72fd-4dff-8a49-fbb193aa5273
 # ╟─80915922-3f67-4d2f-853d-6e50c1e8706e

### 2021-12-10 notebooks/02-mean-sea-level.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.0
+# v0.17.2
 
 using Markdown
 using InteractiveUtils
@@ -16,6 +16,7 @@ begin
 	using NetCDF
 	using PlutoUI
 	using Plots
+	using Plots: mm
 	using Statistics
 	using StatsBase
 	using StatsPlots
@@ -76,7 +77,12 @@ Next, let's plot 95% Confidence Intervals for a couple of representative models
 
 # ╔═╡ 5f912aeb-c9e4-485a-897f-a71e84ec6038
 md"""
-Last, we can show individual scenarios with lines
+Alternatively, we can show individual scenarios with lines
+"""
+
+# ╔═╡ f815e33f-97fa-4a1b-b8fc-77241d81d172
+md"""
+Last, we can create boxplots
 """
 
 # ╔═╡ 48b37aa3-b756-4628-be82-5289ed202aa2
@@ -344,6 +350,39 @@ begin
 	p3
 end
 
+# ╔═╡ 84182c9e-1b97-4e3e-b3dc-b1c8c50d899c
+begin
+	function make_boxplots()
+
+		all_trajs = get_norfolk_brick()
+		noaa = get_noaa_scenarios()
+		
+		rows = [
+			DataFrame(
+				"scenario" => "RCP $(traj.rcp), $(traj.dynamics) dynamics",
+				"msl_2100" => ustrip(u"ft", get_year_data(traj, 2100))
+			)
+			for traj in all_trajs
+		]
+		lsl_2100 = vcat(rows...)
+
+		p = plot(
+			ylabel = "MSL at Norfolk, VA in 2100 [ft]",
+			bottom_margin = 10mm,
+			left_margin = 5mm,
+			size = (700, 300),
+			xtickfontsize = 7,
+			ylabelfontsize = 9,
+		)
+		@df lsl_2100 violin!(p, :scenario, :msl_2100, xrotation=30, legend=false)
+		
+		return p
+	end
+	p4 = make_boxplots()
+	savefig(p4, plotsdir("msl_boxplots.pdf"))
+	plot(p4)
+end
+
 # ╔═╡ Cell order:
[…]

### 2021-12-10 notebooks/03-sow-outcomes.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.0
+# v0.17.2
 
 using Markdown
 using InteractiveUtils
@@ -73,7 +73,7 @@ We'll discretize this and look at results over all the discrete values for simpl
 """
 
 # ╔═╡ 37a5ab0d-db74-4bff-aa12-f79a6404368a
-𝐱 = (0:0.5:14)u"ft";
+𝐱 = (0:0.125:14)u"ft";
 
 # ╔═╡ 26eb91da-a71b-4eb9-9a8c-a9a336b34446
 md"""
@@ -86,7 +86,7 @@ md"""
 We'll start by defining a house.
 We'll use the same house for all our calculations.
 A house is characterized by its floor plan, in square feet, its dollar value, and its initial elevation relative to the gauge.
-Like Zarekarizi et al (2020), we'll consider a house that was at the BFE in 2000.
+We'll consider a house that was about 1ft below the BFE in 2000.
 """
 
 # ╔═╡ fcb1ddc0-e6b8-4847-b994-fce762046f8e
@@ -145,7 +145,7 @@ K = 4; # number of outcomes to track
 md"""
 ## Calculate Outcomes
 
-We'll use some notation magic with the `Tullio.jl` package to create an array of outcomes indexed by [scenario, decision, outcome variable].
+We'll use some notation magic with the `Tullio.jl` package to create an array of outcomes $u$ indexed by [scenario, decision, outcome variable].
 """
 
 # ╔═╡ 0bd6b397-59ad-429d-bc84-c5ab6bc91ba2
@@ -277,7 +277,7 @@ end;
 
 # ╔═╡ a0a7288a-a838-4f05-950a-68088258d7db
 function get_default_house()
-	return HouseStructure(1_000u"ft^2", 185_000.0, 8.5u"ft")
+	return HouseStructure(1_000u"ft^2", 185_000.0, 7.5u"ft")
 end;
 
 # ╔═╡ 01f44422-a92f-42f7-a0c6-86e797f09c3d
@@ -333,9 +333,15 @@ begin
[…]

### 2021-12-10 notebooks/04-outcomes-tradeoffs.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.0
+# v0.17.2
 
 using Markdown
 using InteractiveUtils
@@ -8,20 +8,22 @@ using InteractiveUtils
 begin
 	using Pkg
 	Pkg.activate("..")
-	using ColorSchemes
-	using CSV
-	using DataFrames
-	using Distributions
-	using Downloads
-	using DrWatson
-	using KernelDensity
-	using PlutoUI
-	using StaticArrays
-	using StatsBase
-	using StatsPlots
-	using Unitful
-	using Turing
-	TableOfContents()
+    using ColorSchemes
+    using CSV
+    using DataFrames
+    using Distributions
+    using Downloads
+    using DrWatson
+    using KernelDensity
+	using Plots
+	using Plots: mm
+    using PlutoUI
+    using StaticArrays
+    using StatsBase
+    using StatsPlots
+    using Unitful
+    using Turing
+    TableOfContents()
 end
 
 # ╔═╡ 3c2d6a02-1f68-48ef-9c9c-3b29a653b2c5
@@ -34,7 +36,7 @@ md"""
 """
 
 # ╔═╡ c57017ea-d32d-4fb0-8cf7-88ffc45edc93
-colors = ColorSchemes.okabe_ito;
+colors = ColorSchemes.okabe_ito
 
 # ╔═╡ 24a51c14-9046-42d0-8999-2b0c5e6ee519
 md"""
@@ -45,10 +47,10 @@ We start by loading the scenarios (𝐬; `\bfs`), decisions (𝐱; `\bfx`), and 
 
 # ╔═╡ d92e3059-9bc5-4993-a38d-6dfad3854ae5
 begin
-	cachefile = datadir("processed", "sows_outcomes.jld2")
-	u = DrWatson.load(cachefile, "u")
-	𝐬 = DrWatson.load(cachefile, "𝐬")
-	𝐱 = DrWatson.load(cachefile, "𝐱")
+    cachefile = datadir("processed", "sows_outcomes.jld2")
+    u = DrWatson.load(cachefile, "u")
+    𝐬 = DrWatson.load(cachefile, "𝐬")
+    𝐱 = DrWatson.load(cachefile, "𝐱")
 end;
 
 # ╔═╡ 4544a2aa-0926-45c9-9082-5f706f7b9273
@@ -60,20 +62,16 @@ md"""
 
 # ╔═╡ 37b2cda6-7d57-404a-a235-fb4ec04f26c4
 begin
[…]

### 2021-12-10 scripts/runall.jl

@@ -1,49 +1,0 @@
-using DrWatson
-using PlutoSliderServer: export_notebook
-using Pluto
-
-NB_DIR = projectdir("notebooks")
-OUTPUT_DIR = projectdir("nb_out")
-
-
-"""Helper function to get the extension of a file"""
-function file_extension(filename)
-    return filename[findlast(isequal('.'), filename):end]
-end
-
-"""Clean up all processed data, plots, and outputs"""
-function clean_outputs()
-    for dir in [datadir("processed"), plotsdir(), OUTPUT_DIR]
-        if isdir(dir)
-            rm(dir; recursive = true)
-        end
-        mkdir(dir)
-    end
-end
-
-"""Helper function to get all the notebooks"""
-function get_all_notebooks()
-    return [
-        joinpath(NB_DIR, file) for file in readdir(NB_DIR) if file_extension(file) == ".jl"
-    ]
-end
-
-"""run Pluto"""
-function run_pluto(; port = 8888)
-    Pluto.run(port = port)
-end
-
-"""run all notebooks and convert to HTML files"""
-function convert_notebooks()
-    if !isdir(OUTPUT_DIR)
-        mkpath(OUTPUT_DIR)
-    end
-    for nb in get_all_notebooks()
-        export_notebook(nb, Export_output_dir = OUTPUT_DIR)
-    end
-end
-
-# run one of the following
-# clean_outputs()
-# run_pluto()
-# convert_notebooks()

### 2021-12-10 papers/agu21/combining.tex
Some more edits from AGU

@@
     If there are only a few PDFs, then qualitative comparisions (\eg, \cref{fig:tradeoffs}) may be sufficient.
-    With \emph{many} PDFs (\eg, \cref{fig:boxplots}), further synthesis is needed.
-
+    With \emph{many} PDFs (\eg, \cref{fig:boxplots}, refs.~\cite{ruckert_coastal:2019,sharma_rcp:2021}), further synthesis is needed.\\[0.5em]
     We consider the \emph{common case} of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
+    A system model $f$ quantifies the outcomes $u$ of taking decision $x$ under SOW $s$.
+    Examining all SOWs created by a given model yields \emph{scenario-conditional optimization} (\cref{fig:tradeoffs}).
+    To synthesize across PDFs, a weighting approach based on Bayesian Model Averaging \cite{wong_surge:2018,Yao:2018bu} is used.
     \begin{figure}
         \centering
-        \includegraphics[width=\textwidth]{bayes-rdm.pdf}
+        \includegraphics[width=0.9\textwidth]{bayes-rdm.pdf}
         \caption{

### 2021-12-10 papers/agu21/doesnt-fit-all.tex
Some more edits from AGU

@@
             \caption{
-                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (L) \SI{2.5}{ft} and (R) \SI{0.5}{ft} below the \gls{bfe} under \gls{rcp} 6.0 with fast BRICK dynamics.
+                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (L) \SI{2.5}{ft} and (R) \SI{0.5}{ft} below the \gls{bfe} under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics.
                 Note $y$-axes are not equal.

### 2021-12-10 papers/agu21/not-equally-plausible.tex
Some more edits from AGU

@@
-\begin{block}{we're always }
+\begin{block}{different beliefs emphasizes different possibilities}
+    The prior model averaging approach (\cref{fig:flowchart}) assigns a weight to each PDF based on their
+    \begin{itemize}
+        \item The prior can't be ``right''  \cite{gelman_workflow:2020,gelman_philosophy:2013} or ``scenario neutral'' \cite{quinn_exploratory:2020}
+        \item let's at least \textbf{make assumptions transparent}!
+    \end{itemize}
     \begin{figure}
@@
         \caption{
-            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution.
-            This may or may not accurately reflect available information.
-        }
+            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution; this may or may not reflect available information.        }
     \end{figure}
@@
         \caption{
-            The prior model averaging approach (\cref{fig:flowchart}) re-weights PDFs based on their distance, and that of other PDFs available, from the (ideally coproduced!) prior.
+            Weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
+            Horizontal line shows the weight given to each under an implicit uniform assumption.
         }

### 2021-12-10 notebooks/03-sow-outcomes.jl

@@ -623,7 +623,7 @@ DrWatson.wsave(datadir("processed", "sows_outcomes.jld2"),
 # ╟─26eb91da-a71b-4eb9-9a8c-a9a336b34446
 # ╟─cf4c37c4-0755-4af3-857c-529cce25583d
 # ╠═a0a7288a-a838-4f05-950a-68088258d7db
-# ╟─fcb1ddc0-e6b8-4847-b994-fce762046f8e
+# ╠═fcb1ddc0-e6b8-4847-b994-fce762046f8e
 # ╠═ef2d2c8d-afda-461e-beab-6ce07b8d8650
 # ╟─ba716c60-0b2e-4abe-b195-afa49dbe63d8
 # ╟─9f8016bd-c104-4d0c-9e5f-61a495f4fdbb

### 2021-12-10 notebooks/04-outcomes-tradeoffs.jl

@@ -86,7 +86,7 @@ begin
             xlabel="MSL at Norfolk, VA in 2100 [ft]",
             ylabel="Probability Density",
             xlims=(0, 8),
-            size=(600, 300),
+            size=(700, 300),
         )
         for (i, prior) in enumerate(priors)
             plot!(p, prior; label=prior_names[i], c=colors[i], linewidth=2)
@@ -236,7 +236,7 @@ end;
 # ╔═╡ 95add889-a398-4bae-a3cb-37ca7c620644
 begin
     pt3 = plot_tradeoff(scenario_weights, 1, 4; size=(500, 500))
-    savefig(pt3, plotsdir("tradeoffs_height_totalcost_byscenario.pdf"))
+	savefig(pt3, plotsdir("tradeoffs_height_totalcost_byscenario.pdf"))
     pt3
 end
 
@@ -333,7 +333,7 @@ begin
 		bottom_margin=10mm,
 		left_mm=4mm,
 		yaxis = "Weight Assigned to Each PDF",
-		size = (600, 300),
+		size = (700, 300),
 	)
 	groupedbar!(
 		p,

### 2021-12-10 papers/agu21/case-study.tex
As printed for AGU!

@@
 \begin{block}{case study: norfolk, va}
-    We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard, following the approach of \cite{zarekarizi_suboptimal:2020}.
+    We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard, following the approach of ref.~\cite{zarekarizi_suboptimal:2020}.
+    \begin{framed}
         \begin{figure}
@@
             \caption{
-            Time series of annual maximum of storm surge (after subtracting mean sea level) at Sewells Point, VA (NAVD datum).
+                Time series of annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+                Purple (orange) arrows denote notable tropical cyclones (nor'easters).
             }
@@
         \end{figure}
+    \end{framed}
+    \vspace*{-1em}
+    \begin{framed}
         \begin{figure}
@@
             \caption{
-            Neglecting hydrodynamics, we add sea level rise (a; see ref.~\cite{ruckert_coastal:2019}) to a Bayesian GEV model of storm surge (b) to get flood hazard.
-            The convolution of hazard and fragility (c; see ref.~\cite{zarekarizi_suboptimal:2020}) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
+                Neglecting hydrodynamics, we add sea level rise (a) to a Bayesian GEV model of storm surge (b) to get flood hazard.
+                The convolution of hazard and fragility (c) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
             }
         \end{figure}
+    \end{framed}
 \end{block}

### 2021-12-10 papers/agu21/combining.tex
As printed for AGU!

--- 2021-12-10 papers/agu21/combining.tex
@@
 \begin{block}{synthesizing pdfs for decision relevance}
-    If there are only a few PDFs, then qualitative comparisions (\eg, \cref{fig:tradeoffs}) may be sufficient.
-    With \emph{many} PDFs (\eg, \cref{fig:boxplots}), further synthesis is needed.
-
-    We consider the \emph{common case} of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
+    While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability (see refs.~\cite{schneider_dangerous:2001,schneider_scenarios:2002}).
+    \begin{framed}
         \begin{figure}
             \centering
-        \includegraphics[width=\textwidth]{bayes-rdm.pdf}
+            \includegraphics[width=0.8\textwidth]{bayes-rdm.pdf}
             \caption{
-            Sketch of the ``Prior Model Averaging'' framework.
+                ``Prior Model Averaging'' framework.
             }\label{fig:flowchart}
         \end{figure}
+    \end{framed}
+    Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
+    A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
+    To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \cite{wong_surge:2018,Yao:2018bu}.
 \end{block}

### 2021-12-10 papers/agu21/doesnt-fit-all.tex
As printed for AGU!

--- 2021-12-10 papers/agu21/doesnt-fit-all.tex
@@
-\begin{block}{one size doesn't fit all!}
+\begin{block}{one size does not fit all!}
     \begin{itemize}
         \item Tailoring guidance to specific building characteristics can improve outcomes
-        \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \cite{xian_elevation:2017,zarekarizi_suboptimal:2020}
+        \item Elevating to ``\acrshort{bfe} plus a foot'' is not always optimal \cite{xian_elevation:2017,zarekarizi_suboptimal:2020}
         \item Both over- and under-building can be costly \cite{ansar_bigisfragile:2017,DossGollin:2019}
@@
             \centering
-            \includegraphics[width=0.475\textwidth]{5.5.pdf}%
-            \includegraphics[width=0.475\textwidth]{7.5.pdf}
+            \begin{subfigure}{0.48\textwidth}
+                \subcaption{\SI{2.5}{ft} below BFE}
+                \includegraphics[width=\textwidth]{5.5.pdf}
+            \end{subfigure}
+            \hfill
+            \begin{subfigure}{0.48\textwidth}
+                \subcaption{\SI{0.5}{ft} below BFE}
+                \includegraphics[width=\textwidth]{7.5.pdf}
+            \end{subfigure}
             \caption{
-                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (L) \SI{2.5}{ft} and (R) \SI{0.5}{ft} below the \gls{bfe} under \gls{rcp} 6.0 with fast BRICK dynamics.
-                Note $y$-axes are not equal.
+                Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics).
+                Note unequal $y$-axis scales.
             }

### 2021-12-10 papers/agu21/house-elevation.tex
As printed for AGU!

@@
     Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \cite{zarekarizi_suboptimal:2020,xian_elevation:2017}.
-    \begin{framed}
+    \vspace*{-1em}
     \begin{description}
@@
         \item[Q2] How does nonstationary hazard change guidance?
-            \item[Q3] Which model of nonstationary hazard should be used?
+        \item[Q3] Which model of nonstationary hazard to use?
     \end{description}
-    \end{framed}
+    \begin{framed}
         \begin{figure}
@@
             \subcaptionbox{Selinsgrove, PA}{\includegraphics[height=3.35in]{img_selinsgrove.jpeg}}%
-        \hfill
+            \hfill%
             \subcaptionbox{Houston, TX}{\includegraphics[height=3.35in]{img_braeswood.png}}\\
             \subcaptionbox{Plaquemines Parish, LA}{\includegraphics[height=3.3in]{epstein-2020-la.jpeg}}%
-        \hfill
+            \hfill%
             \subcaptionbox{Dare County, NC}{\includegraphics[height=3.3in]{nichols-ocracoke.jpeg}}%
             \caption{
-            \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols.
+                \faIcon{camera}: (a) JDG, (b) Google Maps, (c) Mitch Epstein /  New York Times (d) Rob Nichols / PSU.
             }
@@
         \end{figure}
+    \end{framed}
+    These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
 \end{block}

### 2021-12-10 papers/agu21/multiple-pdf.tex
As printed for AGU!

@@
 \begin{block}{the multiple pdf problem}
-    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie the \textbf{\itshape{Multiple PDF Problem}}.
+    Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
+    Including additional models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
     \begin{framed}
@@
             \caption{
-                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017}.
+                Sea level rise projections use the BRICK model \cite{wong_brick0.2:2017,ruckert_coastal:2019}.
                 We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
@@
     \end{framed}
-    Considering other models \cite{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} would \emph{compound} this problem.
-    \textbf{Which PDF should long-lasting infrastructure like housing be designed for?}
 \end{block}

### 2021-12-10 papers/agu21/not-equally-plausible.tex
As printed for AGU!

--- 2021-12-10 papers/agu21/not-equally-plausible.tex
@@
-\begin{block}{we're always }
+\begin{block}{key message}
+    \vspace*{-1em}
+    \begin{framed}
         \begin{figure}
             \centering
-        \includegraphics[width=\textwidth]{lsl_priors.pdf}
+            \includegraphics[width=0.48\textwidth]{lsl_priors.pdf}~
+            \includegraphics[width=0.48\textwidth]{tradeoffs_height_totalcost_byprior.pdf}~
             \caption{
-            If we weight all scenarios equally, then our projection of MSL in 2100 follows the sampling distribution.
-            This may or may not accurately reflect available information.
+                The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
+                Alternative priors can provide robustness checks.
             }
         \end{figure}
+    \end{framed}
+    \begin{itemize}
+        \item Natural extensions to Bayesian updating
+        \item Belief $p(s)$ can't be ``right''  \cite{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \cite{quinn_exploratory:2020}
+        \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
+    \end{itemize}
+    \begin{framed}
         \begin{figure}
@@
             \caption{
-            The prior model averaging approach (\cref{fig:flowchart}) re-weights PDFs based on their distance, and that of other PDFs available, from the (ideally coproduced!) prior.
+                The weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
             }
         \end{figure}
+    \end{framed}
+    ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \cite{hausfather_scenarios:2020}.
 \end{block}

### 2021-12-10 papers/agu21/other-applications.tex
As printed for AGU!

@@
         \item More complex models to capture more relevant metrics
-        \item Better priors over nonstationary hazard
+        \item Better models of nonstationary hazards
         \item Interacting, sequential decisions
-        \item Inclusive assessment of ethical questions around scenario weighting
     \end{itemize}
@@
             \caption{
-                House elevation is just one of many design problems where nonstationary hazards, subject to scenario and model structure uncertainties, drive outcomes.
+                House elevation is just one of many design problems where nonstationary hazards, subject to scenario nd model structure uncertainties, drive outcomes.
                 \faIcon{camera}: (a) JDG (b) D. Howard / Wikipedia.
             }
-            \label{fig:images}
+            \label{fig:stormwater-bridge}
         \end{figure}

### 2021-12-10 papers/agu21/scenario-matters.tex
As printed for AGU!

@@
 \begin{block}{it matters which pdf you choose!}
-    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes (see ref.~\cite{gelman_philosophy:2013} for a more philosophical discussion).
+    Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
     \begin{framed}
@@
             \caption{
-                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the model selected.
+                Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
             }

### 2021-12-10 notebooks/01-surge-modeling.jl

@@ -1,5 +1,5 @@
 ### A Pluto.jl notebook ###
-# v0.17.3
+# v0.17.2
 
 using Markdown
 using InteractiveUtils
@@ -810,7 +810,7 @@ begin
 			legend = :topleft,
 			xticks = (xticks, string.(xticks)),
 			dpi=250, # for saving
-			size=(500, 500),
+			size=(400, 400),
 		)
 		
 		# start with Bayesian posterior draws

### 2021-12-10 notebooks/02-mean-sea-level.jl

@@ -176,7 +176,7 @@ begin
 			legend=:topleft,
 			xlabel="Year",
 			ylabel="Mean Sea Level at Norfolk, VA [ft]",
-			size = (500, 500),
+			size = (400, 400),
 		)
 		models = ((2.6, "slow"), (4.5, "slow"), (6.0, "fast"), (8.5, "fast"))
 		for (model, color) in zip(models, colors)
@@ -396,7 +396,7 @@ end
 # ╟─2022a380-290f-45fb-930c-e0b00d1806ee
 # ╟─37d98257-5de6-456a-a542-36d0dc1c7a96
 # ╟─f251e057-4409-4c8b-a2d7-4603c8af6750
-# ╟─c73ddb13-50d4-4e6d-ac54-c171c66b3786
+# ╠═c73ddb13-50d4-4e6d-ac54-c171c66b3786
 # ╠═5f912aeb-c9e4-485a-897f-a71e84ec6038
 # ╟─eece84cd-3f44-4d8b-851e-933d22151c19
 # ╟─f815e33f-97fa-4a1b-b8fc-77241d81d172

### 2021-12-10 notebooks/04-outcomes-tradeoffs.jl

@@ -86,7 +86,8 @@ begin
             xlabel="MSL at Norfolk, VA in 2100 [ft]",
             ylabel="Probability Density",
             xlims=(0, 8),
-            size=(700, 300),
+            size=(400, 400),
+			bottom_margin=5mm,
         )
         for (i, prior) in enumerate(priors)
             plot!(p, prior; label=prior_names[i], c=colors[i], linewidth=2)
@@ -210,7 +211,6 @@ function plot_tradeoff(weights, kx, ky; size = (400, 400), kwargs...)
     p = plot(;
 		xlabel = mnames[kx],
 		ylabel = mnames[ky],
-		top_margin = 10mm,
 		size = size,
 		kwargs...
 	)
@@ -235,21 +235,21 @@ end;
 
 # ╔═╡ 95add889-a398-4bae-a3cb-37ca7c620644
 begin
-    pt3 = plot_tradeoff(scenario_weights, 1, 4; size=(500, 500))
+    pt3 = plot_tradeoff(scenario_weights, 1, 4; legend=:topright)
 	savefig(pt3, plotsdir("tradeoffs_height_totalcost_byscenario.pdf"))
     pt3
 end
 
 # ╔═╡ 37f159ec-1cd2-4463-8862-e39bae53e9cb
 begin
-    pt4 = plot_tradeoff(scenario_weights, 1, 3; size=(500, 500), legend=:topright)
+    pt4 = plot_tradeoff(scenario_weights, 1, 3; legend=:topright)
     savefig(pt4, plotsdir("tradeoffs_height_damages_byscenario.pdf"))
     pt4
 end
 
 # ╔═╡ 1fd21800-92cb-4b8b-bd4b-3c4afa73b43a
 begin
-    pt6 = plot_tradeoff(scenario_weights, 1, 5; size=(500, 500), legend=:topright)
+    pt6 = plot_tradeoff(scenario_weights, 1, 5; legend=:topright)
     savefig(pt6, plotsdir("tradeoffs_height_iqr_byscenario.pdf"))
     pt6
 end
@@ -278,21 +278,21 @@ prior_weights = [(name=n, w=w) for (n, w) in zip(prior_names, weights)];
 
[…]

### 2022-01-09 papers/paper/preprint.tex
Get the stuff from poster into preprint -- now need to outline and flesh out

--- 2021-10-21 papers/operationalizing-bayesian-checking.tex
@@
 \usepackage{graphicx}
-\usepackage{wrapfig}
-\graphicspath{{../plots/}{../tikz/}{../img/}}
+\usepackage[list=true]{subcaption}
+\graphicspath{{../../plots/}{../../tikz/}{../../img/}}
 \usepackage[section]{placeins} % require floats to appear in the section they are defined
@@
 \usepackage[ttscale=0.8]{libertine}
+\usepackage{fontawesome5}
 
@@
 \def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
+\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
+\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
+\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+
+\usepackage{xspace}
+\makeatletter
+\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
+\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
 \newcommand{\usd}[1]{\SI{#1}[\$]{}}
@@
 \AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
-%\addbibresource{../_bibliography/library.bib}
 \addbibresource{library.bib}
@@
 
-\section*{Front Matter}
+\clearpage
 
-\subsection*{Submission Plan}
-We should revisit this!
-\begin{enumerate}
-    \item Earth's Future
-    \item Climatic Change
-    \item IOP Environmental Research: Infrastructure and Sustainability (new journal)
-    \item Risk analysis
-\end{enumerate}
-\subsection*{Internal peer review}
-\begin{enumerate}
-    \item Vivek
-    \item Courtney
-    \item Sanjib
-    \item Kelsey to replicate code
-\end{enumerate}
+\section{Introduction}
+
+Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \citep{zarekarizi_suboptimal:2020,xian_elevation:2017}.
+\begin{description}
+    \item[Q1] How to adapt guidance to building characteristics or household preferences?
+    \item[Q2] How does nonstationary hazard change guidance?
+    \item[Q3] Which model of nonstationary hazard to use?
+\end{description}
+These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
+
+The paper is organized as follows.
+In \cref{sec:case-study} we describe our case study in Norfolk, VA and present results to illustrate the ``Multiple PDF problem'' and motivate our methodological advances.
+Then in \cref{sec:prior-averaging}, we present a method for combining multiple \glspl{pdf}.
+Finally in \cref{sec:conclusions} we summarize our findings, discuss implications for engineering practice, and identify needs for future research.
+
+\section{Case study}\label{sec:case-study}
+
+We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.
+
+\subsection{Model structure and data}
+
+We follow the general modeling approach of \citet{zarekarizi_suboptimal:2020} to model expected annual damages for a particular house.
+
+\subsubsection{Multiple PDFs}
+
+\subsubsection{Candidate decisions}
+
+\subsubsection{States of the World (SOWs)}
+
+\subsubsection{System model}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{historic_surge.pdf}
+    \caption{
+        Time series of annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
+    }
+    \label{fig:observations}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{hazard-tradeoffs.pdf}
+    \caption{
+        Neglecting hydrodynamics, we add sea level rise (a) to a stationary Bayesian GEV model of storm surge (b) to get flood hazard.
+        The convolution of hazard and fragility (c) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
+    }
+\end{figure}
+
+\subsection{One size does not fit all}
+
+\begin{itemize}
+    \item Tailoring guidance to specific building characteristics can improve outcomes
+    \item Elevating to ``\acrshort{bfe} plus a foot'' is not always optimal \citep{xian_elevation:2017,zarekarizi_suboptimal:2020}
+    \item Both over- and under-building can be costly \citep{ansar_bigisfragile:2017,DossGollin:2019}
+\end{itemize}
+\begin{figure}
+    \centering
+    \begin{subfigure}{0.3\textwidth}
+        \subcaption{\SI{2.5}{ft} below BFE}
+        \includegraphics[width=\textwidth]{5.5.pdf}
+    \end{subfigure}~
+    \begin{subfigure}{0.3\textwidth}
+        \subcaption{\SI{0.5}{ft} below BFE}
+        \includegraphics[width=\textwidth]{7.5.pdf}
+    \end{subfigure}
+    \caption{
+        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \citep{wong_brick0.2:2017} dynamics).
+        Note unequal $y$-axis scales.
+    }
+\end{figure}
+
+\subsection{The multiple PDF problem}
+
+Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
+Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
+\begin{figure}
+    \centering
+    \includegraphics[width=5in]{msl_boxplots.pdf}
+    \caption{
+        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
+        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
+    }\label{fig:boxplots}
+\end{figure}
+
+\subsection{It matters which PDF you choose}
+
+Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
+\begin{figure}
+    \centering
+    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
+    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
+    \caption{
+        Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
+    }
+    \label{fig:tradeoffs}
+\end{figure}
+Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.
+
+\section{Synthesizing PDFs for decision relevance}\label{sec:prior-averaging}
+
+\subsection{Methods}
+
+While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{bayes-rdm.pdf}
+    \caption{
+        ``Prior Model Averaging'' framework.
+    }\label{fig:flowchart}
+\end{figure}
+Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
+A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
+To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \citep{wong_surge:2018,Yao:2018bu}.
+
+\subsection{Findings}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{inference_weights.pdf}
+    \caption{
+        The weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
+    }
+\end{figure}
+``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.
+
+\section{Conclusions}\label{sec:conclusions}
+
+To summarize:
+\begin{itemize}
+    \item Natural extensions to Bayesian updating
+    \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
+    \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
+\end{itemize}
+
+\subsection{Implications for policy}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=3in]{lsl_priors.pdf}~
+    \includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
+    \caption{
+        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
+        Alternative priors can provide robustness checks.
+    }
+\end{figure}
+
+\subsection{Future needs}
+This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
+\textbf{\scshape let's collaborate on:}
+\begin{itemize}
+    \item More complex models to capture more relevant metrics
+    \item Better models of nonstationary hazards
+    \item Interacting, sequential decisions
+\end{itemize}
 
@@
     \centering
-    \includegraphics[height=2in]{uconn-elevateA}
+    %\includegraphics[height=2in]{uconn-elevateA}
     \caption{
@@
     \centering
-    \includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
+    %\includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
     \caption{
@@
     \centering
-    \includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
+    %\includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
     \caption{

### 2022-01-18 papers/paper/preprint.tex
working in comments and filling in new organization

@@
 \fancyhf{}
-\fancyhead[L]{DRAFT IN PROGRESS -- Doss-Gollin \& Keller}
+\fancyhead[L]{CONFIDENTIAL DRAFT -- Doss-Gollin \& Keller}
 \fancyhead[R]{\DTMnow}
@@
 \usepackage{authblk}
-\title{Which Scenario Should We Design For? Insights from House Elevation on the Multiple PDF Problem}
+\title{Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem}
 \author[1]{James Doss-Gollin}
@@
 \newacronym{gcm}{GCM}{general circulation model}
+\newacronym[]{gev}{GEV}{generalized extreme value}
 \newacronym{lsl}{LSL}{local mean sea level}
 \newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
-\newacronym{mordm}{MORDM}{multiobjective \gls{rdm}}
 \newacronym{pdf}{PDF}{probability density function}
 \newacronym{rcp}{RCP}{representative concentration pathway}
-\newacronym{rslr}{RSLR}{relative sea level rise}
+\newacronym{slr}{SLR}{relative sea level rise}
 \newacronym{rdm}{RDM}{robust decision making}
@@
 \begin{abstract}
-    Many homeowners in the coastal zone elevate their houses to manage flood risk.
-    Federal guidance for this decision relies on floodplain maps that are silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards.
-    This uncertainty poses challenges for the design of decision-support systems.\james{If this is about RDM then let's lead with that.}
-
-    Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
-    Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
-    We introduce an approach to bridge this divide through a computationally efficient method that weights each state of the world based on a probability distribution over possible futures.
-    Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
-    This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.
+    TBD
 \end{abstract}
@@
 
-\section{Introduction}
+\section{Introduction}\label{sec:introduction}
 
-Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \citep{zarekarizi_suboptimal:2020,xian_elevation:2017}.
+Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face far into the future.
+On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
+For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
+On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
+For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
+
+TODO: set up flood risk management as a context where over- or under-design could be relevant, then raise house elevation in the coastal zone as a specific example.
+Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
+Be sure to note that this has parallels to lots of other engineering design guidance.
+Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve significantly on a standards-based approach.
+
+Additionally, the floodplain maps used for decision supportare silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
+Both risk-based design and standards-based approaches are complicated by nonstationarity.
+On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
+On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[\eg,][]{keller_management:2021}.
+
+This suggests that either nonstationarity should be ignored, or subjective approaches for designing for nonstationarity should be adapted.
+Given that physical theories consistent with the observational record and models robustly project changes to climate hazard under warning, ignoring nonstationarity seems suboptimal.
+Thus decision-makers need some subjective way to choose which possible futures to design for.
+
+In this paper, we use a didactic case study of house elevation in Norfolk, VA as a lens through which to shed light on the following research questions.
 \begin{description}
-    \item[Q1] How to adapt guidance to building characteristics or household preferences?
-    \item[Q2] How does nonstationary hazard change guidance?
-    \item[Q3] Which model of nonstationary hazard to use?
+    \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
+    \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
+    \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
 \end{description}
-These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
+Specifically, we examine this problem through three lenses.
 
 The paper is organized as follows.
-In \cref{sec:case-study} we describe our case study in Norfolk, VA and present results to illustrate the ``Multiple PDF problem'' and motivate our methodological advances.
-Then in \cref{sec:prior-averaging}, we present a method for combining multiple \glspl{pdf}.
-Finally in \cref{sec:conclusions} we summarize our findings, discuss implications for engineering practice, and identify needs for future research.
+In \cref{sec:case-study} we describe our case study in Norfolk, VA.
+First, in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
+Next, in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
+Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
+Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify needs for future research.
 
@@
 
-\subsection{Model structure and data}
+For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
+We consider a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
+We evaluate this decision for each of many possible simulations of \gls{slr} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
+In the remainder of this section we discuss data sources and the system model.
 
-We follow the general modeling approach of \citet{zarekarizi_suboptimal:2020} to model expected annual damages for a particular house.
-
-\subsubsection{Multiple PDFs}
-
-\subsubsection{Candidate decisions}
-
-\subsubsection{States of the World (SOWs)}
-
-\subsubsection{System model}
+\begin{table}
+    \centering
+    \begin{tabular}{l l p{3in} l}
+        \toprule
+        Name           & Symbol           & Conceptual framework                                                                 & Section               \\
+        \midrule
+        \Gls{slr}      & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
+        Storm surge    & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
+        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
+        Depth-damage   & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
+        Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
+        Initial height & $h_0$            & Deterministic: equal to \gls{bfe} unless otherwise noted                             & TBD                   \\
+        \bottomrule
+    \end{tabular}
+    \caption{Uncertainties modeled}\label{tab:uncertainties}
+\end{table}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{historic_surge.pdf}
+    \includegraphics[width=4in]{xlrm.pdf}
     \caption{
-        Time series of annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
-        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
+        Conceptual diagram of the decision problem.
+        A \gls{sow} consists of a description of the uncertain factors (red).
+        We model a single decision (yellow).
+        For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
     }
-    \label{fig:observations}
 \end{figure}
+
+\subsection{Sea level rise}\label{sec:sea-level}
+
+Deeply uncertain sea level rise is the focus of our model.
+Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
+Specifically, we use experiments using the BRICK model version 0.2 \citep{wong_brick0.2:2017} simulations published in \citet{ruckart_hazardous:2008}.
+There are eight \glspl{pdf}, each with ZXX\james{count} simulations: four emissions pathways (\gls{rcp} scenarios) $\times$ two models of ice sheet dynamics.
+
+For consistency with literature on decision making under deep uncertainty \citep[see][]{marchau:2019}, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.
+
+\subsection{Storm surges}\label{sec:storm-surge}
+
+Neglecting hydrodynamics, we model annual maximum floods $y_t$ as the sum of sea level $\overline{y}_t$ and storm surges $y'_t$.
+
+Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite}
+Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
+For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
+This time series is shown in \cref{fig:observed-surges}(a).\james{This figure needs a bit of work: subpanel plots, equalizing axes, narrower time series, and no lines between the points}
 
@@
     \centering
-    \includegraphics[width=6in]{hazard-tradeoffs.pdf}
+    \includegraphics[height=1.5in]{historic_surge.pdf}%
+    \includegraphics[height=1.5in]{surge_return_period.png}
     \caption{
-        Neglecting hydrodynamics, we add sea level rise (a) to a stationary Bayesian GEV model of storm surge (b) to get flood hazard.
-        The convolution of hazard and fragility (c) yields an assessment of damages and tradeoffs (\cref{fig:tradeoffs}).
+        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+        (a):
+        time series of historic storms.
+        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
+        (b):
+        return periods.
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull rule.
+        Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
     }
+    \label{fig:observed-surges}
 \end{figure}
+
+We model future storm surge risk through a probabilistic lens.
+To do this, we develop a stationary Bayesian model for storm surge:
+\begin{equation}\label{eq:surge-model}
+    y'_t \sim \text{GEV}\left(\mu, \sigma, \xi \right),
+\end{equation}
+where $y'_t$ is the storm surge (above sea level) in year $t$.
+
+When GEV distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
+Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
+To constrain estimates, we add weakly informative priors.
+However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
+Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).
+These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}
+
+To evaluate a single \gls{sow}, we integrate over $J$ samples from the posterior distribution (of \cref{eq:surge-model}).
+
+\subsection{Expected annual damage}\label{sec:ead}
+
+There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
+The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
+Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
+\begin{equation}
+    \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
+\end{equation}
+where $y_t = \overline{y}_t + y'_t$ is the annual maximum flood, relative to the gauge, in year $t$, and $D(h - y_t)$ is the damage as a function of depth relative to the house.
+Following \citet{zarekarizi_suboptimal:2020} we use the HAZUS deterministic depth-damage relationship to parameterize $D(h-y_t)$ (see fig. XYZ(c)).\james{Cite HAZUS}
+
+The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
+However, the convolution of the HAZUS depth-damage equation with the GEV posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
+\begin{enumerate}
+    \item For $k=1, \ldots, K$:
+          \begin{enumerate}
+              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
+              \item simulate a single storm surge from the stationary distribution (equation XYZ) and add the mean sea level to get total flood depth
+              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_t$) into  the deterministic HAZUS depth-damage relationship
+              \item store this as the $k$th damage
+          \end{enumerate}
+    \item Estimate expected annual damages as the sample mean of the $K$ estimates
+\end{enumerate}
+
+However, evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
+Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
+However, noticing that this function dependsonly on the elevation of the house relative to \gls{slr}, we develop a simuple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
+To do this, we  precompute expected annual damage for all height differences in \SI{0.1}{ft} increments from -XY to YX ft and fit a piecewise linear interpolation to this data.\james{check increements, piecewise linear}
+We use $K=\num{1e6}$ samples to fit this emulator for each of the NUMBER increments.
+This model is shown in supplemental figure SX.\james{supplemental figure}
+Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
+
+\subsection{Lifetime expected damages}\label{sec:led}
+
+The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the upfront cost plus the discounted sum of expected annual damages
+\begin{equation}\label{eq:led}
+    \textrm{LED} = \sum_{t=1}^T \gamma^{t-1} \mathbb{E}[d_t | \Delta h],
+\end{equation}
+where $\gamma$ is the 1 minus the discount rate.
+For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2, thus $\gamma = 0.98$.
+Neglecting uncertainty in house lifetime, we model flood damages from 2022 to 2071, thus $T=50$.\james{This has not been the case thus far}
+We calculate lifetime expected damages \emph{for each \gls{sow} separately}.
+
+\subsection{Assessing performance}\label{sec:sow-metrics}
+
+To assess the performance of a given decision for a specific \gls{sow}, we calculate three metrics for each \gls{sow}.
+\begin{enumerate}
+    \item The first metric considered is ``upfront cost,'' which reflects the cost of elevating a house.
+Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
+We normalize this cost by house value.\james{Normalize!}
+    \item The second metric is the lifetime expected damages (\cref{eq:led})
+    \item The third metric is lifetime expected costs, which is the sum of lifetime expected damages and upfront costs
+    \item The third metric is the expected benefit-to-cost ratio.  We define this ratio in terms of the ratio of construction cost to damages averted relative to not elevating at all:
+          \begin{equation}
+              \mathrm{BCR}(\Delta h) = \begin{cases} \Delta h = 0 & 1 \\ \Delta h \neq 0 & \frac{\mathbb{E}[d_t | \Delta h = 0] - \mathbb{E}[d_t | \Delta h = \Delta h]}{C(\Delta h)}\end{cases}
+          \end{equation}
+\end{enumerate}
+
+\section{Exploring possible states of the world}\label{sec:multiple-simulation}
+
+The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \acrfullpl{sow}.
+
+In \cref{fig:scenario-map}, we plot each possible future as a single dot for each of several possible amounts by which to heighten the house.
+On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
+On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
+Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ.
+
+\begin{figure}
+    \includegraphics[width=\textwidth]{scenario_maps.png}
+    \caption{
+        Scenario maps: one dot represents one \gls{sow}.
+    }\label{fig:scenario-map}
+\end{figure}
+
+This plot sheds light on the question: ``how well does a given decision perform under different possible futures?''
+The best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
+However, the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
+When $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases.
+This plot suggests that $\Delta h = \SI{8}{ft}$ dominates $\Delta h = \SI{14}{ft}$ because \SI{8}{ft} is sufficient to make expected annual damages negligible at a substantially lower construction cost.
+
+Yet important questions remain -- since the \glspl{sow} are not draws from the full \gls{pdf}
+
+\section{The multiple PDF problem}\label{sec:multiple-pdf}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
     \item Tailoring guidance to specific building characteristics can improve outcomes
-    \item Elevating to ``\acrshort{bfe} plus a foot'' is not always optimal \citep{xian_elevation:2017,zarekarizi_suboptimal:2020}
+    \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \citep[this is consistent with][]{zarekarizi_suboptimal:2020,xian_elevation:2017}
     \item Both over- and under-building can be costly \citep{ansar_bigisfragile:2017,DossGollin:2019}
-\end{itemize}
+\end{enumerate}
+
 \begin{figure}
     \centering
-    \begin{subfigure}{0.3\textwidth}
+    \begin{subfigure}{0.48\textwidth}
         \subcaption{\SI{2.5}{ft} below BFE}
         \includegraphics[width=\textwidth]{5.5.pdf}
-    \end{subfigure}~
-    \begin{subfigure}{0.3\textwidth}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.48\textwidth}
         \subcaption{\SI{0.5}{ft} below BFE}
@@
     \caption{
-        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \citep{wong_brick0.2:2017} dynamics).
+        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics).
         Note unequal $y$-axis scales.
@@
 
-\subsection{The multiple PDF problem}
+\subsection{It matters which PDF you design for}
 
@@
 
-\subsection{It matters which PDF you choose}
+Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
 
-Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
 \begin{figure}
@@
 \end{figure}
-Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.
 
-\section{Synthesizing PDFs for decision relevance}\label{sec:prior-averaging}
+Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.\james{TODO: create a figure showing expected outcomes if you design for scenario $x$ and get scenario $y$}
+
+\section{Synthesizing across PDFs for decision relevance}\label{sec:synthesizing}
 
@@
 
-\section{Conclusions}\label{sec:conclusions}
-
-To summarize:
-\begin{itemize}
-    \item Natural extensions to Bayesian updating
-    \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
-    \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
-\end{itemize}
-
-\subsection{Implications for policy}
-
 \begin{figure}
@@
 
-\subsection{Future needs}
+\section{Conclusions}\label{sec:conclusions}
+
+To summarize:
+\begin{itemize}
+    \item Natural extensions to Bayesian updating
+    \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
+    \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
+\end{itemize}
+
 This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
@@
 \clearpage
-\section{Introduction}
-
-\paragraph{Why this problem is important}
-Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face far into the future.
-On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
-On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
-For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}.}
-
-\paragraph{Set up the ``Multiple PDF Problem''}
-While classical utility- or regret-based frameworks offer a structure for weighing these risks, they are not designed to handle uncertainties that are \emph{deep} in the sense that stakeholders may not agree upon their probabilities \citep{lempert_complex:2002} or \emph{irreducible} in the sense that there is not an objective truth that better models and data can identify \citep{DossGollin:2019}.\footnote{Unless we believe everything is pre-determined, in which case why bother with climate adaptation\ldots}
-Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S.
-federal government will continue to subsidize flood insurance for homeowners.\james{Add some references}
-In response, a wide range of fields have developed scenario-based methods\ldots.\klaus{Which arguments in \citet{savage:1954} did you think should be cited here?}
-This leads to the ``Multiple PDF Problem'' \citep{sharma_rcp:2021}\ldots.
-A key question, then, facing the designers, builders, regulators, owners, and users of built facilities is \emph{to which scenario should infrastructure be designed?}
-
-\paragraph{Set up house elevation as an example}
-One example is building codes and guidelines\ldots.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
-Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
-\citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
-Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
-
-\paragraph{Nonstationarity motivates the case study}
-Nonstationarity is particularly relevant in coastal communities where \gls{rslr} is expected to drive future flood risk.
-As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch},\james{Need to revisit this picture} given stochastic flood peaks and deeply uncertain \gls{rslr}.
-\Gls{rslr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots\klaus{You wrote something here that I can't read -- I see something that looks like ICM.
-I assume Tony and Kelsey papers are relevant as well?}) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
-Two key uncertainties identified by previous studies are
-\begin{enumerate*}[label=(\roman*)]
-    \item which emissions pathway (\eg, \gls{rcp} scenario) will we experience and
-    \item how will polar ice sheets respond to warming \citep{wong_brick0.2:2017,ruckert_coastal:2019,wong_nola:2017,deconto_antarctica:2016}.
-\end{enumerate*}
-Simulations of \gls{rslr} are available to decision makers grouped by \gls{rcp} scenario and model structure.
-For a given combination of \gls{rcp} scenario and model, results are typically presented and interpreted as probabilistic, but this is not the same as giving decision-makers a single \gls{pdf}.
-This is the ``Multiple \gls{pdf} Problem'' described in, \eg, \citet{sharma_rcp:2021}.
-Thus, \emph{which scenario/\gls{pdf} should we design for?}
-
-\paragraph{Needs work}
-Lacking a crystal ball capable of telling us precisely which \gls{pdf} we need to design for, we have to consider several as plausible.
-Any decision will perform differently in different scenarios.
-There is no objective answer to this question.
-Since we can't be ojectively correct, we might as well be transparent and make our assumptions explicit (and subject to critique) rather than implicit.
-
-\paragraph{Roadmap}
-We first present a general framework for\ldots.
-We then illustrate this method through a didactic case study of deciding if, and how high, to elevate a single house in Norfolk, VA.
-
-\begin{figure}
-    \centering
-    %\includegraphics[height=2in]{uconn-elevateA}
-    \caption{
-        Conceptual sketch illustrating the path from historical data to FEMA's guidance.
-        This figure \emph{needs to be revised} (KATERINA?) to have three panels, of which this is the third.
-        (a): A time series of historical floods.
-        (b): A \gls{pdf} of the flood peak.
-        (c): as shown, but with waves eliminated.
-        This sketch figure is from \citet{patten_elevate:2016}.
-    }\label{fig:house-sketch}
-\end{figure}
-
-\section{Decision framework}\label{sec:framework}
 
@@
 In \cref{sec:methods-iterative} we discuss how this framework can be integrated into frameworks for iterative engagement with experts and stakeholders to expand and frame the model.
-
-\begin{figure}
-    \centering
-    %\includegraphics[width=\textwidth, height=4.5in, keepaspectratio=true]{conceptual.pdf}
-    \caption{
-        Schematic of our decision support framework.
-        Nodes in green are modeling choices that the modeler must make while those in yellow are computed quantities.
-        Teal text illustrates a non-exhaustive set of opportunities to perform model checks and/or engage with stakeholders to iterate on design of the decision support system.
-        Feedbacks between these questions and the green nodes are omitted for clarity.
-        Our approach could be extended by coupling the evaluation framework to an optimization or policy search tool that updates the candidate decisions.
-        This approach could also be extended to a stochastic system model.
-    }\label{fig:framework}
-\end{figure}
 
@@
 
-\gls{rslr}:
+\gls{slr}:
 \begin{enumerate}
@@
 \paragraph{Prior on outputs only}
-Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{rslr} in 2100.\klaus{You suggest looking at another paper.
+Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{slr} in 2100.\klaus{You suggest looking at another paper.
 Are you referring to this?\\\fullcite{sriver_sealevel:2018}}
@@
 The idea of embracing subjective choices -- and making them explicit and open to critique -- draws heavily from philosophies of iterative workflow in statistics \citep{box:1976,gelman_workflow:2020,gelman_philosophy:2013}.
-              \item This approach can be used in a multiobjective context and is an alternative way to measure robustness in \gls{rdm} and \gls{mordm}
+              \item This approach can be used in a multiobjective context and is an alternative way to measure robustness in \gls{rdm} and MORDM
           \end{enumerate}
@@
 
+\section{Old material}
+
+\clearpage
+\newpage
+
+Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
+Be sure to note that this has parallels to lots of other engineering design guidance.
+Then note the work of \citet{zarekarizi_suboptimal:2020} and \citet{xian_elevation:2017} who show that a risk-based approach may improve significantly on a standards-based approach.\james{Need to addd some citations of ``risk-based design''}
+
+Additionally, the floodplain maps used for decision supportare silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
+Both risk-based design and standards-based approaches are complicated by nonstationarity.
+On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
+On the other, efforts to find objective methods for quantifying nonstationarity, \eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[see][]{keller_management:2021}.
+
+This suggests that either nonstationarity should be ignored, or subjective approaches for designing for nonstationarity should be adapted.
+Given that physical theories consistent with the observational record and models robustly project changes to climate hazard under warning,
+
+Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
+Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
+We introduce an approach to bridge this divide through a computationally efficient method that weights each state of the world based on a probability distribution over possible futures.
+Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
+This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.
+
+While classical utility- or regret-based frameworks offer a structure for weighing these risks, they are not designed to handle uncertainties that are \emph{deep} in the sense that stakeholders may not agree upon their probabilities \citep{lempert_complex:2002} or \emph{irreducible} in the sense that there is not an objective truth that better models and data can identify \citep{DossGollin:2019}.\footnote{Unless we believe everything is pre-determined, in which case why bother with climate adaptation\ldots}
+Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S.
+federal government will continue to subsidize flood insurance for homeowners.\james{Add some references}
+In response, a wide range of fields have developed scenario-based methods\ldots.\klaus{Which arguments in \citet{savage:1954} did you think should be cited here?}
+This leads to the ``Multiple PDF Problem'' \citep{sharma_rcp:2021}\ldots.
+A key question, then, facing the designers, builders, regulators, owners, and users of built facilities is \emph{to which scenario should infrastructure be designed?}
+
+One example is building codes and guidelines\ldots.
+Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
+Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
+For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
+\citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
+Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
+
+Nonstationarity is particularly relevant in coastal communities where \gls{slr} is expected to drive future flood risk.
+As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch},\james{Need to revisit this picture} given stochastic flood peaks and deeply uncertain \gls{slr}.
+\Gls{slr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots\klaus{You wrote something here that I can't read -- I see something that looks like ICM.
+I assume Tony and Kelsey papers are relevant as well?}) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
+Two key uncertainties identified by previous studies are
+\begin{enumerate*}[label=(\roman*)]
+    \item which emissions pathway (\eg, \gls{rcp} scenario) will we experience and
+    \item how will polar ice sheets respond to warming \citep{wong_brick0.2:2017,ruckert_coastal:2019,wong_nola:2017,deconto_antarctica:2016}.
+\end{enumerate*}
+Simulations of \gls{slr} are available to decision makers grouped by \gls{rcp} scenario and model structure.
+For a given combination of \gls{rcp} scenario and model, results are typically presented and interpreted as probabilistic, but this is not the same as giving decision-makers a single \gls{pdf}.
+This is the ``Multiple \gls{pdf} Problem'' described in, \eg, \citet{sharma_rcp:2021}.
+Thus, \emph{which scenario/\gls{pdf} should we design for?}
+
+Lacking a crystal ball capable of telling us precisely which \gls{pdf} we need to design for, we have to consider several as plausible.
+Any decision will perform differently in different scenarios.
+There is no objective answer to this question.
+Since we can't be ojectively correct, we might as well be transparent and make our assumptions explicit (and subject to critique) rather than implicit.
+
+Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \citep{zarekarizi_suboptimal:2020,xian_elevation:2017}.
+\begin{description}
+    \item[Q1] How to adapt guidance to building characteristics or household preferences?
+    \item[Q2] How does nonstationary hazard change guidance?
+    \item[Q3] Which model of nonstationary hazard to use?
+\end{description}
+These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
+
 \begin{enumerate}

### 2022-01-19 papers/paper/preprint.tex
major facelift for the flowchart

@@
 % fonts and appearance
-\usepackage{amsmath, amsfonts, physics, siunitx, fleqn, nicefrac}
+\usepackage{amsmath, amsfonts, physics, siunitx}
 \usepackage[american]{babel} 
@@
 \usepackage{fontawesome5}
+\usepackage[format=plain, textfont=it]{caption}
 
@@
         Elevation cost & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
-        Initial height & $h_0$            & Deterministic: equal to \gls{bfe} unless otherwise noted                             & TBD                   \\
+        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
         \bottomrule
     \end{tabular}
-    \caption{Uncertainties modeled}\label{tab:uncertainties}
+    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
 \end{table}
@@
         For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
-    }
+    }\label{fig:xlrm}
 \end{figure}
@@
 Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
-Specifically, we use experiments using the BRICK model version 0.2 \citep{wong_brick0.2:2017} simulations published in \citet{ruckart_hazardous:2008}.
-There are eight \glspl{pdf}, each with ZXX\james{count} simulations: four emissions pathways (\gls{rcp} scenarios) $\times$ two models of ice sheet dynamics.
+Specifically, we use simulations published in \citet{ruckart_hazardous:2008} that use the BRICK model version 0.2 \citep{wong_brick0.2:2017}.
 
-For consistency with literature on decision making under deep uncertainty \citep[see][]{marchau:2019}, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.
+\Cref{fig:boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
+Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
+This is also the \gls{pdf} with the greatest within-model variance.
+\begin{figure}
+    \centering
+    \includegraphics[width=5in]{msl_boxplots.pdf}
+    \caption{
+        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
+        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
+    }\label{fig:boxplots}
+\end{figure}
+
+For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.
 
@@
         return periods.
-        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull rule.
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula .
         Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
@@
 
-In \cref{fig:scenario-map}, we plot each possible future as a single dot for each of several possible amounts by which to heighten the house.
+The first that we consider is exploratory modeling.
+In general, we consider $J$ simulated \glspl{sow} $\qty{s_1, \ldots, S_J}$ and $I$ possible decisions $\qty{x_1, \ldots, x_I}$.
+As described in the previous section, these $J$ \glspl{sow} represent time series of mean sea level and the $I$ decisions represent how high to elevate the house.
+For each combination of these, we calculate the vector of metrics described in \cref{sec:sow-metrics}: $u_{ij} = f(x_i, s_j)$ using the system model $f$, which includes storm surge (\cref{sec:storm-surge}), damage (\cref{sec:ead,sec:led}), and calculation of additional metrics (\cref{sec:sow-metrics}).
+This framework is illustrated in \cref{fig:flowchart} inside the box labeled Section 3\klaus{Any suggestions for how I can draw these regions better?}; we return in \cref{sec:multiple-pdf,sec:synthesizing} to add additional steps to the model.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{bayes-rdm.pdf}
+    \caption{
+        Outline of the generally applicable decision theoretic framework developed.
+        We begin in \cref{sec:multiple-simulation} by presenting results using exploratory modeling, \ie nodes  (a), (b), and (c).
+        In \cref{sec:multiple-pdf} we discuss nodes (d) and (e) to understand scenario-conditional probability distributions.
+        Finally in \cref{sec:synthesizing} we add nodes (f) and (g) to synthesize across multiple \glspl{pdf}.
+    }\label{fig:flowchart}
+\end{figure}
+
+Analysis of these $u_{ij}$ sheds light on the question: ``how well does a given decision perform under different possible futures?''
+\Cref{fig:scenario-map} shows each possible future as a single dot for each of several possible values for $\Delta h$.
 On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
@@
 
-This plot sheds light on the question: ``how well does a given decision perform under different possible futures?''
-The best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
+\Cref{fig:scenario-map} shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
 However, the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
-When $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases.
-This plot suggests that $\Delta h = \SI{8}{ft}$ dominates $\Delta h = \SI{14}{ft}$ because \SI{8}{ft} is sufficient to make expected annual damages negligible at a substantially lower construction cost.
-
-Yet important questions remain -- since the \glspl{sow} are not draws from the full \gls{pdf}
+As $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases, up to a point.
+\emph{No decision dominates} any other, including $\Delta h = 0$, in the formal sense of being optimal over all \glspl{sow}.
 
@@
 
-\subsection{One size does not fit all}
+Whereas the analysis in \cref{sec:multiple-simulation} shed light on how decisions perform under different \glspl{sow}, the metrics calculated represent the performance of a given decision-\gls{sow} combination, and \emph{do not shed light on the performance of the decisions themselves}.
 
-\begin{enumerate}
-    \item Tailoring guidance to specific building characteristics can improve outcomes
-    \item Elevating to ``\gls{bfe} plus a foot'' is not always optimal \citep[this is consistent with][]{zarekarizi_suboptimal:2020,xian_elevation:2017}
-    \item Both over- and under-building can be costly \citep{ansar_bigisfragile:2017,DossGollin:2019}
-\end{enumerate}
+An alternative framework is to interpret the \glspl{sow} as draws from each of eight \glspl{pdf}.
+This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as IID draws from the distribution of outcomes.
 
-\begin{figure}
-    \centering
-    \begin{subfigure}{0.48\textwidth}
-        \subcaption{\SI{2.5}{ft} below BFE}
-        \includegraphics[width=\textwidth]{5.5.pdf}
-    \end{subfigure}
-    \hfill
-    \begin{subfigure}{0.48\textwidth}
-        \subcaption{\SI{0.5}{ft} below BFE}
-        \includegraphics[width=\textwidth]{7.5.pdf}
-    \end{subfigure}
-    \caption{
-        Tradeoffs between construction cost and expected NPV lifetime costs for houses initially situated (a) \SI{2.5}{ft} and (b) \SI{0.5}{ft} below the \gls{bfe} (under \gls{rcp} 6.0 with fast BRICK \cite{wong_brick0.2:2017} dynamics).
-        Note unequal $y$-axis scales.
-    }
-\end{figure}
+Some general notation is shown in \cref{fig:flowchart} (additional terms are added in \cref{sec:synthesizing}).
+We have $K=8$ models of sea level rise, which we refer to as $\qty{\mathcal{M}_1, \ldots, \mathcal{M}_K}$.
 
-\subsection{It matters which PDF you design for}
+\subsection{Metrics}
 
-Uncertainties from multiple model structures (\eg, ice sheet dynamics) and/or scenarios (\eg, emissions pathways) create multiple estimates of time-varing hazard, \ie, the \textbf{\itshape{Multiple PDF Problem}}.
-Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
-\begin{figure}
-    \centering
-    \includegraphics[width=5in]{msl_boxplots.pdf}
-    \caption{
-        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
-        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
-    }\label{fig:boxplots}
-\end{figure}
+This allows us to reason formally about the \gls{pdf} of outcomes for each candidate decision, conditional on a model.
+We can use these \glspl{pdf} of outcomes to calculate metrics describing the performance of decisions.
 
@@
 
+Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
+
 Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.\james{TODO: create a figure showing expected outcomes if you design for scenario $x$ and get scenario $y$}
@@
 While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{bayes-rdm.pdf}
-    \caption{
-        ``Prior Model Averaging'' framework.
-    }\label{fig:flowchart}
-\end{figure}
+
 Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
@@
 
-In this section we present a general framework for decision making under deep uncertainty.
-We begin with terms and definitions that are broadly consistent with common adaptive frameworks\james{add refs}\klaus{did I read this right?}.
-In \cref{sec:methods-weight} we present a Bayesian model for weighting \glspl{sow}.
-In \cref{sec:methods-iterative} we discuss how this framework can be integrated into frameworks for iterative engagement with experts and stakeholders to expand and frame the model.
-
-We start by modeling the performance of a decision $x \in \mathcal{X}$.
-Outcomes depend both on the decision and on a state of the world $s \in \Omega$ that describes all relevant uncertainties.
-To model outcomes $u \in \mathcal{U}$, we use a system model $f: \mathcal{X}, \Omega \rightarrow \mathcal{U}$ \citep[we generally follow the notation of][]{mcphail_robustness:2019} which we assume to be deterministic.
-If $f$ is a stochastic function, we can define its metrics in terms of expectations, \eg $u_{ij} = \mathbb{E}\qty[u | x=x_i, s=s_j]$, which can be approximated through techniques like Monte Carlo sampling.
-
-The literature on decision making under deep uncertainty \citep[see][for a review]{marchau:2019} emphasizes the importance of using system models like $f$ in an exploratory context by sampling widely from plausible futures $s \in \Omega$ and studying their implications \citep{bankes:1993,lempert_shaping:2003,kwakkel:2019,lamontagne_discovery:2018}.
-This is often described as ``bottom-up'' modeling \citep{Brown:2012kb} and has parallels to model-based stress tests and vulnerability assessments.
-
-Given our system model $f$ and a sampling of $J$ possible states of the world $s_1, s_2, \ldots, s_J$, we can calculate a set of \emph{\gls{sow} metrics} $u_{ij}$ that tell us ``how will decision $x_i$ perform under state of the world $s_j$'' for many $s_j$, and one or more $x_i$.
-For an apples to apples comparison of how different decisions perform on the $j$th \gls{sow}, we can directly compare $u_{1j}, u_{2j}, \ldots, u_{Ij}$.
-However, if we wish to compare the performance of two or more decisions across \emph{all} scenarios, we will need some way of aggregating the \gls{sow} metrics into \emph{decision metrics} $m_1, m_2, \ldots, m_I$ such that for every $x_i$ there is a single (scalar or vector) $m_i$.\footnote{I think the link back to our original motivation -- the multiple PDF problem}
-
-\citet{mcphail_robustness:2019} split the problem of aggregating \gls{sow} metrics into decision metrics into two steps: ``subsetting'' (\eg, single value, all values, subset of values) and ``metric calculation'' (\eg, mean, variance, skew, kurtosis).\klaus{You say something about see discussion in \citet{mcinerney_robust:2012} -- please elaborate}
-We generalize this to
-\begin{equation}
-    m_i = g(\qty{u_{i1}, u_{i2}, \ldots, u_{iJ}}, \qty{w_1, w_2, \ldots, w_J}),
-\end{equation}
-where $g$ is an \emph{aggregating function} describing the ``metric calculation'' step and the $w_j$ describe weights given to each \gls{sow}.
-A optional, but likely useful, interpretation of the weighted samples $\qty{w_j, u_{ij} \forall j=1, \ldots, J}$ is an empirical \gls{pdf}, in which case $g$ maps a \gls{pdf} of future outcomes to the metrics $m_i$.\footnote{perhaps this belongs in the Discussion}
-This is a flexible framework; \cref{tab:literature-weighting} shows some methods for robust decision making from the literature using our notation.\james{Add some references to the table}
-
-\begin{table}
-    \centering
-    \footnotesize
-    \begin{tabular}{llll}
-        \toprule
-        Citation        & Explanation                     & $w_j$                                                                             & $g(u, w)$          \\
-        \midrule
-        Many            & All $s_j$ are equally plausible & $w_j = \nicefrac{1}{J}$                                                           & any                \\
-        Many            & Worst case                      & $w_j = \begin{cases}1 & u_{ij} = \min_{j'} u_{ij'} \\ 0 & \text{else}\end{cases}$ & $u$                \\
-        TBD             & Average over \glspl{sow}        & $w_j = \nicefrac{1}{J}$                                                           & $\sum_j w_j u_{j}$ \\
-        Characklis/Reed & 95th percentile                 & $w_j = \nicefrac{1}{J}$                                                           & weighted quantile  \\
-        \bottomrule
-    \end{tabular}
-    \caption{
-        A non-exhaustive summary of methods for setting $w_j$ and $g$ from the literature.
-        (\emph{What other examples should we add?})
-    }\label{tab:literature-weighting}
-\end{table}
-
-In the next section we propose an alternative scheme for weighting scenarios.
-
-\paragraph{To think  about}
-I have two flavors of this idea presented here!
-\Cref{fig:framework}\james{Updates for \cref{fig:framework}: Don't need to say here that $f$ is deterministic.
-Larger text and darker color for questions.
-Rename $p(s)$ as ``prior over parameters and model structures''} takes a more Bayesian focus: the goal is to reason about properties of the \emph{distribution} over outcomes $p(u | x=x_i)$.
-The text follows \citet{mcphail_robustness:2019} more.
-I'm not sure which is easier to understand.
-
-\subsection{Weighting \glsentryplural{sow}}\label{sec:methods-weight}
-
-If the common practice of weighting all scenarios equally is applied ($w_j \propto 1$ in \cref{tab:literature-weighting}), a tension emerges: we want to \emph{explore} highly unlikely but theoretically possible trajectories to understand their consequences, but we don't want extremely unlikely scenarios to dominate our decision.
-
-Some aims:
-\begin{enumerate}
-    \item Assumptions should be as transparent and visible as possible to facilitate critique and dialogue\klaus{I moved this up}
-    \item We should be invariant to the parameterization of $s$ -- if we sample population uniformly from 1 million to 10 million or $\log_{10}$ population uniformly from \num6 to 7, we should get the same answer
-    \item We should be invariant to the sampling of $s$:
-          \begin{enumerate}
-              \item if we include scientifically interesting but extremely low-probability scenarios (\eg, if someone created \gls{rcp} 45), our esimtates of $m_i$ should not change meaningfully.\james{Explain why}
-              \item if we include nearly redundant scenarios (\eg, if we are studying \gls{rcp} 2.6, 4.5, 6.0, and 8.5 and we then add results from a \gls{rcp} 6.01 experiment), our estimates of $m_i$ should not change meaningfully.\james{Explain why}
-          \end{enumerate}
-\end{enumerate}
-We can do this with probability if we have a target distribution $p(s)$.
-We return to the question of how to choose this in \cref{sec:methods-prior}.
-
-\subsection{Melding prior over \glsentryplural{sow}}\label{sec:methods-prior}
-
-Often (including in the case study described in \cref{sec:case}) our \glspl{sow} are the outputs of complicated model chains.
-To set our prior $p(s)$ we might like to incorporate prior external knowledge about both the
-
-\subsection{Opportunities for iterative model expansion}\label{sec:methods-iterative}
-
-\paragraph{Robustness checks}
-The computationally expensive piece is computing the $u_{ij}$.
-Once you do that, you can look at several $p(\Omega)$ and compute the $m_i$ multiple times.
-
-\paragraph{Links to DMDU Methods}
-You can use scenario discovery on the $(x_i, \mu_{ij}$).
-(Others to note?)
-
-\paragraph{Bayesian model checks}
-Let's just cite \citet{gelman_workflow:2020} and \citet{gabry_visualization:2019} and point out that you can and should conduct sample-based checks on your prior $p(\Omega)$ and your posterior $p(u | x=x_i)$.\footnote{See discussion above about whether we want to emphasize that we are treating this as a posterior or not}
-
-\section{Case study: house elevation}\label{sec:case}
-
-We start with a bare-bones version of the \citet{zarekarizi_suboptimal:2020} model.
-For interpretability, we fix the discount rate, house lifetime, and depth-damage curve to illustrate how our framework can be used to model nonstationary flood risk.
-Real-world decision makers should consider additional uncertainties and stakeholder-identified metrics.
-
-\begin{figure}
-    \centering
-    %\includegraphics[height=2.5in, width=0.6\textwidth, keepaspectratio=true]{xlrm}
-    \caption{
-        Diagram showing uncertainties ($X$; red), levers ($L$; yellow), system relationships ($R$; gray), and metrics ($M$; blue).
-    }\label{fig:xlrm}
-\end{figure}
-
-\subsection{Scenarios}
-
-Coastal floods are often decomposed into changes in local sea level \citep{kopp_evolving:2017,kopp_probabilistic:2014,wong_brick0.2:2017} plus storm surge \citep{garner_slrise:2018,cagigal_emulator:2020,rueda_surge:2016}.
-We neglect waves and hydrodynamics for simplicity.
-
-\gls{slr}:
-\begin{enumerate}
-    \item We have 2 models of ice sheet dynamics (3? Vivek and Tony are trying to tell me what models the data from \citet{ruckert_coastal:2019} actually comes from)
-    \item We have 4 \gls{rcp} scenario
-    \item We use the BRICK model \citep{wong_brick0.2:2017} simulations from \citet{ruckert_coastal:2019} to define $J$ \glspl{sow}.
-\end{enumerate}
-Storm surge:
-\begin{enumerate}
-    \item We have a $N$ year observational record
-    \item We separate out storm surge from the background trend in some standard way
-    \item We model storm surge as a stationary GEV model
-    \item We place a prior on the return level (data space) rather than a prior on the parameters themselves (except the shape parameter -- we put a restrictive prior for it to be positive)\james{arguably we should stick to boring priors here -- the question being whether it's better to say we use iterative checks to improve the priors, or whether we want to keep things simple}
-    \item We draw $K$ sets of parameters $\qty(\mu_k, \sigma_k, \xi_k)$
-\end{enumerate}
-Thus a \gls{sow} is defined just an 81 year time series of \gls{lsl}.
-We will see what we do with the $K$ draws of storm surge parameters shortly.
-\begin{figure}
-    \caption{
-        INSERT HERE: figure showing time series of observed surge and \gls{lsl} with notable storms indicated, as well as projections of future \gls{lsl} under all eight models (two BRICK, four \gls{rcp})
-    }
-\end{figure}
-
-\subsection{Decisions}
-
-We consider only one lever: how high to elevate the house today.
-This is the same lever considered in \citet{zarekarizi_suboptimal:2020}
-
-\subsection{Metrics}
-
-Three scenario metrics $u$:
-\begin{enumerate}
-    \item Construction cost
-    \item Discounted expected flood damages
-    \item Total cost (construction plus flood)
-\end{enumerate}
-The decision metrics $m$ are just the expected value of $u$.
-
-\subsection{A simplified didactic system model}
-
-\begin{enumerate}
-    \item Costs of elevating the house are the same as \citet{zarekarizi_suboptimal:2020}
-    \item Depth-damage: use HAZUS model from \citet{zarekarizi_suboptimal:2020}
-    \item Discount: fix at some appropriate rate (2\%?)
-    \item House lifetime: fix at some reasonable amount (80 years)
-\end{enumerate}
-
-Recall that we have $K$ simulations of the storm surge parameters.
-For each year of each \gls{sow} we would like to estimate the expected value of the convolution of hazard (storm surge plus \gls{lsl}) with fragility (depth-damage).
-A simple algorithm to estimate this is, for each \gls{pdf} of storm surge $k \in 1, \ldots, K$, to draw $N$ Monte Carlo sample values of storm surge, calculate the damage associated with each, and take the expectation with respect to $N$.
-Then, take the expectation again with respect to the $K$ storm surge \glspl{pdf}.
-Since this is computationally expensive, we build an emulator that estimates expected damage as a function of mean sea level only by fitting a spline to the result of systematically calculating expected damages for different values of \gls{lsl}.\footnote{A supplemental figure would be good here.
-This is probably a bit dense of a discussion.}
-
-\begin{figure}
-    \caption{
-        INSERT HERE: figure showing the convolution of storm surge hazard through the depth-damage function.
-        We fix \gls{lsl} constant but have many \glspl{pdf} of \gls{lsl}.
-        This is the ``hazard, damage, risk'' graph that Klaus has sketched.
-    }
-\end{figure}
-
-\subsection{Belief model}
-
-\paragraph{Prior on outputs only}
-Following \citet{garner_slrise:2018}, \citet{oddo_coastal:2017}, and \citet{lempert_slr:2012} we place a prior on \gls{slr} in 2100.\klaus{You suggest looking at another paper.
-Are you referring to this?\\\fullcite{sriver_sealevel:2018}}
-
-\paragraph{Prior on inputs only}
-Alternatively, we can place a prior over scenarios.
-Let's say that we think that emissions forcing in 2100 is has some distribution.
-This means we are placing a weight over \gls{rcp} scenarios.
-We can assume both BRICK models are equally likely.
-
-\paragraph{Melding}
-We can do cool stuff with Bayesian melding \citep{poole_melding:2000,sevcikova_melding:2007}.
-This combines the prior over inputs and outputs geometrically.
-\begin{equation}\label{eq:melding}
-    \text{melding equation here}
-\end{equation}
-This lets us express information about some inputs being more or less likely than others, and also some outputs being more or less likely than others.
-THIS IS USEFUL because often we would like to incorporate relevant knowledge from many different sources on both inputs and outputs (\eg, IPCC reports, National Climate Assessments, \etc).
-
-\paragraph{Implementation}
-Code will be available on GitHub.
-We use the Julia programming language.
-
-\section{Results}\label{sec:results}
-
-We will need a figure or two showing our prior and posterior checks, as well as some results.
-Results should show the tradeoff between cost and damage under different assumptions.
-
-\section{Discussion}
-
 Most fundamentally, we are thinking about robustness a bit differently than usual.
@@
 We are thinking about robustness to different modeling structures and assumptions.
-This has links to the ``rival framings'' literature that Pat has been working on.
+This has links to the ``rival framings'' literature.
 Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).

### 2022-01-19 tikz/bayes-rdm.tex
major facelift for the flowchart

@@
 
+\pgfdeclarelayer{bg}    % declare background
+\pgfdeclarelayer{bg2}    % declare background
+\pgfdeclarelayer{bg3}    % declare background
+\pgfsetlayers{bg3, bg2, bg, main}  % order of layers (main = standard layer)
+
 \begin{tikzpicture}[auto]
 
-    \node [block] (decisions) at (0, 3) {
+    \node [block, label={[]135:{\footnotesize (d)}}] at (-4, 3) (models) {
+            Multiple models\\
+            $\qty{M_1, \ldots, M_K}$
+        };
+    \node [block, label={[]135:{\footnotesize (a)}}] (decisions) at (0, 3) {
             Candidate decisions\\
-        $\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$
+            $\qty{x_1, x_2, \ldots, x_I}$
         };
-    \node [block] at (-4, 3) (models) {
-        Multiple PDFs\\
-        $\qty{M_1, \ldots, M_K} \in \mathcal{M}$
-    };
-    \node [block] at (-4, 0) (simulations) {
-        Simulated SOWs\\
-        $\qty{s_1, \ldots, s_{N_k}}_k$
+    \node [block, label={[]135:{\footnotesize (b)}}] at (-4, 0) (simulations) {
+            Simulated futures\\
+            $\qty{s_1, \ldots, s_J} \in \Omega$
         };
     \path [line] (models) -- (simulations);
-    \node [function] (model) at (0, 0) {
+    \node [function, label={[]45:{\footnotesize (c)}}] (model) at (0, 0) {
             System model\\
@@
 
-    \node [derived] (combined) at (4, -3) {
-        Conditional distributions\\
-        over outcomes\\
-        $\hat{p} \qty(u | x_i, M_k)$
+    \node [function, label={[]135:{\footnotesize (f)}}] (prior) at (-4, -3.5) {
+            Prior over\\
+            possible futures\\
+            $p\qty(\Omega)$
         };
-    \path [line] (model) -- (combined);
-
-    \node [function] (prior) at (-4, -3) {
-        Prior\\
-        over SOWs\\
-        $p\qty(s)$
-    };
-    \node [derived] (weights) at (0, -3) {
-        Model weights\\
-        $w_1, \ldots, w_K$
+    \node [derived, label={[]90:{\footnotesize (g)}}] (weights) at (0, -3.5) {
+            Weights\\
+            $w_1, \ldots, w_J$
         };
@@
 
-    \node [derived] (aggregated) at (4, -6) {
-        Aggregated distribution\\
+    \node [derived, label={[]135:{\footnotesize (e)}}] (combined) at (4, 0) {
+            Conditional\\
+            distribution\\
             over outcomes\\
-        $\hat{p}(u | x_i)$
+            $\hat{p}(u | x_i, M_k)$
         };
-    \path [line] (combined) -- (aggregated);
-    \path [line] (weights) -- (aggregated);
+    \path [line] (model) -- (combined);
+    \path [line] (weights) -- (combined);
+
+    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\large \textbf{Section 3}};
+    \begin{pgfonlayer}{bg}    % select background
+        \path [dotted, thick, color=thirds, fill=grigio!20] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- (-6, -1.5);
+    \end{pgfonlayer}
+
+    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\large \textbf{Section 4}};
+    \begin{pgfonlayer}{bg2}    % select background
+        \path [dotted, thick, color=thirds, fill=rosso!20] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- (-6.5, 4.5);
+    \end{pgfonlayer}
+
+    \node [color=thirds] (s5) at (4, -3.5) {\large \textbf{Section 5}};
+    \begin{pgfonlayer}{bg3}    % select background
+        \path [dotted, thick, color=thirds, fill=giallo!20] (-6.75, 4.8) -- (-6.75, -5) -- (5.75, -5) -- (5.75, 4.8) -- (-6.75, 4.8);
+    \end{pgfonlayer}
+
 \end{tikzpicture}

### 2022-01-20 papers/paper/preprint.tex
Send to Klaus for comments / feedback

@@
 
-\Cref{fig:boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
+\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
 Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
@@
         We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
-    }\label{fig:boxplots}
+    }\label{fig:msl-boxplots}
 \end{figure}
@@
 
-To assess the performance of a given decision for a specific \gls{sow}, we calculate three metrics for each \gls{sow}.
-\begin{enumerate}
-    \item The first metric considered is ``upfront cost,'' which reflects the cost of elevating a house.
+To assess the performance of a given decision for a specific \gls{sow}, we calculate the following metrics \emph{for each decision-\gls{sow} combination}.
+\begin{description}
+    \item[upfront cost] reflects the cost of elevating a house.
 Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
 We normalize this cost by house value.\james{Normalize!}
-    \item The second metric is the lifetime expected damages (\cref{eq:led})
-    \item The third metric is lifetime expected costs, which is the sum of lifetime expected damages and upfront costs
-    \item The third metric is the expected benefit-to-cost ratio.  We define this ratio in terms of the ratio of construction cost to damages averted relative to not elevating at all:
+    \item[lifetime expected damages] is discussed in \cref{sec:led}
+    \item[lifetime expected costs] is the sum of lifetime expected damages and upfront costs
+    \item[the expected benefit-to-cost ratio] is defined as the ratio of construction cost to damages averted, relative to not elevating at all:
         \begin{equation}
@@
         \end{equation}
-\end{enumerate}
+\end{description}
+In \cref{sec:multiple-simulation} we introduce the notation $u$ for these metrics.
 
-\section{Exploring possible states of the world}\label{sec:multiple-simulation}
+\subsection{Decision analytical framework}
 
-The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \acrfullpl{sow}.
-
-The first that we consider is exploratory modeling.
-In general, we consider $J$ simulated \glspl{sow} $\qty{s_1, \ldots, S_J}$ and $I$ possible decisions $\qty{x_1, \ldots, x_I}$.
-As described in the previous section, these $J$ \glspl{sow} represent time series of mean sea level and the $I$ decisions represent how high to elevate the house.
-For each combination of these, we calculate the vector of metrics described in \cref{sec:sow-metrics}: $u_{ij} = f(x_i, s_j)$ using the system model $f$, which includes storm surge (\cref{sec:storm-surge}), damage (\cref{sec:ead,sec:led}), and calculation of additional metrics (\cref{sec:sow-metrics}).
-This framework is illustrated in \cref{fig:flowchart} inside the box labeled Section 3\klaus{Any suggestions for how I can draw these regions better?}; we return in \cref{sec:multiple-pdf,sec:synthesizing} to add additional steps to the model.
+In \cref{fig:flowchart} we introduce a general notation for model-based decision analysis which we develop incrementally.
+We consider $J$ \glspl{sow} representing time series of mean sea level and $I$ discrete decisions representing the choice for how high to elevate the house $\Delta h$.
+For each combination of these, we calculate the vector of metrics described in \cref{sec:sow-metrics}, which we denote $u_{ij} = f(x_i, s_j)$, using a system model $f$ (box c).
+This system model represents storm surge, building costs, and damages to output the metrics $u$ (which in this case is a vector).
+In \cref{sec:multiple-simulation} we consider ``the multiple \gls{pdf} problem'' by introducing boxes (d) and (e).
+Finally in \cref{sec:synthesizing} we introduce box (f) as a tool for synthesizing uncertainty across the multiple \glspl{pdf}.\klaus{Any feedback on the is figure and/or section?}
 
@@
         Outline of the generally applicable decision theoretic framework developed.
-        We begin in \cref{sec:multiple-simulation} by presenting results using exploratory modeling, \ie nodes  (a), (b), and (c).
-        In \cref{sec:multiple-pdf} we discuss nodes (d) and (e) to understand scenario-conditional probability distributions.
-        Finally in \cref{sec:synthesizing} we add nodes (f) and (g) to synthesize across multiple \glspl{pdf}.
+        Boxes are labeled in the order in which they are presented in the text: we begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
+        In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
+        Finally in \cref{sec:synthesizing} we add boxes (f) and (g) to synthesize across multiple \glspl{pdf}.
     }\label{fig:flowchart}
@@
 
-Analysis of these $u_{ij}$ sheds light on the question: ``how well does a given decision perform under different possible futures?''
-\Cref{fig:scenario-map} shows each possible future as a single dot for each of several possible values for $\Delta h$.
+\section{Exploratory analysis}\label{sec:multiple-simulation}
+
+The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \acrfullpl{sow}.
+The first that we consider is exploratory modeling, which sheds light the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''
+
+\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\klaus{Would it be better here to have a figure that has $\Delta h$ on the $y$ axis and uses expected lifetime cost to color the points?}
 On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
 On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
-Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ.
+Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ (\ie, the 1-dimensional projection of the \glspl{sow} does not represent all variance between them).
 
@@
         Scenario maps: one dot represents one \gls{sow}.
+        Some work is still needed -- I think that changing to four values of $\Delta h$ and showing two or three different values of $h_0$ might be useful (alternatively we could vary the discount rate $\gamma$).
+        Also need to rescale the $y$-axis by house value.
+        It would also be cool to create an interactive online version of this figure that lets you scroll over a particular dot and learn which scenario it's from!
+        %TODO
     }\label{fig:scenario-map}
@@
 
-\Cref{fig:scenario-map} shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible.
-However, the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
+The exploratory modeling approach shown in \cref{fig:scenario-map} shows that shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible, but the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
 As $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases, up to a point.
 \emph{No decision dominates} any other, including $\Delta h = 0$, in the formal sense of being optimal over all \glspl{sow}.
+\james{Revisit this paragraph once changes are made to \cref{fig:scenario-map}.}
 
-\section{The multiple PDF problem}\label{sec:multiple-pdf}
+\section{Scenario-conditional probabilistic analysis: the multiple PDF problem}\label{sec:multiple-pdf}
 
@@
 
-An alternative framework is to interpret the \glspl{sow} as draws from each of eight \glspl{pdf}.
+An alternative interpretation of the available \glspl{sow} is as IID draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
 This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as IID draws from the distribution of outcomes.
 
-Some general notation is shown in \cref{fig:flowchart} (additional terms are added in \cref{sec:synthesizing}).
-We have $K=8$ models of sea level rise, which we refer to as $\qty{\mathcal{M}_1, \ldots, \mathcal{M}_K}$.
-
-\subsection{Metrics}
-
-This allows us to reason formally about the \gls{pdf} of outcomes for each candidate decision, conditional on a model.
-We can use these \glspl{pdf} of outcomes to calculate metrics describing the performance of decisions.
-
-Any optimal strategy or Pareto frontier is conditional upon an (explicit or implicit) model of future outcomes.
+A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
+Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be IID samples from the distribution of outcomes (box e).
+Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
+We can then use this conditional distribution over outcomes to create metrics that depend only on the decision, such as averages or quantiles of outcomes.
 
@@
 
-Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
+For example, \cref{fig:tradeoffs}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
+Similarly, \cref{fig:tradeoffs} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
+This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp} 8.6 with fast ice sheet dynamics).
+Consistent with \cref{fig:scenario-map}, the variability of lifetime costs increases as $\Delta h$ increases.
+Once $\Delta h$ reaches \SIrange[]{3}{7}{ft}, depending on the model considered, construction costs start to dominate flood losses, and thus higher values of $\Delta h$ increase average lifetime costs.
 
-Weights assigned to PDFs or simulations, \textbf{including implicit uniform weights}, should be communicated transparently to facilitate critique and improvement.\james{TODO: create a figure showing expected outcomes if you design for scenario $x$ and get scenario $y$}
+Calculating either of the quantities shown in \cref{fig:tradeoffs} \emph{requires synthesizing across \glspl{sow}}.
+However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
+This means that any assessment of decision performance (\eg, an optimal strategy, a Pareto frontier, or a cost-benefit analysis) is conditional upon an (explicit or implicit) model of future outcomes.
+Where these models are implicit, they should be made more explicit to facilitate iterative model critique and improvement.
 
-\section{Synthesizing across PDFs for decision relevance}\label{sec:synthesizing}
+Second, this approach presents what we call ``the multiple \gls{pdf} problem'' because it leaves decision makers with many probability distributions to choose from
+Cite \citet{sharma_rcp:2021} as another example.
+Add a sentence or two explaining why this is a problem for decision makers\ldots\james{Need to bring in a bit of discussion from \citet{schneider_scenarios:2002} and from engineering docs (ASCE, FEMA, \ldots).}\klaus{Would appreciate your help putting together a succinct summary of the multiple \gls{pdf} problem}
 
-\subsection{Methods}
+\section{Synthesizing PDFs for decision relevance}\label{sec:synthesizing}
 
-While comparing results across many PDFs is a valuable exploratory exercise (see, \eg, \cref{fig:tradeoffs}), providing end users (\eg, households practicing engineers, and planning departments of local governments) with many PDFs may lead to confusion, inconsistency, and potentially liability \citep[see][]{schneider_dangerous:2001,schneider_scenarios:2002}.
+Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.
 
-Our approach to synthesizing PDFs is designed for the common case of assessing decisions using simulations (SOWs) from each of $K$ PDFs.
-A system model $(f)$ quantifies the performance $(u)$ of a decision ($x$) under a single SOW ($s$).
-To synthesize, we weight PDFs using an approach based on Bayesian Model Averaging \citep{wong_surge:2018,Yao:2018bu}.
+Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{empirical}(\Omega)$, does not reflect our belief about their true distribution, $p_\mathrm{true}(\Omega)$.
+Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.
+This is analagous to the deliberate over- or under-sampling of certain groups in survey design, and so following standard practice in survey design we develop a re-weighting approach to correct for the over- or under-sampling of different regions of the survey space.
 
-\subsection{Findings}
+To calculate these weights, we apply a grid-based discretization scheme.
+First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
+As before, we take this space to be the mean sea level in 2100.
+Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
+This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
+Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
+$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
+we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:weighting}.
+\begin{equation}\label{eq:weights}
+    w_j = \begin{cases}
+        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
+        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
+        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
+    \end{cases}
+\end{equation}
+This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.
+
+The primary difficulty with \cref{eq:weights} is that the true probability distribution $p_\mathrm{true}(\phi)$ is not only unknown but deeply uncertain.
+Because it is unknown, it must be approximated with a model.
+And because it is deeply uncertain, multiple models for $p_(\Omega)$ should be used.
+But whereas the $M_k$ defined in \cref{sec:multiple-pdf} represented the raw input data (four \gls{rcp} scenarios times two models of ice sheet dynamics), we define the $M_k$ here to be subjective models for mean sea level in 2100.
+
+Need a few sentences here on prior uses of expert judgement and subjective belief.\klaus{Any suggested citations?}
+The key point to get across is what a subjective belief means in the Bayesian context \citep{savage:1954,gelman_philosophy:2013,gelman_subjectiveobjective:2017}.
+David Spiegelhalter had a nice comment on a podcast where he explained that in medicine, there isn't a ``true'' probability that, for example, you will get cancer; this is not some real quantity that can be measured \citep[hence the famous ``probability isn't real'' admonisment of][]{definetti_probability:1972}.
+Similarly, the probability distribution of sea level rise in 2100 isn't something that can be measured.
+However, probability provides a clear and self-consistent mathematical language for reasoning about uncertainty, which is why we're using it.
+A key advantage is that \textbf{since we can't be right, we can at least be transparent} about the assumed values of sea level in 2100, rather than hiding behind implicit assumptions.
+Thus we refer to the $M_k$ here as \emph{priors} to reflect that they consistently represent our current knowledge.
 
@@
     \centering
-    \includegraphics[width=6in]{inference_weights.pdf}
+    \includegraphics[width=4in]{../img/weighting.pdf}
     \caption{
-        The weights assigned to each PDF (from \cref{fig:boxplots}) under each prior considered.
-    }
+        Schematic of re-weighting scheme defined in \cref{eq:weights} for $J=4$.
+        Top panel shows $p_\mathrm{true}$ and bottom panel shows $F_\mathrm{true}$.
+        Vertical black lines show $\frac{1}{2}\qty[\psi_j + \psi_{j+1}]$ for $j=1, \ldots, J-1$.
+        Green text shows $w_1, \ldots, w_4$ and how they are calculated from $F_\mathrm{true}$.
+        Need to update this figure using real data.
+        %TODO
+    }\label{fig:weighting}
 \end{figure}
-``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.
+
+Need to explain how we came up with our two or three subjective priors.\klaus{Would love suggestions here!}
+Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
+Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
+\Cref{fig:tradeoffs} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
 
@@
         Alternative priors can provide robustness checks.
-    }
+        Need to add panel labels, update data, \etc.
+        %TODO
+    }\label{fig:tradeoffs}
+\end{figure}
+
+This technique can also be used as a diagnostic tool.
+In \cref{fig:model-average-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=6in]{inference_weights.pdf}
+    \caption{
+        The average weight assigned to \glspl{sow} from each model (\gls{rcp} scenario and ice sheet parameterization) for both priors considered.
+        If the sampling distribution (green line) is used, then the weight assigned to each \gls{sow} is the same, regardless of whether it is over- or under-sampled.
+    }\label{fig:model-average-weights}
 \end{figure}
@@
     \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
-    \item Instead, we can ask \textbf{what assumptions are different priors consistent with?}
 \end{itemize}
@@
 This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
-\textbf{\scshape let's collaborate on:}
+
+Future research needs:
 \begin{itemize}
@@
 
-\clearpage
+\section{Bonus (need to sort through)}
+
+This section is a work in progress -- it includes some text that used to be in the paper as well as some bullet points that may be useful for the discussion section.\klaus{Can you highlight any picese here that you find especially salient?}
+
+Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
 
@@
 Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).
+
+Since we have priors, there are clear links to how we would update our belief (and thus decision) once more information arrives.
 
@@
 
-\section{Conclusions}
-
-\begin{enumerate}
-    \item We started with the problem of ``which scenario should we design to?'' We then devised a cop-out: design to a weighted combination of all the scenarios.
-(So how should we wight the scenarios?
-Come up with a model $p(\Omega)$ that tells you.)
-    \item There is no objective way to choose $p(\Omega)$ because we don't have data on the future!
-    \item We can't get around from choosing $p(\Omega)$ implicitly\footnote{I'm not sure we have demonstrated this rigorously, or whether we want to argue it, but I think there's a way to frame this point that is valid and helpful}
-    \item Ultimately, engineering design needs to consider not only how to weight different future scenarios, but also which metrics are important, how risk aversion should be addressed, \etc.
-    \item We used a didactic case study to present a toolkit (decision framework, melding priors, prior over return levels) for making transparent and critique-friendly assessments.
-\end{enumerate}
-
-\section*{End Matter}
-
-\subsection*{Acknowledgements}
-
-\begin{enumerate}
-    \item MARISA\james{details}
-    \item PSIRC\james{details}
-    \item Rice University (for startup)
-    \item Tor Erlend Fjelde for helpful advice on implementing the GEV prior using \texttt{Turing.jl}.
-    \item For many journals, a code and data statement will go here.
-\end{enumerate}
-
-\subsection*{Author Contributions}
-
-See journal requirements and format
-
-\subsection*{Supplementary Materials}
-
-\section{Old material}
-
-\clearpage
-\newpage
-
 Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
@@
 
+\section*{End Matter}
+
 \begin{enumerate}
@@
 
+\subsection*{Acknowledgements}
+
+\begin{enumerate}
+    \item MARISA\james{details}
+    \item PSIRC\james{details}
+    \item Rice University (for startup)
+    \item Tor Erlend Fjelde for helpful advice on implementing the GEV prior using \texttt{Turing.jl}.
+    \item For many journals, a code and data statement will go here.
+\end{enumerate}
+
+\subsection*{Author Contributions}
+
+See journal requirements and format
+
 \printbibliography

### 2022-01-20 tikz/bayes-rdm.tex
Send to Klaus for comments / feedback

--- 2021-12-10 tikz/bayes-rdm.tex
@@
 \tikzstyle{function} = [circle, draw, fill=grigio, text=rosso, minimum width=5.5em, text centered, align=center]
-\tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=center]
-\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=rosso, text=grigio, minimum width=6em, text centered,  minimum height=4em, align=center]
-\tikzstyle{check} = [text=thirds, align=center]
+\tikzstyle{block} = [rectangle, draw, fill=giallo, text=nero, minimum width=6em, text centered,  minimum height=4em, align=left]
+\tikzstyle{derived} = [rectangle, draw, rounded corners, fill=rosso, text=grigio, minimum width=6em, text centered,  minimum height=4em, align=left]
 
@@
 
-    \node [block] (decisions) at (0, 3) {
+    % label the various sections
+    \path [fill=grigio!20] (-6.75, 4.8) -- (5.75, 4.8) -- (5.75, -4) -- (-6.75, -4) -- cycle;
+    \node [color=thirds] (s5) at (-4, -3) {\textbf{Synthesize (\S 5)}};
+
+    \path [fill=giallo!20] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- cycle;
+    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{Condition (\S 4)}};
+
+    \path [fill=thirds!20] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- cycle;
+    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\textbf{Explore (\S 3)}};
+
+    \node [block, label={[]135:{\footnotesize (d)}}] at (-4, 3) (models) {
+            Multiple models\\
+            for~$p(\Omega | M_k)$:\\
+            $\qty{M_1, \ldots, M_K}$
+        };
+    \node [block, label={[]135:{\footnotesize (a)}}] (decisions) at (0, 3) {
             Candidate decisions\\
-        $\qty{x_1, x_2, \ldots, x_I} \in \mathcal{X}$
+            $\qty{x_1, x_2, \ldots, x_I}$
         };
-    \node [block] at (-4, 3) (models) {
-        Multiple PDFs\\
-        $\qty{M_1, \ldots, M_K} \in \mathcal{M}$
-    };
-    \node [block] at (-4, 0) (simulations) {
-        Simulated SOWs\\
-        $\qty{s_1, \ldots, s_{N_k}}_k$
+    \node [block, label={[]135:{\footnotesize (b)}}] at (-4, 0) (simulations) {
+            Many\\
+            Simulated futures\\
+            $\qty{s_1, \ldots, s_J} \in \Omega$
         };
     \path [line] (models) -- (simulations);
-    \node [function] (model) at (0, 0) {
+    \node [function, label={[]45:{\footnotesize (c)}}] (model) at (0, 0) {
             System model\\
@@
 
-    \node [derived] (combined) at (4, -3) {
-        Conditional distributions\\
+    \node [derived, label={[]90:{\footnotesize (f)}}] (weights) at (0, -3) {
+            Weights\\
+            $w_1, \ldots, w_J$
+        };
+    \path [line] (simulations) -- (weights);
+
+    \node [derived, label={[]135:{\footnotesize (e)}}] (combined) at (4, 0) {
+            Conditional\\
+            distribution\\
             over outcomes\\
-        $\hat{p} \qty(u | x_i, M_k)$
+            $\hat{p}(u | x_i, M_k)$
         };
     \path [line] (model) -- (combined);
+    \path [line] (weights) -- (combined);
 
-    \node [function] (prior) at (-4, -3) {
-        Prior\\
-        over SOWs\\
-        $p\qty(s)$
-    };
-    \node [derived] (weights) at (0, -3) {
-        Model weights\\
-        $w_1, \ldots, w_K$
-    };
-    \path [line] (simulations) -- (weights);
-    \path [line] (prior) -- (weights);
-
-    \node [derived] (aggregated) at (4, -6) {
-        Aggregated distribution\\
-        over outcomes\\
-        $\hat{p}(u | x_i)$
-    };
-    \path [line] (combined) -- (aggregated);
-    \path [line] (weights) -- (aggregated);
 \end{tikzpicture}

### 2022-01-27 papers/paper/preprint.tex
Track some suggestions from Klaus

@@
 \fancyhf{}
-\fancyhead[L]{CONFIDENTIAL DRAFT -- Doss-Gollin \& Keller}
+\fancyhead[L]{CONFIDENTIAL DRAFT by J. Doss-Gollin \& K. Keller}
 \fancyhead[R]{\DTMnow}
@@
 \author[1]{James Doss-Gollin}
-\author[2,3,4]{Klaus Keller}
+\author[2]{Klaus Keller}
 \affil[1]{Department of Civil and Environmental Engineering, Rice University}
-\affil[2]{Department of Geosciences, the Pennsylvania State University}
-\affil[3]{Earth and Environmental Systems Institute, the Pennsylvania State University}
-\affil[4]{Thayer School of Engineering, Dartmouth College}
+\affil[2]{Thayer School of Engineering, Dartmouth College}
 \renewcommand*{\Affilfont}{\normalsize\normalfont}
@@
 \newacronym[]{gev}{GEV}{generalized extreme value}
+\newacronym{iid}{IID}{independent and identically distributed}
 \newacronym{lsl}{LSL}{local mean sea level}
@@
   sorting=nyt,
+  giveninits=true,
 ]{biblatex}
@@
 \begin{abstract}
-    TBD
+    \textbf{Draft in preparation to Earth's Future}.
 \end{abstract}
@@
 
-Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face far into the future.
+Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face decades into the future.
 On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
+For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.\james{Too fast on intro.
+Need to slow down and define concepts}
 On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
@@
 Be sure to note that this has parallels to lots of other engineering design guidance.
-Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve significantly on a standards-based approach.
+Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve considerably on a standards-based approach.
 
@@
 
-This suggests that either nonstationarity should be ignored, or subjective approaches for designing for nonstationarity should be adapted.
-Given that physical theories consistent with the observational record and models robustly project changes to climate hazard under warning, ignoring nonstationarity seems suboptimal.
+This motivates subjective approaches for designing for nonstationarity.
 Thus decision-makers need some subjective way to choose which possible futures to design for.
@@
 \end{description}
-Specifically, we examine this problem through three lenses.
+Specifically, we examine this problem through three lenses.\james{Which ones?}
 
-The paper is organized as follows.
+The paper is organized as follows.\james{This is confusing to the reader}
 In \cref{sec:case-study} we describe our case study in Norfolk, VA.
@@
 Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
-Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify needs for future research.
+Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.
 
@@
 We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.
-
 For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
@@
     \centering
+    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
     \begin{tabular}{l l p{3in} l}
@@
     \end{tabular}
-    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
 \end{table}
@@
 
-\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways (\gls{rcp} scenarios) times two models of ice sheet dynamics.
+\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
 Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
@@
 
-Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite}
+Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
 For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
-This time series is shown in \cref{fig:observed-surges}(a).\james{This figure needs a bit of work: subpanel plots, equalizing axes, narrower time series, and no lines between the points}
+This time series is shown in \cref{fig:observed-surges}(a).
+In (b) we use the Weibull plotting position
+\begin{equation}\label{eq:plot-pos}
+    \frac{r}{N + 1}
+\end{equation}
+where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
 
@@
     \caption{
-        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
         (a):
-        time series of historic storms.
+        Time series of historic storms.
         Purple (orange) arrows denote notable tropical cyclones (nor'easters).
         (b):
-        return periods.
-        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula .
-        Blue line shows a maximum likelihood GEV fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
+        Return periods.
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
+        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
     }
@@
 We model future storm surge risk through a probabilistic lens.
-To do this, we develop a stationary Bayesian model for storm surge:
+To do this, we develop a stationary\james{discuss / validate this -- it's cool, we have the supplemental info and can emphpasize our focus here is on SLR for didactic purposes} Bayesian model for storm surge:
 \begin{equation}\label{eq:surge-model}
@@
 \end{equation}
-where $y'_t$ is the storm surge (above sea level) in year $t$.
+where $y'_t$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively.
 
-When GEV distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
+When \gls{gev} distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
 Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
@@
 However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
-Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).
+Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).\james{Expand on this ``probabilistic inversion''}
 These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}
@@
 The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
-However, the convolution of the HAZUS depth-damage equation with the GEV posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
+However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
 \begin{enumerate}
@@
 \end{equation}
-where $\gamma$ is the 1 minus the discount rate.
-For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2, thus $\gamma = 0.98$.
+where the discount rate is $1 - \gamma$.
+For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2\%, thus $\gamma = 0.98$.
 Neglecting uncertainty in house lifetime, we model flood damages from 2022 to 2071, thus $T=50$.\james{This has not been the case thus far}
@@
     \centering
-    \includegraphics[width=4in]{bayes-rdm.pdf}
+    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
     \caption{
-        Outline of the generally applicable decision theoretic framework developed.
-        Boxes are labeled in the order in which they are presented in the text: we begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
+        Outline of the proposed decision theoretic framework.
+        Boxes are labeled in the order in which they are presented in the text.
+        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
         In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
@@
 
-The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \acrfullpl{sow}.
-The first that we consider is exploratory modeling, which sheds light the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''
+The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \glspl{sow}.
+We first consider exploratory modeling, which sheds light on the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''\james{Maybe this doesn't need to be a questino}
 
-\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\klaus{Would it be better here to have a figure that has $\Delta h$ on the $y$ axis and uses expected lifetime cost to color the points?}
+\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\james{Add after updating this figure}
 On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
@@
     \caption{
-        Scenario maps: one dot represents one \gls{sow}.
-        Some work is still needed -- I think that changing to four values of $\Delta h$ and showing two or three different values of $h_0$ might be useful (alternatively we could vary the discount rate $\gamma$).
-        Also need to rescale the $y$-axis by house value.
-        It would also be cool to create an interactive online version of this figure that lets you scroll over a particular dot and learn which scenario it's from!
+        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
+        One dot represents one \gls{sow}.
         %TODO
@@
 
-An alternative interpretation of the available \glspl{sow} is as IID draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
-This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as IID draws from the distribution of outcomes.
+An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
+This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.
 
 A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
-Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be IID samples from the distribution of outcomes (box e).
+Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
 Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
@@
     \caption{
-        Tradeoffs between (L) construction cost and expected lifetime costs or (R) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
+        Need a summary sentence.
+        %TODO
+        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
     }
@@
 
-Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.
+Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.\james{This also needs to go in the abstract}
 
 Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{empirical}(\Omega)$, does not reflect our belief about their true distribution, $p_\mathrm{true}(\Omega)$.
-Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.
+Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.\james{Tighten up this writing}
 This is analagous to the deliberate over- or under-sampling of certain groups in survey design, and so following standard practice in survey design we develop a re-weighting approach to correct for the over- or under-sampling of different regions of the survey space.
 
-To calculate these weights, we apply a grid-based discretization scheme.
+We apply a grid-based discretization scheme to calculate these weights.
 First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
@@
 
-The primary difficulty with \cref{eq:weights} is that the true probability distribution $p_\mathrm{true}(\phi)$ is not only unknown but deeply uncertain.
+The primary difficulty with \cref{eq:weights} is that the true probability distribution $p_\mathrm{true}(\phi)$ is deeply uncertain.
 Because it is unknown, it must be approximated with a model.
-And because it is deeply uncertain, multiple models for $p_(\Omega)$ should be used.
-But whereas the $M_k$ defined in \cref{sec:multiple-pdf} represented the raw input data (four \gls{rcp} scenarios times two models of ice sheet dynamics), we define the $M_k$ here to be subjective models for mean sea level in 2100.
+And because it is deeply uncertain, multiple models\james{multiple priors?} for $p_(\Omega)$ should be used.
+But whereas the $M_k$ defined in \cref{sec:multiple-pdf} represented the raw input data (four \gls{rcp} scenarios times two models of ice sheet dynamics), we define the $M_k$ here to be subjective models for mean sea level in 2100.\james{Are these models or priors?}
 
-Need a few sentences here on prior uses of expert judgement and subjective belief.\klaus{Any suggested citations?}
+Need a few sentences here on prior uses of expert judgement and subjective belief.
 The key point to get across is what a subjective belief means in the Bayesian context \citep{savage:1954,gelman_philosophy:2013,gelman_subjectiveobjective:2017}.
@@
 
-Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.
+Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.\james{Lee et al: Lee, Haran, Pollard, Keller}
 
 Most fundamentally, we are thinking about robustness a bit differently than usual.
-The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.
+The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.\james{This is a simplification}
 We are thinking about robustness to different modeling structures and assumptions.
 This has links to the ``rival framings'' literature.
-Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the tradeoff curves as shown in figures XXXX (in \cref{sec:results}).
+Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the trade-off curves as shown in figures XXXX (in \cref{sec:results}).
 
@@
               \item We should also cite \citet{quinn_exploratory:2020}, which makes the point that it matters how you sample \glspl{sow}.
-              \item Exploratory modeling can demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
+              \item Exploratory modeling can help to demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
               \item You can do scenario discovery
@@
           \begin{enumerate}
-              \item The observation that many of these mechanisms cannot be represented by a single objective \gls{pdf} has motivated many criticisms of the application of \gls{bdt} to planning problems.
+              \item The observation that many of these mechanisms cannot be represented by a single objective \gls{pdf} has motivated many criticisms of the application of \gls{bdt} to planning problems.\james{Cite some (from thesis)}
                     Yet \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
@@
                     More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
-                    Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
-                    Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
+                    Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.\james{This should go to intro}
+                    Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.\james{Explain this to the audience}
                     Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
@@
               \item Our prior $p(s)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
-              \item We neglect true ignorance \citep{knight_risk:1921}.
+              \item We neglect true ignorance \citep{knight_risk:1921}.\james{Perhaps ``unknown unknowns''}
               \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
@@
           \begin{enumerate}
-              \item A goal is to make -- intrinsically flawed and subjective -- modeling choices as transparent as possible.
+              \item A goal is to make imperfect and subjective modeling choices as transparent as possible.
 Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions\james{Focus on this!}
@@
                     \begin{enumerate}
-                        \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!
+                        \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!\james{Let's emphasize this -- really important to policy analysis under deep uncertainty}
                         \item Model structure uncertainties
@@
 
-Additionally, the floodplain maps used for decision supportare silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
+Additionally, the floodplain maps used for decision support are silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
 Both risk-based design and standards-based approaches are complicated by nonstationarity.
 On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
-On the other, efforts to find objective methods for quantifying nonstationarity, \eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[see][]{keller_management:2021}.
+On the other, efforts to find objective methods for quantifying nonstationarity, \eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[see][]{keller_management:2021}.\james{This goes earlier -- see Klaus's added text}
 
@@
     \item Rice University (for startup)
-    \item Tor Erlend Fjelde for helpful advice on implementing the GEV prior using \texttt{Turing.jl}.
-    \item For many journals, a code and data statement will go here.
+    \item Dartmouth College (for startup)
+    \item Tor Erlend Fjelde for helpful advice on implementing the \gls{gev} prior using \texttt{Turing.jl}.
+    \item For many journals, a code and data statement will go here.\james{Keep an eye on this}
 \end{enumerate}
@@
 
-See journal requirements and format
+See journal requirements and format\james{Keep an eye on this}
 

### 2022-01-27 HouseElevation/src/lsl.jl
Start putting the source code in its own package so that code is more flexible and reliable

@@ -1,0 +1,123 @@
+using Base: vcat
+using CSV
+using DataFrames, DataFramesMeta
+using Dates
+using Downloads
+using ProgressBars
+using StatsBase
+using Unitful
+
+"""A `WaterLevelStation` stores the basic information about a Tides and Currents station data query"""
+Base.@kwdef struct TidesAndCurrentsRecord
+    datum::String = "NAVD"
+    gage_id::String = "8638610"
+    gage_name::String = "Sewells Point, VA"
+    units::String = "metric"
+    time_zone::String = "GMT"
+end
+
+"""
+An `HourlyGageRecord` stores all the data from a tide gauge
+"""
+struct HourlyGageRecord
+    t::Vector{<:Dates.DateTime}
+    lsl::Vector{<:Unitful.Length}
+    stn::TidesAndCurrentsRecord
+end
+
+"""
+An `AnnualGageRecord` stores mean sea level and annual maximum storm surge for each year
+"""
+Base.@kwdef struct AnnualGageRecord
+    t::Vector{<:Int}
+    max_surge::Vector{<:Unitful.Length}
+    msl::Vector{<:Unitful.Length}
+    stn::TidesAndCurrentsRecord
+end;
+
+"""
+Get the URL for the NOAA Tides and Currents data; sensible defaults are applied
+"""
+function get_url(stn::TidesAndCurrentsRecord, year::Int)
+    return "https://api.tidesandcurrents.noaa.gov" *
+           "/api/prod/datagetter?product=hourly_height&application=NOS.COOPS.TAC.WL" *
+           "&begin_date=$(year)0101" *
+           "&end_date=$(year)1231" *
+           "&datum=$(stn.datum)" *
+           "&station=$(stn.gage_id)" *
+           "&time_zone=$(stn.time_zone)" *
+           "&units=$(stn.units)" *
+           "&format=csv"
+end
+
+"""
+Get the filename in which a given storm surge dataset is to be stored
+"""
+function get_fname(stn::TidesAndCurrentsRecord, year::Int)
+    cache_dir = abspath(datadir("processed", "gage"))
+    if !isdir(cache_dir)
+        mkpath(cache_dir)
+    end
+    return joinpath(
+        cache_dir,
+        "tidesandcurrents-$(stn.gage_id)-$(year)-$(stn.datum)-$(stn.time_zone)-$(stn.units).csv",
+    )
+end
+
+"""
[…]

### 2022-02-13 papers/paper/preprint.tex
Working on getting figure captions together

@@
 \usepackage{authblk}
-\title{Which Scenario Should We Design For? Insights from House Elevation for the Multiple PDF Problem}
+\title{Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties}
 \author[1]{James Doss-Gollin}
@@
 \newacronym{bfe}{BFE}{base flood elevation}
+\newacronym[]{cdf}{CDF}{cumulative distribution function}
 \newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
@@
         Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
-        Discount rate  & $1-\gamma$       & Fixed at 2\%                                                                         & \ref{sec:ead}         \\
+        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \ref{sec:ead}         \\
         Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
         Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
-        Initial height & $h_0$            & Deterministic: \SI{0.5}{ft} above \gls{bfe} unless otherwise noted                   & TBD                   \\
+        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 & TBD                   \\
+        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                & TBD                   \\
+        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   & TBD                   \\
+        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       & TBD                   \\
         \bottomrule
@@
     \centering
-    \includegraphics[width=5in]{msl_boxplots.pdf}
+    %\includegraphics[width=5in]{msl_boxplots.pdf}
     \caption{
@@
     \centering
-    \includegraphics[height=1.5in]{historic_surge.pdf}%
-    \includegraphics[height=1.5in]{surge_return_period.png}
+    %\includegraphics[height=1.5in]{historic_surge.pdf}%
+    %\includegraphics[height=1.5in]{surge_return_period.png}
     \caption{
@@
 \begin{figure}
-    \includegraphics[width=\textwidth]{scenario_maps.png}
+    %\includegraphics[width=\textwidth]{scenario_maps.png}
     \caption{
@@
     \centering
-    \includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
-    \includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
+    %\includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
+    %\includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=3in]{lsl_priors.pdf}~
-    \includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
+    %\includegraphics[width=3in]{lsl_priors.pdf}~
+    %\includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
     \caption{
@@
     \centering
-    \includegraphics[width=6in]{inference_weights.pdf}
+    %\includegraphics[width=6in]{inference_weights.pdf}
     \caption{
@@
 \printbibliography
+
+\appendix
+\newcommand{\hbAppendixPrefix}{S}
+\renewcommand{\thefigure}{\hbAppendixPrefix\arabic{figure}}
+\setcounter{figure}{0}
+\renewcommand{\thetable}{\hbAppendixPrefix\arabic{table}}
+\setcounter{table}{0}
+\renewcommand{\theequation}{\hbAppendixPrefix\arabic{equation}}
+\setcounter{equation}{0}
+
+\newpage
+\section{Supplemental Figures}
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{cost-depth-damage}
+    \caption{
+        Depth-damage relationship.
+        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA.
+        Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
+    }\label{fig:cost-depth-damage}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{cost-expected-damage-emulator}
+    \caption{
+        As discussed in XXX, we model expected annual damages as a function of the house's elevation relative to \gls{lsl}. % TODO: ref equation
+    }\label{fig:cost-expected-damage-emulator}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{cost-up-front}
+    \caption{
+        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
+        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial fee.
+        Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
+    }\label{fig:cost-up-front}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{grid-sketch}
+    \caption{
+        Schematic of equation XXX used to calculate weights for each sample. % TODO: reference equation
+        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
+        As discussed in section XXX, the weights (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution  (horizontal dotted lines) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+% TODO: reference section
+    }\label{fig:grid-sketch}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{lsl-evolution}
+    \caption{
+        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
+        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
+        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
+    }\label{fig:lsl-evolution}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{lsl-priors-weights}
+    \caption{
+        Subjective priors for local sea level.
+        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
+        The \glspl{pdf} of these subjective priors are shown in panel (A).
+        In panels (B)-(D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
+        Specifically, (B)-(D) show the average weight given to each model by each of the three subjective priors.
+        The ``Slow SLR'' prior places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
+        The ``Fast SLR'' prior places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
+        The ``Intermediate SLR'' scenario places approximately equal weight on all models.
+    }\label{fig:lsl-priors-weights}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{scenario-map-height-slr}
+    \caption{
+        Expected lifetime total cost (eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
+        The best outcomes (yellow) arise when the house is not elevated and sea level rise is minimal (bottom left corner).
+        The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
+        In general, a height increase of $<\SI{3}{ft}$ can be considered unrealistic \citep{zarekarizi_suboptimal:2020}.
+        Values are sensitive to model constants; see \cref{tab:uncertainties}.
+    }\label{fig:scenario-map-height-slr}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
+    \caption{scenario-map-slr-cost}\label{fig:scenario-map-slr-cost}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{surge-gev-priors}
+    \caption{surge-gev-priors}\label{fig:surge-gev-priors}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-obs-return}
+    \caption{surge-obs-return}\label{fig:surge-obs-return}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-posterior-chains}
+    \caption{surge-posterior-chains}\label{fig:surge-posterior-chains}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-prior-chains}
+    \caption{surge-prior-chains}\label{fig:surge-prior-chains}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{surge-synthetic-data-experiment}
+    \caption{surge-synthetic-data-experiment}\label{fig:surge-synthetic-data-experiment}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-test-statistics}
+    \caption{surge-test-statistics}\label{fig:surge-test-statistics}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
+    \caption{tradeoffs-by-prior}\label{fig:tradeoffs-by-prior}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
+    \caption{tradeoffs-by-rcp}\label{fig:tradeoffs-by-rcp}
+\end{figure}
+
 \end{document}

### 2022-02-13 scripts/06-combining.jl

@@ -16,7 +16,7 @@ function plot_grid_scheme()
         xticks=(vcat(ψ, grids), vcat(ψ_labels, grid_labels)),
         linewidth=3,
         color=colors[1],
-        label="True Distribution",
+        label="Target Distribution",
         xlims=xlimits,
         ylabel="CDF",
         xlabel="Mean Sea Level in 2100 [ft]",

### 2022-02-13 papers/paper/preprint.tex
Caption all figures and make a few tweaks to figures

@@
 
-The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the upfront cost plus the discounted sum of expected annual damages
+The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the up-front cost plus the discounted sum of expected annual damages
 \begin{equation}\label{eq:led}
@@
 \begin{description}
-    \item[upfront cost] reflects the cost of elevating a house.
+    \item[up-front cost] reflects the cost of elevating a house.
 Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
@@
     \item[lifetime expected damages] is discussed in \cref{sec:led}
-    \item[lifetime expected costs] is the sum of lifetime expected damages and upfront costs
+    \item[lifetime expected costs] is the sum of lifetime expected damages and up-front costs
     \item[the expected benefit-to-cost ratio] is defined as the ratio of construction cost to damages averted, relative to not elevating at all:
@@
 \newpage
-\section{Supplemental Figures}
+\section{Supplemental figures}
 \begin{figure}
@@
     \caption{
-        Expected lifetime total cost (eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
+        Expected total lifetime cost (damages plus up-front cost; eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
         The best outcomes (yellow) arise when the house is not elevated and sea level rise is minimal (bottom left corner).
@@
     \includegraphics[width=\textwidth]{scenario-map-slr-cost}
-    \caption{scenario-map-slr-cost}\label{fig:scenario-map-slr-cost}
+    \caption{
+        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost; eq.XXX) as a function of \gls{slr} in 2100 for several values of $\Delta h$ and $h_0$.
+        Colors indicate the density of \glspl{sow}; this data is often presented as a scatterplot (CITE) but we bin to a 2D histogram. % TODO: cite scenario maps
+        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
+        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
+        In all cases, elevating the house reduces the variance in total lifetime cost.
+        Values are sensitive to model constants; see \cref{tab:uncertainties}.
+    }\label{fig:scenario-map-slr-cost}
 \end{figure}
@@
     \includegraphics[width=4in]{surge-gev-priors}
-    \caption{surge-gev-priors}\label{fig:surge-gev-priors}
+    \caption{
+        Prior distributions for annual maximum storm surge.
+        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following REFERENCE.
+% TODO: cite
+        See ex.~XYQ for details. % TODO: cross-ref equation
+        We apply priors over the 2, 10, 100, and 500 year events.
+    }\label{fig:surge-gev-priors}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{surge-obs-return}
-    \caption{surge-obs-return}\label{fig:surge-obs-return}
+    \caption{
+        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE). % TODO: cite source
+        (A):
+        Time series of historic storms.
+        Blue (green) arrows denote notable tropical cyclones (nor'easters).
+        (B):
+        Return periods.
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
+        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian GEV fit.
+    }\label{fig:surge-obs-return}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{surge-posterior-chains}
-    \caption{surge-posterior-chains}\label{fig:surge-posterior-chains}
+    \caption{
+        \Gls{mcmc} plots for posterior draws from the storm surge model.
+        We draw \num{10000} samples by running four chains of \num{3500} iterations each and discarding the first \num{1000}.
+        The mixing of the chains is consistent with, though does not guarantee, convergence.
+    }\label{fig:surge-posterior-chains}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{surge-prior-chains}
-    \caption{surge-prior-chains}\label{fig:surge-prior-chains}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{surge-synthetic-data-experiment}
-    \caption{surge-synthetic-data-experiment}\label{fig:surge-synthetic-data-experiment}
+    \caption{
+        As \cref{fig:surge-posterior-chains} but for draws from the prior distribution.
+    }\label{fig:surge-prior-chains}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{surge-test-statistics}
-    \caption{surge-test-statistics}\label{fig:surge-test-statistics}
+    \caption{
+        Posterior predictive checks for the stationary GEV storm surge model (section XXX). % TODO: reference section
+        Each panel shows a different test statistic: partial autocorrelation at lags 1 and 2; sample maximum; sample minimum; sample median; and Mann-Kendall trend test statistic.
+        The histograms show the distribution of each test statistic from the posterior predictive distribution.
+        Orange lines show the test statistic's value in the observed data.
+        Observed values near the mode of the posterior predictive distribution are consistent with, but do not guarantee, a good fit.
+        For further discussion of posterior predictive checks, see Chapter 6 of \citet{Gelman:2014tc}.
+    }\label{fig:surge-test-statistics}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
+    \caption{
+        Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
+        (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs (eq.~XX). % TODO: ref equation
+        (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
+        Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
+        Colored lines highlight three representative models for emphasis.
+    }\label{fig:tradeoffs-by-rcp}
 \end{figure}
@@
     \includegraphics[width=\textwidth]{tradeoffs-by-prior}
-    \caption{tradeoffs-by-prior}\label{fig:tradeoffs-by-prior}
+    \caption{
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the synthesizing Bayesian models.
+    }\label{fig:tradeoffs-by-prior}
 \end{figure}
 
-\begin{figure}
+\section{Supplemental tables}
+
+\begin{table}[h]
     \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
-    \caption{tradeoffs-by-rcp}\label{fig:tradeoffs-by-rcp}
-\end{figure}
+    \caption{
+        Diagnostic statistics for the \gls{mcmc} sampling for the storm surge posterior draws.
+        Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.
+        In general, a $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence.
+    }\label{tab:surge-posterior-mcmc-diagnostics}
+    \input{../../plots/surge-posterior-mcmc-diagnostics.tex}
+\end{table}
+
+\begin{table}[h]
+    \centering
+    \caption{As \cref{tab:surge-posterior-mcmc-diagnostics} but for draws from the prior distribution.}\label{tab:surge-prior-mcmc-diagnostics}
+    \input{../../plots/surge-prior-mcmc-diagnostics.tex}
+\end{table}
 

### 2022-02-13 HouseElevation/src/surge-bayes.jl

@@ -1,7 +1,9 @@
 using Distributions
 using DynamicPPL
-using Optim
+using Latexify
+using LaTeXStrings
 using MCMCChains
+using Optim
 using StatsBase
 using Turing
 
@@ -85,7 +87,22 @@ end
 """Save your diagnostics to file"""
 function write_diagnostics(fit, fname::String)
     df = DataFrame(MCMCChains.summarize(fit))
-    return CSV.write(fname, df)
+    rename!(
+        df,
+        "parameters" => L"\textrm{Parameters}",
+        "mean" => L"\textrm{Mean}",
+        "std" => L"\textrm{Stdev.}",
+        "naive_se" => L"\textrm{Naive SE}",
+        "mcse" => L"\textrm{MCSE}",
+        "ess" => L"\textrm{ESS}",
+        "rhat" => L"$\hat{R}$",
+        "ess_per_sec" => L"\textrm{ESS per second}",
+    )
+    tex_str = latexify(df; env=:table, fmt="%.2f")
+    open(fname, "w") do io
+        write(io, tex_str)
+    end
+    return true
 end
 
 """Sample from the posterior predictive of the GEV"""

### 2022-02-13 scripts/01-surge-modeling.jl

@@ -81,7 +81,7 @@ function plot_surge_prior_chains()
     display(prior_fits)
 
     HouseElevation.write_diagnostics(
-        prior_fits, plots_dir("surge-prior-mcmc-diagnostics.csv")
+        prior_fits, plots_dir("surge-prior-mcmc-diagnostics.tex")
     )
 
     p = plot(prior_fits)
@@ -171,7 +171,7 @@ function plot_surge_posterior_chains(fits::T) where {T<:HouseElevation.MCMCChain
     display(fits)
 
     HouseElevation.write_diagnostics(
-        fits, plots_dir("surge-posterior-mcmc-diagnostics.csv")
+        fits, plots_dir("surge-posterior-mcmc-diagnostics.tex")
     )
 
     p = plot(fits)
@@ -310,7 +310,6 @@ function plot_surge_posterior_return(
 
     xp, ys = HouseElevation.weibull_plot_pos(surge_ft)
     scatter!(p, 1 ./ xp, ys; label="Obs (Weibull Plot Pos.)", color=colors[1], alpha=1)
-    savefig(p, plots_dir("surge-synthetic-data-experiment.pdf"))
     return p
 end
 

### 2022-02-13 scripts/04-scenario-maps.jl

@@ -130,6 +130,8 @@ function plot_scenario_map_slr_cost(;
     fake_plot[1][:yaxis][:showaxis]
     l = @layout [grid(N_row, N_col) a{0.1w}]
 
+    add_panel_letters!(plots)
+
     p = plot(
         vcat([plots[row, col] for row in 1:N_row for col in 1:N_col], fake_plot)...;
         layout=l,

### 2022-02-13 scripts/05-tradeoffs-by-rcp.jl

@@ -10,12 +10,14 @@ function plot_rcp_tradeoffs(
 ) where {T<:Real,A<:Unitful.Area}
 
     # define constant
-    models = (
+    model_rcp_plot = (
         (rcp=2.6, model="BRICK Slow"),
-        (rcp=4.5, model="BRICK Fast"),
         (rcp=6.0, model="K14"),
         (rcp=8.5, model="DP16"),
     )
+
+    all_rcp = unique([si.rcp for si in s])
+    all_models = unique([si.model for si in s])
 
     total_cost = map(ui -> (ui.led_usd + ui.upfront_cost_usd) / house_value_usd, u)
     Δh_ft = ustrip.(u"ft", x)
@@ -28,13 +30,15 @@ function plot_rcp_tradeoffs(
 
     # what we're going to save
     p_archive = []
+    varnames = [
+        "Expected Lifetime Cost [% House Value]",
+        "Lifetime Expected Damages [% House Value]",
+    ]
+    vars = [total_cost, led]
 
     for (var, varname) in zip(
-        [total_cost, led],
-        [
-            "Expected Lifetime Cost [% House Value]",
-            "Lifetime Expected Damages [% House Value]",
-        ],
+        vars,
+        varnames,
     )
         p = plot(;
             xlabel=L"Height Increase $\Delta h$ [ft]",
@@ -45,26 +49,26 @@ function plot_rcp_tradeoffs(
             top_margin=12.5Plots.mm,
             left_margin=7.5Plots.mm,
             bottom_margin=7.5Plots.mm,
-            legend=:topright,
+            legend=ifelse(var == first(vars), :topright, false),
         )
-        for (model, color) in zip(models, colors)
-            w = weights([(si.rcp == model.rcp) & (si.model == model.model) for si in s])
-            cond = vec(mean(var, w; dims=1))
-            plot!(
-                p,
-                Δh_ft,
-                cond;
-                label="RCP $(model.rcp), $(model.model)",
-                color=color,
-                linewidth=3,
-            )
-        end
-        if var == total_cost
-            for (model, color) in zip(models, colors)
-                w = weights([(si.rcp == model.rcp) & (si.model == model.model) for si in s])
+
+        i = 1 # counter for colors
+        for rcp in all_rcp
+            for model in all_models
+                plot_in_color = (rcp=rcp, model=model) in model_rcp_plot
+                w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
[…]

### 2022-02-13 scripts/06-combining.jl

@@ -148,13 +148,13 @@ function plot_prior_tradeoffs(
     # what we're going to save
     p_archive = []
 
-    for (var, varname) in zip(
-        [total_cost, led],
-        [
-            "Expected Lifetime Cost [% House Value]",
-            "Lifetime Expected Damages [% House Value]",
-        ],
-    )
+    # to plot
+    vars = [total_cost, led]
+    varnames = [
+        "Expected Lifetime Cost [% House Value]",
+        "Lifetime Expected Damages [% House Value]",
+    ]
+    for (var, varname) in zip(vars, varnames)
         p = plot(;
             xlabel=L"Height Increase $\Delta h$ [ft]",
             ylabel=varname,
@@ -164,21 +164,12 @@ function plot_prior_tradeoffs(
             top_margin=12.5Plots.mm,
             left_margin=7.5Plots.mm,
             bottom_margin=7.5Plots.mm,
-            legend=:topright,
+            legend=ifelse(var == first(vars), :topright, false),
         )
         for (prior, color) in zip(priors, colors)
             w = HouseElevation.make_weights(prior.dist)
             cond = vec(mean(var, w; dims=1))
             plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=3)
-        end
-
-        if var == total_cost
-            for (prior, color) in zip(priors, colors)
-                w = HouseElevation.make_weights(prior.dist)
-                cond = vec(mean(var, w; dims=1))
-                idx = argmin(cond)
-                scatter!(p, [Δh_ft[idx]], [cond[idx]]; label=false, color=color)
-            end
         end
 
         # add the cost on upper x axis

### 2022-02-24 papers/paper/preprint.tex
better figures color schemes and some updates to text

@@
 \begin{abstract}
-    \textbf{Draft in preparation to Earth's Future}.
+    The policy and engineering tools used to inform climate adaptation etc rely on standards and cost-benefit analyses that depend on risk probabilities.
+    One example is house elevation: the recommended elevation is a standard (BFE plus a foot) and funding decisions rely on CBA.
+    Under stationarity, the probabilities for this can almost be agreed upon.
+    Under nonstationarity, there are deep uncertainties.
+    We show that the way these deep uncertainties are synthesized into CBA matters a lot.
+    Thus, static decisions are subject to deep uncertainties
 \end{abstract}
@@
     \centering
-    \caption{Summary of uncertainties modeled and their treatment.}\label{tab:uncertainties}
+    \caption{
+        Summary of parameters, their notation, and how their uncertainty is modeled.
+    }\label{tab:uncertainties}
     \begin{tabular}{l l p{3in} l}
         \toprule
-        Name             & Symbol           & Conceptual framework                                                                 & Section               \\
+        Name             & Symbol           & Uncertainty                                                                          & See more                     \\
         \midrule
-        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \ref{sec:sea-level}   \\
-        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \ref{sec:storm-surge} \\
-        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \ref{sec:ead}         \\
-        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \ref{sec:ead}         \\
-        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \ref{sec:sow-metrics} \\
-        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 & TBD                   \\
-        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                & TBD                   \\
-        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   & TBD                   \\
-        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       & TBD                   \\
+        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \cref{sec:sea-level}         \\
+        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \cref{sec:storm-surge}       \\
+        Discount rate    & $1-\gamma$       & Fixed at 3\%                                                                         & \cref{sec:led}               \\
+        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
+        Elevation cost   & $C(\Delta h)$    & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
+        Initial height   & $h_0$            & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
+        House floor area &                  & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
+        Structural value &                  & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
+        House lifetime   &                  & Deterministic: 70 years unless otherwise noted                                       &                              \\
         \bottomrule
@@
         We model a single decision (yellow).
-        For each combination of \gls{sow} and policy, the system model (gray) is used to calculate metrics describing the performance of the policy over the selected \gls{sow}.
+        For each combination of \gls{sow} and policy, the system model (blue) is used to calculate metrics (orange) describing the performance of the policy over the selected \gls{sow}.
     }\label{fig:xlrm}
@@
 
-\Cref{fig:msl-boxplots} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
+\Cref{fig:lsl-evolution} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
 Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
@@
     \centering
-    %\includegraphics[width=5in]{msl_boxplots.pdf}
+    \includegraphics[width=\textwidth]{lsl-evolution}
     \caption{
-        Sea level rise projections use the BRICK model \citep{wong_brick0.2:2017,ruckert_coastal:2019}.
-        We consider \emph{eight time-dependent PDFs of sea level rise}: four \gls{rcp} scenarios $\times$ two parameterizations of ice sheet processes.
-    }\label{fig:msl-boxplots}
+        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
+        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
+        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
+    }\label{fig:lsl-evolution}
 \end{figure}
 
-For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \acrfull{sow}}.
+For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \gls{sow}}.
 
@@
 For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
-This time series is shown in \cref{fig:observed-surges}(a).
+This time series is shown in \cref{fig:surge-obs-return}(a).
 In (b) we use the Weibull plotting position
@@
     \centering
-    %\includegraphics[height=1.5in]{historic_surge.pdf}%
-    %\includegraphics[height=1.5in]{surge_return_period.png}
+    \includegraphics[width=\textwidth]{surge-obs-return}
     \caption{
-        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE).
-        (a):
+        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+        (A):
         Time series of historic storms.
-        Purple (orange) arrows denote notable tropical cyclones (nor'easters).
-        (b):
+        Blue (green) arrows denote notable tropical cyclones (nor'easters).
+        (B):
         Return periods.
         Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
-        Blue line shows a maximum likelihood \gls{gev} fit and the gray lines show Monte Carlo draws from the full posterior distribution of the Bayesian GEV fit.
-    }
-    \label{fig:observed-surges}
+        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit.
+    }\label{fig:surge-obs-return}
 \end{figure}
@@
 Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
-\begin{equation}
+\begin{equation}\label{eq:ead}
     \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
@@
     \centering
-    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
+    \includegraphics[width=4in]{bayes-rdm.pdf}
     \caption{
         Outline of the proposed decision theoretic framework.
-        Boxes are labeled in the order in which they are presented in the text.
-        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling, \ie boxes  (a), (b), and (c), to analyze the elevation problem.
-        In \cref{sec:multiple-pdf} we discuss Boxes (d) and (e) to understand scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem''.
-        Finally in \cref{sec:synthesizing} we add boxes (f) and (g) to synthesize across multiple \glspl{pdf}.
+        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
+        Next in \cref{sec:multiple-pdf} we use scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem'' (region 2).
+        Finally in \cref{sec:synthesizing} propose a method for weighting \glspl{sow} to synthesize across multiple \glspl{pdf} (region 3).
     }\label{fig:flowchart}
@@
 
-\begin{figure}
-    %\includegraphics[width=\textwidth]{scenario_maps.png}
-    \caption{
-        Scenario maps showing the dependence of average lifetime cost as a function of \gls{slr} in 2100 for several values of $\Delta h$.
-        One dot represents one \gls{sow}.
-        %TODO
-    }\label{fig:scenario-map}
-\end{figure}
-
 The exploratory modeling approach shown in \cref{fig:scenario-map} shows that shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible, but the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
@@
 
-An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:msl-boxplots}.
+An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:lsl-evolution}.
 This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.
@@
     \centering
-    %\includegraphics[width=2.5in]{tradeoffs_height_totalcost_byscenario.pdf}~
-    %\includegraphics[width=2.5in]{tradeoffs_height_iqr_byscenario.pdf}
+    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
     \caption{
-        Need a summary sentence.
-        %TODO
-        Trade-offs between (a) construction cost and expected lifetime costs or (b) construction cost and the uncertainty of future lifetime costs depend on the PDF selected.
-    }
-    \label{fig:tradeoffs}
+        Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
+        (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
+        (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
+        Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
+        Colored lines highlight three representative models for emphasis.
+    }\label{fig:tradeoffs-by-rcp}
 \end{figure}
 
-For example, \cref{fig:tradeoffs}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
-Similarly, \cref{fig:tradeoffs} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
+Similarly, \cref{fig:tradeoffs-by-rcp} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
 This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp} 8.6 with fast ice sheet dynamics).
@@
 
-Calculating either of the quantities shown in \cref{fig:tradeoffs} \emph{requires synthesizing across \glspl{sow}}.
+Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} \emph{requires synthesizing across \glspl{sow}}.
 However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
@@
 $$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
-we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:weighting}.
+we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:grid-sketch}.
 \begin{equation}\label{eq:weights}
@@
     \centering
-    \includegraphics[width=4in]{../img/weighting.pdf}
+    \includegraphics[width=\textwidth]{grid-sketch}
     \caption{
-        Schematic of re-weighting scheme defined in \cref{eq:weights} for $J=4$.
-        Top panel shows $p_\mathrm{true}$ and bottom panel shows $F_\mathrm{true}$.
-        Vertical black lines show $\frac{1}{2}\qty[\psi_j + \psi_{j+1}]$ for $j=1, \ldots, J-1$.
-        Green text shows $w_1, \ldots, w_4$ and how they are calculated from $F_\mathrm{true}$.
-        Need to update this figure using real data.
-        %TODO
-    }\label{fig:weighting}
+        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
+        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
+        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+    }\label{fig:grid-sketch}
 \end{figure}
@@
 Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
-\Cref{fig:tradeoffs} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
+\Cref{fig:tradeoffs-by-prior} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
 
@@
     \centering
-    %\includegraphics[width=3in]{lsl_priors.pdf}~
-    %\includegraphics[width=3in]{tradeoffs_height_totalcost_byprior.pdf}
+    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
     \caption{
-        The prior model averaging approach (\cref{fig:flowchart}), uses a probability distribution representing subjective belief to average insights across each PDF available.
-        Alternative priors can provide robustness checks.
-        Need to add panel labels, update data, \etc.
-        %TODO
-    }\label{fig:tradeoffs}
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the synthesizing Bayesian models.
+    }\label{fig:tradeoffs-by-prior}
 \end{figure}
@@
 This technique can also be used as a diagnostic tool.
-In \cref{fig:model-average-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}
+In \cref{fig:lsl-priors-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}
 
@@
     \centering
-    %\includegraphics[width=6in]{inference_weights.pdf}
+    \includegraphics[width=\textwidth]{lsl-priors-weights}
     \caption{
-        The average weight assigned to \glspl{sow} from each model (\gls{rcp} scenario and ice sheet parameterization) for both priors considered.
-        If the sampling distribution (green line) is used, then the weight assigned to each \gls{sow} is the same, regardless of whether it is over- or under-sampled.
-    }\label{fig:model-average-weights}
+        Subjective priors for local sea level.
+        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
+        The \glspl{pdf} of these subjective priors are shown in panel (A).
+        In panels (B-D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
+        Specifically, (B-D) show the average weight given to each model by each of the three subjective priors.
+        The ``Slow SLR'' prior (B) places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
+        The ``Intermediate SLR'' (C) prior places approximately equal weight on all models.
+        The ``Fast SLR'' prior (D) places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
+    }\label{fig:lsl-priors-weights}
 \end{figure}
@@
 Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
-We introduce an approach to bridge this divide through a computationally efficient method that weights each state of the world based on a probability distribution over possible futures.
+We introduce an approach to bridge this divide through a computationally efficient method that weights each \gls{sow} based on a probability distribution over possible futures.
 Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
@@
 
-\begin{enumerate}
-    \item Supplementary figures and analysis online
-    \item Link to live repository on GitHub
-    \item Link to code DOI on Zenodo
-\end{enumerate}
-
 \subsection*{Acknowledgements}
 
-\begin{enumerate}
-    \item MARISA\james{details}
-    \item PSIRC\james{details}
-    \item Rice University (for startup)
-    \item Dartmouth College (for startup)
-    \item Tor Erlend Fjelde for helpful advice on implementing the \gls{gev} prior using \texttt{Turing.jl}.
-    \item For many journals, a code and data statement will go here.\james{Keep an eye on this}
-\end{enumerate}
+This work was supported by the National Oceanic and Atmospheric Administration (NOAA) through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
+JDG thanks Rice University for support.
+KK thanks Dartmouth College for support.
+The authors thank Tor Erlend Fjelde for helpful comments.
+
+\subsection*{Code and data availability}
+
+All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
+This code is written in the open source Julia programming language.
 
@@
 
-See journal requirements and format\james{Keep an eye on this}
+JDG and KK designed the research.
+JDG wrote the codes and ran simulations.
+JDG wrote the manuscript with support from KK.
+JDG wrote the manuscript in consulation and collaboration with KK.
 
@@
     \caption{
-        As discussed in XXX, we model expected annual damages as a function of the house's elevation relative to \gls{lsl}. % TODO: ref equation
+        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{lsl}.
+        Damages ($y$ axis) are shown as a percentage of house value.
     }\label{fig:cost-expected-damage-emulator}
@@
         Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
-        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial fee.
+        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial cost.
         Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
     }\label{fig:cost-up-front}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{grid-sketch}
-    \caption{
-        Schematic of equation XXX used to calculate weights for each sample. % TODO: reference equation
-        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
-        As discussed in section XXX, the weights (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution  (horizontal dotted lines) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
-% TODO: reference section
-    }\label{fig:grid-sketch}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{lsl-evolution}
-    \caption{
-        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
-        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
-        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
-    }\label{fig:lsl-evolution}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{lsl-priors-weights}
-    \caption{
-        Subjective priors for local sea level.
-        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
-        The \glspl{pdf} of these subjective priors are shown in panel (A).
-        In panels (B)-(D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
-        Specifically, (B)-(D) show the average weight given to each model by each of the three subjective priors.
-        The ``Slow SLR'' prior places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
-        The ``Fast SLR'' prior places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
-        The ``Intermediate SLR'' scenario places approximately equal weight on all models.
-    }\label{fig:lsl-priors-weights}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{scenario-map-height-slr}
-    \caption{
-        Expected total lifetime cost (damages plus up-front cost; eq.XXX) as a function of height increase $\Delta h$ and sea level rise over the house lifetime. % TODO: reference equation
-        The best outcomes (yellow) arise when the house is not elevated and sea level rise is minimal (bottom left corner).
-        The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
-        In general, a height increase of $<\SI{3}{ft}$ can be considered unrealistic \citep{zarekarizi_suboptimal:2020}.
-        Values are sensitive to model constants; see \cref{tab:uncertainties}.
-    }\label{fig:scenario-map-height-slr}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
-    \caption{
-        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost; eq.XXX) as a function of \gls{slr} in 2100 for several values of $\Delta h$ and $h_0$.
-        Colors indicate the density of \glspl{sow}; this data is often presented as a scatterplot (CITE) but we bin to a 2D histogram. % TODO: cite scenario maps
-        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
-        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
-        In all cases, elevating the house reduces the variance in total lifetime cost.
-        Values are sensitive to model constants; see \cref{tab:uncertainties}.
-    }\label{fig:scenario-map-slr-cost}
 \end{figure}
@@
         Prior distributions for annual maximum storm surge.
-        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following REFERENCE.
-% TODO: cite
-        See ex.~XYQ for details. % TODO: cross-ref equation
-        We apply priors over the 2, 10, 100, and 500 year events.
+        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
+        See \cref{sec:storm-surge} for details.
+        For the 2, 10, 100, and 500 year events we apply Normal distributions, truncated at zero, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
     }\label{fig:surge-gev-priors}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-obs-return}
-    \caption{
-        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA (CITE SOURCE). % TODO: cite source
-        (A):
-        Time series of historic storms.
-        Blue (green) arrows denote notable tropical cyclones (nor'easters).
-        (B):
-        Return periods.
-        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
-        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian GEV fit.
-    }\label{fig:surge-obs-return}
 \end{figure}
@@
     \caption{
-        Posterior predictive checks for the stationary GEV storm surge model (section XXX). % TODO: reference section
+        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:storm-surge}).
         Each panel shows a different test statistic: partial autocorrelation at lags 1 and 2; sample maximum; sample minimum; sample median; and Mann-Kendall trend test statistic.
@@
     }\label{fig:surge-test-statistics}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
-    \caption{
-        Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
-        (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs (eq.~XX). % TODO: ref equation
-        (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
-        Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
-        Colored lines highlight three representative models for emphasis.
-    }\label{fig:tradeoffs-by-rcp}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
-    \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the synthesizing Bayesian models.
-    }\label{fig:tradeoffs-by-prior}
 \end{figure}

### 2022-02-24 tikz/xlrm.tex
a major reorganization is underway, this is just an intermediate saved point
better figures color schemes and some updates to text

@@
 \usepackage{xcolor}
-\definecolor{TurquoiseGreen}{HTML}{B8D5B8}
-\definecolor{TealBlue}{HTML}{598392}
-\definecolor{YellowOrange}{HTML}{FFB140}
-\definecolor{RichBlack}{HTML}{011627}
+\definecolor{giallo}{HTML}{F0BC42}
+\definecolor{rosso}{HTML}{8E1F2F}
+\definecolor{grigio}{HTML}{CACACC}
+\definecolor{nero}{HTML}{000000}
+\definecolor{thirds}{HTML}{31467A}
+\definecolor{orange}{HTML}{F18101}
+\definecolor{deepblue}{HTML}{CA3CFF}
 
@@
 
-\tikzstyle{uncertainty} = [rectangle, draw, fill=red!30, text=RichBlack, text width=6.5em, text centered]
-\tikzstyle{lever} = [rectangle, draw, fill=yellow!20, text=RichBlack, text width=6.5em, text centered]
-\tikzstyle{relation} = [rectangle, draw, fill=gray!20, text=RichBlack, text width=6.5em, text centered]
-\tikzstyle{metric} = [rectangle, draw, fill=blue!20, text=RichBlack, text width=6.5em, text centered]
+\tikzstyle{uncertainty} = [rectangle, draw, fill=rosso, text=grigio, text width=6.5em, text centered]
+\tikzstyle{lever} = [rectangle, draw, fill=giallo, text=nero, text width=6.5em, text centered]
+\tikzstyle{relation} = [rectangle, draw, fill=thirds, text=grigio, text width=6.5em, text centered]
+\tikzstyle{metric} = [rectangle, draw, fill=orange, text=nero, text width=6.5em, text centered]
 
-\tikzstyle{line} = [draw, very thick, color=RichBlack, ->]
+\tikzstyle{line} = [draw, very thick, color=nero, ->]
 \tikzset{>=latex}
@@
 
-    \node [relation] (relationships) at (0, 0) {System\\Relationships\\(R)};
-    \node [uncertainty] (surge) at (-3.5, 2.5) {Storm Surge\\{\footnotesize ($J$ draws from 1 posterior)}};
+    \node [font=\color{rosso}\selectfont] at (-4.5, 3.5) {\Large Uncertainties};
+    \node [relation] (relationships) at (0, -0.5) {Cost and damage models};
+    \node [uncertainty] (surge) at (-3.25, 2.5) {Storm surge\\{\footnotesize (draws from 1 posterior)}};
     \path [line] (surge) to (relationships);
-    \node [uncertainty] (slr) at (-3.5, -1) {Sea Level Rise\\{\footnotesize ($J$ Simulations)}};
+
+    \node [font=\color{thirds}\selectfont] at (0, -1.75) {\Large Relationships};
+    \node [uncertainty] (slr) at (-3.5, -1) {Sea level rise\\{\footnotesize (16 PDFs)}};
     \path [line] (slr) to (relationships);
-    \node [lever] (elevate) at (0, 2.5) {How high to elevate?};
+
+    \node [font=\color{giallo}\selectfont] at (0, 3) {\Large Levers};
+    \node [lever] (elevate) at (0, 2) {How high to elevate?};
     \path [line] (elevate) to (relationships);
-    \node [metric] (ccost) at (3.5, 1.25) {Construction Costs};
+
+    \node [font=\color{giallo}\selectfont] at (3.5, 2.25) {\Large Metrics};
+    \node [metric] (ccost) at (3.5, 1) {Up-front\\(construction)\\cost};
     \path [line] (relationships) to (ccost);
-    \node [metric] (fcost) at (3.5, -1.25) {Flood Damages};
+    \node [metric] (fcost) at (3.5, -1) {Lifetime expected damages};
     \path [line] (relationships) to (fcost);
 
-    \node[uncertainty] (rcp) at (-4.5, -2.5) {RCP Scenario\\{\footnotesize (4 Scenarios)}};
-    \node[uncertainty] (ice) at (-4.5, 0.75) {Ice Sheet Dynamics\\{\footnotesize (3 Models)}};
+    \node[uncertainty] (rcp) at (-4.5, -2.5) {RCP scenario\\{\footnotesize (4 Scenarios)}};
+    \node[uncertainty] (ice) at (-4.5, 0.75) {Ice sheet dynamics\\{\footnotesize (4 Models)}};
     \path [line] (rcp) to (slr);

### 2022-02-24 HouseElevation/src/surge-bayes.jl

@@ -98,7 +98,7 @@ function write_diagnostics(fit, fname::String)
         "rhat" => L"$\hat{R}$",
         "ess_per_sec" => L"\textrm{ESS per second}",
     )
-    tex_str = latexify(df; env=:table, fmt="%.2f")
+    tex_str = Latexify.latexify(df; env=:table, fmt="%.3f")
     open(fname, "w") do io
         write(io, tex_str)
     end

### 2022-02-24 scripts/03-cost.jl

@@ -14,6 +14,7 @@ function plot_depth_damage()
         legend=:topleft,
         xlabel="Flood Depth [ft relative to bottom of house]",
         ylabel="Damage [% House Value]",
+        yformatter=pct_formatter,
     )
     plot!(
         p,
@@ -51,6 +52,7 @@ function plot_cost_expected_damage()
         leftmargin=5mm,
         label="HAZUS (Used)",
         color=colors[1],
+        yformatter=pct_formatter,
     )
     plot!(
         p,

### 2022-02-24 scripts/main.jl

@@ -5,11 +5,10 @@ using ColorSchemes
 using Unitful
 
 # HouseElevation.clear_cache() # to clean all the processed data and plots
-#colors = ColorSchemes.seaborn_bright6 # colorblind friendly and consistent scheme
-colors = ColorSchemes.okabe_ito
+include("plotutils.jl")
+colors = get_colormap()
 
 # these files just provide functions that will make plots -- nothing happens except defining functions
-include("plotutils.jl")
 include("01-surge-modeling.jl")
 include("02-mean-sea-level.jl")
 include("03-cost.jl")
@@ -22,10 +21,10 @@ function main()
     # define some constants
     house_floor_area = 1500u"ft^2"
     house_value_usd = 200_000.0 # HOUSE NOT LAND VALUE
-    discount_rate = 0.03 # mortgages going for 3-5%
+    discount_rate = 0.03 # mortgages going for 3-4.5% at the moment
     syear = 2022
     eyear = 2092
-    # for a **VERY** vague idea of prices see
+    # for a **VERY** rough idea of prices see
     # https://www.zillow.com/homedetails/9638-Selby-Pl-Norfolk-VA-23503/79223088_zpid/
 
     # get the storm surge posterior samples

### 2022-02-24 scripts/plotutils.jl

@@ -1,3 +1,4 @@
+using Colors
 using Distributions
 using Plots
 
@@ -32,3 +33,13 @@ twiny(plt::Plots.Plot=current()) = twiny(plt[1])
 
 pct_formatter(yi) = "$(Int.(round.(yi * 100)))%"
 blank_formatter(yi) = ""
+
+function get_colormap()
+    giallo = colorant"#F0BC42"
+    rosso = colorant"#8E1F2F"
+    grigio = colorant"#CACACC"
+    nero = colorant"#000000"
+    thirds = colorant"#31467A"
+    orange = colorant"#F18101"
+    return [rosso, giallo, thirds, grigio, orange, nero]
+end

### 2022-02-26 papers/paper/preprint.tex
some edits by JDG, going to send to Klaus for feedback

@@
 % fonts and appearance
-\usepackage{amsmath, amsfonts, physics, siunitx}
+\usepackage{amsmath, amsfonts, physics, siunitx, nicefrac}
 \usepackage[american]{babel} 
@@
 \usepackage{geometry}
-\geometry{letterpaper,top=1in, bottom=1in, left=0.5in, right=2in}
+\geometry{letterpaper,top=1in, bottom=1in, left=1in, right=2in}
 
@@
 \newacronym{iid}{IID}{independent and identically distributed}
+\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
 \newacronym{lsl}{LSL}{local mean sea level}
@@
 \newacronym{rcp}{RCP}{representative concentration pathway}
-\newacronym{slr}{SLR}{relative sea level rise}
+\newacronym{msl}{MSL}{mean relative sea level}
 \newacronym{rdm}{RDM}{robust decision making}
@@
 \begin{abstract}
-    The policy and engineering tools used to inform climate adaptation etc rely on standards and cost-benefit analyses that depend on risk probabilities.
-    One example is house elevation: the recommended elevation is a standard (BFE plus a foot) and funding decisions rely on CBA.
-    Under stationarity, the probabilities for this can almost be agreed upon.
-    Under nonstationarity, there are deep uncertainties.
-    We show that the way these deep uncertainties are synthesized into CBA matters a lot.
-    Thus, static decisions are subject to deep uncertainties
+    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
+    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
+    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
+    However, growing recognition of nonstationarity motivates a forward-looking approach.
+    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
+    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
+    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-sea level rise scenarios.
+    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future sea level rise to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
+    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future sea level rise.
+    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
+    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
 \end{abstract}
 
+Key points
+\begin{enumerate}
+    \item
+\end{enumerate}
+
 \clearpage
-
 \section{Introduction}\label{sec:introduction}
 
-Since infrastructure built today is likely to persist for several decades, infrastructure designs must take into account the stresses that a project is likely to face decades into the future.
-On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.\james{Too fast on intro.
-Need to slow down and define concepts}
-On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
-For example, \ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
-
-TODO: set up flood risk management as a context where over- or under-design could be relevant, then raise house elevation in the coastal zone as a specific example.
-Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
-Be sure to note that this has parallels to lots of other engineering design guidance.
-Then note the work of \citet{zarekarizi_suboptimal:2020} or \citet{xian_elevation:2017}, who show that a risk-based approach may improve considerably on a standards-based approach.
-
-Additionally, the floodplain maps used for decision supportare silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
-Both risk-based design and standards-based approaches are complicated by nonstationarity.
-On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
-On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[\eg,][]{keller_management:2021}.
-
-This motivates subjective approaches for designing for nonstationarity.
-Thus decision-makers need some subjective way to choose which possible futures to design for.
-
-In this paper, we use a didactic case study of house elevation in Norfolk, VA as a lens through which to shed light on the following research questions.
+Estimates of flood hazard affect peoples' lives
+\begin{enumerate}
+    \item Floods affect lots of people and property
+    \item Climate change and exposure making floods worse
+    \item Floodplain management: $T$ year flood
+    \item Some cost-benefit analysis and lots of design standards / heuristics
+\end{enumerate}
+One example is floodproofing
+\begin{enumerate}
+    \item Literature on floodproofing and building-scale vulnerability reduction measures
+    \item This works great!
+    \item But who should use it? We're still using standards.
+    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
+          Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
+          For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
+          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
+          Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
+\end{enumerate}
+Nonstationarity
+\begin{enumerate}
+    \item These analyses all rely on \glspl{pdf}
+    \item Growing evidence for nonstationarity in a wide variety of hydroclimate processes
+    \item We often focus on climate change as a driver but sometimes local modifications matter just as much \citep{Merz:2014gf}
+    \item Even where risk-based analyses are used, this problem recommendations
+    \item Yet stationarity remains the dominant assumption because it is immortal \citep{Montanari:2014hl} or undead \citep{Serinaldi:2015bq}
+    \item Bonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{msl} is clear.
+    \item On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification
+\end{enumerate}
+It matters how you prepare
+\begin{enumerate}
+    \item Equifinality problem: different models diverge even if they fit the historical record well
+    \item Over- and under-design can each be costly \citep{DossGollin:2019}
+    \item On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
+          For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
+    \item On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
+For example,\ldots\ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
+\end{enumerate}
+A quick philosophical literature review of how people have approached and interpreted this problem
+\begin{enumerate}
+    \item This motivates subjective approaches for designing for nonstationarity.
+    \item There's a little bit of RDM-style work on floods \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
+    \item Model choice in $\mathcal{M}$-open case
+    \item Philosophy of Bayesian statistics and decision making -- explain subjective probability
+\end{enumerate}
+Case study
+\begin{enumerate}
+    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment could be a problem
+    \item Federal guidance uses both standards (floodplain, \gls{bfe}) and cost-benefit analysis (does it pass test?)
+    \item Be sure to note that this has parallels to lots of other engineering design guidance.
+    \item \citet{xian_elevation:2017}: one size does not fit all
+    \item \citet{zarekarizi_suboptimal:2020}: neglecting uncertainty leads to bad outcomes
+\end{enumerate}
+Plan d'attaque
+\begin{enumerate}
+    \item Research questions
           \begin{description}
@@
           \end{description}
-Specifically, we examine this problem through three lenses.\james{Which ones?}
-
-The paper is organized as follows.\james{This is confusing to the reader}
-In \cref{sec:case-study} we describe our case study in Norfolk, VA.
-First, in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
-Next, in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
-Then in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
-Finally in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.
+    \item In \cref{sec:case-study} we describe our case study in Norfolk, VA.
+    \item  in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
+    \item  in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
+    \item  in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
+    \item  in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.
+\end{enumerate}
 
@@
 
-We model a \emph{hypothetical} house in Norfolk, VA, where sea level rise drives nonstationary future flood hazard.
-For interpretability, we focus on deep uncertainty in \gls{slr} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
-We consider a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
-We evaluate this decision for each of many possible simulations of \gls{slr} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
-In the remainder of this section we discuss data sources and the system model.
+We model a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
+For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
+We evaluate this decision for each of many possible simulations of \gls{msl} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
+In the remainder of this section we discuss data sources and models.
 
@@
         \midrule
-        \Gls{slr}        & $\overline{y}_t$ & Deeply uncertain: multiple methods considered                                        & \cref{sec:sea-level}         \\
-        Storm surge      & $y'_t$           & Probabilistic: a single stationary \acrshort{gev} model                              & \cref{sec:storm-surge}       \\
+        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
+        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
+        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
         Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
-        Depth-damage     & $D(h-y_t)$       & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
+        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
         Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
@@
 
-Deeply uncertain sea level rise is the focus of our model.
-Our analysis considers the common case where a decision maker or analyst has access to an existing set of simulations.
-Specifically, we use simulations published in \citet{ruckart_hazardous:2008} that use the BRICK model version 0.2 \citep{wong_brick0.2:2017}.
+Deeply uncertain sea level rise is the dominant source of deep uncertainty in our analysis.
+Decision makers and analysts often have access to simulations of physical variables like \gls{msl}, but usefully incorporating these simulations into decision analyses remains an area of active research.
 
-\Cref{fig:lsl-evolution} shows the \glspl{pdf} of mean sea level in 2100 under each of the eight models considered: four emissions pathways \citep[\gls{rcp} scenarios][]{moss_rcp:2010} and two models of ice sheet dynamics.
-Variation between different \glspl{pdf} is dominated by the \gls{rcp} 8.5 scenario with fast ice sheet dynamics \citep[see][for a discussion of these dynamics]{wong_brick0.2:2017}.
-This is also the \gls{pdf} with the greatest within-model variance.
+To inform the house elevation decision, we develop an ensemble of simulations of future \gls{msl}.
+Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic models using data published in \citet{ruckert_coastal:2019}.
+The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
+The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
+These four models paramaterize physical processes, particularly of ice sheet dynamics, in different ways.
+For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, or \citet{bamber_slrise:2019}.
+
+Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
+To represent this uncertainty, we sample four \gls{rcp} scenarios.
+This gives us sixteen time-varying probability density functions of \gls{msl}.
+We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \gls{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.
+
 \begin{figure}
@@
         Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
-        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for three probabilistic models.
-        (B): probability distribution of mean sea level at Sewells Point, VA in the year 2100 for all probabilistic models considered.
+        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
+        (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
     }\label{fig:lsl-evolution}
@@
 
-For consistency with previous studies, we refer to one time series of \gls{slr} ($\overline{y}_t$) as \emph{a \gls{sow}}.
+\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
+The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
+\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
+Both physical model and \gls{rcp} scenario strongly inform the modeled probability distribution.\klaus{Do you want to add a sentence here comparing BRICK to the Kopp et al models?}
 
@@
 
-Neglecting hydrodynamics, we model annual maximum floods $y_t$ as the sum of sea level $\overline{y}_t$ and storm surges $y'_t$.
+Coastal flood hazard does not depend only on \gls{msl} but also on storm surges.
+Neglecting hydrodynamics, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
+In this subsection we describe our representation of storm surge.
 
-Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).\james{cite NOAA data source}
+Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
+The NOAA tides and currents dataset is available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
-For each year we remove the annual mean then extract a single maximum; we refer the set of these maxima for all years as the time series of annual maximum storm surges.
-This time series is shown in \cref{fig:surge-obs-return}(a).
-In (b) we use the Weibull plotting position
-\begin{equation}\label{eq:plot-pos}
-    \frac{r}{N + 1}
-\end{equation}
-where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
+For each year we first remove the annual mean, then extract the  maximum water level.
+We refer these maxima as the annual maximum storm surges $y'(t)$.
+
+\Cref{fig:surge-obs-return}(a) shows the time series of historical annual maximum storm surges.
+Major historical surges have been driven by two different mechanisms: tropical cyclones and nor'easters.
+The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and nor'easters have caused surges above \SI{6}{ft}.
 
@@
         (A):
-        Time series of historic storms.
-        Blue (green) arrows denote notable tropical cyclones (nor'easters).
+        time series of historic storms.
+        Red (yellow) arrows denote notable tropical cyclones (nor'easters).
         (B):
-        Return periods.
-        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\cref{eq:plot-pos}).
-        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit.
+        return periods.
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
+        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
     }\label{fig:surge-obs-return}
@@
 
-We model future storm surge risk through a probabilistic lens.
-To do this, we develop a stationary\james{discuss / validate this -- it's cool, we have the supplemental info and can emphpasize our focus here is on SLR for didactic purposes} Bayesian model for storm surge:
+We model future storm surge using a stationary \gls{gev} model:
 \begin{equation}\label{eq:surge-model}
-    y'_t \sim \text{GEV}\left(\mu, \sigma, \xi \right),
+    y'(t) \sim \text{GEV}\left(\mu, \sigma, \xi \right),
 \end{equation}
-where $y'_t$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively.
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively, of the \gls{gev} distribution.
+This model is stationary because there is no time dependence; surges in different years are modeled as \gls{iid}.
 
-When \gls{gev} distributions are fit to time series data without strong prior information, the resulting uncertainties can be unrealistically large.\james{cite}
-Thus, information from other locations, time periods, and/or from physical principles may help to reduce uncertainties \citep{Merz:2008bl,Merz:2008eh}.
-To constrain estimates, we add weakly informative priors.
-However, rather than adding prior information over the parameters themselves, we follow \citet{coles_evd:1996} and apply a prior over extreme quantiles of the distribution.
-Specifically, we apply LogNormal priors over the 90th, 99th, and 99.9th percentiles of the distribution (\ie, the 10, 100, and 1000 year return period events).\james{Expand on this ``probabilistic inversion''}
-These values were chosen to represent plausible physical ranges (the assumed 90\% confidence interval for the 10 year event is XY years, for the 100 year event is XZ years, and for the 1000 year event ZY years) and are plotted in supporting figure SX.\james{return to this sentence}
+To improve estimates, we add twp forms of prior information.
+First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$: when $\xi < 0$, $y' \in \qty(-\infty, \mu - \nicefrac{\sigma}{\xi})$  and when $\xi > 0$, $y' \in \qty(\mu-\nicefrac{\sigma}{\xi}, \infty)$.
+Since storm surges cannot be negative, only the latter is physically defensible.
+Second, we add weakly informative priors.
+Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition is difficult, we apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
+Specifically we apply half Normal (\ie, Normal distributions truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
 
-To evaluate a single \gls{sow}, we integrate over $J$ samples from the posterior distribution (of \cref{eq:surge-model}).
+For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} through the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
+Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
+We evaluate the model's fit using posterior predictive checks \citep[see section 2.4 of][and references therein]{gelman_workflow:2020}.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly \cref{fig:surge-test-statistics}.
+Since \cref{fig:surge-test-statistics}(b) suggests the possibility of meaningful temporal structure not captured by our stationary \gls{iid} model, future efforts  could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020}.
+
+\Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
+The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position
+\begin{equation}\label{eq:plot-pos}
+    \frac{r}{N + 1}
+\end{equation}
+where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
+The good fit of the Bayesian fit (gray shading) to the data (points) suggests that the stationary approximation is reasonable for this dataset.
 
@@
 The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
-Specifically, expected annual damages in year $t$ is defined as the expected damages in a given year, conditional on the house's height relative to the gauge after elevation ($h = h_0 + \Delta h$):
+Specifically, we define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
+This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
+The expcted annual damage is thus
 \begin{equation}\label{eq:ead}
-    \textrm{EAD}_t = \mathbb{E}[d_t | h] = \int_{y_t} p(y_t) D(h - y_t) dy_t,
+    \textrm{EAD} = \mathbb{E}[D(h-\overline{y})] = \int_{y'} p(y') D(h - (\overline{y} + y')) \dd{y'}
 \end{equation}
-where $y_t = \overline{y}_t + y'_t$ is the annual maximum flood, relative to the gauge, in year $t$, and $D(h - y_t)$ is the damage as a function of depth relative to the house.
-Following \citet{zarekarizi_suboptimal:2020} we use the HAZUS deterministic depth-damage relationship to parameterize $D(h-y_t)$ (see fig. XYZ(c)).\james{Cite HAZUS}
+where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
+The term $\overline{y}$ in \cref{eq:ead} is time-dependent but we have omitted the time dependence for readability.
+Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
+(HAZUS) depth-damage curves provided by FEMA to parameterize $D(h-\overline{y})$; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+For comparison, \cref{fig:cost-depth-damage} also shows results using the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
+Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for concision.
 
-The expected annual damages is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
+The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
 However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
@@
               \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
-              \item simulate a single storm surge from the stationary distribution (equation XYZ) and add the mean sea level to get total flood depth
-              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_t$) into  the deterministic HAZUS depth-damage relationship
-              \item store this as the $k$th damage
+              \item simulate a single storm surge from the stationary distribution (\cref{eq:surge-model}) and add the mean sea level to get total flood depth $y^\mathrm{sim}_k$
+              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $k$th damage.
           \end{enumerate}
@@
 \end{enumerate}
-
-However, evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
+Evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
 Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
-However, noticing that this function dependsonly on the elevation of the house relative to \gls{slr}, we develop a simuple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
-To do this, we  precompute expected annual damage for all height differences in \SI{0.1}{ft} increments from -XY to YX ft and fit a piecewise linear interpolation to this data.\james{check increements, piecewise linear}
-We use $K=\num{1e6}$ samples to fit this emulator for each of the NUMBER increments.
-This model is shown in supplemental figure SX.\james{supplemental figure}
+However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
+To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
+We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
+This model is shown in \cref{fig:cost-expected-damage-emulator}.
 Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
@@
 where the discount rate is $1 - \gamma$.
-For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate of 2\%, thus $\gamma = 0.98$.
-Neglecting uncertainty in house lifetime, we model flood damages from 2022 to 2071, thus $T=50$.\james{This has not been the case thus far}
+For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate (see \cref{tab:uncertainties}).
+Neglecting uncertainty in house lifetime, we model flood damages over a fixed house lifetime (\cref{tab:uncertainties}).
 We calculate lifetime expected damages \emph{for each \gls{sow} separately}.
-
-\subsection{Assessing performance}\label{sec:sow-metrics}
 
@@
 \begin{description}
-    \item[up-front cost] reflects the cost of elevating a house.
+    \item[up-front cost] is the cost of elevating a house.
 Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
-We normalize this cost by house value.\james{Normalize!}
-    \item[lifetime expected damages] is discussed in \cref{sec:led}
-    \item[lifetime expected costs] is the sum of lifetime expected damages and up-front costs
-    \item[the expected benefit-to-cost ratio] is defined as the ratio of construction cost to damages averted, relative to not elevating at all:
-        \begin{equation}
-            \mathrm{BCR}(\Delta h) = \begin{cases} \Delta h = 0 & 1 \\ \Delta h \neq 0 & \frac{\mathbb{E}[d_t | \Delta h = 0] - \mathbb{E}[d_t | \Delta h = \Delta h]}{C(\Delta h)}\end{cases}
-        \end{equation}
+We normalize this cost by house value.
+This cost curve is shown in \cref{fig:cost-up-front}.
+    \item[lifetime expected damages] is discussed in \cref{eq:led}
+    \item[expected lifetime costs] is the sum of lifetime expected damages and up-front costs
 \end{description}
-In \cref{sec:multiple-simulation} we introduce the notation $u$ for these metrics.
 
-\subsection{Decision analytical framework}
+\subsection{Experiment design}
 
-In \cref{fig:flowchart} we introduce a general notation for model-based decision analysis which we develop incrementally.
-We consider $J$ \glspl{sow} representing time series of mean sea level and $I$ discrete decisions representing the choice for how high to elevate the house $\Delta h$.
-For each combination of these, we calculate the vector of metrics described in \cref{sec:sow-metrics}, which we denote $u_{ij} = f(x_i, s_j)$, using a system model $f$ (box c).
-This system model represents storm surge, building costs, and damages to output the metrics $u$ (which in this case is a vector).
-In \cref{sec:multiple-simulation} we consider ``the multiple \gls{pdf} problem'' by introducing boxes (d) and (e).
-Finally in \cref{sec:synthesizing} we introduce box (f) as a tool for synthesizing uncertainty across the multiple \glspl{pdf}.\klaus{Any feedback on the is figure and/or section?}
+In \cref{fig:flowchart} we introduce a general notation for simulation-based decision analysis.
+We consider $J$ \glspl{sow} representing time series of \gls{msl} and $I$ discrete decisions $x_1, \ldots, x_I$ representing the choice for how high to elevate the house.
+For each combination of these, we calculate the vector of metrics described in \cref{sec:led}, which we denote $u_{ij} = f(x_i, s_j)$, using a system model $f$.
+This system model caculates the metrics $u$ (which in this case is a vector) using the equations and relationships described in the preceding subsections.
+We begin in \cref{sec:multiple-simulation} by using the system model to explore the drivers of desirable and undesirable outcomes (blue box in \cref{fig:flowchart}).
+Next in \cref{sec:multiple-simulation} we consider scenario-conditional optimization and introduce ``the multiple \gls{pdf} problem'' (red box in \cref{fig:flowchart}).
+Finally in \cref{sec:synthesizing} we introduce a metnod for synthesizing uncertainty across the multiple \glspl{pdf} (orange box in \cref{fig:flowchart}).
 
@@
 
-The literature on decision making under deep uncertainty \citep[see][]{marchau:2019} offers some insight as to how to identify decisions that perform well over many plausible \glspl{sow}.
-We first consider exploratory modeling, which sheds light on the questions: ``how well does a given decision perform under different possible futures, and to which possible futures are different decisions vulnerable?''\james{Maybe this doesn't need to be a questino}
+We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions \citep[see][]{bankes:1993}.
+Exploratory modeling is used to explore the consequences of different possibilities, and does not require any assessment of how likely a given \gls{sow} may be.
 
-\Cref{fig:scenario-map} shows each possible future as a single point for each of several values of $\Delta h$.\james{Add after updating this figure}
-On the $y$ axis is expected lifetime costs (see \cref{sec:sow-metrics}).
-On the $x$ axis is mean sea level in 2100, which is an interpretable 1-dimensional projection of the \glspl{sow}.
-Although our model for expected discounted total costs is deterministic, points with similar $x$-values may have different $y$-values because their sea levels before 2100 can differ (\ie, the 1-dimensional projection of the \glspl{sow} does not represent all variance between them).
+\begin{figure}
+    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
+    \caption{
+        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of $\Delta h$ and $h_0$.
+        Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
+        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
+        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
+        In all cases, elevating the house reduces the variance in total lifetime cost.
+        Values are sensitive to model constants; see \cref{tab:uncertainties}.
+    }\label{fig:scenario-map-slr-cost}
+\end{figure}
 
-The exploratory modeling approach shown in \cref{fig:scenario-map} shows that shows that the best outcome is when $\Delta h = 0$ and when \gls{slr} is negligible, but the worst outcome is also when $\Delta h = 0$ and when \gls{slr} is substantial.
-As $\Delta h$ increases, the construction cost increases but the variability between different \glspl{sow} decreases, up to a point.
-\emph{No decision dominates} any other, including $\Delta h = 0$, in the formal sense of being optimal over all \glspl{sow}.
-\james{Revisit this paragraph once changes are made to \cref{fig:scenario-map}.}
+\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs) as a function of height increase ($\Delta h$) and sea level rise over the house lifetime.
+The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and sea level rise is minimal (bottom left corners).
+The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
+As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if sea level rise is substantial, damages will be negligible.
+Decisions about if and, if so, how high to elevate may thus depend on the decision-maker's risk aversion.
+Additionally, since exploratory modeling does not shed light on the likelihood of different \glspl{sow}, decisions about if and, if so, how high to elevate will also depend on an assessment of sea level rise likelihoods.
 
@@
 
-Whereas the analysis in \cref{sec:multiple-simulation} shed light on how decisions perform under different \glspl{sow}, the metrics calculated represent the performance of a given decision-\gls{sow} combination, and \emph{do not shed light on the performance of the decisions themselves}.
-
-An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of eight \glspl{pdf} of sea level rise, as shown in \cref{fig:lsl-evolution}.
+The exploreatory analysis of the previous subsection considered each realization of future sea level to be an exploratory sample from the space of possible futures.
+An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of the sixteen models of sea level rise, as shown in \cref{fig:lsl-evolution}.
 This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.
@@
 A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
-Specifically, for each of $K$ models (box d), which here are our eight models for \gls{slr},\james{need to go through whole document and be consistent about MSL, SLS, SRL, RSLR, \etc} the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
+Specifically, for each of $K=16$ models, the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
 Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
@@
 
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected lifetime cost\james{update figure} as a function of $\Delta h$\james{Add $\Delta h$ to the plot} for four of the eight models available.
-Similarly, \cref{fig:tradeoffs-by-rcp} plots the 80\% confidence interval width (\ie, the 90th percentile minus the 10th percentile) of lifetime cost.
-This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp} 8.6 with fast ice sheet dynamics).
-Consistent with \cref{fig:scenario-map}, the variability of lifetime costs increases as $\Delta h$ increases.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the total lifetime cost as a function of $\Delta h$ the sixteen models considered (three are highlighted).
+Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages.
+This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
+Consistent with \cref{fig:scenario-map-slr-cost}, the variability of lifetime costs decreases as $\Delta h$ increases.
 Once $\Delta h$ reaches \SIrange[]{3}{7}{ft}, depending on the model considered, construction costs start to dominate flood losses, and thus higher values of $\Delta h$ increase average lifetime costs.
 
-Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} \emph{requires synthesizing across \glspl{sow}}.
+Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} requires synthesizing across \glspl{sow}.
 However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
@@
 
-Second, this approach presents what we call ``the multiple \gls{pdf} problem'' because it leaves decision makers with many probability distributions to choose from
-Cite \citet{sharma_rcp:2021} as another example.
-Add a sentence or two explaining why this is a problem for decision makers\ldots\james{Need to bring in a bit of discussion from \citet{schneider_scenarios:2002} and from engineering docs (ASCE, FEMA, \ldots).}\klaus{Would appreciate your help putting together a succinct summary of the multiple \gls{pdf} problem}
+Second, this approach presents what we call ``the multiple \gls{pdf} problem'' because it leaves decision makers with many probability distributions from which to choose.
+The multiple \gls{pdf} problem has been shown in other contexts.
+For example, \citet{sharma_rcp:2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
+Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
+Further, once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.
 
@@
 
-Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.\james{This also needs to go in the abstract}
+\citet{schneider_scenarios:2002} wrote, in the context of \gls{ipcc} assessments, that ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
+Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.
 
-Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{empirical}(\Omega)$, does not reflect our belief about their true distribution, $p_\mathrm{true}(\Omega)$.
-Often this is by design: exploratory modeling emphasizes the value of sampling low-probability futures as a way to learn about the system and develop insight \citep{bankes:1993}.\james{Tighten up this writing}
-This is analagous to the deliberate over- or under-sampling of certain groups in survey design, and so following standard practice in survey design we develop a re-weighting approach to correct for the over- or under-sampling of different regions of the survey space.
+\subsection{Re-weighting}
 
-We apply a grid-based discretization scheme to calculate these weights.
-First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional space, which we denote $\psi_1, \ldots, \psi_J$.
-As before, we take this space to be the mean sea level in 2100.
+Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{sampling}(\Omega)$, does not reflect our subjective belief about their true distribution, $p_\mathrm{belief}(\Omega)$.
+Indeed, this is often by design: it is often advantageous to sample low-probability but high-consequence \glspl{sow} to learn about a system's behavior \citep{bankes:1993}.
+To address this, we apply a scheme to weight each \gls{sow} based on $p_\mathrm{belief}(\Omega)$.
+First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional representation, which we denote $\psi_1, \ldots, \psi_J$.
+As discussed in \citet{csillery_abc:2010} and other research on approximate Bayesian computation, the choice of summary statistic is an important one.
+We take the summary statistic to be sea level rise from 2022 to 2100, recognizing that future work could consider multidimensional summaries.
+
 Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-This allows us to simplify $p_\mathrm{true}(s_j) = p_\mathrm{true}(\psi_j)$.\klaus{This notation seems bad, suggestions?}
-Defining $F_\mathrm{true}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{true}$, \ie
-$$F_\mathrm{true}(s) = \int_{-\infty}^{s'} p_\mathrm{true}(s) \dd{s},$$
+This allows us to define $p_\mathrm{belief}(\psi \in \Psi)$ instead of $p_\mathrm{belief}(s \in \Omega)$.
+Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, \ie
+$$
+    F_\mathrm{belief}(\psi) = \int_{-\infty}^{s'} p_\mathrm{belief}(\psi) \dd{\psi},
+$$
 we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:grid-sketch}.
@@
     w_j = \begin{cases}
-        j = 1     & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
-        1 < j < J & F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
-        j = J     & 1 - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
+        j = 1     & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
+        1 < j < J & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
+        j = J     & 1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
     \end{cases}
 \end{equation}
-This grid-based weighting scheme can be used with higher-dimensional projections of the scenarios, as is done for example in the joint probability method used for hurricanes \citep[see, \eg,][]{johnson_clara:2013,toro_jpm-os:2010}.
 
-The primary difficulty with \cref{eq:weights} is that the true probability distribution $p_\mathrm{true}(\phi)$ is deeply uncertain.
-Because it is unknown, it must be approximated with a model.
-And because it is deeply uncertain, multiple models\james{multiple priors?} for $p_(\Omega)$ should be used.
-But whereas the $M_k$ defined in \cref{sec:multiple-pdf} represented the raw input data (four \gls{rcp} scenarios times two models of ice sheet dynamics), we define the $M_k$ here to be subjective models for mean sea level in 2100.\james{Are these models or priors?}
-
-Need a few sentences here on prior uses of expert judgement and subjective belief.
+TODO\james{This paragraph needs some work}
 The key point to get across is what a subjective belief means in the Bayesian context \citep{savage:1954,gelman_philosophy:2013,gelman_subjectiveobjective:2017}.
@@
 
-Need to explain how we came up with our two or three subjective priors.\klaus{Would love suggestions here!}
+Need to explain how we came up with our two or three subjective priors.\klaus{Are there published assessments we should refer to}
 Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
@@
 \Cref{fig:tradeoffs-by-prior} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
+
+\subsection{Implications}
 
@@
 
-\section{Conclusions}\label{sec:conclusions}
+\section{Discussion and conclusions}\label{sec:conclusions}
 
-To summarize:
-\begin{itemize}
-    \item Natural extensions to Bayesian updating
-    \item Belief $p(s)$ can't be ``right''  \citep{gelman_workflow:2020,gelman_philosophy:2013} or ``neutral'' \citep{quinn_exploratory:2020}
-\end{itemize}
-
-This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
-
-Future research needs:
-\begin{itemize}
-    \item More complex models to capture more relevant metrics
-    \item Better models of nonstationary hazards
-    \item Interacting, sequential decisions
-\end{itemize}
-
-\section{Bonus (need to sort through)}
-
-This section is a work in progress -- it includes some text that used to be in the paper as well as some bullet points that may be useful for the discussion section.\klaus{Can you highlight any picese here that you find especially salient?}
-
-Including additional models \citep[\eg,][]{kopp_evolving:2017,deconto_antarctica:2016,ruckert_coastal:2019} \emph{would compound this challenge}.\james{Lee et al: Lee, Haran, Pollard, Keller}
-
-Most fundamentally, we are thinking about robustness a bit differently than usual.
-The typical way that the DMDU literature thinks about it is robustness to a \gls{sow}.\james{This is a simplification}
-We are thinking about robustness to different modeling structures and assumptions.
-This has links to the ``rival framings'' literature.
-Given a subjective prior $p(s)$, I can compute expectations of the decision metrics $m$; if there are a lot of decisions, I can start to sketch out the trade-off curves as shown in figures XXXX (in \cref{sec:results}).
-
-Since we have priors, there are clear links to how we would update our belief (and thus decision) once more information arrives.
-
+Quick summary of what we have done
 \begin{enumerate}
-    \item Links to DMDU literature
+    \item OK
+\end{enumerate}
+Quick summary of what our motivation was
+\begin{enumerate}
+    \item Explore every \gls{sow}
+    \item Be transparent about how we are calculating trade-offs
+\end{enumerate}
+Philosophical links to Bayesian philosophy and model building in the $\mathcal{M}$-open case (NB: this is copied and pasted from my thesis -- need to rework)
+\begin{enumerate}
+    \item  \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
+          \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
+    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
+    \item More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
+    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.\james{This should go to intro}
+    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.\james{Explain this to the audience}
+    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
+          More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
+    \item Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}
+    \item Our approach has interesting theoretical links to Bayesian model averaging or stacking \citep{Yao:2018bu}
+\end{enumerate}
+Links to DMDU literature
 \begin{enumerate}
@@
     \item We are in effect calculating robustness metrics \citep{mcphail_robustness:2019,herman:2015}
+    \item \citet{schneider_scenarios:2002} notes that a RDM analysis ``still requires the construction of ranges of future outcomes -- itself a subjective assessment.
+Subjectivity is simply inherent in all future projections, whether in natural or social sciences.''
 \end{enumerate}
-    \item Philosophy of Bayesian stats / decision making
-          \begin{enumerate}
-              \item The observation that many of these mechanisms cannot be represented by a single objective \gls{pdf} has motivated many criticisms of the application of \gls{bdt} to planning problems.\james{Cite some (from thesis)}
-                    Yet \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
-                    \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
-                    The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
-                    More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
-                    Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.\james{This should go to intro}
-                    Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.\james{Explain this to the audience}
-                    Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
-                    More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
-                    Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}\footnote{this paragraph is copied from my dissertation, needs to be re-worked!}
-          \end{enumerate}
-    \item There are interesting links to Bayesian model averaging
-          \begin{enumerate}
-              \item Lots of examples to cite if we want
-              \item I'm partial to something like stacking \citep{Yao:2018bu}
-              \item The main difference is we are using only the prior to average the models!
-          \end{enumerate}
-    \item Limitations of the case study
+
+\subsection{Limitations and future work}
+
+Limitations of the case study
 \begin{enumerate}
@@
 \end{enumerate}
-    \item Limitations of the method
+Limitations of the method
 \begin{enumerate}
-              \item Our prior $p(s)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
+    \item Our prior $p(\Omega)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
     \item We neglect true ignorance \citep{knight_risk:1921}.\james{Perhaps ``unknown unknowns''}
@@
 \end{enumerate}
-    \item The benefits of this approach:
+Future research needs:
 \begin{enumerate}
-              \item A goal is to make imperfect and subjective modeling choices as transparent as possible.
-Since we can't be right, we should make it as easy as possible for others to understand and critique our assumptions\james{Focus on this!}
-              \item Because of how we weight realizations of the future, you can apply a different weighting function and immediately assess performance.
-Thus qualitative and quantitative sensitivity analyses are simple.
-              \item Of course the choice of $p(\mathcal{Z})$ is subjective, but so is every other aspect of how we frame the problem.
-Modeling is intrinsically subjective.
-The idea of embracing subjective choices -- and making them explicit and open to critique -- draws heavily from philosophies of iterative workflow in statistics \citep{box:1976,gelman_workflow:2020,gelman_philosophy:2013}.
-              \item This approach can be used in a multiobjective context and is an alternative way to measure robustness in \gls{rdm} and MORDM
+    \item This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
+    \item More complex models to capture more relevant metrics
+    \item Better models of nonstationary hazards
+    \item Interacting, sequential decisions
 \end{enumerate}
-    \item This approach can be used in many other contexts
+
+\subsection{Implications}
+
+This approach can be used in many other contexts
 \begin{enumerate}
@@
 \end{enumerate}
-\end{enumerate}.
-
-Federal guidance for this decision relies on standards (the $T$-year flood) rather than risk-based design.
-Be sure to note that this has parallels to lots of other engineering design guidance.
-Then note the work of \citet{zarekarizi_suboptimal:2020} and \citet{xian_elevation:2017} who show that a risk-based approach may improve significantly on a standards-based approach.\james{Need to addd some citations of ``risk-based design''}
-
-Additionally, the floodplain maps used for decision support are silent on projected future changes and neglect the deep and dynamic uncertainties surrounding the projected hazards\ldots
-Both risk-based design and standards-based approaches are complicated by nonstationarity.
-On the one hand, nonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{slr} is clear.
-On the other, efforts to find objective methods for quantifying nonstationarity, \eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification \citep[see][]{keller_management:2021}.\james{This goes earlier -- see Klaus's added text}
-
-This suggests that either nonstationarity should be ignored, or subjective approaches for designing for nonstationarity should be adapted.
-Given that physical theories consistent with the observational record and models robustly project changes to climate hazard under warning,
-
-Many methods for decision making under deep uncertainty first evaluate candidate decisions over many plausible states of the world, then aggregate performance over these possible futures.
-Through a didactic case study of determining how high to elevate a single home in Norfolk, VA, we demonstrate that the common practice of weighting all scenarios equally in this aggregation creates a tension between (a) fully exploring the parameter space, including unlikely regions, and (b) accurately representing available scientific knowledge.
-We introduce an approach to bridge this divide through a computationally efficient method that weights each \gls{sow} based on a probability distribution over possible futures.
-Since the distribution of deeply uncertain variables is necessarily subjective, we turn to frameworks for model critique from applied Bayesian statistics to compare and contrast different modeling assumptions.
-This approach can help to improve decision-making by facilitating iterative and collaborative model improvement.
-
-While classical utility- or regret-based frameworks offer a structure for weighing these risks, they are not designed to handle uncertainties that are \emph{deep} in the sense that stakeholders may not agree upon their probabilities \citep{lempert_complex:2002} or \emph{irreducible} in the sense that there is not an objective truth that better models and data can identify \citep{DossGollin:2019}.\footnote{Unless we believe everything is pre-determined, in which case why bother with climate adaptation\ldots}
-Examples of such uncertainties include how much greenhouse gasses humans will emit in the future or whether the U.S.
-federal government will continue to subsidize flood insurance for homeowners.\james{Add some references}
-In response, a wide range of fields have developed scenario-based methods\ldots.\klaus{Which arguments in \citet{savage:1954} did you think should be cited here?}
-This leads to the ``Multiple PDF Problem'' \citep{sharma_rcp:2021}\ldots.
-A key question, then, facing the designers, builders, regulators, owners, and users of built facilities is \emph{to which scenario should infrastructure be designed?}
-
-One example is building codes and guidelines\ldots.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
-Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
-\citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
-Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
-
-Nonstationarity is particularly relevant in coastal communities where \gls{slr} is expected to drive future flood risk.
-As a didactic case study, we consider the problem of how high a homeowner should elevate their structure to mitigate future flood losses, illustrated in \cref{fig:house-sketch},\james{Need to revisit this picture} given stochastic flood peaks and deeply uncertain \gls{slr}.
-\Gls{slr} is driven mainly by physical processes like (\ldots, West Antarctic Ice Sheet, \ldots\klaus{You wrote something here that I can't read -- I see something that looks like ICM.
-I assume Tony and Kelsey papers are relevant as well?}) \citep{kopp_probabilistic:2014,kopp_evolving:2017}.
-Two key uncertainties identified by previous studies are
-\begin{enumerate*}[label=(\roman*)]
-    \item which emissions pathway (\eg, \gls{rcp} scenario) will we experience and
-    \item how will polar ice sheets respond to warming \citep{wong_brick0.2:2017,ruckert_coastal:2019,wong_nola:2017,deconto_antarctica:2016}.
-\end{enumerate*}
-Simulations of \gls{slr} are available to decision makers grouped by \gls{rcp} scenario and model structure.
-For a given combination of \gls{rcp} scenario and model, results are typically presented and interpreted as probabilistic, but this is not the same as giving decision-makers a single \gls{pdf}.
-This is the ``Multiple \gls{pdf} Problem'' described in, \eg, \citet{sharma_rcp:2021}.
-Thus, \emph{which scenario/\gls{pdf} should we design for?}
-
-Lacking a crystal ball capable of telling us precisely which \gls{pdf} we need to design for, we have to consider several as plausible.
-Any decision will perform differently in different scenarios.
-There is no objective answer to this question.
-Since we can't be ojectively correct, we might as well be transparent and make our assumptions explicit (and subject to critique) rather than implicit.
-
-Households elevate their homes to manage flood risks, but regulations and guidance are silent on key questions \citep{zarekarizi_suboptimal:2020,xian_elevation:2017}.
-\begin{description}
-    \item[Q1] How to adapt guidance to building characteristics or household preferences?
-    \item[Q2] How does nonstationary hazard change guidance?
-    \item[Q3] Which model of nonstationary hazard to use?
-\end{description}
-These challenges also affect other adaptation plans and engineering designs (\cref{fig:stormwater-bridge}).
 
@@
 JDG wrote the manuscript with support from KK.
-JDG wrote the manuscript in consulation and collaboration with KK.
+JDG wrote the manuscript in consulation with KK.
+JDG and KK revised the manuscript.
 
@@
 \section{Supplemental figures}
+
 \begin{figure}
@@
 
+\begin{figure}
+    \includegraphics[width=\textwidth]{scenario-map-height-slr}
+    \caption{
+        Expected total lifetime cost (damages plus up-front cost) as a function of sea level rise over the house lifetime and height increase $\Delta h$.
+        Initial house elevation is fixed to \SI{1}{ft} below the \gls{bfe}.
+    }\label{fig:scenario-map-height-slr}
+\end{figure}
+
 \section{Supplemental tables}

### 2022-02-26 tikz/xlrm.tex
some edits by JDG, going to send to Klaus for feedback

@@
 \definecolor{orange}{HTML}{F18101}
-\definecolor{deepblue}{HTML}{CA3CFF}
 
@@
     \node [font=\color{rosso}\selectfont] at (-4.5, 3.5) {\Large Uncertainties};
-    \node [relation] (relationships) at (0, -0.5) {Cost and damage models};
+    \node [relation] (relationships) at (0, -0.5) {Engineering and economic models};
     \node [uncertainty] (surge) at (-3.25, 2.5) {Storm surge\\{\footnotesize (draws from 1 posterior)}};
@@
     \node [font=\color{thirds}\selectfont] at (0, -1.75) {\Large Relationships};
-    \node [uncertainty] (slr) at (-3.5, -1) {Sea level rise\\{\footnotesize (16 PDFs)}};
+    \node [uncertainty] (slr) at (-3.5, -1) {Sea level\\{\footnotesize (16 PDFs)}};
     \path [line] (slr) to (relationships);

### 2022-02-26 HouseElevation/src/cost.jl

@@ -50,7 +50,7 @@ end
 
 """Fit the expected damage emulator"""
 function fit_expected_damage_emulator(key::Symbol=:hazus; N::Int=1_000_000)
-    clearances = collect(-30:0.5:30)u"ft" # house above MSL
+    clearances = collect(-30:0.25:30)u"ft" # house above MSL
     surge_fits = get_surge_posterior()
     surges = sample_predictive_GEV(surge_fits, N)u"ft"
 

### 2022-02-26 scripts/01-surge-modeling.jl

@@ -37,7 +37,6 @@ function plot_annmax_floods(annual::HouseElevation.AnnualGageRecord)
         (name="Chesapeake-Potomac", year=1933, is_tc=true, Δx=7.5, Δy=0.75),
         (name="Outer Banks hurricane", year=1936, is_tc=true, Δx=10, Δy=0.75),
         (name="Isabel", year=2003, is_tc=true, Δx=-7.5, Δy=0.5),
-        (name="Irene", year=2015, is_tc=true, Δx=2.5, Δy=1),
         (name="Nor'Ida", year=2009, is_tc=false, Δx=0, Δy=1),
         (name="Ash Wednesday", year=1962, is_tc=false, Δx=10, Δy=0.75),
     ]
@@ -65,7 +64,7 @@ function plot_annmax_floods(annual::HouseElevation.AnnualGageRecord)
         x1 = x_text + 0.9 * (storm.year - x_text)
         y0 = y_text + 0.2 * (yobs - y_text)
         y1 = y_text + 0.9 * (yobs - y_text)
-        color = storm.is_tc ? colors[2] : colors[3]
+        color = storm.is_tc ? colors[1] : colors[2]
         plot!([x0, x1], [y0, y1]; color=color, linewidth=1.5, label="")
         annotate!(p, x_text, y_text, text(storm.name, :center, 7; color=color))
     end
@@ -328,7 +327,7 @@ function plot_surge_obs_return(
         link=:y,
         ylims=(2, 9),
         layout=grid(1, 2; widths=[0.6, 0.35]),
-        size=[1000, 350] .* 1.35,
+        size=[1000, 350] .* 1.15,
         bottommargin=8mm,
         leftmargin=8mm,
     )

### 2022-02-26 scripts/05-tradeoffs-by-rcp.jl

@@ -89,7 +89,7 @@ function plot_rcp_tradeoffs(
         push!(p_archive, p)
     end
     add_panel_letters!(p_archive; fontsize=12, loc=(0.1, 0.95))
-    p = plot(p_archive...; layout=(1, 2), link=:x, size=(1200, 600))
+    p = plot(p_archive...; layout=(1, 2), link=:x, size=(1100, 550))
 
     savefig(plots_dir("tradeoffs-by-rcp.pdf"))
     return p

### 2022-02-26 scripts/06-combining.jl

@@ -81,7 +81,7 @@ function plot_priors()
         xlabel="Sea Level Rise, 2022-2100, at Sewells Point, VA [ft]",
         ylabel="Probability Density",
     )
-    for (prior, color) in zip(priors, colors)
+    for (prior, color) in zip(priors, [:red, :blue, :green])
         plot!(p, prior.dist, 0, 12.5; label=prior.name, color=color, linewidth=3)
     end
     p

### 2022-02-26 scripts/plotutils.jl

@@ -34,7 +34,7 @@ twiny(plt::Plots.Plot=current()) = twiny(plt[1])
 pct_formatter(yi) = "$(Int.(round.(yi * 100)))%"
 blank_formatter(yi) = ""
 
-function get_colormap()
+function get_roma_colormap()
     giallo = colorant"#F0BC42"
     rosso = colorant"#8E1F2F"
     grigio = colorant"#CACACC"
@@ -43,3 +43,5 @@ function get_colormap()
     orange = colorant"#F18101"
     return [rosso, giallo, thirds, grigio, orange, nero]
 end
+
+get_colormap() = get_roma_colormap()

### 2022-03-08 HouseElevation/src/surge-bayes.jl
Better Bayesian inference

@@ -1,34 +1,37 @@
 using Distributions
+using DynamicHMC
 using DynamicPPL
+using HDF5
 using Latexify
 using LaTeXStrings
 using MCMCChains
+using MCMCChainsStorage
 using Optim
 using StatsBase
 using Turing
 
 # stationary GEV model
 gev_priors = [
-    (rt=2, dist=Distributions.TruncatedNormal(4, 1.5, 0, Inf)),
-    (rt=10, dist=Distributions.TruncatedNormal(6, 1.75, 0, Inf)),
-    (rt=100, dist=Distributions.TruncatedNormal(10, 2.25, 0, Inf)),
-    (rt=500, dist=Distributions.TruncatedNormal(15, 2.75, 0, Inf)),
+    (rt=2, dist=Distributions.TruncatedNormal(6, 3, 0, Inf)),
+    (rt=10, dist=Distributions.TruncatedNormal(8, 4, 0, Inf)),
+    (rt=100, dist=Distributions.TruncatedNormal(12, 6, 0, Inf)),
+    (rt=500, dist=Distributions.TruncatedNormal(16, 8, 0, Inf)),
 ]
 
 @model function StationaryGEV(y)
 
     # have to put a prior to define parameters in Turing
-    ξ ~ Distributions.TruncatedNormal(0, 0.5, 0, Inf) # surge has lower bound => ξ>0
-    μ ~ Distributions.TruncatedNormal(0, 10, 0, Inf) # lower bound should not be negative
-    σ ~ Distributions.TruncatedNormal(0, 4, 0, Inf) # σ > 0 by definition
+    μ ~ Distributions.TruncatedNormal(0, 25, 0, Inf) # lower bound should not be negative
+    σ ~ Distributions.TruncatedNormal(0, 10, 0.01, Inf) # σ > 0 by definition
+    ξ ~ Distributions.TruncatedNormal(0, 2.5, 0, Inf) # surge has lower bound => ξ>0
 
-    dist = Distributions.GeneralizedExtremeValue(μ, σ, ξ)
+    dist = GeneralizedExtremeValue(μ, σ, ξ)
 
     # implement the prior on quantile levels
     for prior in gev_priors
         xp = 1.0 - 1.0 / prior.rt
         rl = quantile(dist, xp)
-        #Turing.@addlogprob!(logpdf(prior.dist, rl))
+        Turing.@addlogprob!(loglikelihood(prior.dist, rl))
     end
 
     # data model
[…]

### 2022-03-08 scripts/01-surge-modeling.jl

@@ -101,6 +101,7 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
     # get fake data
     dist = Distributions.GeneralizedExtremeValue(4, 0.5, 0.15)
     fake_data = make_fake_data(dist, N)
+    rts = 1 ./ (1 .- cdf.(dist, fake_data))
 
     # fit the model
     posterior = HouseElevation.get_posterior(
@@ -111,56 +112,13 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
         GeneralizedExtremeValue(row[:μ], row[:σ], row[:ξ]) for row in eachrow(posterior_df)
     ]
 
-    # set up the return period plot
-    rts = range(1.25, 300; length=250) # return periods
-    aeps = 1 .- 1 ./ rts # annual exceedance probability
-    xticks = [2, 5, 10, 25, 50, 100, 250]
+    # plot the quantiles in gray
+    p = plot_return_period(post_gevs)
 
-    ub1 = [quantile([quantile(d, xi) for d in post_gevs], 0.95) for xi in aeps]
-    lb1 = [quantile([quantile(d, xi) for d in post_gevs], 0.05) for xi in aeps]
-    ub2 = [quantile([quantile(d, xi) for d in post_gevs], 0.9) for xi in aeps]
-    lb2 = [quantile([quantile(d, xi) for d in post_gevs], 0.1) for xi in aeps]
-    ub3 = [quantile([quantile(d, xi) for d in post_gevs], 0.75) for xi in aeps]
-    lb3 = [quantile([quantile(d, xi) for d in post_gevs], 0.25) for xi in aeps]
+    # add true line
+    scatter!(p, rts, fake_data; label="Obs (Known Return Period)", color=colors[1], alpha=1)
 
-    p = plot(;
-        xlabel="Return Period [years]",
-        ylabel="Return Level [ft]",
-        xscale=:log,
-        legend=:topleft,
-        xticks=(xticks, string.(xticks)),
-        dpi=250, # for saving
-        size=(500, 500),
-        title="Fake Data Experiment: N=$N",
-    )
-    plot!(
-        rts,
-        ub1;
-        fillbetween=lb1,
-        fillcolor=:gray,
-        fillalpha=0.35,
-        linecolor=false,
-        label="95% Posterior CI",
-    )
-    plot!(
-        rts,
-        ub2;
[…]

### 2022-03-08 scripts/main.jl

@@ -42,6 +42,7 @@ function main()
         plot_surge_posterior_teststats(annual, fits)
         plot_surge_posterior_return(annual, fits)
         plot_surge_obs_return(annual, fits)
+        plot_surge_prior_return()
     end
 
     # make some plots of the sea level (BRICK) data
@@ -80,7 +81,7 @@ function main()
     let
         s = HouseElevation.get_lsl(; syear=syear, eyear=eyear)
         bfe = calc_bfe(fits, s, syear)
-        x = collect(0:0.25:10)u"ft"
+        x = collect(0:0.25:12)u"ft"
         elevation_init = bfe - 1u"ft"
         u = get_outcomes(
             x;

### 2022-03-08 scripts/plotutils.jl

@@ -45,3 +45,53 @@ function get_roma_colormap()
 end
 
 get_colormap() = get_roma_colormap()
+
+function plot_return_period(gevs::Vector{<:Distributions.GeneralizedExtremeValue}) where {T<:HouseElevation.MCMCChains.Chains}
+    
+        rts = range(1.25, 275; length=250) # return periods
+        aeps = 1 .- 1 ./ rts # annual exceedance probability
+        xticks = [2, 5, 10, 25, 50, 100, 250]
+    
+        ub1 = [quantile([quantile(d, xi) for d in gevs], 0.95) for xi in aeps]
+        lb1 = [quantile([quantile(d, xi) for d in gevs], 0.05) for xi in aeps]
+        ub2 = [quantile([quantile(d, xi) for d in gevs], 0.9) for xi in aeps]
+        lb2 = [quantile([quantile(d, xi) for d in gevs], 0.1) for xi in aeps]
+        ub3 = [quantile([quantile(d, xi) for d in gevs], 0.75) for xi in aeps]
+        lb3 = [quantile([quantile(d, xi) for d in gevs], 0.25) for xi in aeps]
+    
+        p = plot(;
+            xlabel="Return Period [years]",
+            ylabel="Return Level [ft]",
+            xscale=:log,
+            legend=:bottomright,
+            xticks=(xticks, string.(xticks)),
+        )
+        plot!(
+            rts,
+            ub1;
+            fillbetween=lb1,
+            fillcolor=:gray,
+            fillalpha=0.35,
+            linecolor=false,
+            label="95% Posterior CI",
+        )
+        plot!(
+            rts,
+            ub2;
+            fillbetween=lb2,
+            fillcolor=:gray,
+            fillalpha=0.35,
+            linecolor=false,
+            label="80% Posterior CI",
+        )
+        plot!(
+            rts,
+            ub3;
+            fillbetween=lb3,
+            fillcolor=:gray,
+            fillalpha=0.35,
+            linecolor=false,
+            label="50% Posterior CI",
+        )
+        return p
+    end

### 2022-03-16 papers/paper/preprint.tex
Better Bayesian inference
edits esp base SLR priors on NOAA report

@@
-\documentclass[12pt]{article}
+\documentclass[11pt]{article}
 
@@
 \usepackage{authblk}
-\title{Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties}
+\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
 \author[1]{James Doss-Gollin}
@@
 \usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
+\newacronym{abc}{ABC}{approximate Bayesian computation}
 \newacronym{bdt}{BDT}{Bayesian decision theory}
@@
 \newacronym{ipcc}{IPCC}{International Panel on Climate Change}
-\newacronym{lsl}{LSL}{local mean sea level}
 \newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
+\newacronym{msl}{MSL}{mean relative sea level}
+\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
 \newacronym{pdf}{PDF}{probability density function}
 \newacronym{rcp}{RCP}{representative concentration pathway}
-\newacronym{msl}{MSL}{mean relative sea level}
 \newacronym{rdm}{RDM}{robust decision making}
+\newacronym{slr}{SLR}{sea level rise}
+\newacronym{ssp}{SSP}{shared socio-economic pathway}
 \newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
@@
     In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
-    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-sea level rise scenarios.
-    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future sea level rise to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
-    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future sea level rise.
+    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-\gls{slr} scenarios.
+    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future \gls{slr} to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
+    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future \gls{slr}.
     This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
@@
 
-Estimates of flood hazard affect peoples' lives
+Flood risk management relies on probabilistic estimates of flood hazard:
 \begin{enumerate}
-    \item Floods affect lots of people and property
-    \item Climate change and exposure making floods worse
-    \item Floodplain management: $T$ year flood
-    \item Some cost-benefit analysis and lots of design standards / heuristics
+    \item In many countries including the United States, flood risk management involves many different public and private sector actors but most rely on probabilistic estimates of flood hazard.
+    \item Local governments can build infrastructure to reduce flood risk; they often prioritize these by considering a reduction in damages from a design storm (\eg, the 100 year storm) \citep{hcfcd_prioritization:2019}
+    \item There is a role for industry organizations that set standards that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
+    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by FEMA.
+At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).
+\citep{kousky_voucher:2014}
+    \item Beyond flood insurance and disaster recovery there has been a federal retreat from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}
 \end{enumerate}
-One example is floodproofing
+Future flood hazard is deeply uncertain
 \begin{enumerate}
-    \item Literature on floodproofing and building-scale vulnerability reduction measures
-    \item This works great!
-    \item But who should use it? We're still using standards.
-    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
-          Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-          For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
-          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
-          Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
+    \item ``Along the U.S.
+coastline, public infrastructure and \$1 trillion in national wealth held in coastal real estate are threatened by rising sea levels, higher storm surges, and the ongoing increase in high tide flooding'' \citep{reidmiller_reportinbrief:2018}
+    \item Nonstationarity also affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
+    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}
+    \item For coastal flooding, \gls{fema} guidance suggests using up to date estimates of \gls{msl} \citep{fema_slr:2016}, but no projections, and stationary estimates of storm surge \citep{fema_ffa:2016}, although the analyst at each location has considerable flexibility
+    \item Ditto extreme rainfall \citep{atlas14_texas:2018} or floods \citep{bulletin17c:2019}
+    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}
 \end{enumerate}
-Nonstationarity
+Methods for decision making under deep uncertainty emphasize exploring many possible futures
 \begin{enumerate}
-    \item These analyses all rely on \glspl{pdf}
-    \item Growing evidence for nonstationarity in a wide variety of hydroclimate processes
-    \item We often focus on climate change as a driver but sometimes local modifications matter just as much \citep{Merz:2014gf}
-    \item Even where risk-based analyses are used, this problem recommendations
-    \item Yet stationarity remains the dominant assumption because it is immortal \citep{Montanari:2014hl} or undead \citep{Serinaldi:2015bq}
-    \item Bonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{msl} is clear.
-    \item On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification
+    \item deep uncertainty describes ``known unknowns'' where there is not a single objectively true model \citep{walker_deep:2013,Walker:2013gi,lempert_complex:2002}
+    \item This is related to equifinality where different models fit the historical record equally well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
+    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that
+    \item \citet{bankes:1993} draws a distinction between exploratory and consolidative modeling
+    \item Decision scaling \citep{Steinschneider:2015kk}
+    \item global sensitivity analysis \citep{saltelli_sensitivity:2010,herman_salib:2017,sobol_sensitivity:2001} use feed large scenario ensembles into a black box model, then analyze outputs to assess the relative importance of each input (conditional on an assumed joint distribution over inputs)
+    \item MORDM \citep{kasprzyk:2013} combines exploratory modeling (exploring performance over many scenarios) and policy search (using optimization tools)  \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+Typically ``robustness metric'' is used \citep{herman:2015,mcphail_robustness:2019}
+    \item For a more complete review see \citet{marchau:2019}
+    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
+    \item Crucially, these approaches all emphasize exploring performance over many scenarios, then describing this performance into some set of metrics.
 \end{enumerate}
-It matters how you prepare
+Some critiques
 \begin{enumerate}
-    \item Equifinality problem: different models diverge even if they fit the historical record well
-    \item Over- and under-design can each be costly \citep{DossGollin:2019}
-    \item On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
-          For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
-    \item On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
-For example,\ldots\ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
+    \item There are always subjective  modeling decisions, but sometimes they are opaque which makes it hard to communicate them to the end user
+    \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}
+    \item Technical detail: joint probability distributions over high dimensional spaces are confusing.
+Bayesian analysis teaches us that posterior distributions of parameters are often highly correlated.
+If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data!
+    \item Some scenarios are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019} yet this is generally not acknowledged
 \end{enumerate}
-A quick philosophical literature review of how people have approached and interpreted this problem
+The question of whether probability is a useful tool for decision making under deep uncertainty depends on how one interprets probability; we argue that a subjective Bayesian interpretation of probability is useful for many DMDU problems
 \begin{enumerate}
-    \item This motivates subjective approaches for designing for nonstationarity.
-    \item There's a little bit of RDM-style work on floods \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
-    \item Model choice in $\mathcal{M}$-open case
-    \item Philosophy of Bayesian statistics and decision making -- explain subjective probability
+    \item We are inspired by methods and philosophy of Bayesian model choice in the ``$\mathcal{M}$-open'' case where there is no true model to identify
+    \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
+This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critiqued
+    \item \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
+\citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
+    \item We know that choosing a single ``best'' model under-estimates uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}
+    \item In the absence of data that would tell us how accurate each model is via some likelihood function, we have to use the prior.
+    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
+    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
+    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
+    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
+More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
+    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}
 \end{enumerate}
-Case study
+Our aim is to illustrate how methods for model selection in the $\mathcal{M}$-open case can improve decision making under deep uncertainty.
+\begin{enumerate}
+    \item We want to combine multiple models in a way that is transparent, and that can be changed with low computational cost (cheap robustness checks).
+    \item We are particularly motivated by problems for which simulations of some key uncertainties (particularly climate-driven ones but also population, economic growth, etc) already exist.
+We assume only that there is no coupling between the system model and these simulations, so they can be treated as fully exogenous.
+\end{enumerate}
+We illustrate our approach with a didactic case study\ldots.
 \begin{enumerate}
@@
     \item \citet{zarekarizi_suboptimal:2020}: neglecting uncertainty leads to bad outcomes
+    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
+    \item Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
+    \item For example, \citet{xian_elevation:2017} uses a cost-benefit analysis to demonstrate that tailoring recommendations to the initial elevation of a structure can reduce expected costs.
+          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
+    \item Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
+    \item FEMA's Hazard Mitigation Grant Program, authorized under Section 404 of the 1988 Stafford Act, which awards grants after a disaster.
+Under the Stafford Act, projects funded by HMGP, including buyouts, must be shown to be cost-effective; typcally, up-front costs are compared to avoided losses \citep{bendor_buyouts:2020}
 \end{enumerate}
-Plan d'attaque
+Our primary research question is ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
+We proceed as follows.
 \begin{enumerate}
-    \item Research questions
-          \begin{description}
-              \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
-              \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
-              \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
-          \end{description}
     \item In \cref{sec:case-study} we describe our case study in Norfolk, VA.
-    \item  in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
+    \item  in \cref{sec:exploratory} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
     \item  in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
@@
 
-We model a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
-For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
-We evaluate this decision for each of many possible simulations of \gls{msl} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
-In the remainder of this section we discuss data sources and models.
+We model one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
+For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
 
@@
     \caption{
-        Summary of parameters, their notation, and how their uncertainty is modeled.
+        Summary of parameters, their notation, and how their uncertainty is represented.
     }\label{tab:uncertainties}
@@
         House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
+        First year       & $t_i$             & Deterministic: 2022                                                                  &                              \\
+        Last year        & $t_f$             & Deterministic: 2091 unless otherwise noted                                           &                              \\
         \bottomrule
@@
     \centering
-    \includegraphics[width=4in]{xlrm.pdf}
+    \includegraphics[width=\textwidth]{xlrm.pdf}
     \caption{
@@
         A \gls{sow} consists of a description of the uncertain factors (red).
-        We model a single decision (yellow).
-        For each combination of \gls{sow} and policy, the system model (blue) is used to calculate metrics (orange) describing the performance of the policy over the selected \gls{sow}.
+        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
+        For each \acrshort{sow} and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (orange).
     }\label{fig:xlrm}
@@
 
-\subsection{Sea level rise}\label{sec:sea-level}
+\subsection{\gls{slr}}\label{sec:sea-level}
 
-Deeply uncertain sea level rise is the dominant source of deep uncertainty in our analysis.
-Decision makers and analysts often have access to simulations of physical variables like \gls{msl}, but usefully incorporating these simulations into decision analyses remains an area of active research.
-
-To inform the house elevation decision, we develop an ensemble of simulations of future \gls{msl}.
-Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic models using data published in \citet{ruckert_coastal:2019}.
+To inform the house elevation decision, we develop an ensemble \gls{msl} time series.
+Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
 The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
 The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
-These four models paramaterize physical processes, particularly of ice sheet dynamics, in different ways.
-For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, or \citet{bamber_slrise:2019}.
+These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
+For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, and \citet{bamber_slrise:2019}.
 
 Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
-To represent this uncertainty, we sample four \gls{rcp} scenarios.
-This gives us sixteen time-varying probability density functions of \gls{msl}.
-We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \gls{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.
+To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probability density functions of \gls{msl}.
+We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \acrfull{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.
 
@@
     \caption{
-        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
-        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
+        Projections of future mean sea level depend strongly on the choices of physical model and forcing.
+        (A): 90\% confidence intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
         (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
@@
 
+The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
 \Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
@@
 \Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-Both physical model and \gls{rcp} scenario strongly inform the modeled probability distribution.\klaus{Do you want to add a sentence here comparing BRICK to the Kopp et al models?}
+We return in \cref{sec:multiple-pdf} to the problem that these multiple models poses to decision makers.
 
@@
 
-Coastal flood hazard does not depend only on \gls{msl} but also on storm surges.
-Neglecting hydrodynamics, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
-In this subsection we describe our representation of storm surge.
+We model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
+In this subsection we describe the representation of storm surge in our model.
 
 Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
-The NOAA tides and currents dataset is available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
-Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
-For each year we first remove the annual mean, then extract the  maximum water level.
+The NOAA tides and currents dataset is freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
+For each calendar year we first remove the annual mean, then extract the maximum water level.
 We refer these maxima as the annual maximum storm surges $y'(t)$.
 
-\Cref{fig:surge-obs-return}(a) shows the time series of historical annual maximum storm surges.
-Major historical surges have been driven by two different mechanisms: tropical cyclones and nor'easters.
-The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and nor'easters have caused surges above \SI{6}{ft}.
+In the historical record, major surges have been driven by two different mechanisms: tropical cyclones and Nor'easters.
+This time series of annual maxima is shown in \cref{fig:surge-obs-return}(a).
+The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.
 
@@
         time series of historic storms.
-        Red (yellow) arrows denote notable tropical cyclones (nor'easters).
+        Red (yellow) arrows denote notable tropical cyclones (Nor'easters).
         (B):
@@
 \begin{equation}\label{eq:surge-model}
-    y'(t) \sim \text{GEV}\left(\mu, \sigma, \xi \right),
+    y'(t) \sim \text{GEV}\qty(\mu, \sigma, \xi),
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively, of the \gls{gev} distribution.
-This model is stationary because there is no time dependence; surges in different years are modeled as \gls{iid}.
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location $\mu$, scale $\sigma$, and shape $\xi$ has the probability density function
+\begin{equation*}
+    f(x | \mu, \sigma, \xi)= \begin{cases}
+        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
+        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
+    \end{cases}
+\end{equation*}
+This model assumes stationarity, neglecting any potential time dependence.
 
-To improve estimates, we add twp forms of prior information.
-First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$: when $\xi < 0$, $y' \in \qty(-\infty, \mu - \nicefrac{\sigma}{\xi})$  and when $\xi > 0$, $y' \in \qty(\mu-\nicefrac{\sigma}{\xi}, \infty)$.
+Our approach to model building follows a principled workflow for model building and checking \citep[see][for details]{gelman_workflow:2020}.
+One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
+We include two forms of prior information.
+First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$:
+\begin{equation*}
+    \mathrm{supp}~ y' =
+    \begin{cases}
+        \xi < 0: & y' \in \qty(-\infty,~ \mu - \nicefrac{\sigma}{\xi}) \\
+        \xi > 0: & y' \in \qty(\mu-\nicefrac{\sigma}{\xi},~ \infty).
+    \end{cases}
+\end{equation*}
 Since storm surges cannot be negative, only the latter is physically defensible.
+This is related to the approach of \citet{martins_estimators:2000}, who use physical reasoning to place an informative prior $\xi$.
 Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition is difficult, we apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
-Specifically we apply half Normal (\ie, Normal distributions truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition may be difficult, we instead apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
+Specifically we apply Normal (truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
 These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
 
-For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} through the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
+For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
 Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
-We evaluate the model's fit using posterior predictive checks \citep[see section 2.4 of][and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly \cref{fig:surge-test-statistics}.
-Since \cref{fig:surge-test-statistics}(b) suggests the possibility of meaningful temporal structure not captured by our stationary \gls{iid} model, future efforts  could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020}.
+We evaluate the model's fit using posterior predictive checks \citep[section 2.4 and references therein]{gelman_workflow:2020}.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly (\cref{fig:surge-test-statistics}) although \cref{fig:surge-test-statistics}(b) suggests the possibility of temporal structure not captured by our stationary \gls{iid} model.
+Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017}.
 
@@
 where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
-The good fit of the Bayesian fit (gray shading) to the data (points) suggests that the stationary approximation is reasonable for this dataset.
+The good fit of the Bayesian fit (gray shading) to the data (points) suggest the modeling choices made are reasonable and sufficient for this didactic example.
+Finally, a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
 
-\subsection{Expected annual damage}\label{sec:ead}
+\subsection{Damages and metrics}\label{sec:ead-led}
 
-There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
-The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
-Specifically, we define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
+The system model (``relationships'' in \cref{fig:xlrm}) is comprised of two key pieces.
+The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
+The second converts model converts a time series of annual expected damages into lifetime expected damages.
+
+We define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
 This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
-The expcted annual damage is thus
+The expected annual damage is thus
 \begin{equation}\label{eq:ead}
-    \textrm{EAD} = \mathbb{E}[D(h-\overline{y})] = \int_{y'} p(y') D(h - (\overline{y} + y')) \dd{y'}
+    \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'}
 \end{equation}
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-The term $\overline{y}$ in \cref{eq:ead} is time-dependent but we have omitted the time dependence for readability.
 Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
@@
 For comparison, \cref{fig:cost-depth-damage} also shows results using the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
-Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for concision.
+Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
 
 The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
-However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
-\begin{enumerate}
-    \item For $k=1, \ldots, K$:
-          \begin{enumerate}
-              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
-              \item simulate a single storm surge from the stationary distribution (\cref{eq:surge-model}) and add the mean sea level to get total flood depth $y^\mathrm{sim}_k$
-              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $k$th damage.
-          \end{enumerate}
-    \item Estimate expected annual damages as the sample mean of the $K$ estimates
-\end{enumerate}
-Evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
-Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
-However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
-To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
-We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
-This model is shown in \cref{fig:cost-expected-damage-emulator}.
-Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
+However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
+Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible values $h - \overline{y}(t)$ and use this to train a computationally efficient a surrogate model (see \cref{sec:surrogate-ead}).
 
-\subsection{Lifetime expected damages}\label{sec:led}
-
-The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the up-front cost plus the discounted sum of expected annual damages
+The second component of the system model converts a time series of $\mathrm{EAD}$ into lifetime expected damages, which we define as the up-front  discounted sum of expected annual damages:
 \begin{equation}\label{eq:led}
-    \textrm{LED} = \sum_{t=1}^T \gamma^{t-1} \mathbb{E}[d_t | \Delta h],
+    \textrm{LED} = \sum_{t=t_i}^{t_f} \gamma^{\qty(t-t_i)-1} \textrm{EAD}(t),
 \end{equation}
@@
 For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate (see \cref{tab:uncertainties}).
-Neglecting uncertainty in house lifetime, we model flood damages over a fixed house lifetime (\cref{tab:uncertainties}).
 We calculate lifetime expected damages \emph{for each \gls{sow} separately}.
 
-To assess the performance of a given decision for a specific \gls{sow}, we calculate the following metrics \emph{for each decision-\gls{sow} combination}.
-\begin{description}
-    \item[up-front cost] is the cost of elevating a house.
+To assess the performance of a given decision for a specific \gls{sow} (``Metrics'' in \cref{fig:xlrm}), we calculate the following metrics for each decision-\gls{sow} combination:
+\begin{enumerate}
+    \item ``Up-front cost'' is the cost of elevating a house.
 Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
@@
 This cost curve is shown in \cref{fig:cost-up-front}.
-    \item[lifetime expected damages] is discussed in \cref{eq:led}
-    \item[expected lifetime costs] is the sum of lifetime expected damages and up-front costs
-\end{description}
+    \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
+    \item ``Expected lifetime costs'' is the sum of lifetime expected damages and up-front costs.\james{Make sure the figures all use this term}
+\end{enumerate}
 
@@
 
-In \cref{fig:flowchart} we introduce a general notation for simulation-based decision analysis.
-We consider $J$ \glspl{sow} representing time series of \gls{msl} and $I$ discrete decisions $x_1, \ldots, x_I$ representing the choice for how high to elevate the house.
-For each combination of these, we calculate the vector of metrics described in \cref{sec:led}, which we denote $u_{ij} = f(x_i, s_j)$, using a system model $f$.
-This system model caculates the metrics $u$ (which in this case is a vector) using the equations and relationships described in the preceding subsections.
-We begin in \cref{sec:multiple-simulation} by using the system model to explore the drivers of desirable and undesirable outcomes (blue box in \cref{fig:flowchart}).
-Next in \cref{sec:multiple-simulation} we consider scenario-conditional optimization and introduce ``the multiple \gls{pdf} problem'' (red box in \cref{fig:flowchart}).
-Finally in \cref{sec:synthesizing} we introduce a metnod for synthesizing uncertainty across the multiple \glspl{pdf} (orange box in \cref{fig:flowchart}).
+In the remainder of this article we incrementally develop a general framework, illustrated in \cref{fig:flowchart}, for analyzing climate risk management strategies under deep uncertainty.
+
+Our framework considers $J$ \glspl{sow} $s_1, \ldots, s_J$ (here representing time series of \gls{msl}) and $I$ discrete decisions $x_1, \ldots, x_I$ (here representing the choice for how high to elevate the house).
+For each combination of these, we calculate the vector of metrics described in \cref{sec:ead-led}, which we denote $u$.
+Letting $f$ be a system model (here: the economic and engineering loss models in \cref{sec:ead-led}) we define $u_{ij} := f(x_i, s_j)$.
+
+We begin in \cref{sec:exploratory} with an exploratory analysis (blue box in \cref{fig:flowchart}) that uses the system model to identify the conditions under which particularly desirable or undesirable outcomes occur.
+This exploratory modeling does not require any assessment of the relative likelihood of the \glspl{sow}, but can only be interpreted as a ``what-if'' analysis.
+We continue in \cref{sec:exploratory} with a scenario-conditional analysis (red box in \cref{fig:flowchart}) that integrates outcomes over all the \glspl{sow} corresponding to a particular model (here \gls{rcp} scenario and physical model structure).
+While this analysis can be used to quantify the conditional distribution of outcomes, this conditional distribution will be different for each model considered.
+We refer to this as ``the multiple \gls{pdf} problem.''
+Finally in \cref{sec:synthesizing} we introduce a method for synthesizing uncertainty across the multiple \glspl{pdf} (orange box in \cref{fig:flowchart}) by weighting each \gls{sow}, given some subjective belief.
 
@@
     \centering
-    \includegraphics[width=4in]{bayes-rdm.pdf}
+    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
     \caption{
-        Outline of the proposed decision theoretic framework.
-        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
-        Next in \cref{sec:multiple-pdf} we use scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem'' (region 2).
-        Finally in \cref{sec:synthesizing} propose a method for weighting \glspl{sow} to synthesize across multiple \glspl{pdf} (region 3).
+        Outline of the proposed decision analytic framework.
+        We begin in \cref{sec:exploratory} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
+        Next we use scenario-conditional optimization to illustrate the ``multiple \acrshort{pdf} problem'' (region 2).
+        Finally we propose a method for weighting \glspl{sow} to synthesize uncertainties across multiple \glspl{pdf} (region 3).
     }\label{fig:flowchart}
@@
 
-\section{Exploratory analysis}\label{sec:multiple-simulation}
+\section{Exploratory analysis}\label{sec:exploratory}
 
-We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions \citep[see][]{bankes:1993}.
-Exploratory modeling is used to explore the consequences of different possibilities, and does not require any assessment of how likely a given \gls{sow} may be.
+We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions.
+Exploratory modeling has been widely used in water resources management \citep{Brown:2012kb,Poff:2015jn,quinn_exploratory:2020,Steinschneider:2015kk}, analysis of coupled human-natural systems \citep{moallemi_exploratory:2020,moallemi_decisionsupport:2020}, and coastal adaptation \citep{sriver_sealevel:2018} and does not require any \emph{a priori} assessment of how likely a given \gls{sow} may be \citep{bankes:1993,kwakkel_exploratory:2013,reed_multisectordynamics:2022}.
 
@@
     \caption{
-        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of $\Delta h$ and $h_0$.
+        Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of $\Delta h$ and $h_0$.
         Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
-        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
-        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
+        The lowest-cost outcomes occur when both exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
+        The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
         In all cases, elevating the house reduces the variance in total lifetime cost.
@@
 
-\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs) as a function of height increase ($\Delta h$) and sea level rise over the house lifetime.
-The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and sea level rise is minimal (bottom left corners).
-The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
-As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if sea level rise is substantial, damages will be negligible.
-Decisions about if and, if so, how high to elevate may thus depend on the decision-maker's risk aversion.
-Additionally, since exploratory modeling does not shed light on the likelihood of different \glspl{sow}, decisions about if and, if so, how high to elevate will also depend on an assessment of sea level rise likelihoods.
+\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs) as a function of height increase ($\Delta h$) and \gls{slr} over the house lifetime.
+The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
+The worst outcomes (blue) arise when the house is elevated only slightly and \gls{slr} is rapid.
+As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
 
-\section{Scenario-conditional probabilistic analysis: the multiple PDF problem}\label{sec:multiple-pdf}
+This exploratory analysis can be used to quantify the regions of parameter space under which certain measure of performance are met \citep{herman:2015}.
+For example, \cref{fig:scenario-map-height-slr} shows expected total costs as a function of $\Delta h$ and \gls{slr} from 2022-2092 for a single value of $h_0$.
+As in \cref{fig:scenario-map-slr-cost}, for high values of $\Delta h$ the expected total costs are dominated by up-front costs, and are thus relatively insensitive to deep uncertainty in \gls{slr}.
+However, for lower values of $\Delta h$ the expected total costs are highly sensitive to the modeled \gls{slr}.
 
-The exploreatory analysis of the previous subsection considered each realization of future sea level to be an exploratory sample from the space of possible futures.
-An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of the sixteen models of sea level rise, as shown in \cref{fig:lsl-evolution}.
-This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.
+Because this exploratory analysis makes no assumptions about the relative likelihood of different \glspl{sow}, no probabilistic conclusions can be drawn.
+For example, while an analyst can report the fraction of \glspl{sow} that meet some criterion (often called a robustness metric), this fraction is in general highly sensitive to the strategy used to sample \glspl{sow} \citep{quinn_exploratory:2020} and the choice of robustness measure used \citep{mcphail_robustness:2019}, and may not correspond to the expected probability of this criterion being met in the future.
+For this reason many analyses first use exploratory modeling to identify the conditions under which different decisions are optimal, then look to qualitative insight from models and expert judgement as to which scenarios are most likely \citep[\eg,][]{Brown:2012kb,Steinschneider:2015kk}.
+To quantify this insight, however, a more formal approach is needed.\james{Revisit this wording perhaps}
 
-A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
-Specifically, for each of $K=16$ models, the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
-Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
-We can then use this conditional distribution over outcomes to create metrics that depend only on the decision, such as averages or quantiles of outcomes.
+\section{Scenario-conditional analysis: the multiple PDF problem}\label{sec:multiple-pdf}
+
+In this section we build on the exploratory analysis of \cref{sec:exploratory} by aggregating insights across \glspl{sow} to quantify the probability distribution of outcomes, conditional on a particular model.
+This probability distribution over outcomes can be used to quantify costs and benefits, map Pareto frontiers, or formulate optimal decisions.
+
+To develop a set of \glspl{pdf}, we next interpret the available \glspl{sow} is as \gls{iid} draws from each of the sixteen models of \gls{slr}, as shown in \cref{fig:lsl-evolution}.
+A formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}: for each of $K=16$ models, we interpret all \glspl{sow} generated with that model as \gls{iid} draws from the distribution of outcomes.
+We can then use this conditional distribution over outcomes to define metrics that depend only on the decision, such as averages or quantiles of outcomes.
 
@@
 The multiple \gls{pdf} problem has been shown in other contexts.
-For example, \citet{sharma_rcp:2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
+For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{If we think of other examples let's put them here}
+
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
@@
 
-\citet{schneider_scenarios:2002} wrote, in the context of \gls{ipcc} assessments, that ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
-Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.
+\citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
+In the context of climate risk management in general, and physical floodproofing of buildings in the coastal zone more specifically, an unwillingness to assign probabilities to \gls{rcp} scenarios defers to the end user (\ie, the homeowner, property manager, or local government) the choice of which scenario to design for.
+Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
+Thus, motivated by the multiple \gls{pdf} problem illustrated in the previous section, we present here a framework for synthesizing information across multiple \glspl{pdf} using a subjective Bayesian interpretation of probability.
 
@@
 
-Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{sampling}(\Omega)$, does not reflect our subjective belief about their true distribution, $p_\mathrm{belief}(\Omega)$.
-Indeed, this is often by design: it is often advantageous to sample low-probability but high-consequence \glspl{sow} to learn about a system's behavior \citep{bankes:1993}.
-To address this, we apply a scheme to weight each \gls{sow} based on $p_\mathrm{belief}(\Omega)$.
-First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional representation, which we denote $\psi_1, \ldots, \psi_J$.
-As discussed in \citet{csillery_abc:2010} and other research on approximate Bayesian computation, the choice of summary statistic is an important one.
-We take the summary statistic to be sea level rise from 2022 to 2100, recognizing that future work could consider multidimensional summaries.
+Consider a set of \glspl{sow} $\qty{s_1, s_2, \ldots, s_j}$.
+In general these \glspl{sow} may deliberately over-sample low-probability but potentially high-consequence futures, and so the distribution they are drawn from will not match our subjective belief about their true distribution, $p_\mathrm{belief}(\Omega)$.
+To address this, we draw from approaches to sampling and poststratification.
 
+First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a low-dimensional representation, which we denote $\psi_1, \ldots, \psi_J$.
+This step is commonly used in \gls{abc} for the purpose of performing inference over parameters using a black-box model \citep[see][]{csillery_abc:2010}.
+We take this low-dimensional statistic ($\phi$) to be \gls{slr} from 2022 to 2100, recognizing that future work could consider multidimensional summaries.
+
+This projection simplifies analysis considerably by allowing us to define $p_\mathrm{belief}(\psi \in \Psi)$ instead of $p_\mathrm{belief}(s \in \Omega)$ as reasoning about a single quantity may be more straightforward than exploring a full probability distribution over time series.
 Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-This allows us to define $p_\mathrm{belief}(\psi \in \Psi)$ instead of $p_\mathrm{belief}(s \in \Omega)$.
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, \ie
-$$
-    F_\mathrm{belief}(\psi) = \int_{-\infty}^{s'} p_\mathrm{belief}(\psi) \dd{\psi},
-$$
-we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:grid-sketch}.
+Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as illustrated in \cref{fig:grid-sketch}:
 \begin{equation}\label{eq:weights}
@@
 David Spiegelhalter had a nice comment on a podcast where he explained that in medicine, there isn't a ``true'' probability that, for example, you will get cancer; this is not some real quantity that can be measured \citep[hence the famous ``probability isn't real'' admonisment of][]{definetti_probability:1972}.
-Similarly, the probability distribution of sea level rise in 2100 isn't something that can be measured.
+Similarly, the probability distribution of \gls{slr} in 2100 isn't something that can be measured.
 However, probability provides a clear and self-consistent mathematical language for reasoning about uncertainty, which is why we're using it.
@@
 
-Need to explain how we came up with our two or three subjective priors.\klaus{Are there published assessments we should refer to}
-Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
-Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
-\Cref{fig:tradeoffs-by-prior} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
+\subsection{Expert assessment}
+
+We construct two probabilistic models for $p_\text{belief}$, based on expert assessments published by \gls{noaa}.
+Following \citet{fuller_inversion:2017}, we convert published ranges into Normal probability distributions by interpreting the bounds of the published range as $\pm 2$ standard deviations from the mean.
+\begin{enumerate}
+    \item Table~2.4 of \citet{sweet_slr:2022} provides a range of \gls{slr} scenarios for different regions of the country; for the Northeast this range is \SIrange{0.6}{2.1}{\meter} (\SIrange{2.0}{6.9}{ft}) for the low and high scenarios, respectively.
+These ranges closely those of \citet{pfeffer_sealevel:2008} used in \citet{fuller_inversion:2017}.
+We refer to this model for $p_\text{belief}$ as ``All \gls{ssp}.''
+    \item \Cref{fig:lsl-evolution} indicates that the highest (lowest) sea level rises occur under \gls{rcp} 8.5 (2.6), which are unlikely under current global pathways \citep{hausfather_scenarios:2020}.
+Table~2.4 of \citet{sweet_slr:2022} gives projections of total \gls{slr}, relative to a 2005 baseline under different \glspl{ssp}.
+Discarding \gls{ssp} 1 (corresponding to \gls{rcp} 2.6) and \gls{ssp} 5 (corresponding to \gls{rcp} 8.5), a plausible range of \SIrange{0.4}{0.92}{\meter} (\SIrange{1.3}{3.0}{ft}) is given.
+We refer to this model for as ``Likely SSPs.''
+\end{enumerate}
+\Cref{fig:lsl-priors-weights}(A) shows a visual representation of the models for $p_\text{belief}$ in panel.
+
+A practical advantage of our weighting method (\cref{eq:weights}) is that changing the weights does not require re-sampling the \glspl{sow} or recomputing outcomes.
+Thus it is computationally cheap to perform robustness checks that compare the effect of different weighting schemes
 
 \subsection{Implications}
+
+A few sentences explaining this trade-off curve\ldots\james{Add more}
+Notably, they give different guidance.
+If $p_\text{belief}$ is based on all \glspl{ssp}, elevating the house by approximately \SI{6}{ft} saves approximately 15\% of the house value relative to not elevating.
+Conversely if $p_\text{belief}$ is based on the likely \gls{ssp} scenarios, then elevating the house costs approximately 15\% of the house value relative to not elevating.
+Comparing the colored lines that synthesize uncertainty to the gray lines that consider a single scenario indicates the importance of integrating across uncertainties.
 
@@
     \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the synthesizing Bayesian models.
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for two models of $p_\text{belief}$ (colors).
+        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
     }\label{fig:tradeoffs-by-prior}
@@
 
-This technique can also be used as a diagnostic tool.
-In \cref{fig:lsl-priors-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}
+This technique can also be used as a diagnostic tool\ldots\james{To do}
+Specifically: what would I need to believe in order to justify some assumption?
 
@@
         Subjective priors for local sea level.
-        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
+        We develop two distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
         The \glspl{pdf} of these subjective priors are shown in panel (A).
-        In panels (B-D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
-        Specifically, (B-D) show the average weight given to each model by each of the three subjective priors.
-        The ``Slow SLR'' prior (B) places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
-        The ``Intermediate SLR'' (C) prior places approximately equal weight on all models.
-        The ``Fast SLR'' prior (D) places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
+        In panels (B-C) we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
+        Specifically, (B-C) show the average weight given to each model by each of the subjective priors.
     }\label{fig:lsl-priors-weights}
@@
 
-Quick summary of what we have done
-\begin{enumerate}
-    \item OK
-\end{enumerate}
 Quick summary of what our motivation was
@@
     \item Be transparent about how we are calculating trade-offs
+    \item Work with existing physical model ensembles
 \end{enumerate}
-Philosophical links to Bayesian philosophy and model building in the $\mathcal{M}$-open case (NB: this is copied and pasted from my thesis -- need to rework)
+Quick summary of what we have done
 \begin{enumerate}
-    \item  \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
-          \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
-    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
-    \item More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
-    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.\james{This should go to intro}
-    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.\james{Explain this to the audience}
-    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
-          More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
-    \item Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}
-    \item Our approach has interesting theoretical links to Bayesian model averaging or stacking \citep{Yao:2018bu}
-\end{enumerate}
-Links to DMDU literature
-\begin{enumerate}
-    \item We should also cite \citet{quinn_exploratory:2020}, which makes the point that it matters how you sample \glspl{sow}.
-    \item Exploratory modeling can help to demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
-    \item You can do scenario discovery
-    \item We are in effect calculating robustness metrics \citep{mcphail_robustness:2019,herman:2015}
-    \item \citet{schneider_scenarios:2002} notes that a RDM analysis ``still requires the construction of ranges of future outcomes -- itself a subjective assessment.
-Subjectivity is simply inherent in all future projections, whether in natural or social sciences.''
+    \item Will fill in.
 \end{enumerate}
@@
 
+Limitations of the re-weighting
+\begin{enumerate}
+    \item Our prior $p(\Omega)$ is limited -- we are just using \gls{msl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
+    \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
+    \item The Uniform and Normal approaches used in \citet{fuller_inversion:2017} were applied for their simplicity and interpretability, but we note that they may not represent a fully calibrated prior belief about \gls{slr}; we refer the reader to \citet{gelman_workflow:2020} for a discussion of model checking steps that be applied.
+\end{enumerate}
 Limitations of the case study
@@
     \item Better data on cost of elevation and depth-damage
-    \item Robustness: ptimize separately for different kinds of house structures and locations
+    \item Scalability: optimize separately for different kinds of house structures and locations
     \item Timing
-\end{enumerate}
-Limitations of the method
-\begin{enumerate}
-    \item Our prior $p(\Omega)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
-    \item We neglect true ignorance \citep{knight_risk:1921}.\james{Perhaps ``unknown unknowns''}
-    \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
+    \item More physics -- we neglect hydrodynamics and tides
 \end{enumerate}
@@
 
-\subsection{Implications}
+\subsection{Policy relevance}
 
@@
 \begin{enumerate}
-    \item Direct
-          \begin{enumerate}
     \item Stormwater management \citep{sharma_rcp:2021,lopez-cantu:2018}
     \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}
-          \end{enumerate}
-    \item Indirect
-          \begin{enumerate}
     \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!\james{Let's emphasize this -- really important to policy analysis under deep uncertainty}
     \item Model structure uncertainties
+    \item Local government revenues are sensitive to flood losses; this approach can be used to structure updates to building codes \citep{shi_municipal:2020}
 \end{enumerate}
-\end{enumerate}
+Key summary about the importance of integrating across uncertainties
 
@@
 
-This work was supported by the National Oceanic and Atmospheric Administration (NOAA) through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
+This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
 JDG thanks Rice University for support.
@@
 JDG wrote the codes and ran simulations.
-JDG wrote the manuscript with support from KK.
-JDG wrote the manuscript in consulation with KK.
+JDG wrote the manuscript in consultation with KK.
 JDG and KK revised the manuscript.
@@
 \newpage
+\section{Supplemental methods}
+
+\subsection{Algorithm to estimate expected damages}\label{sec:alg-ead}
+
+\begin{enumerate}
+    \item For $k=1, \ldots, K$:
+          \begin{enumerate}
+              \item sample a draw from the posterior distribution of flood hazard to get $\qty{ \mu_k, \sigma_k, \xi_k }$
+              \item simulate a single storm surge from the stationary distribution (\cref{eq:surge-model}) and add the mean sea level to get total flood depth $y^\mathrm{sim}_k$
+              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $k$th damage.
+          \end{enumerate}
+    \item Estimate expected annual damages as the sample mean of the $K$ estimates
+\end{enumerate}
+
+\subsection{Surrogate model for expected annual damages}\label{sec:surrogate-ead}
+Evaluating expected annual damages for each of $N$ simulations of \gls{slr}, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
+Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
+However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
+To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
+We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
+This model is shown in \cref{fig:cost-expected-damage-emulator}.
+Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
+
 \section{Supplemental figures}
@@
     \caption{
-        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{lsl}.
+        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{msl}.
         Damages ($y$ axis) are shown as a percentage of house value.
@@
     }\label{fig:surge-gev-priors}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-prior-return}
+    \caption{
+        Surge prior
+    }\label{fig:surge-prior-return}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{surge-synthetic-data-experiment}
+    \caption{
+        Synthetic data experiment as a positive control test for the \gls{gev} model of storm surge.
+        A synthetic record was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively (dots).
+        These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:storm-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
+        The blue line shows the true quantiles of the (known) \gls{gev} distribution.
+        By random chance the sample maximum has a true return period of $\gg 250$ years, which increases the upper confidence interval of the estimated return probabilities, but the true value is nevertheless within the 50\% posterior confidence interval.
+        This experiment yields similar results for alternative values of the known \gls{gev} distribution, and for different random seeds (not shown).
+    }\label{fig:surge-synthetic-data-experiment}
 \end{figure}
@@
     \caption{
-        Expected total lifetime cost (damages plus up-front cost) as a function of sea level rise over the house lifetime and height increase $\Delta h$.
+        Expected total lifetime cost (damages plus up-front cost) as a function of \gls{slr} over the house lifetime and height increase $\Delta h$.
         Initial house elevation is fixed to \SI{1}{ft} below the \gls{bfe}.

### 2022-03-16 tikz/xlrm.tex
edits esp base SLR priors on NOAA report

@@
 
-    \node [font=\color{rosso}\selectfont] at (-4.5, 3.5) {\Large Uncertainties};
+    \node [font=\color{rosso}\selectfont] at (-6, 2.5) {\Large Uncertainties};
     \node [relation] (relationships) at (0, -0.5) {Engineering and economic models};
-    \node [uncertainty] (surge) at (-3.25, 2.5) {Storm surge\\{\footnotesize (draws from 1 posterior)}};
+    \node [uncertainty] (surge) at (-3, 1.5) {Storm surge};
     \path [line] (surge) to (relationships);
 
-    \node [font=\color{thirds}\selectfont] at (0, -1.75) {\Large Relationships};
-    \node [uncertainty] (slr) at (-3.5, -1) {Sea level\\{\footnotesize (16 PDFs)}};
+    \node [font=\color{thirds}\selectfont] at (0, -2) {\Large Relationships};
+    \node [uncertainty] (slr) at (-3.5, -0.5) {Sea level};
     \path [line] (slr) to (relationships);
 
-    \node [font=\color{giallo}\selectfont] at (0, 3) {\Large Levers};
-    \node [lever] (elevate) at (0, 2) {How high to elevate?};
+    \node [font=\color{giallo}\selectfont] at (0, 2.5) {\Large Levers};
+    \node [lever] (elevate) at (0, 1.5) {How high to elevate?};
     \path [line] (elevate) to (relationships);
 
-    \node [font=\color{giallo}\selectfont] at (3.5, 2.25) {\Large Metrics};
+    \node [font=\color{giallo}\selectfont] at (3.5, 2.5) {\Large Metrics};
     \node [metric] (ccost) at (3.5, 1) {Up-front\\(construction)\\cost};
@@
 
-    \node[uncertainty] (rcp) at (-4.5, -2.5) {RCP scenario\\{\footnotesize (4 Scenarios)}};
-    \node[uncertainty] (ice) at (-4.5, 0.75) {Ice sheet dynamics\\{\footnotesize (4 Models)}};
+    \node[uncertainty] (rcp) at (-6, -1.5) {RCP scenario\\{\footnotesize (4 Scenarios)}};
+    \node[uncertainty] (ice) at (-6, 0.75) {Ice sheet dynamics\\{\footnotesize (4 Models)}};
     \path [line] (rcp) to (slr);

### 2022-03-16 scripts/01-surge-modeling.jl

@@ -22,7 +22,7 @@ function plot_surge_gev_priors()
     # loop through all the distributions
     for (prior, c) in zip(HouseElevation.gev_priors, colors)
         plot!(
-            p, 0:0.025:25, prior.dist; label="$(prior.rt) Year Surge", linewidth=3, color=c
+            p, 0:0.05:30, prior.dist; label="$(prior.rt) Year Surge", linewidth=3, color=c
         )
     end
     savefig(p, plots_dir("surge-gev-priors.pdf"))
@@ -101,7 +101,6 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
     # get fake data
     dist = Distributions.GeneralizedExtremeValue(4, 0.5, 0.15)
     fake_data = make_fake_data(dist, N)
-    rts = 1 ./ (1 .- cdf.(dist, fake_data))
 
     # fit the model
     posterior = HouseElevation.get_posterior(
@@ -116,7 +115,15 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
     p = plot_return_period(post_gevs)
 
     # add true line
-    scatter!(p, rts, fake_data; label="Obs (Known Return Period)", color=colors[1], alpha=1)
+    rts = range(1.25, 275; length=250) # return periods
+    aeps = 1 .- 1 ./ rts # annual exceedance probability
+    plot!(p, rts, quantile.(dist, aeps); linewidth=2, color=:blue, label="True")
+
+    # add data
+    xp, ys = HouseElevation.weibull_plot_pos(fake_data)
+    scatter!(p, 1 ./ xp, ys; label="Obs (Weibull Plot Pos.)", color=colors[1], alpha=1)
+
+    plot!(p, legend=:topleft)
 
     # save
     savefig(p, plots_dir("surge-synthetic-data-experiment.pdf"))

### 2022-03-16 scripts/06-combining.jl

@@ -1,5 +1,7 @@
+using Distributions
 using LaTeXStrings
 using Plots
+using Unitful
 
 function plot_grid_scheme()
     dist = Distributions.Normal()
@@ -67,22 +69,41 @@ function plot_grid_scheme()
 end
 
 function get_priors()
-    priors = [
-        (name="Slow SLR", dist=Gamma(1.75, 0.5)),
-        (name="Intermediate SLR", dist=Gamma(1.75, 1.25)),
-        (name="Fast SLR", dist=Gamma(3.5, 1.25)),
-    ]
+    all_ssp = let
+        lb = Unitful.ustrip(u"ft", 0.6u"m") # Table 2.5
+        ub = Unitful.ustrip(u"ft", 2.1u"m") # Table 2.5
+        μ = (ub + lb) / 2
+        σ = (ub - lb) / 4
+        (name="All SSPs", dist=Normal(μ, σ))
+    end
+    intermediate_ssp = let
+        lb = Unitful.ustrip(u"ft", 0.40u"m") # Table 2.4
+        ub = Unitful.ustrip(u"ft", 0.92u"m") # Table 2.4
+        μ = (ub + lb) / 2
+        σ = (ub - lb) / 4
+        (name="Likely SSPs", dist=Normal(μ, σ))
+    end
+
+    priors = [all_ssp, intermediate_ssp]
     return priors
 end
 
 function plot_priors()
     priors = get_priors()
     p = plot(;
-        xlabel="Sea Level Rise, 2022-2100, at Sewells Point, VA [ft]",
-        ylabel="Probability Density",
+        xlabel="SLR 2022-2100, at Sewells Point, VA [ft]", ylabel="Probability Density"
     )
     for (prior, color) in zip(priors, [:red, :blue, :green])
-        plot!(p, prior.dist, 0, 12.5; label=prior.name, color=color, linewidth=3)
+        plot!(
+            p,
+            prior.dist,
+            0,
+            9;
+            label=prior.name,
+            color=color,
+            linewidth=5,
+            legend=:topright,
+        )
     end
     p
     #savefig(p, plots_dir("lsl-priors.pdf"))
@@ -105,9 +126,9 @@ function plot_weight(s::Vector{<:HouseElevation.LSLSim})
             :weight,
             group=:rcp,
             palette=colors,
-            legendtitle="RCP",
-            legend=i == 1 ? :topright : false,
-            ylabel="Implicit Weights: $(prior.name)",
[…]

### 2022-03-16 scripts/main.jl

@@ -77,7 +77,6 @@ function main()
         )
     end
 
-    # second scenario map
     let
         s = HouseElevation.get_lsl(; syear=syear, eyear=eyear)
         bfe = calc_bfe(fits, s, syear)

### 2022-03-19 papers/paper/preprint.tex
incorporate some suggestions from klaus; still figuring out a few things before sending for detailed comments

@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
 
@@
 \begin{abstract}
-    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
-    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
-    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
-    However, growing recognition of nonstationarity motivates a forward-looking approach.
-    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
-    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
-    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-\gls{slr} scenarios.
-    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future \gls{slr} to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
-    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future \gls{slr}.
-    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
-    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
+    Climate risk management involves comparing candidate decisions under deep and dynamic uncertainties\ldots
 \end{abstract}
@@
     \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}
+    \item \citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
+          In the context of climate risk management in general, and physical floodproofing of buildings in the coastal zone more specifically, an unwillingness to assign probabilities to \gls{rcp} scenarios defers to the end user (\ie, the homeowner, property manager, or local government) the choice of which scenario to design for.
+          Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
+          Thus, motivated by the multiple \gls{pdf} problem illustrated in the previous section, we present here a framework for synthesizing information across multiple \glspl{pdf} using a subjective Bayesian interpretation of probability.
     \item Technical detail: joint probability distributions over high dimensional spaces are confusing.
@@
 If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data!
-    \item Some scenarios are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019} yet this is generally not acknowledged
+    \item Some scenarios are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019,srikrishnan_probabilistic:2022} yet this is generally not acknowledged in, \eg, \gls{ipcc} analysis
+    \item Technical detail: sampling over ranges is not invariant to parameterization: if I sample $\theta \sim \mathrm{Uniform}(a, b)$ and then resample $\log theta \sim \mathrm{Uniform}(\log a, \log b)$, I'll get different samples and thus different estimates
 \end{enumerate}
@@
 Our primary research question is ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
-We proceed as follows.
-\begin{enumerate}
-    \item In \cref{sec:case-study} we describe our case study in Norfolk, VA.
-    \item  in \cref{sec:exploratory} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
-    \item  in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
-    \item  in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
-    \item  in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.
-\end{enumerate}
+We proceed as follows\ldots
+
+\section{Decision analytic framework}\label{sec:analysis}
+
+In this section we outline our framework for decision analysis under deep uncertainty, maintaining a high level of generality.
+In the next we discuss application to the house elevation case study.
+
+Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $s_1, \ldots, s_J \in \Omega$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $x_1, \ldots, x_I$ (box a in \cref{fig:flowchart}).
+For each scenario $s_j$ and decision $x_i$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
+While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
+    \caption{
+        Outline of the proposed decision analytic framework.
+    }\label{fig:flowchart}
+\end{figure}
+
+\subsection{Explore}\label{sec:analysis-explore}
+
+A first analytical step is to use the model in an ``exploratory'' mode.
+Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through, for example, existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.
+Common applications include the ``discovery'' of particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping of a system's response to \glspl{sow} and decisions \citep{Brown:2012kb,Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Care should be taken when applying these methods, as not only the choice of system model and candidate decisions, but also the choice of how to sample \glspl{sow} strongly influence results \citep[see][]{quinn_exploratory:2020}.
+
+\subsection{Condition}\label{sec:analysis-condition}
+
+Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
+For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``what decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric'' requires estimating the expectation of some function, which has the form
+\begin{equation}
+    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) f(s | x_i) \dd{s}.
+\end{equation}
+Because expectations are defined relative to a probability distribution, \emph{answering these questions necessitates defining a probability distribution over \glspl{sow}}.
+
+In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
+Given some model $M$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\qty{s_1, s_2, \ldots, s_J} \sim p(\Omega | M)$.
+Since the \glspl{sow} are drawn \gls{iid} from $M$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $\hat{p}(u | x_i, M)$.
+Thus, conditional on model $M$, we can compute expectations, global sensitivity analyses, quantiles, \etc.\james{wording}
+
+Many approaches for modeling climate risk management fall into this framework.
+A common approach (to which we return in the case study) is to use probabilistic models with deeply uncertain boundary conditions.
+For example, \ldots\james{Add a water resources and a flood example}
+In this case, $M$ is the deeply uncertain boundary condition (\eg, choice of model and \gls{rcp} scenario).
+This notation also applies to studies that generate an ensemble of \glspl{sow} by sampling parameters across boundaries.
+For example, \citet{lempert_slr:2012} model key parameters as a uniform distribution over a plausible range and \citet{trindade_waterpathways:2020} check the performance of their candidate solutions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
+In our notation, we can write this as the product of $L$ uniform distributions over the plausible range.
+\begin{equation}
+    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell, \mathrm{min}}, \theta_{\ell, \mathrm{max}}).
+\end{equation}
+This uniform distribution allows us to treat the sample as probabilistic so to answer the quiestions described in the first paragraph of this subsection.
+If this model represents at true subjective Bayesian belief then great!
+But most likely it doesn't!
+
+These robust decision analyses often purport to be free from making probabilistic assumptions; we show that they in fact make strong probabilistic assumptions.
+Our bone to pick is not that the subjective assumptions about the likelihood of \glspl{sow} are bad, but that these assumptions are often opaque and presented without critique or validation.\james{Writing needs work}
+
+\subsection{Synthesize}\label{sec:analysis-synthesize}
+
+If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
+One possibility is to discard the $\qty{s_1, s_2, \ldots, s_J}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
+In practice this is impractical because the \glspl{sow} often rely on complicated models that are computationally expensive to re-run.
+Consequently, climate risk management often uses the outputs of structured simulations and analyses \citep[\eg, from the coupled model intercomparison project's climate model simulations][]{meehl_cmip:2000}; it is neither practical nor feasible for local decision-makers to re-run climate models every time they want to sample \glspl{sow} from a different distribution.
+
+Instead of attempting to re-sample the \glspl{sow}, we instead consider re-weighting the \glspl{sow} that are available.
+To do this we first project the \glspl{sow} $\qty{s_1, s_2, \ldots, s_J} \in \Omega$ onto a low-dimensional representation, which we denote $\psi_1, \ldots, \psi_J \in \Psi$.
+This projection step has parallels in in \gls{abc}, where it is used for inferential purposes \citep[see][]{csillery_abc:2010}.
+Projecting the data onto $\Psi$ allows us to specify $p_\mathrm{belief}(\Psi)$ instead of $p_\mathrm{belief}(\Omega)$.
+We then calculate a probabilistic weight $w_j$ for each \gls{sow} $s_j$ that
+
+We present here the 1D case.
+Without loss of generality, assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
+Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as illustrated in \cref{fig:grid-sketch}:
+\begin{equation}\label{eq:weights}
+    w_j = \begin{cases}
+        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
+        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
+        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
+    \end{cases}
+\end{equation}
+For higher dimensional projections, this equation can be extended by (i) partitioning the parameter space into a Voronoi diagram, and (ii) integrating the $p_\mathrm{belief}$ (not $F_\mathrm{belief}$) over the boundaries of each region.\klaus{is this too hand-wavey?}
+From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{grid-sketch}
+    \caption{
+        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
+        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
+        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+    }\label{fig:grid-sketch}
+\end{figure}
+
+The key insight is that we can deliberately over-sample the tails for exploratory modeling but still retain a valid probabilistic interpretation.
+We are still conditioning on a model, but now this is a model that synthesizes across deep uncertainties rather than one that conditions upon them.\james{Need to elaborate on this a bit.}
+We do not think that stakeholders or experts will necessarily agree on $p_\mathrm{belief}$, but since we have to write down our $p_\mathrm{belief}$, it is transparent and can be critiqued and improved \citep{gelman_workflow:2020}.
+
+A practical advantage of our weighting method (\cref{eq:weights}) is that changing the weights does not require re-sampling the \glspl{sow} or recomputing outcomes.
+Thus it is computationally cheap to perform robustness checks that compare the effect of different weighting schemes.
 
@@
         Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
-        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
+        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:ead-led}           \\
         Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
@@
         Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
-        House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
         First year       & $t_i$             & Deterministic: 2022                                                                  &                              \\
-        Last year        & $t_f$             & Deterministic: 2091 unless otherwise noted                                           &                              \\
+        Last year        & $t_f$             & Deterministic: 2092 unless otherwise noted                                           &                              \\
         \bottomrule
@@
 
-\subsection{\gls{slr}}\label{sec:sea-level}
+\subsection{\Glsentrylong{slr}}\label{sec:sea-level}
 
-To inform the house elevation decision, we develop an ensemble \gls{msl} time series.
+To inform the house elevation decision, we develop an ensemble of \gls{msl} time series.
 Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
@@
 
-\subsection{Experiment design}
+\subsection{Prior over \glsentrylong{slr}}\label{sec:priors}
 
-In the remainder of this article we incrementally develop a general framework, illustrated in \cref{fig:flowchart}, for analyzing climate risk management strategies under deep uncertainty.
+We construct two probabilistic models for $p_\text{belief}$ based on expert assessments published by \gls{noaa}.
+Following \citet{fuller_inversion:2017}, we convert published ranges into Normal probability distributions by interpreting the bounds of the published range as $\pm 2$ standard deviations from the mean.
 
-Our framework considers $J$ \glspl{sow} $s_1, \ldots, s_J$ (here representing time series of \gls{msl}) and $I$ discrete decisions $x_1, \ldots, x_I$ (here representing the choice for how high to elevate the house).
-For each combination of these, we calculate the vector of metrics described in \cref{sec:ead-led}, which we denote $u$.
-Letting $f$ be a system model (here: the economic and engineering loss models in \cref{sec:ead-led}) we define $u_{ij} := f(x_i, s_j)$.
+Table~2.4 of \citet{sweet_slr:2022} provides a range of \gls{slr} scenarios for different regions of the country; for the Northeast this range is \SIrange{0.6}{2.1}{\meter} (\SIrange{2.0}{6.9}{ft}) for the low and high scenarios, respectively.
+These ranges closely those of \citet{pfeffer_sealevel:2008} used in \citet{fuller_inversion:2017}.
+We refer to this model for $p_\text{belief}$ as ``All \gls{ssp}.''
 
-We begin in \cref{sec:exploratory} with an exploratory analysis (blue box in \cref{fig:flowchart}) that uses the system model to identify the conditions under which particularly desirable or undesirable outcomes occur.
-This exploratory modeling does not require any assessment of the relative likelihood of the \glspl{sow}, but can only be interpreted as a ``what-if'' analysis.
-We continue in \cref{sec:exploratory} with a scenario-conditional analysis (red box in \cref{fig:flowchart}) that integrates outcomes over all the \glspl{sow} corresponding to a particular model (here \gls{rcp} scenario and physical model structure).
-While this analysis can be used to quantify the conditional distribution of outcomes, this conditional distribution will be different for each model considered.
-We refer to this as ``the multiple \gls{pdf} problem.''
-Finally in \cref{sec:synthesizing} we introduce a method for synthesizing uncertainty across the multiple \glspl{pdf} (orange box in \cref{fig:flowchart}) by weighting each \gls{sow}, given some subjective belief.
+\Cref{fig:lsl-evolution} indicates that the highest (lowest) sea level rises occur under \gls{rcp} 8.5 (2.6), which are unlikely under current global pathways \citep{hausfather_scenarios:2020}.
+Table~2.4 of \citet{sweet_slr:2022} gives projections of total \gls{slr}, relative to a 2005 baseline under different \glspl{ssp}.
+Discarding \gls{ssp} 1 (corresponding to \gls{rcp} 2.6) and \gls{ssp} 5 (corresponding to \gls{rcp} 8.5), a plausible range of \SIrange{0.4}{0.92}{\meter} (\SIrange{1.3}{3.0}{ft}) is given.
+We refer to this model for as ``Likely SSPs.''
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
-    \caption{
-        Outline of the proposed decision analytic framework.
-        We begin in \cref{sec:exploratory} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
-        Next we use scenario-conditional optimization to illustrate the ``multiple \acrshort{pdf} problem'' (region 2).
-        Finally we propose a method for weighting \glspl{sow} to synthesize uncertainties across multiple \glspl{pdf} (region 3).
-    }\label{fig:flowchart}
-\end{figure}
+\Cref{fig:lsl-priors-weights}(A) shows a visual representation of the models for $p_\text{belief}$.
 
-\section{Exploratory analysis}\label{sec:exploratory}
+\section{Results and discussion}\label{sec:results}
+
+We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed re-weighting synthesis method in \cref{sec:results-synthesis}.
+
+\subsection{Exploratory modeling}\label{sec:results-exploratory}
 
 We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions.
-Exploratory modeling has been widely used in water resources management \citep{Brown:2012kb,Poff:2015jn,quinn_exploratory:2020,Steinschneider:2015kk}, analysis of coupled human-natural systems \citep{moallemi_exploratory:2020,moallemi_decisionsupport:2020}, and coastal adaptation \citep{sriver_sealevel:2018} and does not require any \emph{a priori} assessment of how likely a given \gls{sow} may be \citep{bankes:1993,kwakkel_exploratory:2013,reed_multisectordynamics:2022}.
 
@@
 
-\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs) as a function of height increase ($\Delta h$) and \gls{slr} over the house lifetime.
+\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
 The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
@@
 
-This exploratory analysis can be used to quantify the regions of parameter space under which certain measure of performance are met \citep{herman:2015}.
-For example, \cref{fig:scenario-map-height-slr} shows expected total costs as a function of $\Delta h$ and \gls{slr} from 2022-2092 for a single value of $h_0$.
-As in \cref{fig:scenario-map-slr-cost}, for high values of $\Delta h$ the expected total costs are dominated by up-front costs, and are thus relatively insensitive to deep uncertainty in \gls{slr}.
-However, for lower values of $\Delta h$ the expected total costs are highly sensitive to the modeled \gls{slr}.
+This analysis answers ``what-if'' questions like ``given $h_0$ and $\Delta h$, what is the range of total costs a homeowner could face if \gls{slr} over the house lifetime is \SI{1}{ft} or \SI{10}{ft}.''
+For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \citep[\eg, those of][]{sweet_slr:2022} may prove sufficient for decision making.
+This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
+However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions.
 
-Because this exploratory analysis makes no assumptions about the relative likelihood of different \glspl{sow}, no probabilistic conclusions can be drawn.
-For example, while an analyst can report the fraction of \glspl{sow} that meet some criterion (often called a robustness metric), this fraction is in general highly sensitive to the strategy used to sample \glspl{sow} \citep{quinn_exploratory:2020} and the choice of robustness measure used \citep{mcphail_robustness:2019}, and may not correspond to the expected probability of this criterion being met in the future.
-For this reason many analyses first use exploratory modeling to identify the conditions under which different decisions are optimal, then look to qualitative insight from models and expert judgement as to which scenarios are most likely \citep[\eg,][]{Brown:2012kb,Steinschneider:2015kk}.
-To quantify this insight, however, a more formal approach is needed.\james{Revisit this wording perhaps}
-
-\section{Scenario-conditional analysis: the multiple PDF problem}\label{sec:multiple-pdf}
-
-In this section we build on the exploratory analysis of \cref{sec:exploratory} by aggregating insights across \glspl{sow} to quantify the probability distribution of outcomes, conditional on a particular model.
-This probability distribution over outcomes can be used to quantify costs and benefits, map Pareto frontiers, or formulate optimal decisions.
-
-To develop a set of \glspl{pdf}, we next interpret the available \glspl{sow} is as \gls{iid} draws from each of the sixteen models of \gls{slr}, as shown in \cref{fig:lsl-evolution}.
-A formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}: for each of $K=16$ models, we interpret all \glspl{sow} generated with that model as \gls{iid} draws from the distribution of outcomes.
-We can then use this conditional distribution over outcomes to define metrics that depend only on the decision, such as averages or quantiles of outcomes.
+\subsection{Scenario-conditional optimization}\label{sec:results-conditional}
+We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
+Whereas the exploratory analysis in the previous subsection modeled each time series of future sea level as a sample from the space of possible futures, we now model the \glspl{sow} as \gls{iid} draws from each of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
+As discussed in \cref{sec:analysis-condition}, this allows a probabilistic interpretation: conditional on a particular \gls{rcp} scenario and physical model, we can reason about the distribution of outcomes.
 
@@
 
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the total lifetime cost as a function of $\Delta h$ the sixteen models considered (three are highlighted).
-Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages.
+This probabilistic interpretation allows us to plot trade-offs between different metrics.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the total lifetime cost as a function of $\Delta h$ the sixteen models considered (we highlight three representative models).
+Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
 This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
@@
 
-Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} requires synthesizing across \glspl{sow}.
-However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
-This means that any assessment of decision performance (\eg, an optimal strategy, a Pareto frontier, or a cost-benefit analysis) is conditional upon an (explicit or implicit) model of future outcomes.
-Where these models are implicit, they should be made more explicit to facilitate iterative model critique and improvement.
-
-Second, this approach presents what we call ``the multiple \gls{pdf} problem'' because it leaves decision makers with many probability distributions from which to choose.
-The multiple \gls{pdf} problem has been shown in other contexts.
+This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
+This approach is attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
+However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from.
+The multiple \gls{pdf} problem has also been shown in other contexts.
 For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{If we think of other examples let's put them here}
@@
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
-Further, once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.
 
-\section{Synthesizing PDFs for decision relevance}\label{sec:synthesizing}
+\subsection{Synthesizing deep uncertainties for decision analysis}\label{sec:results-synthesis}
 
-\citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
-In the context of climate risk management in general, and physical floodproofing of buildings in the coastal zone more specifically, an unwillingness to assign probabilities to \gls{rcp} scenarios defers to the end user (\ie, the homeowner, property manager, or local government) the choice of which scenario to design for.
-Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
-Thus, motivated by the multiple \gls{pdf} problem illustrated in the previous section, we present here a framework for synthesizing information across multiple \glspl{pdf} using a subjective Bayesian interpretation of probability.
-
-\subsection{Re-weighting}
-
-Consider a set of \glspl{sow} $\qty{s_1, s_2, \ldots, s_j}$.
-In general these \glspl{sow} may deliberately over-sample low-probability but potentially high-consequence futures, and so the distribution they are drawn from will not match our subjective belief about their true distribution, $p_\mathrm{belief}(\Omega)$.
-To address this, we draw from approaches to sampling and poststratification.
-
-First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a low-dimensional representation, which we denote $\psi_1, \ldots, \psi_J$.
-This step is commonly used in \gls{abc} for the purpose of performing inference over parameters using a black-box model \citep[see][]{csillery_abc:2010}.
-We take this low-dimensional statistic ($\phi$) to be \gls{slr} from 2022 to 2100, recognizing that future work could consider multidimensional summaries.
-
-This projection simplifies analysis considerably by allowing us to define $p_\mathrm{belief}(\psi \in \Psi)$ instead of $p_\mathrm{belief}(s \in \Omega)$ as reasoning about a single quantity may be more straightforward than exploring a full probability distribution over time series.
-Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as illustrated in \cref{fig:grid-sketch}:
-\begin{equation}\label{eq:weights}
-    w_j = \begin{cases}
-        j = 1     & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
-        1 < j < J & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
-        j = J     & 1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
-    \end{cases}
-\end{equation}
-
-TODO\james{This paragraph needs some work}
-The key point to get across is what a subjective belief means in the Bayesian context \citep{savage:1954,gelman_philosophy:2013,gelman_subjectiveobjective:2017}.
-David Spiegelhalter had a nice comment on a podcast where he explained that in medicine, there isn't a ``true'' probability that, for example, you will get cancer; this is not some real quantity that can be measured \citep[hence the famous ``probability isn't real'' admonisment of][]{definetti_probability:1972}.
-Similarly, the probability distribution of \gls{slr} in 2100 isn't something that can be measured.
-However, probability provides a clear and self-consistent mathematical language for reasoning about uncertainty, which is why we're using it.
-A key advantage is that \textbf{since we can't be right, we can at least be transparent} about the assumed values of sea level in 2100, rather than hiding behind implicit assumptions.
-Thus we refer to the $M_k$ here as \emph{priors} to reflect that they consistently represent our current knowledge.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{grid-sketch}
-    \caption{
-        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
-        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
-    }\label{fig:grid-sketch}
-\end{figure}
-
-\subsection{Expert assessment}
-
-We construct two probabilistic models for $p_\text{belief}$, based on expert assessments published by \gls{noaa}.
-Following \citet{fuller_inversion:2017}, we convert published ranges into Normal probability distributions by interpreting the bounds of the published range as $\pm 2$ standard deviations from the mean.
-\begin{enumerate}
-    \item Table~2.4 of \citet{sweet_slr:2022} provides a range of \gls{slr} scenarios for different regions of the country; for the Northeast this range is \SIrange{0.6}{2.1}{\meter} (\SIrange{2.0}{6.9}{ft}) for the low and high scenarios, respectively.
-These ranges closely those of \citet{pfeffer_sealevel:2008} used in \citet{fuller_inversion:2017}.
-We refer to this model for $p_\text{belief}$ as ``All \gls{ssp}.''
-    \item \Cref{fig:lsl-evolution} indicates that the highest (lowest) sea level rises occur under \gls{rcp} 8.5 (2.6), which are unlikely under current global pathways \citep{hausfather_scenarios:2020}.
-Table~2.4 of \citet{sweet_slr:2022} gives projections of total \gls{slr}, relative to a 2005 baseline under different \glspl{ssp}.
-Discarding \gls{ssp} 1 (corresponding to \gls{rcp} 2.6) and \gls{ssp} 5 (corresponding to \gls{rcp} 8.5), a plausible range of \SIrange{0.4}{0.92}{\meter} (\SIrange{1.3}{3.0}{ft}) is given.
-We refer to this model for as ``Likely SSPs.''
-\end{enumerate}
-\Cref{fig:lsl-priors-weights}(A) shows a visual representation of the models for $p_\text{belief}$ in panel.
-
-A practical advantage of our weighting method (\cref{eq:weights}) is that changing the weights does not require re-sampling the \glspl{sow} or recomputing outcomes.
-Thus it is computationally cheap to perform robustness checks that compare the effect of different weighting schemes
-
-\subsection{Implications}
-
-A few sentences explaining this trade-off curve\ldots\james{Add more}
-Notably, they give different guidance.
-If $p_\text{belief}$ is based on all \glspl{ssp}, elevating the house by approximately \SI{6}{ft} saves approximately 15\% of the house value relative to not elevating.
-Conversely if $p_\text{belief}$ is based on the likely \gls{ssp} scenarios, then elevating the house costs approximately 15\% of the house value relative to not elevating.
-Comparing the colored lines that synthesize uncertainty to the gray lines that consider a single scenario indicates the importance of integrating across uncertainties.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
-    \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for two models of $p_\text{belief}$ (colors).
-        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
-    }\label{fig:tradeoffs-by-prior}
-\end{figure}
-
-This technique can also be used as a diagnostic tool\ldots\james{To do}
-Specifically: what would I need to believe in order to justify some assumption?
+To overcome the limitations of scenario-conditional analysis, in this subsection we illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
+We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:priors}: a low-confidence model based on \gls{noaa} \gls{slr} projections under all \gls{ssp} scenarios, and a higher-confidence model that considers \gls{noaa} \gls{slr} projections only under intermediate \gls{ssp} scenarios.
+These two distributions are shown in \cref{fig:lsl-priors-weights}(A).
 
@@
 
-\section{Discussion and conclusions}\label{sec:conclusions}
+This method can be used in a diagnostic framework.
+\Cref{fig:lsl-priors-weights}(B-C) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
+This diagnostic analysis sheds light on the assumptions that are consistent with each $p_\mathrm{belief}$.
+The model for $p_\mathrm{belief}$ that considers a wider distribution over \gls{slr} rates places high weights on the DP16 simulations, particularly \gls{rcp} 8.5, whereas the more narrow distribution over \gls{slr} places more equal weights across physical models and \gls{rcp} scenarios.
+
+This method can also be used to calculate expectations, allowing us to revisit the trade-off diagrams of \cref{fig:tradeoffs-by-rcp}.
+\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) for under both models.
+Notably, they give different guidance.
+If $p_\text{belief}$ is based on all \glspl{ssp}, elevating the house by approximately \SI{6}{ft} saves approximately 15\% of the house value relative to not elevating.
+Conversely if $p_\text{belief}$ is based on the likely \gls{ssp} scenarios, then elevating the house costs approximately 15\% of the house value relative to not elevating.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
+    \caption{
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for two models of $p_\text{belief}$ (colors).
+        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
+    }\label{fig:tradeoffs-by-prior}
+\end{figure}
+
+\section{Conclusions}\label{sec:conclusions}
 
@@
 \begin{enumerate}
+    \item We are motivated by the work of \citet{sriver_sealevel:2018} and \citet{lempert_slr:2012}, which points out that while for the particular case study at hand probabilistic and RDM analyses gave similar results (unsurprising since the probabilistic analysis used uniform priors over ranges), the motivation of RDM was an interactive,  problem-focused and iterative exploration with stakeholders.
+We are motivated to develop a method with more formal probabilistic interpretations that can also be used in this iterative framework.
     \item Explore every \gls{sow}
@@
 \begin{enumerate}
-    \item Will fill in.
+    \item It matters a lot which scenario you design for
+    \item Robustness metrics and robust decision analyses make implicit assumptions about the plausibility of future scenarios
 \end{enumerate}
-
-\subsection{Limitations and future work}
-
 Limitations of the re-weighting
@@
     \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
-    \item The Uniform and Normal approaches used in \citet{fuller_inversion:2017} were applied for their simplicity and interpretability, but we note that they may not represent a fully calibrated prior belief about \gls{slr}; we refer the reader to \citet{gelman_workflow:2020} for a discussion of model checking steps that be applied.
+    \item Normal distributions used in \citet{fuller_inversion:2017} were applied for their simplicity and interpretability, but we note that they may not represent a well-calibrated prior belief about \gls{slr}; we refer the reader to \citet{gelman_workflow:2020} for a discussion of prior predictive checks that could be applied.
+    \item Perhaps discarding the edge \gls{ssp} scenarios is unhelpful;
 \end{enumerate}
@@
     \item Interacting, sequential decisions
+    \item Once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.
+We hypothesize this will lead to under-valuation of adaptive measures.
+    \item Ultimately, developing probabilistic scenarios of future emissions \citep[as in][]{srikrishnan_probabilistic:2022}, temperature, and sea level rise would be valuable.
 \end{enumerate}
-
-\subsection{Policy relevance}
-
-This approach can be used in many other contexts
+Another clarification about how to interpret the $p_\mathrm{belief}$
+\begin{enumerate}
+    \item As statistitian David Spiegelhalter discusses, this subjective belief should not be interpreted as a population distribution because there is no true population distribution that could be answered by collecting more data \citep{andorra_spiegelhalter:2021}.\james{Contextualize this}
+    \item
+\end{enumerate}
+This approach can be applied to lots of important problems
 \begin{enumerate}
@@
 \end{enumerate}
-Key summary about the importance of integrating across uncertainties
+We call for researchers studying climate risk management to make their implicit probability distributions more transparent, and we call for coordinated national and international guidance to assist local governments and engineering bodies determine appropriate subjective beliefs.
 

### 2022-03-19 scripts/01-surge-modeling.jl

@@ -51,8 +51,9 @@ function plot_annmax_floods(annual::HouseElevation.AnnualGageRecord)
         surge_ft;
         label=false,
         markercolor=:gray,
-        ylabel="Ann-Max Storm Surge at $(annual.stn.gage_name) [ft]",
+        ylabel="Ann-Max Storm Surge [ft]",
         left_margin=5mm,
+        markerstrokewidth=0,
     )
 
     # show historic storms
@@ -65,7 +66,7 @@ function plot_annmax_floods(annual::HouseElevation.AnnualGageRecord)
         y0 = y_text + 0.2 * (yobs - y_text)
         y1 = y_text + 0.9 * (yobs - y_text)
         color = storm.is_tc ? colors[1] : colors[2]
-        plot!([x0, x1], [y0, y1]; color=color, linewidth=1.5, label="")
+        plot!([x0, x1], [y0, y1]; color=color, linewidth=1.5, label="", markerstrokewidth=0)
         annotate!(p, x_text, y_text, text(storm.name, :center, 7; color=color))
     end
     return p
@@ -123,7 +124,7 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
     xp, ys = HouseElevation.weibull_plot_pos(fake_data)
     scatter!(p, 1 ./ xp, ys; label="Obs (Weibull Plot Pos.)", color=colors[1], alpha=1)
 
-    plot!(p, legend=:topleft)
+    plot!(p; legend=:topleft)
 
     # save
     savefig(p, plots_dir("surge-synthetic-data-experiment.pdf"))
@@ -258,7 +259,7 @@ function plot_surge_obs_return(
     xlabel!(p1, "Time [year]")
     p2 = plot_surge_posterior_return(annual, fits)
     p2 = plot(p2; ylabel="")
-    add_panel_letters!([p1]; fontsize=14)
+    add_panel_letters!([p1]; fontsize=12)
     annotate!(p2, [1], [8.5], text("(B)", :left, 14))
     p = plot(
         p1,
@@ -266,7 +267,7 @@ function plot_surge_obs_return(
         link=:y,
         ylims=(2, 9),
         layout=grid(1, 2; widths=[0.6, 0.35]),
-        size=[1000, 350] .* 1.15,
+        size=[1000, 350],
         bottommargin=8mm,
         leftmargin=8mm,
     )

### 2022-03-19 scripts/02-mean-sea-level.jl

@@ -67,7 +67,6 @@ function plot_lsl_boxplots_2100(all_trajs::Vector{<:HouseElevation.LSLSim})
         palette=colors,
         legend=(0.15, 0.8),
         legendtitle="  RCP  ",
-        yformatter=blank_formatter,
         ylabel="Mean Sea Level in 2100",
         xrotation=30,
     )

### 2022-03-19 scripts/04-scenario-maps.jl

@@ -82,7 +82,7 @@ function plot_scenario_map_slr_cost(;
             if col == 1 # LHS
                 push!(
                     kwargs,
-                    :ylabel => "Total Cost [% House Value]",
+                    :ylabel => "Total Cost\n[% House Value]",
                     :leftmargin => 7.5mm,
                     :yformatter => pct_formatter,
                 )
@@ -110,7 +110,7 @@ function plot_scenario_map_slr_cost(;
         end
     end
 
-    fake_plot = scatter(
+    cbar_fake_plot = scatter(
         [0, 0],
         [0, 1];
         zcolor=[0, 3],
@@ -118,6 +118,7 @@ function plot_scenario_map_slr_cost(;
         xshowaxis=false,
         yshowaxis=false,
         colorbar_title="Density of SOWs",
+        colorbar_titlefontrotation=90,
         grid=false,
         clims=clims,
         c=cmap,
@@ -126,14 +127,55 @@ function plot_scenario_map_slr_cost(;
         yticks=[],
     )
 
-    fake_plot[1][:xaxis][:showaxis]
-    fake_plot[1][:yaxis][:showaxis]
-    l = @layout [grid(N_row, N_col) a{0.1w}]
+    cbar_fake_plot[1][:xaxis][:showaxis]
+    cbar_fake_plot[1][:yaxis][:showaxis]
+
+    harrow = plot(
+        [0.2, 0.9],
+        [0, 0];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:black,
+        xlims=(0, 1),
+        ylims=(-0.05, 0.25),
+    )
+    annotate!(
+        harrow, [0.5], [0.1], text(L"Increasing House Elevation ($\Delta h$)", :center, 14)
+    )
+
+    varrow = plot(
+        [0, 0],
+        [0.1, 0.9];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:black,
+        xlims=(-0.25, 0.05),
+        ylims=(0, 1),
+    )
+    annotate!(
+        varrow,
+        [-0.1],
+        [0.5],
+        text(L"Increasing Initial Exposure ($h_0$)", :center, 14; rotation=90),
+    )
+
+    l = @layout [
+        a{0.075h}
+        [b{0.05w} grid(N_row, N_col) c{0.125w}]
+    ]
 
     add_panel_letters!(plots)
-
     p = plot(
[…]

### 2022-03-19 scripts/05-tradeoffs-by-rcp.jl

@@ -24,7 +24,7 @@ function plot_rcp_tradeoffs(
     led = map(ui -> ui.led_usd / house_value_usd, u)
 
     # we need ticks to plot
-    x_ticks = 0:2:14 # in feet
+    x_ticks = 0:2:12 # in feet
     cost_fn = HouseElevation.get_elevation_cost_function()
     prop_cost = cost_fn.(x_ticks .* 1u"ft", house_floor_area) ./ house_value_usd
 

### 2022-03-19 scripts/06-combining.jl

@@ -20,7 +20,7 @@ function plot_grid_scheme()
         color=colors[1],
         label="Target Distribution",
         xlims=xlimits,
-        ylabel="CDF",
+        ylabel=L"$F_\mathrm{belief}(\psi)$",
         xlabel="Mean Sea Level in 2100 [ft]",
         legend=:right,
         bottom_margin=5Plots.mm,
@@ -199,23 +199,15 @@ function plot_prior_tradeoffs(
         )
 
         # plot all trade-off lines in light gray
-    for rcp in all_rcp
-        for model in all_models
-            w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
-            cond = vec(mean(var, w; dims=1))
-            plot!(
-                p,
-                Δh_ft,
-                cond;
-                label="",
-                color=:gray,
-                linewidth=0.5,
-                alpha=0.5,
-            )
+        for rcp in all_rcp
+            for model in all_models
+                w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
+                cond = vec(mean(var, w; dims=1))
+                plot!(p, Δh_ft, cond; label="", color=:gray, linewidth=0.5, alpha=0.5)
+            end
         end
-    end
 
-    # plot the conditional beliefs
+        # plot the conditional beliefs
 
         # plot the trade-off for the priors
         for (prior, color) in zip(priors, colors)
@@ -223,7 +215,7 @@ function plot_prior_tradeoffs(
             cond = vec(mean(var, w; dims=1))
             plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=3)
         end
-        
+
         # add the cost on upper x axis
         p = plot!(
             twiny(p),

### 2022-03-19 scripts/main.jl

@@ -35,7 +35,6 @@ function main()
         stn = HouseElevation.TidesAndCurrentsRecord()
         annual = HouseElevation.get_annual_data(stn)
         plot_surge_gev_priors()
-        plot_annmax_floods(annual)
         plot_surge_prior_chains()
         plot_surge_synthetic_experiment(annual)
         plot_surge_posterior_chains(fits)

### 2022-03-19 scripts/plotutils.jl

@@ -51,14 +51,8 @@ function plot_return_period(gevs::Vector{<:Distributions.GeneralizedExtremeValue
         rts = range(1.25, 275; length=250) # return periods
         aeps = 1 .- 1 ./ rts # annual exceedance probability
         xticks = [2, 5, 10, 25, 50, 100, 250]
-    
-        ub1 = [quantile([quantile(d, xi) for d in gevs], 0.95) for xi in aeps]
-        lb1 = [quantile([quantile(d, xi) for d in gevs], 0.05) for xi in aeps]
-        ub2 = [quantile([quantile(d, xi) for d in gevs], 0.9) for xi in aeps]
-        lb2 = [quantile([quantile(d, xi) for d in gevs], 0.1) for xi in aeps]
-        ub3 = [quantile([quantile(d, xi) for d in gevs], 0.75) for xi in aeps]
-        lb3 = [quantile([quantile(d, xi) for d in gevs], 0.25) for xi in aeps]
-    
+        ranges = [0.95, 0.80, 0.5]
+
         p = plot(;
             xlabel="Return Period [years]",
             ylabel="Return Level [ft]",
@@ -66,32 +60,26 @@ function plot_return_period(gevs::Vector{<:Distributions.GeneralizedExtremeValue
             legend=:bottomright,
             xticks=(xticks, string.(xticks)),
         )
-        plot!(
-            rts,
-            ub1;
-            fillbetween=lb1,
-            fillcolor=:gray,
-            fillalpha=0.35,
-            linecolor=false,
-            label="95% Posterior CI",
-        )
-        plot!(
-            rts,
-            ub2;
-            fillbetween=lb2,
-            fillcolor=:gray,
-            fillalpha=0.35,
-            linecolor=false,
-            label="80% Posterior CI",
-        )
-        plot!(
-            rts,
-            ub3;
-            fillbetween=lb3,
-            fillcolor=:gray,
-            fillalpha=0.35,
-            linecolor=false,
-            label="50% Posterior CI",
-        )
+
+        for range in ranges
+            qup = 1 - (1 - range) / 2
+            qlow = (1 - range) / 2
+            ub = [quantile([quantile(d, xi) for d in gevs], qup) for xi in aeps]
+            lb = [quantile([quantile(d, xi) for d in gevs], qlow) for xi in aeps]
+            range_pct = Int(range * 100)
+            plot!(p,
+                rts,
+                ub;
+                fillbetween=lb,
+                fillcolor=:gray,
[…]

### 2022-03-21 papers/paper/preprint.tex
As sent to Klaus for comments

@@
 
-\clearpage
 \section{Introduction}\label{sec:introduction}
@@
 \begin{enumerate}
-    \item In many countries including the United States, flood risk management involves many different public and private sector actors but most rely on probabilistic estimates of flood hazard.
+    \item Formal flood risk management in the United States is implemented through a mix of floodplain management policies, design standards, and insurance programs.
     \item Local governments can build infrastructure to reduce flood risk; they often prioritize these by considering a reduction in damages from a design storm (\eg, the 100 year storm) \citep{hcfcd_prioritization:2019}
-    \item There is a role for industry organizations that set standards that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
+    \item There is a role for industry organizations that set standards (\eg, reliability) that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
     \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by FEMA.
@@
     \item This is related to equifinality where different models fit the historical record equally well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
-    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that
+    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used
     \item \citet{bankes:1993} draws a distinction between exploratory and consolidative modeling
     \item Decision scaling \citep{Steinschneider:2015kk}
-    \item global sensitivity analysis \citep{saltelli_sensitivity:2010,herman_salib:2017,sobol_sensitivity:2001} use feed large scenario ensembles into a black box model, then analyze outputs to assess the relative importance of each input (conditional on an assumed joint distribution over inputs)
+    \item global sensitivity analysis \citep{saltelli_sensitivity:2010,herman_salib:2017,sobol_sensitivity:2001} feeds large scenario ensembles into a black box model, then analyzes outputs to assess the relative importance of each input (conditional on an assumed joint distribution over inputs)
     \item MORDM \citep{kasprzyk:2013} combines exploratory modeling (exploring performance over many scenarios) and policy search (using optimization tools)  \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
 Typically ``robustness metric'' is used \citep{herman:2015,mcphail_robustness:2019}
+    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
     \item For a more complete review see \citet{marchau:2019}
-    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
     \item Crucially, these approaches all emphasize exploring performance over many scenarios, then describing this performance into some set of metrics.
@@
     \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}
-    \item \citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
-          In the context of climate risk management in general, and physical floodproofing of buildings in the coastal zone more specifically, an unwillingness to assign probabilities to \gls{rcp} scenarios defers to the end user (\ie, the homeowner, property manager, or local government) the choice of which scenario to design for.
-          Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
-          Thus, motivated by the multiple \gls{pdf} problem illustrated in the previous section, we present here a framework for synthesizing information across multiple \glspl{pdf} using a subjective Bayesian interpretation of probability.
+    \item \citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.'' Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
     \item Technical detail: joint probability distributions over high dimensional spaces are confusing.
 Bayesian analysis teaches us that posterior distributions of parameters are often highly correlated.
-If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data!
-    \item Some scenarios are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019,srikrishnan_probabilistic:2022} yet this is generally not acknowledged in, \eg, \gls{ipcc} analysis
+If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!
+    \item Some outcomes of deeply uncertain processes are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019,srikrishnan_probabilistic:2022} yet this is generally not acknowledged in, \eg, \gls{ipcc} analysis
     \item Technical detail: sampling over ranges is not invariant to parameterization: if I sample $\theta \sim \mathrm{Uniform}(a, b)$ and then resample $\log theta \sim \mathrm{Uniform}(\log a, \log b)$, I'll get different samples and thus different estimates
@@
 
-Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $s_1, \ldots, s_J \in \Omega$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $x_1, \ldots, x_I$ (box a in \cref{fig:flowchart}).
-For each scenario $s_j$ and decision $x_i$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
-While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
+Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
+For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
+While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.\klaus{Is there better mathematical notation for panel (b)?}
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through, for example, existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.
-Common applications include the ``discovery'' of particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping of a system's response to \glspl{sow} and decisions \citep{Brown:2012kb,Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Care should be taken when applying these methods, as not only the choice of system model and candidate decisions, but also the choice of how to sample \glspl{sow} strongly influence results \citep[see][]{quinn_exploratory:2020}.
+Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.
+Common applications include ``discovering'' particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Brown:2012kb,Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``what decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric'' requires estimating the expectation of some function, which has the form
-\begin{equation}
-    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) f(s | x_i) \dd{s}.
-\end{equation}
+For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric''  requires estimating the expectation of some function $\pi$, which has the form
+\begin{equation*}
+    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
+\end{equation*}
 Because expectations are defined relative to a probability distribution, \emph{answering these questions necessitates defining a probability distribution over \glspl{sow}}.
+The general intractability of these integrals motivates the use of Monte Carlo methods by sampling across many \glspl{sow} \citep{Betancourt:2017vd}.
 
 In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
-Given some model $M$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\qty{s_1, s_2, \ldots, s_J} \sim p(\Omega | M)$.
-Since the \glspl{sow} are drawn \gls{iid} from $M$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $\hat{p}(u | x_i, M)$.
-Thus, conditional on model $M$, we can compute expectations, global sensitivity analyses, quantiles, \etc.\james{wording}
+Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
+Since the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
 
-Many approaches for modeling climate risk management fall into this framework.
-A common approach (to which we return in the case study) is to use probabilistic models with deeply uncertain boundary conditions.
+Many decision analytic tools used in climate risk management can be interpreted through this lens.
+A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
 For example, \ldots\james{Add a water resources and a flood example}
-In this case, $M$ is the deeply uncertain boundary condition (\eg, choice of model and \gls{rcp} scenario).
-This notation also applies to studies that generate an ensemble of \glspl{sow} by sampling parameters across boundaries.
-For example, \citet{lempert_slr:2012} model key parameters as a uniform distribution over a plausible range and \citet{trindade_waterpathways:2020} check the performance of their candidate solutions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
-In our notation, we can write this as the product of $L$ uniform distributions over the plausible range.
+In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
+As we shall illustrate in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only as a.\ldots\klaus{I'm stuck, can you help me finish this sentence?}
+
+Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
+For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
+Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
+Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
 \begin{equation}
-    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell, \mathrm{min}}, \theta_{\ell, \mathrm{max}}).
+    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
 \end{equation}
-This uniform distribution allows us to treat the sample as probabilistic so to answer the quiestions described in the first paragraph of this subsection.
-If this model represents at true subjective Bayesian belief then great!
-But most likely it doesn't!
+If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][for some directions]{gelman_workflow:2020}.
 
-These robust decision analyses often purport to be free from making probabilistic assumptions; we show that they in fact make strong probabilistic assumptions.
-Our bone to pick is not that the subjective assumptions about the likelihood of \glspl{sow} are bad, but that these assumptions are often opaque and presented without critique or validation.\james{Writing needs work}
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to sub-optimal decisions.
 
@@
 If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
-One possibility is to discard the $\qty{s_1, s_2, \ldots, s_J}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
-In practice this is impractical because the \glspl{sow} often rely on complicated models that are computationally expensive to re-run.
-Consequently, climate risk management often uses the outputs of structured simulations and analyses \citep[\eg, from the coupled model intercomparison project's climate model simulations][]{meehl_cmip:2000}; it is neither practical nor feasible for local decision-makers to re-run climate models every time they want to sample \glspl{sow} from a different distribution.
+One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
+In practice this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
+Consequently, climate risk management often uses the outputs of structured simulations and analyses \citep[\eg, from the Coupled Model Intercomparison Project's simulations;][]{meehl_cmip:2000}; it is neither practical nor feasible for local decision-makers to re-run climate models every time they want to sample \glspl{sow} from a different distribution.
 
-Instead of attempting to re-sample the \glspl{sow}, we instead consider re-weighting the \glspl{sow} that are available.
-To do this we first project the \glspl{sow} $\qty{s_1, s_2, \ldots, s_J} \in \Omega$ onto a low-dimensional representation, which we denote $\psi_1, \ldots, \psi_J \in \Psi$.
-This projection step has parallels in in \gls{abc}, where it is used for inferential purposes \citep[see][]{csillery_abc:2010}.
-Projecting the data onto $\Psi$ allows us to specify $p_\mathrm{belief}(\Psi)$ instead of $p_\mathrm{belief}(\Omega)$.
-We then calculate a probabilistic weight $w_j$ for each \gls{sow} $s_j$ that
-
+Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
+To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
+This allows us to specify $p_\mathrm{belief}(\Psi)$ instead of $p_\mathrm{belief}(\Omega)$.
+We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ such that $\sum_{j=1}^J w_j \psi_j$ approximates $p_\mathrm{belief}(\Psi)$.\klaus{Really we are setting the weights so that $\sum_{j=1}^J w_j \pi(s_j)$ approximates $\int_{s} p(s) \pi(s) \dd{s}$ -- is this too hand-wavey?}
 We present here the 1D case.
@@
 \end{equation}
-For higher dimensional projections, this equation can be extended by (i) partitioning the parameter space into a Voronoi diagram, and (ii) integrating the $p_\mathrm{belief}$ (not $F_\mathrm{belief}$) over the boundaries of each region.\klaus{is this too hand-wavey?}
+For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over the boundaries of each region.
 From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
@@
 
-The key insight is that we can deliberately over-sample the tails for exploratory modeling but still retain a valid probabilistic interpretation.
-We are still conditioning on a model, but now this is a model that synthesizes across deep uncertainties rather than one that conditions upon them.\james{Need to elaborate on this a bit.}
-We do not think that stakeholders or experts will necessarily agree on $p_\mathrm{belief}$, but since we have to write down our $p_\mathrm{belief}$, it is transparent and can be critiqued and improved \citep{gelman_workflow:2020}.
-
-A practical advantage of our weighting method (\cref{eq:weights}) is that changing the weights does not require re-sampling the \glspl{sow} or recomputing outcomes.
-Thus it is computationally cheap to perform robustness checks that compare the effect of different weighting schemes.
+The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
+As in \cref{sec:analysis-condition} we must condition on a model, but whereas that analysis conditions upon deep uncertainties, this approach synthesizes them.
+It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
+However, reasoning using subjective probability allows\ldots\james{Finish sentence}
 
@@
 
-We model one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
+We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
 For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
+To use the notation developed in the previous section, we consider the following.
+\begin{enumerate}
+    \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
+    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T$ year house lifetime ($T=71$ generally; see \cref{tab:uncertainties}) so $\vb{s} \in \mathbb{R}^T$
+    \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
+\end{enumerate}
+In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).
 
@@
     }\label{tab:uncertainties}
-    \begin{tabular}{l l p{3in} l}
+    \small
+    \begin{tabular}{l l p{3.5in}}
         \toprule
-        Name             & Symbol            & Uncertainty                                                                          & See more                     \\
+        Name             & Symbol            & Uncertainty                                                                          \\
         \midrule
-        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
-        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
-        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
-        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:ead-led}           \\
-        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
-        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
-        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
-        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
-        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
-        First year       & $t_i$             & Deterministic: 2022                                                                  &                              \\
-        Last year        & $t_f$             & Deterministic: 2092 unless otherwise noted                                           &                              \\
+        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        \\
+        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
+        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
+        Discount rate    & $1-\gamma$        & Fixed at 2.5\%                                                                       \\
+        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  \\
+        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
+        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 \\
+        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                \\
+        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   \\
+        House lifespan   & $T$               & Deterministic: 71 years unless otherwise noted                                       \\
         \bottomrule
@@
 
-\subsection{\Glsentrylong{slr}}\label{sec:sea-level}
+\subsection{\Glsentrylong{slr}}\label{sec:case-slr}
 
-To inform the house elevation decision, we develop an ensemble of \gls{msl} time series.
-Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
+We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
 The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
@@
 Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
-To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probability density functions of \gls{msl}.
-We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \acrfull{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.
+To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
 
@@
 \Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-We return in \cref{sec:multiple-pdf} to the problem that these multiple models poses to decision makers.
+We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.
 
-\subsection{Storm surges}\label{sec:storm-surge}
+\subsection{Storm surges}\label{sec:case-surge}
 
-We model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
-In this subsection we describe the representation of storm surge in our model.
+We model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.
 
-Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
-The NOAA tides and currents dataset is freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
-For each calendar year we first remove the annual mean, then extract the maximum water level.
-We refer these maxima as the annual maximum storm surges $y'(t)$.
+For each calendar year we first remove the annual mean, then calculate the maximum water level.
+We refer this time series of annual maximum storm surges as $y'(t)$.
 
@@
         return periods.
-        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
+        Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula $\nicefrac{r}{N+1}$.
         Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
@@
 \end{equation*}
-Since storm surges cannot be negative, only the latter is physically defensible.
-This is related to the approach of \citet{martins_estimators:2000}, who use physical reasoning to place an informative prior $\xi$.
+Since storm surges cannot be negative, only the latter is physically defensible so we constrain $\xi > 0$.
 Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition may be difficult, we instead apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
-Specifically we apply Normal (truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
+Specifically, we apply Normal (truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
 These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
@@
 Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
-We evaluate the model's fit using posterior predictive checks \citep[section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly (\cref{fig:surge-test-statistics}) although \cref{fig:surge-test-statistics}(b) suggests the possibility of temporal structure not captured by our stationary \gls{iid} model.
+We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
 Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017}.
@@
 \Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position
-\begin{equation}\label{eq:plot-pos}
-    \frac{r}{N + 1}
-\end{equation}
-where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
+The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position, $\nicefrac{r}{N + 1}$, where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
 The good fit of the Bayesian fit (gray shading) to the data (points) suggest the modeling choices made are reasonable and sufficient for this didactic example.
-Finally, a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
+A positive control test (\cref{fig:surge-synthetic-data-experiment}) also validates the model's ability to recover known parameter values.
 
-\subsection{Damages and metrics}\label{sec:ead-led}
+\subsection{Damages and metrics}\label{sec:case-metrics}
 
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
-(HAZUS) depth-damage curves provided by FEMA to parameterize $D(h-\overline{y})$; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
-For comparison, \cref{fig:cost-depth-damage} also shows results using the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
+Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
 Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
@@
 However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
-Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible values $h - \overline{y}(t)$ and use this to train a computationally efficient a surrogate model (see \cref{sec:surrogate-ead}).
+Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient a surrogate model (see \cref{sec:surrogate-ead}).
 
@@
 \end{equation}
-where the discount rate is $1 - \gamma$.
-For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate (see \cref{tab:uncertainties}).
-We calculate lifetime expected damages \emph{for each \gls{sow} separately}.
+where the discount rate is $1 - \gamma$, the initial time $t_i=2022$, and the end time $t_f = t_i + T$.
+Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this study.
+For a more theoretical discussion see \citet{arrow_discount:2013}.
 
@@
 
-\subsection{Prior over \glsentrylong{slr}}\label{sec:priors}
+\subsection{Prior over \glsentrylong{slr}}\label{sec:case-priors}
 
-We construct two probabilistic models for $p_\text{belief}$ based on expert assessments published by \gls{noaa}.
-Following \citet{fuller_inversion:2017}, we convert published ranges into Normal probability distributions by interpreting the bounds of the published range as $\pm 2$ standard deviations from the mean.
+We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
+We use a Gamma distribution for all three priors, parameterized as
+\begin{equation*}
+    f(x | \alpha, \theta) = \frac{x^{\alpha-1} e^{-x/\theta}}{\Gamma(\alpha) \theta^\alpha},
+    \quad x > 0.
+\end{equation*}
+\Cref{tab:slr-priors} specifies these priors, as well as some quantiles of the distributions.
+Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
 
-Table~2.4 of \citet{sweet_slr:2022} provides a range of \gls{slr} scenarios for different regions of the country; for the Northeast this range is \SIrange{0.6}{2.1}{\meter} (\SIrange{2.0}{6.9}{ft}) for the low and high scenarios, respectively.
-These ranges closely those of \citet{pfeffer_sealevel:2008} used in \citet{fuller_inversion:2017}.
-We refer to this model for $p_\text{belief}$ as ``All \gls{ssp}.''
+\begin{table}[h]
+    \centering
+    \caption{
+        Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\text{belief}(\psi)$.
+        The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles.
+    }\label{tab:slr-priors}
+    \input{../../plots/lsl_priors.tex}
+\end{table}
 
-\Cref{fig:lsl-evolution} indicates that the highest (lowest) sea level rises occur under \gls{rcp} 8.5 (2.6), which are unlikely under current global pathways \citep{hausfather_scenarios:2020}.
-Table~2.4 of \citet{sweet_slr:2022} gives projections of total \gls{slr}, relative to a 2005 baseline under different \glspl{ssp}.
-Discarding \gls{ssp} 1 (corresponding to \gls{rcp} 2.6) and \gls{ssp} 5 (corresponding to \gls{rcp} 8.5), a plausible range of \SIrange{0.4}{0.92}{\meter} (\SIrange{1.3}{3.0}{ft}) is given.
-We refer to this model for as ``Likely SSPs.''
+We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
+We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
 
-\Cref{fig:lsl-priors-weights}(A) shows a visual representation of the models for $p_\text{belief}$.
+Future work could develop improved, policy-relevant priors, either by combining probabilistic models for emissions \citep[as in][]{srikrishnan_probabilistic:2022}, climate sensitivity, and the \gls{slr}, or through inversion of expert opinion \citep[as in][]{fuller_inversion:2017}.
+Future work could also consider more sophisticated priors, either by considering higher-dimensional $\psi$ or by modeling a joint prior over model inputs and model outputs through melding \citep{poole_melding:2000,sevcikova_melding:2007}.
 
@@
 
-We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed re-weighting synthesis method in \cref{sec:results-synthesis}.
+We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed synthesis method (\cref{sec:results-synthesis}).
 
@@
 We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
-Whereas the exploratory analysis in the previous subsection modeled each time series of future sea level as a sample from the space of possible futures, we now model the \glspl{sow} as \gls{iid} draws from each of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
+Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draws from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
 As discussed in \cref{sec:analysis-condition}, this allows a probabilistic interpretation: conditional on a particular \gls{rcp} scenario and physical model, we can reason about the distribution of outcomes.
@@
 
-This probabilistic interpretation allows us to plot trade-offs between different metrics.
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the total lifetime cost as a function of $\Delta h$ the sixteen models considered (we highlight three representative models).
+This probabilistic interpretation allows us to compute, for example, expected values.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen models considered (we highlight three representative models).
 Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
@@
 For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{If we think of other examples let's put them here}
-
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
@@
 To overcome the limitations of scenario-conditional analysis, in this subsection we illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
-We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:priors}: a low-confidence model based on \gls{noaa} \gls{slr} projections under all \gls{ssp} scenarios, and a higher-confidence model that considers \gls{noaa} \gls{slr} projections only under intermediate \gls{ssp} scenarios.
-These two distributions are shown in \cref{fig:lsl-priors-weights}(A).
+We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
+These three distributions are shown in \cref{fig:lsl-priors-weights}(A).
 
@@
 
-This method can be used in a diagnostic framework.
-\Cref{fig:lsl-priors-weights}(B-C) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-This diagnostic analysis sheds light on the assumptions that are consistent with each $p_\mathrm{belief}$.
-The model for $p_\mathrm{belief}$ that considers a wider distribution over \gls{slr} rates places high weights on the DP16 simulations, particularly \gls{rcp} 8.5, whereas the more narrow distribution over \gls{slr} places more equal weights across physical models and \gls{rcp} scenarios.
+One application of this method is to diagnose which assumptions different $p_\text{belief}$ are consistent with.
+\Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
+For example, the rapid \gls{slr} scenario places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
+Conversely, the slow \gls{slr} scenario places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
+The uncertain \gls{slr} scenario places approximately equal weight across models.
 
 This method can also be used to calculate expectations, allowing us to revisit the trade-off diagrams of \cref{fig:tradeoffs-by-rcp}.
-\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) for under both models.
+\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
 Notably, they give different guidance.
-If $p_\text{belief}$ is based on all \glspl{ssp}, elevating the house by approximately \SI{6}{ft} saves approximately 15\% of the house value relative to not elevating.
-Conversely if $p_\text{belief}$ is based on the likely \gls{ssp} scenarios, then elevating the house costs approximately 15\% of the house value relative to not elevating.
+Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} saves approximately 25\% of the house value relative to not elevating.
+Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} costs approximately 25\% of the house value relative to not elevating.
+Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house.
+Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
 
@@
     \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for two models of $p_\text{belief}$ (colors).
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\text{belief}$ (colors).
         Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
@@
     \item Work with existing physical model ensembles
+    \item We want to be robust to different plausible beliefs rather than robust to different outcomes
 \end{enumerate}

### 2022-03-21 tikz/bayes-rdm.tex
Track some suggestions from Klaus
As sent to Klaus for comments

@@
 \definecolor{thirds}{HTML}{31467A}
+\definecolor{orange}{HTML}{F18101}
+\definecolor{deepblue}{HTML}{CA3CFF}
 
@@
     % label the various sections
-    \path [fill=grigio!20] (-6.75, 4.8) -- (5.75, 4.8) -- (5.75, -4) -- (-6.75, -4) -- cycle;
-    \node [color=thirds] (s5) at (-4, -3) {\textbf{Synthesize (\S 5)}};
+    \path [fill=orange!30] (-6.75, 4.8) -- (5.75, 4.8) -- (5.75, -4) -- (-6.75, -4) -- cycle;
+    \node [color=thirds] (s5) at (-4, -3) {\textbf{3) Synthesize}};
 
-    \path [fill=giallo!20] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- cycle;
-    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{Condition (\S 4)}};
+    \path [fill=rosso!30] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- cycle;
+    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{2) Condition}};
 
-    \path [fill=thirds!20] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- cycle;
-    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\textbf{Explore (\S 3)}};
+    \path [fill=thirds!30] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- cycle;
+    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\textbf{1) Explore}};
 
@@
             Candidate decisions\\
-            $\qty{x_1, x_2, \ldots, x_I}$
+            $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$
         };
     \node [block, label={[]135:{\footnotesize (b)}}] at (-4, 0) (simulations) {
-            Many\\
-            Simulated futures\\
-            $\qty{s_1, \ldots, s_J} \in \Omega$
+            Many possible futures\\
+            $\vb{s} \in \Omega = \qty{s_1, \ldots, s_J}$
         };
@@
             Weights\\
-            $w_1, \ldots, w_J$
+            $\vb{w} = \qty{w_1, \ldots, w_J}$
         };
@@
             over outcomes\\
-            $\hat{p}(u | x_i, M_k)$
+            $p(u | x_i, M_k)$
         };

### 2022-03-21 HouseElevation/src/HouseElevation.jl

@@ -7,6 +7,7 @@ include("io_lsl.jl") # read in the raw data
 include("cost.jl")
 include("system.jl")
 include("aggregate.jl")
+include("priors.jl")
 
 export data_dir,
     plots_dir,
@@ -17,6 +18,7 @@ export data_dir,
     get_surge_posterior,
     get_outcomes,
     get_metrics,
-    make_weights
+    make_weights,
+    get_priors
 
 end # module

### 2022-03-21 HouseElevation/src/priors.jl

@@ -1,0 +1,43 @@
+using DataFrames
+using Distributions
+using Latexify
+
+"""
+Get the ubjective priors for sea level rise
+
+The values give SLR from 2022 to 2100, in ft
+"""
+function get_priors()
+    priors = [
+        (name="Slow SLR", dist=Gamma(1.75, 0.5)),
+        (name="Uncertain SLR", dist=Gamma(1.75, 1.25)),
+        (name="Rapid SLR", dist=Gamma(3.5, 1.25)),
+    ]
+    return priors
+end
+
+"""Write the priors to file"""
+function write_priors(fname::String)
+    priors = get_priors()
+    df = vcat(
+        [
+            DataFrame(
+                [
+                    "Name" => String(p.name),
+                    L"$\alpha$" => p.dist.α,
+                    L"$\theta$" => p.dist.θ,
+                    [
+                        "Q$(q*100)" => round(quantile(p.dist, q); digits=2) for
+                        q in [0.025, 0.25, 0.5, 0.75, 0.975]
+                    ]...,
+                ],
+            ) for p in priors
+        ]...,
+    )
+    tex_str = Latexify.latexify(
+        df; env=:table, fmt="%.2f", booktabs=true, adjustment=:l, latex=false
+    )
+    open(fname, "w") do io
+        write(io, tex_str)
+    end
+end

### 2022-03-21 HouseElevation/src/surge-bayes.jl

@@ -120,7 +120,7 @@ function write_diagnostics(fit, fname::String)
         "ess" => L"\textrm{ESS}",
         "rhat" => L"$\hat{R}$",
     )
-    tex_str = Latexify.latexify(df; env=:table, fmt="%.3f")
+    tex_str = Latexify.latexify(df; env=:table, fmt="%.3f", booktabs=true, index=:subscript)
     open(fname, "w") do io
         write(io, tex_str)
     end

### 2022-03-21 scripts/06-combining.jl

@@ -1,3 +1,4 @@
+using DataFrames
 using Distributions
 using LaTeXStrings
 using Plots
@@ -68,26 +69,6 @@ function plot_grid_scheme()
     return p
 end
 
-function get_priors()
-    all_ssp = let
-        lb = Unitful.ustrip(u"ft", 0.6u"m") # Table 2.5
-        ub = Unitful.ustrip(u"ft", 2.1u"m") # Table 2.5
-        μ = (ub + lb) / 2
-        σ = (ub - lb) / 4
-        (name="All SSPs", dist=Normal(μ, σ))
-    end
-    intermediate_ssp = let
-        lb = Unitful.ustrip(u"ft", 0.40u"m") # Table 2.4
-        ub = Unitful.ustrip(u"ft", 0.92u"m") # Table 2.4
-        μ = (ub + lb) / 2
-        σ = (ub - lb) / 4
-        (name="Likely SSPs", dist=Normal(μ, σ))
-    end
-
-    priors = [all_ssp, intermediate_ssp]
-    return priors
-end
-
 function plot_priors()
     priors = get_priors()
     p = plot(;
@@ -98,7 +79,7 @@ function plot_priors()
             p,
             prior.dist,
             0,
-            9;
+            10;
             label=prior.name,
             color=color,
             linewidth=5,
@@ -127,7 +108,7 @@ function plot_weight(s::Vector{<:HouseElevation.LSLSim})
             group=:rcp,
             palette=colors,
             legendtitle="  RCP",
-            legend=i == 1 ? :left : false,
+            legend=i == 1 ? :topright : false,
             ylabel="$(prior.name)",
             xrotation=30,
         )
@@ -140,11 +121,11 @@ end
 
 function plot_priors_weights(s::Vector{<:HouseElevation.LSLSim})
     plots = vcat(plot_priors(), plot_weight(s)...)
-    add_panel_letters!(plots; fontsize=11)
+    add_panel_letters!(plots; loc=(0.15, 0.99), fontsize=11)
     p = plot(
         plots...;
-        layout=(1, 3),
-        size=(1200, 400),
+        layout=(2, 2),
+        size=(1000, 750),
         left_margin=5Plots.mm,
         bottom_margin=6Plots.mm,
     )

### 2022-03-21 scripts/main.jl

@@ -21,7 +21,7 @@ function main()
     # define some constants
     house_floor_area = 1500u"ft^2"
     house_value_usd = 200_000.0 # HOUSE NOT LAND VALUE
-    discount_rate = 0.03 # mortgages going for 3-4.5% at the moment
+    discount_rate = 0.025 # mortgages going for 3-4.5% at the moment
     syear = 2022
     eyear = 2092
     # for a **VERY** rough idea of prices see
@@ -98,6 +98,9 @@ function main()
         plot_rcp_tradeoffs(
             u, s, x; house_value_usd=house_value_usd, house_floor_area=house_floor_area
         )
+
+        # write the priors to a table
+        HouseElevation.write_priors(plots_dir("lsl_priors.tex"))
 
         # plot the implicit weight
         plot_priors_weights(s)

### 2022-03-21 papers/paper/preprint.tex
Some tweaks to clean things up

@@
 \pagestyle{fancy}
+\setlength{\headheight}{15pt}
 
@@
 % ACRONYMS
-\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
-\newacronym{abc}{ABC}{approximate Bayesian computation}
-\newacronym{bdt}{BDT}{Bayesian decision theory}
+\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
 \newacronym{bfe}{BFE}{base flood elevation}
-\newacronym[]{cdf}{CDF}{cumulative distribution function}
 \newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
-\newacronym{gcm}{GCM}{general circulation model}
-\newacronym[]{gev}{GEV}{generalized extreme value}
+\newacronym{gev}{GEV}{generalized extreme value}
 \newacronym{iid}{IID}{independent and identically distributed}
 \newacronym{ipcc}{IPCC}{International Panel on Climate Change}
-\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
 \newacronym{msl}{MSL}{mean relative sea level}
@@
 \newacronym{rcp}{RCP}{representative concentration pathway}
-\newacronym{rdm}{RDM}{robust decision making}
 \newacronym{slr}{SLR}{sea level rise}
@@
 \newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
-\newacronym{tc}{TC}{tropical cyclone}
+\makeglossaries % required to make the list of acronyms
 
@@
     \item There is a role for industry organizations that set standards (\eg, reliability) that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
-    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by FEMA.
+    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by \gls{fema}.
 At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).
@@
 This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critiqued
-    \item \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
+    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
 \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
@@
         This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the cumulative distribution function of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
     }\label{fig:grid-sketch}
@@
         Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula $\nicefrac{r}{N+1}$.
-        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
+        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
     }\label{fig:surge-obs-return}
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
+(HAZUS) depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
 For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
@@
 \newpage
+
+\printglossary[type=\acronymtype,title=List of Abbreviations]
+
 \section{Supplemental methods}
@@
 
+We use a Monte Carlo integration to estimate expected damages as a function of house elevation and \gls{msl}, $D(h-y(t))$.
 \begin{enumerate}
-    \item For $k=1, \ldots, K$:
+    \item For $m=1, \ldots, M$:
           \begin{enumerate}
-              \item sample a draw from the posterior distribution of flood hazard to get $\qty{ \mu_k, \sigma_k, \xi_k }$
-              \item simulate a single storm surge from the stationary distribution (\cref{eq:surge-model}) and add the mean sea level to get total flood depth $y^\mathrm{sim}_k$
-              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $k$th damage.
+              \item draw a sample from the posterior distribution of storm surge (see \cref{eq:surge-model})  $\qty{ \mu_m, \sigma_n, \xi_m}$
+              \item simulate a single storm surge from this stationary \gls{gev} distribution and add the mean sea level to get total flood depth $y^\mathrm{sim}_m$
+              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $m$th damage.
           \end{enumerate}
-    \item Estimate expected annual damages as the sample mean of the $K$ estimates
+    \item Estimate expected annual damages as the sample mean of the $M$ estimates
 \end{enumerate}
@@
 \subsection{Surrogate model for expected annual damages}\label{sec:surrogate-ead}
-Evaluating expected annual damages for each of $N$ simulations of \gls{slr}, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
-Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
+Evaluating expected annual damages for each of $N$ simulations of \gls{slr}, each of $J$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
+In our model, we have $T=71$, $J=\num{10000}$, $N=\num{179232}$, and in general .
 However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
@@
         Depth-damage relationship.
-        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA.
+        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by \gls{fema}.
         Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
@@
     \caption{
-        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{msl}.
+        As discussed in \cref{sec:case-metrics}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{msl}.
         Damages ($y$ axis) are shown as a percentage of house value.
@@
         Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
-        See \cref{sec:storm-surge} for details.
+        See \cref{sec:case-surge} for details.
         For the 2, 10, 100, and 500 year events we apply Normal distributions, truncated at zero, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
@@
         A synthetic record was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively (dots).
-        These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:storm-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
+        These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:case-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
         The blue line shows the true quantiles of the (known) \gls{gev} distribution.
@@
     \caption{
-        \Gls{mcmc} plots for posterior draws from the storm surge model.
+        Markov chain diagnostic plots for posterior draws from the storm surge model.
         We draw \num{10000} samples by running four chains of \num{3500} iterations each and discarding the first \num{1000}.
@@
     \caption{
-        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:storm-surge}).
+        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:case-surge}).
         Each panel shows a different test statistic: partial autocorrelation at lags 1 and 2; sample maximum; sample minimum; sample median; and Mann-Kendall trend test statistic.
@@
         Initial house elevation is fixed to \SI{1}{ft} below the \gls{bfe}.
+        Expectations were computed for discrete values of $\Delta h$ ($x$ axis) by discretizing \glspl{sow} ($y$ axis), then taking the sample mean over each grid cell.
     }\label{fig:scenario-map-height-slr}
@@
     \caption{
-        Diagnostic statistics for the \gls{mcmc} sampling for the storm surge posterior draws.
+        Diagnostic statistics for the Hamiltonian Monte Carlo sampling for the storm surge posterior draws.
         Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.

### 2022-03-21 HouseElevation/src/surge-bayes.jl

@@ -110,9 +110,8 @@ end
 """Save your diagnostics to file"""
 function write_diagnostics(fit, fname::String)
     df = DataFrame(MCMCChains.summarize(fit))
-    rename!(
-        df,
-        "parameters" => L"\textrm{Parameters}",
+    name_replacement = Dict(
+        "parameters" => L"\textrm{Parameter}",
         "mean" => L"\textrm{Mean}",
         "std" => L"\textrm{Stdev.}",
         "naive_se" => L"\textrm{Naive SE}",
@@ -120,6 +119,7 @@ function write_diagnostics(fit, fname::String)
         "ess" => L"\textrm{ESS}",
         "rhat" => L"$\hat{R}$",
     )
+    rename!(df, name_replacement...)
     tex_str = Latexify.latexify(df; env=:table, fmt="%.3f", booktabs=true, index=:subscript)
     open(fname, "w") do io
         write(io, tex_str)

### 2022-03-30 papers/paper/preprint.tex
Update based on Klaus comments

@@
 
-Flood risk management relies on probabilistic estimates of flood hazard:
+Flood risk management relies on probabilistic estimates of flood hazard:\james{Maybe broader, start generally and then give the example to flood risk management}
 \begin{enumerate}
@@
     \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by \gls{fema}.
-At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).
-\citep{kousky_voucher:2014}
-    \item Beyond flood insurance and disaster recovery there has been a federal retreat from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}
+At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).\james{Relate to annual exceedance probability} \citep{kousky_voucher:2014}
+    \item Beyond flood insurance and disaster recovery there has been a federal retreat\james{Word choice heads up} from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}
 \end{enumerate}
-Future flood hazard is deeply uncertain
+Future flood hazard is deeply uncertain\james{Define deep uncertainty}
 \begin{enumerate}
@@
 coastline, public infrastructure and \$1 trillion in national wealth held in coastal real estate are threatened by rising sea levels, higher storm surges, and the ongoing increase in high tide flooding'' \citep{reidmiller_reportinbrief:2018}
-    \item Nonstationarity also affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
-    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}
+    \item Nonstationarity affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
+    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to\james{them's fighting words} identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}
     \item For coastal flooding, \gls{fema} guidance suggests using up to date estimates of \gls{msl} \citep{fema_slr:2016}, but no projections, and stationary estimates of storm surge \citep{fema_ffa:2016}, although the analyst at each location has considerable flexibility
     \item Ditto extreme rainfall \citep{atlas14_texas:2018} or floods \citep{bulletin17c:2019}
-    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}
+    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}\james{This is used below, avoid redundancy}
 \end{enumerate}
@@
     \item deep uncertainty describes ``known unknowns'' where there is not a single objectively true model \citep{walker_deep:2013,Walker:2013gi,lempert_complex:2002}
-    \item This is related to equifinality where different models fit the historical record equally well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
-    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used
+    \item This is related to equifinality where different models fit the historical record equally\james{Add a qualifier} well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
+    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used\james{Ref needed}
     \item \citet{bankes:1993} draws a distinction between exploratory and consolidative modeling
@@
 Typically ``robustness metric'' is used \citep{herman:2015,mcphail_robustness:2019}
-    \item RDM-like methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
+    \item RDM-like\james{Explain what qualifies as RDM and how this relates to MORDM} methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}
     \item For a more complete review see \citet{marchau:2019}
@@
 \begin{enumerate}
-    \item There are always subjective  modeling decisions, but sometimes they are opaque which makes it hard to communicate them to the end user
-    \item The range you choose is also subjective! \citep{schneider_dangerous:2001,schneider_scenarios:2002}
+    \item There are always subjective  modeling decisions, but sometimes these decisions are opaque which makes it hard to communicate them to the end user
+    \item The range you choose is also subjective!
+\citep{schneider_dangerous:2001,schneider_scenarios:2002}\james{Broader Issue: You are either explicitly or implicitly a Bayesian in climate change research}
     \item \citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.'' Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
-    \item Technical detail: joint probability distributions over high dimensional spaces are confusing.
-Bayesian analysis teaches us that posterior distributions of parameters are often highly correlated.
-If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!
+    \item Technical detail: joint probability distributions over high dimensional spaces can be confusing.
+Bayesian analyses show that posterior distributions of parameters are often highly correlated.\james{Refs} If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!
     \item Some outcomes of deeply uncertain processes are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019,srikrishnan_probabilistic:2022} yet this is generally not acknowledged in, \eg, \gls{ipcc} analysis
@@
     \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
-This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critiqued
-    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
+This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critique
+    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citep{definetti_probability:1972}.
 \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
-    \item We know that choosing a single ``best'' model under-estimates uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}
+    \item We know that choosing a single ``best'' model can drastically under-estimate uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}
     \item In the absence of data that would tell us how accurate each model is via some likelihood function, we have to use the prior.
-    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
+    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.\james{Wording}
     \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
@@
 More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
-    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}
+    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}
 \end{enumerate}
@@
 \begin{enumerate}
-    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment could be a problem
+    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment can lead to poor outcomes
     \item Federal guidance uses both standards (floodplain, \gls{bfe}) and cost-benefit analysis (does it pass test?)
+    \item Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
     \item Be sure to note that this has parallels to lots of other engineering design guidance.
@@
     \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
-    \item Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
     \item For example, \citet{xian_elevation:2017} uses a cost-benefit analysis to demonstrate that tailoring recommendations to the initial elevation of a structure can reduce expected costs.
@@
 For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
-While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.\klaus{Is there better mathematical notation for panel (b)?}
+While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
 
@@
     \caption{
-        Outline of the proposed decision analytic framework.
+        Outline of the proposed decision-analytic framework.
+        ADD SOME EXPLANATORY TEXT. What are the boxes and panels. % TODO
     }\label{fig:flowchart}
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.
-Common applications include ``discovering'' particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Brown:2012kb,Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
+Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.\james{May be redundant}
+Common applications include discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.\james{Link back to topic.
+In the topic, there is no link to observations.
+Make this clear and use this as a transition to the next section}
 
@@
 In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
-As we shall illustrate in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only as a.\ldots\klaus{I'm stuck, can you help me finish this sentence?}
+As we shall discuss in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only using a scenario approach (\ie, choices of model structures or factors such as forcings we show as cases) without any explicit probability statement \citep[see][]{wong_nola:2017}.\james{Wording}
 
@@
 \end{equation}
-If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][for some directions]{gelman_workflow:2020}.
+If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to sub-optimal decisions.
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
 
@@
 One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
-In practice this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
-Consequently, climate risk management often uses the outputs of structured simulations and analyses \citep[\eg, from the Coupled Model Intercomparison Project's simulations;][]{meehl_cmip:2000}; it is neither practical nor feasible for local decision-makers to re-run climate models every time they want to sample \glspl{sow} from a different distribution.
+In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
+Consequently, climate risk management often uses the outputs of structured simulations and analyses.
+For example, many studies use outputs from the Coupled Model Intercomparison Project's simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
+It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.
 
@@
 To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
-This allows us to specify $p_\mathrm{belief}(\Psi)$ instead of $p_\mathrm{belief}(\Omega)$.
-We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ such that $\sum_{j=1}^J w_j \psi_j$ approximates $p_\mathrm{belief}(\Psi)$.\klaus{Really we are setting the weights so that $\sum_{j=1}^J w_j \pi(s_j)$ approximates $\int_{s} p(s) \pi(s) \dd{s}$ -- is this too hand-wavey?}
-We present here the 1D case.
-Without loss of generality, assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as illustrated in \cref{fig:grid-sketch}:
+This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
+We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ so that the weighted distribution of $\psi_j$ closely approximates $p_\mathrm{belief}(\Psi)$.
+
+We present here the case where the $\psi_j$ are one-dimensional.
+We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
+Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
 \begin{equation}\label{eq:weights}
@@
 \end{equation}
-For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over the boundaries of each region.
+This step is illustrated in \cref{fig:grid-sketch}.
+For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
+This approach is widely used in analysis of tropical cyclones: first an ensemble of cyclones is simulated, and then probabilistic weights are calculated for each based on attributes such as intensity, landfall location, and radius \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
@@
     \caption{
-        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
-        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the cumulative distribution function of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
+        This method is illustrated for a hypothetical target distribution (orange line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (orange dots).
+        As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
     }\label{fig:grid-sketch}
@@
 The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
-As in \cref{sec:analysis-condition} we must condition on a model, but whereas that analysis conditions upon deep uncertainties, this approach synthesizes them.
+As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
 It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
@@
 
-We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
+We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
+We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
 For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
-To use the notation developed in the previous section, we consider the following.
+We use the notation developed in the previous section to describe the case study.
+Specifically,
 \begin{enumerate}
     \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
-    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T$ year house lifetime ($T=71$ generally; see \cref{tab:uncertainties}) so $\vb{s} \in \mathbb{R}^T$
+    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
     \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
@@
         Summary of parameters, their notation, and how their uncertainty is represented.
+        Symbols describing the decision-analytic framework are described in \cref{fig:flowchart}.
     }\label{tab:uncertainties}
-    \small
-    \begin{tabular}{l l p{3.5in}}
+    \footnotesize
+    \begin{tabular}{p{1.25in} p{0.75in} p{3in}}
         \toprule
@@
         Storm surge          & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
-        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
-        Discount rate    & $1-\gamma$        & Fixed at 2.5\%                                                                       \\
-        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  \\
+        Annual maximum flood & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
+        Discount rate        & $1-\gamma$        & Fixed at 2.5\% per year                                                              \\
+        Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
         Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
-        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 \\
-        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                \\
-        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   \\
-        House lifespan   & $T$               & Deterministic: 71 years unless otherwise noted                                       \\
+        Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
+        House floor area     &                   & Deterministic: \SI{1500}{ft^2}                                                       \\
+        Structural value     &                   & Deterministic: \usd{200000}                                                          \\
+        House lifespan       & $T$               & Deterministic: 71 years                                                              \\
         \bottomrule
     \end{tabular}
+    %TODO: check if there are other symbols used
 \end{table}
@@
     \caption{
-        Conceptual diagram of the decision problem.
-        A \gls{sow} consists of a description of the uncertain factors (red).
+        Conceptual diagram of the considered example.
+        A \acrfull{sow} consists of a description of the uncertain factors (red).
         We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
@@
 These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
-For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, and \citet{bamber_slrise:2019}.
+For a discussion of these model outputs we refer the reader to \citet{ruckert_coastal:2019}.
 
-Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
+Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \citep{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
 To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
@@
 The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
-\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
+\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
 The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
@@
 
-We model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.
+Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\james{Try to figure out a better way to cite this dataset (it's hard)}
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
 We refer this time series of annual maximum storm surges as $y'(t)$.
-
-In the historical record, major surges have been driven by two different mechanisms: tropical cyclones and Nor'easters.
-This time series of annual maxima is shown in \cref{fig:surge-obs-return}(a).
+We display this time series of annual maxima storm surges in \cref{fig:surge-obs-return}(a).
 The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.
@@
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location $\mu$, scale $\sigma$, and shape $\xi$ has the probability density function
-\begin{equation*}
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density functionL
+\begin{equation}\label{eq:gev-dist}
     f(x | \mu, \sigma, \xi)= \begin{cases}
@@
     \end{cases}
-\end{equation*}
+\end{equation}
 This model assumes stationarity, neglecting any potential time dependence.
@@
 We include two forms of prior information.
-First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$:
+First, we constrain the shape parameter to be positive, $\xi > 0$, to reflect knowledge about the support of $y'$, which for a variable distributed according to \cref{eq:gev-dist} is:
 \begin{equation*}
@@
 \end{equation*}
-Since storm surges cannot be negative, only the latter is physically defensible so we constrain $\xi > 0$.
+Since storm surges cannot be negative, only the latter is physically defensible, justifying our choice to constrain the shape parameter to be positive.
 Second, we add weakly informative priors.
@@
 We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
 Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017}.
@@
 \begin{equation}\label{eq:ead}
-    \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'}
+    \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'},
 \end{equation}
@@
 where the discount rate is $1 - \gamma$, the initial time $t_i=2022$, and the end time $t_f = t_i + T$.
-Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this study.
+Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
 For a more theoretical discussion see \citet{arrow_discount:2013}.
@@
 
-We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
+We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.\james{Expand a bit on other work and priors for SLR.}
+Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}, and other studies.
+
 We use a Gamma distribution for all three priors, parameterized as
@@
     \input{../../plots/lsl_priors.tex}
+    % TODO: add another column: Parameter and Quantiles [in feet]
 \end{table}
@@
     \caption{
-        Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of $\Delta h$ and $h_0$.
+        Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
         Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
@@
 
-\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
-The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
-The worst outcomes (blue) arise when the house is elevated only slightly and \gls{slr} is rapid.
+\james{Start with a summary sentence that outlines the paragraph}\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
+The outcomes with lowest total lifetime costs arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
+The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
 As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
@@
 This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
-However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions.
+However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions.\james{Strong statement.
+Correct?
+What if we make or use strong implicit assumptions?}
 
@@
         Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
+        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kolpin_introduction:2017} under \gls{rcp} 8.5.
         (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
@@
 The multiple \gls{pdf} problem has also been shown in other contexts.
-For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{If we think of other examples let's put them here}
+For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{Other examples: Wong et al 2017 -- levees.
+There are more!}
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
@@
 
-To overcome the limitations of scenario-conditional analysis, in this subsection we illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
+The proposed approach can help overcome the limitations of scenario-conditional analysis.
+We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
 We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
@@
 \Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
-Notably, they give different guidance.
+Notably, they give different guidance.\james{Say what this guidance is}
 Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} saves approximately 25\% of the house value relative to not elevating.
@@
 \begin{enumerate}
-    \item We are motivated by the work of \citet{sriver_sealevel:2018} and \citet{lempert_slr:2012}, which points out that while for the particular case study at hand probabilistic and RDM analyses gave similar results (unsurprising since the probabilistic analysis used uniform priors over ranges), the motivation of RDM was an interactive,  problem-focused and iterative exploration with stakeholders.
+    \item We are motivated by the work of \citet{sriver_sealevel:2018} and \citet{lempert_slr:2012}, which points out that while for the particular case study at hand probabilistic and RDM analyses gave similar results (unsurprising since the probabilistic analysis used uniform priors over ranges)\james{Spell out the deficient guidance}, the motivation of RDM was an interactive,  problem-focused and iterative exploration with stakeholders.
 We are motivated to develop a method with more formal probabilistic interpretations that can also be used in this iterative framework.
-    \item Explore every \gls{sow}
+    \item Explore every \gls{sow} considered
     \item Be transparent about how we are calculating trade-offs
@@
     \item It matters a lot which scenario you design for
-    \item Robustness metrics and robust decision analyses make implicit assumptions about the plausibility of future scenarios
+    \item Robustness metrics and robust decision analyses typically make implicit assumptions about the plausibility of future scenarios
 \end{enumerate}
@@
     \item Normal distributions used in \citet{fuller_inversion:2017} were applied for their simplicity and interpretability, but we note that they may not represent a well-calibrated prior belief about \gls{slr}; we refer the reader to \citet{gelman_workflow:2020} for a discussion of prior predictive checks that could be applied.
-    \item Perhaps discarding the edge \gls{ssp} scenarios is unhelpful;
+    \item Perhaps discarding the edge \gls{ssp} scenarios is unhelpful;\james{this is slang}
 \end{enumerate}
@@
 \begin{enumerate}
-    \item Objectives: real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house, and more
+    \item Objectives: real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house, and more objectives
     \item More uncertainty (damage functions, cost of construction, lifespan, discount rate, \etc)
@@
     \item Better models of nonstationary hazards
-    \item Interacting, sequential decisions
+    \item Interacting, sequential decisions\james{Commonly }
     \item Once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.

### 2022-03-30 tikz/bayes-rdm.tex
Update based on Klaus comments

@@
 \definecolor{deepblue}{HTML}{CA3CFF}
+
+% https://tex.stackexchange.com/questions/200838/how-do-i-make-dots-more-compact-or-are-there-alternatives
+\newcommand{\myDots}{\ifmmode\mathinner{\ldotp\kern-0.2em\ldotp\kern-0.2em\ldotp}\else.\kern-0.13em.\kern-0.13em.\fi}
 
@@
     \path [fill=orange!30] (-6.75, 4.8) -- (5.75, 4.8) -- (5.75, -4) -- (-6.75, -4) -- cycle;
-    \node [color=thirds] (s5) at (-4, -3) {\textbf{3) Synthesize}};
+    \node [color=thirds] (s5) at (-4, -3) {\textbf{III. Synthesize}};
 
     \path [fill=rosso!30] (-6.5, 4.6) -- (-6.5, -1.75) -- (5.5, -1.75) -- (5.5, 4.6) -- cycle;
-    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{2) Condition}};
+    \node [rotate=0, color=thirds] (s4) at (4, 3.75) {\textbf{II. Condition}};
 
     \path [fill=thirds!30] (-6, -1.5) -- (2, -1.5) -- (2, 4.25) -- (-2, 4.25) -- (-2, 1.25) -- (-6, 1.25) -- cycle;
-    \node [rotate=45, color=thirds] (s3) at (-1.25, 1.25) {\textbf{1) Explore}};
+    \node [color=thirds] (s3) at (0, 4) {\textbf{I. Explore}};
 
-    \node [block, label={[]135:{\footnotesize (d)}}] at (-4, 3) (models) {
+    \node [block, label={[]225:{\footnotesize (d)}}] at (-4, 3) (models) {
             Multiple models\\
-            for~$p(\Omega | M_k)$:\\
-            $\qty{M_1, \ldots, M_K}$
+            $\vb{M} = \qty{M_1, \ldots, M_K}$\\
+            for $p(\Omega | M_k \in \mathcal{M})$
         };
-    \node [block, label={[]135:{\footnotesize (a)}}] (decisions) at (0, 3) {
+    \node [block, label={[]225:{\footnotesize (a)}}] (decisions) at (0, 3) {
             Candidate decisions\\
@@
         };
-    \node [block, label={[]135:{\footnotesize (b)}}] at (-4, 0) (simulations) {
+    \node [block, label={[]225:{\footnotesize (b)}}] at (-4, 0) (simulations) {
             Many possible futures\\
-            $\vb{s} \in \Omega = \qty{s_1, \ldots, s_J}$
+            $\vb{s}=\qty{s_1, \ldots, s_J}$\\
+            where $s_j \in \Omega$
         };
@@
 
-    \node [derived, label={[]90:{\footnotesize (f)}}] (weights) at (0, -3) {
+    \node [derived, label={[]0:{\footnotesize (f)}}] (weights) at (0, -3) {
             Weights\\
-            $\vb{w} = \qty{w_1, \ldots, w_J}$
+            $\vb{w} = \qty{w_1, \ldots, w_J}$\\
+            where $w_j \in (0, 1)$
         };
@@
 
-    \node [derived, label={[]135:{\footnotesize (e)}}] (combined) at (4, 0) {
+    \node [derived, label={[]270:{\footnotesize (e)}}] (combined) at (4, 0) {
             Conditional\\

### 2022-04-17 papers/paper/preprint.tex
get a draft of the intro together!

@@
 \newacronym{bfe}{BFE}{base flood elevation}
+\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
 \newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
@@
 \newacronym{ssp}{SSP}{shared socio-economic pathway}
+\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
+\newacronym[]{usgs}{USGS}{United States Geological Survey}
 \newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
@@
 
-Flood risk management relies on probabilistic estimates of flood hazard:\james{Maybe broader, start generally and then give the example to flood risk management}
-\begin{enumerate}
-    \item Formal flood risk management in the United States is implemented through a mix of floodplain management policies, design standards, and insurance programs.
-    \item Local governments can build infrastructure to reduce flood risk; they often prioritize these by considering a reduction in damages from a design storm (\eg, the 100 year storm) \citep{hcfcd_prioritization:2019}
-    \item There is a role for industry organizations that set standards (\eg, reliability) that local governments often incorporate into building codes \citep{asce_infrastructure_climate:2021}
-    \item The National Flood Insurance Program was designed as a partnership between the federal government and local communities; communities can voluntarily join the program by adopting a floodplain ordinance based on the most up-to-date flood hazard maps provided by \gls{fema}.
-At a minimum, communities that choose to join must require that new development and substantially improved or damaged properties in high hazard areas be built at or above the level of a flood with a return period of one in 100 years (i.e., the 100-year flood).\james{Relate to annual exceedance probability} \citep{kousky_voucher:2014}
-    \item Beyond flood insurance and disaster recovery there has been a federal retreat\james{Word choice heads up} from flood risk management leaving local governments and individuals with most of the responsibility for flood risk management: ``the federal government recognizes its risks as a major owner and operator of infrastructure, facilities, \ldots, and insurer of properties and crops \ldots leaving states, localities, and individual agencies to make their own plans'' \citep{shi_transformative:2021}
-\end{enumerate}
-Future flood hazard is deeply uncertain\james{Define deep uncertainty}
-\begin{enumerate}
-    \item ``Along the U.S.
-coastline, public infrastructure and \$1 trillion in national wealth held in coastal real estate are threatened by rising sea levels, higher storm surges, and the ongoing increase in high tide flooding'' \citep{reidmiller_reportinbrief:2018}
-    \item Nonstationarity affects inland and pluvial flooding \citep{Milly:2008dg} and derives from a combination of climate change, land use change, and local river modifications \citep{Merz:2014gf}
-    \item Federal guidelines for flood risk management that directly inform policy-making recognize the presence of nonstationarity but have failed to\james{them's fighting words} identify objectively valid approaches for modeling nonstationarity, \citep{Montanari:2014hl,Serinaldi:2015bq}
-    \item For coastal flooding, \gls{fema} guidance suggests using up to date estimates of \gls{msl} \citep{fema_slr:2016}, but no projections, and stationary estimates of storm surge \citep{fema_ffa:2016}, although the analyst at each location has considerable flexibility
-    \item Ditto extreme rainfall \citep{atlas14_texas:2018} or floods \citep{bulletin17c:2019}
-    \item This is because where historic flood risk can be constrained by estimates, future flood risk depends on intrinsically unpredictable human decisions like greenhouse gas emissions, and is therefore deeply uncertain \citep{keller_management:2021}\james{This is used below, avoid redundancy}
-\end{enumerate}
-Methods for decision making under deep uncertainty emphasize exploring many possible futures
-\begin{enumerate}
-    \item deep uncertainty describes ``known unknowns'' where there is not a single objectively true model \citep{walker_deep:2013,Walker:2013gi,lempert_complex:2002}
-    \item This is related to equifinality where different models fit the historical record equally\james{Add a qualifier} well but diverge as one looks into the future \citep{beven_equifinality:2006,DossGollin:2019}
-    \item Applying classical optimization methods to problems with deep uncertainty can yield highly fragile solutions in the sense that guidance might change substantially if a different plausible model were used\james{Ref needed}
-    \item \citet{bankes:1993} draws a distinction between exploratory and consolidative modeling
-    \item Decision scaling \citep{Steinschneider:2015kk}
-    \item global sensitivity analysis \citep{saltelli_sensitivity:2010,herman_salib:2017,sobol_sensitivity:2001} feeds large scenario ensembles into a black box model, then analyzes outputs to assess the relative importance of each input (conditional on an assumed joint distribution over inputs)
-    \item MORDM \citep{kasprzyk:2013} combines exploratory modeling (exploring performance over many scenarios) and policy search (using optimization tools)  \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-Typically ``robustness metric'' is used \citep{herman:2015,mcphail_robustness:2019}
-    \item RDM-like\james{Explain what qualifies as RDM and how this relates to MORDM} methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that meet some predetermined satisficing criteria \citep{mcphail_robustness:2019} over a large fraction of possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}
-    \item For a more complete review see \citet{marchau:2019}
-    \item Crucially, these approaches all emphasize exploring performance over many scenarios, then describing this performance into some set of metrics.
-\end{enumerate}
-Some critiques
-\begin{enumerate}
-    \item There are always subjective  modeling decisions, but sometimes these decisions are opaque which makes it hard to communicate them to the end user
-    \item The range you choose is also subjective!
-\citep{schneider_dangerous:2001,schneider_scenarios:2002}\james{Broader Issue: You are either explicitly or implicitly a Bayesian in climate change research}
-    \item \citet{schneider_scenarios:2002} criticized the reluctance of the \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, writing ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.'' Different end users will, in general, have different preferences that motivate different design choices given the same information, it is unlikely that these end users have expertise in global emissions pathways, climate sensitivity, or ice sheet dynamics.
-    \item Technical detail: joint probability distributions over high dimensional spaces can be confusing.
-Bayesian analyses show that posterior distributions of parameters are often highly correlated.\james{Refs} If we neglect that by sampling over the marginal distribution of each parameter separately, we may be sampling almost entirely from regions of the parameter space that are not consistent with our data \citep[see][for a detailed illustration]{carpenter_dimensionality:2017}!
-    \item Some outcomes of deeply uncertain processes are more likely than others \citep{hausfather_scenarios:2020,ho_scenarios:2019,srikrishnan_probabilistic:2022} yet this is generally not acknowledged in, \eg, \gls{ipcc} analysis
-    \item Technical detail: sampling over ranges is not invariant to parameterization: if I sample $\theta \sim \mathrm{Uniform}(a, b)$ and then resample $\log theta \sim \mathrm{Uniform}(\log a, \log b)$, I'll get different samples and thus different estimates
-\end{enumerate}
-The question of whether probability is a useful tool for decision making under deep uncertainty depends on how one interprets probability; we argue that a subjective Bayesian interpretation of probability is useful for many DMDU problems
-\begin{enumerate}
-    \item We are inspired by methods and philosophy of Bayesian model choice in the ``$\mathcal{M}$-open'' case where there is no true model to identify
-    \item Many modern discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.
-This is in many ways closer to the exploratory modeling proposed by \citet{bankes:1993} than the top-down, deterministic, fragile central planning approaches that it (and \cite{rittel:1973} and others) critique
-    \item Bayesian decision theory was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citep{definetti_probability:1972}.
-\citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
-    \item We know that choosing a single ``best'' model can drastically under-estimate uncertainty so inferences need to be combined across models \citep[\eg, as in stacking;][]{Yao:2018bu}
-    \item In the absence of data that would tell us how accurate each model is via some likelihood function, we have to use the prior.
-    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.\james{Wording}
-    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.
-    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.
-    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
-More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
-    \item Since, by definition, models are not ``true,'' this iterative workflow \citep{gelman_workflow:2020} aims to identify models that are useful and promote a dialog amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}
-\end{enumerate}
-Our aim is to illustrate how methods for model selection in the $\mathcal{M}$-open case can improve decision making under deep uncertainty.
-\begin{enumerate}
-    \item We want to combine multiple models in a way that is transparent, and that can be changed with low computational cost (cheap robustness checks).
-    \item We are particularly motivated by problems for which simulations of some key uncertainties (particularly climate-driven ones but also population, economic growth, etc) already exist.
-We assume only that there is no coupling between the system model and these simulations, so they can be treated as fully exogenous.
-\end{enumerate}
-We illustrate our approach with a didactic case study\ldots.
-\begin{enumerate}
-    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment can lead to poor outcomes
-    \item Federal guidance uses both standards (floodplain, \gls{bfe}) and cost-benefit analysis (does it pass test?)
-    \item Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-    \item Be sure to note that this has parallels to lots of other engineering design guidance.
-    \item \citet{xian_elevation:2017}: one size does not fit all
-    \item \citet{zarekarizi_suboptimal:2020}: neglecting uncertainty leads to bad outcomes
-    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
-    \item For example, \citet{xian_elevation:2017} uses a cost-benefit analysis to demonstrate that tailoring recommendations to the initial elevation of a structure can reduce expected costs.
-          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
-    \item Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
-    \item FEMA's Hazard Mitigation Grant Program, authorized under Section 404 of the 1988 Stafford Act, which awards grants after a disaster.
-Under the Stafford Act, projects funded by HMGP, including buyouts, must be shown to be cost-effective; typcally, up-front costs are compared to avoided losses \citep{bendor_buyouts:2020}
-\end{enumerate}
-Our primary research question is ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
-We proceed as follows\ldots
+Climate risk management, as applied in the United States and around the world, relies heavily upon cost benefit analyses and engineering standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
+For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.\james{cite}
+Anyone purchasing a home inside the floodplain using a federally backed mortgage is nominally required to purchase flood insurance \citep{kousky_voucher:2014}.
+Additionally, local building regulations \citep[based on guidance such as][]{asce_24-05:2006}, and the regulations of major institutions involved in flood risk management \citep[\eg,][]{FEMA_p-55:2011} recommends all houses in particular flood zones to be elevated some margin above a nominal \gls{bfe}.
+Homeowners facing high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.\james{cite}
+Similar methods also apply to the design of massive infrastructure projects.
+For example, the design height of a seawall, the key component of a \$29 billion coastal protection project proposed by the \gls{usace} for  Galveston Bay, was selected by setting the annual probability that overtopping exceeds a safe threshold to 1\%, neglecting the potential for sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+And while many stakeholder objectives contributed to proposed design, the 1\% annual exceedance probability threshold was selected in large part because the benfit to cost ratio of this design was higher than alternatives considered \citep[\S 2.1.3]{USACE_coastal:2021}.
+A key commonality between these engineering criteria based on design events and these project comparison tools based on cost-benefit analysis is a reliance on probabilistic descriptions of relevant hazard.
+
+Where climate risk management is directly informed by design events (\ie, quantiles of a distribution) or cost-benefit analysis (\ie, integrating over a distribution), the estimated \gls{pdf} will strongly inform the resulting design.
+Current methods focus on data-driven, objective frameworks that can be applied consistently across locations.
+For example, \gls{usgs} Bulletin 17C specifies the procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity duration and frequency of extreme rainfall.
+One of the key assumptions underlying most of these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
+Yet changes in both global climate and local environmental conditions challenge this assumption in many cases \citep{Merz:2014gf,Milly:2008dg}.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
+
+More fundamentally, while models for estimating and predicting specific forms of nonstationarity have been proposed \citep[see][for a review]{Salas:2018ge}, many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
+Recognizing the challenges of decision making under deep uncertainty, many frameworks for identifying robust decisions have been proposed.
+Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
+For example, robust decision methods \citep{lempert_shaping:2003} run models over large ensembles of possible futures to assess the performance of different policies, then apply statistical analysis to identify the conditions under which particular policies perform well or poorly.
+Such methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
+When many possible levers are available, and exhausting sampling therefore infeasible, multiobjective robust decision methods \citep{kasprzyk:2013} add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+Many studies formally quantify robustness \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for multiobjective comparison of candidate policies.
+
+Although these methods for decision making under deep uncertainty have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
+One common approach is to systematically sample deeply uncertain parameters from a specified range, then feed these parameters into a system model in order to quantify some metric of aggregate performance \citep{herman:2015,mcphail_robustness:2019,hall_comparison:2012,lempert:2007}.
+A challenge here is that assumptions about the range and independence of uncertain factors can strongly influence which factors are found to be most important and which policies are most robust \citep{quinn_exploratory:2020}.
+Aggregating measures of performance across uncertain factors (\eg, to quantify expectations, quantiles, or robustness metrics of particular decisions) increases sensitivity to these implicit assumptions.
+These assumptions are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002} and, indeed, can be interpreted  as a specific choice of probability distribution (\ie, uniform distribution with fixed upper and lower bound).
+
+\textbf{Ultimately, deeply uncertain nonstationary hazard challenges not only the existing stationary estimates of hazard but more fundamentally the premise that objective estimates of future hazard exist and can be estimated empirically}.
+Though seeking robustness can add value to many decision analyses, sensitivity to model assumptions about the likelihood and plausibility of different futures cannot be avoided.
+Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
+This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a family of candidate models, and we find many parallels to deep uncertainty.
+The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, fr which probability distributions offer a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a discussion of methods for model selection).
+From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
+Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}
+
+The conceptual similarities between this statistical philosophy and the literature on decision making under deep uncertainty lead us to reject the premise, articulated in some studies,\klaus{is it ok to not cite any here} that probability should not be used to inform decision making under deep uncertainty.
+Our subjectivist perspective has a rich history in the Earth sciences.
+For example, \citet{schneider_scenarios:2002} criticizes the reluctance of an \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, arguing that ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
+More specifically,
+When even sophisticated \citep[or so we like to think; see][and references therein]{fuller_inversion:2017} consumers of climate information incorporate published plausible ranges into adaptation planning, information that experts may hold the shape of particular distributions can be lost \citep[as warned in][]{moss_uncertainties:2000}.
+
+In this paper we a first conceptual step towards bridging the fields of decision making under deep uncertainty and subjective Bayesian philosophy through  a didactic case study.
+We consider the problem studied in \citet{zarekarizi_suboptimal:2020} of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
+Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
+Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
+Prior research has shown that this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}, and that neglecting uncertainty can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
+Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
+
+We proceed as follows\ldots\james{Revisit}
+In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-explore}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
+In \cref{sec:case-study} we describe the case study in detail.
+Next in \cref{sec:results} we discuss results for each of the three theoretical lenses.
+Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.
 
@@
         Outline of the proposed decision-analytic framework.
-        ADD SOME EXPLANATORY TEXT. What are the boxes and panels. % TODO
+        In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance (c) of candidate decisions (a) under a large ensemble of possible futures (b).
+        In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes (e) that are conditional upon specific models describing the likelihood of different futures (d).
+        In our case study, these are \gls{rcp} scenarios and physical models of sea level rise.
+        Finally in \cref{sec:analysis-synthesize} we propose a subjective Bayesian framework for synthesizing across deep uncertainties by re-weighting sampled futures (f).
     }\label{fig:flowchart}
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption  about the likelihood different \glspl{sow} and instead seeks to generate new knowledge through existence proofs, hypothesis generation, and vulnerability identification \citep{bankes:1993}.\james{May be redundant}
-Common applications include discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.\james{Link back to topic.
-In the topic, there is no link to observations.
-Make this clear and use this as a transition to the next section}
+Exploratory modeling avoids making explicit assumption about the likelihood different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
+\klaus{Does this address your comment about linking back to observations?}
 
@@
 Consequently, climate risk management often uses the outputs of structured simulations and analyses.
-For example, many studies use outputs from the Coupled Model Intercomparison Project's simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
+For example, many studies use outputs from \gls{cmip} simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
 It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.
@@
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\james{Try to figure out a better way to cite this dataset (it's hard)}
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\klaus{I can't find a better reference for this dataset!}
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density functionL
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function:
 \begin{equation}\label{eq:gev-dist}
@@
     \item More uncertainty (damage functions, cost of construction, lifespan, discount rate, \etc)
+    \item Better damage functions \citep{nofal_fragility:2020,Rozer:2019}
     \item Better data on cost of elevation and depth-damage
@@
 Another clarification about how to interpret the $p_\mathrm{belief}$
-\begin{enumerate}
-    \item As statistitian David Spiegelhalter discusses, this subjective belief should not be interpreted as a population distribution because there is no true population distribution that could be answered by collecting more data \citep{andorra_spiegelhalter:2021}.\james{Contextualize this}
-    \item
-\end{enumerate}
 This approach can be applied to lots of important problems

### 2022-04-17 scripts/01-surge-modeling.jl

@@ -22,7 +22,7 @@ function plot_surge_gev_priors()
     # loop through all the distributions
     for (prior, c) in zip(HouseElevation.gev_priors, colors)
         plot!(
-            p, 0:0.05:30, prior.dist; label="$(prior.rt) Year Surge", linewidth=3, color=c
+            p, 0:0.05:40, prior.dist; label="$(prior.rt) Year Surge", linewidth=3, color=c
         )
     end
     savefig(p, plots_dir("surge-gev-priors.pdf"))
@@ -229,6 +229,8 @@ function plot_surge_posterior_return(
 
     # add gray lines
     p = plot_return_period(post_gevs)
+    y_ticks = [5, 7.5, 10, 12.5, 15]
+    plot!(p; yaxis=:log, yticks=(y_ticks, string.(y_ticks)))
 
     # add weibul plot pls
     xp, ys = HouseElevation.weibull_plot_pos(surge_ft)
@@ -259,13 +261,17 @@ function plot_surge_obs_return(
     xlabel!(p1, "Time [year]")
     p2 = plot_surge_posterior_return(annual, fits)
     p2 = plot(p2; ylabel="")
-    add_panel_letters!([p1]; fontsize=12)
-    annotate!(p2, [1], [8.5], text("(B)", :left, 14))
+
+    y_ticks = collect(3:10)
+    p1 = plot!(p1; yaxis=:log, yticks=(y_ticks, string.(y_ticks)))
+    p2 = plot!(p2; yaxis=:log, yticks=(y_ticks, string.(y_ticks)))
+
+    add_panel_letters!([p1, p2]; fontsize=12, loc=(0.05, 0.95))
     p = plot(
         p1,
         p2;
         link=:y,
-        ylims=(2, 9),
+        ylims=(2.5, 10),
         layout=grid(1, 2; widths=[0.6, 0.35]),
         size=[1000, 350],
         bottommargin=8mm,

### 2022-04-17 scripts/06-combining.jl

@@ -8,7 +8,8 @@ function plot_grid_scheme()
     dist = Distributions.Normal()
     ψ = quantile.(dist, [0.05, 0.3, 0.7, 0.95])
     ψ_labels = [L"$\psi_%$i$" for i in 1:length(ψ)]
-    xlimits = (-3, 3)
+    xlimits = (-3, 2.5)
+    y_ticks = [0, 0.25, 0.5, 0.75, 1]
 
     ymax = pdf(dist, 0)
     grids = [0.5 * (ψ[j] + ψ[j + 1]) for j in 1:(length(ψ) - 1)]
@@ -27,8 +28,9 @@ function plot_grid_scheme()
         bottom_margin=5Plots.mm,
         left_margin=5Plots.mm,
         size=(800, 350),
+        yticks=(y_ticks, string.(y_ticks)),
     )
-    scatter!(p, ψ, zeros(size(ψ)); label="Observations", color=colors[2], markersize=5)
+    scatter!(p, ψ, zeros(size(ψ)); label="Sampled Points", color=colors[2], markersize=5)
 
     ymax_prev = 0.0
     x_arrow = range(minimum(xlimits), 0; length=length(ψ) + 2)[2:(end - 1)]
@@ -180,6 +182,7 @@ function plot_prior_tradeoffs(
         )
 
         # plot all trade-off lines in light gray
+        #=
         for rcp in all_rcp
             for model in all_models
                 w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
@@ -187,11 +190,10 @@ function plot_prior_tradeoffs(
                 plot!(p, Δh_ft, cond; label="", color=:gray, linewidth=0.5, alpha=0.5)
             end
         end
-
-        # plot the conditional beliefs
+        =#
 
         # plot the trade-off for the priors
-        for (prior, color) in zip(priors, colors)
+        for (prior, color) in zip(priors, [:red, :blue, :green])
             w = HouseElevation.make_weights(prior.dist)
             cond = vec(mean(var, w; dims=1))
             plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=3)

### 2022-04-17 scripts/plotutils.jl

@@ -31,7 +31,7 @@ function twiny(sp::Plots.Subplot)
 end
 twiny(plt::Plots.Plot=current()) = twiny(plt[1])
 
-pct_formatter(yi) = "$(Int.(round.(yi * 100)))%"
+pct_formatter(yi) = "$(Int(round(yi * 100))) %"
 blank_formatter(yi) = ""
 
 function get_roma_colormap()
@@ -46,40 +46,42 @@ end
 
 get_colormap() = get_roma_colormap()
 
-function plot_return_period(gevs::Vector{<:Distributions.GeneralizedExtremeValue}) where {T<:HouseElevation.MCMCChains.Chains}
-    
-        rts = range(1.25, 275; length=250) # return periods
-        aeps = 1 .- 1 ./ rts # annual exceedance probability
-        xticks = [2, 5, 10, 25, 50, 100, 250]
-        ranges = [0.95, 0.80, 0.5]
+function plot_return_period(
+    gevs::Vector{<:Distributions.GeneralizedExtremeValue}
+) where {T<:HouseElevation.MCMCChains.Chains}
+    rts = range(1.25, 500; length=250) # return periods
+    aeps = 1 .- 1 ./ rts # annual exceedance probability
+    xticks = [2, 5, 10, 25, 50, 100, 250, 500]
+    ranges = [0.95, 0.80, 0.5]
 
-        p = plot(;
-            xlabel="Return Period [years]",
-            ylabel="Return Level [ft]",
-            xscale=:log,
-            legend=:bottomright,
-            xticks=(xticks, string.(xticks)),
+    p = plot(;
+        xlabel="Return Period [years]",
+        ylabel="Return Level [ft]",
+        xscale=:log,
+        legend=:bottomright,
+        xticks=(xticks, string.(xticks)),
+    )
+
+    for range in ranges
+        qup = 1 - (1 - range) / 2
+        qlow = (1 - range) / 2
+        ub = [quantile([quantile(d, xi) for d in gevs], qup) for xi in aeps]
+        lb = [quantile([quantile(d, xi) for d in gevs], qlow) for xi in aeps]
+        range_pct = Int(range * 100)
+        plot!(
+            p,
+            rts,
+            ub;
+            fillbetween=lb,
+            fillcolor=:gray,
+            fillalpha=0.35,
+            linecolor=false,
+            label="$(range_pct)% Credible Interval",
         )
-
-        for range in ranges
-            qup = 1 - (1 - range) / 2
-            qlow = (1 - range) / 2
[…]

### 2022-04-21 papers/paper/preprint.tex
GEtting close, just need a bit of work on abstract / intro

@@
 \begin{abstract}
-    Climate risk management involves comparing candidate decisions under deep and dynamic uncertainties\ldots
+    Methods for doing so, whether classical standards-based approaches or bottom-up robust methods, make implicit or explicit assumptions about the likelihood of different futures.
+    These assumptions may be inconsistent with expert belief.
+    In this paper we propose a probabilsitic method for synthesizing across deep uncertainties, using subjective prior beliefs.
+    The proposed method can be used in tandem with bottom-up exploratory methods and is consistent with fully probabilistic design criteria.
+    Through the lens of a didactic case study of house elevation in Norfolk, VA, we illustrate how this approach can improve the transparency of modeling assumptions.\klaus{Abstract needs a little work}
 \end{abstract}
@@
 \begin{enumerate}
-    \item
+    \item Estimated trade-offs between adaptation objectives are sensitive to assumptions about the likelihood of future conditions.
+    \item We propose a probabilistic method for transparently synthesizing across deep uncertainty in future conditions.
+    \item The proposed method identifies strategies for house elevation that are robust to assumptions in sea level rise.
 \end{enumerate}
@@
 Climate risk management, as applied in the United States and around the world, relies heavily upon cost benefit analyses and engineering standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
-For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.\james{cite}
-Anyone purchasing a home inside the floodplain using a federally backed mortgage is nominally required to purchase flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building regulations \citep[based on guidance such as][]{asce_24-05:2006}, and the regulations of major institutions involved in flood risk management \citep[\eg,][]{FEMA_p-55:2011} recommends all houses in particular flood zones to be elevated some margin above a nominal \gls{bfe}.
-Homeowners facing high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.\james{cite}
-Similar methods also apply to the design of massive infrastructure projects.
-For example, the design height of a seawall, the key component of a \$29 billion coastal protection project proposed by the \gls{usace} for  Galveston Bay, was selected by setting the annual probability that overtopping exceeds a safe threshold to 1\%, neglecting the potential for sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+
+For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
+One consequence is that those purchasing a home inside the designated floodplain, if using federally backed mortgage porgrams, are nominally required to purchase flood insurance \citep{kousky_voucher:2014}.
+Additionally, local building regulations \citep[based on guidance such as][]{asce_24-05:2006}, and the regulations of major institutions involved in flood risk management \citep[\eg,][]{FEMA_p-55:2011} recommend all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
+And homeowners subject high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.
+Similar methods are used to inform the design of massive infrastructure projects.
+For example, the design height of a seawall, the key component of a \$29 billion coastal protection project proposed by the \gls{usace} for  Galveston Bay, was selected by setting the annual probability that overtopping exceeds a safe threshold to 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 And while many stakeholder objectives contributed to proposed design, the 1\% annual exceedance probability threshold was selected in large part because the benfit to cost ratio of this design was higher than alternatives considered \citep[\S 2.1.3]{USACE_coastal:2021}.
@@
 For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
+
+The theory of rational decision making evolved in the middle of the 20th century in tandem with mathematical theories of probability and economics.
+In particular, \citet{vonneumann_games:1944} proposed four axioms for rational choice:
+completeness of preference;
+transitivity of preference;
+continuity of preference, and
+independence of preference from the existence of irrelevant alternatives.
+\citeauthor{vonneumann_games:1944} then proved that an individual or firm that follows these actions behaves as though maximizing the expectation of some function, which they called a utility function.
+Similarly, individuals and firms that do not follow these axioms can be systematically exploited \citep[see the ``Dutch book'' arguments of][]{definetti_probability:1972}.
+
+Bayesian decision theory \citep{savage:1954}
+Over the next decades, the \emph{descriptive} model of rationality used to study firms and individuals \citep[see][]{savage:1954,debreu:1959} became widely applied as a \emph{normative} tool for planning.
+Efforts such as the Harvard Water Program \citep[see][]{maass_hwp:1962} popularized the use of top-down rational planning approaches for water management \citep{howe_bca:1971} and other social planning problems.
+These methods used the vocabulary and mathematical framework of Bayesian decision theory \citep{Gelman:2014tc,savage:1954}, bringing a formal mathematical treatment of these problems.
 
@@
 Aggregating measures of performance across uncertain factors (\eg, to quantify expectations, quantiles, or robustness metrics of particular decisions) increases sensitivity to these implicit assumptions.
-These assumptions are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002} and, indeed, can be interpreted  as a specific choice of probability distribution (\ie, uniform distribution with fixed upper and lower bound).
+These assumptions are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
 
-\textbf{Ultimately, deeply uncertain nonstationary hazard challenges not only the existing stationary estimates of hazard but more fundamentally the premise that objective estimates of future hazard exist and can be estimated empirically}.
+\textbf{Ultimately, recognition of deeply uncertain nonstationary hazard challenges not only the existing stationary estimates of hazard but more fundamentally the premise that objective estimates of future hazard exist and can be estimated empirically}.
 Though seeking robustness can add value to many decision analyses, sensitivity to model assumptions about the likelihood and plausibility of different futures cannot be avoided.
@@
 Prior research has shown that this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}, and that neglecting uncertainty can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
+Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
 
-We proceed as follows\ldots\james{Revisit}
+We proceed as follows.
 In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-explore}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
@@
         Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
-        House floor area     &                   & Deterministic: \SI{1500}{ft^2}                                                       \\
-        Structural value     &                   & Deterministic: \usd{200000}                                                          \\
+        House floor area     & --                & Deterministic: \SI{1500}{ft^2}                                                       \\
+        Structural value     & --                & Deterministic: \usd{200000}                                                          \\
         House lifespan       & $T$               & Deterministic: 71 years                                                              \\
@@
         We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
-        For each \acrshort{sow} and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (orange).
+        For each \acrshort{sow} (red) and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (gray).
     }\label{fig:xlrm}
@@
 We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.\james{Expand a bit on other work and priors for SLR.}
-Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}, and other studies.
+Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.
 
@@
 
-Quick summary of what our motivation was
+The motivation for this study was to develop a framework for increasing the transparency of assumptions about the likelihood of different \glspl{sow}.
+Additionally, we sought to develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
+Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
+In this paper we proposed a decision analytic framework for synthesizing across deep uncertainties that is compatible with exploratory modeling.
+Through a case study of house elevation in the coastal zone, we show that:\klaus{I'd actually love feedback on these -- they can more or less parallel the key points but it would be really helpful for me to see what you think the key points are}
 \begin{enumerate}
-    \item We are motivated by the work of \citet{sriver_sealevel:2018} and \citet{lempert_slr:2012}, which points out that while for the particular case study at hand probabilistic and RDM analyses gave similar results (unsurprising since the probabilistic analysis used uniform priors over ranges)\james{Spell out the deficient guidance}, the motivation of RDM was an interactive,  problem-focused and iterative exploration with stakeholders.
-We are motivated to develop a method with more formal probabilistic interpretations that can also be used in this iterative framework.
-    \item Explore every \gls{sow} considered
-    \item Be transparent about how we are calculating trade-offs
-    \item Work with existing physical model ensembles
-    \item We want to be robust to different plausible beliefs rather than robust to different outcomes
+    \item estimated design trade-offs are highly sensitive to the choice of scenario;\klaus{This seems self-evident, though}
+    \item implicit or explicit assumptions about the likelihood of different scenarios are inevitable; and
+    \item the method proposed in \cref{sec:analysis-synthesize} can transparently synthesize across deep uncertainties.
 \end{enumerate}
-Quick summary of what we have done
-\begin{enumerate}
-    \item It matters a lot which scenario you design for
-    \item Robustness metrics and robust decision analyses typically make implicit assumptions about the plausibility of future scenarios
-\end{enumerate}
-Limitations of the re-weighting
-\begin{enumerate}
-    \item Our prior $p(\Omega)$ is limited -- we are just using \gls{msl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
-    \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
-    \item Normal distributions used in \citet{fuller_inversion:2017} were applied for their simplicity and interpretability, but we note that they may not represent a well-calibrated prior belief about \gls{slr}; we refer the reader to \citet{gelman_workflow:2020} for a discussion of prior predictive checks that could be applied.
-    \item Perhaps discarding the edge \gls{ssp} scenarios is unhelpful;\james{this is slang}
-\end{enumerate}
-Limitations of the case study
-\begin{enumerate}
-    \item Objectives: real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house, and more objectives
-    \item More uncertainty (damage functions, cost of construction, lifespan, discount rate, \etc)
-    \item Better damage functions \citep{nofal_fragility:2020,Rozer:2019}
-    \item Better data on cost of elevation and depth-damage
-    \item Scalability: optimize separately for different kinds of house structures and locations
-    \item Timing
-    \item More physics -- we neglect hydrodynamics and tides
-\end{enumerate}
-Future research needs:
-\begin{enumerate}
-    \item This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
-    \item More complex models to capture more relevant metrics
-    \item Better models of nonstationary hazards
-    \item Interacting, sequential decisions\james{Commonly }
-    \item Once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.
-We hypothesize this will lead to under-valuation of adaptive measures.
-    \item Ultimately, developing probabilistic scenarios of future emissions \citep[as in][]{srikrishnan_probabilistic:2022}, temperature, and sea level rise would be valuable.
-\end{enumerate}
-Another clarification about how to interpret the $p_\mathrm{belief}$
-This approach can be applied to lots of important problems
-\begin{enumerate}
-    \item Stormwater management \citep{sharma_rcp:2021,lopez-cantu:2018}
-    \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}
-    \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!\james{Let's emphasize this -- really important to policy analysis under deep uncertainty}
-    \item Model structure uncertainties
-    \item Local government revenues are sensitive to flood losses; this approach can be used to structure updates to building codes \citep{shi_municipal:2020}
-\end{enumerate}
-We call for researchers studying climate risk management to make their implicit probability distributions more transparent, and we call for coordinated national and international guidance to assist local governments and engineering bodies determine appropriate subjective beliefs.
+
+There are several limitations to our study that merit further discussion here.
+The first category has to do with limitations of the underlying methodology proposed for reweighting state of the world.
+For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
+Although this is a low-dimensional projection of the full time series, it is not a sufficient statistic.
+It is known that in approximate Bayesian computation, calibrating models on low dimensional statistics that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
+Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences.
+For example, \citet{lempert_slr:2012} considers a five parameter model for sea level rise that takes into account rates of change for different physical mechanisms; it could make sense to put a prior over this higher dimensional space.
+A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent the .
+Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
+Finally, and more fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
+An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
+Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.\klaus{Is this too vague?}
+
+The second category of limitations has to do with the case study and our interpretation of the house elevation decision problem.
+This problem intersects with decisions about where to live and how to manage household finances, both of which are highly complex.
+We could extend our analysis, for example, by considering additional decision objectives representing risk aversion, usable space created under the house, non-financial consequences of flooding, or other factors.
+We could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
+Finally, while here we consider the decision to be a one-time decision, it can also be framed as a sequential decision.
+Analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions.
+Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \citep{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \citep{herman_control:2020}.
+
+These limitations motivate several directions for future research.
+From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \citep[see fig.~1 of][]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
+From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
+Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that scale to different house characteristics may add value.
+
+The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate adaptation.
+For example, much of the stormwater infrastructure in the United States is inadequate for current climate \citep{lopez-cantu:2018}, let alone future changes.
+Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \citep{sharma_rcp:2021} and \gls{rcp} scenarios.
+Another example is the design and operation of coastal infrastructure.
+Problems such as levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
+
+Ultimately, the proposed framework rests on the argument that if you cannot be right, it is best to be transparent.
+We call for researchers studying climate risk management to make their implicit assumptions about  distributions more transparent, and we call for coordinated national and international guidance to assist local governments and engineering bodies determine appropriate subjective beliefs.\klaus{OK with this?}
 
@@
         See \cref{sec:case-surge} for details.
-        For the 2, 10, 100, and 500 year events we apply Normal distributions, truncated at zero, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+        For the 2, 10, 100, and 500 year events we apply Inverse Gamma distributions, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
     }\label{fig:surge-gev-priors}

### 2022-04-21 HouseElevation/src/surge-bayes.jl

@@ -9,21 +9,32 @@ using MCMCChainsStorage
 using Optim
 using StatsBase
 using Turing
+using Distributions: InverseGamma
+
+"""
+This function converts a Normal distribution to an InverseGamma distribution,
+preserving the mean and variance analytically.
+"""
+function InverseGamma(d::Distributions.Normal)
+    α = 2 + d.μ^2 / d.σ^2
+    β = (α - 1) * d.μ
+    return Distributions.InverseGamma(α, β)
+end
 
 # stationary GEV model
 gev_priors = [
-    (rt=2, dist=Distributions.TruncatedNormal(6, 3, 0, Inf)),
-    (rt=10, dist=Distributions.TruncatedNormal(8, 4, 0, Inf)),
-    (rt=100, dist=Distributions.TruncatedNormal(12, 6, 0, Inf)),
-    (rt=500, dist=Distributions.TruncatedNormal(16, 8, 0, Inf)),
+    (rt=2, dist=InverseGamma(Distributions.Normal(6, 3))),
+    (rt=10, dist=InverseGamma(Distributions.Normal(8, 4))),
+    (rt=100, dist=InverseGamma(Distributions.Normal(12, 6))),
+    (rt=500, dist=InverseGamma(Distributions.Normal(16, 8))),
 ]
 
 @model function StationaryGEV(y)
 
-    # have to put a prior to define parameters in Turing
-    μ ~ Distributions.TruncatedNormal(0, 25, 0, Inf) # lower bound should not be negative
-    σ ~ Distributions.TruncatedNormal(0, 10, 0.01, Inf) # σ > 0 by definition
-    ξ ~ Distributions.TruncatedNormal(0, 2.5, 0, Inf) # surge has lower bound => ξ>0
+    # very weak prior -- need _a_ prior to define variables in Turing
+    μ ~ Distributions.TruncatedNormal(0, 25, 0, Inf)
+    σ ~ Distributions.TruncatedNormal(0, 10, 0.001, Inf)
+    ξ ~ Distributions.TruncatedNormal(0, 2.5, 0, Inf)
 
     dist = GeneralizedExtremeValue(μ, σ, ξ)
 

### 2022-04-21 scripts/03-cost.jl

@@ -88,6 +88,7 @@ function plot_cost_upfront(
         label=false,
         xticks=0:2:14,
         color=colors[1],
+        yformatter=pct_formatter,
     )
     # add in-line label
     annotate!(p, [8], [0.675], text("Cost of Elevating", :center, 10; color=colors[1]))

### 2022-04-21 scripts/04-scenario-maps.jl

@@ -57,7 +57,7 @@ function plot_scenario_map_slr_cost(;
             kwargs = Dict{Symbol,Any}(
                 :label => false,
                 :xformatter => x -> "",
-                :bins => (LinRange(0, maximum(msl_rise_ft), 60), LinRange(0, 4.5, 60)),
+                :bins => (LinRange(0, maximum(msl_rise_ft), 40), LinRange(0, 4.5, 40)),
                 :c => cmap,
                 :clims => clims,
                 :colorbar => false,
@@ -117,7 +117,7 @@ function plot_scenario_map_slr_cost(;
         xlims=(1, 1.1),
         xshowaxis=false,
         yshowaxis=false,
-        colorbar_title="Density of SOWs",
+        colorbar_title="Count of SOWs",
         colorbar_titlefontrotation=90,
         grid=false,
         clims=clims,
@@ -177,7 +177,7 @@ function plot_scenario_map_slr_cost(;
             cbar_fake_plot,
         )...;
         layout=l,
-        size=(1500, 1000),
+        size=(1200, 750),
         dpi=250,
         link=:y,
     )
@@ -222,20 +222,24 @@ function plot_scenario_map_height_slr(;
     )
 
     x_ft = ustrip.(u"ft", x)
+
     p = heatmap(
         x_ft,
         msl_plot,
         expected_cost;
         ylabel="LSLR: $syear to $eyear [ft]",
         colorbar_title="Expected Total Costs\n[% House Value]",
+        colorbar_formatter=pct_formatter,
         c=cgrad(:plasma; rev=true),
         linewidth=0,
-        levels=30,
+        levels=LinRange(0.25, 3.25, 31),
+        #clim=(0.25, 3.25),
         xticks=(x_ticks, string.(x_ticks)),
         xlabel=L"Height Increase $\Delta h$ [ft]",
-        size=(900, 500),
+        size=(800, 450),
         bottom_margin=5mm,
         left_margin=5mm,
+        right_margin=7.5mm,
     )
     contour!(p, x_ft, msl_plot, expected_cost; linewidth=0.5, levels=25)
     savefig(p, plots_dir("scenario-map-height-slr.pdf"))

### 2022-04-21 scripts/05-tradeoffs-by-rcp.jl

@@ -45,7 +45,7 @@ function plot_rcp_tradeoffs(
             ylabel=varname,
             linewidth=2,
             xticks=(x_ticks, string.(x_ticks)),
-            yformatter=y -> pct_formatter(y),
+            yformatter=pct_formatter,
             top_margin=12.5Plots.mm,
             left_margin=7.5Plots.mm,
             bottom_margin=7.5Plots.mm,

### 2022-04-21 scripts/06-combining.jl

@@ -18,7 +18,7 @@ function plot_grid_scheme()
     p = plot(
         x -> cdf(dist, x);
         xticks=(vcat(ψ, grids), vcat(ψ_labels, grid_labels)),
-        linewidth=3,
+        linewidth=4,
         color=colors[1],
         label="Target Distribution",
         xlims=xlimits,
@@ -81,7 +81,7 @@ function plot_priors()
             p,
             prior.dist,
             0,
-            10;
+            12.5;
             label=prior.name,
             color=color,
             linewidth=5,
@@ -172,9 +172,9 @@ function plot_prior_tradeoffs(
         p = plot(;
             xlabel=L"Height Increase $\Delta h$ [ft]",
             ylabel=varname,
-            linewidth=2,
+            linewidth=4.5,
             xticks=(x_ticks, string.(x_ticks)),
-            yformatter=y -> pct_formatter(y),
+            yformatter=pct_formatter,
             top_margin=12.5Plots.mm,
             left_margin=7.5Plots.mm,
             bottom_margin=7.5Plots.mm,
@@ -182,7 +182,6 @@ function plot_prior_tradeoffs(
         )
 
         # plot all trade-off lines in light gray
-        #=
         for rcp in all_rcp
             for model in all_models
                 w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
@@ -190,10 +189,9 @@ function plot_prior_tradeoffs(
                 plot!(p, Δh_ft, cond; label="", color=:gray, linewidth=0.5, alpha=0.5)
             end
         end
-        =#
 
         # plot the trade-off for the priors
-        for (prior, color) in zip(priors, [:red, :blue, :green])
+        for (prior, color) in zip(priors, colors)
             w = HouseElevation.make_weights(prior.dist)
             cond = vec(mean(var, w; dims=1))
             plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=3)

### 2022-04-21 scripts/main.jl

@@ -61,8 +61,8 @@ function main()
 
     # create scenario maps
     let
-        elevation_init_bfe = [-2, -1, 0, 1]u"ft" # height relative to gauge
-        x_plot = [0, 3, 6, 9]u"ft"
+        elevation_init_bfe = [-2, -0.5, 1]u"ft" # height relative to gauge
+        x_plot = [0, 4, 8]u"ft" # elevations to plot
         plot_scenario_map_slr_cost(;
             x_plot=x_plot,
             elevation_init_bfe=elevation_init_bfe,

### 2022-04-22 papers/paper/preprint.tex
Making good progress on the intro, still needs a bit of polishing

@@
 \newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
+\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
 \newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
@@
 \newacronym[]{usgs}{USGS}{United States Geological Survey}
-\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
+\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}
 \makeglossaries % required to make the list of acronyms
@@
 
+\clearpage
 \section{Introduction}\label{sec:introduction}
 
-Climate risk management, as applied in the United States and around the world, relies heavily upon cost benefit analyses and engineering standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
+Critical infrastructure services in America are aging and inadequate, leading to high-impact failures even during conditions with substantial precedent in the historical record \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
+Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
+A critical question, then, is what conditions or futures particular infrastructure systems and components should be designed to withstand.
+The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
+However, this decision is also subject to assumptions regarding the likelihhod and relevance of possible futures.
 
+\subsection{Current practice}
+
+Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
+These are often, though not always, based on probabilistic analysis of relevant data.
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-One consequence is that those purchasing a home inside the designated floodplain, if using federally backed mortgage porgrams, are nominally required to purchase flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building regulations \citep[based on guidance such as][]{asce_24-05:2006}, and the regulations of major institutions involved in flood risk management \citep[\eg,][]{FEMA_p-55:2011} recommend all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
+These trigger specific floodplain regulations, such as a nominal requirement for homes purchased through using federally backed mortgage porgrams to be covered by flood insurance \citep{kousky_voucher:2014}.
+Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
+Design standards area also used to design large-scale infrastructure.
+For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is $<\nicefrac{1}{4000}$ years \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project proposed for  Galveston Bay, was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+
+Risk-based design \citep{eijgenraam_flooding:2014,vandantzig_dike:1956}\ldots\james{Maybe add a bit here -- this is a low hanging critique}
 And homeowners subject high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.
-Similar methods are used to inform the design of massive infrastructure projects.
-For example, the design height of a seawall, the key component of a \$29 billion coastal protection project proposed by the \gls{usace} for  Galveston Bay, was selected by setting the annual probability that overtopping exceeds a safe threshold to 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
-And while many stakeholder objectives contributed to proposed design, the 1\% annual exceedance probability threshold was selected in large part because the benfit to cost ratio of this design was higher than alternatives considered \citep[\S 2.1.3]{USACE_coastal:2021}.
-A key commonality between these engineering criteria based on design events and these project comparison tools based on cost-benefit analysis is a reliance on probabilistic descriptions of relevant hazard.
+Like standards-based approaches, cost-benefit analyses or risk-based designs rely on probabilistic descriptions of relevant hazard.
 
-Where climate risk management is directly informed by design events (\ie, quantiles of a distribution) or cost-benefit analysis (\ie, integrating over a distribution), the estimated \gls{pdf} will strongly inform the resulting design.
-Current methods focus on data-driven, objective frameworks that can be applied consistently across locations.
+Current methods for estimating the probability distributions that underpin cost-benefit analysis or design events focus on data-driven, objective methods that can be applied consistently across locations.
 For example, \gls{usgs} Bulletin 17C specifies the procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity duration and frequency of extreme rainfall.
-One of the key assumptions underlying most of these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
+A key assumption underlying these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
 Yet changes in both global climate and local environmental conditions challenge this assumption in many cases \citep{Merz:2014gf,Milly:2008dg}.
 For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
+While methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of the trend which may not represent the physical mechanisms that lead to nonstationarity or their uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
 
-The theory of rational decision making evolved in the middle of the 20th century in tandem with mathematical theories of probability and economics.
-In particular, \citet{vonneumann_games:1944} proposed four axioms for rational choice:
-completeness of preference;
-transitivity of preference;
-continuity of preference, and
-independence of preference from the existence of irrelevant alternatives.
-\citeauthor{vonneumann_games:1944} then proved that an individual or firm that follows these actions behaves as though maximizing the expectation of some function, which they called a utility function.
-Similarly, individuals and firms that do not follow these axioms can be systematically exploited \citep[see the ``Dutch book'' arguments of][]{definetti_probability:1972}.
+\subsection{Emerging paradigms}
 
-Bayesian decision theory \citep{savage:1954}
-Over the next decades, the \emph{descriptive} model of rationality used to study firms and individuals \citep[see][]{savage:1954,debreu:1959} became widely applied as a \emph{normative} tool for planning.
-Efforts such as the Harvard Water Program \citep[see][]{maass_hwp:1962} popularized the use of top-down rational planning approaches for water management \citep{howe_bca:1971} and other social planning problems.
-These methods used the vocabulary and mathematical framework of Bayesian decision theory \citep{Gelman:2014tc,savage:1954}, bringing a formal mathematical treatment of these problems.
+Many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
+These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
 
-More fundamentally, while models for estimating and predicting specific forms of nonstationarity have been proposed \citep[see][for a review]{Salas:2018ge}, many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
-Recognizing the challenges of decision making under deep uncertainty, many frameworks for identifying robust decisions have been proposed.
+Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
 Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
 For example, robust decision methods \citep{lempert_shaping:2003} run models over large ensembles of possible futures to assess the performance of different policies, then apply statistical analysis to identify the conditions under which particular policies perform well or poorly.
-Such methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
-When many possible levers are available, and exhausting sampling therefore infeasible, multiobjective robust decision methods \citep{kasprzyk:2013} add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-Many studies formally quantify robustness \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for multiobjective comparison of candidate policies.
+These methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
+When many possible levers are available, and exhaustive sampling therefore infeasible, multiobjective robust decision methods add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+In addition, many studies formally quantify robustness to deep unccertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
 
-Although these methods for decision making under deep uncertainty have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
-One common approach is to systematically sample deeply uncertain parameters from a specified range, then feed these parameters into a system model in order to quantify some metric of aggregate performance \citep{herman:2015,mcphail_robustness:2019,hall_comparison:2012,lempert:2007}.
-A challenge here is that assumptions about the range and independence of uncertain factors can strongly influence which factors are found to be most important and which policies are most robust \citep{quinn_exploratory:2020}.
-Aggregating measures of performance across uncertain factors (\eg, to quantify expectations, quantiles, or robustness metrics of particular decisions) increases sensitivity to these implicit assumptions.
-These assumptions are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
+Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
+When analyses integrate performance over many possible futures, whether to compute sensitivity of responses to inputs, estimate expectations of functions, or compute robustness metrics, they necessarily make assumptions about likelihood of different scenarios.
+Although many analyses contend that specifying a joint probability distribution function over inputs is inappropriate under deep uncertainty, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.
 
-\textbf{Ultimately, recognition of deeply uncertain nonstationary hazard challenges not only the existing stationary estimates of hazard but more fundamentally the premise that objective estimates of future hazard exist and can be estimated empirically}.
-Though seeking robustness can add value to many decision analyses, sensitivity to model assumptions about the likelihood and plausibility of different futures cannot be avoided.
+\subsection{Research gaps and objectives}
+
 Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
@@
 
-The conceptual similarities between this statistical philosophy and the literature on decision making under deep uncertainty lead us to reject the premise, articulated in some studies,\klaus{is it ok to not cite any here} that probability should not be used to inform decision making under deep uncertainty.
-Our subjectivist perspective has a rich history in the Earth sciences.
-For example, \citet{schneider_scenarios:2002} criticizes the reluctance of an \gls{ipcc} special report on emission scenarios \citep{nakicenovic_scenarios:2000} to assign probabilities to scenarios, arguing that ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
-More specifically,
-When even sophisticated \citep[or so we like to think; see][and references therein]{fuller_inversion:2017} consumers of climate information incorporate published plausible ranges into adaptation planning, information that experts may hold the shape of particular distributions can be lost \citep[as warned in][]{moss_uncertainties:2000}.
-
-In this paper we a first conceptual step towards bridging the fields of decision making under deep uncertainty and subjective Bayesian philosophy through  a didactic case study.
-We consider the problem studied in \citet{zarekarizi_suboptimal:2020} of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
+In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model choice.
+We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
 Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
 Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-Prior research has shown that this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}, and that neglecting uncertainty can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``how can deep uncertainties be transparently and consistently synthesized for decision analysis?''
+For example, this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}.
+Further, that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
+Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''
 
@@
 For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
-While we assume here that the \glspl{sow} and decisions are known and discrete, this approach could be  coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
+While we assume for simplicity that the decision space is are known and finite, this approach could be coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
 
@@
 \end{equation*}
-Because expectations are defined relative to a probability distribution, \emph{answering these questions necessitates defining a probability distribution over \glspl{sow}}.
-The general intractability of these integrals motivates the use of Monte Carlo methods by sampling across many \glspl{sow} \citep{Betancourt:2017vd}.
+Because expectations are defined relative to a probability distribution, \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}.
 
@@
         return periods.
-        Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula $\nicefrac{r}{N+1}$.
+        Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula (eq.~\ref{eq:weibull}).
         Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
@@
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function:
-\begin{equation}\label{eq:gev-dist}
-    f(x | \mu, \sigma, \xi)= \begin{cases}
-        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
-        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
-    \end{cases}
-\end{equation}
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
 This model assumes stationarity, neglecting any potential time dependence.
@@
 Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
-Specifically, we apply Normal (truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
+Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+The parameters of the Inverse Gamma distribution can be calculated from these moments (see \cref{eq:inv-gamma-params}).
+These means and standard deviations were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
 
@@
 \Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position, $\nicefrac{r}{N + 1}$, where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
+The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position (see \cref{eq:weibull}).
 The good fit of the Bayesian fit (gray shading) to the data (points) suggest the modeling choices made are reasonable and sufficient for this didactic example.
@@
 Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences.
-For example, \citet{lempert_slr:2012} considers a five parameter model for sea level rise that takes into account rates of change for different physical mechanisms; it could make sense to put a prior over this higher dimensional space.
-A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent the .
+A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
 Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
-Finally, and more fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
+Another important point is that the gridding approach of \cref{eq:weights} implicitly assumes that the space is well spanned; this is not necessarily true but is a goal of exploratory modeling.\james{revisit}
+More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
 An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
 Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.\klaus{Is this too vague?}
+Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
 
@@
 
-The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate adaptation.
+The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate risk management.
+An obvious area of application is to the design of infrastructure for climate risk management.
 For example, much of the stormwater infrastructure in the United States is inadequate for current climate \citep{lopez-cantu:2018}, let alone future changes.
 Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \citep{sharma_rcp:2021} and \gls{rcp} scenarios.
-Another example is the design and operation of coastal infrastructure.
-Problems such as levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
+Similarly, decisions like  levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
+This framework can also be applied to uncertainties in non-physical variables.
+For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \citep{trindade_deeplyuncertainpathways:2019}.
+Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways,\james{cite} are conditional on probabilistic models for inputs like technology prices and population.
 
-Ultimately, the proposed framework rests on the argument that if you cannot be right, it is best to be transparent.
-We call for researchers studying climate risk management to make their implicit assumptions about  distributions more transparent, and we call for coordinated national and international guidance to assist local governments and engineering bodies determine appropriate subjective beliefs.\klaus{OK with this?}
+Ultimataely, all models of the future are flawed.
+Seeking decisions that perform well even when assumptions are broken through robust design and flexibility generally improves outcomes.
+Yet whenever decisions are comppared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
+We call for researchers studying climate risk management to make these implicit assumptions, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\klaus{OK with this?}
 
@@
 \section{Supplemental methods}
+
+\subsection{Storm surge statistical modeling}
+
+We parameterize the \gls{gev} distribution as
+\begin{equation}\label{eq:gev-dist}
+    f(x | \mu, \sigma, \xi)= \begin{cases}
+        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
+        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
+    \end{cases}
+\end{equation}
+
+Parameterizing the Inverse Gamma distribution as
+\begin{equation}\label{eq:inv-gamma-dist}
+    f(x | \alpha, \theta) = \frac{\theta^\alpha x^{-(\alpha + 1)}}{\Gamma(\alpha)} e^{-\frac{\theta}{x}}, \quad x > 0,
+\end{equation}
+the parameters $\alpha$ and $\theta$ can be computed from the desired mean $\mu$ and standard deviation $\sigma$ as
+\begin{align}\label{eq:inv-gamma-params}
+    \begin{split}
+        \alpha &= 2 + \frac{\mu^2}{\sigma^2} \\
+        \theta &= \mu \qty(\alpha - 1).
+    \end{split}
+\end{align}
+
+The plotting position used in \cref{fig:surge-obs-return} is the Weibull (``empirical'') plotting position
+\begin{equation}\label{eq:weibull}
+    \nicefrac{r}{N + 1}
+\end{equation}
+where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
 

### 2022-04-22 scripts/05-tradeoffs-by-rcp.jl
fix upper x axis labels

@@ -77,7 +77,7 @@ function plot_rcp_tradeoffs(
             twiny(p),
             Δh_ft,
             [mean(var[:, i]) for (i, _) in enumerate(Δh_ft)];
-            xticks=(x_ticks, string.(round.(prop_cost, digits=2))),
+            xticks=(x_ticks, pct_formatter.(prop_cost)),
             xlabel="Up-Front Cost [% House Value]",
             linewidth=0,
             alpha=0,

### 2022-04-22 scripts/06-combining.jl

@@ -202,7 +202,7 @@ function plot_prior_tradeoffs(
             twiny(p),
             Δh_ft,
             [mean(var[:, i]) for (i, _) in enumerate(Δh_ft)];
-            xticks=(x_ticks, string.(round.(prop_cost, digits=2))),
+            xticks=(x_ticks, pct_formatter.(prop_cost)),
             xlabel="Up-Front Cost [% House Value]",
             linewidth=0,
             alpha=0,

### 2022-04-22 tikz/bayes-rdm.tex
remove extraneous text

@@
             $\vb{M} = \qty{M_1, \ldots, M_K}$\\
-            for $p(\Omega | M_k \in \mathcal{M})$
+            for $p(\Omega | M_k)$
         };

### 2022-04-22 papers/paper/preprint.tex
Edits to print out and work on

@@
 
-Critical infrastructure services in America are aging and inadequate, leading to high-impact failures even during conditions with substantial precedent in the historical record \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
+Critical infrastructure services in America are aging and inadequate, leading to high-impact failures under conditions known to be possible \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
 Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
@@
 
-Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
+Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
 These are often, though not always, based on probabilistic analysis of relevant data.
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a nominal requirement for homes purchased through using federally backed mortgage porgrams to be covered by flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
-Design standards area also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is $<\nicefrac{1}{4000}$ years \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project proposed for  Galveston Bay, was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+These trigger specific floodplain regulations, such as a nominal requirement for homes purchased with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
+Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require new construction in flood zones to be elevated some margin above a nominal \gls{bfe}.
+Design standards are also used to design large-scale infrastructure.
+For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 
-Risk-based design \citep{eijgenraam_flooding:2014,vandantzig_dike:1956}\ldots\james{Maybe add a bit here -- this is a low hanging critique}
-And homeowners subject high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.
-Like standards-based approaches, cost-benefit analyses or risk-based designs rely on probabilistic descriptions of relevant hazard.
+There are many advantages to standards-based design.
+In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.\klaus{I understand this to mean that they are the same for everyone, even if they don't necessarily lead to equitable outcomes -- is that how you read this?}
+However, there are also limitations.
+In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
+This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
+Like standards-based approaches, quantitative cost-benefit analyses rely on probabilistic descriptions of relevant hazard.
 
-Current methods for estimating the probability distributions that underpin cost-benefit analysis or design events focus on data-driven, objective methods that can be applied consistently across locations.
-For example, \gls{usgs} Bulletin 17C specifies the procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity duration and frequency of extreme rainfall.
-A key assumption underlying these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
-Yet changes in both global climate and local environmental conditions challenge this assumption in many cases \citep{Merz:2014gf,Milly:2008dg}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
-While methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of the trend which may not represent the physical mechanisms that lead to nonstationarity or their uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
+Currently used methods for estimating these \glspl{pdf} emphasize data-driven, objective methods that can be applied consistently across locations.
+For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
+Among several statistical assumptions these analyses make that have been recently called into question \citep[see][for a review]{Merz:2014gf} is stationarity.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are apparent or anticipated \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
+Although some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
+Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
 
@@
 
-Many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
+Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016}.
+In other words, they are deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
 These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
@@
 Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
-For example, robust decision methods \citep{lempert_shaping:2003} run models over large ensembles of possible futures to assess the performance of different policies, then apply statistical analysis to identify the conditions under which particular policies perform well or poorly.
-These methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
-When many possible levers are available, and exhaustive sampling therefore infeasible, multiobjective robust decision methods add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-In addition, many studies formally quantify robustness to deep unccertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
+For example, robust decision making \citep{lempert_shaping:2003} runs models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
+When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+In addition, many studies formally quantify robustness to deep uncertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
 
@@
 Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
-When analyses integrate performance over many possible futures, whether to compute sensitivity of responses to inputs, estimate expectations of functions, or compute robustness metrics, they necessarily make assumptions about likelihood of different scenarios.
-Although many analyses contend that specifying a joint probability distribution function over inputs is inappropriate under deep uncertainty, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
+Although some have argued\klaus{Can I get away w/o calling out a particular paper here?} that specifying a joint \gls{pdf} over inputs is inappropriate under deeply uncertain conditions, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
 This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.
@@
 Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
-This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a family of candidate models, and we find many parallels to deep uncertainty.
-The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, fr which probability distributions offer a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a discussion of methods for model selection).
+This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a candidate space.
+The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
 From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
-Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}
+Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders, decisionmakers, and analysts \citep{gelman_philosophy:2013}.
+Within this philosophical framework we find many parallels to the motivations and philosophies of \gls{dmdu}.
 
-In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model choice.
-We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
+In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
+We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
 Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
-Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-For example, this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}.
-Further, that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
+Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
 Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''
@@
 We proceed as follows.
-In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-explore}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
+In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
 In \cref{sec:case-study} we describe the case study in detail.
-Next in \cref{sec:results} we discuss results for each of the three theoretical lenses.
+Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
 Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.
@@
 For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
-While we assume for simplicity that the decision space is are known and finite, this approach could be coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
+While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
+Similarly, we assume that a fixed ensemble of \glspl{sow} are available for analysis, but in some cases additional simulations may add value.
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption about the likelihood different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
+Exploratory modeling avoids making explicit assumption about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or mapping a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
 \klaus{Does this address your comment about linking back to observations?}
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric''  requires estimating the expectation of some function $\pi$, which has the form
-\begin{equation*}
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires estimating the expectation of some function $\pi$, which has the form
+\begin{equation}\label{eq:expectation-is-probability}
     \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
-\end{equation*}
-Because expectations are defined relative to a probability distribution, \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}.
+\end{equation}
+In other words, \cref{eq:expectation-is-probability} shows that \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}, $p(s)$.
 
@@
 Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
-Since the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
+If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing Monte Carlo estimation of \cref{eq:expectation-is-probability} \citep{betancourt:2018}.
 
-Many decision analytic tools used in climate risk management can be interpreted through this lens.
-A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
-For example, \ldots\james{Add a water resources and a flood example}
-In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
-As we shall discuss in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only using a scenario approach (\ie, choices of model structures or factors such as forcings we show as cases) without any explicit probability statement \citep[see][]{wong_nola:2017}.\james{Wording}
-
-Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
+This framework can be applied to understand many existing analyses.
+For example, if we define $M$ to be a probabilistic scenario (\eg, a choice of physical model and of an \gls{rcp} scenario, as we shall see in the next section), then we can rigorously quantify uncertainty \emph{within models}, but can only characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
+We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
 For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
 Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
-Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
-\begin{equation}
-    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
-\end{equation}
-If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.
+Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$.
 
 Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
+For example, if this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.
 
@@
 If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
-One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
+One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
 In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
-Consequently, climate risk management often uses the outputs of structured simulations and analyses.
-For example, many studies use outputs from \gls{cmip} simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
-It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.
 
@@
 For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
-This approach is widely used in analysis of tropical cyclones: first an ensemble of cyclones is simulated, and then probabilistic weights are calculated for each based on attributes such as intensity, landfall location, and radius \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
+This weighting approach parallels the joint probability method used in analysis of tropical cyclones \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
@@
 
-The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
+The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis, even when the \glspl{sow} deliberately over- or under-sample some regions of the parameter space.
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
 It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
-However, reasoning using subjective probability allows\ldots\james{Finish sentence}
+However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
 
@@
 
-Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.
+Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
 
@@
 Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
+Rather than applying prior information directly over the joint distribution of the parameters $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
 Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
@@
 Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017}.
+Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017,ossandon_semibayesian:2021}.
 
 \Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position (see \cref{eq:weibull}).
-The good fit of the Bayesian fit (gray shading) to the data (points) suggest the modeling choices made are reasonable and sufficient for this didactic example.
+The return period of the data (dots) is shown using the Weibull plotting position (see \cref{eq:weibull}).
+The good fit of the Bayesian model to the data suggest the modeling choices made are reasonable and sufficient for this didactic example.
 A positive control test (\cref{fig:surge-synthetic-data-experiment}) also validates the model's ability to recover known parameter values.
@@
 The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
-The second converts model converts a time series of annual expected damages into lifetime expected damages.
+The second model converts a time series of annual expected damages into lifetime expected damages.
 
@@
 However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
-Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient a surrogate model (see \cref{sec:surrogate-ead}).
+Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).
 
@@
     \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
-    \item ``Expected lifetime costs'' is the sum of lifetime expected damages and up-front costs.\james{Make sure the figures all use this term}
+    \item ``Expected lifetime costs'' is the sum of lifetime expected damages and up-front costs.
 \end{enumerate}
@@
 
-We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.\james{Expand a bit on other work and priors for SLR.}
-Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.
+We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
 
-We use a Gamma distribution for all three priors, parameterized as
-\begin{equation*}
-    f(x | \alpha, \theta) = \frac{x^{\alpha-1} e^{-x/\theta}}{\Gamma(\alpha) \theta^\alpha},
-    \quad x > 0.
-\end{equation*}
-\Cref{tab:slr-priors} specifies these priors, as well as some quantiles of the distributions.
+We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
+\Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
 Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
@@
 We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
-We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
-
-Future work could develop improved, policy-relevant priors, either by combining probabilistic models for emissions \citep[as in][]{srikrishnan_probabilistic:2022}, climate sensitivity, and the \gls{slr}, or through inversion of expert opinion \citep[as in][]{fuller_inversion:2017}.
-Future work could also consider more sophisticated priors, either by considering higher-dimensional $\psi$ or by modeling a joint prior over model inputs and model outputs through melding \citep{poole_melding:2000,sevcikova_melding:2007}.
+We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}, or to probabilistic models used in prior \gls{dmdu} studies \citep{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.
 
@@
 
-\james{Start with a summary sentence that outlines the paragraph}\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
+One application of explotatory analysis is to reveal uncertainty and variance in outcome under particular decisions.
+\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
 The outcomes with lowest total lifetime costs arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
@@
 This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
-However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions.\james{Strong statement.
-Correct?
-What if we make or use strong implicit assumptions?}
+However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
 
 \subsection{Scenario-conditional optimization}\label{sec:results-conditional}
+
 We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
@@
 This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
-Consistent with \cref{fig:scenario-map-slr-cost}, the variability of lifetime costs decreases as $\Delta h$ increases.
-Once $\Delta h$ reaches \SIrange[]{3}{7}{ft}, depending on the model considered, construction costs start to dominate flood losses, and thus higher values of $\Delta h$ increase average lifetime costs.
+For example, under the most pessimistic model (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
+Converseley, under the most optimistic model (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost of 73\% of house value yet reduces lifetime expected damages by less than 50\% of house value.
 
@@
 The multiple \gls{pdf} problem has also been shown in other contexts.
-For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{Other examples: Wong et al 2017 -- levees.
-There are more!}
-Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
+For example, \citet{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
+Similarly, \citet{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
+Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
 
@@
 \Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
-Notably, they give different guidance.\james{Say what this guidance is}
-Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} saves approximately 25\% of the house value relative to not elevating.
-Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} costs approximately 25\% of the house value relative to not elevating.
-Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house.
+Notably, they give different guidance.
+Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} costs 73\% of house value and reduces damages by over 100\% of house value, yielding a benefit to cost ratio of approximately 1.25.
+Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} reduces damages only by 50\% of house value, yielding a benefit to cost ratio of approximately 0.7.
+Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house, and thus the benefit to cost ratio is nearly 1.
 Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
@@
 Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
-Another important point is that the gridding approach of \cref{eq:weights} implicitly assumes that the space is well spanned; this is not necessarily true but is a goal of exploratory modeling.\james{revisit}
 More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
@@
 This problem intersects with decisions about where to live and how to manage household finances, both of which are highly complex.
-We could extend our analysis, for example, by considering additional decision objectives representing risk aversion, usable space created under the house, non-financial consequences of flooding, or other factors.
+We could extend our analysis, for example, by considering additional decision objectives.
+In particular, we hypothesize that incorporating improved representations of risk aversion into decision support may substantially improve their usability.
 We could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
@@
 For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \citep{trindade_deeplyuncertainpathways:2019}.
-Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways,\james{cite} are conditional on probabilistic models for inputs like technology prices and population.
+Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
 
@@
 
-\subsection{Storm surge statistical modeling}
+\subsection{Statistical distributions}
 
@@
 \end{equation}
-
 Parameterizing the Inverse Gamma distribution as
@@
 \end{align}
-
+We parameterize the Gamma distribution as
+\begin{equation}\label{eq:gamma-dist}
+    f(x | \alpha, \theta) = \frac{x^{\alpha-1} e^{-x/\theta}}{\Gamma(\alpha) \theta^\alpha},
+    \quad x > 0.
+\end{equation}
 The plotting position used in \cref{fig:surge-obs-return} is the Weibull (``empirical'') plotting position

### 2022-04-24 papers/paper/preprint.tex
getting close, still needs a bit of work on abstract and section 2.2

@@
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a nominal requirement for homes purchased with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require new construction in flood zones to be elevated some margin above a nominal \gls{bfe}.
+These trigger specific floodplain regulations, such as a requirement for homes in the 100 year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
+Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
 Design standards are also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the annual probability of overtopping to 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 
@@
 For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are apparent or anticipated \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
-Although some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
+While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
 Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
@@
 Although some have argued\klaus{Can I get away w/o calling out a particular paper here?} that specifying a joint \gls{pdf} over inputs is inappropriate under deeply uncertain conditions, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
-This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.
+This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
 
@@
 Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
-This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a candidate space.
+This construct is a theoretical contrast to the classical approach to model selection in the ``$\mathcal{M}$-closed'' case, in which methods seek to asymptotically identify the ``true'' model from a candidate space.
 The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
@@
 
-Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
-For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
+Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
+For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
 While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
-Similarly, we assume that a fixed ensemble of \glspl{sow} are available for analysis, but in some cases additional simulations may add value.
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or mapping a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
-\klaus{Does this address your comment about linking back to observations?}
+Exploratory modeling are averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993}by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
+Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system modle, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires estimating the expectation of some function $\pi$, which has the form
-\begin{equation}\label{eq:expectation-is-probability}
-    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
-\end{equation}
-In other words, \cref{eq:expectation-is-probability} shows that \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}, $p(s)$.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes.\klaus{I could try to pull in some Monte Carlo equations here, but I think that would just be adding confusion}
 
-In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
-Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
-If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing Monte Carlo estimation of \cref{eq:expectation-is-probability} \citep{betancourt:2018}.
+One way to interpret an ensemble of \glspl{sow} is as \gls{iid} draws from some true data generating process.
+We call this scenario-conditional probabilistic analysis, which commonly arises when a probabilistically calibrated model for relevant outputs (\eg, a physical model for \gls{msl}; see \cref{sec:case-slr}) is forced by a deterministic scenario (\eg, a \gls{rcp} scenario).
+In this case, \emph{the set of available \glspl{sow} is assumed to come from the true (often called population) distribution, and thus the sample statistics are assumed to reflect the true values.}
+We add this concept to our framework with boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario (\ie, the assumed input) $M_k$.
+If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
+This approach allows for rigorous quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
 
-This framework can be applied to understand many existing analyses.
-For example, if we define $M$ to be a probabilistic scenario (\eg, a choice of physical model and of an \gls{rcp} scenario, as we shall see in the next section), then we can rigorously quantify uncertainty \emph{within models}, but can only characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
 We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
 For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
-Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
-Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$.
+Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
+Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
 
 Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
-For example, if this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.
 
@@
 
-If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
+If the ensemble of \glspl{sow} is not drawn \gls{iid} from the true distribution (\ie, if we deliberately over-sample low-probability but high-impact \glspl{sow}) or if we wish to synthesize \emph{across} multiple scenarios $M_1, M_2, \ldots, M_K$, then additional measures must be taken.
 One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
@@
 Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
-To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
+A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through kernel density estimation or other methods.
+Instead, paralleling joint probability methods for statistical analysis of tropical cycloines, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
+
+First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
 This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
@@
 For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
-This weighting approach parallels the joint probability method used in analysis of tropical cyclones \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
@@
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
+We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained (even with better models and data).
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
@@
 
-We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
+To illustrate the proposed decision analytic framework we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
 We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
@@
 
-Our approach to model building follows a principled workflow for model building and checking \citep[see][for details]{gelman_workflow:2020}.
+Our approach to model validation is based on the concept of principled workflow design for model building and checking \citep[see][for details]{gelman_workflow:2020}.
 One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
@@
 Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-The parameters of the Inverse Gamma distribution can be calculated from these moments (see \cref{eq:inv-gamma-params}).
+The parameters of the Inverse Gamma distribution can be calculated from these moments (see eq.~\ref{eq:inv-gamma-params}).
 These means and standard deviations were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
@@
 
-We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
+We construct three probabilistic models for $p_\mathrm{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
 
@@
     \caption{
-        Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\text{belief}(\psi)$.
-        The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles.
+        Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\mathrm{belief}(\psi)$.
+        The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles (values in \si{ft}).
     }\label{tab:slr-priors}
@@
 We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
-We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}, or to probabilistic models used in prior \gls{dmdu} studies \citep{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.
+We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
+We can also compare to the analyses of \citet{lempert_slr:2012,sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
+Our samples bound all of these estimates.
 
@@
 
-One application of explotatory analysis is to reveal uncertainty and variance in outcome under particular decisions.
+One application of explotatory analysis is to reveal the range and variation in outcomes, conditional on taking a particular decision.
 \Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
@@
 The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
-As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
+As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because,.
+even if \gls{slr} is substantial, damages will be negligible.
 
@@
 We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
-Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draws from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
-As discussed in \cref{sec:analysis-condition}, this allows a probabilistic interpretation: conditional on a particular \gls{rcp} scenario and physical model, we can reason about the distribution of outcomes.
+Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draw from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
+As discussed in \cref{sec:analysis-condition}, this allows a formal estimation of decision metrics.
 
@@
         Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
-        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kolpin_introduction:2017} under \gls{rcp} 8.5.
+        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
         (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
@@
 
-One application of this method is to diagnose which assumptions different $p_\text{belief}$ are consistent with.
+One application of this method is to diagnose which assumptions different $p_\mathrm{belief}$ are consistent with.
 \Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
@@
     \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\text{belief}$ (colors).
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\mathrm{belief}$ (colors).
         Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
@@
 
+To avoid ambiguity, we define here some statistical distributions used in the main text.
+Except where otherwise indicated, we follow the default notation of the Distributions.jl Julia package \citep{besancon_distributions.jl:2021}.
+
+\subsubsection{Generalized Extreme Value distribution}
 We parameterize the \gls{gev} distribution as
@@
 \end{equation}
+
+\subsubsection{Inverse Gamma distribution}
 Parameterizing the Inverse Gamma distribution as
@@
 \end{align}
+
+\subsubsection{Gamma distribution}
 We parameterize the Gamma distribution as
@@
 \end{equation}
+
+\subsubsection{Plotting position}
 The plotting position used in \cref{fig:surge-obs-return} is the Weibull (``empirical'') plotting position
@@
 
-We use a Monte Carlo integration to estimate expected damages as a function of house elevation and \gls{msl}, $D(h-y(t))$.
-\begin{enumerate}
-    \item For $m=1, \ldots, M$:
+As discussed in \cref{sec:case-metrics}, we use a Monte Carlo integration to estimate expected damages $D$ as a function of house elevation $h = h_0 + \delta h$ and \gls{msl}, $y(t)$.
+Specifically, for $m=1, \ldots, M$:
 \begin{enumerate}
@@
     \item simulate a single storm surge from this stationary \gls{gev} distribution and add the mean sea level to get total flood depth $y^\mathrm{sim}_m$
-              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $m$th damage.
+    \item calculate the flood damages for this draw by plugging the annual maximum flood depth relative to the house,$h - y'(t)$, into  the deterministic HAZUS depth-damage relationship, storing this as the $m$th damage.
 \end{enumerate}
-    \item Estimate expected annual damages as the sample mean of the $M$ estimates
-\end{enumerate}
+We then estimate expected annual damages as the sample mean of the $M$ estimates.
 
 \subsection{Surrogate model for expected annual damages}\label{sec:surrogate-ead}
-Evaluating expected annual damages for each of $N$ simulations of \gls{slr}, each of $J$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
-In our model, we have $T=71$, $J=\num{10000}$, $N=\num{179232}$, and in general .
-However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
+Evaluating expected annual damages for each of $J$ simulations of \gls{slr}, each of $N$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
+In our model, we have $T=71$, $N=\num{10000}$, $J=\num{179232}$, and so exhaustive sampling may be prohibitive.
+
+Noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
 To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.

### 2022-05-24 papers/paper/preprint.tex
Updates from Overleaf

--- 2022-04-22 papers/paper/preprint.tex
@@
-\documentclass[11pt]{article}
+\documentclass[12pt]{article}
 
 % Figures
-\usepackage{graphicx}
+\usepackage[demo]{graphicx}
 \usepackage[list=true]{subcaption}
@@
 \pagestyle{fancy}
-\setlength{\headheight}{15pt}
 
@@
 \usepackage{authblk}
-\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
+\title{Mainstreaming Climate Adaptation Requires Subjective Integration of Deep Uncertainties}
 \author[1]{James Doss-Gollin}
@@
 % ACRONYMS
-\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
+\usepackage[acronym,nopostdot,nonumberlist,shortcuts,]{glossaries}
+\newacronym{bdt}{BDT}{Bayesian decision theory}
 \newacronym{bfe}{BFE}{base flood elevation}
-\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
-\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
+\newacronym[]{cdf}{CDF}{cumulative distribution function}
 \newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
-\newacronym{gev}{GEV}{generalized extreme value}
+\newacronym{gcm}{GCM}{general circulation model}
+\newacronym[]{gev}{GEV}{generalized extreme value}
 \newacronym{iid}{IID}{independent and identically distributed}
 \newacronym{ipcc}{IPCC}{International Panel on Climate Change}
-\newacronym{msl}{MSL}{mean relative sea level}
-\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
+\newacronym{lsl}{LSL}{local mean sea level}
+\newacronym{mcmc}{MCMC}{Markov Chain Monte Carlo}
 \newacronym{pdf}{PDF}{probability density function}
 \newacronym{rcp}{RCP}{representative concentration pathway}
-\newacronym{slr}{SLR}{sea level rise}
-\newacronym{ssp}{SSP}{shared socio-economic pathway}
-\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
-\newacronym[]{usgs}{USGS}{United States Geological Survey}
-\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}
-\makeglossaries % required to make the list of acronyms
+\newacronym{msl}{MSL}{mean relative sea level}
+\newacronym{rdm}{RDM}{robust decision making}
+\newacronym[plural=SOWs,descriptionplural=states of the world]{sow}{SOW}{state of the world}
+\newacronym{tc}{TC}{tropical cyclone}
 
@@
 \begin{abstract}
-    Methods for doing so, whether classical standards-based approaches or bottom-up robust methods, make implicit or explicit assumptions about the likelihood of different futures.
-    These assumptions may be inconsistent with expert belief.
-    In this paper we propose a probabilsitic method for synthesizing across deep uncertainties, using subjective prior beliefs.
-    The proposed method can be used in tandem with bottom-up exploratory methods and is consistent with fully probabilistic design criteria.
-    Through the lens of a didactic case study of house elevation in Norfolk, VA, we illustrate how this approach can improve the transparency of modeling assumptions.\klaus{Abstract needs a little work}
+    National agencies, regional governments, and individual households manage flood risk through a combination of standards and cost-benefit analyses that rely on estimates of flood hazard.
+    One example is house elevation: \acrlong{fema} recommends a house located in the floodplain be elevated to the \acrlong{bfe} (typically the 100 year flood level) plus a freeboard (typically \SI{1}{ft}) if doing so would pass a cost-benefit test.
+    Under stationarity, the probabilities used for this analysis (\ie to estimate the 100 year flood height or to perform the cost-benefit analysis) can be estimated using empirical frequencies from the observational record.
+    However, growing recognition of nonstationarity motivates a forward-looking approach.
+    Efforts to incorporate nonstationarity hazard into planning have struggled to overcome deep uncertainties in projections of future hazard because different models  may fit the observed data equally well but yield very different estimates of future hazard, and there is often no objective way to choose between them (\ie, equifinality).
+    In this paper we use a didactic case study of deciding if and how high to elevate a single house in the coastal zone to illustrate a decision theoretic framework that combines exploratory modeling, robust decision making, and subjective Bayesian philosophy.
+    We first use exploratory modeling to illustrate that there is no dominant strategy; instead, there are instrinsic tradeoffs between performance under low- and high-sea level rise scenarios.
+    Next, we map the tradeoff between up-front cost and expected damages for each of sixteen models of future sea level rise to illustrate that the choice of which model to use has a strong influence on estimated trade-offs, cost-benefit analyses, or policy prescriptions.
+    To overcome these shortcomings, we propose and illustrate a computationally efficient method for scenario weighting conditional on a subjective prior belief over future sea level rise.
+    This method provides a transparent framework for synthesizing deep uncertainties, facilitating critique and dialogue around necessarily subjective decisions.
+    It can be applied beyond nontationary flood risk management to other problems in climate adaptation, calculating the social cost of carbon, and more.
 \end{abstract}
@@
 \begin{enumerate}
-    \item Estimated trade-offs between adaptation objectives are sensitive to assumptions about the likelihood of future conditions.
-    \item We propose a probabilistic method for transparently synthesizing across deep uncertainty in future conditions.
-    \item The proposed method identifies strategies for house elevation that are robust to assumptions in sea level rise.
+    \item
 \end{enumerate}
@@
 
-Critical infrastructure services in America are aging and inadequate, leading to high-impact failures even during conditions with substantial precedent in the historical record \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
-Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
-A critical question, then, is what conditions or futures particular infrastructure systems and components should be designed to withstand.
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
-However, this decision is also subject to assumptions regarding the likelihhod and relevance of possible futures.
-
-\subsection{Current practice}
-
-Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular ``design events'' that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
-These are often, though not always, based on probabilistic analysis of relevant data.
-For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a nominal requirement for homes purchased through using federally backed mortgage porgrams to be covered by flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011}, may require all houses in flood zones to be elevated some margin above a nominal \gls{bfe}.
-Design standards area also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is $<\nicefrac{1}{4000}$ years \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project proposed for  Galveston Bay, was designed for the annual probability of overtopping to equal 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
-
-Risk-based design \citep{eijgenraam_flooding:2014,vandantzig_dike:1956}\ldots\james{Maybe add a bit here -- this is a low hanging critique}
-And homeowners subject high flood risk can apply to programs that fund house elevation; decisions about which projects to fund are made primarily on the basis of cost-benefit ratio tests.
-Like standards-based approaches, cost-benefit analyses or risk-based designs rely on probabilistic descriptions of relevant hazard.
-
-Current methods for estimating the probability distributions that underpin cost-benefit analysis or design events focus on data-driven, objective methods that can be applied consistently across locations.
-For example, \gls{usgs} Bulletin 17C specifies the procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity duration and frequency of extreme rainfall.
-A key assumption underlying these approaches is stationarity (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
-Yet changes in both global climate and local environmental conditions challenge this assumption in many cases \citep{Merz:2014gf,Milly:2008dg}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}.
-While methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of the trend which may not represent the physical mechanisms that lead to nonstationarity or their uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
-
-\subsection{Emerging paradigms}
-
-Many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016} and are therefore deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
-These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
-
-Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
-Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
-For example, robust decision methods \citep{lempert_shaping:2003} run models over large ensembles of possible futures to assess the performance of different policies, then apply statistical analysis to identify the conditions under which particular policies perform well or poorly.
-These methods have been applied to coastal infrastructure planning by sampling \glspl{sow} from a \gls{pdf} of \gls{slr}, then identifying decisions that perform well over many possible futures \citep{sriver_sealevel:2018,lempert_slr:2012,wong_nola:2017}.
-When many possible levers are available, and exhaustive sampling therefore infeasible, multiobjective robust decision methods add policy search (using optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-In addition, many studies formally quantify robustness to deep unccertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
-
-Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
-Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
-When analyses integrate performance over many possible futures, whether to compute sensitivity of responses to inputs, estimate expectations of functions, or compute robustness metrics, they necessarily make assumptions about likelihood of different scenarios.
-Although many analyses contend that specifying a joint probability distribution function over inputs is inappropriate under deep uncertainty, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
-This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the necessity of modeling the likelihood of possible futures.
-
-\subsection{Research gaps and objectives}
-
-Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
-This construct is a theoretical contrast to the ``$\mathcal{M}$-closed'' case, where the goal is to select the ``true'' model from a family of candidate models, and we find many parallels to deep uncertainty.
-The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, fr which probability distributions offer a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a discussion of methods for model selection).
-From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
-Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders and decisionmakers and analysts \citep{gelman_philosophy:2013}
-
-In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model choice.
-We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem where both over- and under-investment can lead to undesired outcomes.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
-Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
-For example, this ``one size fits all'' approach can deviate substantially from ``optimal'' guidance tailored to specific homes \citep{xian_elevation:2017}.
-Further, that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''
-
-We proceed as follows.
-In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-explore}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
-In \cref{sec:case-study} we describe the case study in detail.
-Next in \cref{sec:results} we discuss results for each of the three theoretical lenses.
-Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.
-
-\section{Decision analytic framework}\label{sec:analysis}
-
-In this section we outline our framework for decision analysis under deep uncertainty, maintaining a high level of generality.
-In the next we discuss application to the house elevation case study.
-
-Following \cref{fig:flowchart}, consider using $J$ \glspl{sow} $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$ (box b in \cref{fig:flowchart}) to evaluate $I$ discrete decisions $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$ (box a in \cref{fig:flowchart}).
-For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$ (box c in \cref{fig:flowchart}).
-While we assume for simplicity that the decision space is are known and finite, this approach could be coupled to a policy search model that proposes candidate decisions $x_i$ from some decision space $\mathcal{X}$.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
-    \caption{
-        Outline of the proposed decision-analytic framework.
-        In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance (c) of candidate decisions (a) under a large ensemble of possible futures (b).
-        In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes (e) that are conditional upon specific models describing the likelihood of different futures (d).
-        In our case study, these are \gls{rcp} scenarios and physical models of sea level rise.
-        Finally in \cref{sec:analysis-synthesize} we propose a subjective Bayesian framework for synthesizing across deep uncertainties by re-weighting sampled futures (f).
-    }\label{fig:flowchart}
-\end{figure}
-
-\subsection{Explore}\label{sec:analysis-explore}
-
-A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling avoids making explicit assumption about the likelihood different \glspl{sow} and instead seeks to generate new knowledge \citep{bankes:1993}, for example by discovering particularly relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} and mapping a system's response to decisions and \glspl{sow} \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Although exploratory modeling seeks to avoid explicit assumptions about the likelihood of different futures, subjective modeling decisions including the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} \citep[see][]{quinn_exploratory:2020} strongly influence results.
-\klaus{Does this address your comment about linking back to observations?}
-
-\subsection{Condition}\label{sec:analysis-condition}
-
-Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of some metric under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of failing to satisfy a critical metric''  requires estimating the expectation of some function $\pi$, which has the form
-\begin{equation*}
-    \mathbb{E}\qty[f(s | x_i)] = \int_{s \in \Omega} p(s) \pi(s | x_i) \dd{s}.
-\end{equation*}
-Because expectations are defined relative to a probability distribution, \emph{estimating any such expectation requires defining a probability distribution over \glspl{sow}}.
-
-In boxes (d) and (e) of \cref{fig:flowchart} we add this concept to our decision analytic framework.
-Given some model $M_k$ specifying a probability over \gls{sow} $p(\Omega)$, we sample $\vb{s} \sim p(\Omega | M)$.
-Since the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
-
-Many decision analytic tools used in climate risk management can be interpreted through this lens.
-A common application (to which we return in our case study) is when using probabilistic models with deeply uncertain boundary conditions.
-For example, \ldots\james{Add a water resources and a flood example}
-In this case, $M$ describes both the choice of model and the choice of boundary condition (\eg, choice of physical model and the \gls{rcp} scenario).
-As we shall discuss in \cref{sec:results-conditional}, this analysis models uncertainty \emph{within models} probabilistically, but models uncertainty \emph{between models} only using a scenario approach (\ie, choices of model structures or factors such as forcings we show as cases) without any explicit probability statement \citep[see][]{wong_nola:2017}.\james{Wording}
-
-Another application is to studies that generate an ensemble of \glspl{sow} by sampling parameters across fixed ranges.
-For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
-Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, calculating robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
-Since sampling over a range is equivalent to sampling from a Uniform distribution, this method makes an implicit assumption that $M$ is the product of $L$ independent Uniform distributions (one for each parameter) over the plausible ranges $\qty(\theta_{\ell,\min}, \theta_{\ell,\max}) \forall \ell \in \qty{1, \ldots, L}$:
-\begin{equation}
-    p(\Omega | M) = \prod_{\ell=1}^L \mathrm{Uniform} \qty(\theta_{\ell,\min}, \theta_{\ell,\max}).
-\end{equation}
-If this product of independent Uniform distributions truly represents a well-calibrated belief about the joint probability distribution over \glspl{sow} $p(\Omega)$ then it should be used, but analysts should be clear that, like it or not, they are making probabilistic assumptions, and should apply methods for model critique and validation \citep[see][]{gelman_workflow:2020}.
-
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
-
-\subsection{Synthesize}\label{sec:analysis-synthesize}
-
-If the ensemble of \glspl{sow} available does not represent our subjective belief about the plausibility of future conditions, then some corrective measure should be taken.
-One possibility is to discard the $\vb{s}$ and resample from some ``true'' distribution $p_\mathrm{belief}$.
-In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
-Consequently, climate risk management often uses the outputs of structured simulations and analyses.
-For example, many studies use outputs from \gls{cmip} simulations \citep{meehl_cmip:2000} to assess risks and design adaptation strategies.\james{Add some refs}
-It is not practical or feasible for end users to re-run such complex Earth system models every time they want to sample \glspl{sow} from a different distribution.
-
-Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
-To do this we first project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
-This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
-We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ so that the weighted distribution of $\psi_j$ closely approximates $p_\mathrm{belief}(\Psi)$.
-
-We present here the case where the $\psi_j$ are one-dimensional.
-We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
-\begin{equation}\label{eq:weights}
-    w_j = \begin{cases}
-        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
-        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
-        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
-    \end{cases}
-\end{equation}
-This step is illustrated in \cref{fig:grid-sketch}.
-For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
-This approach is widely used in analysis of tropical cyclones: first an ensemble of cyclones is simulated, and then probabilistic weights are calculated for each based on attributes such as intensity, landfall location, and radius \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
-From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{grid-sketch}
-    \caption{
-        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
-        This method is illustrated for a hypothetical target distribution (orange line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (orange dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
-    }\label{fig:grid-sketch}
-\end{figure}
-
-The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis.
-As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-It is crucially important to understand that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be estimated if only we had better models and data.
-However, reasoning using subjective probability allows\ldots\james{Finish sentence}
+Estimates of flood hazard affect peoples' lives
+\begin{enumerate}
+    \item Floods affect lots of people and property
+    \item Climate change and exposure making floods worse
+    \item Floodplain management: $T$ year flood
+    \item Some cost-benefit analysis and lots of design standards / heuristics
+\end{enumerate}
+One example is floodproofing
+\begin{enumerate}
+    \item Literature on floodproofing and building-scale vulnerability reduction measures
+    \item This works great!
+    \item But who should use it? We're still using standards.
+    \item Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}.
+          Official guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent suggests scope for improvement.
+          For example, \citet{xian_elevation:2017} used a cost-benefit analysis to demonstrate that tailoring recommendations to the inital elevation of a structure can reduce expected costs.
+          \citet{zarekarizi_suboptimal:2020} show that neglecting uncertainty in discount rate, house lifespan, flood risk, and depth-damage curves can lead to maladaptation.
+          Yet these studies are silent on the question of how nonstationary flood hazard should factor into this decision.
+\end{enumerate}
+Nonstationarity
+\begin{enumerate}
+    \item These analyses all rely on \glspl{pdf}
+    \item Growing evidence for nonstationarity in a wide variety of hydroclimate processes
+    \item We often focus on climate change as a driver but sometimes local modifications matter just as much \citep{Merz:2014gf}
+    \item Even where risk-based analyses are used, this problem recommendations
+    \item Yet stationarity remains the dominant assumption because it is immortal \citep{Montanari:2014hl} or undead \citep{Serinaldi:2015bq}
+    \item Bonstationarity of flood hazard is already evident in many places, especially in coastal areas where \gls{msl} is clear.
+    \item On the other, efforts to find objective methods for quantifying nonstationarity, eg of precipitation \citep{atlas14_texas:2018} or flood frequency \citep{bulletin17c:2019}, have struggled to find a satisfactory approach because factors like future emissions and localized climate response create deep uncertainties that resist objective quantification
+\end{enumerate}
+It matters how you prepare
+\begin{enumerate}
+    \item Equifinality problem: different models diverge even if they fit the historical record well
+    \item Over- and under-design can each be costly \citep{DossGollin:2019}
+    \item On the one hand, overly optimistic assumptions of future risk can lead to unacceptably high risk of failure in the future if additional investments are not made.
+          For example, a failure to prepare Texas's interconnected natural gas, electricity, and water systems for temperatures that had been observed in the past 30 years caused cascading failures with disparate impacts \citep{doss-gollin_txtreme:2021,busby_cascadingrisks:2021}.
+    \item On the other, unless opportunity costs are negligible, designing for a too-pessimistic scenario can contribute to maladptation by limiting resource availability in the future.
+For example,\ldots\ldots\james{add an example of a maladaptive megaproject -- look to sustainability science or the arguments in \citet{ansar_bigisfragile:2017}, something by Flyvbjerg, \etc}
+\end{enumerate}
+A quick philosophical literature review of how people have approached and interpreted this problem
+\begin{enumerate}
+    \item This motivates subjective approaches for designing for nonstationarity.
+    \item There's a little bit of RDM-style work on floods \citep{sriver_sealevel:2018,garner_slrise:2018,lempert_slr:2012}
+    \item Model choice in $\mathcal{M}$-open case
+    \item Philosophy of Bayesian statistics and decision making -- explain subjective probability
+\end{enumerate}
+Case study
+\begin{enumerate}
+    \item House elevation in the coastal zone as a specific example of a problem where both over- or under-investment could be a problem
+    \item Federal guidance uses both standards (floodplain, \gls{bfe}) and cost-benefit analysis (does it pass test?)
+    \item Be sure to note that this has parallels to lots of other engineering design guidance.
+    \item \citet{xian_elevation:2017}: one size does not fit all
+    \item \citet{zarekarizi_suboptimal:2020}: neglecting uncertainty leads to bad outcomes
+\end{enumerate}
+Plan d'attaque
+\begin{enumerate}
+    \item Research questions
+          \begin{description}
+              \item[Q1] What are the potential benefits to homeowners of changing threshold-based guidance to risk-based guidance under nonstationary flood hazard?
+              \item[Q2] How does risk-based guidance change under different assumptions of physical model structure and human emissions forcings?
+              \item[Q3] How can deep uncertainties be transparently and consistently synthesized for decision analysis?
+          \end{description}
+    \item In \cref{sec:case-study} we describe our case study in Norfolk, VA.
+    \item  in \cref{sec:multiple-simulation} we analyze each simulation (\ie, realizations from these \glspl{pdf}) independently and illustrate .
+    \item  in \cref{sec:multiple-pdf} we consider the design problem as a set of ``multiple \glspl{pdf}'' and use these findings to motivate theoretical advances.
+    \item  in \cref{sec:synthesizing} we present a probabilistic method for synthesizing simulations from multiple \glspl{pdf} and illustrate the advantages and potential pitfalls of this approach.
+    \item  in \cref{sec:conclusions} we briefly summarize our findings, discuss implications for engineering practice, and identify  future research needs.
+\end{enumerate}
 
@@
 
-We model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
-We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
-For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
-We use the notation developed in the previous section to describe the case study.
-Specifically,
-\begin{enumerate}
-    \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
-    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
-    \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
-\end{enumerate}
-In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).
+We model a single, one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}), for a \emph{hypothetical} house in Norfolk, VA.
+For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as fixed quantities or as outputs from a single probabilsitic model, as shown in \cref{tab:uncertainties}.
+We evaluate this decision for each of many possible simulations of \gls{msl} using a combined engineering-economic model based on that of \citet{zarekarizi_suboptimal:2020}.
+In the remainder of this section we discuss data sources and models.
 
@@
     \caption{
-        Summary of parameters, their notation, and how their uncertainty is represented.
-        Symbols describing the decision-analytic framework are described in \cref{fig:flowchart}.
+        Summary of parameters, their notation, and how their uncertainty is modeled.
     }\label{tab:uncertainties}
-    \footnotesize
-    \begin{tabular}{p{1.25in} p{0.75in} p{3in}}
+    \begin{tabular}{l l p{3in} l}
         \toprule
-        Name                 & Symbol            & Uncertainty                                                                          \\
+        Name             & Symbol            & Uncertainty                                                                          & See more                     \\
         \midrule
-        \Gls{msl}            & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        \\
-        Storm surge          & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
-        Annual maximum flood & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
-        Discount rate        & $1-\gamma$        & Fixed at 2.5\% per year                                                              \\
-        Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
-        Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
-        Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
-        House floor area     & --                & Deterministic: \SI{1500}{ft^2}                                                       \\
-        Structural value     & --                & Deterministic: \usd{200000}                                                          \\
-        House lifespan       & $T$               & Deterministic: 71 years                                                              \\
+        \Gls{msl}        & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        & \cref{sec:sea-level}         \\
+        Storm surge      & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               & \cref{sec:storm-surge}       \\
+        Ann-max flood    & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          &                              \\
+        Discount rate    & $1-\gamma$        & Fixed at 3\%                                                                         & \cref{sec:led}               \\
+        Depth-damage     & $D(h-y)$          & Deterministic: based on HAZUS model                                                  & \cref{fig:cost-depth-damage} \\
+        Elevation cost   & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} & \cref{fig:cost-up-front}     \\
+        Initial height   & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe} unless otherwise noted                 &                              \\
+        House floor area &                   & Deterministic: \SI{1500}{ft^2} unless otherwise noted                                &                              \\
+        Structural value &                   & Deterministic: \usd{200000} unless otherwise noted                                   &                              \\
+        House lifetime   &                   & Deterministic: 70 years unless otherwise noted                                       &                              \\
         \bottomrule
     \end{tabular}
-    %TODO: check if there are other symbols used
 \end{table}
@@
     \centering
-    \includegraphics[width=\textwidth]{xlrm.pdf}
+    \includegraphics[width=4in]{xlrm.pdf}
     \caption{
-        Conceptual diagram of the considered example.
-        A \acrfull{sow} consists of a description of the uncertain factors (red).
-        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
-        For each \acrshort{sow} (red) and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (gray).
+        Conceptual diagram of the decision problem.
+        A \gls{sow} consists of a description of the uncertain factors (red).
+        We model a single decision (yellow).
+        For each combination of \gls{sow} and policy, the system model (blue) is used to calculate metrics (orange) describing the performance of the policy over the selected \gls{sow}.
     }\label{fig:xlrm}
@@
 
-\subsection{\Glsentrylong{slr}}\label{sec:case-slr}
+\subsection{Sea level rise}\label{sec:sea-level}
 
-We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
+Deeply uncertain sea level rise is the dominant source of deep uncertainty in our analysis.
+Decision makers and analysts often have access to simulations of physical variables like \gls{msl}, but usefully incorporating these simulations into decision analyses remains an area of active research.
+
+To inform the house elevation decision, we develop an ensemble of simulations of future \gls{msl}.
+Specifically, we analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic models using data published in \citet{ruckert_coastal:2019}.
 The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
 The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
-These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
-For a discussion of these model outputs we refer the reader to \citet{ruckert_coastal:2019}.
+These four models paramaterize physical processes, particularly of ice sheet dynamics, in different ways.
+For a discussion of these methods, their similarities, their differences, and their consistency with other analyses see \citet{ruckert_coastal:2019}, \citet{kopp_evolving:2017}, or \citet{bamber_slrise:2019}.
 
-Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \citep{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
-To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
+Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain.
+To represent this uncertainty, we sample four \gls{rcp} scenarios.
+This gives us sixteen time-varying probability density functions of \gls{msl}.
+We refer to each time series of \gls{msl}, $\overline{y}(t)$, as \emph{a \gls{sow}} and to each combination of physical model and \gls{rcp} scenario as \emph{a model} for $p(\overline{y}|t)$.
 
@@
     \caption{
-        Projections of future mean sea level depend strongly on the choices of physical model and forcing.
-        (A): 90\% confidence intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
+        Projections of future mean sea level depend strongly on the model (\gls{rcp} scenario and physical representation) used.
+        (A): 90\% intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
         (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
@@
 
-The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
-\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
+\Cref{fig:lsl-evolution}(a) shows the time-varying \glspl{pdf} of \gls{msl} for three representative models.
 The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
 \Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.
+Both physical model and \gls{rcp} scenario strongly inform the modeled probability distribution.\klaus{Do you want to add a sentence here comparing BRICK to the Kopp et al models?}
 
-\subsection{Storm surges}\label{sec:case-surge}
+\subsection{Storm surges}\label{sec:storm-surge}
 
-Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$.
+Coastal flood hazard does not depend only on \gls{msl} but also on storm surges.
+Neglecting hydrodynamics, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$ and annual maximum storm surges $y'(t)$.
+In this subsection we describe our representation of storm surge.
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\klaus{I can't find a better reference for this dataset!}
-Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
-For each calendar year we first remove the annual mean, then calculate the maximum water level.
-We refer this time series of annual maximum storm surges as $y'(t)$.
-We display this time series of annual maxima storm surges in \cref{fig:surge-obs-return}(a).
-The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.
+Data on storm surge comes from Sewells Point, VA (NOAA gauge 8638610).
+The NOAA tides and currents dataset is available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+Hourly recordings of water level are available from 1928 to the present; we use data from the period January 01 1928 to December 31 2021.
+For each year we first remove the annual mean, then extract the  maximum water level.
+We refer these maxima as the annual maximum storm surges $y'(t)$.
+
+\Cref{fig:surge-obs-return}(a) shows the time series of historical annual maximum storm surges.
+Major historical surges have been driven by two different mechanisms: tropical cyclones and nor'easters.
+The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and nor'easters have caused surges above \SI{6}{ft}.
 
@@
         time series of historic storms.
-        Red (yellow) arrows denote notable tropical cyclones (Nor'easters).
+        Red (yellow) arrows denote notable tropical cyclones (nor'easters).
         (B):
         return periods.
-        Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula (eq.~\ref{eq:weibull}).
-        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
+        Dots indicate observed values; their $x$-value (``plotting position'') is calculated using the Weibull formula (eq.~\ref{eq:plot-pos}).
+        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:storm-surge}).
     }\label{fig:surge-obs-return}
@@
 \begin{equation}\label{eq:surge-model}
-    y'(t) \sim \text{GEV}\qty(\mu, \sigma, \xi),
+    y'(t) \sim \text{GEV}\left(\mu, \sigma, \xi \right),
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
-This model assumes stationarity, neglecting any potential time dependence.
+where $y'(t)$ is the storm surge (above sea level) in year $t$ and $\mu$, $\sigma$, and $\xi$ are the location, scale, and shape parameters, respectively, of the \gls{gev} distribution.
+This model is stationary because there is no time dependence; surges in different years are modeled as \gls{iid}.
 
-Our approach to model building follows a principled workflow for model building and checking \citep[see][for details]{gelman_workflow:2020}.
-One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
-We include two forms of prior information.
-First, we constrain the shape parameter to be positive, $\xi > 0$, to reflect knowledge about the support of $y'$, which for a variable distributed according to \cref{eq:gev-dist} is:
-\begin{equation*}
-    \mathrm{supp}~ y' =
-    \begin{cases}
-        \xi < 0: & y' \in \qty(-\infty,~ \mu - \nicefrac{\sigma}{\xi}) \\
-        \xi > 0: & y' \in \qty(\mu-\nicefrac{\sigma}{\xi},~ \infty).
-    \end{cases}
-\end{equation*}
-Since storm surges cannot be negative, only the latter is physically defensible, justifying our choice to constrain the shape parameter to be positive.
+To improve estimates, we add twp forms of prior information.
+First, we constrain $\xi > 0$ to reflect knowledge about the support of $y'$: when $\xi < 0$, $y' \in \qty(-\infty, \mu - \nicefrac{\sigma}{\xi})$  and when $\xi > 0$, $y' \in \qty(\mu-\nicefrac{\sigma}{\xi}, \infty)$.
+Since storm surges cannot be negative, only the latter is physically defensible.
 Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
-Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-The parameters of the Inverse Gamma distribution can be calculated from these moments (see \cref{eq:inv-gamma-params}).
-These means and standard deviations were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
+Rather than applying prior information directly over the joint distribution of the priors $\qty{\mu,\sigma,\xi}$, for which intuition is difficult, we apply a prior over extreme quantiles of the distribution as in \citet{coles_evd:1996}.
+Specifically we apply half Normal (\ie, Normal distributions truncated at zero) priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+These values were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
 
-For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
+For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} through the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
 Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
-We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017}.
+We evaluate the model's fit using posterior predictive checks \citep[see section 2.4 of][and references therein]{gelman_workflow:2020}.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesisan test statistics, we found that draws from the posterior predictive distribution matched the observed test statistic credibly \cref{fig:surge-test-statistics}.
+Since \cref{fig:surge-test-statistics}(b) suggests the possibility of meaningful temporal structure not captured by our stationary \gls{iid} model, future efforts  could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020}.
 
 \Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position (see \cref{eq:weibull}).
-The good fit of the Bayesian fit (gray shading) to the data (points) suggest the modeling choices made are reasonable and sufficient for this didactic example.
-A positive control test (\cref{fig:surge-synthetic-data-experiment}) also validates the model's ability to recover known parameter values.
+The return period of the data (dots) is shown using the Weibull (``empirical'') plotting position
+\begin{equation}\label{eq:plot-pos}
+    \frac{r}{N + 1}
+\end{equation}
+where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
+The good fit of the Bayesian fit (gray shading) to the data (points) suggests that the stationary approximation is reasonable for this dataset.
 
-\subsection{Damages and metrics}\label{sec:case-metrics}
+\subsection{Expected annual damage}\label{sec:ead}
 
-The system model (``relationships'' in \cref{fig:xlrm}) is comprised of two key pieces.
-The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
-The second converts model converts a time series of annual expected damages into lifetime expected damages.
-
-We define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
+There are two components of the system model (``relationships'' in \cref{fig:xlrm}).
+The first is a fragility model that estimates the expected flood loss for a particular year, given the elevation of the house and the mean sea level for that year.
+Specifically, we define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
 This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
-The expected annual damage is thus
+The expcted annual damage is thus
 \begin{equation}\label{eq:ead}
-    \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'},
+    \textrm{EAD} = \mathbb{E}[D(h-\overline{y})] = \int_{y'} p(y') D(h - (\overline{y} + y')) \dd{y'}
 \end{equation}
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
+The term $\overline{y}$ in \cref{eq:ead} is time-dependent but we have omitted the time dependence for readability.
 Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
-(HAZUS) depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
-For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
-Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
+(HAZUS) depth-damage curves provided by FEMA to parameterize $D(h-\overline{y})$; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+For comparison, \cref{fig:cost-depth-damage} also shows results using the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
+Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for concision.
 
 The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
-However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
-Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient a surrogate model (see \cref{sec:surrogate-ead}).
+However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method:
+\begin{enumerate}
+    \item For $k=1, \ldots, K$:
+          \begin{enumerate}
+              \item sample a draw from the posterior distribution of flood hazard to get $\left\{ \mu_k, \sigma_k, \xi_k \right\}$
+              \item simulate a single storm surge from the stationary distribution (\cref{eq:surge-model}) and add the mean sea level to get total flood depth $y^\mathrm{sim}_k$
+              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $k$th damage.
+          \end{enumerate}
+    \item Estimate expected annual damages as the sample mean of the $K$ estimates
+\end{enumerate}
+Evaluating expected annual damages for each of $N$ simulations of sea level rise, each of $J$ draw from the posterior distribution of storm surge, and each of $T$ time steps requires $N \times J \times T \times K$ simulations.
+Since $K$ must be large in order to accurately approximate the integral, this requires incurring a heavy computational cost.
+However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
+To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
+We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
+This model is shown in \cref{fig:cost-expected-damage-emulator}.
+Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
 
-The second component of the system model converts a time series of $\mathrm{EAD}$ into lifetime expected damages, which we define as the up-front  discounted sum of expected annual damages:
+\subsection{Lifetime expected damages}\label{sec:led}
+
+The second component of the system model converts a time series of annual expected damages (\cref{sec:ead}) into lifetime expected damages, which we define as the up-front cost plus the discounted sum of expected annual damages
 \begin{equation}\label{eq:led}
-    \textrm{LED} = \sum_{t=t_i}^{t_f} \gamma^{\qty(t-t_i)-1} \textrm{EAD}(t),
+    \textrm{LED} = \sum_{t=1}^T \gamma^{t-1} \mathbb{E}[d_t | \Delta h],
 \end{equation}
-where the discount rate is $1 - \gamma$, the initial time $t_i=2022$, and the end time $t_f = t_i + T$.
-Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
-For a more theoretical discussion see \citet{arrow_discount:2013}.
+where the discount rate is $1 - \gamma$.
+For the didactic purposes of this study we neglect uncertainty in the discount rate \citep{zarekarizi_suboptimal:2020,arrow_discount:2013} and use a fixed discount rate (see \cref{tab:uncertainties}).
+Neglecting uncertainty in house lifetime, we model flood damages over a fixed house lifetime (\cref{tab:uncertainties}).
+We calculate lifetime expected damages \emph{for each \gls{sow} separately}.
 
-To assess the performance of a given decision for a specific \gls{sow} (``Metrics'' in \cref{fig:xlrm}), we calculate the following metrics for each decision-\gls{sow} combination:
-\begin{enumerate}
-    \item ``Up-front cost'' is the cost of elevating a house.
+To assess the performance of a given decision for a specific \gls{sow}, we calculate the following metrics \emph{for each decision-\gls{sow} combination}.
+\begin{description}
+    \item[up-front cost] is the cost of elevating a house.
 Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
@@
 This cost curve is shown in \cref{fig:cost-up-front}.
-    \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
-    \item ``Expected lifetime costs'' is the sum of lifetime expected damages and up-front costs.\james{Make sure the figures all use this term}
-\end{enumerate}
+    \item[lifetime expected damages] is discussed in \cref{eq:led}
+    \item[expected lifetime costs] is the sum of lifetime expected damages and up-front costs
+\end{description}
 
-\subsection{Prior over \glsentrylong{slr}}\label{sec:case-priors}
+\subsection{Experiment design}
 
-We construct three probabilistic models for $p_\text{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.\james{Expand a bit on other work and priors for SLR.}
-Refer to \citet{lempert_slr:2012,sriver_sealevel:2018,garner_slrise:2018}.
+In \cref{fig:flowchart} we introduce a general notation for simulation-based decision analysis.
+We consider $J$ \glspl{sow} representing time series of \gls{msl} and $I$ discrete decisions $x_1, \ldots, x_I$ representing the choice for how high to elevate the house.
+For each combination of these, we calculate the vector of metrics described in \cref{sec:led}, which we denote $u_{ij} = f(x_i, s_j)$, using a system model $f$.
+This system model caculates the metrics $u$ (which in this case is a vector) using the equations and relationships described in the preceding subsections.
+We begin in \cref{sec:multiple-simulation} by using the system model to explore the drivers of desirable and undesirable outcomes (blue box in \cref{fig:flowchart}).
+Next in \cref{sec:multiple-simulation} we consider scenario-conditional optimization and introduce ``the multiple \gls{pdf} problem'' (red box in \cref{fig:flowchart}).
+Finally in \cref{sec:synthesizing} we introduce a metnod for synthesizing uncertainty across the multiple \glspl{pdf} (orange box in \cref{fig:flowchart}).
 
-We use a Gamma distribution for all three priors, parameterized as
-\begin{equation*}
-    f(x | \alpha, \theta) = \frac{x^{\alpha-1} e^{-x/\theta}}{\Gamma(\alpha) \theta^\alpha},
-    \quad x > 0.
-\end{equation*}
-\Cref{tab:slr-priors} specifies these priors, as well as some quantiles of the distributions.
-Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
+\begin{figure}
+    \centering
+    \includegraphics[width=4in]{bayes-rdm.pdf}
+    \caption{
+        Outline of the proposed decision theoretic framework.
+        We begin in \cref{sec:multiple-simulation} using a lens of exploratory modeling (region 1) to analyze the elevation problem.
+        Next in \cref{sec:multiple-pdf} we use scenario-conditional probability distributions and the ``multiple \acrshort{pdf} problem'' (region 2).
+        Finally in \cref{sec:synthesizing} propose a method for weighting \glspl{sow} to synthesize across multiple \glspl{pdf} (region 3).
+    }\label{fig:flowchart}
+\end{figure}
 
-\begin{table}[h]
-    \centering
-    \caption{
-        Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\text{belief}(\psi)$.
-        The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles.
-    }\label{tab:slr-priors}
-    \input{../../plots/lsl_priors.tex}
-    % TODO: add another column: Parameter and Quantiles [in feet]
-\end{table}
+\section{Exploratory analysis}\label{sec:multiple-simulation}
 
-We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
-We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
-
-Future work could develop improved, policy-relevant priors, either by combining probabilistic models for emissions \citep[as in][]{srikrishnan_probabilistic:2022}, climate sensitivity, and the \gls{slr}, or through inversion of expert opinion \citep[as in][]{fuller_inversion:2017}.
-Future work could also consider more sophisticated priors, either by considering higher-dimensional $\psi$ or by modeling a joint prior over model inputs and model outputs through melding \citep{poole_melding:2000,sevcikova_melding:2007}.
-
-\section{Results and discussion}\label{sec:results}
-
-We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed synthesis method (\cref{sec:results-synthesis}).
-
-\subsection{Exploratory modeling}\label{sec:results-exploratory}
-
-We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions.
+We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions \citep[see][]{bankes:1993}.
+Exploratory modeling is used to explore the consequences of different possibilities, and does not require any assessment of how likely a given \gls{sow} may be.
 
@@
     \caption{
-        Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
+        Scenario maps showing the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of $\Delta h$ and $h_0$.
         Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
-        The lowest-cost outcomes occur when both exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
-        The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
+        The lowest-cost outcomes occur when $h_0$ is large, the house is not elevated (no up-front cost), and sea level rise is minimal.
+        The highest-cost outcomes arise when the house is not elevated (no up-front cost) and sea level rise is rapid.
         In all cases, elevating the house reduces the variance in total lifetime cost.
@@
 
-\james{Start with a summary sentence that outlines the paragraph}\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
-The outcomes with lowest total lifetime costs arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
-The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
-As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
+\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs) as a function of height increase ($\Delta h$) and sea level rise over the house lifetime.
+The best outcomes (yellow) arise when the house is not elevated ($\Delta h = 0$) and sea level rise is minimal (bottom left corners).
+The worst outcomes (blue) arise when the house is elevated only slightly and sea level rise is rapid.
+As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if sea level rise is substantial, damages will be negligible.
+Decisions about if and, if so, how high to elevate may thus depend on the decision-maker's risk aversion.
+Additionally, since exploratory modeling does not shed light on the likelihood of different \glspl{sow}, decisions about if and, if so, how high to elevate will also depend on an assessment of sea level rise likelihoods.
 
-This analysis answers ``what-if'' questions like ``given $h_0$ and $\Delta h$, what is the range of total costs a homeowner could face if \gls{slr} over the house lifetime is \SI{1}{ft} or \SI{10}{ft}.''
-For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \citep[\eg, those of][]{sweet_slr:2022} may prove sufficient for decision making.
-This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
-However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions.\james{Strong statement.
-Correct?
-What if we make or use strong implicit assumptions?}
+\section{Scenario-conditional probabilistic analysis: the multiple PDF problem}\label{sec:multiple-pdf}
 
-\subsection{Scenario-conditional optimization}\label{sec:results-conditional}
-We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
-Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draws from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
-As discussed in \cref{sec:analysis-condition}, this allows a probabilistic interpretation: conditional on a particular \gls{rcp} scenario and physical model, we can reason about the distribution of outcomes.
+The exploreatory analysis of the previous subsection considered each realization of future sea level to be an exploratory sample from the space of possible futures.
+An alternative interpretation of the available \glspl{sow} is as \gls{iid} draws from each of the sixteen models of sea level rise, as shown in \cref{fig:lsl-evolution}.
+This has the approach of a probabilistic interpretation: \emph{conditional on a particular model} (\gls{rcp} scenario and ice sheet model structure), we can view corresponding \glspl{sow} as \gls{iid} draws from the distribution of outcomes.
+
+A more formal notation is introduced in boxes (d) and (e) of \cref{fig:flowchart}.
+Specifically, for each of $K=16$ models, the samples of metrics are taken to be \gls{iid} samples from the distribution of outcomes (box e).
+Since we define these models to estimate $p(\Omega)$, our models $M_1, \ldots, M_K$ return a probability of $1 / J_K$ for each of the $J_K$ simulations from the $K$th model, and $0$ for simulations from all other models.
+We can then use this conditional distribution over outcomes to create metrics that depend only on the decision, such as averages or quantiles of outcomes.
 
@@
         Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
-        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kolpin_introduction:2017} under \gls{rcp} 8.5.
         (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
@@
 
-This probabilistic interpretation allows us to compute, for example, expected values.
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen models considered (we highlight three representative models).
-Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the total lifetime cost as a function of $\Delta h$ the sixteen models considered (three are highlighted).
+Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages.
 This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
@@
 
-This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
-This approach is attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
-However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from.
-The multiple \gls{pdf} problem has also been shown in other contexts.
-For example, \citet{sharma_rcp:2021}\klaus{According to Earth's Future this is in fact 2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.\james{Other examples: Wong et al 2017 -- levees.
-There are more!}
+Calculating either of the quantities shown in \cref{fig:tradeoffs-by-rcp} requires synthesizing across \glspl{sow}.
+However, these metrics are necessarily dependent on the model used (\ie, \gls{rcp} scenario and model of ice sheet dynamics).
+This means that any assessment of decision performance (\eg, an optimal strategy, a Pareto frontier, or a cost-benefit analysis) is conditional upon an (explicit or implicit) model of future outcomes.
+Where these models are implicit, they should be made more explicit to facilitate iterative model critique and improvement.
+
+Second, this approach presents what we call ``the multiple \gls{pdf} problem'' because it leaves decision makers with many probability distributions from which to choose.
+The multiple \gls{pdf} problem has been shown in other contexts.
+For example, \citet{sharma_rcp:2021} modeled the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that \emph{it places the burden for deciding which model to design for onto the end user.}
+Further, once a model is chosen, it will under-represent total uncertainty, since this depends on both with- and between-model variability.
 
-\subsection{Synthesizing deep uncertainties for decision analysis}\label{sec:results-synthesis}
+\section{Synthesizing PDFs for decision relevance}\label{sec:synthesizing}
 
-The proposed approach can help overcome the limitations of scenario-conditional analysis.
-We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
-We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
-These three distributions are shown in \cref{fig:lsl-priors-weights}(A).
+\citet{schneider_scenarios:2002} wrote, in the context of \gls{ipcc} assessments, that ``if we in the scientific assessment business do not offer some explicit notions of the likelihood of projected events, then the users of our products -- policy analysts and policy makers -- must guess what we think these likelihood estimates are.''
+Motivated by the multiple \gls{pdf} problem, we develop in this section a framework for synthesizing information across \glspl{pdf} using subjective Bayesian modeling.
+
+\subsection{Re-weighting}
+
+Specifically, we frame the set of \glspl{sow} through a sampling lens and consider the possibility that the distribution from which they are sampled, $p_\textrm{sampling}(\Omega)$, does not reflect our subjective belief about their true distribution, $p_\mathrm{belief}(\Omega)$.
+Indeed, this is often by design: it is often advantageous to sample low-probability but high-consequence \glspl{sow} to learn about a system's behavior \citep{bankes:1993}.
+To address this, we apply a scheme to weight each \gls{sow} based on $p_\mathrm{belief}(\Omega)$.
+First, we project the \glspl{sow} $s_1, \ldots, s_J$ onto a one-dimensional representation, which we denote $\psi_1, \ldots, \psi_J$.
+As discussed in \citet{csillery_abc:2010} and other research on approximate Bayesian computation, the choice of summary statistic is an important one.
+We take the summary statistic to be sea level rise from 2022 to 2100, recognizing that future work could consider multidimensional summaries.
+
+Without loss of generality, we assume the $\psi_j$ to be sorted from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
+This allows us to define $p_\mathrm{belief}(\psi \in \Psi)$ instead of $p_\mathrm{belief}(s \in \Omega)$.
+Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, \ie
+$$
+    F_\mathrm{belief}(\psi) = \int_{-\infty}^{s'} p_\mathrm{belief}(\psi) \dd{\psi},
+$$
+we calculate weights using \cref{eq:weights} as illustrated in \cref{fig:grid-sketch}.
+\begin{equation}\label{eq:weights}
+    w_j = \begin{cases}
+        j = 1     & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                   \\
+        1 < j < J & F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{true}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) \\
+        j = J     & 1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J]).
+    \end{cases}
+\end{equation}
+
+TODO\james{This paragraph needs some work}
+The key point to get across is what a subjective belief means in the Bayesian context \citep{savage:1954,gelman_philosophy:2013,gelman_subjectiveobjective:2017}.
+David Spiegelhalter had a nice comment on a podcast where he explained that in medicine, there isn't a ``true'' probability that, for example, you will get cancer; this is not some real quantity that can be measured \citep[hence the famous ``probability isn't real'' admonisment of][]{definetti_probability:1972}.
+Similarly, the probability distribution of sea level rise in 2100 isn't something that can be measured.
+However, probability provides a clear and self-consistent mathematical language for reasoning about uncertainty, which is why we're using it.
+A key advantage is that \textbf{since we can't be right, we can at least be transparent} about the assumed values of sea level in 2100, rather than hiding behind implicit assumptions.
+Thus we refer to the $M_k$ here as \emph{priors} to reflect that they consistently represent our current knowledge.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{grid-sketch}
+    \caption{
+        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights} for $J=4$.
+        This method is illustrated for a hypothetical target distribution (orange line) and four samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
+        As shown in \cref{eq:weights}, the weights $w_j$ (vertical blue lines) are calculated based on the \gls{cdf} of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+    }\label{fig:grid-sketch}
+\end{figure}
+
+Need to explain how we came up with our two or three subjective priors.\klaus{Are there published assessments we should refer to}
+Another advantage of this method is that it is computationally cheap to re-weight the \glspl{sow}.
+Thus, exploring performance under alternative priors is computationally cheap and easy to implement.
+\Cref{fig:tradeoffs-by-prior} shows the priors used in panel (a) and in panel (b) shows average lifetime cost as a function of the decision ($\Delta h$) after integrating over all possible \glspl{sow} (\ie, after synthesizing the ``multiple \glspl{pdf}.'')
+
+\subsection{Implications}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
+    \caption{
+        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the synthesizing Bayesian models.
+    }\label{fig:tradeoffs-by-prior}
+\end{figure}
+
+This technique can also be used as a diagnostic tool.
+In \cref{fig:lsl-priors-weights}, we can see that the ``Pessimistic'' (``Optimistic'') prior heavily weights \gls{rcp} 8.5 (2.6), which is unlikely given current policies \citep{hausfather_scenarios:2020}.\james{Discuss a bit more}
 
@@
         Subjective priors for local sea level.
-        We develop two distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
+        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{lsl} at Sewells Point, VA in 2100, relative to the present.
         The \glspl{pdf} of these subjective priors are shown in panel (A).
-        In panels (B-C) we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
-        Specifically, (B-C) show the average weight given to each model by each of the subjective priors.
+        In panels (B-D), we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
+        Specifically, (B-D) show the average weight given to each model by each of the three subjective priors.
+        The ``Slow SLR'' prior (B) places large weight on BRICK model \citep{wong_brick0.2:2017}, particularly the scenarios with no fast ice sheet dynamics and on low-emissions scenarios.
+        The ``Intermediate SLR'' (C) prior places approximately equal weight on all models.
+        The ``Fast SLR'' prior (D) places most of its weight on the \citet{deconto_antarctica:2016} model, and in particular on high-emissions scenarios.
     }\label{fig:lsl-priors-weights}
@@
 
-One application of this method is to diagnose which assumptions different $p_\text{belief}$ are consistent with.
-\Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-For example, the rapid \gls{slr} scenario places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
-Conversely, the slow \gls{slr} scenario places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
-The uncertain \gls{slr} scenario places approximately equal weight across models.
+\section{Discussion and conclusions}\label{sec:conclusions}
 
-This method can also be used to calculate expectations, allowing us to revisit the trade-off diagrams of \cref{fig:tradeoffs-by-rcp}.
-\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
-Notably, they give different guidance.\james{Say what this guidance is}
-Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} saves approximately 25\% of the house value relative to not elevating.
-Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} costs approximately 25\% of the house value relative to not elevating.
-Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house.
-Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
-    \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\text{belief}$ (colors).
-        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
-    }\label{fig:tradeoffs-by-prior}
-\end{figure}
-
-\section{Conclusions}\label{sec:conclusions}
-
-The motivation for this study was to develop a framework for increasing the transparency of assumptions about the likelihood of different \glspl{sow}.
-Additionally, we sought to develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
-Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-In this paper we proposed a decision analytic framework for synthesizing across deep uncertainties that is compatible with exploratory modeling.
-Through a case study of house elevation in the coastal zone, we show that:\klaus{I'd actually love feedback on these -- they can more or less parallel the key points but it would be really helpful for me to see what you think the key points are}
+Quick summary of what we have done
 \begin{enumerate}
-    \item estimated design trade-offs are highly sensitive to the choice of scenario;\klaus{This seems self-evident, though}
-    \item implicit or explicit assumptions about the likelihood of different scenarios are inevitable; and
-    \item the method proposed in \cref{sec:analysis-synthesize} can transparently synthesize across deep uncertainties.
+    \item OK
+\end{enumerate}
+Quick summary of what our motivation was
+\begin{enumerate}
+    \item Explore every \gls{sow}
+    \item Be transparent about how we are calculating trade-offs
+\end{enumerate}
+Philosophical links to Bayesian philosophy and model building in the $\mathcal{M}$-open case (NB: this is copied and pasted from my thesis -- need to rework)
+\begin{enumerate}
+    \item  \gls{bdt} was conceived as a calculus for reasoning rather than for identifying objective truth; \citeauthor{definetti_probability:1972} often said that ``probability does not exist'' \citeyear{definetti_probability:1972}.
+          \citet{savage:1954} and \citet{ramsey_probability:2016}, among others, also viewed probability as ``subjective,'' representing the state of belief of the decision-maker.
+    \item The famous phrase ``all models are wrong, but some are useful'' \citep[generally attributed to][]{box:1976} also suggests that probability distributions and predictions ought to be viewed subjectively.
+    \item More recent discussions of Bayesian philosophy \citep{jaynes_probability:2003,McElreath:2016vu,Gelman:2014tc,bernardo_bayesian:1994} also emphasize a philosophical view of probability as a language with which to reason about the unknown rather than a statement of objective truth \citep[see][for a thorough discussion of Bayesian philosophy]{gelman_philosophy:2013}.\klaus{Am I reading right that you say ``maybve have this in intro to motivate the analysis?''}
+    \item Although the true data generating process is not known and inference should not be represented as objective truth, probability gives a transparent and consistent language for reasoning about uncertainty.\james{This should go to intro}
+    \item Since modeling assumptions cannot overcome epistemic uncertainty, even with better models and more data, we draw from the literature on statistical model selection in the $\mathcal{M}$-open setting, which provides a theoretical background for choosing between models when the true data generating process is not among the models considered \citep[see][]{Piironen:2017eh}.\james{Explain this to the audience}
+    \item Often, combining inferences from multiple models is more effective than seeking a single ``best'' model \citep{Yao:2018bu}.
+          More fundamentally, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and critiquing them \citep{gelman_workflow:2020}.
+    \item Since, by definition, models are not ``true,'' this iterative workflow aims to identify models that are useful and promote a dialog amongst stakeholders \citep{gelman_philosophy:2013}
+    \item Our approach has interesting theoretical links to Bayesian model averaging or stacking \citep{Yao:2018bu}
+\end{enumerate}
+Links to DMDU literature
+\begin{enumerate}
+    \item We should also cite \citet{quinn_exploratory:2020}, which makes the point that it matters how you sample \glspl{sow}.
+    \item Exploratory modeling can help to demonstrate the existence of particular outcomes, generate hypotheses, build qualitative insight, and identify scenarios worthy of further study \citep[see][]{bankes:1993}.
+    \item You can do scenario discovery
+    \item We are in effect calculating robustness metrics \citep{mcphail_robustness:2019,herman:2015}
+    \item \citet{schneider_scenarios:2002} notes that a RDM analysis ``still requires the construction of ranges of future outcomes -- itself a subjective assessment.
+Subjectivity is simply inherent in all future projections, whether in natural or social sciences.''
 \end{enumerate}
 
-There are several limitations to our study that merit further discussion here.
-The first category has to do with limitations of the underlying methodology proposed for reweighting state of the world.
-For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional projection of the full time series, it is not a sufficient statistic.
-It is known that in approximate Bayesian computation, calibrating models on low dimensional statistics that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
-Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences.
-A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
-Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
-Another important point is that the gridding approach of \cref{eq:weights} implicitly assumes that the space is well spanned; this is not necessarily true but is a goal of exploratory modeling.\james{revisit}
-More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
-An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
-Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.\klaus{Is this too vague?}
-Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
+\subsection{Limitations and future work}
 
-The second category of limitations has to do with the case study and our interpretation of the house elevation decision problem.
-This problem intersects with decisions about where to live and how to manage household finances, both of which are highly complex.
-We could extend our analysis, for example, by considering additional decision objectives representing risk aversion, usable space created under the house, non-financial consequences of flooding, or other factors.
-We could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
-Finally, while here we consider the decision to be a one-time decision, it can also be framed as a sequential decision.
-Analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions.
-Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \citep{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \citep{herman_control:2020}.
+Limitations of the case study
+\begin{enumerate}
+    \item Objectives: real people might care about uncertainty (risk aversion), probability of experiencing flooding at all (disruptions are hard to quantify), usable space created under the house, and more
+    \item More uncertainty (damage functions, cost of construction, lifespan, discount rate, \etc)
+    \item Better data on cost of elevation and depth-damage
+    \item Robustness: ptimize separately for different kinds of house structures and locations
+    \item Timing
+\end{enumerate}
+Limitations of the method
+\begin{enumerate}
+    \item Our prior $p(\Omega)$ is limited -- we are just using \gls{lsl} in 2100 but we could be looking at more parameters of it including rate of change, \etc
+    \item We neglect true ignorance \citep{knight_risk:1921}.\james{Perhaps ``unknown unknowns''}
+    \item The real world is in a state of ``unknown unknowns'' \citep[level 5 as defined in][fig.~1]{walker_deep:2013} so trying to represent \emph{all} uncertainty is futile; we must make subjective modeling choices and assumptions about what is most important
+\end{enumerate}
+Future research needs:
+\begin{enumerate}
+    \item This didactic example illustrates a need for better synthesis and communication of deep uncertainties for decision making.
+    \item More complex models to capture more relevant metrics
+    \item Better models of nonstationary hazards
+    \item Interacting, sequential decisions
+\end{enumerate}
 
-These limitations motivate several directions for future research.
-From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \citep[see fig.~1 of][]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
-From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
-Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that scale to different house characteristics may add value.
+\subsection{Implications}
 
-The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate risk management.
-An obvious area of application is to the design of infrastructure for climate risk management.
-For example, much of the stormwater infrastructure in the United States is inadequate for current climate \citep{lopez-cantu:2018}, let alone future changes.
-Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \citep{sharma_rcp:2021} and \gls{rcp} scenarios.
-Similarly, decisions like  levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
-This framework can also be applied to uncertainties in non-physical variables.
-For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \citep{trindade_deeplyuncertainpathways:2019}.
-Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways,\james{cite} are conditional on probabilistic models for inputs like technology prices and population.
-
-Ultimataely, all models of the future are flawed.
-Seeking decisions that perform well even when assumptions are broken through robust design and flexibility generally improves outcomes.
-Yet whenever decisions are comppared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
-We call for researchers studying climate risk management to make these implicit assumptions, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\klaus{OK with this?}
+This approach can be used in many other contexts
+\begin{enumerate}
+    \item Direct
+          \begin{enumerate}
+              \item Stormwater management \citep{sharma_rcp:2021,lopez-cantu:2018}
+              \item Levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956}
+          \end{enumerate}
+    \item Indirect
+          \begin{enumerate}
+              \item Lots of economic models like social cost of carbon or effect of policy on metrics of interest make assumptions about deep uncertainties!\james{Let's emphasize this -- really important to policy analysis under deep uncertainty}
+              \item Model structure uncertainties
+          \end{enumerate}
+\end{enumerate}
 
@@
 
-This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
+This work was supported by the National Oceanic and Atmospheric Administration (NOAA) through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
 JDG thanks Rice University for support.
@@
 JDG wrote the codes and ran simulations.
-JDG wrote the manuscript in consultation with KK.
+JDG wrote the manuscript with support from KK.
+JDG wrote the manuscript in consulation with KK.
 JDG and KK revised the manuscript.
@@
 \newpage
-
-\printglossary[type=\acronymtype,title=List of Abbreviations]
-
-\section{Supplemental methods}
-
-\subsection{Storm surge statistical modeling}
-
-We parameterize the \gls{gev} distribution as
-\begin{equation}\label{eq:gev-dist}
-    f(x | \mu, \sigma, \xi)= \begin{cases}
-        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
-        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
-    \end{cases}
-\end{equation}
-
-Parameterizing the Inverse Gamma distribution as
-\begin{equation}\label{eq:inv-gamma-dist}
-    f(x | \alpha, \theta) = \frac{\theta^\alpha x^{-(\alpha + 1)}}{\Gamma(\alpha)} e^{-\frac{\theta}{x}}, \quad x > 0,
-\end{equation}
-the parameters $\alpha$ and $\theta$ can be computed from the desired mean $\mu$ and standard deviation $\sigma$ as
-\begin{align}\label{eq:inv-gamma-params}
-    \begin{split}
-        \alpha &= 2 + \frac{\mu^2}{\sigma^2} \\
-        \theta &= \mu \qty(\alpha - 1).
-    \end{split}
-\end{align}
-
-The plotting position used in \cref{fig:surge-obs-return} is the Weibull (``empirical'') plotting position
-\begin{equation}\label{eq:weibull}
-    \nicefrac{r}{N + 1}
-\end{equation}
-where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
-
-\subsection{Algorithm to estimate expected damages}\label{sec:alg-ead}
-
-We use a Monte Carlo integration to estimate expected damages as a function of house elevation and \gls{msl}, $D(h-y(t))$.
-\begin{enumerate}
-    \item For $m=1, \ldots, M$:
-          \begin{enumerate}
-              \item draw a sample from the posterior distribution of storm surge (see \cref{eq:surge-model})  $\qty{ \mu_m, \sigma_n, \xi_m}$
-              \item simulate a single storm surge from this stationary \gls{gev} distribution and add the mean sea level to get total flood depth $y^\mathrm{sim}_m$
-              \item calculate the flood damages for this draw by plugging the annual maximum flood depth ($h - y_k$) into  the deterministic HAZUS depth-damage relationship, storing this as the $m$th damage.
-          \end{enumerate}
-    \item Estimate expected annual damages as the sample mean of the $M$ estimates
-\end{enumerate}
-
-\subsection{Surrogate model for expected annual damages}\label{sec:surrogate-ead}
-Evaluating expected annual damages for each of $N$ simulations of \gls{slr}, each of $J$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
-In our model, we have $T=71$, $J=\num{10000}$, $N=\num{179232}$, and in general .
-However, noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
-To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
-We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
-This model is shown in \cref{fig:cost-expected-damage-emulator}.
-Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
-
 \section{Supplemental figures}
@@
         Depth-damage relationship.
-        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by \gls{fema}.
+        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by FEMA.
         Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
@@
     \caption{
-        As discussed in \cref{sec:case-metrics}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{msl}.
+        As discussed in \cref{sec:ead}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{lsl}.
         Damages ($y$ axis) are shown as a percentage of house value.
@@
         Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
-        See \cref{sec:case-surge} for details.
-        For the 2, 10, 100, and 500 year events we apply Inverse Gamma distributions, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+        See \cref{sec:storm-surge} for details.
+        For the 2, 10, 100, and 500 year events we apply Normal distributions, truncated at zero, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
     }\label{fig:surge-gev-priors}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-prior-return}
-    \caption{
-        Surge prior
-    }\label{fig:surge-prior-return}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-synthetic-data-experiment}
-    \caption{
-        Synthetic data experiment as a positive control test for the \gls{gev} model of storm surge.
-        A synthetic record was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively (dots).
-        These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:case-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
-        The blue line shows the true quantiles of the (known) \gls{gev} distribution.
-        By random chance the sample maximum has a true return period of $\gg 250$ years, which increases the upper confidence interval of the estimated return probabilities, but the true value is nevertheless within the 50\% posterior confidence interval.
-        This experiment yields similar results for alternative values of the known \gls{gev} distribution, and for different random seeds (not shown).
-    }\label{fig:surge-synthetic-data-experiment}
 \end{figure}
@@
     \caption{
-        Markov chain diagnostic plots for posterior draws from the storm surge model.
+        \Gls{mcmc} plots for posterior draws from the storm surge model.
         We draw \num{10000} samples by running four chains of \num{3500} iterations each and discarding the first \num{1000}.
@@
     \caption{
-        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:case-surge}).
+        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:storm-surge}).
         Each panel shows a different test statistic: partial autocorrelation at lags 1 and 2; sample maximum; sample minimum; sample median; and Mann-Kendall trend test statistic.
@@
     \caption{
-        Expected total lifetime cost (damages plus up-front cost) as a function of \gls{slr} over the house lifetime and height increase $\Delta h$.
+        Expected total lifetime cost (damages plus up-front cost) as a function of sea level rise over the house lifetime and height increase $\Delta h$.
         Initial house elevation is fixed to \SI{1}{ft} below the \gls{bfe}.
-        Expectations were computed for discrete values of $\Delta h$ ($x$ axis) by discretizing \glspl{sow} ($y$ axis), then taking the sample mean over each grid cell.
     }\label{fig:scenario-map-height-slr}
@@
     \caption{
-        Diagnostic statistics for the Hamiltonian Monte Carlo sampling for the storm surge posterior draws.
+        Diagnostic statistics for the \gls{mcmc} sampling for the storm surge posterior draws.
         Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.
@@
     }\label{tab:surge-posterior-mcmc-diagnostics}
-    \input{../../plots/surge-posterior-mcmc-diagnostics.tex}
+    %\input{../../plots/surge-posterior-mcmc-diagnostics.tex}
 \end{table}
@@
     \caption{As \cref{tab:surge-posterior-mcmc-diagnostics} but for draws from the prior distribution.}\label{tab:surge-prior-mcmc-diagnostics}
-    \input{../../plots/surge-prior-mcmc-diagnostics.tex}
+    %\input{../../plots/surge-prior-mcmc-diagnostics.tex}
 \end{table}

### 2022-05-24 papers/paper/preprint.tex
commit the plots

--- 2022-04-24 papers/paper/preprint.tex
@@
 \begin{abstract}
-    Methods for doing so, whether classical standards-based approaches or bottom-up robust methods, make implicit or explicit assumptions about the likelihood of different futures.
-    These assumptions may be inconsistent with expert belief.
-    In this paper we propose a probabilsitic method for synthesizing across deep uncertainties, using subjective prior beliefs.
-    The proposed method can be used in tandem with bottom-up exploratory methods and is consistent with fully probabilistic design criteria.
-    Through the lens of a didactic case study of house elevation in Norfolk, VA, we illustrate how this approach can improve the transparency of modeling assumptions.\klaus{Abstract needs a little work}
+    Needs work
+    \vspace*{10em}
 \end{abstract}
@@
 \begin{enumerate}
-    \item Estimated trade-offs between adaptation objectives are sensitive to assumptions about the likelihood of future conditions.
-    \item We propose a probabilistic method for transparently synthesizing across deep uncertainty in future conditions.
-    \item The proposed method identifies strategies for house elevation that are robust to assumptions in sea level rise.
+    \item
+    \item
+    \item
 \end{enumerate}
@@
 Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
-A critical question, then, is what conditions or futures particular infrastructure systems and components should be designed to withstand.
+This motivates the question of which of many possible futures futures infrastructure systems and components should be designed for.
 The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
-However, this decision is also subject to assumptions regarding the likelihhod and relevance of possible futures.
+However, this decision is also subject to assumptions regarding the likelihood and relevance of possible futures.\klaus{Vivek suggests spelling this out in a bit more detail}
 
@@
 Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
-These are often, though not always, based on probabilistic analysis of relevant data.
+These are often, though not always, informed by probabilistic analysis of relevant data.
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a requirement for homes in the 100 year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
+These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
 Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
 Design standards are also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the annual probability of overtopping to 1\%, neglecting the effects of sea level rise \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 
 There are many advantages to standards-based design.
-In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.\klaus{I understand this to mean that they are the same for everyone, even if they don't necessarily lead to equitable outcomes -- is that how you read this?}
+In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
 However, there are also limitations.
@@
 This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
-Like standards-based approaches, quantitative cost-benefit analyses rely on probabilistic descriptions of relevant hazard.
+These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.
 
-Currently used methods for estimating these \glspl{pdf} emphasize data-driven, objective methods that can be applied consistently across locations.
+Currently used methods for estimating these \glspl{pdf} emphasize data-driven, nominally objective methods that can be applied consistently across locations.
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
-Among several statistical assumptions these analyses make that have been recently called into question \citep[see][for a review]{Merz:2014gf} is stationarity.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are apparent or anticipated \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
+Among several statistical assumptions these analyses make that have been recently called into question \citep[see][for a review]{Merz:2014gf} is stationarity (the assumption that past and future hazard come from the same \gls{pdf}).
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
 While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
@@
 Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
-For example, robust decision making \citep{lempert_shaping:2003} runs models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
+For example, robust decision making \citep{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
 When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
@@
 When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
-Although some have argued\klaus{Can I get away w/o calling out a particular paper here?} that specifying a joint \gls{pdf} over inputs is inappropriate under deeply uncertain conditions, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
+While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
 This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
@@
 
-Recognizing that the assumptions underpinning any decision analysis will necessarily be subjective and imperfect, we draw inspiration from the philosophy of Bayesian model selection, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify (\ie, all models are approximations of the truth).
-This construct is a theoretical contrast to the classical approach to model selection in the ``$\mathcal{M}$-closed'' case, in which methods seek to asymptotically identify the ``true'' model from a candidate space.
-The theory of model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, this literature emphasizes the importance of iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
-Since, by definition, the models considered in an $\mathcal{M}$-open world are not ``true,'' this iterative workflow aims to identify models that are ``useful'' \citep{Box:1976gu} and promote a dialogue amongst stakeholders, decisionmakers, and analysts \citep{gelman_philosophy:2013}.
-Within this philosophical framework we find many parallels to the motivations and philosophies of \gls{dmdu}.
+We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
+\citet{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
+In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
+This challenge is described in the Bayesian literature on model selection, that is the problem of using competing statistical models to draw inferences, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify and all models are imperfect approximations of reality.
+Just as the literature on decision making under deep uncertainty contrasts with classical methods for decision analysis that identify an objectively ``best'' decision (or decision set) under a ``true'' description of uncertaintty, the literature on Bayesian model selection in the $\mathcal{M}$-open case contrasts with classical methods that seek to asymptotically identify a ``true'' model from a candidate space.
+In particular, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
+From a practical perspective, emphasis is placed on iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
 
@@
 We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,Rozer:2016dn,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
+Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
 Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently and consistently synthesized for decision analysis?}''
+Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling are averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993}by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
+Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993}by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes.\klaus{I could try to pull in some Monte Carlo equations here, but I think that would just be adding confusion}
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes.\klaus{Vivek suggests citing Jaynes here, anything in particular?}
 
@@
 If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
-This approach allows for rigorous quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
+This approach allows for quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
+% Vivek raises an interesting point -- what if we incorporate the choice of model into the data generating process and use a posterior probability of each model being true -- but I think that's basically what we are calling step #3
 
 We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
-For example, \citet{lempert_slr:2012} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
+For example, \citet{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
 Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
@@
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.\james{Build this out -- it's not so much the potential for poor decisions so much as that it can be hard to undwind the decision and understand what's driving it.
+We cant know if critical assumptions turned out to be poor if we don't know what they are or how changing them would result in a different analysis}
 
@@
 A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through kernel density estimation or other methods.
-Instead, paralleling joint probability methods for statistical analysis of tropical cycloines, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
+Instead, paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 
@@
 This step is illustrated in \cref{fig:grid-sketch}.
-For higher dimensional projections, this equation can be extended by partitioning the parameter space into a Voronoi diagram, and then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
+For higher dimensional projections, this equation can be extended by partitioning the parameter space, then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
 From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
+Diagnostic checks, such as examining the histogram of weights (not shown), may be valuable protections against degeneracy.
 
@@
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its uncertainty is epistemic -- that is, there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained (even with better models and data).
+We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained \citep{oreskes_verification:1994,walker_deep:2013}.
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
@@
 \begin{enumerate}
-    \item The decision vector $\vb{x}$ is a scalar describing how high to elevate a house ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
+    \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
     \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
@@
 Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,Farnham:2016tw,farnham_jetstream:2017,ossandon_semibayesian:2021}.
+Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
 
@@
 Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-In this paper we proposed a decision analytic framework for synthesizing across deep uncertainties that is compatible with exploratory modeling.
-Through a case study of house elevation in the coastal zone, we show that:\klaus{I'd actually love feedback on these -- they can more or less parallel the key points but it would be really helpful for me to see what you think the key points are}
-\begin{enumerate}
-    \item estimated design trade-offs are highly sensitive to the choice of scenario;\klaus{This seems self-evident, though}
-    \item implicit or explicit assumptions about the likelihood of different scenarios are inevitable; and
-    \item the method proposed in \cref{sec:analysis-synthesize} can transparently synthesize across deep uncertainties.
-\end{enumerate}
+In this paper we used a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
 
@@
 An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
-Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.\klaus{Is this too vague?}
+Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.
 Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
@@
 
-Ultimataely, all models of the future are flawed.
+Ultimately, all models of the future are flawed.
 Seeking decisions that perform well even when assumptions are broken through robust design and flexibility generally improves outcomes.
-Yet whenever decisions are comppared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
-We call for researchers studying climate risk management to make these implicit assumptions, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\klaus{OK with this?}
+Yet whenever decisions are compared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
+We call for researchers studying climate risk management to make these implicit assumptions explicit, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\james{This ends abruptly.
+What does it mean for design criteria to be appropriate given the subjective assessments?
+What are the implications?}
 
@@
 KK thanks Dartmouth College for support.
-The authors thank Tor Erlend Fjelde for helpful comments.
+The authors thank Tor Erlend Fjelde and Vivek Srikrishnan for helpful comments.
 
@@
         Observed values near the mode of the posterior predictive distribution are consistent with, but do not guarantee, a good fit.
-        For further discussion of posterior predictive checks, see Chapter 6 of \citet{Gelman:2014tc}.
+        For further discussion of posterior predictive checks, see Chapter 6 of \citet{gelman_bda3:2014}.
     }\label{fig:surge-test-statistics}

### 2022-06-17 papers/paper/preprint.tex
Klaus edits

@@
 \author[2]{Klaus Keller}
-\affil[1]{Department of Civil and Environmental Engineering, Rice University}
+\affil[1]{Department of Civil and Environmental Engineering, Rice University}https://www.overleaf.com/project/6165ef2754a6517ec4648164
 \affil[2]{Thayer School of Engineering, Dartmouth College}

### 2022-06-17 papers/paper/preprint.tex
start incorporating written edits from klaus

@@
 % Figures
-\usepackage{graphicx}
+\usepackage[demo]{graphicx}
 \usepackage[list=true]{subcaption}
@@
 \author[2]{Klaus Keller}
-\affil[1]{Department of Civil and Environmental Engineering, Rice University}https://www.overleaf.com/project/6165ef2754a6517ec4648164
+\affil[1]{Department of Civil and Environmental Engineering, Rice University}
 \affil[2]{Thayer School of Engineering, Dartmouth College}
@@
 \begin{abstract}
-    Needs work
-    \vspace*{10em}
+    Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand.
+    The
 \end{abstract}
@@
 
-Critical infrastructure services in America are aging and inadequate, leading to high-impact failures under conditions known to be possible \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020}.
-Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, and climate will further stress infrastructure systems.
-This motivates the question of which of many possible futures futures infrastructure systems and components should be designed for.
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021}.
-However, this decision is also subject to assumptions regarding the likelihood and relevance of possible futures.\klaus{Vivek suggests spelling this out in a bit more detail}
+Many critical infrastructure services around the world are aging and inadequate for a changing climate \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
+Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
+This motivates the question: which possible futures should infrastructure systems and components be designed for?
+The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021} and are often regulated by statue or industry guidelines.
+For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
+At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different future conditions.
 
@@
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
-Among several statistical assumptions these analyses make that have been recently called into question \citep[see][for a review]{Merz:2014gf} is stationarity (the assumption that past and future hazard come from the same \gls{pdf}).
+Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{milly_stationarity:2008,merz_review:2014}.
 For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.

### 2022-06-17 papers/paper/preprint.tex
address klaus handwritten edits

@@
 % Figures
-\usepackage[demo]{graphicx}
+\usepackage{graphicx}
 \usepackage[list=true]{subcaption}
@@
 \newacronym{gev}{GEV}{generalized extreme value}
+\newacronym{hazus}{HAZUS}{Hazard U.S.}
 \newacronym{iid}{IID}{independent and identically distributed}
@@
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
-Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{milly_stationarity:2008,merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
+Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{merz_review:2014}.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.\klaus{I'm citing the technical summary which is 96 pages -- there are about 100 authors listed.
+Any suggestion for how to cite best?}
 While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
@@
 When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
-For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
+For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \citep[see][p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
 While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
@@
 In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
-This challenge is described in the Bayesian literature on model selection, that is the problem of using competing statistical models to draw inferences, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify and all models are imperfect approximations of reality.
-Just as the literature on decision making under deep uncertainty contrasts with classical methods for decision analysis that identify an objectively ``best'' decision (or decision set) under a ``true'' description of uncertaintty, the literature on Bayesian model selection in the $\mathcal{M}$-open case contrasts with classical methods that seek to asymptotically identify a ``true'' model from a candidate space.
-In particular, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, emphasis is placed on iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
+We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
+This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \citep{bernardo_bayesian:1994}.
+Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013}, \cite{ramsey_probability:2016}, or \cite{jaynes_probability:2003} for a discussion of Bayesian philosophy and \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
+From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \citep{gelman_workflow:2020}.
 
 In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
-We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
+We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
 Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
 Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
+Focusing on deep uncertainty in \gls{slr} over the 70\james{Need to re-run codes with 70 rather than 71} year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
 
@@
 Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
-For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
+For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
 While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993}by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
+Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system modle, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.
+Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes.\klaus{Vivek suggests citing Jaynes here, anything in particular?}
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \citep{schneider_scenarios:2002}.
 
@@
 
-We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
+We can also use this approach to understand \gls{dmdu} methods that sample a set of parameters from fixed ranges.
 For example, \citet{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
@@
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.\james{Build this out -- it's not so much the potential for poor decisions so much as that it can be hard to undwind the decision and understand what's driving it.
-We cant know if critical assumptions turned out to be poor if we don't know what they are or how changing them would result in a different analysis}
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
+\james{Build this out -- it's not so much the potential for poor decisions so much as that it can be hard to undwind the decision and understand what's driving it.
+We cant know if critical assumptions turned out to be poor if we don't know what they are or how changing them would result in a different analysis.
+That said, I think they can also lead to poor decisions!}
 
@@
 
-If the ensemble of \glspl{sow} is not drawn \gls{iid} from the true distribution (\ie, if we deliberately over-sample low-probability but high-impact \glspl{sow}) or if we wish to synthesize \emph{across} multiple scenarios $M_1, M_2, \ldots, M_K$, then additional measures must be taken.
+If the \gls{sow} ensemble is not drawn \gls{iid} from the true distribution (\eg, if one deliberately over-samples low-probability but high-impact \glspl{sow}) or if one wishes to synthesize \emph{across} multiple scenarios $M_1, M_2, \ldots, M_K$, then additional measures must be taken.
 One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
@@
 Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
-A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through kernel density estimation or other methods.
-Instead, paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
-
+A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through methods such as kernel density estimation.
+Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
@@
         As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
+        % TODO: add the weights to the legend
+        % TODO: clarify colors! (NO, Klaus, you're not color blind -- but I am a little bit)
     }\label{fig:grid-sketch}
@@
 
-\section{Case study}\label{sec:case-study}
+\section{Demonstrating the concept with a case study}\label{sec:case-study}
 
-To illustrate the proposed decision analytic framework we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
+To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
 We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
-For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
+For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
 We use the notation developed in the previous section to describe the case study.
@@
     \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
-    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
+    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
     \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
@@
         Annual maximum flood & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
-        Discount rate        & $1-\gamma$        & Fixed at 2.5\% per year                                                              \\
+        Discount rate        & $\rho$            & Determinitic: 2.5\% per year                                                         \\
         Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
@@
         Structural value     & --                & Deterministic: \usd{200000}                                                          \\
-        House lifespan       & $T$               & Deterministic: 71 years                                                              \\
+        House lifespan       & $T$               & Deterministic: 70 years                                                              \\
         \bottomrule
@@
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\klaus{I can't find a better reference for this dataset!}
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
     \caption{
-        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
+        Annual maximum storm surges (after subtracting \acrlong{msl}) at Sewells Point, VA from NOAA tides and currents dataset available at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
         (A):
@@
         Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
+        % TODO: change caption for Obs (Weibull Plot Pos.) => Observations
+        % TODO: make the different confidence intervals slightly different colors instead of changing opacity
+        % TODO: move (B) to more logical location
     }\label{fig:surge-obs-return}
@@
 \end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
+where $y'(t)$ is the storm surge (above \gls{msl}) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
 This model assumes stationarity, neglecting any potential time dependence.
 
-Our approach to model validation is based on the concept of principled workflow design for model building and checking \citep[see][for details]{gelman_workflow:2020}.
+Our approach to model assessment is based on the concept of principled workflow design for model building and checking \citep[see][for details]{gelman_workflow:2020}.
 One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
@@
 Rather than applying prior information directly over the joint distribution of the parameters $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
-Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
+Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means of \SIlist{4;6;10;15}{ft} and standard deviations of \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
 The parameters of the Inverse Gamma distribution can be calculated from these moments (see eq.~\ref{eq:inv-gamma-params}).
-These means and standard deviations were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
+These means and standard deviations were chosen to represent plausible physical ranges (\cref{fig:surge-gev-priors}).
 
@@
 We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
 Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
 
-\Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull plotting position (see \cref{eq:weibull}).
-The good fit of the Bayesian model to the data suggest the modeling choices made are reasonable and sufficient for this didactic example.
-A positive control test (\cref{fig:surge-synthetic-data-experiment}) also validates the model's ability to recover known parameter values.
+Other model validations lend confidence to the stationary \gls{gev} model selected.
+For example, \cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges; the estimated return period of the (shading) matches the (dots) the empirical plotting position dots.
+Further, a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
 
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
-(HAZUS) depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+Following \citet{zarekarizi_suboptimal:2020}, we use the we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
 For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
@@
 \begin{equation}\label{eq:led}
-    \textrm{LED} = \sum_{t=t_i}^{t_f} \gamma^{\qty(t-t_i)-1} \textrm{EAD}(t),
+    \textrm{LED} = \sum_{t=t_i}^{t_f} \gamma^{\qty(t-t_i)} \textrm{EAD}(t),
 \end{equation}
-where the discount rate is $1 - \gamma$, the initial time $t_i=2022$, and the end time $t_f = t_i + T$.
+where  $\gamma = 1- \rho$ ($\rho$ being the discount rate), the initial time $t_i=2022$, and the end time $t_f = t_i + T - 1$.
 Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
@@
 The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
-As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because,.
-even if \gls{slr} is substantial, damages will be negligible.
+As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because even if \gls{slr} is substantial, damages will be negligible.
 
@@
         Colored lines highlight three representative models for emphasis.
+        % TODO: add arrows and star
     }\label{fig:tradeoffs-by-rcp}
@@
 Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
-This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
+For small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics; red lines) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model; blue lines).
 For example, under the most pessimistic model (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
@@
     \caption{
-        Subjective priors for local sea level.
-        We develop two distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
+        Impact of subjective priors for local sea level on implicit weight given to each \gls{rcp} scenario and physical model.
+        We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
         The \glspl{pdf} of these subjective priors are shown in panel (A).
@@
 \Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-For example, the rapid \gls{slr} scenario places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
-Conversely, the slow \gls{slr} scenario places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
-The uncertain \gls{slr} scenario places approximately equal weight across models.
+For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
+Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
+The uncertain \gls{slr} scenario (blue line) places approximately equal weight across models.
 
@@
     \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\mathrm{belief}$ (colors).
+        As \cref{fig:tradeoffs-by-rcp}, but with Pareto frontiers for the full distribution of outcomes using the three models of $p_\mathrm{belief}$ (colors).
         Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
+        % TODO: maybe we don't need the gray lines, make the red/yellow/blue lines more bold.
     }\label{fig:tradeoffs-by-prior}
@@
 
-The motivation for this study was to develop a framework for increasing the transparency of assumptions about the likelihood of different \glspl{sow}.
-Additionally, we sought to develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
+This study develops a framework to increase the transparency of assumptions about the likelihood of different \glspl{sow}.
+Additionally, we develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
 Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-In this paper we used a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
+We use a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
 
-There are several limitations to our study that merit further discussion here.
+There are several limitations to our study that merit further discussion.
 The first category has to do with limitations of the underlying methodology proposed for reweighting state of the world.
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional projection of the full time series, it is not a sufficient statistic.
-It is known that in approximate Bayesian computation, calibrating models on low dimensional statistics that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
-Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences.
+Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
+Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences not represented in our model.
 A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
-Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
-More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
-An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
-Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.
+Although this is appropriate for our didactic illustration, recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} can be applied to improve decision making in real world case studies.
+More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response \citep{morgan_elicitation:2014}.
+An alternative approach would be to build a probabilistic model for each of these steps, and to use each as input to the next to develop a fully probabilistic model for \gls{slr}.
+Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022,wong_brick0.2:2017}, this remains a computational and conceptual challenge.
 Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
@@
 This problem intersects with decisions about where to live and how to manage household finances, both of which are highly complex.
-We could extend our analysis, for example, by considering additional decision objectives.
+One extension of our analysis would be to consider additional decision objectives.
 In particular, we hypothesize that incorporating improved representations of risk aversion into decision support may substantially improve their usability.
-We could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
-Finally, while here we consider the decision to be a one-time decision, it can also be framed as a sequential decision.
-Analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions.
+One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
+Finally, while here we consider the decision to be a one-time decision, one could also frame this as problem a sequential decision.
+The aaalysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \citep{herman_control:2020}.
 Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \citep{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \citep{herman_control:2020}.
@@
 
-The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate risk management.
+The proposed scenario weighting framework can be applied to inform critical challenges in climate risk management.
 An obvious area of application is to the design of infrastructure for climate risk management.
@@
 
-Ultimately, all models of the future are flawed.
-Seeking decisions that perform well even when assumptions are broken through robust design and flexibility generally improves outcomes.
-Yet whenever decisions are compared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
-We call for researchers studying climate risk management to make these implicit assumptions explicit, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\james{This ends abruptly.
-What does it mean for design criteria to be appropriate given the subjective assessments?
-What are the implications?}
+Of course, all models are XYZ.\james{Get feedback from Klaus here}
+Seeking decisions that perform well across a range of assumptions through robust design and flexibility can improve outcomes.
+Yet whenever decisions are compared quantitatively, assumptions about the probability of different possible futures are necessarily made.
+We call for researchers studying climate risk management to make these implicit assumptions explicit, and we suggest that coordinated guidance can help practitioners determine better design criteria.
 
@@
 Evaluating expected annual damages for each of $J$ simulations of \gls{slr}, each of $N$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
-In our model, we have $T=71$, $N=\num{10000}$, $J=\num{179232}$, and so exhaustive sampling may be prohibitive.
+In our model, we have $T=70$, $N=\num{10000}$, $J=\num{179232}$, hence exhaustive sampling may be prohibitive.
 
@@
         Depth-damage relationship.
-        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by \gls{fema}.
+        Following \citet{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves.
         Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
@@
     \caption{
-        Surge prior
+        Implicit prior over storm surge represents large uncertainty but is consistent with basic physical reasoning (see \cref{sec:case-surge}).
+        The prior median (blue line) and uncertainty quantiles (shading) of the return period for different return levels of storm surge at Sewells Point, VA using prior predictive sampling \citep{gelman_workflow:2020}.
+        This illustrates the weakly informative prior information (\cref{sec:case-surge}) used in the analysis.
+        These weakly informative priors were selected to be consistent with basic physical principles (\eg, storm surges most years are $>$ \SI{2}{ft} but rarely exceed \SI{20}{ft}).
+        % TODO: fix legend to say Prior Median
     }\label{fig:surge-prior-return}
@@
         Synthetic data experiment as a positive control test for the \gls{gev} model of storm surge.
-        A synthetic record was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively (dots).
+        A synthetic record (dots) was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively.
         These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:case-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
@@
     \caption{
-        As \cref{fig:surge-posterior-chains} but for draws from the prior distribution.
+        As \cref{fig:surge-posterior-chains}, but for draws from the prior distribution.
     }\label{fig:surge-prior-chains}
@@
         Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.
-        In general, a $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence.
+        The naive standard error (SE) returns the standard error of the mean.
+        The Monte Carlo standard error (MCSE) is calculated using the initial monotone sequence estimator \citep[][pp.~473-483]{geyer_practical:1992}.
+        The effective sample size (ESS) is a crude estimate of the number of independent samples and is calculated following \citet[][pp.~473-483]{geyer_practical:1992}.
+        The $\hat{R}$ is an indicator of the convergence of the Markov chains to the target distribution; a value $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence \citep{mcelreath_rethinking2:2020}.
+        For additional details see the \texttt{MCMCDiagnosticTools} package documentation.
     }\label{tab:surge-posterior-mcmc-diagnostics}
-    %\input{../../plots/surge-posterior-mcmc-diagnostics.tex}
+    \input{../../plots/surge-posterior-mcmc-diagnostics.tex}
 \end{table}
@@
     \caption{As \cref{tab:surge-posterior-mcmc-diagnostics} but for draws from the prior distribution.}\label{tab:surge-prior-mcmc-diagnostics}
-    %\input{../../plots/surge-prior-mcmc-diagnostics.tex}
+    \input{../../plots/surge-prior-mcmc-diagnostics.tex}
 \end{table}

### 2022-06-17 papers/paper/preprint.tex
Updates from Overleaf
work on follow up email with klaus

@@
 \newacronym{rcp}{RCP}{representative concentration pathway}
+\newacronym{rdm}{RDM}{robust decision making}
 \newacronym{slr}{SLR}{sea level rise}
@@
 
-% up to 1250 words
 \begin{document}
@@
 
-Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
+Following prior work \citep[\eg,][]{garner_slrise:2018,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the freely available NOAA tides and currents data archive \citep{noaa_tidesandcurrents:2022}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
     \caption{
-        Annual maximum storm surges (after subtracting \acrlong{msl}) at Sewells Point, VA from NOAA tides and currents dataset available at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.
+        Annual maximum storm surges (after subtracting \acrlong{msl}) at Sewells Point, VA from the freely available NOAA tides and currents data archive \citep{noaa_tidesandcurrents:2022}.
         (A):
@@
 Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
+Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
 
@@
 We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
-We can also compare to the analyses of \citet{lempert_slr:2012,sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
+We can also compare to the analyses of \citet{sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
 Our samples bound all of these estimates.
@@
 This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
+Scenario mapping is also commonly used in \gls{rdm} analyses \citep{lempert_rdm:2019,sriver_sealevel:2018}.
 However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
@@
 Similarly, \citet{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
+And \citet{lempert_rdm:2019}
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
@@
 
-Of course, all models are XYZ.\james{Get feedback from Klaus here}
-Seeking decisions that perform well across a range of assumptions through robust design and flexibility can improve outcomes.
+Of course, all models are ultimately wrong \citep{box:1976}.
+Thus seeking decisions that perform well across a range of assumptions, and improving the deccision space through robust design and flexibility, can improve outcomes.
 Yet whenever decisions are compared quantitatively, assumptions about the probability of different possible futures are necessarily made.
@@
 
+% up to 1250 words
 \end{document}

### 2022-06-22 papers/earths-future/submission.tex
Format to prepare for Earth's Future submission

--- 2022-06-17 papers/paper/preprint.tex
@@
-\documentclass[11pt]{article}
+\documentclass[draft]{agujournal2019}
+
+%% required packages
+\usepackage{lineno}
+\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
 
@@
 \usepackage{amsmath, amsfonts, physics, siunitx, nicefrac}
-\usepackage[american]{babel} 
-\usepackage[T1]{fontenc} % improved font encoding
-\usepackage[ttscale=0.8]{libertine}
-\usepackage{fontawesome5}
-\usepackage[format=plain, textfont=it]{caption}
-
-% page size and margins
-\usepackage{geometry}
-\geometry{letterpaper,top=1in, bottom=1in, left=1in, right=2in}
-
-% footer
-\usepackage{fancyhdr}
-\usepackage{lastpage}
-\usepackage[en-US]{datetime2}
-\fancyhf{}
-\fancyhead[L]{CONFIDENTIAL DRAFT by J. Doss-Gollin \& K. Keller}
-\fancyhead[R]{\DTMnow}
-\fancyfoot[R]{page~\thepage~of~\pageref{LastPage}}
-\pagestyle{fancy}
-\setlength{\headheight}{15pt}
-
-% TO DO NOTES
-\usepackage{xcolor} % list of colors at https://en.wikibooks.org/wiki/LaTeX/Colors
-\definecolor{giallo}{HTML}{F0BC42} % https://teamcolorcodes.com/a-s-roma-color-codes/
-\definecolor{rosso}{HTML}{8E1F2F}
-\definecolor{grigio}{HTML}{CACACC}
-\definecolor{nero}{HTML}{000000}
-\usepackage[textsize=scriptsize]{todonotes}
-\setlength{\marginparwidth}{1.5in}
-\newcommand{\james}[1]{\todo[color=giallo, textcolor=nero]{\textbf{ATTN James:~}#1}} % if desired create a custom command for each author
-\newcommand{\klaus}[1]{\todo[color=rosso, textcolor=grigio]{\textbf{ATTN Klaus:~}#1}}
 
@@
 \newcolumntype{L}[1]{>{\PreserveBackslash\raggedright}p{#1}}
-
-% better lists
-\usepackage[inline]{enumitem}
-\setlist{nosep}
-
-% authors
-\usepackage{authblk}
-\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
-\author[1]{James Doss-Gollin}
-\author[2]{Klaus Keller}
-\affil[1]{Department of Civil and Environmental Engineering, Rice University}
-\affil[2]{Thayer School of Engineering, Dartmouth College}
-\renewcommand*{\Affilfont}{\normalsize\normalfont}
 
@@
 \newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}
-\makeglossaries % required to make the list of acronyms
 
@@
 \def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+\newcommand{\usd}[1]{\SI{#1}[\$]{}}
 
-\usepackage{xspace}
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\newcommand{\usd}[1]{\SI{#1}[\$]{}}
-\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
-\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
-\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+% cite things in supplemental info file
+\usepackage{xr}
+\externaldocument{supplemental}
 
-% use biblatex
-\usepackage{csquotes}
-\usepackage[
-  backend=biber,
-  doi=true,
-  url=false,
-  isbn=false,
-  style=authoryear-comp,
-  natbib=true,
-  backref=false,
-  maxbibnames=10,
-  maxcitenames=2,
-  uniquename=false,
-  uniquelist=false,
-  sorting=nyt,
-  giveninits=true,
-]{biblatex}
-\renewbibmacro{in:}{}
-\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
-\addbibresource{library.bib}
-
-% load this last
-\usepackage[hidelinks]{hyperref}
+% this goes last
 \usepackage{cleveref}
 
+\linenumbers
+%%%%%%%
+% As of 2018 we recommend use of the TrackChanges package to mark revisions.
+% The trackchanges package adds five new LaTeX commands:
+%
+%  \note[editor]{The note}
+%  \annote[editor]{Text to annotate}{The note}
+%  \add[editor]{Text to add}
+%  \remove[editor]{Text to remove}
+%  \change[editor]{Text to remove}{Text to add}
+%
+% complete documentation is here: http://trackchanges.sourceforge.net/
+%%%%%%%
+
+\draftfalse
+
+%% Enter journal name below.
+%% Choose from this list of Journals:
+%
+% JGR: Atmospheres
+% JGR: Biogeosciences
+% JGR: Earth Surface
+% JGR: Oceans
+% JGR: Planets
+% JGR: Solid Earth
+% JGR: Space Physics
+% Global Biogeochemical Cycles
+% Geophysical Research Letters
+% Paleoceanography and Paleoclimatology
+% Radio Science
+% Reviews of Geophysics
+% Tectonics
+% Space Weather
+% Water Resources Research
+% Geochemistry, Geophysics, Geosystems
+% Journal of Advances in Modeling Earth Systems (JAMES)
+% Earth's Future
+% Earth and Space Science
+% Geohealth
+%
+% ie, \journalname{Water Resources Research}
+
+\journalname{Earth's Future}
+
 \begin{document}
-\maketitle
-\thispagestyle{empty}
+
+%% ------------------------------------------------------------------------ %%
+%  Title
+%
+% (A title should be specific, informative, and brief. Use
+% abbreviations only if they are defined in the abstract. Titles that
+% start with general keywords then specific terms are optimized in
+% searches)
+%
+%% ------------------------------------------------------------------------ %%
+
+% Example: \title{This is a test title}
+
+\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  AUTHORS AND AFFILIATIONS
+%
+%% ------------------------------------------------------------------------ %%
+
+% Authors are individuals who have significantly contributed to the
+% research and preparation of the article. Group authors are allowed, if
+% each author in the group is separately identified in an appendix.)
+
+% List authors by first name or initial followed by last name and
+% separated by commas. Use \affil{} to number affiliations, and
+% \thanks{} for author notes.
+% Additional author notes should be indicated with \thanks{} (for
+% example, for current addresses).
+
+% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
+% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
+
+\authors{James Doss-Gollin\affil{1}, Klaus Keller\affil{2}}
+\affiliation{1}{Department of Civil and Environmental Engineering, Rice University}
+\affiliation{2}{Thayer School of Engineering, Dartmouth College}
+
+%% Corresponding Author:
+% Corresponding author mailing address and e-mail address:
+
+% (include name and email addresses of the corresponding author.  More
+% than one corresponding author is allowed in this LaTeX file and for
+% publication; but only one corresponding author is allowed in our
+% editorial system.)
+
+% Example: \correspondingauthor{First and Last Name}{email@address.edu}
+
+\correspondingauthor{James Doss-Gollin}{jdossgollin@rice.edu}
+
+%% Keypoints, final entry on title page.
+
+%  List up to three key points (at least one is required)
+%  Key Points summarize the main points and conclusions of the article
+%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+
+% Example:
+% \begin{keypoints}
+% \item	List up to three key points (at least one is required)
+% \item	Key Points summarize the main points and conclusions of the article
+% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
+% \end{keypoints}
+
+\begin{keypoints}
+  \item Decisions about how to manage climate risks are often made under deep uncertainty
+  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
+  \item We demonstrate the framework using a simple case study of coastal flood risk management
+\end{keypoints}
+
+%% ------------------------------------------------------------------------ %%
+%
+%  ABSTRACT and PLAIN LANGUAGE SUMMARY
+%
+% A good Abstract will begin with a short description of the problem
+% being addressed, briefly describe the new data or analyses, then
+% briefly states the main conclusion(s) and how they are supported and
+% uncertainties.
+
+% The Plain Language Summary should be written for a broad audience,
+% including journalists and the science-interested public, that will not have 
+% a background in your field.
+%
+% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
+% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
+% see http://sharingscience.agu.org/creating-plain-language-summary/)
+%
+%% ------------------------------------------------------------------------ %%
+
+%% \begin{abstract} starts the second page
 
 \begin{abstract}
-    Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand.
-    The
+  Projections of future climate risks can vary considerably between sources without a  consensus about their  relative probability.
+  This deep uncertainty poses considerable communication and decision-analytical challenges.
+  One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
+  Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
+  These approaches (i) can underestimate risks, (ii) provide little insights into which assumptions drive decision-relevant outcomes, and (iii) can nudge how people choose their decisions-making frameworks.
+  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform decision support.
+  The first step of this workflow is to create an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
+  Next, a small set of probability distributions representing subjective beliefs over possible futures is used to weight the scenarios.
+  Finally, these weights are used to compute and characterize trade-offs, conduct robustness checks, and reveal implicit assumptions.
+  We demonstrate the framework through a didactic case study analyzing how high to elevate a house to manage coastal flood risks.
 \end{abstract}
 
-Key points
-\begin{enumerate}
-    \item
-    \item
-    \item
-\end{enumerate}
+\section*{Plain Language Summary}
 
-\clearpage
+What are sound strategies to manage risks driven by climatic changes?
+Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems.
+These uncertainties often arise from choices experts have to make, for example about technical detail on how to model future water levels.
+Different experts can disagree about these choices and provide different plausible projections.
+Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
+For one, just picking one of these projections or just an average can lead to poor decisions.
+On the other extreme, communicating results separately for all projections can overwhelm decision-makers.
+To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
+We develop and demonstrate a framework to address these challenges.
+The framework provides a transparent approach (i) to combine a large number of deeply uncertain projections to a more understandable and smaller sample and (ii) to provide insights as to how assumptions and modeling choices made by experts influence decisions.
+We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections about future water levels.
+
+%% ------------------------------------------------------------------------ %%
+%
+%  TEXT
+%
+%% ------------------------------------------------------------------------ %%
+
 \section{Introduction}\label{sec:introduction}
 
-Many critical infrastructure services around the world are aging and inadequate for a changing climate \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
+Many critical infrastructure services around the world are aging and inadequate for a changing climate \cite<\eg,>[]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
 Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
 This motivates the question: which possible futures should infrastructure systems and components be designed for?
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021} and are often regulated by statue or industry guidelines.
+The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statue or industry guidelines.
 For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
@@
 
-Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
+Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
 These are often, though not always, informed by probabilistic analysis of relevant data.
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
+These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
+Additionally, local building codes \cite<based on guidance such as>[]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
 Design standards are also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 
@@
 In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
-This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
+This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
 These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.
@@
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
-Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.\klaus{I'm citing the technical summary which is 96 pages -- there are about 100 authors listed.
-Any suggestion for how to cite best?}
-While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
+Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
+While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
 Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
@@
 
-Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016}.
-In other words, they are deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
+Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet; see>[]{deconto_antarctica:2016}.
+In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
 These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
@@
 Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
-Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
-For example, robust decision making \citep{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
-When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-In addition, many studies formally quantify robustness to deep uncertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
+Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \cite{bankes:1993}.
+For example, robust decision making \cite{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
+When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+In addition, many studies formally quantify robustness to deep uncertainties \cite{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
 
@@
 When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
-For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \citep[see][p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
-While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
+While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
 This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
@@
 We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
-\citet{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
+\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
 In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
 We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
-This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \citep{bernardo_bayesian:1994}.
+This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \cite{bernardo_bayesian:1994}.
 Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013}, \cite{ramsey_probability:2016}, or \cite{jaynes_probability:2003} for a discussion of Bayesian philosophy and \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \citep{gelman_workflow:2020}.
+From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \cite{gelman_workflow:2020}.
 
@@
 We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
-Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 70\james{Need to re-run codes with 70 rather than 71} year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
+Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \cite{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \cite{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \cite{FEMA_p-55:2011} require elevation in some cases.
+Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \cite{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \cite{zarekarizi_suboptimal:2020}.
+Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
-Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.
+Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
+Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
+Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \cite{quinn_exploratory:2020}.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \citep{schneider_scenarios:2002}.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite{schneider_scenarios:2002}.
 
@@
 If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
-This approach allows for quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
+This approach allows for quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
 % Vivek raises an interesting point -- what if we incorporate the choice of model into the data generating process and use a posterior probability of each model being true -- but I think that's basically what we are calling step #3
@@
 We can also use this approach to understand \gls{dmdu} methods that sample a set of parameters from fixed ranges.
-For example, \citet{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
-Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
+For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
+Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \cite<\ie, robustness metrics; see>[for details]{mcphail_robustness:2019,herman:2015}.
 Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.
-\james{Build this out -- it's not so much the potential for poor decisions so much as that it can be hard to undwind the decision and understand what's driving it.
-We cant know if critical assumptions turned out to be poor if we don't know what they are or how changing them would result in a different analysis.
-That said, I think they can also lead to poor decisions!}
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to inscrutable decision processes and suboptimal decisions.
 
@@
 A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through methods such as kernel density estimation.
-Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
+Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
@@
     Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
-        This method is illustrated for a hypothetical target distribution (orange line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (orange dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
-        % TODO: add the weights to the legend
-        % TODO: clarify colors! (NO, Klaus, you're not color blind -- but I am a little bit)
+    This method is illustrated for a hypothetical target distribution (dark red line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
+    As shown in \cref{eq:weights}, the weights $w_j$ (orange vertical lines) are calculated based on the cumulative distribution function of the target distribution at the halfway points $\frac{1}{2}\qty[\psi_j+\psi_{j+1}]$ (vertical dashed lines).
   }\label{fig:grid-sketch}
@@
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained \citep{oreskes_verification:1994,walker_deep:2013}.
+We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained \cite{oreskes_verification:1994,walker_deep:2013}.
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
@@
     Discount rate        & $\rho$            & Determinitic: 2.5\% per year                                                         \\
-        Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
-        Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
+    Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \cite<see>[]{zarekarizi_suboptimal:2020}         \\
+    Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citeA{zarekarizi_suboptimal:2020} \\
     Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
@@
   \end{tabular}
-    %TODO: check if there are other symbols used
 \end{table}
@@
 
-We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
-The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
-The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
+We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
+The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
+The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
 These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
-For a discussion of these model outputs we refer the reader to \citet{ruckert_coastal:2019}.
+For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.
 
-Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \citep{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
+Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
 To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
@@
 
-Following prior work \citep[\eg,][]{garner_slrise:2018,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
+Following prior work \cite<\eg,>[]{garner_slrise:2018,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the freely available NOAA tides and currents data archive \citep{noaa_tidesandcurrents:2022}.
+Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the freely available NOAA tides and currents data archive \cite{noaa_tidesandcurrents:2022}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
   \caption{
-        Annual maximum storm surges (after subtracting \acrlong{msl}) at Sewells Point, VA from the freely available NOAA tides and currents data archive \citep{noaa_tidesandcurrents:2022}.
+    Annual maximum storm surges (after subtracting \acrlong{msl}) at Sewells Point, VA from the freely available NOAA tides and currents data archive \protect\cite{noaa_tidesandcurrents:2022}.
     (A):
@@
     Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
-        % TODO: change caption for Obs (Weibull Plot Pos.) => Observations
-        % TODO: make the different confidence intervals slightly different colors instead of changing opacity
-        % TODO: move (B) to more logical location
   }\label{fig:surge-obs-return}
@@
 
-Our approach to model assessment is based on the concept of principled workflow design for model building and checking \citep[see][for details]{gelman_workflow:2020}.
+Our approach to model assessment is based on the concept of principled workflow design for model building and checking \cite<see>[for details]{gelman_workflow:2020}.
 One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
@@
 Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the parameters $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
+Rather than applying prior information directly over the joint distribution of the parameters $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citeA{coles_evd:1996}.
 Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means of \SIlist{4;6;10;15}{ft} and standard deviations of \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
@@
 
-For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
+For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
 Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
-We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
+We evaluate the model's fit using posterior predictive checks \cite<see>[section 2.4 and references therein]{gelman_workflow:2020}.
 Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
+Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \cite<as in>[]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
 
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citet{zarekarizi_suboptimal:2020}, we use the we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
-For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
-Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
+Following \citeA{zarekarizi_suboptimal:2020}, we use the we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
+Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
 
-The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
+The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \cite<\eg,>[]{vandantzig_dike:1956}.
 However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
@@
 where  $\gamma = 1- \rho$ ($\rho$ being the discount rate), the initial time $t_i=2022$, and the end time $t_f = t_i + T - 1$.
-Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
-For a more theoretical discussion see \citet{arrow_discount:2013}.
+Although \citeA{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
+For a more theoretical discussion see \citeA{arrow_discount:2013}.
 
@@
   \item ``Up-front cost'' is the cost of elevating a house.
-Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
+Following \citeA{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \cite{fischbach_clara:2012}.
 We normalize this cost by house value.
@@
   }\label{tab:slr-priors}
-    \input{../../plots/lsl_priors.tex}
-    % TODO: add another column: Parameter and Quantiles [in feet]
+  \begin{tabular}{llllllll}
+    \toprule
+    Name          & \multicolumn{2}{c}{Parameters} & \multicolumn{5}{c}{Percentiles (in \si{ft})}                                     \\
+    \cmidrule(lr){2-3}
+    \cmidrule(lr){4-8}
+                  & $\alpha$                       & $\theta$                                     & 2.5  & 25.0 & 50.0 & 75.0 & 97.5  \\
+    \midrule
+    Slow SLR      & 1.75                           & 0.50                                         & 0.08 & 0.39 & 0.72 & 1.19 & 2.57  \\
+    Uncertain SLR & 1.75                           & 1.25                                         & 0.21 & 0.98 & 1.79 & 2.97 & 6.41  \\
+    Rapid SLR     & 3.50                           & 1.25                                         & 1.06 & 2.66 & 3.97 & 5.65 & 10.01 \\
+    \bottomrule
+  \end{tabular}
+
 \end{table}
@@
 We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
-We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
-We can also compare to the analyses of \citet{sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
+We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.
+We can also compare to the analyses of \citeA{sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
 Our samples bound all of these estimates.
@@
 This analysis answers ``what-if'' questions like ``given $h_0$ and $\Delta h$, what is the range of total costs a homeowner could face if \gls{slr} over the house lifetime is \SI{1}{ft} or \SI{10}{ft}.''
-For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \citep[\eg, those of][]{sweet_slr:2022} may prove sufficient for decision making.
-This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
-Scenario mapping is also commonly used in \gls{rdm} analyses \citep{lempert_rdm:2019,sriver_sealevel:2018}.
+For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \cite<\eg, those of>[]{sweet_slr:2022} may prove sufficient for decision making.
+This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \cite{Brown:2012kb,Steinschneider:2015kk}.
+Scenario mapping is also commonly used in \gls{rdm} analyses \cite{lempert_rdm:2019,sriver_sealevel:2018}.
 However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
@@
     Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
-        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
+    For emphasis, we highlight three representative models: the Brick Slow model \protect\cite{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \protect\cite{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \protect\cite{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
     (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
@@
     Colored lines highlight three representative models for emphasis.
-        % TODO: add arrows and star
   }\label{fig:tradeoffs-by-rcp}
@@
 The multiple \gls{pdf} problem has also been shown in other contexts.
-For example, \citet{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
-Similarly, \citet{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
-And \citet{lempert_rdm:2019}
+For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
+Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
+And \citeA{lempert_rdm:2019}
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
@@
 \Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
+For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
 Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
@@
     As \cref{fig:tradeoffs-by-rcp}, but with Pareto frontiers for the full distribution of outcomes using the three models of $p_\mathrm{belief}$ (colors).
-        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
-        % TODO: maybe we don't need the gray lines, make the red/yellow/blue lines more bold.
   }\label{fig:tradeoffs-by-prior}
@@
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
+Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased estimates of the posterior \cite{csillery_abc:2010,marjoram_abc:2006}.
 Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences not represented in our model.
 A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
-Although this is appropriate for our didactic illustration, recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} can be applied to improve decision making in real world case studies.
-More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response \citep{morgan_elicitation:2014}.
+Although this is appropriate for our didactic illustration, recent advances in Bayesian elicitation of expert opinion \cite<see>[and references therein]{mikkola_elicitation:2021} can be applied to improve decision making in real world case studies.
+More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response \cite{morgan_elicitation:2014}.
 An alternative approach would be to build a probabilistic model for each of these steps, and to use each as input to the next to develop a fully probabilistic model for \gls{slr}.
-Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022,wong_brick0.2:2017}, this remains a computational and conceptual challenge.
-Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
+Yet while some progress has been made developing probabilistic models for specific elements of this model chain \cite<\eg,>[]{srikrishnan_probabilistic:2022,wong_brick0.2:2017}, this remains a computational and conceptual challenge.
+Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \cite{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
 
@@
 In particular, we hypothesize that incorporating improved representations of risk aversion into decision support may substantially improve their usability.
-One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
+One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \cite{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \cite[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
 Finally, while here we consider the decision to be a one-time decision, one could also frame this as problem a sequential decision.
-The aaalysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \citep{herman_control:2020}.
-Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \citep{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \citep{herman_control:2020}.
+The aaalysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \cite{herman_control:2020}.
+Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \cite{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \cite{herman_control:2020}.
 
 These limitations motivate several directions for future research.
-From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \citep[see fig.~1 of][]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
+From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \cite<see fig.~1 of>[]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
@@
 An obvious area of application is to the design of infrastructure for climate risk management.
-For example, much of the stormwater infrastructure in the United States is inadequate for current climate \citep{lopez-cantu:2018}, let alone future changes.
-Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \citep{sharma_rcp:2021} and \gls{rcp} scenarios.
-Similarly, decisions like  levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
+For example, much of the stormwater infrastructure in the United States is inadequate for current climate \cite{lopez-cantu:2018}, let alone future changes.
+Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \cite{sharma_rcp:2021} and \gls{rcp} scenarios.
+Similarly, decisions like  levee heightening \cite{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
 This framework can also be applied to uncertainties in non-physical variables.
-For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \citep{trindade_deeplyuncertainpathways:2019}.
-Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
+For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \cite{trindade_deeplyuncertainpathways:2019}.
+Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \cite{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
 
-Of course, all models are ultimately wrong \citep{box:1976}.
+Of course, all models are ultimately wrong \cite{box:1976}.
 Thus seeking decisions that perform well across a range of assumptions, and improving the deccision space through robust design and flexibility, can improve outcomes.
@@
 
-\section*{End Matter}
+\section{Open Research}
 
-\subsection*{Acknowledgements}
+All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
+This code is written in the open source Julia programming language and detailed instructions for reproducing our results are provided.
+A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at TBD. % TODO: create Zenodo release!
 
-This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\acknowledgments
+
+This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under \gls{noaa} grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
 JDG thanks Rice University for support.
@@
 
-\subsection*{Code and data availability}
+%% ------------------------------------------------------------------------ %%
+%% References and Citations
+% don't specify bibliographystyle
+% In the References section, cite the data/software described in the Availability Statement (this includes primary and processed data used for your research).
+For details on data/software citation as well as examples, see the Data & Software Citation section of the Data & Software for Authors guidance
+% https://www.agu.org/Publish-with-AGU/Publish/Author-Resources/Data-and-Software-for-Authors#citation
 
-All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
-This code is written in the open source Julia programming language.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsection*{Author Contributions}
+\bibliography{library-bibtex}
 
-JDG and KK designed the research.
-JDG wrote the codes and ran simulations.
-JDG wrote the manuscript in consultation with KK.
-JDG and KK revised the manuscript.
-
-\printbibliography
-
-\appendix
-\newcommand{\hbAppendixPrefix}{S}
-\renewcommand{\thefigure}{\hbAppendixPrefix\arabic{figure}}
-\setcounter{figure}{0}
-\renewcommand{\thetable}{\hbAppendixPrefix\arabic{table}}
-\setcounter{table}{0}
-\renewcommand{\theequation}{\hbAppendixPrefix\arabic{equation}}
-\setcounter{equation}{0}
-
-\newpage
-
-\printglossary[type=\acronymtype,title=List of Abbreviations]
-
-\section{Supplemental methods}
-
-\subsection{Statistical distributions}
-
-To avoid ambiguity, we define here some statistical distributions used in the main text.
-Except where otherwise indicated, we follow the default notation of the Distributions.jl Julia package \citep{besancon_distributions.jl:2021}.
-
-\subsubsection{Generalized Extreme Value distribution}
-We parameterize the \gls{gev} distribution as
-\begin{equation}\label{eq:gev-dist}
-    f(x | \mu, \sigma, \xi)= \begin{cases}
-        \frac{1}{\sigma} \qty[ 1 + \qty( \frac{x-\mu}{\sigma} ) \xi ]^{-1 / \xi - 1} \exp \qty{- \qty[1+\qty(\frac{x-\mu}{\sigma}) \xi ]^{-1 / \xi} }, & \xi \neq 0 \\
-        \frac{1}{\sigma} \exp \qty{-\frac{x-\mu}{\sigma}} \exp \qty{ -\exp \qty[-\frac{x-\mu}{\sigma}]},                                               & \xi=0.
-    \end{cases}
-\end{equation}
-
-\subsubsection{Inverse Gamma distribution}
-Parameterizing the Inverse Gamma distribution as
-\begin{equation}\label{eq:inv-gamma-dist}
-    f(x | \alpha, \theta) = \frac{\theta^\alpha x^{-(\alpha + 1)}}{\Gamma(\alpha)} e^{-\frac{\theta}{x}}, \quad x > 0,
-\end{equation}
-the parameters $\alpha$ and $\theta$ can be computed from the desired mean $\mu$ and standard deviation $\sigma$ as
-\begin{align}\label{eq:inv-gamma-params}
-    \begin{split}
-        \alpha &= 2 + \frac{\mu^2}{\sigma^2} \\
-        \theta &= \mu \qty(\alpha - 1).
-    \end{split}
-\end{align}
-
-\subsubsection{Gamma distribution}
-We parameterize the Gamma distribution as
-\begin{equation}\label{eq:gamma-dist}
-    f(x | \alpha, \theta) = \frac{x^{\alpha-1} e^{-x/\theta}}{\Gamma(\alpha) \theta^\alpha},
-    \quad x > 0.
-\end{equation}
-
-\subsubsection{Plotting position}
-The plotting position used in \cref{fig:surge-obs-return} is the Weibull (``empirical'') plotting position
-\begin{equation}\label{eq:weibull}
-    \nicefrac{r}{N + 1}
-\end{equation}
-where $N$ is the sample size and $r$ is the order of the $N$ observations ($r=1$ is the largest, $r=N$ is the smallest).
-
-\subsection{Algorithm to estimate expected damages}\label{sec:alg-ead}
-
-As discussed in \cref{sec:case-metrics}, we use a Monte Carlo integration to estimate expected damages $D$ as a function of house elevation $h = h_0 + \delta h$ and \gls{msl}, $y(t)$.
-Specifically, for $m=1, \ldots, M$:
-\begin{enumerate}
-    \item draw a sample from the posterior distribution of storm surge (see \cref{eq:surge-model})  $\qty{ \mu_m, \sigma_n, \xi_m}$
-    \item simulate a single storm surge from this stationary \gls{gev} distribution and add the mean sea level to get total flood depth $y^\mathrm{sim}_m$
-    \item calculate the flood damages for this draw by plugging the annual maximum flood depth relative to the house,$h - y'(t)$, into  the deterministic HAZUS depth-damage relationship, storing this as the $m$th damage.
-\end{enumerate}
-We then estimate expected annual damages as the sample mean of the $M$ estimates.
-
-\subsection{Surrogate model for expected annual damages}\label{sec:surrogate-ead}
-Evaluating expected annual damages for each of $J$ simulations of \gls{slr}, each of $N$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
-In our model, we have $T=70$, $N=\num{10000}$, $J=\num{179232}$, hence exhaustive sampling may be prohibitive.
-
-Noticing that this function depends only on the elevation of the house relative to \gls{msl}, we develop a simple emulator for expected annual damages given this difference: $\hat{D}(h - \overline{y})$.
-To do this, we  precompute expected annual damage for all height differences in \SI{0.25}{ft} increments from \SIrange{-30}{30}{ft} and fit a piecewise linear interpolation to this data.
-We use $K=\num{1e6}$ samples to fit this emulator for each of the 241 increments.
-This model is shown in \cref{fig:cost-expected-damage-emulator}.
-Once this interpolation has been precomputed, calculating expected annual damage for a particular year only requires evaluating a piecewise linear function.
-
-\section{Supplemental figures}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{cost-depth-damage}
-    \caption{
-        Depth-damage relationship.
-        Following \citet{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves.
-        Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
-    }\label{fig:cost-depth-damage}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{cost-expected-damage-emulator}
-    \caption{
-        As discussed in \cref{sec:case-metrics}, we model expected annual damages (eq.~\ref{eq:ead}) as a function of the house's elevation relative to \gls{msl}.
-        Damages ($y$ axis) are shown as a percentage of house value.
-    }\label{fig:cost-expected-damage-emulator}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{cost-up-front}
-    \caption{
-        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
-        According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial cost.
-        Values are sensitive to house floor area and structural value; see \cref{tab:uncertainties}.
-    }\label{fig:cost-up-front}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=4in]{surge-gev-priors}
-    \caption{
-        Prior distributions for annual maximum storm surge.
-        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
-        See \cref{sec:case-surge} for details.
-        For the 2, 10, 100, and 500 year events we apply Inverse Gamma distributions, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-    }\label{fig:surge-gev-priors}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-prior-return}
-    \caption{
-        Implicit prior over storm surge represents large uncertainty but is consistent with basic physical reasoning (see \cref{sec:case-surge}).
-        The prior median (blue line) and uncertainty quantiles (shading) of the return period for different return levels of storm surge at Sewells Point, VA using prior predictive sampling \citep{gelman_workflow:2020}.
-        This illustrates the weakly informative prior information (\cref{sec:case-surge}) used in the analysis.
-        These weakly informative priors were selected to be consistent with basic physical principles (\eg, storm surges most years are $>$ \SI{2}{ft} but rarely exceed \SI{20}{ft}).
-        % TODO: fix legend to say Prior Median
-    }\label{fig:surge-prior-return}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-synthetic-data-experiment}
-    \caption{
-        Synthetic data experiment as a positive control test for the \gls{gev} model of storm surge.
-        A synthetic record (dots) was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively.
-        These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:case-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
-        The blue line shows the true quantiles of the (known) \gls{gev} distribution.
-        By random chance the sample maximum has a true return period of $\gg 250$ years, which increases the upper confidence interval of the estimated return probabilities, but the true value is nevertheless within the 50\% posterior confidence interval.
-        This experiment yields similar results for alternative values of the known \gls{gev} distribution, and for different random seeds (not shown).
-    }\label{fig:surge-synthetic-data-experiment}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-posterior-chains}
-    \caption{
-        Markov chain diagnostic plots for posterior draws from the storm surge model.
-        We draw \num{10000} samples by running four chains of \num{3500} iterations each and discarding the first \num{1000}.
-        The mixing of the chains is consistent with, though does not guarantee, convergence.
-    }\label{fig:surge-posterior-chains}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-prior-chains}
-    \caption{
-        As \cref{fig:surge-posterior-chains}, but for draws from the prior distribution.
-    }\label{fig:surge-prior-chains}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-test-statistics}
-    \caption{
-        Posterior predictive checks for the stationary \gls{gev} storm surge model (\cref{sec:case-surge}).
-        Each panel shows a different test statistic: partial autocorrelation at lags 1 and 2; sample maximum; sample minimum; sample median; and Mann-Kendall trend test statistic.
-        The histograms show the distribution of each test statistic from the posterior predictive distribution.
-        Orange lines show the test statistic's value in the observed data.
-        Observed values near the mode of the posterior predictive distribution are consistent with, but do not guarantee, a good fit.
-        For further discussion of posterior predictive checks, see Chapter 6 of \citet{gelman_bda3:2014}.
-    }\label{fig:surge-test-statistics}
-\end{figure}
-
-\begin{figure}
-    \includegraphics[width=\textwidth]{scenario-map-height-slr}
-    \caption{
-        Expected total lifetime cost (damages plus up-front cost) as a function of \gls{slr} over the house lifetime and height increase $\Delta h$.
-        Initial house elevation is fixed to \SI{1}{ft} below the \gls{bfe}.
-        Expectations were computed for discrete values of $\Delta h$ ($x$ axis) by discretizing \glspl{sow} ($y$ axis), then taking the sample mean over each grid cell.
-    }\label{fig:scenario-map-height-slr}
-\end{figure}
-
-\section{Supplemental tables}
-
-\begin{table}[h]
-    \centering
-    \caption{
-        Diagnostic statistics for the Hamiltonian Monte Carlo sampling for the storm surge posterior draws.
-        Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.
-        The naive standard error (SE) returns the standard error of the mean.
-        The Monte Carlo standard error (MCSE) is calculated using the initial monotone sequence estimator \citep[][pp.~473-483]{geyer_practical:1992}.
-        The effective sample size (ESS) is a crude estimate of the number of independent samples and is calculated following \citet[][pp.~473-483]{geyer_practical:1992}.
-        The $\hat{R}$ is an indicator of the convergence of the Markov chains to the target distribution; a value $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence \citep{mcelreath_rethinking2:2020}.
-        For additional details see the \texttt{MCMCDiagnosticTools} package documentation.
-    }\label{tab:surge-posterior-mcmc-diagnostics}
-    \input{../../plots/surge-posterior-mcmc-diagnostics.tex}
-\end{table}
-
-\begin{table}[h]
-    \centering
-    \caption{As \cref{tab:surge-posterior-mcmc-diagnostics} but for draws from the prior distribution.}\label{tab:surge-prior-mcmc-diagnostics}
-    \input{../../plots/surge-prior-mcmc-diagnostics.tex}
-\end{table}
-
-% up to 1250 words
 \end{document}

### 2022-06-22 papers/earths-future/supplemental.tex
Format to prepare for Earth's Future submission

--- 2022-06-17 papers/paper/preprint.tex
@@
-\documentclass[11pt]{article}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% AGUtmpl.tex: this template file is for articles formatted with LaTeX2e,
+% Modified December 2018
+%
+% This template includes commands and instructions
+% given in the order necessary to produce a final output that will
+% satisfy AGU requirements.
+%
+% FOR FIGURES, DO NOT USE \psfrag
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% IMPORTANT NOTE:
+%
+% SUPPORTING INFORMATION DOCUMENTATION IS NOT COPYEDITED BEFORE PUBLICATION.
+%
+%
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+% Step 1: Set the \documentclass
+%
+%
+% PLEASE USE THE DRAFT OPTION TO SUBMIT YOUR PAPERS.
+% The draft option produces double spaced output.
+%
+% Choose the journal abbreviation for the journal you are
+% submitting to:
 
-% Figures
-\usepackage[demo]{graphicx}
-\usepackage[list=true]{subcaption}
-\graphicspath{{../../plots/}{../../tikz/}{../../img/}}
-\usepackage[section]{placeins} % require floats to appear in the section they are defined
+% jgrga JOURNAL OF GEOPHYSICAL RESEARCH (use for all of them)
+% gbc   GLOBAL BIOCHEMICAL CYCLES
+% grl   GEOPHYSICAL RESEARCH LETTERS
+% pal   PALEOCEANOGRAPHY
+% ras   RADIO SCIENCE
+% rog   REVIEWS OF GEOPHYSICS
+% tec   TECTONICS
+% wrr   WATER RESOURCES RESEARCH
+% gc    GEOCHEMISTRY, GEOPHYSICS, GEOSYSTEMS
+% sw    SPACE WEATHER
+% ms    JAMES
+% ef    EARTH'S FUTURE
+%
+%
+%
+% (If you are submitting to a journal other than jgrga,
+% substitute the initials of the journal for "jgrga" below.)
+
+\documentclass[ef,draft]{agutexSI2019}
 
@@
 
-% better lists
-\usepackage[inline]{enumitem}
-\setlist{nosep}
-
-% authors
-\usepackage{authblk}
-\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
-\author[1]{James Doss-Gollin}
-\author[2]{Klaus Keller}
-\affil[1]{Department of Civil and Environmental Engineering, Rice University}
-\affil[2]{Thayer School of Engineering, Dartmouth College}
-\renewcommand*{\Affilfont}{\normalsize\normalfont}
+\usepackage{xspace}
+\makeatletter
+\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
+\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
+\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
+\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
+\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+\newcommand{\usd}[1]{\SI{#1}[\$]{}}
 
@@
 
-\usepackage{xspace}
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
-\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
-\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%  SUPPORTING INFORMATION TEMPLATE
+%
+%% ------------------------------------------------------------------------ %%
+%
+%
+%Please use this template when formatting and submitting your Supporting Information.
 
-\usepackage{xspace}
-\makeatletter
-\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
-\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
-\newcommand{\usd}[1]{\SI{#1}[\$]{}}
-\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
-\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
-\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+%This template serves as both a “table of contents” for the supporting information for your article and as a summary of files.
+%
+%
+%OVERVIEW
+%
+%Please note that all supporting information will be peer reviewed with your manuscript. It will not be copyedited if the paper is accepted.
+%In general, the purpose of the supporting information is to enable authors to provide and archive auxiliary information such as data tables, method information, figures, video, or computer software, in digital formats so that other scientists can use it.
+%The key criteria are that the data:
+% 1. supplement the main scientific conclusions of the paper but are not essential to the conclusions (with the exception of
+%    including %data so the experiment can be reproducible);
+% 2. are likely to be usable or used by other scientists working in the field;
+% 3. are described with sufficient precision that other scientists can understand them, and
+% 4. are not exe files.
+%
+%USING THIS TEMPLATE
+%
+%***All references should be included in the reference list of the main paper so that they can be indexed, linked, and counted as citations.  The reference section does not count toward length limits.
+%
+%All Supporting text and figures should be included in this document.
+Insert supporting information content into each appropriate section of the template.
+To add additional captions, simply copy and paste each sample as needed.
 
-% use biblatex
-\usepackage{csquotes}
-\usepackage[
-  backend=biber,
-  doi=true,
-  url=false,
-  isbn=false,
-  style=authoryear-comp,
-  natbib=true,
-  backref=false,
-  maxbibnames=10,
-  maxcitenames=2,
-  uniquename=false,
-  uniquelist=false,
-  sorting=nyt,
-  giveninits=true,
-]{biblatex}
-\renewbibmacro{in:}{}
-\AtEveryBibitem{\clearfield{month}\clearfield{day}\clearfield{pages}\clearlist{language}}
-\addbibresource{library.bib}
+%Tables may be included, but can also be uploaded separately, especially if they are larger than 1 page, or if necessary for retaining table formatting.
+Data sets, large tables, movie files, and audio files should be uploaded separately.
+Include their captions in this document and list the file name with the caption.
+You will be prompted to upload these files on the Upload Files tab during the submission process, using file type “Supporting Information (SI)”
 
-% load this last
-\usepackage[hidelinks]{hyperref}
+%IMPORTANT NOTE ON FIGURES AND TABLES
+% Placeholders for figures and tables appear after the \end{article} command, after references.
+% DO NOT USE \psfrag or \subfigure commands.
+%
+ \usepackage{graphicx}
+ \graphicspath{{../../plots/}{../../tikz/}{../../img/}}
+%
+%  Uncomment the following command to allow illustrations to print
+%   when using Draft:
+\setkeys{Gin}{draft=false}
+%
+% You may need to use one of these options for graphicx depending on the driver program you are using. 
+%
+% [xdvi], [dvipdf], [dvipsone], [dviwindo], [emtex], [dviwin],
+% [pctexps],  [pctexwin],  [pctexhp],  [pctex32], [truetex], [tcidvi],
+% [oztex], [textures]
+%
+%
+%% ------------------------------------------------------------------------ %%
+%
+%  ENTER PREAMBLE
+%
+%% ------------------------------------------------------------------------ %%
+
+% Author names in capital letters:
+\authorrunninghead{DOSS-GOLLIN AND KELLER}
+
+% Shorter version of title entered in capital letters:
+\titlerunninghead{Synthesizing Deep Uncertainties}
+
+%Corresponding author mailing address and e-mail address:
+\authoraddr{Corresponding author: J. Doss-Gollin, Department of Civil and Environmental Engineering, Rice University, Houston, TX, USA (jdossgollin@rice.edu)}
+
+% cite things in supplemental info file
+\usepackage{xr}
+\externaldocument{submission}
+
+% this goes last
 \usepackage{cleveref}
+\begin{document}
 
-% up to 1250 words
-\begin{document}
-\maketitle
-\thispagestyle{empty}
+%% ------------------------------------------------------------------------ %%
+%
+%  TITLE
+%
+%% ------------------------------------------------------------------------ %%
 
-\begin{abstract}
-    Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand.
-    The
-\end{abstract}
+%\includegraphics{agu_pubart-white_reduced.eps}
 
-Key points
-\begin{enumerate}
-    \item
-    \item
-    \item
-\end{enumerate}
+\title{Supporting Information for ``A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management''}
+%
+% e.g., \title{Supporting Information for "Terrestrial ring current:
+% Origin, formation, and decay $\alpha\beta\Gamma\Delta$"}
+%
+%DOI: 10.1002/%insert paper number here%
 
-\clearpage
-\section{Introduction}\label{sec:introduction}
+%% ------------------------------------------------------------------------ %%
+%
+%  AUTHORS AND AFFILIATIONS
+%
+%% ------------------------------------------------------------------------ %%
 
-Many critical infrastructure services around the world are aging and inadequate for a changing climate \citep[\eg,][]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
-Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
-This motivates the question: which possible futures should infrastructure systems and components be designed for?
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \citep{keller_management:2021} and are often regulated by statue or industry guidelines.
-For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
-At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different future conditions.
+% List authors by first name or initial followed by last name and
+% separated by commas. Use \affil{} to number affiliations, and
+% \thanks{} for author notes.
+% Additional author notes should be indicated with \thanks{} (for
+% example, for current addresses).
 
-\subsection{Current practice}
+% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
+% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
 
-Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \citep{asce_7-10:2013,bruneau_multihazard:2017}.
-These are often, though not always, informed by probabilistic analysis of relevant data.
-For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \citep{kousky_voucher:2014}.
-Additionally, local building codes \citep[based on guidance such as][]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
-Design standards are also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \citep{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \citep[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+\authors{James Doss-Gollin\affil{1}, Klaus Keller\affil{2}}
+\affiliation{1}{Department of Civil and Environmental Engineering, Rice University}
+\affiliation{2}{Thayer School of Engineering, Dartmouth College}
 
-There are many advantages to standards-based design.
-In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
-However, there are also limitations.
-In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
-This has motivated economically informed approaches, like risk-based design and cost benefit analysis \citep{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \citep{merz_fluvial:2010}.
-These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.
+%% ------------------------------------------------------------------------ %%
+%
+%  BEGIN ARTICLE
+%
+%% ------------------------------------------------------------------------ %%
 
-Currently used methods for estimating these \glspl{pdf} emphasize data-driven, nominally objective methods that can be applied consistently across locations.
-For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
-Among several statistical assumptions these analyses make that have been recently called into question is stationarity \citep[the assumption that past and future hazard come from the same \gls{pdf}; see][for a review]{milly_stationarity:2008,merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \citep{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \citep[see][for a comprehensive summary]{ipcc_impacts:2022}.
-While some methods have been proposed for incorporating trends into these analyses \citep[see][for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \citep{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
-Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
+% The body of the article must start with a \begin{article} command
+%
+% \end{article} must follow the references section, before the figures
+%  and tables.
 
-\subsection{Emerging paradigms}
+\begin{article}
 
-Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \citep[\eg, collapse of the West Antarctic ice sheet; see][]{deconto_antarctica:2016}.
-In other words, they are deeply uncertain \citep{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
-These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
+      This document contains supplementary methods, figures, and tables for the paper ``A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management'' by James Doss-Gollin and Klaus Keller.
 
-Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
-Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \citep{bankes:1993}.
-For example, robust decision making \citep{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
-When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \citep{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-In addition, many studies formally quantify robustness to deep uncertainties \citep{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
-
-Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
-Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
-When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
-For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \citep[p.54]{gelman_bda3:2014}.
-While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \citep{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
-This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
-
-\subsection{Research gaps and objectives}
-
-We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
-\citet{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
-In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
-This challenge is described in the Bayesian literature on model selection, that is the problem of using competing statistical models to draw inferences, particularly in the ``$\mathcal{M}$-open'' case where there is no true model to identify and all models are imperfect approximations of reality.
-Just as the literature on decision making under deep uncertainty contrasts with classical methods for decision analysis that identify an objectively ``best'' decision (or decision set) under a ``true'' description of uncertaintty, the literature on Bayesian model selection in the $\mathcal{M}$-open case contrasts with classical methods that seek to asymptotically identify a ``true'' model from a candidate space.
-In particular, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013} or \cite{ramsey_probability:2016} for a philosophical discussion, \cite{jaynes_probability:2003} for a detailed reference, or \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, emphasis is placed on iteratively building models, simulating the consequences of those models, and subsequently critiquing them \citep{gelman_workflow:2020}.
-
-In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
-We consider the didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
-Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \citep{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \citep{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \citep{FEMA_p-55:2011} require elevation in some cases.
-Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \citep{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \citep{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 71 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
-
-We proceed as follows.
-In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
-In \cref{sec:case-study} we describe the case study in detail.
-Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
-Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.
-
-\section{Decision analytic framework}\label{sec:analysis}
-
-In this section we outline our framework for decision analysis under deep uncertainty, maintaining a high level of generality.
-In the next we discuss application to the house elevation case study.
-
-Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
-For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
-While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{bayes-rdm.pdf}
-    \caption{
-        Outline of the proposed decision-analytic framework.
-        In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance (c) of candidate decisions (a) under a large ensemble of possible futures (b).
-        In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes (e) that are conditional upon specific models describing the likelihood of different futures (d).
-        In our case study, these are \gls{rcp} scenarios and physical models of sea level rise.
-        Finally in \cref{sec:analysis-synthesize} we propose a subjective Bayesian framework for synthesizing across deep uncertainties by re-weighting sampled futures (f).
-    }\label{fig:flowchart}
-\end{figure}
-
-\subsection{Explore}\label{sec:analysis-explore}
-
-A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \citep{bankes:1993}by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \citep{reed_msdbook:2022}.
-Exploratory modeling is often paired with analyses that identify relevant scenarios  \citep{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \citep{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system modle, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \citep{quinn_exploratory:2020}.
-
-\subsection{Condition}\label{sec:analysis-condition}
-
-Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes.\klaus{Vivek suggests citing Jaynes here, anything in particular?}
-
-One way to interpret an ensemble of \glspl{sow} is as \gls{iid} draws from some true data generating process.
-We call this scenario-conditional probabilistic analysis, which commonly arises when a probabilistically calibrated model for relevant outputs (\eg, a physical model for \gls{msl}; see \cref{sec:case-slr}) is forced by a deterministic scenario (\eg, a \gls{rcp} scenario).
-In this case, \emph{the set of available \glspl{sow} is assumed to come from the true (often called population) distribution, and thus the sample statistics are assumed to reflect the true values.}
-We add this concept to our framework with boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario (\ie, the assumed input) $M_k$.
-If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
-This approach allows for quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \citep{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
-% Vivek raises an interesting point -- what if we incorporate the choice of model into the data generating process and use a posterior probability of each model being true -- but I think that's basically what we are calling step #3
-
-We can also use this approach to understand methods for \gls{dmdu} that sample a set of parameters from fixed ranges.
-For example, \citet{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
-Similarly, \citet{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \citep[\ie, robustness metrics; see][for details]{mcphail_robustness:2019,herman:2015}.
-Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
-
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to poor decisions.\james{Build this out -- it's not so much the potential for poor decisions so much as that it can be hard to undwind the decision and understand what's driving it.
-We cant know if critical assumptions turned out to be poor if we don't know what they are or how changing them would result in a different analysis}
-
-\subsection{Synthesize}\label{sec:analysis-synthesize}
-
-If the ensemble of \glspl{sow} is not drawn \gls{iid} from the true distribution (\ie, if we deliberately over-sample low-probability but high-impact \glspl{sow}) or if we wish to synthesize \emph{across} multiple scenarios $M_1, M_2, \ldots, M_K$, then additional measures must be taken.
-One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
-In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
-
-Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
-A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through kernel density estimation or other methods.
-Instead, paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \citep{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
-
-First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
-This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
-We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ so that the weighted distribution of $\psi_j$ closely approximates $p_\mathrm{belief}(\Psi)$.
-
-We present here the case where the $\psi_j$ are one-dimensional.
-We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
-\begin{equation}\label{eq:weights}
-    w_j = \begin{cases}
-        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
-        F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
-        1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
-    \end{cases}
-\end{equation}
-This step is illustrated in \cref{fig:grid-sketch}.
-For higher dimensional projections, this equation can be extended by partitioning the parameter space, then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
-From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
-Diagnostic checks, such as examining the histogram of weights (not shown), may be valuable protections against degeneracy.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{grid-sketch}
-    \caption{
-        Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
-        This method is illustrated for a hypothetical target distribution (orange line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (orange dots).
-        As shown in \cref{eq:weights}, the weights $w_j$ (vertical lines) are calculated based on the cumulative distribution function of the target distribution  (black line) at the halfway points $\frac{1}{2}\qty[\psi_i+\psi_{i+1}]$ (vertical dashed lines).
-    }\label{fig:grid-sketch}
-\end{figure}
-
-The aim of this re-weighting framework is to integrate an ensemble of \glspl{sow} used for exploratory modeling into formal decision analysis, even when the \glspl{sow} deliberately over- or under-sample some regions of the parameter space.
-As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained \citep{oreskes_verification:1994,walker_deep:2013}.
-However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
-
-\section{Case study}\label{sec:case-study}
-
-To illustrate the proposed decision analytic framework we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
-We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
-For interpretability, we focus on deep uncertainty in \gls{msl} and treat other model parameters as more shallow uncertainties as shown in \cref{tab:uncertainties}.
-We use the notation developed in the previous section to describe the case study.
-Specifically,
-\begin{enumerate}
-    \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
-    \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=71$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
-    \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
-\end{enumerate}
-In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).
-
-\begin{table}
-    \centering
-    \caption{
-        Summary of parameters, their notation, and how their uncertainty is represented.
-        Symbols describing the decision-analytic framework are described in \cref{fig:flowchart}.
-    }\label{tab:uncertainties}
-    \footnotesize
-    \begin{tabular}{p{1.25in} p{0.75in} p{3in}}
-        \toprule
-        Name                 & Symbol            & Uncertainty                                                                          \\
-        \midrule
-        \Gls{msl}            & $\overline{y}(t)$ & Deeply uncertain: four physical models $\times$ four \acrshort{rcp} scenarios        \\
-        Storm surge          & $y'(t)$           & Probabilistic: Bayesian inference on a stationary \acrshort{gev} model               \\
-        Annual maximum flood & $y(t)$            & Deterministic: $y(t)=\overline{y}(t)+y'(t)$                                          \\
-        Discount rate        & $1-\gamma$        & Fixed at 2.5\% per year                                                              \\
-        Depth-damage         & $D(h-y)$          & Deterministic: based on HAZUS model \citep[see][]{zarekarizi_suboptimal:2020}        \\
-        Elevation cost       & $C(\Delta h)$     & Deterministic: a piecewise linear model following \citet{zarekarizi_suboptimal:2020} \\
-        Initial height       & $h_0$             & Deterministic: \SI{1}{ft} below the \gls{bfe}, unless otherwise noted                \\
-        House floor area     & --                & Deterministic: \SI{1500}{ft^2}                                                       \\
-        Structural value     & --                & Deterministic: \usd{200000}                                                          \\
-        House lifespan       & $T$               & Deterministic: 71 years                                                              \\
-        \bottomrule
-    \end{tabular}
-    %TODO: check if there are other symbols used
-\end{table}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{xlrm.pdf}
-    \caption{
-        Conceptual diagram of the considered example.
-        A \acrfull{sow} consists of a description of the uncertain factors (red).
-        We model a problem with a single lever (yellow), which is how high to elevate a house ($\Delta h$).
-        For each \acrshort{sow} (red) and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (gray).
-    }\label{fig:xlrm}
-\end{figure}
-
-\subsection{\Glsentrylong{slr}}\label{sec:case-slr}
-
-We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citet{ruckert_coastal:2019}.
-The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \citep{wong_brick0.2:2017}, (iii) the \citet{kopp_probabilistic:2014} model (``K14''), and (iv) the \citet{deconto_antarctica:2016} model (``DP16'').
-The \citet{kopp_probabilistic:2014} and \citet{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
-These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
-For a discussion of these model outputs we refer the reader to \citet{ruckert_coastal:2019}.
-
-Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \citep{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
-To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{lsl-evolution}
-    \caption{
-        Projections of future mean sea level depend strongly on the choices of physical model and forcing.
-        (A): 90\% confidence intervals for mean sea level at Sewells Point, VA as a function of time for a representative subset of three probabilistic models (out of sixteen).
-        (B): probability distribution of \gls{msl} at Sewells Point, VA in the year 2100 for each probabilistic model considered.
-    }\label{fig:lsl-evolution}
-\end{figure}
-
-The choices of physical model and \gls{rcp} scenario jointly determine future \gls{msl} $p(\overline{y}|t)$.
-\Cref{fig:lsl-evolution}(a) shows the time-varying 90\% credible intervals of \gls{msl} for three representative models.
-The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
-\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.
-
-\subsection{Storm surges}\label{sec:case-surge}
-
-Following prior work \citep[\eg,][]{garner_slrise:2018,lempert_slr:2012,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
-
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the NOAA tides and currents dataset,  freely available to the public at \url{https://tidesandcurrents.noaa.gov/waterlevels.html}.\klaus{I can't find a better reference for this dataset!}
-Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
-For each calendar year we first remove the annual mean, then calculate the maximum water level.
-We refer this time series of annual maximum storm surges as $y'(t)$.
-We display this time series of annual maxima storm surges in \cref{fig:surge-obs-return}(a).
-The largest recorded surge was the Chesapeake-Potomac hurricane of 1933, which caused a surge of over \SI{7}{ft} at this gauge, but other hurricanes and Nor'easters have caused surges above \SI{6}{ft}.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{surge-obs-return}
-    \caption{
-        Annual maximum storm surges (after subtracting mean sea level) at Sewells Point, VA.
-        (A):
-        time series of historic storms.
-        Red (yellow) arrows denote notable tropical cyclones (Nor'easters).
-        (B):
-        return periods.
-        Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula (eq.~\ref{eq:weibull}).
-        Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
-    }\label{fig:surge-obs-return}
-\end{figure}
-
-We model future storm surge using a stationary \gls{gev} model:
-\begin{equation}\label{eq:surge-model}
-    y'(t) \sim \text{GEV}\qty(\mu, \sigma, \xi),
-\end{equation}
-where $y'(t)$ is the storm surge (above sea level) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
-This model assumes stationarity, neglecting any potential time dependence.
-
-Our approach to model validation is based on the concept of principled workflow design for model building and checking \citep[see][for details]{gelman_workflow:2020}.
-One model choice, analogous to the choice of statistical distribution or the assumption of stationarity, is the choice of how to represent prior information.
-We include two forms of prior information.
-First, we constrain the shape parameter to be positive, $\xi > 0$, to reflect knowledge about the support of $y'$, which for a variable distributed according to \cref{eq:gev-dist} is:
-\begin{equation*}
-    \mathrm{supp}~ y' =
-    \begin{cases}
-        \xi < 0: & y' \in \qty(-\infty,~ \mu - \nicefrac{\sigma}{\xi}) \\
-        \xi > 0: & y' \in \qty(\mu-\nicefrac{\sigma}{\xi},~ \infty).
-    \end{cases}
-\end{equation*}
-Since storm surges cannot be negative, only the latter is physically defensible, justifying our choice to constrain the shape parameter to be positive.
-Second, we add weakly informative priors.
-Rather than applying prior information directly over the joint distribution of the parameters $\qty{\mu,\sigma,\xi}$, we instead apply a prior over extreme quantiles of the distribution, as in \citet{coles_evd:1996}.
-Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means \SIlist{4;6;10;15}{ft} and standard deviations \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-The parameters of the Inverse Gamma distribution can be calculated from these moments (see eq.~\ref{eq:inv-gamma-params}).
-These means and standard deviations were chosen to represent plausible physical ranges and are plotted in \cref{fig:surge-gev-priors}.
-
-For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \citep{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \citep{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
-Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
-We evaluate the model's fit using posterior predictive checks \citep[see][section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we found that draws from the posterior predictive distribution matched the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
-Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \citep[as in][]{wong_structural:2020,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
-
-\Cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges.
-The return period of the data (dots) is shown using the Weibull plotting position (see \cref{eq:weibull}).
-The good fit of the Bayesian model to the data suggest the modeling choices made are reasonable and sufficient for this didactic example.
-A positive control test (\cref{fig:surge-synthetic-data-experiment}) also validates the model's ability to recover known parameter values.
-
-\subsection{Damages and metrics}\label{sec:case-metrics}
-
-The system model (``relationships'' in \cref{fig:xlrm}) is comprised of two key pieces.
-The first is a fragility model that estimates the expected flood damages for a particular year (``expected annual damages''), given the elevation of the house and the mean sea level for that year.
-The second model converts a time series of annual expected damages into lifetime expected damages.
-
-We define expected annual damages in year $t$ as the expectation of the damage function with respect to storm surge.
-This expectation depends on the house's height ($h = h_0 + \Delta h$) where $h_0$ is the initial height relative to the gauge and $\Delta h$ is the amount by which the house is elevated.
-The expected annual damage is thus
-\begin{equation}\label{eq:ead}
-    \textrm{EAD}(t) = \mathbb{E}[D \qty(h-\overline{y}(t))] = \int_{y'} p(y') D \qty(h - \qty(\overline{y}(t) + y')) \dd{y'},
-\end{equation}
-where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citet{zarekarizi_suboptimal:2020} we use the we use the Hazard U.S.
-(HAZUS) depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
-For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
-Although \citet{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
-
-The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \citep[\eg][]{vandantzig_dike:1956}.
-However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
-Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).
-
-The second component of the system model converts a time series of $\mathrm{EAD}$ into lifetime expected damages, which we define as the up-front  discounted sum of expected annual damages:
-\begin{equation}\label{eq:led}
-    \textrm{LED} = \sum_{t=t_i}^{t_f} \gamma^{\qty(t-t_i)-1} \textrm{EAD}(t),
-\end{equation}
-where the discount rate is $1 - \gamma$, the initial time $t_i=2022$, and the end time $t_f = t_i + T$.
-Although \citet{zarekarizi_suboptimal:2020} show that uncertainty in the discount rate is important for decision support, we use a fixed discount rate (see \cref{tab:uncertainties}) for the purposes of this didactic study.
-For a more theoretical discussion see \citet{arrow_discount:2013}.
-
-To assess the performance of a given decision for a specific \gls{sow} (``Metrics'' in \cref{fig:xlrm}), we calculate the following metrics for each decision-\gls{sow} combination:
-\begin{enumerate}
-    \item ``Up-front cost'' is the cost of elevating a house.
-Following \citet{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \citep{fischbach_clara:2012}.
-We normalize this cost by house value.
-This cost curve is shown in \cref{fig:cost-up-front}.
-    \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
-    \item ``Expected lifetime costs'' is the sum of lifetime expected damages and up-front costs.
-\end{enumerate}
-
-\subsection{Prior over \glsentrylong{slr}}\label{sec:case-priors}
-
-We construct three probabilistic models for $p_\mathrm{belief}(\psi)$, which represents the amount of \gls{slr} from 2022 to 2100.
-
-We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
-\Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
-Their \glspl{pdf} are also plotted in \cref{fig:lsl-priors-weights}(A).
-
-\begin{table}[h]
-    \centering
-    \caption{
-        Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\mathrm{belief}(\psi)$.
-        The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles (values in \si{ft}).
-    }\label{tab:slr-priors}
-    \input{../../plots/lsl_priors.tex}
-    % TODO: add another column: Parameter and Quantiles [in feet]
-\end{table}
-
-We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
-We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \citep[table.~2.4]{sweet_slr:2022}.
-We can also compare to the analyses of \citet{lempert_slr:2012,sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
-Our samples bound all of these estimates.
-
-\section{Results and discussion}\label{sec:results}
-
-We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed synthesis method (\cref{sec:results-synthesis}).
-
-\subsection{Exploratory modeling}\label{sec:results-exploratory}
-
-We begin by using our model in an ``exploratory'' mode with an aim of learning about interactions between system dynamics and decisions.
-
-\begin{figure}
-    \includegraphics[width=\textwidth]{scenario-map-slr-cost}
-    \caption{
-        Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
-        Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
-        The lowest-cost outcomes occur when both exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
-        The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
-        In all cases, elevating the house reduces the variance in total lifetime cost.
-        Values are sensitive to model constants; see \cref{tab:uncertainties}.
-    }\label{fig:scenario-map-slr-cost}
-\end{figure}
-
-One application of explotatory analysis is to reveal the range and variation in outcomes, conditional on taking a particular decision.
-\Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
-The outcomes with lowest total lifetime costs arise when the house is not elevated ($\Delta h = 0$) and \gls{slr} is minimal (bottom left corners).
-The outcomes with highest total lifetime costs arise when the house is elevated only slightly and \gls{slr} is rapid.
-As $\Delta h$ increases, the best-case scenario becomes more expensive because up-front costs increase, but worst-case scenarios become less expensive because,.
-even if \gls{slr} is substantial, damages will be negligible.
-
-This analysis answers ``what-if'' questions like ``given $h_0$ and $\Delta h$, what is the range of total costs a homeowner could face if \gls{slr} over the house lifetime is \SI{1}{ft} or \SI{10}{ft}.''
-For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \citep[\eg, those of][]{sweet_slr:2022} may prove sufficient for decision making.
-This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \citep{Brown:2012kb,Steinschneider:2015kk}.
-However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
-
-\subsection{Scenario-conditional optimization}\label{sec:results-conditional}
-
-We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
-Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draw from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
-As discussed in \cref{sec:analysis-condition}, this allows a formal estimation of decision metrics.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-rcp}
-    \caption{
-        Each probabilistic model or scenario leads to a different estimate of the Pareto frontier.
-        For emphasis, we highlight three representative models: the Brick Slow model \citep{wong_brick0.2:2017} under \gls{rcp} 2.6, the K14 \citep{kopp_probabilistic:2014} model under \gls{rcp} 6.0 and the DP16 model \citep{deconto_antarctica:2016,kopp_evolving:2017} under \gls{rcp} 8.5.
-        (A): trade-off between up-front cost (which is a monotonic function of height increase) and expected lifetime costs.
-        (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
-        Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
-        Colored lines highlight three representative models for emphasis.
-    }\label{fig:tradeoffs-by-rcp}
-\end{figure}
-
-This probabilistic interpretation allows us to compute, for example, expected values.
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen models considered (we highlight three representative models).
-Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
-This figure shows that for small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model).
-For example, under the most pessimistic model (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
-Converseley, under the most optimistic model (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost of 73\% of house value yet reduces lifetime expected damages by less than 50\% of house value.
-
-This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
-This approach is attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
-However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from.
-The multiple \gls{pdf} problem has also been shown in other contexts.
-For example, \citet{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
-Similarly, \citet{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
-Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
-
-\subsection{Synthesizing deep uncertainties for decision analysis}\label{sec:results-synthesis}
-
-The proposed approach can help overcome the limitations of scenario-conditional analysis.
-We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
-We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
-These three distributions are shown in \cref{fig:lsl-priors-weights}(A).
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{lsl-priors-weights}
-    \caption{
-        Subjective priors for local sea level.
-        We develop two distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
-        The \glspl{pdf} of these subjective priors are shown in panel (A).
-        In panels (B-C) we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
-        Specifically, (B-C) show the average weight given to each model by each of the subjective priors.
-    }\label{fig:lsl-priors-weights}
-\end{figure}
-
-One application of this method is to diagnose which assumptions different $p_\mathrm{belief}$ are consistent with.
-\Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-For example, the rapid \gls{slr} scenario places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which is unlikely given current policy \citep{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
-Conversely, the slow \gls{slr} scenario places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
-The uncertain \gls{slr} scenario places approximately equal weight across models.
-
-This method can also be used to calculate expectations, allowing us to revisit the trade-off diagrams of \cref{fig:tradeoffs-by-rcp}.
-\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
-Notably, they give different guidance.
-Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} costs 73\% of house value and reduces damages by over 100\% of house value, yielding a benefit to cost ratio of approximately 1.25.
-Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} reduces damages only by 50\% of house value, yielding a benefit to cost ratio of approximately 0.7.
-Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house, and thus the benefit to cost ratio is nearly 1.
-Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{tradeoffs-by-prior}
-    \caption{
-        As \cref{fig:tradeoffs-by-rcp}, but Pareto frontiers are shown for the full distribution of outcomes, for three models of $p_\mathrm{belief}$ (colors).
-        Thin gray lines show the trade-off curves for each \gls{rcp} scenario and model separately as in \cref{fig:tradeoffs-by-rcp}.
-    }\label{fig:tradeoffs-by-prior}
-\end{figure}
-
-\section{Conclusions}\label{sec:conclusions}
-
-The motivation for this study was to develop a framework for increasing the transparency of assumptions about the likelihood of different \glspl{sow}.
-Additionally, we sought to develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
-Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-In this paper we used a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
-
-There are several limitations to our study that merit further discussion here.
-The first category has to do with limitations of the underlying methodology proposed for reweighting state of the world.
-For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional projection of the full time series, it is not a sufficient statistic.
-It is known that in approximate Bayesian computation, calibrating models on low dimensional statistics that are not sufficient statistics can lead to biased estimates of the posterior \citep{csillery_abc:2010,marjoram_abc:2006}.
-Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences.
-A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
-Although this is appropriate for our didactic illustration, we suggest that recent advances in Bayesian elicitation of expert opinion \citep[see][and references therein]{mikkola_elicitation:2021} should be applied to formal guidance.
-More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response.
-An alternative approach would be to fully couple these models together to generate samples from the full distribution of \gls{slr}.
-Yet while some progress has been made developing probabilistic models for specific elements of this model chain \citep[\eg,][]{srikrishnan_probabilistic:2022}, coupling these models remains a computational and conceptual challenge.
-Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \citep{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
-
-The second category of limitations has to do with the case study and our interpretation of the house elevation decision problem.
-This problem intersects with decisions about where to live and how to manage household finances, both of which are highly complex.
-We could extend our analysis, for example, by considering additional decision objectives.
-In particular, we hypothesize that incorporating improved representations of risk aversion into decision support may substantially improve their usability.
-We could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \citep{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \citep[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
-Finally, while here we consider the decision to be a one-time decision, it can also be framed as a sequential decision.
-Analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions.
-Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \citep{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \citep{herman_control:2020}.
-
-These limitations motivate several directions for future research.
-From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \citep[see fig.~1 of][]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
-From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
-Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that scale to different house characteristics may add value.
-
-The scenario weighting framework described in \cref{sec:analysis-synthesize} and illustrated in this paper can be applied to inform critical challenges in climate risk management.
-An obvious area of application is to the design of infrastructure for climate risk management.
-For example, much of the stormwater infrastructure in the United States is inadequate for current climate \citep{lopez-cantu:2018}, let alone future changes.
-Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \citep{sharma_rcp:2021} and \gls{rcp} scenarios.
-Similarly, decisions like  levee heightening \citep{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
-This framework can also be applied to uncertainties in non-physical variables.
-For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \citep{trindade_deeplyuncertainpathways:2019}.
-Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \citep{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
-
-Ultimately, all models of the future are flawed.
-Seeking decisions that perform well even when assumptions are broken through robust design and flexibility generally improves outcomes.
-Yet whenever decisions are compared quantitatively, assumptions about the likelihood of different possible futures are necessarily made.
-We call for researchers studying climate risk management to make these implicit assumptions explicit, and we suggest that coordinated national and international guidance can help local governments and engineering bodies determine appropriate design criteria.\james{This ends abruptly.
-What does it mean for design criteria to be appropriate given the subjective assessments?
-What are the implications?}
-
-\section*{End Matter}
-
-\subsection*{Acknowledgements}
-
-This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under NOAA grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
-JDG thanks Rice University for support.
-KK thanks Dartmouth College for support.
-The authors thank Tor Erlend Fjelde and Vivek Srikrishnan for helpful comments.
-
-\subsection*{Code and data availability}
-
-All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
-This code is written in the open source Julia programming language.
-
-\subsection*{Author Contributions}
-
-JDG and KK designed the research.
-JDG wrote the codes and ran simulations.
-JDG wrote the manuscript in consultation with KK.
-JDG and KK revised the manuscript.
-
-\printbibliography
-
-\appendix
+      % Add an S to figures, tables, and sections
       \newcommand{\hbAppendixPrefix}{S}
+      \renewcommand{\thesection}{\hbAppendixPrefix \arabic{section}}
       \renewcommand{\thefigure}{\hbAppendixPrefix\arabic{figure}}
@@
 
-\newpage
-
-\printglossary[type=\acronymtype,title=List of Abbreviations]
-
       \section{Supplemental methods}
@@
       To avoid ambiguity, we define here some statistical distributions used in the main text.
-Except where otherwise indicated, we follow the default notation of the Distributions.jl Julia package \citep{besancon_distributions.jl:2021}.
+      Except where otherwise indicated, we follow the default notation of the Distributions.jl Julia package \cite{besancon_distributions.jl:2021}.
 
@@
       Evaluating expected annual damages for each of $J$ simulations of \gls{slr}, each of $N$ draws from the posterior distribution of storm surge, and each of $T$ time steps for $K$ models requires $N \times J \times T \times K$ simulations.
-In our model, we have $T=71$, $N=\num{10000}$, $J=\num{179232}$, and so exhaustive sampling may be prohibitive.
+      In our model, we have $T=70$, $N=\num{10000}$, $J=\num{179232}$, hence exhaustive sampling may be prohibitive.
 
@@
 
-\section{Supplemental figures}
+      %Repeat for any additional Supporting audio files
+
+      %%% End of body of article:
+      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+      %
+      % Optional Notation section goes here
+      %
+      % Notation -- End each entry with a period.
+      % \begin{notation}
+      % Term & definition.\\
+      % Second term & second definition.\\
+      % \end{notation}
+      %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+      %% ------------------------------------------------------------------------ %%
+      %%  REFERENCE LIST AND TEXT CITATIONS
+
+      \bibliography{library-bibtex}
+
+      %Reference citation instructions and examples:
+      %
+      % Please use ONLY \cite and \citeA for reference citations.
+      % \cite for parenthetical references
+      % ...as shown in recent studies (Simpson et al., 2019)
+      % \citeA for in-text citations
+      % ...Simpson et al (2019) have shown...
+      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
+      %
+      %
+      %...as shown by \citeA{jskilby}.
+      %...as shown by \citeA{lewin76}, \citeA{carson86}, \citeA{bartoldy02}, and \citeA{rinaldi03}.
+      %...has been shown \cite<e.g.,>{jskilbye}.
+      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
+      %...has been shown \cite{lewin76,carson86,bartoldy02,rinaldi03}.
+      %
+      % apacite uses < > for prenotes, not [ ]
+      % DO NOT use other cite commands (e.g., \citet, \citep, \citeyear, \nocite, \citealp, etc.).
+      %
+
+      %% ------------------------------------------------------------------------ %%
+      %
+      %  END ARTICLE
+      %
+      %% ------------------------------------------------------------------------ %%
+\end{article}
+\clearpage
+
+% Copy/paste for multiples of each file type as needed.
+
+% enter figures and tables below here: %%%%%%%
 
@@
             Depth-damage relationship.
-        Following \citet{zarekarizi_suboptimal:2020}, we use the Hazard U.S. (HAZUS) depth-damage curves provided by \gls{fema}.
-        Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \citep{huizinga_depthdamage:2016}.
+            Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves.
+            Since results are sensitive to choice of depth-damage equation, we illustrate (for comparison only) the ``Europa'' depth-damage relationship developed by the Joint Research Center (JRC) of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
       }\label{fig:cost-depth-damage}
@@
       \caption{
-        Following \citet{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \citep{johnson_clara:2013}.
+            Following \citeA{zarekarizi_suboptimal:2020}, we model the cost of elevating a single-family house by interpolating estimates from the Coastal Louisiana Risk Assessment Model \cite{johnson_clara:2013}.
             According to this model, the unit cost of elevating a house by 3-7, 7-10, and 10-14 feet is \usd{82.50}, \usd{86.25}, and \usd{103.75} per square foot, respectively, with a \usd{20745} initial cost.
@@
             Prior distributions for annual maximum storm surge.
-        Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citet{coles_evd:1996}.
+            Rather than apply a prior over model parameters directly, we apply a weakly informative prior over quantiles of the resulting distribution (that is, over a function of the model parameters) following \citeA{coles_evd:1996}.
             See \cref{sec:case-surge} for details.
@@
       \caption{
-        Surge prior
+            Implicit prior over storm surge represents large uncertainty but is consistent with basic physical reasoning (see \cref{sec:case-surge}).
+            The prior median (blue line) and uncertainty quantiles (shading) of the return period for different return levels of storm surge at Sewells Point, VA using prior predictive sampling \cite{gelman_workflow:2020}.
+            This illustrates the weakly informative prior information (\cref{sec:case-surge}) used in the analysis.
+            These weakly informative priors were selected to be consistent with basic physical principles (\eg, storm surges most years are $>$ \SI{2}{ft} but rarely exceed \SI{20}{ft}).
       }\label{fig:surge-prior-return}
@@
             Synthetic data experiment as a positive control test for the \gls{gev} model of storm surge.
-        A synthetic record was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively (dots).
+            A synthetic record (dots) was sampled from a \gls{gev} distribution with location, scale, and shape parameters of 4, 0.5, and 0.15, respectively.
             These samples were used to fit the Bayesian \gls{gev} model described in \cref{sec:case-surge}; the gray shading indicates the 50, 80, and 95\% posterior confidence intervals.
@@
       \caption{
-        As \cref{fig:surge-posterior-chains} but for draws from the prior distribution.
+            As \cref{fig:surge-posterior-chains}, but for draws from the prior distribution.
       }\label{fig:surge-prior-chains}
@@
             Observed values near the mode of the posterior predictive distribution are consistent with, but do not guarantee, a good fit.
-        For further discussion of posterior predictive checks, see Chapter 6 of \citet{gelman_bda3:2014}.
+            For further discussion of posterior predictive checks, see Chapter 6 of \citeA{gelman_bda3:2014}.
       }\label{fig:surge-test-statistics}
@@
 
-\section{Supplemental tables}
-
 \begin{table}[h]
@@
             Statistics include the mean and standard deviation of each parameter, the naive standard error and Monte Carlo standard error (which measure uncertainty in the mean), the effective sample size, $\hat{R}$ diagnostic, and effective samples per second, which describes sampling speed.
-        In general, a $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence.
+            The naive standard error (SE) returns the standard error of the mean.
+            The Monte Carlo standard error (MCSE) is calculated using the initial monotone sequence estimator \cite[pp.~473-483]{geyer_practical:1992}.
+            The effective sample size (ESS) is a crude estimate of the number of independent samples and is calculated following \citeA[pp.~473-483]{geyer_practical:1992}.
+            The $\hat{R}$ is an indicator of the convergence of the Markov chains to the target distribution; a value $\hat{R}$ value close to one is consistent with, though does not guarantee, convergence \cite{mcelreath_rethinking2:2020}.
+            For additional details see the \texttt{MCMCDiagnosticTools} package documentation.
       }\label{tab:surge-posterior-mcmc-diagnostics}
-    %\input{../../plots/surge-posterior-mcmc-diagnostics.tex}
+      \begin{tabular}{ccccccc}
+            \toprule
+            $\textrm{Parameter}$ & $\textrm{Mean}$ & $\textrm{Stdev.}$ & $\textrm{Naive SE}$ & $\textrm{MCSE}$ & $\textrm{ESS}$ & $\hat{R}$ \\
+            \midrule
+            $\mu$                & $3.610$         & $0.059$           & $0.001$             & $0.001$         & $4922.462$     & $1.001$   \\
+            $\xi$                & $0.231$         & $0.079$           & $0.001$             & $0.001$         & $4609.415$     & $1.001$   \\
+            $\sigma$             & $0.504$         & $0.049$           & $0.000$             & $0.001$         & $4547.881$     & $1.001$   \\
+            \bottomrule
+      \end{tabular}
+
 \end{table}
@@
       \caption{As \cref{tab:surge-posterior-mcmc-diagnostics} but for draws from the prior distribution.}\label{tab:surge-prior-mcmc-diagnostics}
-    %\input{../../plots/surge-prior-mcmc-diagnostics.tex}
+      \begin{tabular}{ccccccc}
+            \toprule
+            $\textrm{Parameter}$ & $\textrm{Mean}$ & $\textrm{Stdev.}$ & $\textrm{Naive SE}$ & $\textrm{MCSE}$ & $\textrm{ESS}$ & $\hat{R}$ \\
+            \midrule
+            $\mu$                & $4.774$         & $1.702$           & $0.005$             & $0.011$         & $26657.405$    & $1.000$   \\
+            $\xi$                & $0.246$         & $0.215$           & $0.001$             & $0.002$         & $11238.145$    & $1.000$   \\
+            $\sigma$             & $0.682$         & $0.531$           & $0.002$             & $0.004$         & $24720.263$    & $1.000$   \\
+            \bottomrule
+      \end{tabular}
+
 \end{table}

### 2022-06-22 scripts/01-surge-modeling.jl

@@ -122,7 +122,7 @@ function plot_surge_synthetic_experiment(annual::HouseElevation.AnnualGageRecord
 
     # add data
     xp, ys = HouseElevation.weibull_plot_pos(fake_data)
-    scatter!(p, 1 ./ xp, ys; label="Obs (Weibull Plot Pos.)", color=colors[1], alpha=1)
+    scatter!(p, 1 ./ xp, ys; label="Observations", color=:gray, alpha=1)
 
     plot!(p; legend=:topleft)
 
@@ -234,7 +234,7 @@ function plot_surge_posterior_return(
 
     # add weibul plot pls
     xp, ys = HouseElevation.weibull_plot_pos(surge_ft)
-    scatter!(p, 1 ./ xp, ys; label="Obs (Weibull Plot Pos.)", color=colors[1], alpha=1)
+    scatter!(p, 1 ./ xp, ys; label="Observations", color=:gray, alpha=1)
     return p
 end
 
@@ -249,7 +249,7 @@ function plot_surge_prior_return() where {T<:HouseElevation.MCMCChains.Chains}
     df = DataFrames.DataFrame(prior_fits)
     gevs = [GeneralizedExtremeValue(row[:μ], row[:σ], row[:ξ]) for row in eachrow(df)]
 
-    p = plot_return_period(gevs)
+    p = plot_return_period(gevs; type="Prior")
     savefig(p, plots_dir("surge-prior-return.pdf"))
     return p
 end
@@ -266,7 +266,9 @@ function plot_surge_obs_return(
     p1 = plot!(p1; yaxis=:log, yticks=(y_ticks, string.(y_ticks)))
     p2 = plot!(p2; yaxis=:log, yticks=(y_ticks, string.(y_ticks)))
 
-    add_panel_letters!([p1, p2]; fontsize=12, loc=(0.05, 0.95))
+    fontsize = 10
+    annotate!(p1, (0.05, 0.95), text("(A)", :left, fontsize))
+    annotate!(p2, (1/1000, 0.95), text("(B)", :left, fontsize))
     p = plot(
         p1,
         p2;

### 2022-06-22 scripts/06-combining.jl

@@ -30,7 +30,7 @@ function plot_grid_scheme()
         size=(800, 350),
         yticks=(y_ticks, string.(y_ticks)),
     )
-    scatter!(p, ψ, zeros(size(ψ)); label="Sampled Points", color=colors[2], markersize=5)
+    scatter!(p, ψ, zeros(size(ψ)); label="Sampled Points", color=colors[3], markersize=5)
 
     ymax_prev = 0.0
     x_arrow = range(minimum(xlimits), 0; length=length(ψ) + 2)[2:(end - 1)]
@@ -51,20 +51,25 @@ function plot_grid_scheme()
             linestyle=:dot,
         )
         plot!(p, [xmax, xmax], [0, ymax]; label=false, color=colors[3], linestyle=:dash)
+        
+        
+        # weights
         plot!(
             p,
             [x_arrow[i], x_arrow[i]],
             [ymax_prev, ymax];
             color=colors[5],
             linewidth=3,
-            label=false,
+            label= i==1 ? "Weights" : false,
         )
+
         annotate!(
             p,
             [x_arrow[i] + 0.2],
             [(ymax_prev + ymax) / 2],
             text(L"$w_%$i$", :center, 10; color=colors[5]),
         )
+        
         ymax_prev = ymax
     end
     savefig(p, plots_dir("grid-sketch.pdf"))
@@ -181,20 +186,11 @@ function plot_prior_tradeoffs(
             legend=ifelse(var == first(vars), :topright, false),
         )
 
-        # plot all trade-off lines in light gray
-        for rcp in all_rcp
-            for model in all_models
-                w = weights([(si.rcp == rcp) & (si.model == model) for si in s])
-                cond = vec(mean(var, w; dims=1))
-                plot!(p, Δh_ft, cond; label="", color=:gray, linewidth=0.5, alpha=0.5)
-            end
-        end
-
         # plot the trade-off for the priors
         for (prior, color) in zip(priors, colors)
             w = HouseElevation.make_weights(prior.dist)
             cond = vec(mean(var, w; dims=1))
-            plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=3)
+            plot!(p, Δh_ft, cond; label=prior.name, color=color, linewidth=4.5)
         end
 
         # add the cost on upper x axis

### 2022-06-22 scripts/main.jl

@@ -23,9 +23,7 @@ function main()
     house_value_usd = 200_000.0 # HOUSE NOT LAND VALUE
     discount_rate = 0.025 # mortgages going for 3-4.5% at the moment
     syear = 2022
-    eyear = 2092
-    # for a **VERY** rough idea of prices see
-    # https://www.zillow.com/homedetails/9638-Selby-Pl-Norfolk-VA-23503/79223088_zpid/
+    eyear = syear + 70 - 1 # 70 year lifetime including first year
 
     # get the storm surge posterior samples
     fits = get_surge_posterior()

### 2022-06-22 scripts/plotutils.jl

@@ -1,4 +1,5 @@
 using Colors
+using ColorSchemesa
 using Distributions
 using Plots
 
@@ -47,7 +48,9 @@ end
 get_colormap() = get_roma_colormap()
 
 function plot_return_period(
-    gevs::Vector{<:Distributions.GeneralizedExtremeValue}
+    gevs::Vector{<:Distributions.GeneralizedExtremeValue};
+    color_scheme=ColorSchemes.algae,
+    type="Posterior",
 ) where {T<:HouseElevation.MCMCChains.Chains}
     rts = range(1.25, 500; length=250) # return periods
     aeps = 1 .- 1 ./ rts # annual exceedance probability
@@ -68,20 +71,21 @@ function plot_return_period(
         ub = [quantile([quantile(d, xi) for d in gevs], qup) for xi in aeps]
         lb = [quantile([quantile(d, xi) for d in gevs], qlow) for xi in aeps]
         range_pct = Int(range * 100)
+        fillcolor = ColorSchemes.get(color_scheme, range - 0.5)
         plot!(
             p,
             rts,
             ub;
             fillbetween=lb,
-            fillcolor=:gray,
-            fillalpha=0.35,
+            fillcolor=fillcolor,
+            fillalpha=1,
             linecolor=false,
             label="$(range_pct)% Credible Interval",
         )
     end
 
     median = [quantile([quantile(d, xi) for d in gevs], 0.50) for xi in aeps]
-    plot!(p, rts, median; label="Posterior Median")
+    plot!(p, rts, median; color=ColorSchemes.get(color_scheme, 1.0), label="$(type) Median", linewidth=2)
 
     return p
 end

### 2022-06-23 papers/earths-future/submission.tex
Make XR package work on Overleaf
James local edits esp w/r/t abstract & intro

@@
 % cite things in supplemental info file
-\usepackage{xr}
-\externaldocument{supplemental}
+% see https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
+% it's a big strange for Overleaf!
+\usepackage{xr-hyper}
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+\myexternaldocument{supplemental}
 
@@
 \begin{keypoints}
-  \item Decisions about how to manage climate risks are often made under deep uncertainty
+  \item Decisions about how to manage climate risks are often made under deep uncertainty.
   \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
-  \item We demonstrate the framework using a simple case study of coastal flood risk management
+  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management
 \end{keypoints}
@@
 \begin{abstract}
-  Projections of future climate risks can vary considerably between sources without a  consensus about their  relative probability.
+  Projections of future climate risks can vary considerably between sources without consensus regarding their relative probability.
   This deep uncertainty poses considerable communication and decision-analytical challenges.
@@
   Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
-  These approaches (i) can underestimate risks, (ii) provide little insights into which assumptions drive decision-relevant outcomes, and (iii) can nudge how people choose their decisions-making frameworks.
+  These approaches can underestimate risks, hide deep uncertainties, and provide little insights into which assumptions drive decision-relevant outcomes.
   Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform decision support.
@@
 These uncertainties often arise from choices experts have to make, for example about technical detail on how to model future water levels.
-Different experts can disagree about these choices and provide different plausible projections.
+Different experts can disagree about these choices, leading to different plausible projections.
 Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
-For one, just picking one of these projections or just an average can lead to poor decisions.
-On the other extreme, communicating results separately for all projections can overwhelm decision-makers.
+At one extreme, just picking a single representative projection can lead to under-estimation of risk and poor decisions.
+At the other extreme, communicating results separately for each projection can overwhelm decision-makers.
 To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
 We develop and demonstrate a framework to address these challenges.
-The framework provides a transparent approach (i) to combine a large number of deeply uncertain projections to a more understandable and smaller sample and (ii) to provide insights as to how assumptions and modeling choices made by experts influence decisions.
+The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller sample and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
 We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections about future water levels.
@@
 
-Many critical infrastructure services around the world are aging and inadequate for a changing climate \cite<\eg,>[]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
-Looking to the future, it is expected that changes to regulation, economics and finance, patterns of population and infrastructure use, as well as climate, will further stress infrastructure systems.
-This motivates the question: which possible futures should infrastructure systems and components be designed for?
+Many critical infrastructure services around the world are aging and inadequate, and anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet societal needs \cite<\eg,>[]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,ho_dams:2017}.
+The need to upgrade existing infrastructure and build new facilities motivates the question: for which possible futures should infrastructure systems and components be designed?
 The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statue or industry guidelines.
@@
 However, there are also limitations.
-In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs.
+In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
 This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
@@
 
-Currently used methods for estimating these \glspl{pdf} emphasize data-driven, nominally objective methods that can be applied consistently across locations.
-For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall.
+Currently methods for estimating the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
+For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
 Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
@@
 While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
-Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations (see \cite{bulletin17c:2019}, p.~2, or \cite{atlas14_texas:2018} p.~A.4-42).
+Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.
 
@@
 
-Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet; see>[]{deconto_antarctica:2016}.
+Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
 In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
@@
 This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \cite{bernardo_bayesian:1994}.
-Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \cite{gelman_philosophy:2013}, \cite{ramsey_probability:2016}, or \cite{jaynes_probability:2003} for a discussion of Bayesian philosophy and \cite{Piironen:2017eh} for a technical discussion of methods for model selection).
+Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
 From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \cite{gelman_workflow:2020}.

### 2022-06-23 papers/earths-future/submission.tex
Updates from Overleaf

@@
 
-All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2021-elevation-robustness}.
+All code, including source code, is available under the GNU Public License (version 3) at \url{https://github.com/jdossgollin/2022-elevation-robustness}.
 This code is written in the open source Julia programming language and detailed instructions for reproducing our results are provided.

### 2022-07-02 papers/earths-future/submission.tex
incorporate james and adam proofreading

@@
 \begin{keypoints}
-  \item Decisions about how to manage climate risks are often made under deep uncertainty.
+  \item Decisions about how to manage climate risks are often made under deep uncertainty
   \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
@@
 
-%% ------------------------------------------------------------------------ %%
-%
-%  ABSTRACT and PLAIN LANGUAGE SUMMARY
-%
-% A good Abstract will begin with a short description of the problem
-% being addressed, briefly describe the new data or analyses, then
-% briefly states the main conclusion(s) and how they are supported and
-% uncertainties.
-
-% The Plain Language Summary should be written for a broad audience,
-% including journalists and the science-interested public, that will not have 
-% a background in your field.
-%
-% A Plain Language Summary is required in GRL, JGR: Planets, JGR: Biogeosciences,
-% JGR: Oceans, G-Cubed, Reviews of Geophysics, and JAMES.
-% see http://sharingscience.agu.org/creating-plain-language-summary/)
-%
-%% ------------------------------------------------------------------------ %%
-
-%% \begin{abstract} starts the second page
-
 \begin{abstract}
-  Projections of future climate risks can vary considerably between sources without consensus regarding their relative probability.
-  This deep uncertainty poses considerable communication and decision-analytical challenges.
+  Projections of future climate risks can vary considerably from one source to another, posing considerable communication and decision-analytical challenges.
   One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
   Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
-  These approaches can underestimate risks, hide deep uncertainties, and provide little insights into which assumptions drive decision-relevant outcomes.
-  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform decision support.
-  The first step of this workflow is to create an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
-  Next, a small set of probability distributions representing subjective beliefs over possible futures is used to weight the scenarios.
+  These approaches can underestimate risks, hide deep uncertainties, and provide little insight into which assumptions drive decision-relevant outcomes.
+  Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform climate risk management.
+  The first step of this workflow is to generate an ensemble of simulations representing possible futures and analyze them through standard exploratory modeling techniques.
+  Next, a small set of probability distributions representing subjective beliefs about the likelihood of possible futures is used to weight the scenarios.
   Finally, these weights are used to compute and characterize trade-offs, conduct robustness checks, and reveal implicit assumptions.
@@
 What are sound strategies to manage risks driven by climatic changes?
-Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems.
-These uncertainties often arise from choices experts have to make, for example about technical detail on how to model future water levels.
+Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
+These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
 Different experts can disagree about these choices, leading to different plausible projections.
 Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
-At one extreme, just picking a single representative projection can lead to under-estimation of risk and poor decisions.
-At the other extreme, communicating results separately for each projection can overwhelm decision-makers.
+At one extreme, picking a single representative projection can lead to under-estimation of risk and poor decisions.
+At the other, communicating results separately for each projection can overwhelm decision-makers.
 To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
 We develop and demonstrate a framework to address these challenges.
-The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller sample and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
-We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections about future water levels.
-
-%% ------------------------------------------------------------------------ %%
-%
-%  TEXT
-%
-%% ------------------------------------------------------------------------ %%
+The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller samples and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
+We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.
 
@@
 
-Many critical infrastructure services around the world are aging and inadequate, and anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet societal needs \cite<\eg,>[]{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,ho_dams:2017}.
+Many critical infrastructure services around the world are aging and inadequate, and anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet societal needs \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,ho_dams:2017}.
 The need to upgrade existing infrastructure and build new facilities motivates the question: for which possible futures should infrastructure systems and components be designed?
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statue or industry guidelines.
+The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines.
 For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
-At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different future conditions.
+At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.
 
@@
 
-Current practice in engineering, infrastructure design, and local governance relies heavily upon standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
+Current practice in engineering, infrastructure design, and local governance relies heavily on standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
 These are often, though not always, informed by probabilistic analysis of relevant data.
@@
 Additionally, local building codes \cite<based on guidance such as>[]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
-Design standards are also used to design large-scale infrastructure.
-For example, some levees in the Netherlands are required to be designed such that the annual probability of failure is less than 1 in \num{4000} \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+Probabilistic analysis also informs the design of large-scale infrastructure.
+For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
 
@@
 
-Currently methods for estimating the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
+The approahces currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
 Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
 While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
@@
 For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
-While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and, indeed, can be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
+While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and can often be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
 This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
@@
 We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
-This is a theoretical contrast to the ``$\mathcal{M}$''-closed case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose as well as the ``$\mathcal{M}$''-complete case in which a true model exists but is inaccessible in the sense that even though it can be conceptualized it cannot be written down \cite{bernardo_bayesian:1994}.
-Whereas model selection in the $\mathcal{M}$-closed case seeks to asymptotically identify the true model from a set of candidate models, model selection in the $\mathcal{M}$-open case emphasizes a subjective view, both of probability itself and of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them \cite{gelman_workflow:2020}.
+This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose \cite{bernardo_bayesian:1994}.
+Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
+By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
+From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model is reached \cite{gelman_workflow:2020}.
 
@@
 We proceed as follows.
-In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs.
+In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs (\cref{sec:analysis-synthesize}).
 In \cref{sec:case-study} we describe the case study in detail.
 Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
-Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for research, policy, and practice.
+Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for policy and practice.
 
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seek to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
+Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
@@
 One way to interpret an ensemble of \glspl{sow} is as \gls{iid} draws from some true data generating process.
-We call this scenario-conditional probabilistic analysis, which commonly arises when a probabilistically calibrated model for relevant outputs (\eg, a physical model for \gls{msl}; see \cref{sec:case-slr}) is forced by a deterministic scenario (\eg, a \gls{rcp} scenario).
-In this case, \emph{the set of available \glspl{sow} is assumed to come from the true (often called population) distribution, and thus the sample statistics are assumed to reflect the true values.}
+This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a fully probabilistic model; we call this scenario-conditional probabilistic analysis.
 We add this concept to our framework with boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario (\ie, the assumed input) $M_k$.
 If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
-This approach allows for quantification of uncertainty \emph{within models}, but can only qualitatively characterize uncertainty \emph{between models} \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
-% Vivek raises an interesting point -- what if we incorporate the choice of model into the data generating process and use a posterior probability of each model being true -- but I think that's basically what we are calling step #3
+This approach allows for quantification of uncertainty \emph{within} models of deeply uncertain processes, but can only qualitatively characterize uncertainty \emph{between} models \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
 
@@
 
-If the \gls{sow} ensemble is not drawn \gls{iid} from the true distribution (\eg, if one deliberately over-samples low-probability but high-impact \glspl{sow}) or if one wishes to synthesize \emph{across} multiple scenarios $M_1, M_2, \ldots, M_K$, then additional measures must be taken.
-One possibility is to discard the \glspl{sow} ($\vb{s}$) and resample from some ``true'' distribution, which we shall denote $p_\mathrm{belief}$.
-In practice, this is impractical because simulating $\vb{s}$ may rely on complicated models that are computationally expensive to re-run.
+As discussed in the previous subsection, estimating the expectation of arbitrary functions under uncertainty is a critical component of decision making under uncertainty.
+If the \glspl{sow} are drawn \gls{iid} from the true distribution, then the expected value of some function  $f$, $\mathbb{E}\qty[f(\vb{s})]$, can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
+However, this is often not the case; for example, a fixed set of simulations may be available for analysis or low-probability but high-impact regions of the parameter space may have been sampled.
+In this case, the formula must be adjusted to
+\begin{equation}
+  \mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{i=j}^N w_j f(\vb{s}_j),
+\end{equation}
+where the $w_j$ are suitably chosen weights such that their sum is equal to 1.
 
-Instead of attempting to re-sample $\vb{s}$, we instead consider re-weighting.
-A simple approach would be to assign each \gls{sow} $s_j$ a weight $w_j$ equivalent to the ratio of the desired (often called target distribution in survey analysis) distribution, $p_\mathrm{belief}$ and the sampling distribution, which could be approximated through methods such as kernel density estimation.
+The challenge then becomes suitable estimation of the $w_j$.
 Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
-This allows us to specify a prior belief over this reduced space $p_\mathrm{belief}(\Psi)$ instead of over the full space $p_\mathrm{belief}(\Omega)$.
-We then calculate a probabilistic weight $w_j \in \qty[0, 1]$ for each \gls{sow} $s_j$ so that the weighted distribution of $\psi_j$ closely approximates $p_\mathrm{belief}(\Psi)$.
+Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
 
-We present here the case where the $\psi_j$ are one-dimensional.
+We present here the case where the $\psi_j$ are one-dimensional; extensions to higher dimensions are possible.
 We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
@@
 This step is illustrated in \cref{fig:grid-sketch}.
-For higher dimensional projections, this equation can be extended by partitioning the parameter space, then integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
-From a practical perspective, many simulation models round outputs, meaning that there is a possibility of having $\psi_j$ that are equal; in this case it may be helpful to add a small amount of noise to the $\psi_j$ before computing the weights.
 Diagnostic checks, such as examining the histogram of weights (not shown), may be valuable protections against degeneracy.
@@
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because its there is no ``true'' value of $p_\mathrm{belief}$ that could be objectively ascertained \cite{oreskes_verification:1994,walker_deep:2013}.
+We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
@@
 To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
-We focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
+Following \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
 For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
@@
 \begin{enumerate}
-  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$ f
+  \item The decision vector $\vb{x}$ is comprised of discrete possible house heightenings ($\Delta h$); we consider $\Delta h = \qty{\SI{0}{ft}, \SI{0.25}{ft}, \ldots, \SI{12}{ft}}$.
   \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
@@
 \Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-We return in \cref{sec:results-conditional} to the problem that these multiple models poses to decision makers.
+We return in \cref{sec:results-conditional} to the challenge of decision making under multiple \glspl{pdf}.
 
@@
 Other model validations lend confidence to the stationary \gls{gev} model selected.
-For example, \cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges; the estimated return period of the (shading) matches the (dots) the empirical plotting position dots.
-Further, a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
+For example, \cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges; the estimated return period  (shading) matches the the empirical plotting position (dots) and a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
 
@@
 Following \citeA{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \cite{fischbach_clara:2012}.
-We normalize this cost by house value.
-This cost curve is shown in \cref{fig:cost-up-front}.
+We normalize this cost by house value; this cost curve is shown in \cref{fig:cost-up-front}.
   \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
@@
 We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.
-We can also compare to the analyses of \citeA{sriver_sealevel:2018} which use a rescaled Beta distribution with bounds of \SIrange{255}{2508}{\milli\meter} (\SIrange{0.83}{8.2}{ft}) and a most plausible estimate of \SI{950}{\milli\meter} (\SI{3.1}{ft}).
+We can also compare to the analyses of \citeA{sriver_sealevel:2018} which uses a rescaled Beta distribution with bounds of \SIrange{0.83}{8.2}{ft} and a most plausible estimate of \\SI{3.1}{ft}.
 Our samples bound all of these estimates.
@@
     Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
-    The lowest-cost outcomes occur when both exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
+    The lowest-cost outcomes occur when exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
     The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
@@
 For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \cite<\eg, those of>[]{sweet_slr:2022} may prove sufficient for decision making.
-This would be analogous to decision scaling methods that plot climate model projections on a scenario map of water system performance as a function of changes in key climate variables \cite{Brown:2012kb,Steinschneider:2015kk}.
-Scenario mapping is also commonly used in \gls{rdm} analyses \cite{lempert_rdm:2019,sriver_sealevel:2018}.
 However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
@@
 Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draw from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
-As discussed in \cref{sec:analysis-condition}, this allows a formal estimation of decision metrics.
+As discussed in \cref{sec:analysis-condition}, this allows a formal estimation of decision metrics, conditional on the chosen scenario.
 
@@
 For example, under the most pessimistic model (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
-Converseley, under the most optimistic model (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost of 73\% of house value yet reduces lifetime expected damages by less than 50\% of house value.
+Under the most optimistic model (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost yet reduces lifetime expected damages by less than 50\% of house value.
 
@@
 This approach is attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
-However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from.
+However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from and hence many trade-off curves to navigate.
 The multiple \gls{pdf} problem has also been shown in other contexts.
@@
 Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
-And \citeA{lempert_rdm:2019}
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
@@
 One application of this method is to diagnose which assumptions different $p_\mathrm{belief}$ are consistent with.
-\Cref{fig:lsl-priors-weights}(B-D) shows the average weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
+\Cref{fig:lsl-priors-weights}(B-D) shows the total weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
 For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
@@
 Under an assumption of rapid \gls{slr}, elevating the house by approximately \SI{6}{ft} costs 73\% of house value and reduces damages by over 100\% of house value, yielding a benefit to cost ratio of approximately 1.25.
-Under an assumption of slow \gls{slr}, elevating the house by \SI{6}{ft} reduces damages only by 50\% of house value, yielding a benefit to cost ratio of approximately 0.7.
+Under an assumption of slow \gls{slr}, the same decision reduces damages only by 50\% of house value, yielding a benefit to cost ratio of approximately 0.7.
 Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house, and thus the benefit to cost ratio is nearly 1.
@@
 
-This study develops a framework to increase the transparency of assumptions about the likelihood of different \glspl{sow}.
-Additionally, we develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
+This study develops a framework to increase the transparency of quantitative decision analysis under deep uncertainty.
+We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
 Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
@@
 
-There are several limitations to our study that merit further discussion.
-The first category has to do with limitations of the underlying methodology proposed for reweighting state of the world.
+Several limitations to our study merit further discussion.
+The first category has to do with limitations of the underlying methodology proposed for reweighting \glspl{sow}.
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
@@
 More fundamentally, our method assumes that there exists an expert capable of integrating over the many processes that drive \gls{slr}, from global greenhouse gas emissions to the global carbon cycle to climate sensitivity and ice sheet response \cite{morgan_elicitation:2014}.
-An alternative approach would be to build a probabilistic model for each of these steps, and to use each as input to the next to develop a fully probabilistic model for \gls{slr}.
+An alternative approach would be to build a probabilistic model for each of these steps, and to use each as an input to the next to develop a fully probabilistic model for \gls{slr}.
 Yet while some progress has been made developing probabilistic models for specific elements of this model chain \cite<\eg,>[]{srikrishnan_probabilistic:2022,wong_brick0.2:2017}, this remains a computational and conceptual challenge.
@@
 Finally, while here we consider the decision to be a one-time decision, one could also frame this as problem a sequential decision.
-The aaalysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \cite{herman_control:2020}.
+The analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \cite{herman_control:2020}.
 Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \cite{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \cite{herman_control:2020}.
@@
 These limitations motivate several directions for future research.
-From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \cite<see fig.~1 of>[]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimates of local hazard, subject to (still necessarily subjective) priors over key parameters.
+From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \cite<see fig.~1 of>[]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimation of local hazard, subject to (still necessarily subjective) priors over key parameters.
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
@@
 The proposed scenario weighting framework can be applied to inform critical challenges in climate risk management.
-An obvious area of application is to the design of infrastructure for climate risk management.
+An obvious area of application is to the design of infrastructure.
 For example, much of the stormwater infrastructure in the United States is inadequate for current climate \cite{lopez-cantu:2018}, let alone future changes.
@@
 Similarly, decisions like  levee heightening \cite{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
-This framework can also be applied to uncertainties in non-physical variables.
-For example, investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \cite{trindade_deeplyuncertainpathways:2019}.
-Similarly, analyses of climate change mitigation options, such as estimates of the social cost of pollutants \cite{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
+Investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \cite{trindade_deeplyuncertainpathways:2019}.
+And analyses of climate change mitigation options, such as estimates of the social cost of pollutants \cite{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
 
 Of course, all models are ultimately wrong \cite{box:1976}.
-Thus seeking decisions that perform well across a range of assumptions, and improving the deccision space through robust design and flexibility, can improve outcomes.
+Thus seeking decisions that perform well across a range of assumptions, and improving the decision space through robust design and flexibility, can improve outcomes.
 Yet whenever decisions are compared quantitatively, assumptions about the probability of different possible futures are necessarily made.
@@
 KK thanks Dartmouth College for support.
-The authors thank Tor Erlend Fjelde and Vivek Srikrishnan for helpful comments.
+The authors thank
+Tor Erlend Fjelde,
+Vivek Srikrishnan,
+Adam Pollack,
+and Sitara Baboolal
+for helpful comments.
 

### 2022-07-02 scripts/main.jl

@@ -18,10 +18,15 @@ include("06-combining.jl")
 
 function main()
 
+    # uses notation from paper:
+    # s = vector of scenarios
+    # x = decision vector
+    # u = metrics
+
     # define some constants
     house_floor_area = 1500u"ft^2"
     house_value_usd = 200_000.0 # HOUSE NOT LAND VALUE
-    discount_rate = 0.025 # mortgages going for 3-4.5% at the moment
+    discount_rate = 0.025
     syear = 2022
     eyear = syear + 70 - 1 # 70 year lifetime including first year
 
@@ -43,7 +48,6 @@ function main()
     end
 
     # make some plots of the sea level (BRICK) data
-    # uses notation from paper: s = set of all scenarios
     let
         s = HouseElevation.get_lsl(; syear=syear, eyear=2125)
         plot_lsl_evolution(s)
@@ -90,6 +94,8 @@ function main()
             house_value_usd=house_value_usd,
             overwrite=false,
         )
+
+        # plot expected costs as a function of Δh annd LSRL
         plot_scenario_map_height_slr(; x=x, s=s, u=u, house_value_usd=house_value_usd)
 
         # plot tradeoffs by RCP scenario

### 2022-07-03 papers/earths-future/submission.tex
Klaus handwritten edits, some questions pending

@@
-\documentclass[draft]{agujournal2019}
+\documentclass{agujournal2019}
 
@@
 \newacronym[]{usgs}{USGS}{United States Geological Survey}
-\newacronym[plural=SOWs,firstplural=states of the world (SOWs),descriptionplural=states of the world]{sow}{SOW}{state of the world}
+\newacronym[\glslongpluralkey={states of the world}]{sow}{SOW}{state of the world}
 
@@
 
-\draftfalse
+%\draftfalse
 
@@
 \begin{keypoints}
-  \item Decisions about how to manage climate risks are often made under deep uncertainty
-  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making
-  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management
+  \item Decisions about how to manage climate risks are often made under deep uncertainty.
+  \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
+  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
 \end{keypoints}
@@
 At one extreme, picking a single representative projection can lead to under-estimation of risk and poor decisions.
-At the other, communicating results separately for each projection can overwhelm decision-makers.
+At the other extreme, communicating results separately for each projection can overwhelm decision-makers.
 To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
@@
 
-Many critical infrastructure services around the world are aging and inadequate, and anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet societal needs \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,ho_dams:2017}.
-The need to upgrade existing infrastructure and build new facilities motivates the question: for which possible futures should infrastructure systems and components be designed?
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines.
-For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings.
+Many critical infrastructure services around the world are aging \cite{ho_dams:2017}.
+Further, anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet their design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
+The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
+The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines \cite{bruneau_multihazard:2017}.
+For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings \cite{asce_7-10:2013}.
 At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.
@@
 These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
-Additionally, local building codes \cite<based on guidance such as>[]{asce_24-05:2006,FEMA_p-55:2011} may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
-Probabilistic analysis also informs the design of large-scale infrastructure.
+Additionally, local building codes (based on guidance such as \citeA{asce_24-05:2006} or \citeA{FEMA_p-55:2011}) may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
+Probabilistic analysis also informs the design of large-scale infrastructure.in particu
 For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
@@
 There are many advantages to standards-based design.
-In particular, these heuristics are scalable and explainable, do not require complex analysis to apply, and are fair in at least a procedural sense.
-However, there are also limitations.
+In particular, these heuristics are scalable and explainable, they reduce the complexity of design analysis, and they are fair in at least a procedural sense.
+However, reliance on these heuristics also faces limitations.
 In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
 This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
-These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and may be particularly sensitive to representation of tail probabilities.
+These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and can be particularly sensitive to representation of tail probabilities \cite{merz_heavytails:2022,wong_floodrisk:2018,garner_slrise:2018}.
 
-The approahces currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
+The approaches currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
 For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
 Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
-While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{DossGollin:2019,Montanari:2014hl,Serinaldi:2015bq}.
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.\footnote{Klaus comment TBD}
+While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,Montanari:2014hl,Serinaldi:2015bq}.
 Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.
@@
 
-Projecting nonstationary hazard is difficult because many future hazards depend on intrinsically unpredictable human decisions (\eg, the rate of future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
+Projecting nonstationary hazard is difficult because many future hazards depend on human decisions (\eg, future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
 In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
@@
 
-Although these methods for \gls{dmdu} have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
+Although these \gls{dmdu} methods have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
 Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
@@
 In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
-We find parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
-This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which a true model can be identified and written down but is one amongst finitely many models from which an analyst has to choose \cite{bernardo_bayesian:1994}.
+There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
+This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
 Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
 By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model is reached \cite{gelman_workflow:2020}.
+From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.
 
@@
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} strongly influence results \cite{quinn_exploratory:2020}.
+Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020}.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite{schneider_scenarios:2002}.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite<see>{schneider_scenarios:2002}.
 
@@
 
-The challenge then becomes suitable estimation of the $w_j$.
+The challenge then becomes to suitably estimate the $w_j$.
 Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
@@
 
-Data on storm surge comes from Sewells Point, VA (gauge 8638610) from the freely available NOAA tides and currents data archive \cite{noaa_tidesandcurrents:2022}.
+We use data on storm surge at Sewells Point, VA (gauge 8638610) from the \gls{noaa} tides and currents data archive \cite{noaa_tidesandcurrents:2022}.
 Hourly recordings of water level are available from 1928 to the present; we use data from the period January 1, 1928 to December 31, 2021.
@@
 
-For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{Betancourt:2017vd,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
+For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{betancourt_hamiltonian:2018,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
 Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
-Following \citeA{zarekarizi_suboptimal:2020}, we use the we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
+Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
 For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
@@
 The expected annual damage is sometimes calculated by assuming analytically tractable functional forms for the depth-damage relationship and for the  distribution of hazard \cite<\eg,>[]{vandantzig_dike:1956}.
-However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution so we instead estimate it through a Monte Carlo method (see \cref{sec:alg-ead} for details).
+However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution.
+Instead, we estimate this convolution through a Monte Carlo method (see \cref{sec:alg-ead} for details).
 Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).
@@
 We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.
-We can also compare to the analyses of \citeA{sriver_sealevel:2018} which uses a rescaled Beta distribution with bounds of \SIrange{0.83}{8.2}{ft} and a most plausible estimate of \\SI{3.1}{ft}.
+We can also compare to the analyses of \citeA{sriver_sealevel:2018} which uses a rescaled Beta distribution with bounds of \SIrange{0.83}{8.2}{ft} and a most plausible estimate of \SI{3.1}{ft}.
 Our samples bound all of these estimates.
@@
   \caption{
-    Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \gls{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
-    Colors indicate the density of \glspl{sow}; the color of each grid box corresponds to the number of \glspl{sow} falling within that box.
-    The lowest-cost outcomes occur when exposure is low ($h_0$ is large and \gls{slr} is minimal) and the house is not elevated (no up-front cost).
+    Scenario maps show the dependence of expected lifetime cost (damages plus up-front cost) as a function of \acrfull{msl} in 2100 for several values of initial height ($h_0$) and house elevation ($\Delta h$).
+    Colors indicate the number of \acrfullpl{sow} of falling within each box.
+    The lowest-cost outcomes occur when exposure is low ($h_0$ is large and \acrfull{slr} is minimal) and the house is not elevated (no up-front cost).
     The highest-cost outcomes arise when exposure is high ($h_0$ is small and \gls{slr} is rapid) and investment is inadequate.
@@
 
-One application of explotatory analysis is to reveal the range and variation in outcomes, conditional on taking a particular decision.
+One application of exploratory analysis is to reveal the range and variation in outcomes, conditional on taking a particular decision.
 \Cref{fig:scenario-map-slr-cost} shows the dependence of expected lifetime costs (damages plus up-front costs; $y$-axis) as a function of  \gls{slr} over the house lifetime ($x$-axis), height increase ($\Delta h$; columns), and initial elevation ($h_0$; rows).
@@
 The proposed approach can help overcome the limitations of scenario-conditional analysis.
-We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can shed light on climate risk management under deep uncertainty.
+We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can help to shed light on climate risk management under deep uncertainty.
 We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
@@
 For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
-Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 (also unlikely given current policy) and \gls{rcp} 4.5.
-The uncertain \gls{slr} scenario (blue line) places approximately equal weight across models.
+Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 \cite<also unlikely given current policy;>{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}) and \gls{rcp} 4.5.
+The uncertain \gls{slr} scenario (blue line) allocates approximately equal weight across models.
 
@@
 
-This study develops a framework to increase the transparency of quantitative decision analysis under deep uncertainty.
+This study develops a framework designed to increase the transparency of quantitative decision analysis under deep uncertainty.
 We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
@@
 
+\subsection{Implications for research and practice}
+
 Several limitations to our study merit further discussion.
-The first category has to do with limitations of the underlying methodology proposed for reweighting \glspl{sow}.
+The first category has to do with limitations of the underlying method proposed for reweighting \glspl{sow}.
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased estimates of the posterior \cite{csillery_abc:2010,marjoram_abc:2006}.
+Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased posterior estimates \cite{csillery_abc:2010,marjoram_abc:2006}.
 Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences not represented in our model.
@@
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
-Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that scale to different house characteristics may add value.
+Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that are applicable across different house characteristics may improve useability.
+
+\subsection{Conclusions and implications}
 
@@
 An obvious area of application is to the design of infrastructure.
-For example, much of the stormwater infrastructure in the United States is inadequate for current climate \cite{lopez-cantu:2018}, let alone future changes.
+For example, much of the stormwater infrastructure in the United States is inadequate for current climates \cite{lopez-cantu:2018}, let alone future changes.
 Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \cite{sharma_rcp:2021} and \gls{rcp} scenarios.
@@
 
-Of course, all models are ultimately wrong \cite{box:1976}.
+Of course, all models are ultimately wrong \cite{box_sciencestatistics:1976}.
 Thus seeking decisions that perform well across a range of assumptions, and improving the decision space through robust design and flexibility, can improve outcomes.
@@
 This code is written in the open source Julia programming language and detailed instructions for reproducing our results are provided.
-A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at TBD. % TODO: create Zenodo release!
+A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at TBD.\footnote{Will create Zenodo release} % TODO: create Zenodo release!
 

### 2022-07-03 scripts/04-scenario-maps.jl

@@ -72,10 +72,15 @@ function plot_scenario_map_slr_cost(;
                     :title_align => :center,
                 )
             elseif row == N_row
+                kwargs[:xformatter] = identity
                 push!(
-                    kwargs, :xlabel => "LSLR: $syear to $eyear [ft]", :bottommargin => 7.5mm
+                    kwargs,  :bottommargin => 7.5mm
                 )
-                kwargs[:xformatter] = x -> x # update it
+                if col == 2
+                    push!(
+                    kwargs, :xlabel => "SLR: $syear to $eyear [ft]"
+                )
+                end
             end
 
             # left column has ylabel, right hgas h₀
@@ -227,7 +232,7 @@ function plot_scenario_map_height_slr(;
         x_ft,
         msl_plot,
         expected_cost;
-        ylabel="LSLR: $syear to $eyear [ft]",
+        ylabel="SLR from $syear to $eyear [ft]",
         colorbar_title="Expected Total Costs\n[% House Value]",
         colorbar_formatter=pct_formatter,
         c=cgrad(:plasma; rev=true),

### 2022-07-05 papers/earths-future/submission.tex
incorporate rest of klaus handwritten suggestions

@@
 \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
-\usepackage{soul}
 
@@
 \usepackage{graphicx}
-\usepackage[list=true]{subcaption}
 \graphicspath{{../../plots/}{../../tikz/}{../../img/}}
@@
 
-Many critical infrastructure services around the world are aging \cite{ho_dams:2017}.
-Further, anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet their design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
+Many critical infrastructure services around the world are aging \cite<\eg,>{ho_dams:2017}.
+Moreover, anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet their design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
 The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
@@
 Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.\footnote{Klaus comment TBD}
+For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
 While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,Montanari:2014hl,Serinaldi:2015bq}.
@@
 Projecting nonstationary hazard is difficult because many future hazards depend on human decisions (\eg, future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
-In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002}.
+In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
 These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
@@
 Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
-Finally in \cref{sec:conclusions} we discuss key findings, future research needs, and implications for policy and practice.
+In \cref{sec:caveats} we discuss limitations of the study and future research needs.
+Finally in \cref{sec:conclusions} we discuss key findings and implications for policy and practice.
 
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite<see>{schneider_scenarios:2002}.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.
 
@@
 To illustrate the proposed decision analytic framework, we model a one-time decision of whether to elevate a house, and if so by how much (\cref{fig:xlrm}).
-Following \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
+Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
 For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
@@
     For each \acrshort{sow} (red) and each value of $\Delta h$, the system model (blue) is used to calculate performance metrics (gray).
+    We also compute a third metric, expected lifetime costs, which is the sum of up-front costs and lifetime expected damages.
   }\label{fig:xlrm}
@@
 
-\begin{table}[h]
+\begin{table}[ht]
   \centering
@@
     Colored lines highlight three representative models for emphasis.
+    Gray lines and the smiling face in the lower left corner indicate the direction of preference.
   }\label{fig:tradeoffs-by-rcp}
@@
 For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
-Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
+Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios and \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different scenarios of \gls{slr}.
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
@@
 
-\section{Conclusions}\label{sec:conclusions}
-
-This study develops a framework designed to increase the transparency of quantitative decision analysis under deep uncertainty.
-We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
-Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-We use a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
-
-\subsection{Implications for research and practice}
+\section{Discussion and research needs}\label{sec:caveats}
 
@@
 
-\subsection{Conclusions and implications}
+\section{Conclusions}\label{sec:conclusions}
+
+This study develops a framework designed to increase the transparency of quantitative decision analysis under deep uncertainty.
+We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
+Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
+We use a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
 
@@
 An obvious area of application is to the design of infrastructure.
-For example, much of the stormwater infrastructure in the United States is inadequate for current climates \cite{lopez-cantu:2018}, let alone future changes.
+For example, much of the stormwater infrastructure in the United States is inadequate for current and anticipated future climates \cite{lopez-cantu:2018}.
 Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \cite{sharma_rcp:2021} and \gls{rcp} scenarios.

### 2022-07-05 scripts/04-scenario-maps.jl

@@ -131,6 +131,7 @@ function plot_scenario_map_slr_cost(;
     cbar_fake_plot[1][:xaxis][:showaxis]
     cbar_fake_plot[1][:yaxis][:showaxis]
 
+    # add the horizontal arrow
     harrow = plot(
         [0.2, 0.9],
         [0, 0];
@@ -146,6 +147,7 @@ function plot_scenario_map_slr_cost(;
         harrow, [0.5], [0.1], text(L"Increasing House Elevation ($\Delta h$)", :center, 14)
     )
 
+    # add the vertical arrow
     varrow = plot(
         [0, 0],
         [0.1, 0.9];

### 2022-07-05 scripts/05-tradeoffs-by-rcp.jl

@@ -41,10 +41,8 @@ function plot_rcp_tradeoffs(
             linewidth=2,
             xticks=(x_ticks, string.(x_ticks)),
             yformatter=pct_formatter,
-            top_margin=12.5Plots.mm,
-            left_margin=7.5Plots.mm,
-            bottom_margin=7.5Plots.mm,
             legend=ifelse(var == first(vars), :topright, false),
+            bottom_margin=7.5Plots.mm,
         )
 
         i = 1 # counter for colors
@@ -83,8 +81,56 @@ function plot_rcp_tradeoffs(
         # add to our plots
         push!(p_archive, p)
     end
+
     add_panel_letters!(p_archive; fontsize=12, loc=(0.1, 0.95))
-    p = plot(p_archive...; layout=(1, 2), link=:x, size=(1100, 550))
+    for p in p_archive
+        annotate!(p, (0.05, 0.05), text("😃", :center, 16))
+    end
+
+    # add horizontal and vertical areas
+    harrow = plot(
+        [0.98, 0.02],
+        [0, 0];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:gray,
+        xlims=(0, 1),
+        ylims=(-0.1, 0.1),
+    )
+
+    # add the vertical arrow
+    varrow = plot(
+        [0, 0],
+        [0.98, 0.02];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:gray,
+        xlims=(-0.1, 0.1),
+        ylims=(0, 1),
+    )
+
+    blankplot = plot(; legend=false, grid=false, foreground_color_subplot=:white)
+
+    # arrange everything
+    l = grid(2, 4; heights=[0.025, 0.975], widths=[0.025, 0.475, 0.025, 0.475])
+
+    p = plot(
+        blankplot,
+        harrow,
+        blankplot,
+        harrow,
+        varrow,
+        p_archive[1],
+        varrow,
+        p_archive[2];
+        layout=l,
+        link=:x,
+        size=(1100, 500),
+    )
 
     savefig(plots_dir("tradeoffs-by-rcp.pdf"))
     return p

### 2022-07-05 scripts/06-combining.jl

@@ -208,8 +208,56 @@ function plot_prior_tradeoffs(
         # add to our plots
         push!(p_archive, p)
     end
-    add_panel_letters!(p_archive; fontsize=12)
-    p = plot(p_archive...; layout=(1, 2), link=:x, size=(1000, 500))
+
+    add_panel_letters!(p_archive; fontsize=12, loc=(0.1, 0.95))
+    for p in p_archive
+        annotate!(p, (0.05, 0.05), text("😃", :center, 16))
+    end
+
+    # add horizontal and vertical areas
+    harrow = plot(
+        [0.98, 0.02],
+        [0, 0];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:gray,
+        xlims=(0, 1),
+        ylims=(-0.1, 0.1),
+    )
+
+    # add the vertical arrow
+    varrow = plot(
+        [0, 0],
+        [0.98, 0.02];
+        axis=false,
+        ticks=false,
+        legend=false,
+        arrow=:closed,
+        color=:gray,
+        xlims=(-0.1, 0.1),
+        ylims=(0, 1),
+    )
+
+    blankplot = plot(; legend=false, grid=false, foreground_color_subplot=:white)
+
+    # arrange everything
+    l = grid(2, 4; heights=[0.025, 0.975], widths=[0.025, 0.475, 0.025, 0.475])
+
+    p = plot(
+        blankplot,
+        harrow,
+        blankplot,
+        harrow,
+        varrow,
+        p_archive[1],
+        varrow,
+        p_archive[2];
+        layout=l,
+        link=:x,
+        size=(1100, 500),
+    )
 
     savefig(plots_dir("tradeoffs-by-prior.pdf"))
     return p

### 2022-07-05 papers/earths-future/submission.tex
Klaus edits

@@
 Many critical infrastructure services around the world are aging \cite<\eg,>{ho_dams:2017}.
-Moreover, anticipated changes to regulation, economics and finance, patterns of population and infrastructure use, and climate challenge their ability to meet their design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
+Moreover, changes in regulations, finance, patterns of population and infrastructure use, and the climate system challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
 The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
@@
     Colored lines highlight three representative models for emphasis.
-    Gray lines and the smiling face in the lower left corner indicate the direction of preference.
+    The gray arrows indicate the direction of preference.  
   }\label{fig:tradeoffs-by-rcp}
@@
 For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
-Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios and \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different scenarios of \gls{slr}.
+Similarly, \citeA{wong_nola:2017} construct 18 probability distribution functions for future flood risk in New Orleans, considering multiple models for ice sheet dynamics and storm surge and multiple \gls{rcp} scenarios.
+As a last example, \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different \gls{slr} scenarios.
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}

### 2022-07-05 scripts/05-tradeoffs-by-rcp.jl
for preprint

@@ -83,9 +83,6 @@ function plot_rcp_tradeoffs(
     end
 
     add_panel_letters!(p_archive; fontsize=12, loc=(0.1, 0.95))
-    for p in p_archive
-        annotate!(p, (0.05, 0.05), text("😃", :center, 16))
-    end
 
     # add horizontal and vertical areas
     harrow = plot(

### 2022-07-05 scripts/06-combining.jl

@@ -61,7 +61,6 @@ function plot_grid_scheme()
             linewidth=3,
             label=i == 1 ? "Weights" : false,
         )
-
         annotate!(
             p,
             [x_arrow[i] + 0.2],
@@ -210,9 +209,6 @@ function plot_prior_tradeoffs(
     end
 
     add_panel_letters!(p_archive; fontsize=12, loc=(0.1, 0.95))
-    for p in p_archive
-        annotate!(p, (0.05, 0.05), text("😃", :center, 16))
-    end
 
     # add horizontal and vertical areas
     harrow = plot(

### 2022-07-05 papers/earths-future/submission.tex
for preprint
final word edits

@@
 
-% Authors are individuals who have significantly contributed to the
-% research and preparation of the article. Group authors are allowed, if
-% each author in the group is separately identified in an appendix.)
-
-% List authors by first name or initial followed by last name and
-% separated by commas. Use \affil{} to number affiliations, and
-% \thanks{} for author notes.
-% Additional author notes should be indicated with \thanks{} (for
-% example, for current addresses).
-
-% Example: \authors{A. B. Author\affil{1}\thanks{Current address, Antartica}, B. C. Author\affil{2,3}, and D. E.
-% Author\affil{3,4}\thanks{Also funded by Monsanto.}}
-
 \authors{James Doss-Gollin\affil{1}, Klaus Keller\affil{2}}
@@
 \correspondingauthor{James Doss-Gollin}{jdossgollin@rice.edu}
-
-%% Keypoints, final entry on title page.
-
-%  List up to three key points (at least one is required)
-%  Key Points summarize the main points and conclusions of the article
-%  Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-
-% Example:
-% \begin{keypoints}
-% \item	List up to three key points (at least one is required)
-% \item	Key Points summarize the main points and conclusions of the article
-% \item	Each must be 140 characters or fewer with no special characters or punctuation and must be complete sentences
-% \end{keypoints}
 
@@
 This code is written in the open source Julia programming language and detailed instructions for reproducing our results are provided.
-A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at TBD.\footnote{Will create Zenodo release} % TODO: create Zenodo release!
+A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at \url{https://doi.org/10.5281/zenodo.6799457}. % Update as needed
 

### 2022-07-06 papers/earths-future/submission.tex
JDG spell edits

@@
 Additionally, local building codes (based on guidance such as \citeA{asce_24-05:2006} or \citeA{FEMA_p-55:2011}) may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
-Probabilistic analysis also informs the design of large-scale infrastructure.in particu
+Probabilistic analysis also informs the design of large-scale infrastructure.
 For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
@@
 However, reliance on these heuristics also faces limitations.
-In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance tradeoffs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
+In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance trade-offs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
 This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
@@
 For example, robust decision making \cite{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
-When the decision space is complex, many analyses use policy search (often using multiobjective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
+When the decision space is complex, many analyses use policy search (often using multi-objective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
 In addition, many studies formally quantify robustness to deep uncertainties \cite{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
@@
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to inscrutable decision processes and suboptimal decisions.
+Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to inscrutable decision processes and sub-optimal decisions.
 
@@
 Several limitations to our study merit further discussion.
-The first category has to do with limitations of the underlying method proposed for reweighting \glspl{sow}.
+The first category has to do with limitations of the underlying method proposed for re-weighting \glspl{sow}.
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
@@
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
-Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that are applicable across different house characteristics may improve useability.
+Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that are applicable across different house characteristics may improve usability.
 

### 2022-07-08 papers/earths-future/submission.tex
Update funding

@@
 
-This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under \gls{noaa} grant NA16OAR4310179 and by the Penn State Center for Climate Risk Management.
+This work was supported by \acrfull{noaa} through the Mid-Atlantic Regional Integrated Sciences and Assessments (MARISA) program under \gls{noaa} grant NA16OAR4310179 and through the Penn State Initiative for Resilient Communities (PSIRC) by a Strategic Plan seed grant from the Penn State Office of the Provost, with co-support from the Center for Climate Risk Management (CLIMA), the Rock Ethics Institute, Penn State Law, and the Hamer Center for Community Design.
 JDG thanks Rice University for support.

### 2022-07-09 papers/earths-future/submission.tex
catalina comments

@@
 From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.
+Thus, where classical statistics, particularly as applied to environmental systems, has often emphasized objectivity, we draw from this Bayesian literature that  emphasizes a focus on building transparency in the presentation, analysis, and critique of models and results.
 
@@
   \item The \glspl{sow} describe annual time series of \gls{msl} over the $T=70$ year house lifetime so $\vb{s} \in \mathbb{R}^T$
-  \item The system model $f$ quantifies up-front costs and lifetime expected damages, given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
+  \item The system model $f$ quantifies up-front costs (the cost of elevating) and lifetime expected damages (the structural cost of experiencing floods), given a decision $x_i$ and \gls{sow} $s_j$, by integrating economic and engineering damage models over a probability distribution for storm surge.
+We elaborate upon these metrics in \cref{sec:case-metrics}.
 \end{enumerate}
@@
 The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
-These four models represent physical processes, particularly of ice sheet dynamics, in different ways, leading to diverging sensitivity of \gls{msl} to forcing.
+These four models represent physical processes, particularly of ice sheet dynamics, in different ways.
+The differences in these models leading to diverging estimates of how \gls{msl} responds to temperature.
 For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.
@@
 The authors thank
+Sitara Baboolal,
 Tor Erlend Fjelde,
-Vivek Srikrishnan,
+Catalina González-Dueñas,
 Adam Pollack,
-and Sitara Baboolal
+and
+Vivek Srikrishnan
 for helpful comments.

### 2022-07-09 papers/earths-future/submission.tex
update the doi in the paper

@@
 This code is written in the open source Julia programming language and detailed instructions for reproducing our results are provided.
-A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at \url{https://doi.org/10.5281/zenodo.6799457}. % Update as needed
+A permanent, citeable archive of the precise version of the codes used in this study is also available on Zenodo at \url{https://doi.org/10.5281/zenodo.6814588}.
 

### 2022-09-12 papers/earths-future/reviewer_response_01.tex
start addressing reviews

\documentclass{ar2rc}

\usepackage{siunitx}
\usepackage{apacite}
\usepackage{physics}

\usepackage[colorinlistoftodos]{todonotes}
\usepackage{color}
\newcommand{\jdg}[1]{\todo[color=red!30]{#1}}
\newcommand{\all}[1]{\todo[color=blue!30]{#1}}

% better lists
\usepackage{enumitem}
\setlist{nosep}

% cross-refs
\usepackage{xr}
%\externaldocument{./submission_02}
\hypersetup{hidelinks}
\usepackage{cleveref}

\title{A subjective Bayesian framework for synthesizing deep uncertainties in climate risk management}
\author{James Doss-Gollin and Klaus Keller}
\journal{Earth's Future}

\begin{document}

\maketitle

We thank the editors and both referees for their thoughtful, detailed, and extremely constructive comments on our initial submission.

Referee 1 particularly encouraged us to clarify the presentation of our methods and results.
As noted in responses to the specific comments given below, we have attempted to incorporate these suggestions into the revised manuscript.

Referee 2 encouraged us to (1) better contextualize our methodological approach relative to previous works, (2) clarify and improve the mathematical formulation of our approach, and (3) to better describe the usefulness of our method.
The specific

In the remainder of this document we address the specific comments made by the two referees.

\section{Referee \#1}

\RC{
    This work introduces a Bayesian framework to synthesize and characterize deep uncertainties to merge standard exploratory modeling techniques with more traditional decision support tools that require statistics like the expected value.
    Such statistics, which require weighting sampled states of the world, are problematic in the presence of deep uncertainty.
    The authors argue that many decision making frameworks commonly used for robust decision-making problems actually make such a weighting, but that is not transparent.
    Decisions about the sampling range, joint sampling distribution, and statistics (such as domain criterion) can have consequences for the robustness of the options presented to stakeholders, but this may not be clear.
    The proposed framework allows for a synthesis across deep uncertainties using subjective Bayesian priors.
    The method is demonstrated for a very simple house-raising example including both sea level rise and a stationary storm surge risk.
}

\AR{
    Thank you for your comments.
    This is an excellent summary of the aim and scope of this paper.
}

\RC{
    Overall, I found the work well written and informative, and I think it is very promising.
    The framework's novelty arises from the way it stitches standard methods from Bayesian statistics and DMDU.
    This is not a negative statement, as I think by employing common-sense methods in a novel way the authors' framework is more likely to gain traction.
    An important contribution is the introduction, which nicely lays out the current state of practice, emerging methods (primarily from the DMDU community), and then research gaps.
    I appreciated the author's care in discussing the difficulty in adopting non-stationary frequency methods for extremes, with key citations to important works by Montanari, Seranaldi, and Bulletin 17C's introduction.
}

\AR{}

\RC{
    The use of the word model to mean different things. Throughout the manuscript, and particularly in Figure 1, the authors utilize the term 'model' to refer to both the combination of RCP-SLR model/assumption and the systems model.
    I found this confusing at multiple places.
    One suggestion is to call the 'models' in box d of Figure 1 'scenarios' or something like that.
    So the scenario is the deep uncertainty (RCP-SLR model) and then the possible futures in box b of Figure 1 are samples drawn from the pdf that's conditioned on the scenario in Box d.
    I think this new terminology would be particularly important if one were to have more than one systems model in box c.
}

\AR{
    This is a tricky distinction to draw and we should be more clear.
    One possibility is to use Model and Probabilistic Scenario.
    Another is Physical Model and Predictive Model. Maybe something else?
}

\RC{
    I understand the intention of the paper is to showcase the method, not necessarily the results.
    This said, some more discussion about the findings would be helpful.
    For example, in Figure 7A there is a spike after 0 ft of increased height. If some detail were given on the construction cost function (taken from another work), and at least a few more sentences on the results, I think it would really improve an already very good manuscript.
}

\AR{
    Good suggestion, can do.
}

\RC{
    Line 270: I think more detail should be given, perhaps in the discussion, of the difficulty projecting SOW to some lower dimensional representation may present in more complicated problems.
    Here it is given little attention, which is fine because this is meant to be a simple didactic problem.
    But I could imagine a much more complicated setting, perhaps in a multisector problem, where collapsing the state space of the world presents problems.
    I suggest the authors provide some more discussion of the challenge this may present and how one might overcome it in the discussion.
}

\AR{
    Yes, we have a bit of discussion about this but more would be helpful.
}

\RC{
    Line 284: One really exciting direction is the opportunity to test the sensitivity of the objective tradeoffs to $p_\textup{belief}$.
    It may be the case that some objectives are not affected by the differences in $p_\textup{belief}$ between stakeholders and so there is no point in getting tied up trying to pin it down.
    I'm not proposing any additional analysis, only that this sort of screening is possible and is consistent with the spirit of exploratory modeling and RDM.
    It would be an interesting exercise, perhaps for follow-on work.
}

\AR{
    Yes, good suggestion and we can elaborate on this ni the discussion.
}

\RC{
    Table 1: I think at least some detail should be given about the elevation cost and depth-damage functions.
    I understand they were taken from another work. This would make interpreting the later results easier.
}

\AR{
    Yes, can do, especially since the shape of these curves matter a lot.
    For example, findings that elevating a little bit is ineffective in nearly all circumstances stem from the shape of the construction cost curve.
}

\RC{
    Figure 5b.
    If the intention is to show the agreement of the fitted model with the annual maximum storm surges, a P-P plot might be clearer.
    I leave this up to the authors.
    I have no doubt the GEV is a good fit.
}

\AR{
    The purpose here is to show the probability distribution using a standard return period plot.
    But a Q-Q or P-P plot in the supplemental material to assess fit would be good.
}

\RC{
    Figure 4b.
    There is remarkable disagreement between the models of mean sea level rise.
    It might be worth at least a few sentences commenting on this.
}

\AR{
    We mainly defer to \citeA{ruckert_coastal:2019} on this, but we can add a bit of discussion.
}

\RC{
    Equation 4: it might be helpful to define $p$ here.
}

\AR{
    Can and should do.
}

\RC{
    Line 416: The use of the gamma distribution is quite sensible but is presented without justification.
    This might be worth just a short sentence on why it was chosen.
}

\AR{
    Good suggestion.
}

\RC{
    Figure 8: The y-axis labels on B-D should be more descriptive.
}

\AR{
    Yes, can do.
}

\section{Referee \#2}

\RC{
    This manuscript addresses an important topic: the role of probabilistic methods in decision-making under deep uncertainty.
    The authors point out that scenario-based DMDU methods implicitly put probabilities on scenarios but do not make these assumptions transparent.
    They address this by developing a method to calculate the weights implicitly applied to each deeply uncertain scenario.
    I applaud this effort -- the tension between scenario-based methods that use optimization and probabilistic methods is an important and understudied one.
    I also see a lot of value in how the authors have framed the problem and described the pros and cons of scenario-based DMDU methods and probabilistic methods for CBA in the context of infrastructure design.
    The simple case study of raising the elevation of a house under flood risk affected by deeply uncertain sea level rise is appropriate for demonstrating the method and is technically sound.
}

\AR{
    Thanks\ldots
}

\RC{
    However, I have three major concerns that prevent me from recommending the manuscript for publication in its current form.
    I defer to the editor on whether, assuming all these concerns are addressed in revision, the contribution warrants publication in Earth's Future.
}

\AR{
    Thanks for the insightful comments.
}

\RC{
    First, the paper is not well grounded in previous literature at the intersection of Bayesian /probabilistic methods and DMDU.
    This has led the authors to make overly broad claims about their contribution that make it difficult to discern what they are adding beyond existing methods.
    It is certainly not true that ``In this paper we offer a first conceptual step towards bridging the fields of DMDU and Bayesian model building'' as claimed in line 183.
    Here are some papers that integrate Bayesian/probabilistic methods into DMDU contexts.
    The first three are especially similar to the authors' contributions; the authors need to make a compelling argument for what their approach adds beyond this existing literature.
    \citeA{reis_distribution:2020},
    \citeA{taner_probabilistic:2019},
    \citeA{hui_adaptive:2018},
    \citeA{fletcher_learning:2019},
    \citeA{fletcher_groundwater:2019}
}

\AR{
    Thanks, can do.
    I haven't read \citeA{reis_distribution:2020}, \citeA{taner_probabilistic:2019}, or \citeA{hui_adaptive:2018} but will do.
    We can add a paragraph to the introduction describing this intersection, possibly with a few additional papers.
}

\RC{
    My second key concern is that the mathematical formulation is not clear and complete, and it is not always clear how the sample application aligns with the general framework.
    Given that the authors suggest the mathematical formulation is a key part of their contribution, this is important.
    Specifically:
    The method is framed both in the introduction (lines 157-182) and methods section (lines 240-250) around Bayesian model selection.
    However, it is not clear if or how the authors are performing Bayesian model selection.
    In figure 1e and line 247, the authors provide a conditional probability of $u$ given $x_i$ and $M_k$, and state this is the result of IID draws from $M_k$.
    This is similar to the $p(y|M_k)$ term in the classical BMA framework from \citeA{raftery_bma:2005} equation 1.
    Does this imply that there is a Bayesian posterior for model weights developed?
    If so, what is it?
    What data is used for the updating and what is the likelihood function? It is unclear if or how Bayes' is used here.
}

\AR{
    This is a great question and one we need to clarify.
    I think specifically that the way we talk about Bayesian model selection is a bit unclear.
    We also could describe the limitations of BMA (which is suited to the M-closed case).
    Our claim here is that there is no updating because everything is based on the prior, but it's true in principle that one could think about how this would be updated; we can discuss this a bit.
}

\RC{
    Shortly later in the paragraph under section 2.3, the authors take an expected value of $f(s)$ and state that it ``can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\bf{s}_j)$.''
    I believe this is inconsistent with the interpretation of the SOW being samples from an underlying distribution.
    We know that: $\mathbb{E}\qty[f(s)]=\sum_{s \in \Omega}f(s)P_s(s)$ where $P_s$ is the probability mass function of $s$.
    This is equivalent to the authors' equation above only when assuming that $P_s(s)$ is a discrete uniform RV where each SOW gets equal probability.
    However, this seems to be at odds with the authors interpretation that the SOW are IID draws from some underlying distribution of models.
    If that is the case, the probability mass or density function shouldn't simply put probability on the available SOW, as these are merely samples from an underlying distribution.
    It conflates the sample with the population.
    If, for example, the SOW were believed to be IID samples from a lognormal distribution, the we would expect an estimate for $\mathbb{E}\qty[f(s)]$ that is asymmetric across the $f(s)$ in different SOW, not an unweighted average.
}

\AR{
    This is another insightful comment, and I think we can frame this a bit better in the literature on non-naive Monte Carlo sampling.
    $\mathbb{E}\qty[f(s)] \approx \frac{1}{N} \sum_{j=1}^N s_j$ is a naive Monte Carlo estimate of the expectation, which is only valid if the SOW are IID draws from $P_s$.
    This is often done in practice, however, and referring to other papers as suggested may be helpful for pointing this out.
}

\RC{
    I am further confused by equation 1 and the following paragraph on choosing the weights.
    First, are the weights “chosen” by the analyst, as stated in line 266, or are they estimated as a way of making clear what weights are implied by a certain selection of SOW?
}

\AR{
    Let's make sure it is clear that they are implied by a certain $p_\text{belief}$ and sampling distribution of SOWs
}

\RC{
    Second, what is the mathematical basis for equation 2?
    The description in 267- 272 describes the method illustrated in figure 2 but doesn't justify it, and I could not find the method in the references provided.
    It seems like the approach is a way of approximating some unknown continuous distribution of $M_k$ using the discrete sampled SOW, but the authors do not state this.
    If this is true, it could potentially be used to address my comment above about conflating sample and population.
}

\AR{
    This is just a definition, but we can cite something like \citeA{johnson_clara:2013} that uses grid-based methods
}

\RC{
    Finally, I am confused how the application in section 4.3 is equivalent to this.
    Line 413 states ``we construct three probabilistic models for $p_\textup{belief}$'' and then refers to these models in line 418 as ``priors''.
    The use of the word “prior” implies a Bayesian model.
    Does this suggest that equation (2) is meant to be a prior distribution? It is not described that way in section 2.3. If so, what is the posterior? What is the likelihood? My interpretation is that the “target distribution” in figure 2 is F belief, but that the weights over SOW are just an approximation of that, not a posterior.
}

\AR{
    Another good point -- need to clarify that refer to priors because we don't have any data about the future so we are relying entirely on prior beliefs.
    But we're not applying Bayes' rule.
}

\RC{
    It is clear the storm surge model in 3.2 uses a Bayesian model to fit a GEV using historical storm surge data. But this is presented as a probabilistic uncertainty, not a deep uncertainty, so it remains unclear how Bayesian methods are used to synthesize deep uncertainties.
}

\AR{
    We can discuss this point.
}

\RC{
    It is not clear to me if there is an error in the mathematical presentation or if it simply needs to be clarified. It is possible also that this is a valid probabilistic framework, but not a Bayesian one, in which case the approach would need to be substantially reframed. If I have misunderstood due to lack of clarity in the presentation, I welcome a clarification from the authors.
}

\AR{
    I think we can clarify and need to.
    In part, at least, the term ``Bayesian'' seems to be suggesting a method of estimation, which is not what we are doing.
}

\RC{
    My final key critique is that I think the authors need to make a more compelling case for the usefulness of their approach relative to existing DMDU methods to warrant
    publication. If I am interpreting correctly, the two new aspects of the method are 1) the conditioning step (section 2.2 which I believe though am not sure is the same as what is applied in section 4.2) and 2) the deep uncertainty synthesis. The key result from the conditioning step is show in Figures 6 and 7. It seems to me that Figure 6 is quite similar to visuals commonly used in exploratory analysis like parallel axis plots or scatter plots showing the results against different objectives across SOWs. Take as an example Figure 6 from Quinn et al. 2020. This plot shows frequency and severity of water shortage as measured by cumulative frequency across SOWs in the experiment, visualized with darker colors for more SOWs. This is a common type of figure in exploratory analysis used to transparently visualize the results across SOW. The authors figure 6 seems quite similar to this. What is more useful or transparent about your conditioning approach? I mean that as a genuine question – I have found the methods difficult to follow and would appreciate clarification from the authors.
}

\AR{
    Thanks, we can add a bit to the discussion.
}

\RC{
    Similarly, Figure 8 shows the key results from the weighting step.
    As I understand it, panel a shows the authors illustrative assumptions for sea level rise distributions, and then panels b-d show what weights across SOWs are consistent with those assumptions. What should a decision maker takeaway from this? How might they use it in their decision process? It would be helpful to see this articulated in the example and expanded on in the discussion or conclusions.
}

\AR{
    This should come across more clearly in the conclusions.
}

\bibliographystyle{apacite}
\bibliography{library-bibtex}

\end{document}

### 2022-09-12 papers/earths-future/submission_02.tex
start addressing reviews

--- 2022-07-09 papers/earths-future/submission.tex
@@
 
-What are sound strategies to manage risks driven by climatic changes?
-Addressing this question is complicated by the large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
+The identification of sound strategies to manage risks driven by climatic changes is challenged by large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
 These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
-Different experts can disagree about these choices, leading to different plausible projections.
+Different experts can disagree about these choices, leading to different projections.
 Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
-At one extreme, picking a single representative projection can lead to under-estimation of risk and poor decisions.
-At the other extreme, communicating results separately for each projection can overwhelm decision-makers.
+For example, picking a single representative projection can lead to under-estimation of risk and poor decisions.
+Similarly, communicating results separately for each projection can overwhelm decision-makers.
 To make matters worse, typical approaches to this problem are mostly silent on what assumptions make a difference for the decisions at hand.
 We develop and demonstrate a framework to address these challenges.
-The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller samples and (ii) provide insights as to how assumptions and modeling choices made by experts influence decisions.
+The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller sample set and (ii) provide insights about which assumptions and modeling choices influence decisions.
 We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.

### 2022-09-20 papers/earths-future/reviewer_response_01.tex
some more progress

@@
 \setlist{nosep}
+
+% some commands
+\usepackage{xspace}
+\makeatletter
+\DeclareRobustCommand\onedot{\futurelet\@let@token\@onedot}
+\def\@onedot{\ifx\@let@token.\else.\null\fi\xspace}
+\def\eg{\emph{e.g}\onedot} \def\Eg{\emph{E.g}\onedot}
+\def\ie{\emph{i.e}\onedot} \def\Ie{\emph{I.e}\onedot}
+\def\etc{\emph{etc}\onedot} \def\vs{\emph{vs}\onedot}
+\newcommand{\usd}[1]{\SI{#1}[\$]{}}
+
+% ACRONYMS
+\usepackage[acronym, nopostdot, nonumberlist, shortcuts, numberedsection, nogroupskip,]{glossaries}
+\newacronym{bfe}{BFE}{base flood elevation}
+\newacronym{cmip}{CMIP}{the Coupled Model Intercomparison Project}
+\newacronym{dmdu}{DMDU}{decision making under deep uncertainty}
+\newacronym{fema}{FEMA}{the Federal Emergency Management Agency}
+\newacronym{gev}{GEV}{generalized extreme value}
+\newacronym{hazus}{HAZUS}{Hazard U.S.}
+\newacronym{iid}{IID}{independent and identically distributed}
+\newacronym{ipcc}{IPCC}{International Panel on Climate Change}
+\newacronym{msl}{MSL}{mean relative sea level}
+\newacronym{noaa}{NOAA}{the National Oceanic and Atmospheric Administration}
+\newacronym{pdf}{PDF}{probability density function}
+\newacronym{rcp}{RCP}{representative concentration pathway}
+\newacronym{rdm}{RDM}{robust decision making}
+\newacronym{slr}{SLR}{sea level rise}
+\newacronym{ssp}{SSP}{shared socio-economic pathway}
+\newacronym[]{usace}{USACE}{United States Army Corps of Engineers}
+\newacronym[]{usgs}{USGS}{United States Geological Survey}
+\newacronym[\glslongpluralkey={states of the world}]{sow}{SOW}{state of the world}
 
@@
     The method is demonstrated for a very simple house-raising example including both sea level rise and a stationary storm surge risk.
+
+    Overall, I found the work well written and informative, and I think it is very promising.
+    The framework's novelty arises from the way it stitches standard methods from Bayesian statistics and DMDU.
+    This is not a negative statement, as I think by employing common-sense methods in a novel way the authors' framework is more likely to gain traction.
+    An important contribution is the introduction, which nicely lays out the current state of practice, emerging methods (primarily from the DMDU community), and then research gaps.
+    I appreciated the author's care in discussing the difficulty in adopting non-stationary frequency methods for extremes, with key citations to important works by Montanari, Seranaldi, and Bulletin 17C's introduction.
 }
@@
 \RC{
-    Overall, I found the work well written and informative, and I think it is very promising.
-    The framework's novelty arises from the way it stitches standard methods from Bayesian statistics and DMDU.
-    This is not a negative statement, as I think by employing common-sense methods in a novel way the authors' framework is more likely to gain traction.
-    An important contribution is the introduction, which nicely lays out the current state of practice, emerging methods (primarily from the DMDU community), and then research gaps.
-    I appreciated the author's care in discussing the difficulty in adopting non-stationary frequency methods for extremes, with key citations to important works by Montanari, Seranaldi, and Bulletin 17C's introduction.
-}
-
-\AR{}
-
-\RC{
     The use of the word model to mean different things.
-Throughout the manuscript, and particularly in Figure 1, the authors utilize the term 'model' to refer to both the combination of RCP-SLR model/assumption and the systems model.
+Throughout the manuscript, and particularly in Figure 1, the authors utilize the term `model' to refer to both the combination of RCP-SLR model/assumption and the systems model.
     I found this confusing at multiple places.
@@
     The simple case study of raising the elevation of a house under flood risk affected by deeply uncertain sea level rise is appropriate for demonstrating the method and is technically sound.
-}
 
-\AR{
-    Thanks\ldots
-}
-
-\RC{
     However, I have three major concerns that prevent me from recommending the manuscript for publication in its current form.
@@
     It is certainly not true that ``In this paper we offer a first conceptual step towards bridging the fields of DMDU and Bayesian model building'' as claimed in line 183.
-    Here are some papers that integrate Bayesian/probabilistic methods into DMDU contexts.
-    The first three are especially similar to the authors' contributions; the authors need to make a compelling argument for what their approach adds beyond this existing literature.
+    Here are some papers that integrate Bayesian/probabilistic methods into DMDU contexts:
     \citeA{reis_distribution:2020},
@@
     \citeA{hui_adaptive:2018},
-    \citeA{fletcher_learning:2019},
-    \citeA{fletcher_groundwater:2019}
+    \citeA{fletcher_learning:2019}, and
+    \citeA{fletcher_groundwater:2019}.
+    The first three are especially similar to the authors' contributions; the authors need to make a compelling argument for what their approach adds beyond this existing literature.
 }
@@
 \AR{
-    Thanks, can do.
-    I haven't read \citeA{reis_distribution:2020}, \citeA{taner_probabilistic:2019}, or \citeA{hui_adaptive:2018} but will do.
+    Thanks very much for this helpful comment.
+    This comment underscores a need for improved clarity in the introduction to the manuscript; as noted, there are numerous papers that use probabilistic tools, including Bayesian methods, in DMDU contexts.
+    We have clarified that there is a long literature on the relative merits of probabilistic vs non-probabilistic information in DMDU contexts, as noted in \citeA{taner_probabilistic:2019}, but that this debate has often centered on a top-down, technocratic, objectivist philosophy of probability in which there is explicitly or implicitly assumed to be a true probability distribution.
+    Our contribution is to outline how a subjective Bayesian philosophical approach can address many of the concerns raised by critics of probabilistic approaches while providing a principled and transparent method for reasoning about uncertainty.
+
     We can add a paragraph to the introduction describing this intersection, possibly with a few additional papers.
+    \citeA{hui_adaptive:2018} is a nice example of the use of SDP applied to a levee heightening problem.
+    There is Bayesian updating through the dynamic programming, but deep uncertainty is addressed only by considering several scenarios of hydrological conditions.
+    \citeA{taner_probabilistic:2019} is conceptually quite similar to our approach.
+    This paper focuses on joint distributions of uncertain parameters rather than how different subjective beliefs about these parameters affect the decision.
+    This paper also focuses on how probabilistic versus exploratory frameworks lead to different robustness values, whereas here we focus more generally on mapping trade-offs.
+    \citeA{reis_distribution:2020} shows that scenario discovery and exploratory modeling results are sensitive to parameterization and choice of distribution.
+    \citeA{fletcher_learning:2019} and \citeA{fletcher_groundwater:2019} apply SDP (which can be interpreted as having an implicit Bayesian updating step) but shallow uncertainties are considered.
 }
+
+\begin{quote}
+    Add a sentence to the intro and link it here
+\end{quote}
 
@@
     This is a great question and one we need to clarify.
-    I think specifically that the way we talk about Bayesian model selection is a bit unclear.
-    We also could describe the limitations of BMA (which is suited to the M-closed case).
-    Our claim here is that there is no updating because everything is based on the prior, but it's true in principle that one could think about how this would be updated; we can discuss this a bit.
+    For reference, equation 1 of \citeA{raftery_bma:2005} is
+    $$p(y) = \frac{1}{K} \sum_{k=1}^K p(y|M_k) p(M_k | y^T),$$
+    ``where $p( y | M_k)$ is the forecast PDF based on model $M_k$ alone, and $p(M_k | y^T)$ is the posterior probability of model $M_k$ being correct given the training data, and reflects how well model $M_k$ fits the training data.''
+
+    This comment is closely related to the previous comment in that we draw inspiration from the literature on Bayesian model selection but do not explicitly perform it.
+    Our approach is conceptually similar to BMA as noted (and even close to stacking; see \cite{Yao:2018bu}).
+    There are two key differences.
+    First, these methods focus on estimating the weights assigned to each model from data; in our context we have no data.
+    Future work could consider adaptive decision making (by revisiting the decision at multiple time steps, \ie stochastic control) in which case there would be data available in the future that could be used to estimate these weights as in several of the papers referenced in the previous comment.
+    Second, these methods assign weights to each model, whereas we assign weights to each SOW.
+    This is motivated by the fact that the SOWs in our applications, and in many climate adaptation applications, are computationally expensive to generate.
 }
+
+\begin{quote}
+    Need to think about how we want to address this in the text
+\end{quote}
 
@@
     I believe this is inconsistent with the interpretation of the SOW being samples from an underlying distribution.
-    We know that: $\mathbb{E}\qty[f(s)]=\sum_{s \in \Omega}f(s)P_s(s)$ where $P_s$ is the probability mass function of $s$.
+    We know that:
+    $$\mathbb{E}\qty[f(s)]=\sum_{s \in \Omega}f(s)P_s(s)$$
+    where $P_s$ is the probability mass function of $s$.
     This is equivalent to the authors' equation above only when assuming that $P_s(s)$ is a discrete uniform RV where each SOW gets equal probability.
@@
     This is another insightful comment, and I think we can frame this a bit better in the literature on non-naive Monte Carlo sampling.
-    $\mathbb{E}\qty[f(s)] \approx \frac{1}{N} \sum_{j=1}^N s_j$ is a naive Monte Carlo estimate of the expectation, which is only valid if the SOW are IID draws from $P_s$.
-    This is often done in practice, however, and referring to other papers as suggested may be helpful for pointing this out.
+    As noted, we write in section 2.3 that
 }
+\begin{quote}
+    If the \glspl{sow} are drawn \gls{iid} from the true distribution, then the expected value of some function  $f$, $\mathbb{E}\qty[f(\vb{s})]$, can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
+    However, this is often not the case; for example, a fixed set of simulations may be available for analysis or low-probability but high-impact regions of the parameter space may have been sampled.
+    In this case, the formula must be adjusted to
+    \begin{equation*}
+        \mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{i=j}^N w_j f(\vb{s}_j),
+    \end{equation*}
+    where the $w_j$ are suitably chosen weights such that their sum is equal to 1.
+\end{quote}
+
+\AR{
+    This does not require assuming that the set of sows, $\vb{s}$, is a discrete uniform RV, only that they are drawn IID from $p(\vb{s})$
+}
+
+\begin{quote}
+    Need to revisit -- the comment about population doesn't make sense to me
+\end{quote}
 
@@
 }
+
+\begin{quote}
+    Let's change it so that instead of saying ``suitably chosen'' weights we just say $\text{s.t.}~ \sum_j w_j = 1$.
+    There's another set of questions.
+\end{quote}
 
@@
 
+% ugly hack to hide month in APACite
+\renewcommand{\APACrefYearMonthDay}[3]{\APACrefYear{#1}}
 \bibliographystyle{apacite}

### 2022-09-29 papers/earths-future/reviewer_response_01.tex
making good progress on intro

@@
 \AR{
-    This is a tricky distinction to draw and we should be more clear.
-    One possibility is to use Model and Probabilistic Scenario.
-    Another is Physical Model and Predictive Model. Maybe something else?
+    Thank you for flagging this as an opportunity to improve clarity.
+    After some discussion we have decided to refer to ``scenarios'' (where it adds clarity, ``probabilistic scenarios'') and ``models'' (or ``physical models'' where it is helpful to draw distinctions from, \eg, system models).
 }
@@
 
+\begin{quote}
+    Copy and paste in the new text where we discuss figure 7A
+\end{quote}
+
 \RC{
@@
 }
+
+\begin{quote}
+    Copy and paste in the new text
+\end{quote}
 
@@
 
+\begin{quote}
+    Copy and paste in the new text
+\end{quote}
+
 \RC{
@@
     Yes, can do.
+    See new figure.
 }

### 2022-09-30 papers/earths-future/submission_02.tex
ready for klaus comments on intro

@@
   \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
+  \item Estimates of performance or robustness under deep uncertainty necessarily involve subjective judgement.
 \end{keypoints}
@@
 \begin{abstract}
-  Projections of future climate risks can vary considerably from one source to another, posing considerable communication and decision-analytical challenges.
+  Projections of nonstationary climate risks can vary considerably from one source to another, posing considerable communication and decision-analytical challenges.
   One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
-  Some common approaches include analyzing a small subset of projections or invoking Laplace's principle of insufficient reason to justify a simple average.
+  Some common approaches include analyzing a small subset of projections or invoking treating all projections as equally likely regardless of how they are sampled.
   These approaches can underestimate risks, hide deep uncertainties, and provide little insight into which assumptions drive decision-relevant outcomes.
@@
 We develop and demonstrate a framework to address these challenges.
-The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more understandable and smaller sample set and (ii) provide insights about which assumptions and modeling choices influence decisions.
+The framework provides a transparent approach to (i) combine a large number of deeply uncertain projections to a more interpretable sample set and (ii) provide insights about which assumptions and modeling choices influence decisions.
 We demonstrate the approach with a relatively simple example question of how high to elevate a house in the face of deeply uncertain projections of future water levels.
 
+\clearpage
 \section{Introduction}\label{sec:introduction}
 
-Many critical infrastructure services around the world are aging \cite<\eg,>{ho_dams:2017}.
-Moreover, changes in regulations, finance, patterns of population and infrastructure use, and the climate system challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020}.
-The need to upgrade and expand infrastructure motivates the question: for which possible futures should infrastructure systems and components be designed?
-The answer to this question depends in part on values: intrinsic trade-offs between safety, performance, cost, and other objectives depend on context and stakeholder preferences \cite{keller_management:2021} and are often regulated by statute or industry guidelines \cite{bruneau_multihazard:2017}.
-For example, hospitals and critical infrastructure are generally designed to a higher standard of risk protection than ordinary buildings \cite{asce_7-10:2013}.
-At the same time, performance depends upon future conditions, and so decisions about which scenario to design for are necessarily subject to implicit or explicit assumptions about the likelihood or possibility of different futures.
+Aging infrastructure and changes in regulations, finance, patterns of population and infrastructure use, and climate challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,asce_infrastructure_climate:2021,ho_dams:2017}.
+To ensure acceptable performance, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
+For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
+Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations \cite{asce_24-05:2006,FEMA_p-55:2011}.
+Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.
 
-\subsection{Current practice}
+Standards-based risk management frameworks have many advantages, including scalability, explainability, and simplicity.
+However, the choice of standard is a complex design and policy choice.
+Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} can provide a quantitative basis for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
+This provides a formal basis for, \eg, protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
+However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021}.
 
-Current practice in engineering, infrastructure design, and local governance relies heavily on standards that specify particular design events or conditions that buildings and infrastructure should safely withstand \cite{asce_7-10:2013,bruneau_multihazard:2017}.
-These are often, though not always, informed by probabilistic analysis of relevant data.
-For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-These trigger specific floodplain regulations, such as a requirement for homes in the 100-year floodplain with federally backed mortgages to be covered by flood insurance \cite{kousky_voucher:2014}.
-Additionally, local building codes (based on guidance such as \citeA{asce_24-05:2006} or \citeA{FEMA_p-55:2011}) may require new construction in flood zones to be elevated freeboard above a nominal \gls{bfe}.
-Probabilistic analysis also informs the design of large-scale infrastructure.
-For example, some levees in the Netherlands are designed for a nominal annual failure probability of $\nicefrac{1}{4000}$ \cite{eijgenraam_flooding:2014}, while a seawall proposed as part of a \$29 billion coastal protection project for  Galveston Bay was designed by setting the nominal annual probability of overtopping to 1\% \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021}.
+Moreover, estimates of the performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
+Current practice emphasizes nominally objective methods that can be applied consistently across locations.
+For example, the \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019}.
+Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
+Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}>{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
+While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
+At least in part because of the challenges associated with developing objective methods to select from diverging projections of future hazard, official guidance continues to rely on the stationarity assumption \cite{bulletin17c:2019,atlas14_texas:2018}.
 
-There are many advantages to standards-based design.
-In particular, these heuristics are scalable and explainable, they reduce the complexity of design analysis, and they are fair in at least a procedural sense.
-However, reliance on these heuristics also faces limitations.
-In particular, this sort of one-size-fits-all guidance may not be an efficient or desirable way to balance trade-offs between metrics of stakeholder values such as sense of place, distributive justice, economic efficiency, and safety \cite{keller_management:2021}.
-This has motivated economically informed approaches, like risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017}, that place ``a strong emphasis upon a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
-These quantitative cost-benefit analyses also rely on probabilistic descriptions of relevant hazard, and can be particularly sensitive to representation of tail probabilities \cite{merz_heavytails:2022,wong_floodrisk:2018,garner_slrise:2018}.
+The limitations of objectivist approaches to projecting risk extend beyond estimating nonstationary climate hazards.
+Human-natural systems are never closed and model results are never unique, and thus validation and verification of models representing these systems is necessarily qualitative and subjective \cite{oreskes_verification:1994}.
+In other words, no model exists that could represent the full truth, and the future is thus deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
+Consequently, a growing literature on \gls{dmdu} emphasizes the value of identifying decisions that are robust, in some sense, to model limitations \cite{moody_robustness:2013,herman:2015,mcphail_robustness:2019,borgomeo_robustness:2018}.
+Within this literature has emerged a debate regarding the value and use of probabilistic information  \cite<see>[and references therein]{taner_probabilistic:2019}.
+On the one hand, scholars have pointed out that predictions are inherently unreliable, and representing deep uncertainties through probability distributions frequently over-estimates predictive skill \cite{groves_scenarios:2007,lempert_robust:2000}.
+On the other, assessments of which decisions are robust depend on subjective choices about how to define robustness and how to sample uncertainties \cite{mcphail_robustness:2019,quinn_exploratory:2020,schneider_scenarios:2002,schneider_dangerous:2001,reis_distribution:2020}.
 
-The approaches currently used to estimate the \glspl{pdf} of climate hazards used in standard-setting and decision-making emphasize nominally objective methods that can be applied consistently across locations.
-For example, \gls{usgs} Bulletin 17C specifies procedures for estimating flood frequency \cite{bulletin17c:2019} and \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018}.
-Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}; see>[for a review]{merz_review:2014}.
-For example, clear trends in extreme rainfall are apparent across much of southeastern Texas \cite{fagnant_spatiotemporal:2020,nielsen-gammon_txrainfall:2020}, and trends in many other hazards for other locations are consistent with observations \cite<see>[for a comprehensive summary]{ipcc_impacts:2022}.
-While some methods have been proposed for incorporating trends into these analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of nonstationarity which may not fully represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,Montanari:2014hl,Serinaldi:2015bq}.
-Ultimately, the difficulty of incorporating nonstationarity into existing frameworks has led to continued reliance on stationarity despite recognition of the associated limitations as discussed in p.~2 of \citeA{bulletin17c:2019} and p.~A.4-42 of \citeA{atlas14_texas:2018}.
+In this paper we offer a first conceptual step towards bridging this divide by presenting a framework that combines the strengths of both approaches.
+In the first step, exploratory or bottom-up modeling is used to build insight and identify potential system vulnerabilities \cite{moallemi_exploratory:2020,bankes:1993,brown_decisionscaling:2012}.
+In the second, we synthesize the results of exploratory modeling to formally estimate performance metrics and trade-offs using probabilistic models for uncertainty.
+Drawing from the literature on building predictive models when ``all models are wrong'' \cite{box_sciencestatistics:1976,gelman_philosophy:2013,Piironen:2017eh}, we interpret these probability distributions not statements of fact, but rather as a self-consistent framework for reasoning about how different assumptions lead to different inferences.
+An advantage of our approach is that it facilitates computationally efficient analysis of how alternative probabilistic models would affect estimated performance metrics and trade-offs.
 
-\subsection{Emerging paradigms}
-
-Projecting nonstationary hazard is difficult because many future hazards depend on human decisions (\eg, future greenhouse gas emissions) or on physical processes that are poorly constrained by existing data \cite<\eg, collapse of the West Antarctic ice sheet;>[]{deconto_antarctica:2016}.
-In other words, they are deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
-These deeply uncertain nonstationary hazards challenge not only the existing stationary estimates but, more fundamentally, the premise that objective estimates of future hazard exist and can be estimated empirically.
-
-Recognizing the challenges of \gls{dmdu}, many frameworks for identifying robust decisions have been proposed.
-Most emphasize the use of models in an exploratory (``what-if'') framework to learn about interactions between decisions and system dynamics \cite{bankes:1993}.
-For example, robust decision making \cite{lempert_shaping:2003} evaluates models over large ensembles of possible futures to assess the performance of different policies under each, then applies statistical analysis to identify the conditions under which particular policies perform well or poorly.
-When the decision space is complex, many analyses use policy search (often using multi-objective optimization tools) to identify promising actions \cite{kasprzyk:2013,kasprzyk_denovo:2012,hadka_mordm:2015}.
-In addition, many studies formally quantify robustness to deep uncertainties \cite{herman:2015,mcphail_robustness:2019} and use this as a criterion for policy comparison.
-
-Although these \gls{dmdu} methods have proven valuable in a wide range of settings, they still require necessarily subjective assessments about the likelihood of future conditions.
-Even exploratory models require the analyst to choose which uncertainties are considered and how they are sampled.
-When analyses use metrics that integrate performance over many possible futures, whether to compute sensitivities, estimate expectations, or compute robustness metrics, they necessarily make assumptions about the likelihood of different futures.
-For example, uniform priors are often justified by deferring to Laplace's principle of insufficient reason \cite<see>[p.~135]{stigler_uncertainty:1986}, but even this choice makes strong assumptions that are sensitive to factors such as parameterization \cite[p.54]{gelman_bda3:2014}.
-While one interpretation of deep uncertainty is that that specifying a joint \gls{pdf} over inputs is inappropriate, assumptions about the ranges and independence of parameters to sample are just as subjective as the choice of probability distribution \cite{schneider_scenarios:2002,quinn_exploratory:2020} and can often be interpreted  as a specific choice of probability distribution (we revisit this point in \cref{sec:analysis-condition}).
-This motivates the development of decision analytic frameworks that draw from the strengths of \gls{dmdu} methods such as exploratory modeling, vulnerability assessment, robustness checks, and iterative stakeholder critique, but that embrace the reality that assumptions about the future are inescapable.
-
-\subsection{Research gaps and objectives}
-
-We are motivated by parallels between decision making under deep uncertainty and the statistical problem of model selection.
-\citeA{oreskes_verification:1994} argues that because natural systems are never closed and model results are never unique, validation and verification of models representing these systems is necessarily qualitative and subjective.
-In the literature on decision making under deep uncertainty, this has led to recognition of the need to develop strategies that are robust to model imperfections.
-There are parallels to this challenge in the Bayesian literature on model selection, particularly in the ``$\mathcal{M}$-open'' case in which there is no ``true'' model to identify and all models are imperfect representations of reality.
-This is a theoretical contrast to the ``$\mathcal{M}$-closed'' case in which the data are assumed to come from a ``true'' model which is a member of the finite set of candidate models available to the analyst \cite{bernardo_bayesian:1994}.
-Commonly used classical methods for model selection implicitly assume $\mathcal{M}$-closed case and are defined by their ability to asymptotically identify the true model from a set of candidate models.
-By contrast, model selection in the $\mathcal{M}$-open case emphasizes a subjective view of the modeling process, for which probability offers a self-consistent language with which to reason about the unknown rather than a statement of objective truth (see \citeA{gelman_philosophy:2013}, \citeA{ramsey_probability:2016}, or \citeA{jaynes_probability:2003} for a discussion of Bayesian philosophy and \citeA{Piironen:2017eh} for a technical discussion of methods for model selection).
-From a practical perspective, model selection in the $\mathcal{M}$-open case emphasizes building models iteratively, simulating the consequences of these models, and subsequently using these simulations to critique and improve them until an acceptable model (in the analyst's subjective judgement) is reached \cite{gelman_workflow:2020}.
-Thus, where classical statistics, particularly as applied to environmental systems, has often emphasized objectivity, we draw from this Bayesian literature that  emphasizes a focus on building transparency in the presentation, analysis, and critique of models and results.
-
-In this paper we offer a first conceptual step towards bridging the fields of \gls{dmdu} and Bayesian model building.
-We consider a didactic case study of whether to elevate a hypothetical house, and if so how high, as a specific example of a decision problem subject to both shallow (storm surge) and deep (\gls{slr}) uncertainties.
+We illustrate our approach through a didactic case study of whether to elevate a hypothetical house, and if so how high.
 Prior studies have found that floodproofing and building-scale vulnerability reduction measures, including house elevation, can effectively reduce local flood damages in many contexts \cite{demoel_reducing:2014,deruig_building:2020,kreibich_building:2005,slotter_floodproofing:2020,rozer_coping:2016,mobley_mitigation:2020,aerts_cost:2018}, and both local building codes \cite{asce_7-10:2013,bruneau_multihazard:2017,asce_24-05:2006} and federal policy \cite{FEMA_p-55:2011} require elevation in some cases.
@@
 Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
+To shed light on this question in a single paper, we necessarily are silent on key issues that would be relevant to real-world decision-makers including alternative decision levers, the potential for adaptive decision rules, and rival problem framings.
 
@@
 In \cref{sec:case-study} we describe the case study in detail.
-Next in \cref{sec:results} we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
+Next, in \cref{sec:results}, we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
 In \cref{sec:caveats} we discuss limitations of the study and future research needs.
@@
 Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
-For each scenario $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
+For each \gls{sow} $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
 While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
@@
 A first analytical step is to use the model in an ``exploratory'' mode.
-Exploratory modeling is averse to making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
+Exploratory modeling strives to avoid making explicit assumptions about the likelihood of different \glspl{sow} and instead seeks to generate new knowledge \cite{bankes:1993} by systematically exploring a large number of possible futures, emphasizing interactions between different uncertainties \cite{reed_msdbook:2022}.
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
@@
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$'' or ``which decision minimizes expected damages'' or ``what is the probability of exceeding a critical threshold'' requires constructing estimators for the probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$''  or ``what is the probability of exceeding a critical threshold'' requires estimating the probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.
 
@@
 \end{equation}
-where the $w_j$ are suitably chosen weights such that their sum is equal to 1.
+where $\sum_{j=1}^J w_j = 1$.
 
@@
 We first sort the $\psi_j$  from least to greatest so that $\psi_{j-1} \leq \psi_j$, ($j \neq 1$).
-Defining $F_\mathrm{belief}(s)$ to be the cumulative distribution corresponding to $p_\mathrm{belief}$, we calculate weights as
+Defining a cumulative distribution function
+$F_\mathrm{belief} \qty(\psi) = \int_{-\infty}^{\psi} p_\mathrm{belief}(\psi') \dd{\psi'}$,
+we calculate weights as
 \begin{equation}\label{eq:weights}
   w_j = \begin{cases}
-    F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_1 + \psi_2])                                                                     & j = 1     \\
-    F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j} + \psi_{j+1}]) - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{j-1} + \psi_j]) & 1 < j < J \\
-    1 - F_\mathrm{belief}\qty(\frac{1}{2}\qty[\psi_{J-1} + \psi_J])                                                             & j = J.
+    F_\mathrm{belief}\qty(\frac{\psi_1 + \psi_2}{2})                                                              & j = 1     \\
+    F_\mathrm{belief}\qty(\frac{\psi_{j} + \psi_{j+1}}{2}) - F_\mathrm{belief}\qty(\frac{\psi_{j-1} + \psi_j}{2}) & 1 < j < J \\
+    1 - F_\mathrm{belief}\qty(\frac{\psi_{J-1} + \psi_J}{2})                                                      & j = J.
   \end{cases}
@@
 
+% TODO: discuss somewhere nearby that creating this low-dimensional projection may be difficult for more complex problems. Refer to R1.
+
 \begin{figure}
@@
   \caption{
-    Schematic of \gls{sow} weighting scheme defined in \cref{eq:weights}.
+    Schematic of \gls{sow} re-weighting scheme defined in \cref{eq:weights}.
     This method is illustrated for a hypothetical target distribution (dark red line) and $J=4$ samples $\psi_1, \psi_2, \psi_3, \psi_4$ (blue dots).
@@
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
+% TODO: R1 -- It may be the case that some objectives are not affected by the differences in pbelief between stakeholders and so there is no point in getting tied up trying to pin it down.
 
@@
 \end{table}
+% TODO: R1 -- Table 1: I think at least some detail should be given about the elevation cost and depth-damage functions.
+I understand they were taken from another work.
+This would make interpreting the later results easier.
 
@@
 \end{figure}
+% TODO -- R1: Figure 4b. There is remarkable disagreement between the models of mean sea level rise. It might be worth at least a few sentences commenting on this.
 
@@
 
-Following prior work \cite<\eg,>[]{garner_slrise:2018,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting hydrodynamic effects.
+Following prior work \cite<\eg,>[]{garner_slrise:2018,sriver_sealevel:2018}, we model annual maximum floods $y(t)$ as the sum of sea level $\overline{y}(t)$, described in the previous subsection, and annual maximum storm surges $y'(t)$, neglecting any potential hydrodynamic interactions.
 
@@
 Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
+% TODO -- R1: Equation 4: it might be helpful to define p here.
 
@@
   \end{tabular}
+\end{table}
 
-\end{table}
+% TODO -- R1: Line 416: The use of the gamma distribution is quite sensible but is presented without justification. This might be worth just a short sentence on why it was chosen.
 
@@
 
+% TODO: Copy and paste in the new text where we discuss figure 7A. Explain why there is a spike at 0ft and link to cost curves. Note in reponse to R1.
+
 This probabilistic interpretation allows us to compute, for example, expected values.
@@
 This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
-This approach is attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
+Scenario-conditional analysis can be attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
 However, conditioning simulations on a set of climate futures and physical models presents what we term ``the multiple \gls{pdf} problem'' because it leaves decision makers with many \glspl{pdf} to choose from and hence many trade-off curves to navigate.
-The multiple \gls{pdf} problem has also been shown in other contexts.
+The multiple \gls{pdf} problem has also been discussed in other contexts.
 For example, \citeA{sharma_rcp:2021} model the reliability of stormwater infrastructure under different climate models and downscaling methods, finding diverging estimates of future rainfall hazard, even under a single \gls{rcp} scenario.
@@
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
+% TODO: add another sentence to wrap up explain the implications of the approach for the case study, for example someone elevating a house or issuing guidance / funds for house elevation
 
@@
 
-The proposed approach can help overcome the limitations of scenario-conditional analysis.
-We illustrate how the re-weighting method described in \cref{sec:analysis-synthesize} can help to shed light on climate risk management under deep uncertainty.
-We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}.
-These three distributions are shown in \cref{fig:lsl-priors-weights}(A).
+The \gls{sow} re-weighting framework described in \cref{sec:analysis-synthesize} can help overcome the limitations of scenario-conditional analysis.
+In this section we illustrate how this approach can help to shed light on climate risk management under deep uncertainty.
+We present results using each of the models for $p_\mathrm{belief}$ outlined in \cref{sec:case-priors}; these three distributions are shown in \cref{fig:lsl-priors-weights}(A).
 
@@
 
-One application of this method is to diagnose which assumptions different $p_\mathrm{belief}$ are consistent with.
+% TODO -- R1: Figure 8: The y-axis labels on B-D should be more descriptive.
+% Suggest to put the curent y label on the title and add y label ''average weight'' or similar
+
+One application of this method is to diagnose the assumptions which which different $p_\mathrm{belief}$ are consistent.
 \Cref{fig:lsl-priors-weights}(B-D) shows the total weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
@@
 Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
+% TODO: add a summary sentence about how this analysis might guide decision-making
 
@@
 
-\section{Discussion and research needs}\label{sec:caveats}
+\section{Limitations and research needs}\label{sec:caveats}
 
@@
 Yet while some progress has been made developing probabilistic models for specific elements of this model chain \cite<\eg,>[]{srikrishnan_probabilistic:2022,wong_brick0.2:2017}, this remains a computational and conceptual challenge.
-Finally, while our aim in this paper has been to demonstrate the value of integrating Bayesian workflow \cite{gelman_workflow:2020} into \gls{dmdu}, further work is needed to improve this integration.
 
@@
 One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \cite{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \cite[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
-Finally, while here we consider the decision to be a one-time decision, one could also frame this as problem a sequential decision.
+Finally, while here we consider the decision to be a one-time decision, one could also frame this as a sequential decision problem.
 The analysis of sequential decision problems applies tools from control theory and reinforcement learning to identify policies that map ``triggers'' (\ie, state variables) to decisions \cite{herman_control:2020}.
-Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \cite{fletcher:2017,garner_slrise:2018}, the optimized policy rules are necessarily sensitive to the characterization of uncertainty, and thus the problem of synthesizing across deep uncertainties remains \cite{herman_control:2020}.
+Yet although framing the decision through a sequential lens can increase adaptability and improve outcomes \cite{fletcher:2017,garner_slrise:2018}, decisions and outcomes remain highly sensitive to the characterization of uncertainty \cite{herman_control:2020}, and thus the problem of synthesizing across deep uncertainties remains relevant.
 
@@
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
-Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that are applicable across different house characteristics may improve usability.
+Additionally, since developing bespoke analyses for each house may be impractical, identifying decision rules that are applicable across different house characteristics may improve usability and guidance.
+Finally, there are many parallels between \gls{dmdu} and subjective Bayesian literature on building predictive models in the ``$\mathcal{M}$-closed'' case when ``all models are wrong'' \cite{gelman_philosophy:2013,box_sciencestatistics:1976}, and thus future work can demonstrate how to incorporate techniques from Bayesian workflow \cite<see>{gelman_workflow:2020} into \gls{dmdu} methodologies.
 
@@
 We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
+% TODO: cite ViMM Manifesto as an example of how exploratory modeling can be stakeholder-driven
+See \cite{helgeson_manifesto:2022}. % ViMM Manifesto
 Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-We use a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
+We use a didactic case study of house xelevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
 
-The proposed scenario weighting framework can be applied to inform critical challenges in climate risk management.
+xThe proposed \gls{sow} re-weighting framework can be applied to inform critical challenges in climate risk management.
 An obvious area of application is to the design of infrastructure.
@@
 Yet upgrading this infrastructure is costly and subject to large uncertainties between rainfall models \cite{sharma_rcp:2021} and \gls{rcp} scenarios.
-Similarly, decisions like  levee heightening \cite{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design (as discussed in the Introduction) are subject to deep uncertainties in sea level rise.
-Investments in water resources planning and management depend on assumptions of future water demand, availability, and technologies \cite{trindade_deeplyuncertainpathways:2019}.
+Similarly, decisions like  levee heightening \cite{garner_slrise:2018,oddo_coastal:2017,vandantzig_dike:1956} and sea wall design \cite[Appendix D.,~p.~2-59]{USACE_coastal:2021} are subject to deep uncertainties including sea level rise.
+Investments in water resources planning and management also depend on assumptions of future water demand, availability, and technologies \cite{trindade_deeplyuncertainpathways:2019}.
 And analyses of climate change mitigation options, such as estimates of the social cost of pollutants \cite{errickson_methane:2021} or cost-minimizing energy transition pathways, are conditional on probabilistic models for inputs like technology prices and population.
@@
 Sitara Baboolal,
+Courtney Cooper,
 Tor Erlend Fjelde,
@@
 Adam Pollack,
-and
-Vivek Srikrishnan
-for helpful comments.
+Vivek Srikrishnan,
+and two anonymous peer reviewers
+for helpful comments that improved this research.
 

### 2022-10-03 papers/earths-future/submission_02.tex
incorporate R1 changes and work on R2

@@
 Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
+Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic (and non-Uniform) priors over functions of parameters and (ii) results can change when the parameterization of a given process is altered \cite{gupta_prior:2022,seaman_noninformative:2012}.
 
@@
 
-% TODO: discuss somewhere nearby that creating this low-dimensional projection may be difficult for more complex problems. Refer to R1.
-
 \begin{figure}
@@
 However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
-% TODO: R1 -- It may be the case that some objectives are not affected by the differences in pbelief between stakeholders and so there is no point in getting tied up trying to pin it down.
+
+Considering multiple probabilistic models for $p_\mathrm{belief}$ can also be useful for understanding the sensitivity of the decision to the choice of $p_\mathrm{belief}$.
+Further, the sensitivity, or lack thereof, of different objectives to the choice of $p_\mathrm{belief}$ may be useful for identifying future research needs.
 
@@
 \end{table}
-% TODO: R1 -- Table 1: I think at least some detail should be given about the elevation cost and depth-damage functions.
-I understand they were taken from another work.
-This would make interpreting the later results easier.
 
@@
 \end{figure}
-% TODO -- R1: Figure 4b. There is remarkable disagreement between the models of mean sea level rise. It might be worth at least a few sentences commenting on this.
 
@@
 \Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
+The differences between models arise primarily due to different representations of Antarctic Ice Sheet contributions to global \gls{slr} and statistical calibration methodologies; for a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
 We return in \cref{sec:results-conditional} to the challenge of decision making under multiple \glspl{pdf}.
@@
 \end{equation}
-where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house).
+where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house) and $p(y')$ is the probability density of storm surge.
 Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
 For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
+Both  models show damage increasing with flood depth before reaching an upper limit.
 Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
-% TODO -- R1: Equation 4: it might be helpful to define p here.
 
@@
 Following \citeA{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \cite{fischbach_clara:2012}.
-We normalize this cost by house value; this cost curve is shown in \cref{fig:cost-up-front}.
+We normalize this cost by house value; this cost curve is shown in \cref{fig:cost-up-front} and shows a large up-front cost plus a piecewise linear marginal cost.
   \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
@@
 We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
+The distributions were chosen to be illustrative, rather than to reflect any particular scientific consensus.
+The Gamma distribution is a flexible distribution that can be used to model skewed, lower-bounded distributions, making it an appropriate choice for modeling subjective uncertainty about \gls{slr}.
 \Cref{tab:slr-priors} specifies the parameters of these distributions, as well as some quantiles of the distributions.
@@
 \end{table}
-
-% TODO -- R1: Line 416: The use of the gamma distribution is quite sensible but is presented without justification. This might be worth just a short sentence on why it was chosen.
 
@@
 
-% TODO: Copy and paste in the new text where we discuss figure 7A. Explain why there is a spike at 0ft and link to cost curves. Note in reponse to R1.
-
 This probabilistic interpretation allows us to compute, for example, expected values.
 For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen models considered (we highlight three representative models).
-Similarly, \cref{fig:tradeoffs-by-rcp}(b) plots the lifetime expected damages as a function of $\Delta h$.
-For small $\Delta h$, expected costs are low under optimistic models (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics; red lines) and high under pessimistic models (\eg, \gls{rcp}8.5 with the DP16 model; blue lines).
-For example, under the most pessimistic model (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
-Under the most optimistic model (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost yet reduces lifetime expected damages by less than 50\% of house value.
+This panel shows lifetime expected damages as a function of $\Delta h$, shown in \cref{fig:tradeoffs-by-rcp}(b,  plus the up-front cost of construction.
+Because there are high fixed costs associated with building (see cost curve in \cref{fig:cost-up-front}), it generally does not make sense to raise the house by only a small amount, since this incurs these fixed costs without providing substantial damage reduction.
+\Cref{fig:tradeoffs-by-rcp} shows that estimates of trade-offs between up-front cost and expected lifetime costs are highly sensitive to the chosen scenario.
+For small $\Delta h$, expected costs are low under optimistic scenarios (\eg, \gls{rcp} 2.6 with slow ice sheet dynamics; red lines) and high under pessimistic scenarios (\eg, \gls{rcp}8.5 with the DP16 model; blue lines).
+Estimates of the optimal decision are highly sensitive to the choice of scenario.
+For example, under the most pessimistic scenario (blue line), the cost-minimizing height increase is \SI{6}{ft}, which incurs an up-front cost of 73\% of the house value but reduces lifetime expected damages by over 150\% of house value.
+Under the most optimistic scenario (gray line), the cost-minimizing decision is to not elevate, as elevating \SI{6}{ft} incurs the same up-front cost yet reduces lifetime expected damages by less than 50\% of house value.
 
@@
 Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
-% TODO: add another sentence to wrap up explain the implications of the approach for the case study, for example someone elevating a house or issuing guidance / funds for house elevation
+Since not all house owners or contractors have expertise in assessing the relative likelihood of different climate futures, they may not be well positioned to make this decision.
 
@@
 
-% TODO -- R1: Figure 8: The y-axis labels on B-D should be more descriptive.
-% Suggest to put the curent y label on the title and add y label ''average weight'' or similar
-
 One application of this method is to diagnose the assumptions which which different $p_\mathrm{belief}$ are consistent.
@@
 Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
-% TODO: add a summary sentence about how this analysis might guide decision-making
+Based on this analysis, we would recommend that the owner if this hypothetical home elevate by approximately 4-6\si{ft}.
+Alternatively, the homeowner could choose to defer the deicision of whether, and how high, to elevate; our analysis did not consider this possibility but there is a rich literature on flexible design and engineering options analysis in climate risk analysis \cite<\eg,>{fletcher_learning:2019,fletcher_groundwater:2019,hui_adaptive:2018,garner_slrise:2018,herman_control:2020,deneufville_eoa_theory:2019}.
 
@@
 For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic; prior studies have shown that using Approximate Bayesian Computation to calibrate models on low dimensional statistics using that are not sufficient statistics can lead to biased posterior estimates \cite{csillery_abc:2010,marjoram_abc:2006}.
-Although we are not performing calibration here, and this is thus not a direct concern, time series with the same \gls{msl} in 2100 may differ in other ways, and experts may have prior information about the likelihood of these differences not represented in our model.
+Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic, meaning that many possible low-dimensional representations are possible; time series with the same \gls{msl} in 2100 may differ in other ways.
+For problems with more sources of uncertainty, such as multisector problems, choosing a low-dimensional representation may prove more challenging.
+In such settings, more detailed diagnostics and sensitivity analyses may shed light on the appropriateness of different modeling choices.
 A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
@@
 This study develops a framework designed to increase the transparency of quantitative decision analysis under deep uncertainty.
-We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling with subjective probabilistic expert assessment.
-% TODO: cite ViMM Manifesto as an example of how exploratory modeling can be stakeholder-driven
-See \cite{helgeson_manifesto:2022}. % ViMM Manifesto
+We develop a framework capable of blending iterative, stakeholder-driven exploratory modeling \cite<see, \eg,>{helgeson_manifesto:2022} with subjective probabilistic expert assessment.
 Such an approach is urgently needed given that deeply uncertain nonstationarity hazards pose a fundamental challenge to classical methods of hazard estimation.
-We use a didactic case study of house xelevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
+We use a didactic case study of house elevation in the coastal zone to illustrate a method for transparently synthesizing across deep uncertainties.
 

### 2022-10-03 scripts/06-combining.jl

@@ -77,7 +77,7 @@ end
 function plot_priors()
     priors = get_priors()
     p = plot(;
-        xlabel="SLR 2022-2100, at Sewells Point, VA [ft]", ylabel="Probability Density"
+        xlabel="SLR 2022-2100, at Sewells Point, VA [ft]", ylabel="Probability Density", title="(a): Implicit Prior"
     )
     for (prior, color) in zip(priors, [:red, :blue, :green])
         plot!(
@@ -92,7 +92,6 @@ function plot_priors()
         )
     end
     p
-    #savefig(p, plots_dir("lsl-priors.pdf"))
     return p
 end
 
@@ -107,6 +106,8 @@ function plot_weight(s::Vector{<:HouseElevation.LSLSim})
         df[!, prior.name] = w
         w_avg = combine(groupby(df, [:model, :rcp]), prior.name => sum => :weight)
         @assert sum(w_avg[!, :weight]) ≈ 1
+
+        letter = collect('a':'z')[i]
         p = @df w_avg groupedbar(
             :model,
             :weight,
@@ -114,7 +115,8 @@ function plot_weight(s::Vector{<:HouseElevation.LSLSim})
             palette=colors,
             legendtitle="  RCP",
             legend=i == 1 ? :topright : false,
-            ylabel="$(prior.name)",
+            title="($letter): $(prior.name)",
+            ylabel="Implicit Weight Assigned",
             xrotation=30,
         )
         return p
@@ -126,7 +128,6 @@ end
 
 function plot_priors_weights(s::Vector{<:HouseElevation.LSLSim})
     plots = vcat(plot_priors(), plot_weight(s)...)
-    add_panel_letters!(plots; loc=(0.15, 0.99), fontsize=11)
     p = plot(
         plots...;
         layout=(2, 2),

### 2022-10-04 papers/earths-future/submission_02.tex
send to klaus,  have everything but model/scenario language

@@
   \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
-  \item We demonstrate the framework using a simple case study of house elevation for coastal flood risk management.
-  \item Estimates of performance or robustness under deep uncertainty necessarily involve subjective judgement.
+  \item We demonstrate the framework using a simple case study of how high to elevate a  house to manage coastal flood risks.
+  \item Estimates of performance or robustness under deep uncertainty necessarily involve subjective judgments.
 \end{keypoints}
@@
   One such challenge is how to present trade-offs under deep uncertainty in a salient and interpretable manner.
-  Some common approaches include analyzing a small subset of projections or invoking treating all projections as equally likely regardless of how they are sampled.
-  These approaches can underestimate risks, hide deep uncertainties, and provide little insight into which assumptions drive decision-relevant outcomes.
+  Some common approaches include analyzing a small subset of projections or treating all considered projections as equally likely.
+  These approaches can underestimate risks, hide deep uncertainties, and are mostly silent on which assumptions drive decision-relevant outcomes.
   Here we introduce and demonstrate a transparent Bayesian framework for synthesizing deep uncertainties to inform climate risk management.
@@
 
-The identification of sound strategies to manage risks driven by climatic changes is challenged by large uncertainties surrounding projections of the coupled natural-human systems that influence projections of climate risk.
+Identifying sound strategies to manage risks driven by climatic changes is a complex task given the large uncertainties surrounding projections of  coupled natural-human systems.
 These uncertainties often arise from choices experts have to make, for example about how to formulate scientific models of future water levels.
@@
 Analyzing decisions in such a situation of deep uncertainty poses nontrivial challenges.
-For example, picking a single representative projection can lead to under-estimation of risk and poor decisions.
+For example, picking a single representative projection can under-estimate risk and result in poor decisions.
 Similarly, communicating results separately for each projection can overwhelm decision-makers.
@@
 Aging infrastructure and changes in regulations, finance, patterns of population and infrastructure use, and climate challenge the ability of critical infrastructures to meet design objectives \cite{doss-gollin_txtreme:2021,doss-gollin_fatalism:2020,chester_reliable:2020,asce_infrastructure_climate:2021,ho_dams:2017}.
-To ensure acceptable performance, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
+To achieve acceptable performance with reasonable planning efforts, current practice in engineering, infrastructure design, and regulation relies heavily on standards that specify design events or conditions that buildings and infrastructure should safely withstand \cite{bruneau_multihazard:2017}.
 For example, \gls{fema}, local governments, and engineering consultants produce local floodplain maps in many communities.
-Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations \cite{asce_24-05:2006,FEMA_p-55:2011}.
+Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations for new construction \cite{asce_24-05:2006,FEMA_p-55:2011}.
+For example, many communities require that new construction be elevated at least \SI{1}{ft} above the nominal 100-year flood level \cite{zarekarizi_suboptimal:2020}.
 Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.
@@
 Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} can provide a quantitative basis for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
-This provides a formal basis for, \eg, protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
-However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021}.
+This provides a formal basis for choices such as protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
+However, these methods are silent on how standards should balance trade-offs, not only between cost and performance but also between other stakeholder values such as sense of place, distributive justice, and safety \cite{keller_management:2021,helgeson_manifesto:2022,quinn_rivalframings:2017,bessette_vimm:2017,vezer_epistemic:2018}.
 
@@
 Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
-Among several statistical assumptions these analyses make that have been recently called into question is stationarity \cite<the assumption that past and future hazard come from the same \gls{pdf}>{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
-While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or true levels of uncertainty \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
+One statistical assumptions these analyses make is stationarity (the assumption that past and future hazard come from the same \gls{pdf}), but global climate change and local environmental changes have cast scrutiny on this assumption \cite{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
+While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or sample only a subset of uncertainties \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
 At least in part because of the challenges associated with developing objective methods to select from diverging projections of future hazard, official guidance continues to rely on the stationarity assumption \cite{bulletin17c:2019,atlas14_texas:2018}.
@@
 In other words, no model exists that could represent the full truth, and the future is thus deeply uncertain \cite{keller_management:2021,walker_deep:2013,lempert_complex:2002,haasnoot_sealevelrise:2021}.
-Consequently, a growing literature on \gls{dmdu} emphasizes the value of identifying decisions that are robust, in some sense, to model limitations \cite{moody_robustness:2013,herman:2015,mcphail_robustness:2019,borgomeo_robustness:2018}.
+Consequently, a growing literature on \gls{dmdu} emphasizes the value of identifying decisions that are robust, in some sense, to deep uncertainties \cite{moody_robustness:2013,herman:2015,mcphail_robustness:2019,borgomeo_robustness:2018}.
 Within this literature has emerged a debate regarding the value and use of probabilistic information  \cite<see>[and references therein]{taner_probabilistic:2019}.
@@
 
-In this paper we offer a first conceptual step towards bridging this divide by presenting a framework that combines the strengths of both approaches.
+In this paper we offer a conceptual step towards bridging this divide by presenting a framework that is designed to combine the strengths of both approaches.
 In the first step, exploratory or bottom-up modeling is used to build insight and identify potential system vulnerabilities \cite{moallemi_exploratory:2020,bankes:1993,brown_decisionscaling:2012}.
-In the second, we synthesize the results of exploratory modeling to formally estimate performance metrics and trade-offs using probabilistic models for uncertainty.
-Drawing from the literature on building predictive models when ``all models are wrong'' \cite{box_sciencestatistics:1976,gelman_philosophy:2013,Piironen:2017eh}, we interpret these probability distributions not statements of fact, but rather as a self-consistent framework for reasoning about how different assumptions lead to different inferences.
+In the second step, we synthesize the results of exploratory modeling to formally estimate performance metrics and trade-offs using probabilistic models for uncertainty.
+Drawing from the literature on building predictive models when all models are wrong \cite{box_sciencestatistics:1976,gelman_philosophy:2013,Piironen:2017eh}, we interpret these probability distributions not as statements of fact, but rather as a self-consistent framework for reasoning about how different assumptions lead to different inferences.
 An advantage of our approach is that it facilitates computationally efficient analysis of how alternative probabilistic models would affect estimated performance metrics and trade-offs.
@@
 Guidance for homeowners, notably from \gls{fema}, recommends elevating to the \gls{bfe} (typically the \SI{100}{year} flood) plus a freeboard \cite{fema_retrofitting:2014,asce_24-14:2015,fema_retrofitting:2014} but recent research has demonstrated that neglecting uncertainty in the cost-benefit analysis can lead to poor decisions \cite{zarekarizi_suboptimal:2020}.
-Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can deep uncertainties be transparently synthesized for decision analysis?}''
+Focusing on deep uncertainty in \gls{slr} over the 70 year design life of a hypothetical house, we seek to answer the research question ``\emph{how can decision analysis transparently synthesize deep uncertainties?}''
 To shed light on this question in a single paper, we necessarily are silent on key issues that would be relevant to real-world decision-makers including alternative decision levers, the potential for adaptive decision rules, and rival problem framings.
@@
 As discussed in the previous subsection, estimating the expectation of arbitrary functions under uncertainty is a critical component of decision making under uncertainty.
-If the \glspl{sow} are drawn \gls{iid} from the true distribution, then the expected value of some function  $f$, $\mathbb{E}\qty[f(\vb{s})]$, can be readily estimated as $\frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
-However, this is often not the case; for example, a fixed set of simulations may be available for analysis or low-probability but high-impact regions of the parameter space may have been sampled.
-In this case, the formula must be adjusted to
+If the \glspl{sow} are drawn \gls{iid} from the target distribution, then the expected value of some function $f(\mathbf{s})$, can be readily estimated following the Monte Carlo estimate $\mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
+However, this is often not the case; for example, a fixed set of simulations not drawn from the target distribution may be available for analysis.
+In this case, the formula may be adjusted to a weighted Monte Carlo estimate:
 \begin{equation}
-  \mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{i=j}^N w_j f(\vb{s}_j),
+  \mathbb{E}\qty[f(\vb{s})] \approx \sum_{i=j}^N w_j f(\vb{s}_j),
 \end{equation}
@@
 Following the approach outlined in \citeA{zarekarizi_suboptimal:2020}, we focus on a case study of a \emph{hypothetical} house in Norfolk, VA.
-For interpretability, we focus on deep uncertainty in \gls{msl} and approximate other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
+For interpretability, we focus on deep uncertainty in \gls{msl} and treat storm surge and other model parameters as shallow uncertainties as shown in \cref{tab:uncertainties}.
 We use the notation developed in the previous section to describe the case study.
@@
 \Cref{fig:lsl-priors-weights}(B-D) shows the total weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
-For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
+Specifically, weights are computed as the sum of weights assigned to each \gls{sow} sampled from that model.
+For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on \glspl{sow} produced by the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
 Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 \cite<also unlikely given current policy;>{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}) and \gls{rcp} 4.5.
 The uncertain \gls{slr} scenario (blue line) allocates approximately equal weight across models.
+Decision analysts can use this approach as a diagnostic to understand the assumptions implicit to their choice of $p_\mathrm{belief}$.
 

### 2022-10-05 papers/earths-future/submission_02.tex
changes made; need to update letter to reviewers

@@
 \newcommand{\usd}[1]{\SI{#1}[\$]{}}
-
-% cite things in supplemental info file
-% see https://www.overleaf.com/learn/how-to/Cross_referencing_with_the_xr_package_in_Overleaf
-% it's a big strange for Overleaf!
-\usepackage{xr-hyper}
-\makeatletter
-\newcommand*{\addFileDependency}[1]{% argument=file name and extension
-  \typeout{(#1)}
-  \@addtofilelist{#1}
-  \IfFileExists{#1}{}{\typeout{No file #1.}}
-}
-\makeatother
-\newcommand*{\myexternaldocument}[1]{%
-    \externaldocument{#1}%
-    \addFileDependency{#1.tex}%
-    \addFileDependency{#1.aux}%
-}
-\myexternaldocument{supplemental}
 
@@
 Buildings in the designated floodplain are subject to specific regulations, such as flood insurance requirements as an eligibility requirement for federally backed mortgages \cite{kousky_voucher:2014} or minimum elevations for new construction \cite{asce_24-05:2006,FEMA_p-55:2011}.
-For example, many communities require that new construction be elevated at least \SI{1}{ft} above the nominal 100-year flood level \cite{zarekarizi_suboptimal:2020}.
 Although this paper focuses on flooding, similar approaches inform mitigation strategies for a wide range of other hazards \cite{asce_7-10:2013}.
@@
 However, the choice of standard is a complex design and policy choice.
-Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} can provide a quantitative basis for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
+Risk-based design and cost benefit analysis \cite{eijgenraam_flooding:2014,vandantzig_dike:1956,xian_elevation:2017} offer a quantitative framework for comparing possible standards by emphasizing ``a proportionate response to risk, so that the amount invested in risk reduction is in proportion to the magnitude of the risk and the cost-effectiveness with which that risk may be reduced'' \cite{merz_fluvial:2010}.
 This provides a formal basis for choices such as protecting hospitals and critical infrastructure to a higher degree than ordinary buildings \cite{asce_7-10:2013}.
@@
 
-Moreover, estimates of the performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
+Moreover, estimates of performance trade-offs require implicit or explicit assumptions about the likelihood of different possible futures.
 Current practice emphasizes nominally objective methods that can be applied consistently across locations.
@@
 Similarly, \gls{noaa} Atlas 14 provides estimates of the intensity, duration, and frequency of extreme rainfall \cite{atlas14_texas:2018,noaa_nonstationary:2022}.
-One statistical assumptions these analyses make is stationarity (the assumption that past and future hazard come from the same \gls{pdf}), but global climate change and local environmental changes have cast scrutiny on this assumption \cite{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
+One statistical assumption these analyses make is stationarity (the assumption that past and future hazard come from the same \gls{pdf}), but global climate change and local environmental changes have cast scrutiny on this assumption \cite{merz_review:2014,milly_stationarity:2008,doss-gollin_robustadaptation:2019}.
 While some methods have been proposed for incorporating nonstationarity into risk analyses \cite<see>[for a review]{Salas:2018ge}, these assume specific forms of a trend which may not adequately represent physical processes or sample only a subset of uncertainties \cite{doss-gollin_robustadaptation:2019,montanari_immortal:2014,serinaldi_undead:2015}.
@@
 We proceed as follows.
-In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling (\cref{sec:analysis-explore}) and scenario analysis (\cref{sec:analysis-condition}) towards a formal Bayesian method for transparently synthesizing deep uncertainty through subjective prior beliefs (\cref{sec:analysis-synthesize}).
-In \cref{sec:case-study} we describe the case study in detail.
+In \cref{sec:analysis} we present three formal decision analytic frameworks for analyzing an ensemble of \gls{slr} simulations, building through existing approaches for exploratory modeling scenario analysis to identify a need for synthesizing across scenarios.
+We present a formal framework for transparently synthesizing deep uncertainties in \cref{sec:reweighting}.
+In \cref{sec:case-study} we describe the didactic case study.
 Next, in \cref{sec:results}, we present results for each of the three decision lenses and discuss the advantages and limitations of each theoretical approach.
@@
 
-\section{Decision analytic framework}\label{sec:analysis}
+\section{Conceptual framework}\label{sec:analysis}
 
-In this section we outline our framework for decision analysis under deep uncertainty, maintaining a high level of generality.
-In the next we discuss application to the house elevation case study.
+In this section we introduce a conceptual framework and notation for decision analysis under deep uncertainty.
+We then apply this framework to understand exploratory modeling (\cref{sec:analysis-explore}) and scenario-conditional probabilistic analysis (\cref{sec:analysis-condition}).
+In \cref{sec:analysis-synthesize} we discuss why synthesizing across multiple scenarios is necessary.
+In \cref{sec:reweighting} we provide a formal method for synthesizing across scenarios.
 
+It is common in climate risk management to work with an ensemble of trajectories of climate, economic, and other key variables, which we refer to as \glspl{sow}.
+Often these \glspl{sow} are computationally expensive to generate, and thus it is important to use them efficiently.
 Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
@@
     Outline of the proposed decision-analytic framework.
-    In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance (c) of candidate decisions (a) under a large ensemble of possible futures (b).
-    In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes (e) that are conditional upon specific models describing the likelihood of different futures (d).
-    In our case study, these are \gls{rcp} scenarios and physical models of sea level rise.
-    Finally in \cref{sec:analysis-synthesize} we propose a subjective Bayesian framework for synthesizing across deep uncertainties by re-weighting sampled futures (f).
+    In \cref{sec:analysis-explore} we use an exploratory framework to quantify the performance of candidate decisions under a large ensemble of possible futures.
+    In \cref{sec:analysis-condition} we illustrate the ``multiple PDF problem'' by creating probability distributions over outcomes that are conditional upon specific probabilistic scenarios.
+    In our case study, these scenarios correspond to combinations of emissions pathways with physical models for sea level rise.
+    Then in \cref{sec:analysis-synthesize} we describe the need for synthesizing insight across scenarios.
+    Finally in \cref{sec:reweighting} we provide a formal framework for doing so.
   }\label{fig:flowchart}
@@
 
-\subsection{Explore}\label{sec:analysis-explore}
+\subsection{Exploratory modeling}\label{sec:analysis-explore}
 
@@
 Exploratory modeling is often paired with analyses that identify relevant scenarios  \cite{lamontagne_discovery:2018,groves_scenarios:2007} or summarize a system's response to forcing \cite{Poff:2015jn,Steinschneider:2015kk,sriver_sealevel:2018}.
-Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020}.
+Despite the aversion to strong assumptions about the likelihood of different futures, subjective modeling decisions such as the choice of system model, the set of candidate decisions, the criteria used to assess outcomes, and the choice of how to sample \glspl{sow} can strongly influence results \cite{quinn_exploratory:2020,quinn_rivalframings:2017,moallemi_decisionsupport:2020}.
 
-\subsection{Condition}\label{sec:analysis-condition}
+\subsection{Scenario-conditional probabilistic analysis}\label{sec:analysis-condition}
 
 Although exploratory modeling is a useful framework for understanding systems, there are many questions that it cannot answer.
-For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$''  or ``what is the probability of exceeding a critical threshold'' requires estimating the probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.
+For example, answering questions like ``what is the 95th percentile of metric $u$ under decision $x$''  or ``what is the probability of exceeding a critical threshold'' requires an implicit or explicit probability distribution over outcomes \cite<see>[for a general discussion]{schneider_scenarios:2002}.
 
-One way to interpret an ensemble of \glspl{sow} is as \gls{iid} draws from some true data generating process.
-This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a fully probabilistic model; we call this scenario-conditional probabilistic analysis.
-We add this concept to our framework with boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario (\ie, the assumed input) $M_k$.
-If the \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$, allowing estimation of descriptive metrics.
-This approach allows for quantification of uncertainty \emph{within} models of deeply uncertain processes, but can only qualitatively characterize uncertainty \emph{between} models \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
+One way to interpret an ensemble of \glspl{sow} is as {iid} draws from some probabilistic data generating process.
+This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
+We illustrate this concept in boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario $M_k$.
+By treating the \glspl{sow} as \gls{iid} draws from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
+This ``scenario-conditional'' probabilistic interpretation of \glspl{sow} allows for fully probabilistic quantification of uncertainty and optimization, conditional on a particular scenario.
+For example, \citeA{fletcher_learning:2019} uses stochastic dynamic programming to quantify the value of flexibility in water resources planning.
+However, only a single \gls{rcp} scenario (\gls{rcp} 8.5) is used.
+While the analysis could be repeated for other \gls{rcp} scenarios, the scenario-conditional analysis framework only qualitatively characterize uncertainty \emph{between} scenarios \cite{wong_nola:2017,ruckert_coastal:2019,sharma_rcp:2021}.
 
-We can also use this approach to understand \gls{dmdu} methods that sample a set of parameters from fixed ranges.
+We can also characterize \gls{dmdu} methods that sample a set of parameters from fixed ranges as scenario-conditional analyses.
 For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
-Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range \cite<\ie, robustness metrics; see>[for details]{mcphail_robustness:2019,herman:2015}.
-Since sampling over a range is equivalent to sampling from a Uniform distribution, this assumption is equivalent to assuming the true inputs to come from independent Uniform distributions (one for each parameter) bounded by the plausible range.
-Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic (and non-Uniform) priors over functions of parameters and (ii) results can change when the parameterization of a given process is altered \cite{gupta_prior:2022,seaman_noninformative:2012}.
+Similarly, \citeA{trindade_waterpathways:2020} checks the performance of candidate decisions against an ensemble of synthetic time series of streamflow, water demand, and other parameters by sampling parameters that transform the available data over a plausible range.
+Analyses that use this methodology are implicitly assuming a single probabilistic model in which different variables are drawn from independent Uniform distributions.
+Many limitations of Uniform and other noninformative priors have been documented in the literature, including (i) that they can induce unrealistic implicit priors over functions of parameters and (ii) results are sensitive to the parameterization of a given process \cite{seaman_noninformative:2012}.
+Yet while replacing Uniform distributions with alternatives such as maximum-entropy distributions can address some of these challenges \cite<\eg,>{gupta_prior:2022}, subjective modeling choices remain necessary.
+Our primary concern here is not that these subjective modeling assumptions are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation \cite<see>[regarding the importance of iterative critique]{gelman_workflow:2020} they may lead to inscrutable decision processes and poor decisions.
 
-Our primary concern is not that subjective assumptions about the likelihood of different futures are wrong -- this is, almost surely, inevitable -- but that when these assumptions are opaque and presented without critique or validation they may lead to inscrutable decision processes and sub-optimal decisions.
+\subsection{Synthesizing deep uncertainties for decision analysis}\label{sec:analysis-synthesize}
 
-\subsection{Synthesize}\label{sec:analysis-synthesize}
+Scenario-conditional probabilistic analysis allows for uncertainty quantification and optimization, and is valuable in many copntexts.
+However, scenarios are often explicitly provided without probabilities or likelihoods \cite<\eg, the \acrlongpl{ssp}>{vanvuuren_conditional:2008}.
+Thus, any such analysis is silent on the question of how to combine information across different scenarios.
+We term this the ``multiple \gls{pdf} problem''.
+Decision making around the multiple \gls{pdf} problem is susceptible to the cognitive biases that interfere with decision-making under uncertainty more generally \cite{tversky_judgment:1974,morgan_uncertainty:1990,srikrishnan_probabilistic:2022}.
+For example, while many analyses treat all scenarios as equally likely, this is often inconsistent with available information and can lead to poor decisions and outcomes \cite{wigley_equallylikely:2001,ho_scenarios:2019,hausfather_scenarios:2020}.
+Other analyses suggest using the worst-case scenario as a conservative measure.
+However, this approach is also problematic, since (i) there are no fundamental limits on what constitutes a worst-case scenario and (ii) improving performance under unlikely worst-case scenarios may lead to substantially impaired performance under more likely scenarios, which may or may not be acceptable to relevant stakeholders.
+There is thus a critical need for synthesizing insights across multiple probabilistic scenarios.
 
-As discussed in the previous subsection, estimating the expectation of arbitrary functions under uncertainty is a critical component of decision making under uncertainty.
-If the \glspl{sow} are drawn \gls{iid} from the target distribution, then the expected value of some function $f(\mathbf{s})$, can be readily estimated following the Monte Carlo estimate $\mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
-However, this is often not the case; for example, a fixed set of simulations not drawn from the target distribution may be available for analysis.
+\section{Re-weighting SOWs to synthesize across scenarios}\label{sec:reweighting}
+
+In this section we provide a formal method for synthesizing across multiple scenarios to inform climate risk management.
+Our objective is to develop a framework that (i) is conceptually and practically amenable to exploratory modeling; (ii) makes subjective modeling choices explicit and transparent; and (iii) allows decision analysts to estimate a probability distribution over outcomes.
+
+A particular need is to estimate the expectation of functions over \glspl{sow} (\ie, box e in \cref{fig:flowchart}).
+If the $J$ \glspl{sow} are drawn \gls{iid} from some distribution that credibly represents the true likelihood of different futures then the expected value of such a function, $f(\mathbf{s})$, can be readily approximated using the Monte Carlo estimate $\mathbb{E}\qty[f(\vb{s})] \approx \frac{1}{N} \sum_{j=1}^N f(\vb{s}_j)$.
+However, this is often not the case.
+For example, in \cref{sec:case-study} we will consider decision analysis where the \glspl{sow} are sampled from multiple physical models and \gls{rcp} scenarios, considering that not all \gls{rcp} scenarios are equally likely and that not all physical models are equally skillful.
 In this case, the formula may be adjusted to a weighted Monte Carlo estimate:
@@
 The challenge then becomes to suitably estimate the $w_j$.
-Paralleling joint probability methods for statistical analysis of tropical cyclones, we use a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
+Many such methods exist; drawing from joint probability methods for statistical analysis of tropical cyclones, we employ a grid-based approach \cite{johnson_clara:2013,resio_probabilities:2007,toro_jpm-os:2010}.
 First, we project the \glspl{sow} $\vb{s} \in \Omega$ onto a low-dimensional representation, which we denote $\qty{\psi_1, \psi_2, \ldots, \psi_J} \in \Psi$.
-Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the \gls{pdf} $p_\mathrm{belief}$ over each region.
+Then, we partition the parameter space into a region corresponding to each \gls{sow} and integrating the probability $p(\psi)$ over each region.
+
+Doing so requires a probabilistic model for this low-dimensional representation of the \glspl{sow},  $p(\psi)$.
+We denote this model $p_\mathrm{belief}$ to emphasize that it represents a subjective belief about the \glspl{sow}, rather than an objectively verifiable choice.
+In general we do not expect that stakeholders and experts will agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
+However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
+Choices for $p_\mathrm{belief}$ can be drawn from many sources, including expert elicitation or  results of previous analyses.
+These models can be interpreted as prior beliefs about \gls{slr} that could be incorporated into a Bayesian analysis as additional data is collected in the future, and thus can draw from literature on Bayesian prior selection and prior predictive checks \cite{gelman_workflow:2020}.
 
@@
 This step is illustrated in \cref{fig:grid-sketch}.
-Diagnostic checks, such as examining the histogram of weights (not shown), may be valuable protections against degeneracy.
+By the definition of cumulative distribution functions, $F_\mathrm{belief}(b) - F_\mathrm{belief}(a) = \int_a^b p_\mathrm{belief}(\psi') \dd{\psi'}$.
+Diagnostic checks, such as examining the histogram of weights to (not shown), may be useful protections against degeneracy.
 
@@
 As in \cref{sec:analysis-condition}, we must condition on a model: where the analysis of \cref{sec:analysis-condition} conditions upon deep uncertainties, the approach outlined in this subsection synthesizes across them.
-We reiterate that stakeholders and experts will not, in general, agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
-However, we posit that since we cannot be ``right,'' it is valuable to maximize the transparency of our implicit probabilistic assumptions, and suggest that writing down an explicit model for $p_\mathrm{belief}$ supports this aim.
-
 Considering multiple probabilistic models for $p_\mathrm{belief}$ can also be useful for understanding the sensitivity of the decision to the choice of $p_\mathrm{belief}$.
@@
 \end{enumerate}
-In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective priors $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).
+In the remainder of this section we describe data sources and treatment of \gls{slr} (\cref{sec:case-slr}), storm surge (\cref{sec:case-surge}), damages and metrics (\cref{sec:case-metrics}), and finally the subjective probabilistic models $p_\mathrm{belief}$ used to apply the re-weighting method described in \cref{sec:analysis-synthesize} to this case study (\cref{sec:case-priors}).
 
@@
 The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
-These four models represent physical processes, particularly of ice sheet dynamics, in different ways.
-The differences in these models leading to diverging estimates of how \gls{msl} responds to temperature.
-For a discussion of these model outputs we refer the reader to \citeA{ruckert_coastal:2019}.
-
 Estimates of nonstationary \gls{msl} also depend on anthropogenic forcing, which is itself deeply uncertain \cite{ho_scenarios:2019,srikrishnan_probabilistic:2022}.
-To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying distributions of \gls{msl}.
+To sample this uncertainty, we use simulations from each physical model under four \gls{rcp} scenarios, yielding sixteen time-varying probabilistic scenarios of \gls{msl}.
 
@@
 The divergence between the the best-case (blue) and worst-case (red) models is small in the early 21st century and increases rapidly thereafter.
-\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen models considered.
-The differences between models arise primarily due to different representations of Antarctic Ice Sheet contributions to global \gls{slr} and statistical calibration methodologies; for a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
-We return in \cref{sec:results-conditional} to the challenge of decision making under multiple \glspl{pdf}.
+\Cref{fig:lsl-evolution}(b) shows the \glspl{pdf} of mean sea level in 2100 (dashed vertical line in panel (a)) under each of the sixteen probabilistic scenarios considered.
+The stark differences between different scenarios of \gls{slr} arise primarily from different representations of Antarctic Ice Sheet contributions to global \gls{slr} and statistical calibration methodologies.
+For a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
+We return in \cref{sec:results-conditional} to the challenge of decision making given multiple \glspl{pdf}.
 
@@
     return periods.
-    Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula (eq.~\ref{eq:weibull}).
+    Dots indicate observed values; their $x$-value is their plotting position using the Weibull formula (eq.~S5).
     Gray lines show the 50, 80, and 95\% posterior confidence intervals from the Bayesian \gls{gev} fit (\cref{sec:case-surge}).
@@
 \end{equation}
-where $y'(t)$ is the storm surge (above \gls{msl}) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in \cref{eq:gev-dist}.
+where $y'(t)$ is the storm surge (above \gls{msl}) in year $t$ and a \gls{gev} distribution with location, shape, and scale parameters $\mu$, $\sigma$, and $\xi$, respectively, has the probability density function given in eq.~(S1).
 This model assumes stationarity, neglecting any potential time dependence.
@@
 We include two forms of prior information.
-First, we constrain the shape parameter to be positive, $\xi > 0$, to reflect knowledge about the support of $y'$, which for a variable distributed according to \cref{eq:gev-dist} is:
+First, we constrain the shape parameter to be positive, $\xi > 0$, to reflect knowledge about the support of $y'$, which for a variable distributed according to eq.~S1 is:
 \begin{equation*}
@@
 Specifically, we apply Inverse Gamma priors over the 2, 10, 100, and 500 year return levels, with means of \SIlist{4;6;10;15}{ft} and standard deviations of \SIlist{1.5;1.75;2.25;2.75}{ft}, respectively.
-The parameters of the Inverse Gamma distribution can be calculated from these moments (see eq.~\ref{eq:inv-gamma-params}).
-These means and standard deviations were chosen to represent plausible physical ranges (\cref{fig:surge-gev-priors}).
+The parameters of the Inverse Gamma distribution can be calculated from these moments (see eq.~S3).
+These means and standard deviations were chosen to represent plausible physical ranges (fig.~S4).
 
 For inference, we draw \num{10000} samples from the posterior distribution $p(\mu,\sigma,\xi | y')$ using Hamiltonian Markov Chain Monte Carlo \cite{betancourt_hamiltonian:2018,hoffman_nuts:2011} implemented in the Turing package of the Julia programming language \cite{perkel_julia:2019,ge_turing:2018,tarek_dynamicppl:2020,besancon_distributions.jl:2021,bezanson_julia:2012}.
-Diagnostics suggest (though cannot guarantee) convergence (see \cref{tab:surge-posterior-mcmc-diagnostics}).
+Diagnostics suggest (though cannot guarantee) convergence (see table S1).
 We evaluate the model's fit using posterior predictive checks \cite<see>[section 2.4 and references therein]{gelman_workflow:2020}.
-Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (\cref{fig:surge-test-statistics}) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
+Using the lag 1 and 2 partial autocorrelations, sample maximum, sample minimum, sample median, and Mann-Kendall test value as Bayesian test statistics, we find that draws from the posterior predictive distribution match the observed test statistics credibly (fig.~S9) although panels (a) and (b) suggest the possibility of temporal structure not captured by our stationary \gls{iid} model.
 Future efforts could represent this structure by conditioning the parameters of the distribution on relevant climate indices \cite<as in>[]{wong_surge:2018,farnham_orbrep:2018,farnham_jetstream:2017,ossandon_semibayesian:2021}.
@@
 Other model validations lend confidence to the stationary \gls{gev} model selected.
-For example, \cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges; the estimated return period  (shading) matches the the empirical plotting position (dots) and a positive control test (\cref{fig:surge-synthetic-data-experiment}) validates the model's ability to recover known parameter values.
+For example, \cref{fig:surge-obs-return}(b) shows the estimated return periods for these storm surges; the estimated return period  (shading) matches the the empirical plotting position (dots) and a positive control test (fig.~S6) validates the model's ability to recover known parameter values.
 
@@
 where $D(h-y)$ is a deterministic function specifying damage as a function of flood depth (relative to the house) and $p(y')$ is the probability density of storm surge.
-Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in \cref{fig:cost-depth-damage}.
-For comparison, \cref{fig:cost-depth-damage} also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
-Both  models show damage increasing with flood depth before reaching an upper limit.
+Following \citeA{zarekarizi_suboptimal:2020}, we use the \gls{hazus} depth-damage curves provided by \gls{fema}; this depth-damage relationship is shown in fig.~S1.
+For comparison, fig.~S1 also shows the ``Europa'' depth-damage relationship developed by the Joint Research Center of the European Commission's science and knowledge service \cite{huizinga_depthdamage:2016}.
+Both models show damage increasing with flood depth before reaching an upper limit but differ in the value of the upper limit and the rate at which damages approach it.
 Although \citeA{zarekarizi_suboptimal:2020} demonstrate that the choice of fragility function is important for informing house elevation, we use only the HAZUS model for simplicity.
@@
 However, the convolution of the HAZUS depth-damage equation with the \gls{gev} posterior does not have a tractable analytic solution.
-Instead, we estimate this convolution through a Monte Carlo method (see \cref{sec:alg-ead} for details).
-Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see \cref{sec:surrogate-ead}).
+Instead, we estimate this convolution through a Monte Carlo method (see section S1.2 for details).
+Then, because the expectation in \cref{eq:ead} depends only on $h-\overline{y}(t)$, we calculate expected annual damages for a wide range of possible heights, then use this to train a computationally efficient surrogate model (using linear interpolation; see section~S1.3).
 
@@
 Following \citeA{zarekarizi_suboptimal:2020}, we use estimates of construction cost from the Coastal Louisiana Risk Assessment \cite{fischbach_clara:2012}.
-We normalize this cost by house value; this cost curve is shown in \cref{fig:cost-up-front} and shows a large up-front cost plus a piecewise linear marginal cost.
+We normalize this cost by house value; this cost curve is shown in fig.~S3 and shows a large up-front cost plus a piecewise linear marginal cost.
   \item ``Lifetime expected damages'' is calculated following \cref{eq:led}.
@@
 
-We use a Gamma distribution for all three priors, parameterized following \cref{eq:gamma-dist}.
+We use a Gamma distribution for all three priors, parameterized following eq.~(S4).
 The distributions were chosen to be illustrative, rather than to reflect any particular scientific consensus.
@@
   \caption{
-    Subjective priors over \gls{slr} from 2022 to 2100, \ie $p_\mathrm{belief}(\psi)$.
+    Subjective probability distributions over \gls{slr} from 2022 to 2100, \ie $p_\mathrm{belief}(\psi)$.
     The name of the distribution, the parameters of the Gamma distribution with shape $\alpha$ and scale $\theta$, and the 2.5, 25, 50, 75, and 97.5th percentiles (values in \si{ft}).
@@
 
-We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through the lenses of exploratory modeling (\cref{sec:results-exploratory}), scenario-conditional analysis (\cref{sec:results-conditional}), and finally the proposed synthesis method (\cref{sec:results-synthesis}).
+We illustrate our approach to synthesizing uncertainties by sequentially analyzing the house elevation problem through multiple lenses for \gls{dmdu}.
+This allows us to demonstrate the advantages and limitations of each approach, and to highlight the value of synthesizing across multiple scenarios.
 
@@
 For some decision-makers, contextualizing this information against a few scenarios of \gls{slr} \cite<\eg, those of>[]{sweet_slr:2022} may prove sufficient for decision making.
-However, this analysis does not shed light on cost-benefit analyses or return periods, nor does it permit quantitative comparison against other possible decisions unless strong implicit assumptions are made.
+However, this analysis is silent on how to estimate cost-benefit comparisons, return periods, and other trade-offs.
 
-\subsection{Scenario-conditional optimization}\label{sec:results-conditional}
+\subsection{Scenario-conditional probabilistic analysis}\label{sec:results-conditional}
 
 We now turn to the scenario-conditional analysis described in \cref{sec:analysis-condition}.
-Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draw from one of the sixteen models of \gls{slr} shown in \cref{fig:lsl-evolution}.
+Whereas the exploratory analysis of the previous subsection interpreted each time series of future sea level as a sample from the space of possible futures, we can also interpret each \gls{sow} as a draw from one of the sixteen probabilistic scenarios of \gls{slr} shown in \cref{fig:lsl-evolution}.
 As discussed in \cref{sec:analysis-condition}, this allows a formal estimation of decision metrics, conditional on the chosen scenario.
@@
 
-This probabilistic interpretation allows us to compute, for example, expected values.
-For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen models considered (we highlight three representative models).
+As discussed in \cref{sec:analysis-condition}, this probabilistic interpretation allows us to compute expected values of functions.
+For example, \cref{fig:tradeoffs-by-rcp}(a) plots the expected total lifetime cost as a function of $\Delta h$ for the sixteen probabilistic scenarios considered (we highlight three representative models).
 This panel shows lifetime expected damages as a function of $\Delta h$, shown in \cref{fig:tradeoffs-by-rcp}(b,  plus the up-front cost of construction.
-Because there are high fixed costs associated with building (see cost curve in \cref{fig:cost-up-front}), it generally does not make sense to raise the house by only a small amount, since this incurs these fixed costs without providing substantial damage reduction.
+Because there are high fixed costs associated with building (see cost curve in fig.~S3), it generally does not make sense to raise the house by only a small amount, since this incurs these fixed costs without providing substantial damage reduction.
 \Cref{fig:tradeoffs-by-rcp} shows that estimates of trade-offs between up-front cost and expected lifetime costs are highly sensitive to the chosen scenario.
@@
 
-The \gls{sow} re-weighting framework described in \cref{sec:analysis-synthesize} can help overcome the limitations of scenario-conditional analysis.
+The \gls{sow} re-weighting framework described in \cref{sec:reweighting} can help overcome the limitations of scenario-conditional analysis.
 In this section we illustrate how this approach can help to shed light on climate risk management under deep uncertainty.
@@
   \caption{
-    Impact of subjective priors for local sea level on implicit weight given to each \gls{rcp} scenario and physical model.
-    We develop three distributions (``subjective priors'') representing plausible probabilistic beliefs about \gls{msl} at Sewells Point, VA in 2100, relative to the present.
-    The \glspl{pdf} of these subjective priors are shown in panel (A).
-    In panels (B-C) we show the relationship between these subjective priors and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
-    Specifically, (B-C) show the average weight given to each model by each of the subjective priors.
+    Impact of different subjective probability distributions for local sea level on implicit weight given to each \gls{rcp} scenario and physical model.
+    We develop three distributions representing plausible probabilistic beliefs ($p_\mathrm{belief}$) over \gls{msl} at Sewells Point, VA in 2100, relative to the present.
+    The \glspl{pdf} of these distributions are shown in panel (A).
+    In panels (B-C) we show the relationship between these distributions and the 16 probabilistic models (four \gls{rcp} scenarios and four physical representations) available.
+    Specifically, (B-C) show the average weight given to each model by each choice for $p_\mathrm{belief}$.
   }\label{fig:lsl-priors-weights}
@@
 One application of this method is to diagnose the assumptions which which different $p_\mathrm{belief}$ are consistent.
-\Cref{fig:lsl-priors-weights}(B-D) shows the total weight that each prior assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
+\Cref{fig:lsl-priors-weights}(B-D) shows the total weight that each choice of $p_\mathrm{belief}$ assigns to \glspl{sow} generated by each \gls{rcp} scenario and physical model.
 Specifically, weights are computed as the sum of weights assigned to each \gls{sow} sampled from that model.
 For example, the rapid \gls{slr} scenario (green line in \cref{fig:lsl-priors-weights}) places most weight on \glspl{sow} produced by the DP16 model, and particularly on \gls{rcp} 8.5 which by some accounts is unlikely given current policy \cite{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}.
-Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 \cite<also unlikely given current policy;>{hausfather_scenarios:2020,srikrishnan_probabilistic:2022}) and \gls{rcp} 4.5.
+Conversely, the slow \gls{slr} scenario (red line) places most weight on the BRICK models, particularly \gls{rcp} 2.6 \cite<also unlikely given current policy;>{hausfather_scenarios:2020,srikrishnan_probabilistic:2022} and \gls{rcp} 4.5.
 The uncertain \gls{slr} scenario (blue line) allocates approximately equal weight across models.
@@
 This method can also be used to calculate expectations, allowing us to revisit the trade-off diagrams of \cref{fig:tradeoffs-by-rcp}.
-\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each model.
+\Cref{fig:tradeoffs-by-prior} shows the total lifetime cost (panel A) and lifetime expected damages (panel B) under each choice of $p_\mathrm{belief}$.
 Notably, they give different guidance.
@@
 Under the intermediate / uncertain \gls{slr} assumption, the expected lifetime costs are similar for elevating or not elevating the house, and thus the benefit to cost ratio is nearly 1.
-Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (\cref{fig:cost-up-front}) but offers relatively little flood reduction.
-Based on this analysis, we would recommend that the owner if this hypothetical home elevate by approximately 4-6\si{ft}.
-Alternatively, the homeowner could choose to defer the deicision of whether, and how high, to elevate; our analysis did not consider this possibility but there is a rich literature on flexible design and engineering options analysis in climate risk analysis \cite<\eg,>{fletcher_learning:2019,fletcher_groundwater:2019,hui_adaptive:2018,garner_slrise:2018,herman_control:2020,deneufville_eoa_theory:2019}.
+Under all assumptions, elevating by only a few feet is impractical because it involves paying the large fixed costs of elevation (fig.~S3) but offers relatively little flood reduction.
+Based on this analysis, we would recommend that the owner if this hypothetical home elevate by approximately 4-6\si{ft} or not at all.
+Alternatively, the homeowner could choose to defer the decision of whether, and how high, to elevate; our analysis did not consider this possibility but there is a rich literature on flexible design and engineering options analysis in climate risk analysis \cite<\eg,>{fletcher_learning:2019,fletcher_groundwater:2019,hui_adaptive:2018,garner_slrise:2018,herman_control:2020,deneufville_eoa_theory:2019}.
 
@@
 The first category has to do with limitations of the underlying method proposed for re-weighting \glspl{sow}.
-For example, we develop a subjective prior belief $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
-Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic, meaning that many possible low-dimensional representations are possible; time series with the same \gls{msl} in 2100 may differ in other ways.
-For problems with more sources of uncertainty, such as multisector problems, choosing a low-dimensional representation may prove more challenging.
-In such settings, more detailed diagnostics and sensitivity analyses may shed light on the appropriateness of different modeling choices.
+For example, we develop a subjective probabilistic model $p_\mathrm{belief}(\Psi)$ over \gls{msl} in the year 2100.
+Although this is a low-dimensional representation of the full time series, it is not a sufficient statistic.
+In other words,  many possible low-dimensional representations are possible and time series with the same \gls{msl} in 2100 may differ in other ways.
+For problems with more sources of uncertainty, such as multisector problems, choosing an appropriate low-dimensional representation may prove challenging.
+In such settings, diagnostics and sensitivity analyses may shed light on the appropriateness of different modeling choices.
 A related concern is that we developed our three distributions for $p_\mathrm{belief}(\Psi)$  in an \emph{ad hoc} fashion that may not represent well-calibrated beliefs.
@@
 In particular, we hypothesize that incorporating improved representations of risk aversion into decision support may substantially improve their usability.
-One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \cite{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \cite[provides a framework for addressing some of these]{zarekarizi_suboptimal:2020}.
+One could also extend the analysis to consider additional sources of uncertainty such as depth-damage relationships \cite{Rozer:2019,nofal_fragility:2020}, the cost of elevating a house, the house lifespan, the effective discount rate, and value of the land on which the house is built \cite{zarekarizi_suboptimal:2020}.
 Finally, while here we consider the decision to be a one-time decision, one could also frame this as a sequential decision problem.
@@
 These limitations motivate several directions for future research.
-From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \cite<see fig.~1 of>[]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimation of local hazard, subject to (still necessarily subjective) priors over key parameters.
+From a methodological perspective, developing model chains that capture uncertainties in global energy and economic pathways, global climate sensitivity, and local hazard response \cite<see fig.~1 of>[]{moss_uncertainties:2000} offers a principled framework for fully probabilistic estimation of local hazard, subject to (still necessarily subjective) probabilistic models for key parameters.
 From a decision support perspective, improved understanding of the conditions under which household-scale strategies for flood risk management, like elevation, achieve relevant objectives could support improved resilience and adaptation.
@@
 
-xThe proposed \gls{sow} re-weighting framework can be applied to inform critical challenges in climate risk management.
+The proposed \gls{sow} re-weighting framework can be applied to inform critical challenges in climate risk management.
 An obvious area of application is to the design of infrastructure.
@@
 Vivek Srikrishnan,
+Skip Wishbone,
 and two anonymous peer reviewers

### 2022-10-05 tikz/bayes-rdm.tex
changes made; need to update letter to reviewers

@@
     \node [block, label={[]225:{\footnotesize (d)}}] at (-4, 3) (models) {
-            Multiple models\\
+            Probabilistic scenarios\\
             $\vb{M} = \qty{M_1, \ldots, M_K}$\\
-            for $p(\Omega | M_k)$
+            for $p(s | M_k)$
         };
@@
             $\vb{w} = \qty{w_1, \ldots, w_J}$\\
-            where $w_j \in (0, 1)$
+            where $\sum_{j=1}^J w_j = 1$
         };

### 2022-11-23 papers/earths-future/submission_03.tex
reviewer response

--- 2022-10-05 papers/earths-future/submission_02.tex
@@
 \begin{keypoints}
-  \item Decisions about how to manage climate risks are often made under deep uncertainty.
   \item We introduce a Bayesian framework to transparently synthesize and characterize deep uncertainties with the goal to support decision-making.
@@
 In the first step, exploratory or bottom-up modeling is used to build insight and identify potential system vulnerabilities \cite{moallemi_exploratory:2020,bankes:1993,brown_decisionscaling:2012}.
-In the second step, we synthesize the results of exploratory modeling to formally estimate performance metrics and trade-offs using probabilistic models for uncertainty.
+In the second step, we integrate exploratory ensembles of deep uncertainties into a single probabilistic representation (we refer to this as ``synthesizing'' deep uncertainties) to formally estimate performance metrics and trade-offs using subjective probability distributions.
 Drawing from the literature on building predictive models when all models are wrong \cite{box_sciencestatistics:1976,gelman_philosophy:2013,Piironen:2017eh}, we interpret these probability distributions not as statements of fact, but rather as a self-consistent framework for reasoning about how different assumptions lead to different inferences.
@@
 In this section we introduce a conceptual framework and notation for decision analysis under deep uncertainty.
-We then apply this framework to understand exploratory modeling (\cref{sec:analysis-explore}) and scenario-conditional probabilistic analysis (\cref{sec:analysis-condition}).
-In \cref{sec:analysis-synthesize} we discuss why synthesizing across multiple scenarios is necessary.
-In \cref{sec:reweighting} we provide a formal method for synthesizing across scenarios.
+Many bottom-up exploratory modeling frameworks used in climate risk management and related fields use a system model ($f$ in \cref{fig:flowchart})  to characterize the system's response to a wide range of plausible futures, often called \glspl{sow} ($\vb{s}$ in \cref{fig:flowchart}).
+This analysis is often used to explore vulnerabilities and build knowledge about the system \cite{bankes:1993}, and in general exploratory modeling frameworks aim to avoid making explicit judgments about the relative likelihood of different futures.
 
-It is common in climate risk management to work with an ensemble of trajectories of climate, economic, and other key variables, which we refer to as \glspl{sow}.
-Often these \glspl{sow} are computationally expensive to generate, and thus it is important to use them efficiently.
-Following \cref{fig:flowchart}, we consider using $J$ \glspl{sow}, $\vb{s} = \qty{s_1, s_2, \ldots, s_J}$, to evaluate $I$ candidate decisions, $\vb{x} = \qty{x_1, x_2, \ldots, x_I}$.
-For each \gls{sow} $s_j \in \vb{s}$ and decision $x_i \in \vb{x}$ we use a system model $f$ (comprised of multiple components) to calculate a set of metrics describing the performance of decision $x_i$ on \gls{sow} $s_j$, which we denote $u_{ij} = f(x_i, s_j)$.
-While we assume for simplicity that the decision space is known and finite, this approach could be coupled to a policy search model that proposes candidate decisions.
+However, as discussed in \cref{sec:introduction}, estimates of trade-offs between desired performance metrics (\eg, cost and reliability) depend on probabilistic models of uncertainty.
+In this paper we present a method for integrating the \glspl{sow} used in exploratory modeling into a formal decision analytic framework using a subjective probability distribution over the space of possible futures, which is used to infer implicit weights over the \glspl{sow}.
+This approach is particularly suited for problems where the \glspl{sow} are generated from or conditioned on specific scenarios (\eg, \gls{rcp} scenarios) or where there are multiple models of the underlying processes (\eg, multiple models of the response of local sea levels to global temperature), which can lead to the ``multiple \gls{pdf} problem'' (discussed in \cref{sec:analysis-condition}).
+A motivating advantage is that it makes assumptions about the likelihoods of different \glspl{sow} transparent to decision-makers.
 
@@
 This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
-We illustrate this concept in boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario $M_k$.
-By treating the \glspl{sow} as \gls{iid} draws from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
+To clarify language, we draw a distinction between a \gls{sow}, which is a single realization of a possible future, and a \emph{scenario}, which is a model that can be used to generate \glspl{sow}.
+These terms are defined relative to the problem at hand; for example, a global emissions trajectory may be a \gls{sow} from an integrated assessment model, but when it is used as a deterministic input to an ice sheet model we would describe it as a scenario.
+
+We illustrate this distinction in boxes (d) and (e) of \cref{fig:flowchart}, denoting the particular scenario $M_k$.
+We assume that each scenario is probabilistic, that is that \glspl{sow} are drawn \gls{iid} from $M_k$, the set of outcomes $u_{i, j}$ can be interpreted as \gls{iid} draws from the conditional distribution over outcomes, $p(u | x_i, M_k)$.
 This ``scenario-conditional'' probabilistic interpretation of \glspl{sow} allows for fully probabilistic quantification of uncertainty and optimization, conditional on a particular scenario.
@@
 
-We can also characterize \gls{dmdu} methods that sample a set of parameters from fixed ranges as scenario-conditional analyses.
+We can also apply this theoretical lens to examine the approach, common in \gls{dmdu} applications, of sampling parameters from a set of fixed ranges.
+The \emph{scenario} in this case is thus the choice of bounds on the parameters; it is consistent with our above definition of a scenario because it provides a probabilistic model that can be used to sample \glspl{sow}.
 For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
@@
 
-Scenario-conditional probabilistic analysis allows for uncertainty quantification and optimization, and is valuable in many copntexts.
+Scenario-conditional probabilistic analysis allows for uncertainty quantification and optimization, and is valuable in many contexts.
 However, scenarios are often explicitly provided without probabilities or likelihoods \cite<\eg, the \acrlongpl{ssp}>{vanvuuren_conditional:2008}.
@@
 
-In this section we provide a formal method for synthesizing across multiple scenarios to inform climate risk management.
+In this section we provide a formal method for integrating exploratory ensembles of deep uncertainties into a single probabilistic representation.
 Our objective is to develop a framework that (i) is conceptually and practically amenable to exploratory modeling; (ii) makes subjective modeling choices explicit and transparent; and (iii) allows decision analysts to estimate a probability distribution over outcomes.
@@
 
-Doing so requires a probabilistic model for this low-dimensional representation of the \glspl{sow},  $p(\psi)$.
-We denote this model $p_\mathrm{belief}$ to emphasize that it represents a subjective belief about the \glspl{sow}, rather than an objectively verifiable choice.
+Implementing this approach requires choosing a probability distribution for this low-dimensional representation of the \glspl{sow},  $p(\psi)$, reflecting subjective belief about the \glspl{sow}.
+We denote this $p_\mathrm{belief}(\psi)$ to emphasize that it represents a subjective belief about the \glspl{sow}, rather than an objectively verifiable choice.
 In general we do not expect that stakeholders and experts will agree on $p_\mathrm{belief}$ because there is not, even conceptually, an objectively correct choice \cite{oreskes_verification:1994,walker_deep:2013}.
@@
 
-\subsection{Prior over \glsentrylong{slr}}\label{sec:case-priors}
+\subsection{Subjective probability distributions for \glsentrylong{slr}}\label{sec:case-priors}
 
@@
 
-We use a Gamma distribution for all three priors, parameterized following eq.~(S4).
+We use a Gamma distribution for all three probability distributions, parameterized following eq.~(S4).
 The distributions were chosen to be illustrative, rather than to reflect any particular scientific consensus.
@@
 
-We developed these priors for didactic purposes, to illustrate a range of possible beliefs.
+We developed these subjective distributions for didactic purposes, to illustrate a range of possible beliefs.
 We can compare them, for example, with analysis published by \gls{noaa}, which project \SIlist{1.94;2.62;4.27;5.25;6.89}{ft} for the low, intermediate, low intermediate, intermediate high, and high scenarios, respectively \cite[table.~2.4]{sweet_slr:2022}.

### 2022-11-23 scripts/06-combining.jl

@@ -79,7 +79,7 @@ function plot_priors()
     p = plot(;
         xlabel="SLR 2022-2100, at Sewells Point, VA [ft]",
         ylabel="Probability Density",
-        title="(a): Implicit Prior",
+        title="(a): Subjective Probability Distribution",
     )
     for (prior, color) in zip(priors, [:red, :blue, :green])
         plot!(
@@ -109,7 +109,7 @@ function plot_weight(s::Vector{<:HouseElevation.LSLSim})
         w_avg = combine(groupby(df, [:model, :rcp]), prior.name => sum => :weight)
         @assert sum(w_avg[!, :weight]) ≈ 1
 
-        letter = collect('a':'z')[i]
+        letter = collect('a':'z')[i+1]
         p = @df w_avg groupedbar(
             :model,
             :weight,

### 2022-12-03 papers/earths-future/submission_03.tex
all done?

@@
 In this paper we present a method for integrating the \glspl{sow} used in exploratory modeling into a formal decision analytic framework using a subjective probability distribution over the space of possible futures, which is used to infer implicit weights over the \glspl{sow}.
-This approach is particularly suited for problems where the \glspl{sow} are generated from or conditioned on specific scenarios (\eg, \gls{rcp} scenarios) or where there are multiple models of the underlying processes (\eg, multiple models of the response of local sea levels to global temperature), which can lead to the ``multiple \gls{pdf} problem'' (discussed in \cref{sec:analysis-condition}).
+This approach is particularly suited for problems where the \glspl{sow} are generated from or conditioned on specific scenarios (\eg, \gls{rcp} scenarios) or where there are multiple models of the underlying processes (\eg, multiple parameterizations of the response of local sea levels to global temperature), which can lead to the ``multiple \gls{pdf} problem'' (discussed in \cref{sec:analysis-condition}).
 A motivating advantage is that it makes assumptions about the likelihoods of different \glspl{sow} transparent to decision-makers.
@@
 This commonly arises when a single deep uncertainty (\eg, an emissions pathway) is used as an input for a stochastic model.
-To clarify language, we draw a distinction between a \gls{sow}, which is a single realization of a possible future, and a \emph{scenario}, which is a model that can be used to generate \glspl{sow}.
-These terms are defined relative to the problem at hand; for example, a global emissions trajectory may be a \gls{sow} from an integrated assessment model, but when it is used as a deterministic input to an ice sheet model we would describe it as a scenario.
+To clarify language, we draw a distinction between a \emph{\gls{sow}}, which is a single realization of a possible future, and a \emph{probabilistic scenario}, which we define loosely as a set of assumptions for which probabilistic projections (\ie, \glspl{sow}) are available or can be simulated.
+For example, in the case study described in \cref{sec:case-slr} we use four physical models of the processes relevant to \gls{slr} and four \gls{rcp} scenarios to generate a total of 16 probabilistic scenarios.
 
@@
 We can also apply this theoretical lens to examine the approach, common in \gls{dmdu} applications, of sampling parameters from a set of fixed ranges.
-The \emph{scenario} in this case is thus the choice of bounds on the parameters; it is consistent with our above definition of a scenario because it provides a probabilistic model that can be used to sample \glspl{sow}.
+The \emph{scenario} in this case is thus the choice of bounds on the parameters; it is consistent with our above definition of a probabilistic scenario because \glspl{sow} can be sampled using probabilistic methods.
 For example, \citeA{sriver_sealevel:2018} sample parameters describing the rate of \gls{slr} across a range of values to inform coastal adaptation.
@@
 We analyze simulations of \gls{msl} at Sewells Point, VA from four probabilistic physical models using data published in \citeA{ruckert_coastal:2019}.
-The four models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
+The four physical process models considered are (i) the BRICK model (version 0.2) with slow (``BRICK Slow'') and (ii) fast (``BRICK Fast'') ice sheet dynamics \cite{wong_brick0.2:2017}, (iii) the \citeA{kopp_probabilistic:2014} model (``K14''), and (iv) the \citeA{deconto_antarctica:2016} model (``DP16'').
 The \citeA{kopp_probabilistic:2014} and \citeA{deconto_antarctica:2016} models have a ten year time step, which we linearly interpolate onto a one year time step for consistency.
@@
 For a more detailed discussion we refer the reader to \citeA{ruckert_coastal:2019}.
-We return in \cref{sec:results-conditional} to the challenge of decision making given multiple \glspl{pdf}.
+We return in \cref{sec:results-conditional} to the challenge of decision making given multiple scenarios.
 
@@
     (B): trade-off between up-front cost and lifetime expected damages (eq.~\ref{eq:led}).
-    Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios times four physical parameterizations) considered.
+    Light gray lines show estimates for all 16 models (four \gls{rcp} scenarios $\times$ four physical process models) considered.
     Colored lines highlight three representative models for emphasis.
@@
 
-This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic models.
+This approach is, in a sense, another form of exploratory modeling: instead of considering a very large ensemble of \glspl{sow}, we consider a much smaller set of probabilistic scenarios.
 Scenario-conditional analysis can be attractive because it allows modelers to focus on their domain expertise (\eg, the response of ice sheets and global sea levels to a particular climate future).
@@
 As a last example, \citeA{haasnoot_sealevelrise:2021} identify global adaptation needs for different \gls{slr} scenarios.
-Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which model to design for onto the end user.}
+Although this scenario-conditional analysis is appropriate for understanding differences between models, its key limitation is that it \textbf{places the burden for deciding which probabilistic scenario to design for onto the end user.}
 Since not all house owners or contractors have expertise in assessing the relative likelihood of different climate futures, they may not be well positioned to make this decision.

### 2022-12-03 scripts/plotutils.jl

@@ -96,3 +96,7 @@ function plot_return_period(
 
     return p
 end
+
+function cummean(x)
+    return cumsum(x) ./ collect(1:length(x))
+end
